### 2018-02-14 paper.tex
init commit

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\begin{document}

%% Title information
\title[Short Title]{Pushdown Control-Flow Analysis via Refunctionalization}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
\subtitle{Filling the Gap between Abstracting Abstract Machine and Abstracting Definitional Interpreter}        %% \subtitle is optional
\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}                    %% \country is recommended
}
\email{first1.last1@inst1.edu}          %% \email is recommended

%% Author with two affiliations and emails.
\author{First2 Last2}
\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position2a}
  \department{Department2a}             %% \department is recommended
  \institution{Institution2a}           %% \institution is required
  \streetaddress{Street2a Address2a}
  \city{City2a}
  \state{State2a}
  \postcode{Post-Code2a}
  \country{Country2a}                   %% \country is recommended
}
\email{first2.last2@inst2a.com}         %% \email is recommended
\affiliation{
  \position{Position2b}
  \department{Department2b}             %% \department is recommended
  \institution{Institution2b}           %% \institution is required
  \streetaddress{Street3b Address2b}
  \city{City2b}
  \state{State2b}
  \postcode{Post-Code2b}
  \country{Country2b}                   %% \country is recommended
}
\email{first2.last2@inst2b.org}         %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Abstracting abstract machines (AAM) is a systematic methodology for constructing abstract interpreter
  from a concrete small-step abstract machine. Recent progress applies the same idea on concrete definitional 
  interpreter and obtains a big-step abstracting definitional interpreter (ADI) written in monadic style.
  But the relation between small-step AAM and big-step ADI is not well-studied.

  In this paper, we show their correspondence and a series of transformations from a small-step 
  abstracting abstract machine to a big-step abstracting definitional interpreter. 
  The transformations include fusing, disentangling, 
  refunctionalizing and then monadification/un-CPS, each of them properly handle the collecting semantics and 
  the non-determinism of abstract interpretation. 
  The transformations also uncover several intermediary forms of abstract interpreter.
  Remarkably, we reveal how computatble precise call/return match in control-flow analysis is obtained by 
  refunctionalizing a small-step abstracting abstract machine with unbounded stack, 
  as well as explain how it been lost by defunctionalizing the higher-order continuations into 
  an algebraic data type.

  TODO: Shiver's formulation of CFA in denotational style.
  %Further, from another direction, we show Shiver's original implementation of control-flow analysis
  %in denotational style can be also transformed into natural semantics by closure conversion.

\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{pushdown analysis, abstract interpretation, static analysis}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}

\subsection{Motivation}

\subsection{Contributions}

\begin{itemize}
  \item We fill the gap between small-step AAM and big-step ADI by developing a series of systematic transformations from small-step AAM to big-step ADI, and also show that the reversed direction is also possible.

  \item We present an abstract interpreter written in CPS obtained from refunctionalization of small-step AAM. This abstract interpreter naturally has pushdown control-flow property, and we analyze how precise call/return match is obtained by representing evaluation context as a continuation function, as well as how it is lost by representing continuation as a data type in the defining language.

  \item We show that the correspondence not only exists between concrete abstract machines and evaluator, and also exists for abstracting abstract machines and abstract evaluators. Galois connection?
\end{itemize}

\subsection{Outline}

\section{Background}

In this section, we first describe the syntax of A-Normal Form (ANF) $\lambda$-calculus and a CESK machine for executing it concretely. Then we review several recent works in control flow analysis: 1) Abstracting Abstract Machine as an abstract counterpart of CESK machine, 2) abstract interpreter written in a monadically-parameterized style, and 3) precise call/return match problem and its solutions.

\subsection{A-Normal Form $\lambda$-Calculus}

\subsection{CESK Machine}

\subsection{Abstracting Abstract Machine}

\subsection{Monadic Abstract Interpreter}

\subsection{Call/Return Match in Control-Flow Analysis}

\subsubsection{AAM with Unbound Stack}

\subsubsection{Computable Solutions}
P4F ANF, not works on direct style,PDCFA,AAC

\section{FROM AAM TO ADI}

We start from the AAM with unbounded stack.

\subsection{Fusing}

\subsection{Disentangling}

\subsection{Refunctionalization}

allocating continuations of defined language on the heap of defining language.

%probably Store Widening \& Abstract Garbage Collection
%monadic
\subsection{Monadification: From Refunctionalized AAM to ADI}

For one direction, we can transform the refunctionalized AAM to abstracting definitional interpreter
by adding monads.

\subsection{Un-CPS: From Refunctionalized AAM to Direct Style}

In another direction, we can transform the refunctionalized AAM to direct style, but explictly use 
side effects such as assignment and mutation to handle collecting semantics and non-determinism.

\section{Refunctionalized AAM}

\subsection{Precise Call/Return Match}

show the precision of stack

\subsection{Termination}

show that number of continuation is finite.

\subsection{Soundness}

\subsection{Complexity}

\section{Detour: Shiver's \textit{k}-CFA}

\section{Perspective}

\subsection{Defunctionalization}

\subsection{Refunctionalization}

\section{Related Work}

\section{Conclusion}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Bibliography
\bibliography{references}

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

\end{document}

### 2018-02-15 paper.tex
misc

@@
 
+  % Mention Danvy's work
   In this paper, we show their correspondence and a series of transformations from a small-step 
@@
   \item We present an abstract interpreter written in CPS obtained from refunctionalization of small-step AAM.
-This abstract interpreter naturally has pushdown control-flow property, and we analyze how precise call/return match is obtained by representing evaluation context as a continuation function, as well as how it is lost by representing continuation as a data type in the defining language.
+This abstract interpreter naturally has pushdown control-flow property, and we show how precise call/return match is obtained by representing evaluation context as a continuation function, as well as how it is lost by representing continuation as a data type in the defining language.
+We also analyze its termination, soundness and complexity property.
 
@@
 In this section, we first describe the syntax of A-Normal Form (ANF) $\lambda$-calculus and a CESK machine for executing it concretely.
-Then we review several recent works in control flow analysis: 1) Abstracting Abstract Machine as an abstract counterpart of CESK machine, 2) abstract interpreter written in a monadically-parameterized style, and 3) precise call/return match problem and its solutions.
+Then we review several recent works in control flow analysis: 1) Abstracting Abstract Machine as an abstract counterpart of CESK machine, 2) abstract interpreter written in a monadically-parameterized style, and 3) precise call/return match problem and existing solutions.
 

### 2018-02-23 paper.tex
update

@@
 %% http://ctan.org/pkg/subcaption
+\usepackage{syntax}
 
@@
 \begin{abstract}
-  Abstracting abstract machines (AAM) is a systematic methodology for constructing abstract interpreter
-  from a concrete small-step abstract machine. Recent progress applies the same idea on concrete definitional 
-  interpreter and obtains a big-step abstracting definitional interpreter (ADI) written in monadic style.
-  But the relation between small-step AAM and big-step ADI is not well-studied.
+  Abstracting abstract machines (AAM) is a systematic methodology 
+  for constructing abstract interpreter that derived from a concrete
+  small-step abstract machine. Recent progress applies the same idea 
+  on concrete definitional interpreter and obtains a big-step 
+  abstracting definitional interpreter (ADI) written in monadic style.
+  Danvy shows the correspondence between abstract machines and interpreters.
+  In this paper, we study the relations between small-step AAM and 
+  big-step ADI.
 
-  % Mention Danvy's work
-  In this paper, we show their correspondence and a series of transformations from a small-step 
-  abstracting abstract machine to a big-step abstracting definitional interpreter. 
-  The transformations include fusing, disentangling, 
-  refunctionalizing and then monadification/un-CPS, each of them properly handle the collecting semantics and 
-  the non-determinism of abstract interpretation. 
-  The transformations also uncover several intermediary forms of abstract interpreter.
-  Remarkably, we reveal how computatble precise call/return match in control-flow analysis is obtained by 
-  refunctionalizing a small-step abstracting abstract machine with unbounded stack, 
-  as well as explain how it been lost by defunctionalizing the higher-order continuations into 
-  an algebraic data type.
+  We show their correspondence and how to syntactically transform a small-step 
+  abstracting abstract machine to a big-step abstracting definitional 
+  interpreter.
+  The transformations include fusing, disentangling, refunctionalizing 
+  and monadification(un-CPS), and all of them properly handle the 
+  collecting semantics and the non-determinism of abstract interpretation. 
+  After each transformation, we also obtain an intermediary form of 
+  abstract interpreter.
+  Remarkably, follow the idea that evaluation contexts are defuntionalized 
+  continuations, we reveal how precise call/return match in control-flow 
+  analysis is obtained by refunctionalizing a small-step abstracting 
+  abstract machine,  % with unbounded stack
+  as well as explain how it been lost by defunctionalizing the higher-order
+  continuations into an algebraic data type. Further, we show how to turn 
+  precise call/return match into computable analysis.
 
-  TODO: Shiver's formulation of CFA in denotational style.
-  %Further, from another direction, we show Shiver's original implementation of control-flow analysis
-  %in denotational style can be also transformed into natural semantics by closure conversion.
+  TODO: Maybe something Shiver's formulation of CFA in denotational style.
 
@@
 
+theoretical and practical
+
 \begin{itemize}
-  \item We fill the gap between small-step AAM and big-step ADI by developing a series of systematic transformations from small-step AAM to big-step ADI, and also show that the reversed direction is also possible.
+  \item We fill the gap between small-step AAM and big-step ADI 
+    by developing a series of systematic transformations from 
+    small-step AAM to big-step ADI, and also the reversed 
+    direction is also possible.
 
-  \item We present an abstract interpreter written in CPS obtained from refunctionalization of small-step AAM.
-This abstract interpreter naturally has pushdown control-flow property, and we show how precise call/return match is obtained by representing evaluation context as a continuation function, as well as how it is lost by representing continuation as a data type in the defining language.
-We also analyze its termination, soundness and complexity property.
+  \item We present an abstract interpreter written in continuation-passing
+    style that is obtained from refunctionalization of a small-step AAM.
+    This abstract interpreter naturally has pushdown control-flow property, 
+    and we show how precise call/return match is obtained by 
+    representing evaluation context as a continuation function, 
+    as well as how it is lost by representing continuation as a 
+    data type in the defining language.
+    As a form of abstract interpreter, We also analyze its termination,
+    soundness and complexity property.
 
-  \item We show that the correspondence not only exists between concrete abstract machines and evaluator, and also exists for abstracting abstract machines and abstract evaluators. Galois connection?
+  \item We show that the correspondence not only exists between 
+    concrete abstract machines and evaluator, and also exists 
+    for abstracting abstract machines and abstract evaluators. 
+    Galois connection?
 \end{itemize}
@@
 
+\subsection{Style}
+
+We use Scala language to demonstrate the idea and each step of transformations.
+
 \section{Background}
 
-In this section, we first describe the syntax of A-Normal Form (ANF) $\lambda$-calculus and a CESK machine for executing it concretely.
-Then we review several recent works in control flow analysis: 1) Abstracting Abstract Machine as an abstract counterpart of CESK machine, 2) abstract interpreter written in a monadically-parameterized style, and 3) precise call/return match problem and existing solutions.
+In this section, we frist describe the target language we will
+analyze and a concrete CESK machine as operational semantics for it.
+Then we review several recent works in control flow analysis for
+higher-order languages:
+1) abstracting abstract machines (AAM) as an abstract interpretation
+counterpart of CESK machines,
+2) definitional abstract interpreter written in a monadically-parameterized
+style, and
+3) precise call/return match problem in control-flow analysis and its existing solutions.
+The abstracting abstract machines form the basis of our transformations, and
+monadic abstract interpreter is the end point. In the meanwhile, the precise
+call/return match is obtained during transformation.
 
-\subsection{A-Normal Form $\lambda$-Calculus}
+\subsection{A-Normal Form $\lambda$-Calculus} \label{anfsyntax}
+
+Tranditionally, continuation-passing style is a popular intermediate representation
+for analyzing functional programs because of it exposes control transfer explicitly
+and simplifies the analysis \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
+Here, we choose to use a direct-style intermediate representation as target for
+clarity but without losing simplicty and generality.
+The transformations we will show in the rest of this paper
+also work on abstract machines for plain $\lambda$-calculus.
+
+To begin, we now present the syntax of a call-by-value $\lambda$-calculus language in
+administrative normal form (ANF) \cite{flanagan1993essence}.
+
+\begin{verbatim}
+e \in Exp ::= ae
+            | (let ([x (f ae)]) e)
+            | (letrec ([x ae] ...) e) (TODO: maybe not include letrec?)
+ae \in AExp ::= x
+              | lam
+lam \in Lam ::= (lambda (x) e)
+x ... are variable names
+\end{verbatim}
+
+In ANF, all the function applications must be administrated in a \texttt{let} expression,
+and then bounded to a variable name under current environment.
+We can use \texttt{letrec} to create mutually recursive function bindings. But \texttt{letrec}
+only accepts atomic expressions that appear at the right-hand side of a binding.
+The atomic expressions $ae$ are either a variable, or a literal \texttt{lambda} term, which
+both can be evaluated in a single step.
+
+The abstract syntax tree can be straightforwardly described in Scala as follows:
+
+\begin{verbatim}
+sealed trait Expr
+case class Binding(x: String, e: Expr)
+
+case class Var(x: String) extends Expr
+case class App(e1: Expr, e2: Expr) extends Expr
+case class Lam(x: String, body: Expr) extends Expr
+case class Let(x: String, e: App, body: Expr) extends Expr
+case class Letrec(bds: List[Binding], body: Expr) extends Expr
+\end{verbatim}
 
 \subsection{CESK Machine}
+
+\textit{C}ontrol \textit{E}nvionment \textit{S}tore \textit{K}ontinuation (CESK) machine is an
+abstract machine for describing semantics of and evaluating $\lambda$-calculus. (TODO: cite)
+CESK machine models program execution as state transitions in small-step fashion. As its name suggested,
+a machine state has four components: 1) Control is the expression currently being evaluated.
+2) Environment is the a map that contains the address of a variable that in the lexical
+scope of control.
+3) Store models the heap of a program, and it is also a map but maps from addresses to values.
+The addresses are infinite numbers starting from 0.
+In our toy language, the only catagory of value is closure, i.e., a function paired with an environment.
+4) Continuation represents the program stack.
+
+We show the Scala representations for the components of CESK machine.
+
+\begin{verbatim}
+type Addr = Int
+type Env = Map[String, Addr]
+type Store = Map[Addr, Storable]
+
+abstract class Storable
+case class Clos(v: Lam, env: Env) extends Storable
+
+case class Frame(x: String, e: Expr, env: Env)
+type Kont = List[Frame]
+
+case class State(e: Expr, env: Env, store: Store, k: Kont)
+\end{verbatim}
+
+It is worth noting that the continuation class \texttt{Kont} is defined as a list of frames.
+The frame class \texttt{Frame} stores the information of call-site, i.e., the information that
+used to resume the interrupted computation.
+A \texttt{Frame} constitutes of a variable name \texttt{x} to be bind later, a control expression
+that resumes to, and its environment.
+
+Before go into describing how the machine evaluates expressions, we firstly define several helper functions.
+As we mentioned in section \ref{anfsyntax}, the atomic expressions are either a variable, or
+a literal \texttt{lambda} term. So the helper function \texttt{atomicEval} handles these two
+cases and evaluates atomic expressions to closures in a straightforward way.
+The \texttt{alloc} function generates a fresh address, and always allocates an unique integer under the domain
+of store.
+The \texttt{isAtomic} function is used as a predicate of whether the expression is atomic.
+
+\begin{verbatim}
+def atomicEval(e: Expr, env: Env, store: Store): Storable = e match {
+  case Var(x) => store(env(x))
+  case lam@Lam(x, body) => Clos(lam, env)
+}
+def alloc(store: Store): Addr = store.keys.size + 1
+def isAtomic(e: Expr): Boolean = e.isInstanceOf[Var] || e.isInstanceOf[Lam]
+\end{verbatim}
+
+Now, we can faithfully describe the state transition function \texttt{step},
+that given a machine state, function \texttt{step} determines its successor state.
+
+\begin{verbatim}
+def step(s: State): State = s match {
+  case State(Let(x, App(f, ae), e), env, store, k) if isAtomic(ae) =>
+    val Clos(Lam(v, body), env_c) = atomicEval(f, env, store)
+    val addr = alloc(store)
+    val new_env = env_c + (v -> addr)
+    val new_store = store + (addr -> atomicEval(ae, env, store))
+    val frame = Frame(x, e, env)
+    State(body, new_env, new_store, frame::k)
+  case State(ae, env, store, k) if isAtomic(ae) =>
+    val Frame(x, e, env_k)::ks = k
+    val addr = alloc(store)
+    val new_env = env_k + (x -> addr)
+    val new_store = store + (addr -> atomicEval(ae, env, store))
+    State(e, new_env, new_store, ks)
+}
+\end{verbatim}
+
+The first case is that the control of current state is a \texttt{Let} expression,
+and its right-hand side is an function application.
+By calling function \texttt{atomicEval}, we obtain the closure that callee \texttt{f} stands for.
+The successor state's control then transfers to the \texttt{body} expression of closure,
+with an updated environment and an update store. The new environment is extended
+from closure's environment and mapped \texttt{v} to a fresh address \texttt{addr}.
+The new store also extends with \texttt{addr} mapping to the value of \texttt{ae},
+which is evaluated from \texttt{atomicEval}.
+In the meanwhile, a new frame \texttt{frame} is pushed onto the stack \texttt{k}.
+This \texttt{frame} contains the left-hand side variable name of \texttt{Let},
+the body expression, and the lexical environment of body expression.
+
+The second case for \texttt{step} is the control is transferred to an atomic expression.
+We extract the top frame of continuations firstly.
+The control (that is an atomic expression) of current state is the evaluated term
+that being bind to the variable \texttt{x} from the top frame.
+The environment and store are updated with \texttt{x} mapping to the closure value of \texttt{ae}.
+Then the successor state is transferred to expression \texttt{e} from the top frame,
+which is the body of a \texttt{Let} expression, with the updated environment, store, and
+the rest of stack \texttt{ks}.
+
+To run the program,
+
+\begin{verbatim}
+def drive(s: State): State = s match {
+  case State(ae, _, _, Nil) if isAtomic(ae) => s
+  case _ => drive(step(s))
+}
+
+def inject(e: Expr): State = State(e, Map(), Map(), Nil)
+def eval(e: Expr): State = drive(inject(e))
+\end{verbatim}
 

### 2018-02-23 paper.tex
update

@@
 
-theoretical and practical
+\begin{itemize}
+\item The main contribution of this paper is we fill the gap
+  between small-step AAM and big-step ADI by developing a series of
+  systematic transformations. The correspondence between abstract
+  machines and interpreters exists not only in concrete semantics artifact,
+  but also in abstract semantics artifact.
+  Galois connection?
 
-\begin{itemize}
-  \item We fill the gap between small-step AAM and big-step ADI 
-    by developing a series of systematic transformations from 
-    small-step AAM to big-step ADI, and also the reversed 
-    direction is also possible.
+\item We present an abstract interpreter obtained from refunctionalizing
+  an small-step abstracting abstract machine. 
+  This abstract interpreter naturally has pushdown control-flow property.
+  We also analyze its termination, soundness and complexity.
 
-  \item We present an abstract interpreter written in continuation-passing
-    style that is obtained from refunctionalization of a small-step AAM.
-    This abstract interpreter naturally has pushdown control-flow property, 
-    and we show how precise call/return match is obtained by 
-    representing evaluation context as a continuation function, 
-    as well as how it is lost by representing continuation as a 
-    data type in the defining language.
-    As a form of abstract interpreter, We also analyze its termination,
-    soundness and complexity property.
+\item Further, we show that defunctionalization and refunctionalization
+  of abstract interpreters play important roles for call stack of analyzed language.
+  Given this insight, we reveal that why small-step AAM has no pushdown
+  control-flow property (if without extra effort), as well as explain
+  that why ADI naturally extends pushdown control-flow property from its
+  defining language.
 
-  \item We show that the correspondence not only exists between 
-    concrete abstract machines and evaluator, and also exists 
-    for abstracting abstract machines and abstract evaluators. 
-    Galois connection?
 \end{itemize}
@@
 analyze and a concrete CESK machine as operational semantics for it.
-Then we review several recent works in control flow analysis for
+Then we review some recent works in control-flow analysis for
 higher-order languages:
 1) abstracting abstract machines (AAM) as an abstract interpretation
-counterpart of CESK machines,
-2) definitional abstract interpreter written in a monadically-parameterized
-style, and
-3) precise call/return match problem in control-flow analysis and its existing solutions.
-The abstracting abstract machines form the basis of our transformations, and
-monadic abstract interpreter is the end point. In the meanwhile, the precise
-call/return match is obtained during transformation.
+counterpart of CESK machines, and
+2) the precise call/return match problem in control-flow analysis and its existing solutions.
+We will show in section 3 that the abstracting abstract machines form the basis of our transformations,
+and call/return match is obtained during transformation.
 
@@
 
-To begin, we now present the syntax of a call-by-value $\lambda$-calculus language in
+To begin, we now present the concrete syntax of a call-by-value $\lambda$-calculus language in
 administrative normal form (ANF) \cite{flanagan1993essence}.
@@
 
-The abstract syntax tree can be straightforwardly described in Scala as follows:
+The abstract syntax can be straightforwardly described in Scala as follows:
 
@@
 \subsection{CESK Machine}
+
+\subsubsection{Machine Components}
 
@@
 The frame class \texttt{Frame} stores the information of call-site, i.e., the information that
-used to resume the interrupted computation.
+can be used to resume the interrupted computation.
 A \texttt{Frame} constitutes of a variable name \texttt{x} to be bind later, a control expression
@@
 
+\subsubsection{Single-step Evaluation}
 Before go into describing how the machine evaluates expressions, we firstly define several helper functions.
@@
 
-To run the program,
+\subsubsection{Valuation}
+
+To run the program, we use function \texttt{inject} to construct an initial machine
+state given an expression. Then the \texttt{drive} function is used to evaluate
+to a final state by iteratively apply \texttt{step} on current state until we reach a state
+that the control is an atomic expression and the continuation stack is empty.
+And of course, we can extract the value from the final state at last.
 
@@
 
-\subsection{Abstracting Abstract Machine}
+\subsection{Abstracting Abstract Machine} \label{aam}
 
-\subsection{Monadic Abstract Interpreter}
+store allocated continuation
+binding store, continuation store
+store maps from addresses to a set of values
+store update is a join
+the step function is non-deterministic
+
+\begin{verbatim}
+case class Store[K,V](map: Map[K, Set[V]]) {
+  def apply(addr: K): List[V] = map(addr).toList
+  def update(addr: K, d: Set[V]): Store[K,V] = {
+    val oldd = map.getOrElse(addr, Set())
+    Store[K, V](map ++ Map(addr -> (d ++ oldd)))
+  }
+  def update(addr: K, sd: V): Store[K,V] = update(addr, Set(sd))
+}
+
+type Time = List[Expr]
+
+case class BAddr(x: String, time: Time)
+type Env = Map[String, BAddr]
+
+abstract class Storable 
+case class Clos(v: Lam, env: Env) extends Storable
+type BStore = Store[BAddr, Storable]
+
+abstract class KAddr
+case object Halt extends KAddr
+case class ContAddr(tgt: Expr, time: Time) extends KAddr
+
+case class Frame(x: String, e: Expr, env: Env)
+case class Cont(frame: Frame, kaddr: KAddr)
+type KStore = Store[KAddr, Cont]
+
+case class State(e: Expr, env: Env, bstore: BStore, 
+                 kstore: KStore, k: KAddr, time: Time)
+\end{verbatim}
+
+\begin{verbatim}
+def allocBind(x: String, time: Time): BAddr = BAddr(x, time)
+def allocKont(tgtExpr: Expr, tgtEnv: Env, tgtStore: BStore, time: Time): KAddr = 
+  ContAddr(tgtExpr, time)
+def k: Int = 0
+def tick(s: State): Time = (s.e::s.time).take(k)
+
+def inject(e: Expr): State = 
+  State(e, Map(), 
+        Store[BAddr, Storable](Map()), 
+        Store[KAddr, Cont](Map(Halt -> Set())), 
+        Halt, List())
+
+def atomicEval(e: Expr, env: Env, bstore: BStore): Set[Storable] = e match {
+  case Var(x) => bstore(env(x)).toSet
+  case lam@Lam(x, body) => Set(Clos(lam, env))
+}
+\end{verbatim}
+
+\begin{verbatim}
+def step(s: State): List[State] = {
+  val newTime = tick(s)
+  s match {
+    case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) if isAtomic(ae)=>
+      for (Clos(Lam(v, body), c_env) <- atomicEval(f, env, bstore).toList) yield {
+        val baddr = allocBind(v, newTime)
+        val newEnv = c_env + (v -> baddr)
+        val newBStore = bstore.update(baddr, atomicEval(ae, env, bstore))
+        val newKAddr = allocKont(body, c_env, newBStore, newTime)                                                          
+        val newKStore = kstore.update(newKAddr, Cont(Frame(x, e, env), kaddr))
+        State(body, newEnv, newBStore, newKStore, newKAddr, newTime)
+      }
+    case State(ae, env, bstore, kstore, kaddr, time) if isAtomic(ae)=>
+      for (Cont(Frame(x, e, f_env), f_kaddr) <- kstore(kaddr).toList) yield {
+        val baddr = allocBind(x, newTime)
+        val newEnv = f_env + (x -> baddr)
+        val newStore = bstore.update(baddr, atomicEval(ae, env, bstore))
+        State(e, newEnv, newStore, kstore, f_kaddr, newTime)
+      }
+  }
+}
+\end{verbatim}
+
+Comparing with concrete CESK machine, the
+\texttt{drive} function performs a collecting semantics rather than valuation
+semantics. That is, for the purpose of analyzing a program, \texttt{drive}
+collects all the intermediate machine states as program is abstractly executing.
+This is a variant of work list algorithm that always applies \texttt{step}
+function to the first element of \texttt{todo} and adds it to \texttt{seen}
+if there is an unexplored one. If the work list is empty, the function just
+returns the set of explored states up to now.
+
+\begin{verbatim}
+def drive(todo: List[State], seen: Set[State]): Set[State] = todo match {
+  case Nil => seen
+  case hd::tl if seen.contains(hd) => drive(tl, seen)
+  case hd::tl => drive(step(hd).toList ++ tl, seen + hd)
+}
+
+def analyze(e: Expr): Set[State] = drive(List(inject(e)), Set())
+\end{verbatim}
+
+%\subsection{Monadic Abstract Interpreter}
+%Maybe not here?
 
@@
 
+describe problem, show some examples, why it is important
+
+In this section, we firstly describe a naive and uncomputable solution to the
+problem; then briefly review several recent research works that enable
+computable and precise call/return match. Even though the first solution is
+uncomputable, but as we will see in section 3, it is the starting point of our
+transformation.
+
 \subsubsection{AAM with Unbound Stack}
 
+uncomputable solutions
+
+\begin{verbatim}
+case class State(e: Expr, env: Env, bstore: BStore, konts: List[Frame], time: Time)
+
+def step(s: State): List[State] = {
+  val newTime = tick(s)
+  s match {
+    case State(Let(x, App(f, ae), e), env, bstore, konts, time) =>
+      for (Clos(Lam(v, body), c_env) <- atomicEval(f, env, bstore).toList) yield {
+        val frame = Frame(x, e, env)
+        val baddr = allocBind(v, newTime)
+        val newEnv = c_env + (v -> baddr)
+        val newStore = bstore.update(baddr, atomicEval(ae, env, bstore))
+        State(body, newEnv, newStore, frame::konts, newTime)
+      }
+    case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
+      konts match {
+        case Nil => List()
+        case Frame(x, e, f_env)::konts =>
+          val baddr = allocBind(x, newTime)
+          val newEnv = f_env + (x -> baddr)
+          val newStore = bstore.update(baddr, atomicEval(ae, env, bstore))
+          List(State(e, newEnv, newStore, konts, newTime))
+      }
+  }
+}
+
+\end{verbatim}
+
 \subsubsection{Computable Solutions}
-P4F ANF, not works on direct style,PDCFA,AAC
+
+Recent years, there are great efforts \cite{vardoulakis2010cfa2,
+  earl2012introspective, gilray2016pushdown}
+to achieve precise call/return match based
+on small-step abstracting abstract machines which is we described in section \ref{aam}.
+
+go over existing computable solutions P4F ANF, not works on direct style,PDCFA,AAC
 
@@
 
-\section{Detour: Shiver's \textit{k}-CFA}
-
 \section{Perspective}
@@
 
+specify higher-order definitional interpreters using first-order means (cite)
+
 \subsection{Refunctionalization}
+
+\section{Detour: Shiver's \textit{k}-CFA}
 

### 2018-02-24 paper.tex
misc
update

@@
 The transformations we will show in the rest of this paper
-also work on abstract machines for plain $\lambda$-calculus.
+also work on abstract machines for plain $\lambda$-calculus languages.
 
@@
 \textit{C}ontrol \textit{E}nvionment \textit{S}tore \textit{K}ontinuation (CESK) machine is an
-abstract machine for describing semantics of and evaluating $\lambda$-calculus. (TODO: cite)
+abstract machine for describing semantics of and evaluating $\lambda$-calculus \cite{felleisen1987calculus}.
 CESK machine models program execution as state transitions in small-step fashion. As its name suggested,
@@
 \subsection{Abstracting Abstract Machine} \label{aam}
+Abstracting abstract machine is a systematic methodology that derives abstract interpreters
+from a concrete abstract machine. TODO
 
-store allocated continuation
-binding store, continuation store
-store maps from addresses to a set of values
-store update is a join
-the step function is non-deterministic
+There are several notable differences of store between AAM and concrete CESK machine.
+In AAM, the store maps addresses to sets of values, so it stores all possible values
+for a particular address. Thus the dereference an address becomes non-deterministic.
+Also, the store performs a join rather than overwriting when updating elements of the store.
+Third, the continuations are allocated on store instead of formed a linked list.
+
+For clarity, we divide the store into binding store \texttt{BStore} and continuation store \texttt{KStore}. 
+The binding store maps addresses of binding to sets of closure values, and continuation store maps
+continuation addresses to sets of continuations.
+We define a generic class \texttt{Store[K,V]} that performs join when updating elements
+in a store. By parametering \texttt{Store[K,V]} with \texttt{[BAddr, Storable]} and 
+\texttt{[KAddr, Cont]}, we obtain \texttt{BStore} and \texttt{KStore}, perspectively.
+
+Note that the store is monotonic. TODO
 
@@
 
-type Time = List[Expr]
+type BStore = Store[BAddr, Storable]
+type KStore = Store[KAddr, Cont]
+\end{verbatim}
 
+As a consequence, the components of a state is also changed: 
+the store is divided into binding store and continuation store;
+the continuation stack becomes an address that maps to a set of continuation.
+By dereferencing this address in continuation store, we can retrieve the 
+actually transfer of controls.
+
+\begin{verbatim}
+case class State(e: Expr, env: Env, bstore: BStore, 
+                 kstore: KStore, k: KAddr, time: Time)
+\end{verbatim}
+
+The definitions of closure \texttt{Clos}, stack frame \texttt{Frame}, and 
+environment \texttt{Env} remain the same as concrete CESK machine.
+
+\begin{verbatim}
 case class BAddr(x: String, time: Time)
@@
 case class Clos(v: Lam, env: Env) extends Storable
-type BStore = Store[BAddr, Storable]
 
@@
 case class Cont(frame: Frame, kaddr: KAddr)
-type KStore = Store[KAddr, Cont]
+\end{verbatim}
 
-case class State(e: Expr, env: Env, bstore: BStore, 
-                 kstore: KStore, k: KAddr, time: Time)
-\end{verbatim}
+To allocate addresses for bindings or continuations, we use \texttt{allocBind}
+and \texttt{allocKont}.
 
@@
   ContAddr(tgtExpr, time)
+type Time = List[Expr]
 def k: Int = 0
@@
 def step(s: State): List[State] = {
-  val newTime = tick(s)
+  val new_time = tick(s)
   s match {
     case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) if isAtomic(ae)=>
-      for (Clos(Lam(v, body), c_env) <- atomicEval(f, env, bstore).toList) yield {
-        val baddr = allocBind(v, newTime)
-        val newEnv = c_env + (v -> baddr)
+      for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
+        val baddr = allocBind(v, new_time)
+        val new_env = env_c + (v -> baddr)
         val newBStore = bstore.update(baddr, atomicEval(ae, env, bstore))
-        val newKAddr = allocKont(body, c_env, newBStore, newTime)                                                          
+        val newKAddr = allocKont(body, env_c, newBStore, new_time)                                                          
         val newKStore = kstore.update(newKAddr, Cont(Frame(x, e, env), kaddr))
-        State(body, newEnv, newBStore, newKStore, newKAddr, newTime)
+        State(body, new_env, newBStore, newKStore, newKAddr, new_time)
       }
@@
       for (Cont(Frame(x, e, f_env), f_kaddr) <- kstore(kaddr).toList) yield {
-        val baddr = allocBind(x, newTime)
-        val newEnv = f_env + (x -> baddr)
-        val newStore = bstore.update(baddr, atomicEval(ae, env, bstore))
-        State(e, newEnv, newStore, kstore, f_kaddr, newTime)
+        val baddr = allocBind(x, new_time)
+        val new_env = f_env + (x -> baddr)
+        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+        State(e, new_env, new_store, kstore, f_kaddr, new_time)
       }
@@
 collects all the intermediate machine states as program is abstractly executing.
-This is a variant of work list algorithm that always applies \texttt{step}
-function to the first element of \texttt{todo} and adds it to \texttt{seen}
+The following code shows a variant of work list algorithm that always applies \texttt{step}
+function to the first element of \texttt{todo} and adds it to \texttt{seen} set
 if there is an unexplored one. If the work list is empty, the function just
@@
 def step(s: State): List[State] = {
-  val newTime = tick(s)
+  val new_time = tick(s)
   s match {
     case State(Let(x, App(f, ae), e), env, bstore, konts, time) =>
-      for (Clos(Lam(v, body), c_env) <- atomicEval(f, env, bstore).toList) yield {
+      for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
         val frame = Frame(x, e, env)
-        val baddr = allocBind(v, newTime)
-        val newEnv = c_env + (v -> baddr)
-        val newStore = bstore.update(baddr, atomicEval(ae, env, bstore))
-        State(body, newEnv, newStore, frame::konts, newTime)
+        val baddr = allocBind(v, new_time)
+        val new_env = env_c + (v -> baddr)
+        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+        State(body, new_env, new_store, frame::konts, new_time)
       }
@@
         case Frame(x, e, f_env)::konts =>
-          val baddr = allocBind(x, newTime)
-          val newEnv = f_env + (x -> baddr)
-          val newStore = bstore.update(baddr, atomicEval(ae, env, bstore))
-          List(State(e, newEnv, newStore, konts, newTime))
+          val baddr = allocBind(x, new_time)
+          val new_env = f_env + (x -> baddr)
+          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+          List(State(e, new_env, new_store, konts, new_time))
       }

### 2018-02-24 paper.tex
misc

--- 2018-02-23 paper.tex
@@
 
+the essence of abstracting defintional interpreter.
+
 \begin{itemize}
@@
   an small-step abstracting abstract machine. 
-  This abstract interpreter naturally has pushdown control-flow property.
+  This abstract interpreter naturally has pushdown control-flow property
+  since we represent the evluation context of analyzed language using
+  evluation context of the meta-language.
   We also analyze its termination, soundness and complexity.
@@
 
-Tranditionally, continuation-passing style is a popular intermediate representation
+Tranditionally, continuation-passing style (CPS) is a popular intermediate representation
 for analyzing functional programs because of it exposes control transfer explicitly

### 2018-02-24 paper.tex
update

@@
 and simplifies the analysis \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
-Here, we choose to use a direct-style intermediate representation as target for
-clarity but without losing simplicty and generality.
+Here, we choose to use administrative normal form (ANF) which is a direct-style intermediate representation 
+as target for clarity but without losing simplicty and generality.
 The transformations we will show in the rest of this paper
-also work on abstract machines for plain $\lambda$-calculus languages.
+also work on abstract machines for plain, direct-style $\lambda$-calculus languages.
 
-To begin, we now present the concrete syntax of a call-by-value $\lambda$-calculus language in
-administrative normal form (ANF) \cite{flanagan1993essence}.
+To begin, we now present the concrete syntax of a call-by-value $\lambda$-calculus language 
+in ANF \cite{flanagan1993essence}.
 
@@
 abstract machine for describing semantics of and evaluating $\lambda$-calculus \cite{felleisen1987calculus}.
-CESK machine models program execution as state transitions in small-step fashion. As its name suggested,
+CESK machine models program execution as state transitions in a small-step fashion. As its name suggested,
 a machine state has four components: 1) Control is the expression currently being evaluated.
@@
 A \texttt{Frame} constitutes of a variable name \texttt{x} to be bind later, a control expression
-that resumes to, and its environment.
+that resumes to, and its environment. (TODO: mention that frame list vs KArg/KFun)
 
@@
 In the meanwhile, a new frame \texttt{frame} is pushed onto the stack \texttt{k}.
-This \texttt{frame} contains the left-hand side variable name of \texttt{Let},
-the body expression, and the lexical environment of body expression.
+This \texttt{frame} contains the variable name \texttt{x} at left-hand side position of \texttt{Let},
+the body expression of \texttt{Let}, and the lexical environment of the body expression.
 
@@
 To run the program, we use function \texttt{inject} to construct an initial machine
-state given an expression. Then the \texttt{drive} function is used to evaluate
+state given an expression \texttt{e}.
+Then the \texttt{drive} function is used to evaluate
 to a final state by iteratively apply \texttt{step} on current state until we reach a state
@@
 \subsection{Abstracting Abstract Machine} \label{aam}
-Abstracting abstract machine is a systematic methodology that derives abstract interpreters
-from a concrete abstract machine. TODO
+Abstracting abstract machine is a systematic methodology that derives sound abstract interpreters
+from a concrete abstract machine \cite{van2012systematic, van2010abstracting}.
 
@@
 def atomicEval(e: Expr, env: Env, bstore: BStore): Set[Storable] = e match {
-  case Var(x) => bstore(env(x)).toSet
+  case Var(x) => bstore(env(x))
   case lam@Lam(x, body) => Set(Clos(lam, env))
@@
   s match {
-    case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) if isAtomic(ae)=>
-      for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
+    case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) 
+      if isAtomic(ae) =>
+      val closures = atomicEval(f, env, bstore).toList
+      for (Clos(Lam(v, body), env_c) <- closures) yield {
         val baddr = allocBind(v, new_time)
@@
       }
-    case State(ae, env, bstore, kstore, kaddr, time) if isAtomic(ae)=>
+    case State(ae, env, bstore, kstore, kaddr, time) 
+      if isAtomic(ae) =>
       for (Cont(Frame(x, e, f_env), f_kaddr) <- kstore(kaddr).toList) yield {
@@
 semantics. That is, for the purpose of analyzing a program, \texttt{drive}
-collects all the intermediate machine states as program is abstractly executing.
+collects all the intermediate machine states as the program is abstractly executing.
 The following code shows a variant of work list algorithm that always applies \texttt{step}
@@
 
-\subsection{Call/Return Match in Control-Flow Analysis}
+\subsection{Pushdown Control-Flow Analysis}
 
-describe problem, show some examples, why it is important
+Pushdown control-flow is a property in analysis that precisely models
+the runtime call-stack of the analyzed program. 
+A pushdown control-flow analysis provides as exact as runtime return-flow 
+when analyzing the program, but traditional control-flow analysis collapse 
+the state space using finite space abstraction thus cause imprecise stack modeling.
+
+To see the problem, we can consider the following example:
+
+\begin{verbatim}
+(let ([x (id 1)])
+  (let ([y (id 2)])
+    x))
+\end{verbatim}
+
+return flow of \texttt{id} is polluted, and returns to both \texttt{x} and
+\texttt{y}. TODO: explain more
 
@@
 }
-
 \end{verbatim}
@@
 
-In another direction, we can transform the refunctionalized AAM to direct style, but explictly use 
+We can transform the refunctionalized AAM to direct style, but explictly use 
 side effects such as assignment and mutation to handle collecting semantics and non-determinism.
@@
 \section{Refunctionalized AAM}
+
+\subsection{Store Widening}
+
+\subsection{Higher-Order Controls?}
+
+TODO: analyze programs with call/cc, shift, reset.
 

### 2018-02-25 paper.tex
update

@@
 %% Title information
-\title[Short Title]{Pushdown Control-Flow Analysis via Refunctionalization}         %% [Short Title] is optional;
+\title[Short Title]{Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
   This abstract interpreter naturally has pushdown control-flow property
-  since we represent the evluation context of analyzed language using
-  evluation context of the meta-language.
+  since we represent the evluation contexts of analyzed language using
+  continuations of the meta-language, and therefore sequentialize the
+  analyzing.
   We also analyze its termination, soundness and complexity.
@@
 We use Scala language to demonstrate the idea and each step of transformations.
+There are two main reasons to do so:
+1) The code does not diminish the accuracy of the material
+than formal and mathematical notations, 
+which are heavily used in other papers in static analysis or semantics.
+The Scala code in this paper can be easily back-translated into formal notations.
+2) The code is also executable with only few changes, which make it particularly fit for 
+presenting topic: syntactical transformations of programs.
 
@@
 
-\subsection{CESK Machine}
+\subsection{CESK Machine} \label{cesk}
 
@@
 \subsection{Abstracting Abstract Machine} \label{aam}
-Abstracting abstract machine is a systematic methodology that derives sound abstract interpreters
-from a concrete abstract machine \cite{van2012systematic, van2010abstracting}.
+Abstracting abstract machine (AAM) is a systematic methodology that derives sound 
+abstract interpreters for higher-order functional languages from concrete 
+abstract machines \cite{van2012systematic, van2010abstracting}. 
+An abstracting abstract machine implements computable abstract semantics that 
+approximates the runtime behaviors of programs.
+Since the state space of concrete execution is possibily infinite,
+so to analyze programs, the key insight of AAM approach is to allocate both bindings
+and continuations on the store, and then bound the addresses space to be finite.
+Thus the approximated machine-state space is finite, and computable.
 
-There are several notable differences of store between AAM and concrete CESK machine.
+In this section, we derive the abstracting abstract machine from concrete
+CESK machines, and also show that how to instantiate
+useful $k$-call-sensitive analysis.
+
+\subsubsection{Machine Components}
+
+Similar with CESK machines, the machine state of the AAM has a control expression,
+an environment, a store and continuation, and in addition with a time-stamp.
+But there are several notable differences of store between AAMs and concrete CESK machines.
 In AAM, the store maps addresses to sets of values, so it stores all possible values
@@
 Also, the store performs a join rather than overwriting when updating elements of the store.
-Third, the continuations are allocated on store instead of formed a linked list.
+Third, the continuations are likewise allocated on store instead of formed a linked list,
+and the continuation in a state then turns into a continuation address.
 
 For clarity, we divide the store into binding store \texttt{BStore} and continuation store \texttt{KStore}. 
-The binding store maps addresses of binding to sets of closure values, and continuation store maps
+The binding store maps binding addresses to sets of closure values, and continuation store maps
 continuation addresses to sets of continuations.
 We define a generic class \texttt{Store[K,V]} that performs join when updating elements
-in a store. By parametering \texttt{Store[K,V]} with \texttt{[BAddr, Storable]} and 
+in a store. By parameterizing \texttt{Store[K,V]} with \texttt{[BAddr, Storable]} and 
 \texttt{[KAddr, Cont]}, we obtain \texttt{BStore} and \texttt{KStore}, perspectively.
@@
 }
-
 type BStore = Store[BAddr, Storable]
@@
 
+The co-domain of binding stores is \texttt{Storable} which is the same
+as we defined for CESK machines.
+The co-domain of continuation stores \texttt{Cont} is comprised of 
+a frame \texttt{Frame} and a continuation address \texttt{KAddr}.
+To mimic the runtime call stack, \texttt{KAddr} plays the role of
+representing the remaining stack frames.
+TODO: nondet of cont?
+
+\begin{verbatim}
+abstract class Storable 
+case class Clos(v: Lam, env: Env) extends Storable
+
+case class Frame(x: String, e: Expr, env: Env)
+case class Cont(frame: Frame, kaddr: KAddr)
+\end{verbatim}
+
 As a consequence, the components of a state is also changed: 
 the store is divided into binding store and continuation store;
-the continuation stack becomes an address that maps to a set of continuation.
+the continuation stack becomes an address.
 By dereferencing this address in continuation store, we can retrieve the 
-actually transfer of controls.
+actual transfer of controls. The definition of environment \texttt{Env}
+remains the same.
 
@@
 
-The definitions of closure \texttt{Clos}, stack frame \texttt{Frame}, and 
-environment \texttt{Env} remain the same as concrete CESK machine.
+Up to now, we did not describe much about the time-stamp \texttt{Time}
+os states and addresses of stores.
+But they are key ingredients to achieve analysis with different sensitivities,
+as well as a finite state space analysis.
+To effectively approximate the runtime behavior, a finite 
+program contour \texttt{time} that encodes the program execution history is introduced.
+Here, we use a list of execution contexts (expressions) to reprent that, and
+as we will see in section \ref{kcfainst}, by using different \texttt{tick}
+function on time we may obtain a family of analyses.
+
+\begin{verbatim}
+type Time = List[Expr]
+\end{verbatim}
+
+Address.
 
@@
 case class BAddr(x: String, time: Time)
-type Env = Map[String, BAddr]
-
-abstract class Storable 
-case class Clos(v: Lam, env: Env) extends Storable
 
@@
 case class ContAddr(tgt: Expr, time: Time) extends KAddr
+\end{verbatim}
 
-case class Frame(x: String, e: Expr, env: Env)
-case class Cont(frame: Frame, kaddr: KAddr)
-\end{verbatim}
+\subsubsection{Single-step Evaluation}
 
@@
 def allocBind(x: String, time: Time): BAddr = BAddr(x, time)
-def allocKont(tgtExpr: Expr, tgtEnv: Env, tgtStore: BStore, time: Time): KAddr = 
-  ContAddr(tgtExpr, time)
-type Time = List[Expr]
-def k: Int = 0
-def tick(s: State): Time = (s.e::s.time).take(k)
-
-def inject(e: Expr): State = 
-  State(e, Map(), 
-        Store[BAddr, Storable](Map()), 
-        Store[KAddr, Cont](Map(Halt -> Set())), 
-        Halt, List())
+def allocKont(tgtExpr: Expr, tgtEnv: Env, 
+              tgtStore: BStore, time: Time): KAddr = ContAddr(tgtExpr, time)
 
@@
         val new_env = env_c + (v -> baddr)
-        val newBStore = bstore.update(baddr, atomicEval(ae, env, bstore))
-        val newKAddr = allocKont(body, env_c, newBStore, new_time)
-        val newKStore = kstore.update(newKAddr, Cont(Frame(x, e, env), kaddr))
-        State(body, new_env, newBStore, newKStore, newKAddr, new_time)
+        val new_bstore = bstore.update(baddr, atomicEval(ae, env, bstore))
+        val new_kaddr = allocKont(body, env_c, new_bstore, new_time)
+        val new_kstore = kstore.update(new_kaddr, Cont(Frame(x, e, env), kaddr))
+        State(body, new_env, new_bstore, new_kstore, new_kaddr, new_time)
       }
@@
       if isAtomic(ae) =>
-      for (Cont(Frame(x, e, f_env), f_kaddr) <- kstore(kaddr).toList) yield {
+      val conts = kstore(kaddr).toList
+      for (Cont(Frame(x, e, env_f), f_kaddr) <- conts) yield {
         val baddr = allocBind(x, new_time)
-        val new_env = f_env + (x -> baddr)
+        val new_env = env_f + (x -> baddr)
         val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
@@
 \end{verbatim}
+
+\subsubsection{$k$-call-sensitive Instantiation} \label{kcfainst}
+
+\begin{verbatim}
+def k: Int = 0
+def tick(s: State): Time = (s.e::s.time).take(k)
+\end{verbatim}
+
+\subsubsection{Collecting Semantics}
 
@@
 \begin{verbatim}
+def inject(e: Expr): State = 
+  State(e, Map(), 
+        Store[BAddr, Storable](Map()), 
+        Store[KAddr, Cont](Map(Halt -> Set())), 
+        Halt, List())
+
 def drive(todo: List[State], seen: Set[State]): Set[State] = todo match {
@@
   s match {
-    case State(Let(x, App(f, ae), e), env, bstore, konts, time) =>
+    case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
       for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
@@
         case Nil => List()
-        case Frame(x, e, f_env)::konts =>
+        case Frame(x, e, env_f)::konts =>
           val baddr = allocBind(x, new_time)
-          val new_env = f_env + (x -> baddr)
+          val new_env = env_f + (x -> baddr)
           val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
@@
 
-Recent years, there are great efforts \cite{vardoulakis2010cfa2,
-  earl2012introspective, gilray2016pushdown}
+Recent years, there are great efforts \cite{vardoulakis2010cfa2, earl2012introspective, gilray2016pushdown}
 to achieve precise call/return match based
-on small-step abstracting abstract machines which is we described in section \ref{aam}.
+on small-step abstracting abstract machines which is what we described in section \ref{aam}.
 
@@
 
-\section{FROM AAM TO ADI}
+\section{From AAM to ADI}
 

### 2018-02-25 paper.tex
update

@@
   Abstracting abstract machines (AAM) is a systematic methodology 
-  for constructing abstract interpreter that derived from a concrete
-  small-step abstract machine. Recent progress applies the same idea 
-  on concrete definitional interpreter and obtains a big-step 
+  for constructing abstract interpreters that derived from concrete
+  small-step abstract machines. Recent progress applies the same idea 
+  on definitional interpreter and obtains a big-step 
   abstracting definitional interpreter (ADI) written in monadic style.
@@
 as target for clarity but without losing simplicty and generality.
+Although we just show the core calculus language, but it can be easily extended
+to support recursive bindings, conditionals, primitive types and operations on primitive types.
 The transformations we will show in the rest of this paper
 also work on abstract machines for plain, direct-style $\lambda$-calculus languages.
+TODO: add letrec?
 
@@
             | (let ([x (f ae)]) e)
-            | (letrec ([x ae] ...) e) (TODO: maybe not include letrec?)
 ae \in AExp ::= x
@@
 case class Let(x: String, e: App, body: Expr) extends Expr
-case class Letrec(bds: List[Binding], body: Expr) extends Expr
 \end{verbatim}
@@
 
-To run the program, we use function \texttt{inject} to construct an initial machine
-state given an expression \texttt{e}.
+To run the program, we first use function \texttt{inject} to construct an initial machine
+state given an expression \texttt{e}. The initial state contains an emtpy environment, 
+store and stack.
+\begin{verbatim}
+def inject(e: Expr): State = State(e, Map(), Map(), Nil)
+\end{verbatim}
+
 Then the \texttt{drive} function is used to evaluate
@@
 }
-
-def inject(e: Expr): State = State(e, Map(), Map(), Nil)
 def eval(e: Expr): State = drive(inject(e))
@@
 
-The co-domain of binding stores is \texttt{Storable} which is the same
+The co-domain of binding stores \texttt{Storable} is the same
 as we defined for CESK machines.
@@
 
-Up to now, we did not describe much about the time-stamp \texttt{Time}
-os states and addresses of stores.
-But they are key ingredients to achieve analysis with different sensitivities,
+\subsubsection{Allocating Addresses}
+Up to now, we did not describe much about the adresses in stores and the 
+time-stamp \texttt{Time}.
+But they are key ingredients to achieve analyses with different sensitivities,
 as well as a finite state space analysis.
@@
 
-Address.
+As we mentioned previously, the space of addresses is finite in AAM.
+Binding addresses are parameterized by variable names and the time that creates this binding.
+Continuation addresses \texttt{KAddr} has two variants: 
+1) \texttt{Halt} which corresponds to the empty stack, and
+2) \texttt{ContAddr} consists of the target expression of callee and the time that
+creates this address. (TODO: why target expression, allocation polyvariance)
 
@@
 
-\subsubsection{Single-step Evaluation}
-
-To allocate addresses for bindings or continuations, we use \texttt{allocBind}
-and \texttt{allocKont}.
+We use two helper functions \texttt{allocBind} and \texttt{allocKont} that will be used later
+to allocate binding addresses and continuation addresses.
 
@@
 def allocBind(x: String, time: Time): BAddr = BAddr(x, time)
-def allocKont(tgtExpr: Expr, tgtEnv: Env, 
-              tgtStore: BStore, time: Time): KAddr = ContAddr(tgtExpr, time)
+def allocKont(tgtExpr: Expr, time: Time): KAddr = ContAddr(tgtExpr, time)
+\end{verbatim}
+
+\subsubsection{Single-step Evaluation}
+
+atomicEval returns a set of storable values, i.e., closures.
+
+\begin{verbatim}
 
@@
         val new_bstore = bstore.update(baddr, atomicEval(ae, env, bstore))
-        val new_kaddr = allocKont(body, env_c, new_bstore, new_time)
+        val new_kaddr = allocKont(body, new_time)
         val new_kstore = kstore.update(new_kaddr, Cont(Frame(x, e, env), kaddr))
@@
 
-Comparing with concrete CESK machine, the
-\texttt{drive} function performs a collecting semantics rather than valuation
-semantics. That is, for the purpose of analyzing a program, \texttt{drive}
-collects all the intermediate machine states as the program is abstractly executing.
-The following code shows a variant of work list algorithm that always applies \texttt{step}
-function to the first element of \texttt{todo} and adds it to \texttt{seen} set
-if there is an unexplored one. If the work list is empty, the function just
-returns the set of explored states up to now.
-
+To run (analyze) a program, as CESK machine, we use \texttt{inject} to construct
+the initial state given a program. 
+Note that the inital continuation store has a special mapping that maps continuation address
+\texttt{Halt} to empty set of continuations.
+We also provide an empty program contour as inital time.
 \begin{verbatim}
@@
         Halt, List())
+\end{verbatim}
 
+However, comparing with concrete CESK machine, the
+\texttt{drive} function performs collecting semantics instead of valuation
+semantics. That is, for the purpose of analyzing programs, the function \texttt{drive}
+collects all the intermediate machine states as the program is abstractly executing.
+The following code shows a variant of work list algorithm that always applies \texttt{step}
+function to the first element \texttt{hd} of the work list \texttt{todo} if \texttt{hd}
+is unseen, then adds the result of \texttt{step} to the work list and also adds \texttt{hd}
+to the explored states set.
+If the work list is empty, the function just
+returns the set of explored states up to now.
+
+\begin{verbatim}
 def drive(todo: List[State], seen: Set[State]): Set[State] = todo match {
@@
 \end{verbatim}
+
+Finally, a user could use function \texttt{analyze} to obtain all possible states 
+for a given program.
 
@@
 
-In this section, we firstly describe a naive and uncomputable solution to the
+In this section, we firstly describe a simple but uncomputable solution to the
 problem; then briefly review several recent research works that enable
@@
 
-Recent years, there are great efforts \cite{vardoulakis2010cfa2, earl2012introspective, gilray2016pushdown}
-to achieve precise call/return match based
-on small-step abstracting abstract machines which is what we described in section \ref{aam}.
+Recent years, there are great efforts \cite{vardoulakis2010cfa2, earl2012introspective, 
+gilray2016pushdown, johnson2015abstracting}
+to achieve precise call/return match based on small-step AAM which is what we 
+described in section \ref{aam}.
 
-go over existing computable solutions P4F ANF, not works on direct style,PDCFA,AAC
+TODO: summarize CFA2, PDCFA, AAC, P4F
+
+On the other side, \citeauthor{darais2017abstracting} construct abstracting 
+definitional interpreters (ADI) that automatically inherit the pushdown control-flow 
+property from the defining language. TODO
+
+In the rest of this section, we take a closer look at the two state-of-art techniques
+to accomplish perfect call-stack precision.
+
+\textbf{Pushdown for Free.}
+\citeauthor{gilray2016pushdown} proposed a continuation address allocator for 
+small-step abstracting abstract machine that is both easy to implement and
+computational inexpensive.
+
+\begin{verbatim}
+case class P4FContAddr(tgt: Expr, tgtEnv: Env, time: Time) extends KAddr
+def allocKont(tgtExpr: Expr, tgtEnv: Env, time: Time): KAddr = 
+  P4FContAddr(tgtExpr, tgtEnv, time)
+\end{verbatim}
+
+\textbf{Abstracting Definitional Interpreters.}
 
@@
 
-\section{Detour: Shiver's \textit{k}-CFA}
+\section{Detour: Shiver's \textit{k}-CFA?}
 

### 2018-02-25 paper.tex
update

@@
   abstract interpreter.
-  Remarkably, follow the idea that evaluation contexts are defuntionalized 
+  Remarkably, follow the idea that evaluation contexts are defunctionalized 
   continuations, we reveal how precise call/return match in control-flow 
@@
 
-the essence of abstracting defintional interpreter.
+the essence of abstracting definitional interpreter.
 
@@
   This abstract interpreter naturally has pushdown control-flow property
-  since we represent the evluation contexts of analyzed language using
+  since we represent the evaluation contexts of analyzed language using
   continuations of the meta-language, and therefore sequentialize the
@@
 
-In this section, we frist describe the target language we will
+In this section, we first describe the target language we will
 analyze and a concrete CESK machine as operational semantics for it.
@@
 
-Tranditionally, continuation-passing style (CPS) is a popular intermediate representation
+Traditionally, continuation-passing style (CPS) is a popular intermediate representation
 for analyzing functional programs because of it exposes control transfer explicitly
@@
 Here, we choose to use administrative normal form (ANF) which is a direct-style intermediate representation 
-as target for clarity but without losing simplicty and generality.
+as target for clarity but without losing simplicity and generality.
 Although we just show the core calculus language, but it can be easily extended
@@
 The addresses are infinite numbers starting from 0.
-In our toy language, the only catagory of value is closure, i.e., a function paired with an environment.
+In our toy language, the only category of value is closure, i.e., a function paired with an environment.
 4) Continuation represents the program stack.
@@
 To run the program, we first use function \texttt{inject} to construct an initial machine
-state given an expression \texttt{e}. The initial state contains an emtpy environment, 
+state given an expression \texttt{e}. The initial state contains an empty environment, 
 store and stack.
@@
 approximates the runtime behaviors of programs.
-Since the state space of concrete execution is possibily infinite,
+Since the state space of concrete execution is possibly infinite,
 so to analyze programs, the key insight of AAM approach is to allocate both bindings
@@
 As we mentioned previously, the space of addresses is finite in AAM.
-Binding addresses are parameterized by variable names and the time that creates this binding.
+Binding addresses are parameterized by variable names and the time that creates this binding,
+which are both finite.
 Continuation addresses \texttt{KAddr} has two variants: 
 1) \texttt{Halt} which corresponds to the empty stack, and
-2) \texttt{ContAddr} consists of the target expression of callee and the time that
-creates this address. (TODO: why target expression, allocation polyvariance)
+2) \texttt{ContAddr} consists of the entry expression of callee and the time that
+creates this address, and they are also finite. 
+(TODO: why target expression, allocation polyvariance)
 
@@
 the initial state given a program. 
-Note that the inital continuation store has a special mapping that maps continuation address
+Note that the initial continuation store has a special mapping that maps continuation address
 \texttt{Halt} to empty set of continuations.
-We also provide an empty program contour as inital time.
+We also provide an empty program contour as initial time.
 \begin{verbatim}
@@
 
-To see the problem, we can consider the following example:
+To see the problem of spurious return-flows, we can consider the following example:
 
 \begin{verbatim}
+(let ([id (lambda (z) z)])
   (let ([x (id 1)])
     (let ([y (id 2)])
-    x))
+      x)))
 \end{verbatim}
 
-return flow of \texttt{id} is polluted, and returns to both \texttt{x} and
-\texttt{y}. TODO: explain more
+In traditional $k$-CFA algorithm and the AAM we showed in section \ref{aam}, 
+the return-flow of invocation \texttt{(id 2)} 
+is going to both call-sites \texttt{(id 1)} and \texttt{(id 2)}. Therefore, the
+returned value \texttt{2} for variable \texttt{y} is also propagated to 
+variable \texttt{x} and then imprecise analysis result is arose.
+This return-flow merging is inevitable even we increase the context-sensitivity.
+If we use the monovariant analysis 0CFA, the analyzing result would be \texttt{x} and 
+\texttt{y} point to set \texttt{\{1, 2\}}, because it not distinguished 
+\texttt{z} at different call-sites.
+Under $1$-CFA, the algorithm is able to distinguish that variable \texttt{z} of function
+\texttt{id} has two different values at two call-sites, so variable \texttt{y} 
+would not be polluted by \texttt{1}, but it is still helplessness for distinguish
+return-flows.
+TODO: do need more details to explain this example? (as in P4F paper)
 
@@
 
+A naive solution to is to use precise call stack modeling as we did in concrete
+CESK machine. Instead of allocating continuations on store and recording an 
+address of continuations, we
+
+In the definition of state \texttt{State}, continuation store disappears, and
+\texttt{konts} becomes a list of frames. Other components keep the same.
+
 \begin{verbatim}
 case class State(e: Expr, env: Env, bstore: BStore, konts: List[Frame], time: Time)
+\end{verbatim}
 
+the only non-determinism in our state transition function is when dereferencing
+the callee \texttt{f} from store at a call-site.
+
+\begin{verbatim}
 def step(s: State): List[State] = {
@@
 
-We can transform the refunctionalized AAM to direct style, but explictly use 
+We can transform the refunctionalized AAM to direct style, but explicitly use 
 side effects such as assignment and mutation to handle collecting semantics and non-determinism.

### 2018-02-26 paper.tex
update

@@
 
-uncomputable solutions
-
 A naive solution to is to use precise call stack modeling as we did in concrete
 CESK machine. Instead of allocating continuations on store and recording an 
-address of continuations, we
+address of continuations, we leave the call stack unabstracted.
 
@@
 
+Unfortunately, even though other components is finite, this AAM with unbounded stack is uncomputable. 
+Because the unbounded stack can grow to arbitrary deep and this implies that the state space is possibly 
+infinite, therefore the analysis may not terminate for all programs if we simply enumerating 
+reachable states.
+To see this, consider a program that has two mutually recursived functions:
+
+\begin{verbatim}
+(letrec ([f1 (lambda (x) 
+               (let ([x1 (f2 x)]) x1))]
+         [f2 (lambda (y)
+               (let ([y1 (f1 y)]) y1))])
+  (let ([z (f1 1)])
+    z))
+\end{verbatim}
+
+\texttt{f1} and \texttt{f2} call each others, and the stack will be interleaved
+pushing \texttt{f1} and \texttt{f2} onto the top of current stack, but not existing 
+two stack components are the same in the state space.
+
 \subsubsection{Computable Solutions}
@@
 Recent years, there are great efforts \cite{vardoulakis2010cfa2, earl2012introspective, 
-gilray2016pushdown, johnson2015abstracting}
-to achieve precise call/return match based on small-step AAM which is what we 
-described in section \ref{aam}.
+gilray2016pushdown, johnson2015abstracting} to achieve precise call/return 
+match based on small-step AAM which is what we described in section \ref{aam}.
 
-TODO: summarize CFA2, PDCFA, AAC, P4F
+CFA2 is the first solution that solves the return-flows problem\cite{vardoulakis2010cfa2}.
+But CFA2 has several drawbacks: it works only on continuation-passing style programs, 
+and only allows monovariant analysis, in addition with an exponential time complexity.
+TODO: summarize PDCFA, AAC, P4F
 
-On the other side, \citeauthor{darais2017abstracting} construct abstracting 
+On the other side, with the insight from Reynold that the defined language will inherit properties
+from the defining language, \citeauthor{darais2017abstracting} construct abstracting 
 definitional interpreters (ADI) that automatically inherit the pushdown control-flow 
-property from the defining language. TODO
+property from the defining language.
+TODO: connection to this paper.
 
-In the rest of this section, we take a closer look at the two state-of-art techniques
-to accomplish perfect call-stack precision.
+Question: although they establish pushdown control-flow analysis in different ways, 
+is there a correspondence between different methods?
+
+In the rest of this section, we take a closer look at the P4F and ADI which are
+two state-of-art techniques to accomplish perfect call-stack precision.
 

### 2018-02-26 paper.tex
update

@@
 
-Note that the store is monotonic. TODO
+TODO(rephrase): Note that the store is monotonic, the store is always growing and never becoming 
+smaller. This property guarantees we eventually will find a fixed point.
 
@@
 representing the remaining stack frames.
-TODO: nondet of cont?
+TODO: talk about nondet of cont deref?
 
@@
 
-atomicEval returns a set of storable values, i.e., closures.
+Since dereferencing an address becomes non-deterministic, our \texttt{atomicEval}
+function also truns into non-deterministic. Given an atomic expression \texttt{e},
+\texttt{atomicEval} returns a set of storable values, i.e., closures, to the caller.
+If the expression is just a \texttt{lambda} term, then the returned set is singleton.
 
 \begin{verbatim}
-
 def atomicEval(e: Expr, env: Env, bstore: BStore): Set[Storable] = e match {
@@
 \end{verbatim}
+
+The structure of function \texttt{step} is similar with concrete CESK machines, 
+except non-determinism which makes \texttt{step} returns a list of possible
+successor states.
+In the beginning of \texttt{step}, we call function \texttt{tick} to update 
+the time.
+Then we have two cases to consider: 1) Current control expression is \texttt{Let},
+and \texttt{App(f, ae)} is being bind to \texttt{x}.
+Fristly we retrieve the set of closures that \texttt{f} may represents.
+For each closure in the set, we do almost the same as what we did in concrete 
+CESK machines. But note that since the continuation is allocated on store (\texttt{kstore}),
+a new continuation address \texttt{new_kaddr} is constructed, and a new frame
+\texttt{Frame(x, e, env)} paired with the current continuation address \texttt{kaddr}
+is joined into this address. Finally, a list of states is generated.
+
+2) An atomic expression \texttt{ae} sits on the control position of the state.
+The value of \texttt{ae} is being returned to its caller. 
+We dereference the continuation address \texttt{kaddr} and obtain a set of
+continuations \texttt{conts}.
+For each continuation in the set, we construct an environment based on the
+environment of the frame (\texttt{env_f}), and bind \texttt{x} to a newly created binding 
+address \texttt{baddr}; also update the store with \texttt{baddr} and 
+the values that \texttt{ae} represents. In every generated states,
+the control becomes the expression \texttt{e} in the frame,
+and as we can tell from the name, continuation address \texttt{f_kaddr} 
+also comes from the frame.
 
@@
 function to the first element \texttt{hd} of the work list \texttt{todo} if \texttt{hd}
-is unseen, then adds the result of \texttt{step} to the work list and also adds \texttt{hd}
-to the explored states set.
-If the work list is empty, the function just
-returns the set of explored states up to now.
+is unseen, then adds the result of \texttt{step} to the rest of work list and in 
+the meanwhile adds \texttt{hd} to the explored states set.
+If the work list is empty, the function just returns the set of reachable states up to now.
 
@@
 If we use the monovariant analysis 0CFA, the analyzing result would be \texttt{x} and 
-\texttt{y} point to set \texttt{\{1, 2\}}, because it not distinguished 
+\texttt{y} point to set \texttt{\{1, 2\}}, because it not distinguishes
 \texttt{z} at different call-sites.
@@
 
-In this section, we firstly describe a simple but uncomputable solution to the
+In this section, we firstly show a simple but uncomputable solution to the
 problem; then briefly review several recent research works that enable
@@
 
-the only non-determinism in our state transition function is when dereferencing
-the callee \texttt{f} from store at a call-site.
+The state transition function \texttt{step} is still non-determinism, but
+the only non-determinism happens when dereferencing the callee \texttt{f} 
+from the store at call-site \texttt{App(f, ae)}. 
+For each closure of \texttt{f} in the set, we do exactly
+the same as what we did in concrete CESK machines: a new frame is constructed
+and pushed onto the stack.
+In the second case, since the continuation \texttt{konts} is a list of 
+frames, we can simply decompose it to differentiate an empty stack
+or a frame exists on the top of stack. 
+If the stack is empty, then empty successors is returned, because the computation
+is finished.
+Otherwise, we matched a frame \texttt{Frame(x, e, env_f)}. 
+This transition is deterministic, because we only have one frame, this frame
+is then popped and the new state uses the rest of stack.
 
@@
 
-On the other side, with the insight from Reynold that the defined language will inherit properties
-from the defining language, \citeauthor{darais2017abstracting} construct abstracting 
-definitional interpreters (ADI) that automatically inherit the pushdown control-flow 
-property from the defining language.
-TODO: connection to this paper.
-
-Question: although they establish pushdown control-flow analysis in different ways, 
-is there a correspondence between different methods?
-
-In the rest of this section, we take a closer look at the P4F and ADI which are
-two state-of-art techniques to accomplish perfect call-stack precision.
+In the rest of this section, we take a closer look at the P4F which is the
+state-of-art techniques to accomplish perfect call-stack precision on small-step
+AAMs, and Abstracting De nitional Interpreters (ADI) that adpots another 
+approach to pushdown control flow analysis.
 
@@
 
+On the other side, with the insight from Reynold that the defined language will inherit properties
+from the defining language, \citeauthor{darais2017abstracting} construct abstracting 
+definitional interpreters (ADI) that automatically inherit the pushdown control-flow 
+property from the defining language.
+TODO: connection to this paper.
+TODO: although they establish pushdown control-flow analysis in different ways, 
+is there a correspondence between different methods?
+
 \section{From AAM to ADI}
@@
 allocating continuations of defined language on the heap of defining language.
+
+\subsubsection{First Try}
+
+simply trun evaluation contexts into higher-order continuations
+
+\subsubsection{Computable Refunctionalization}
+
+fix point cache
 

### 2018-02-26 paper.tex
update

@@
 
-\subsubsection{Single-step Evaluation}
+\subsubsection{Single-step Transition}
 Before go into describing how the machine evaluates expressions, we firstly define several helper functions.
@@
 
-\subsubsection{Single-step Evaluation}
+\subsubsection{Single-step Transition}
 
@@
 
+for plain direct-style, the apply handles more cases.
+
 \subsection{Refunctionalization}

### 2018-02-26 paper.tex
update

@@
 from closure's environment and mapped \texttt{v} to a fresh address \texttt{addr}.
-The new store also extends with \texttt{addr} mapping to the value of \texttt{ae},
+The new store is extended with \texttt{addr} mapping to the value of \texttt{ae},
 which is evaluated from \texttt{atomicEval}.
@@
 The structure of function \texttt{step} is similar with concrete CESK machines, 
-except non-determinism which makes \texttt{step} returns a list of possible
+except non-determinism which makes \texttt{step} returns a list of reachable 
 successor states.
@@
 
-Finally, a user could use function \texttt{analyze} to obtain all possible states 
+Finally, a user could use function \texttt{analyze} to obtain all reachable states 
 for a given program.

### 2018-02-26 paper.tex
update

@@
   abstracting definitional interpreter (ADI) written in monadic style.
-  Danvy shows the correspondence between abstract machines and interpreters.
-  In this paper, we study the relations between small-step AAM and 
-  big-step ADI.
+  Danvy shows the correspondence between concrete abstract machines and interpreters.
+  But the relations between small-step abstracting abstract machines and 
+  big-step abstracting definitional interpreters is not well studied.
 
-  We show their correspondence and how to syntactically transform a small-step 
-  abstracting abstract machine to a big-step abstracting definitional 
-  interpreter.
+  In this paper, we show their correspondence and how to syntactically transform small-step 
+  abstracting abstract machines into big-step abstracting definitional 
+  interpreters.
   The transformations include fusing, disentangling, refunctionalizing 
-  and monadification(un-CPS), and all of them properly handle the 
+  and monadification (or un-CPS to direct-style). All of them properly handle the 
   collecting semantics and the non-determinism of abstract interpretation. 
@@
   continuations, we reveal how precise call/return match in control-flow 
-  analysis is obtained by refunctionalizing a small-step abstracting 
-  abstract machine,  % with unbounded stack
+  analysis can be obtained by refunctionalizing a small-step abstracting 
+  abstract machine,
   as well as explain how it been lost by defunctionalizing the higher-order
-  continuations into an algebraic data type. Further, we show how to turn 
-  precise call/return match into computable analysis.
+  continuations into an firts-order data type.
 

### 2018-02-26 paper.tex
update

@@
 state-of-art techniques to accomplish perfect call-stack precision on small-step
-AAMs, and Abstracting De nitional Interpreters (ADI) that adpots another 
+AAMs, and Abstracting Defnitional Interpreters (ADI) that adpots another 
 approach to pushdown control flow analysis.
@@
 %monadic
-\subsection{Monadification: From Refunctionalized AAM to ADI}
+\subsection{Monadification: to Abstracting Definitional Interpreter}
 
@@
 
-\subsection{Un-CPS: From Refunctionalized AAM to Direct Style}
+\subsection{Un-CPS: to Yet Another Abstracting Definitional Interpreter}
 
@@
 side effects such as assignment and mutation to handle collecting semantics and non-determinism.
+
+TODO: use delimited continuation? 
+Another form of abstracting definitional interpreter.
 

### 2018-02-26 paper.tex
update

@@
 pushing \texttt{f1} and \texttt{f2} onto the top of current stack, but not existing 
-two stack components are the same in the state space.
+two stack components are identical in the state space.
 
@@
 
+\begin{verbatim}
+def drive_step(todo: List[State], seen: Set[State]): Set[State] = todo match {
+  case Nil => seen
+  case hd::tl if seen.contains(hd) => drive_step(tl, seen)
+  case hd::tl =>
+    val new_time = tick(hd)
+    val new_states = hd match {
+      case State(Let(x, App(f, ae), e), env, bstore, konts, time) 
+        if isAtomic(ae) =>
+        for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
+          val frame = Frame(x, e, env)
+          val baddr = allocBind(v, new_time)
+          val new_env = env_c + (v -> baddr)
+          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+          State(body, new_env, new_store, frame::konts, new_time)
+        }
+      case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
+        konts match {
+          case Nil => List()
+          case Frame(x, e, f_env)::konts =>
+            val baddr = allocBind(x, new_time)
+            val new_env = f_env + (x -> baddr)
+            val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+            List(State(e, new_env, new_store, konts, new_time))
+        }
+    }
+    drive_step(new_states ++ tl, seen + hd)
+}
+
+def analyze(e: Expr): Set[State] = drive_step(List(inject(e)), Set())
+\end{verbatim}
+
 \subsection{Disentangling}
 
-for plain direct-style, the apply handles more cases.
+for plain direct-style, the continue function handles more cases.
+
+\begin{verbatim}
+def drive_step(todo: List[State], seen: Set[State]): Set[State] = todo match {
+  case Nil => seen
+  case hd::tl if seen.contains(hd) => drive_step(tl, seen)
+  case hd::tl =>
+    val new_time = tick(hd)
+    hd match {
+      case State(Let(x, App(f, ae), e), env, bstore, konts, time)
+        if isAtomic(ae) =>
+        val closures = atomicEval(f, env, bstore).toList
+        val new_states = for (Clos(Lam(v, body), env_c) <- closures) yield {
+          val frame = Frame(x, e, env)
+          val baddr = allocBind(v, new_time)
+          val new_env = env_c + (v -> baddr)
+          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+          State(body, new_env, new_store, frame::konts, new_time)
+        }
+        drive_step(new_states ++ tl, seen + hd)
+      case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
+        continue(konts, atomicEval(ae, env, bstore), bstore, new_time, tl, seen + hd)
+    }
+}
+
+def continue(konts: List[Frame], ds: Set[Storable], bstore: BStore, time: Time, 
+             todo: List[State], seen: Set[State]): Set[State] = 
+  konts match {
+    case Nil => drive_step(todo, seen)
+    case Frame(x, e, f_env)::konts =>
+      val baddr = allocBind(x, time)
+      val new_env = f_env + (x -> baddr)
+      val new_store = bstore.update(baddr, ds)
+      val new_state = State(e, new_env, new_store, konts, time)
+      drive_step(new_state::todo, seen)
+  }
+\end{verbatim}
 

### 2018-02-26 paper.tex
tweak

--- 2018-02-26 paper.tex
@@
 %% http://ctan.org/pkg/subcaption
-\usepackage{syntax}
+
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage[scaled=0.9]{beramono}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{xcolor,colortbl}
+\usepackage{url}
+\usepackage{listings}
+\usepackage{paralist}
+\usepackage{wrapfig}
+\usepackage{enumitem}
+\usepackage{multicol}
+\usepackage{flushend}
+
+\input{macros}
+
+\lstMakeShortInline[keywordstyle=,%
+              flexiblecolumns=false,%
+              %basewidth={0.56em, 0.52em},%
+              mathescape=false,%
+              basicstyle=\tt]@
 
@@
 
-\begin{verbatim}
-e \in Exp ::= ae
+\begin{lstlisting}
+e $\in$ Exp ::= ae
           | (let ([x (f ae)]) e)
-ae \in AExp ::= x
+ae $\in$ AExp ::= x            
             | lam
-lam \in Lam ::= (lambda (x) e)
+lam $\in$ Lam ::= (lambda (x) e)
 x ... are variable names
-\end{verbatim}
+\end{lstlisting}
 
-In ANF, all the function applications must be administrated in a \texttt{let} expression,
+In ANF, all the function applications must be administrated in a @let@ expression,
 and then bounded to a variable name under current environment.
-We can use \texttt{letrec} to create mutually recursive function bindings. But \texttt{letrec}
+We can use @letrec@ to create mutually recursive function bindings. But @letrec@
 only accepts atomic expressions that appear at the right-hand side of a binding.
-The atomic expressions $ae$ are either a variable, or a literal \texttt{lambda} term, which
+The atomic expressions $ae$ are either a variable, or a literal @lambda@ term, which
 both can be evaluated in a single step.
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 sealed trait Expr
@@
 case class Let(x: String, e: App, body: Expr) extends Expr
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 type Addr = Int
@@
 case class State(e: Expr, env: Env, store: Store, k: Kont)
-\end{verbatim}
+\end{lstlisting}
 
-It is worth noting that the continuation class \texttt{Kont} is defined as a list of frames.
-The frame class \texttt{Frame} stores the information of call-site, i.e., the information that
+It is worth noting that the continuation class @Kont@ is defined as a list of frames.
+The frame class @Frame@ stores the information of call-site, i.e., the information that
 can be used to resume the interrupted computation.
-A \texttt{Frame} constitutes of a variable name \texttt{x} to be bind later, a control expression
+A @Frame@ constitutes of a variable name @x@ to be bind later, a control expression
 that resumes to, and its environment. (TODO: mention that frame list vs KArg/KFun)
@@
 As we mentioned in section \ref{anfsyntax}, the atomic expressions are either a variable, or
-a literal \texttt{lambda} term. So the helper function \texttt{atomicEval} handles these two
+a literal @lambda@ term. So the helper function @atomicEval@ handles these two
 cases and evaluates atomic expressions to closures in a straightforward way.
-The \texttt{alloc} function generates a fresh address, and always allocates an unique integer under the domain
+The @alloc@ function generates a fresh address, and always allocates an unique integer under the domain
 of store.
-The \texttt{isAtomic} function is used as a predicate of whether the expression is atomic.
+The @isAtomic@ function is used as a predicate of whether the expression is atomic.
 
-\begin{verbatim}
+\begin{lstlisting}
 def atomicEval(e: Expr, env: Env, store: Store): Storable = e match {
@@
 def isAtomic(e: Expr): Boolean = e.isInstanceOf[Var] || e.isInstanceOf[Lam]
-\end{verbatim}
+\end{lstlisting}
 
-Now, we can faithfully describe the state transition function \texttt{step},
-that given a machine state, function \texttt{step} determines its successor state.
+Now, we can faithfully describe the state transition function @step@,
+that given a machine state, function @step@ determines its successor state.
 
-\begin{verbatim}
+\begin{lstlisting}
 def step(s: State): State = s match {
@@
 }
-\end{verbatim}
+\end{lstlisting}
 
-The first case is that the control of current state is a \texttt{Let} expression,
+The first case is that the control of current state is a @Let@ expression,
 and its right-hand side is an function application.
-By calling function \texttt{atomicEval}, we obtain the closure that callee \texttt{f} stands for.
-The successor state's control then transfers to the \texttt{body} expression of closure,
+By calling function @atomicEval@, we obtain the closure that callee @f@ stands for.
+The successor state's control then transfers to the @body@ expression of closure,
 with an updated environment and an update store. The new environment is extended
-from closure's environment and mapped \texttt{v} to a fresh address \texttt{addr}.
-The new store is extended with \texttt{addr} mapping to the value of \texttt{ae},
-which is evaluated from \texttt{atomicEval}.
-In the meanwhile, a new frame \texttt{frame} is pushed onto the stack \texttt{k}.
-This \texttt{frame} contains the variable name \texttt{x} at left-hand side position of \texttt{Let},
-the body expression of \texttt{Let}, and the lexical environment of the body expression.
+from closure's environment and mapped @v@ to a fresh address @addr@.
+The new store is extended with @addr@ mapping to the value of @ae@,
+which is evaluated from @atomicEval@.
+In the meanwhile, a new frame @frame@ is pushed onto the stack @k@.
+This @frame@ contains the variable name @x@ at left-hand side position of @Let@,
+the body expression of @Let@, and the lexical environment of the body expression.
 
-The second case for \texttt{step} is the control is transferred to an atomic expression.
+The second case for @step@ is the control is transferred to an atomic expression.
 We extract the top frame of continuations firstly.
 The control (that is an atomic expression) of current state is the evaluated term
-that being bind to the variable \texttt{x} from the top frame.
-The environment and store are updated with \texttt{x} mapping to the closure value of \texttt{ae}.
-Then the successor state is transferred to expression \texttt{e} from the top frame,
-which is the body of a \texttt{Let} expression, with the updated environment, store, and
-the rest of stack \texttt{ks}.
+that being bind to the variable @x@ from the top frame.
+The environment and store are updated with @x@ mapping to the closure value of @ae@.
+Then the successor state is transferred to expression @e@ from the top frame,
+which is the body of a @Let@ expression, with the updated environment, store, and
+the rest of stack @ks@.
 
@@
 
-To run the program, we first use function \texttt{inject} to construct an initial machine
-state given an expression \texttt{e}. The initial state contains an empty environment, 
+To run the program, we first use function @inject@ to construct an initial machine
+state given an expression @e@. The initial state contains an empty environment, 
 store and stack.
-\begin{verbatim}
+\begin{lstlisting}
 def inject(e: Expr): State = State(e, Map(), Map(), Nil)
-\end{verbatim}
+\end{lstlisting}
 
-Then the \texttt{drive} function is used to evaluate
-to a final state by iteratively apply \texttt{step} on current state until we reach a state
+Then the @drive@ function is used to evaluate
+to a final state by iteratively apply @step@ on current state until we reach a state
 that the control is an atomic expression and the continuation stack is empty.
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 def drive(s: State): State = s match {
@@
 def eval(e: Expr): State = drive(inject(e))
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-For clarity, we divide the store into binding store \texttt{BStore} and continuation store \texttt{KStore}. 
+For clarity, we divide the store into binding store @BStore@ and continuation store @KStore@. 
 The binding store maps binding addresses to sets of closure values, and continuation store maps
 continuation addresses to sets of continuations.
-We define a generic class \texttt{Store[K,V]} that performs join when updating elements
-in a store. By parameterizing \texttt{Store[K,V]} with \texttt{[BAddr, Storable]} and 
-\texttt{[KAddr, Cont]}, we obtain \texttt{BStore} and \texttt{KStore}, perspectively.
+We define a generic class @Store[K,V]@ that performs join when updating elements
+in a store. By parameterizing @Store[K,V]@ with @[BAddr, Storable]@ and 
+@[KAddr, Cont]@, we obtain @BStore@ and @KStore@, perspectively.
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 case class Store[K,V](map: Map[K, Set[V]]) {
@@
 type KStore = Store[KAddr, Cont]
-\end{verbatim}
+\end{lstlisting}
 
-The co-domain of binding stores \texttt{Storable} is the same
+The co-domain of binding stores @Storable@ is the same
 as we defined for CESK machines.
-The co-domain of continuation stores \texttt{Cont} is comprised of 
-a frame \texttt{Frame} and a continuation address \texttt{KAddr}.
-To mimic the runtime call stack, \texttt{KAddr} plays the role of
+The co-domain of continuation stores @Cont@ is comprised of 
+a frame @Frame@ and a continuation address @KAddr@.
+To mimic the runtime call stack, @KAddr@ plays the role of
 representing the remaining stack frames.
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 abstract class Storable 
@@
 case class Cont(frame: Frame, kaddr: KAddr)
-\end{verbatim}
+\end{lstlisting}
 
@@
 By dereferencing this address in continuation store, we can retrieve the 
-actual transfer of controls. The definition of environment \texttt{Env}
+actual transfer of controls. The definition of environment @Env@
 remains the same.
 
-\begin{verbatim}
+\begin{lstlisting}
 case class State(e: Expr, env: Env, bstore: BStore, 
                  kstore: KStore, k: KAddr, time: Time)
-\end{verbatim}
+\end{lstlisting}
 
@@
 Up to now, we did not describe much about the adresses in stores and the 
-time-stamp \texttt{Time}.
+time-stamp @Time@.
 But they are key ingredients to achieve analyses with different sensitivities,
@@
 To effectively approximate the runtime behavior, a finite 
-program contour \texttt{time} that encodes the program execution history is introduced.
+program contour @time@ that encodes the program execution history is introduced.
 Here, we use a list of execution contexts (expressions) to reprent that, and
-as we will see in section \ref{kcfainst}, by using different \texttt{tick}
+as we will see in section \ref{kcfainst}, by using different @tick@
 function on time we may obtain a family of analyses.
 
-\begin{verbatim}
+\begin{lstlisting}
 type Time = List[Expr]
-\end{verbatim}
+\end{lstlisting}
 
@@
 which are both finite.
-Continuation addresses \texttt{KAddr} has two variants: 
-1) \texttt{Halt} which corresponds to the empty stack, and
-2) \texttt{ContAddr} consists of the entry expression of callee and the time that
+Continuation addresses @KAddr@ has two variants: 
+1) @Halt@ which corresponds to the empty stack, and
+2) @ContAddr@ consists of the entry expression of callee and the time that
 creates this address, and they are also finite. 
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 case class BAddr(x: String, time: Time)
@@
 case class ContAddr(tgt: Expr, time: Time) extends KAddr
-\end{verbatim}
+\end{lstlisting}
 
-We use two helper functions \texttt{allocBind} and \texttt{allocKont} that will be used later
+We use two helper functions @allocBind@ and @allocKont@ that will be used later
 to allocate binding addresses and continuation addresses.
 
-\begin{verbatim}
+\begin{lstlisting}
 def allocBind(x: String, time: Time): BAddr = BAddr(x, time)
 def allocKont(tgtExpr: Expr, time: Time): KAddr = ContAddr(tgtExpr, time)
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-Since dereferencing an address becomes non-deterministic, our \texttt{atomicEval}
-function also truns into non-deterministic. Given an atomic expression \texttt{e},
-\texttt{atomicEval} returns a set of storable values, i.e., closures, to the caller.
-If the expression is just a \texttt{lambda} term, then the returned set is singleton.
+Since dereferencing an address becomes non-deterministic, our @atomicEval@
+function also truns into non-deterministic. Given an atomic expression @e@,
+@atomicEval@ returns a set of storable values, i.e., closures, to the caller.
+If the expression is just a @lambda@ term, then the returned set is singleton.
 
-\begin{verbatim}
+\begin{lstlisting}
 def atomicEval(e: Expr, env: Env, bstore: BStore): Set[Storable] = e match {
@@
 }
-\end{verbatim}
+\end{lstlisting}
 
-The structure of function \texttt{step} is similar with concrete CESK machines, 
-except non-determinism which makes \texttt{step} returns a list of reachable 
+The structure of function @step@ is similar with concrete CESK machines, 
+except non-determinism which makes @step@ returns a list of reachable 
 successor states.
-In the beginning of \texttt{step}, we call function \texttt{tick} to update 
+In the beginning of @step@, we call function @tick@ to update 
 the time.
-Then we have two cases to consider: 1) Current control expression is \texttt{Let},
-and \texttt{App(f, ae)} is being bind to \texttt{x}.
-Fristly we retrieve the set of closures that \texttt{f} may represents.
+Then we have two cases to consider: 1) Current control expression is @Let@,
+and @App(f, ae)@ is being bind to @x@.
+Fristly we retrieve the set of closures that @f@ may represents.
 For each closure in the set, we do almost the same as what we did in concrete 
-CESK machines. But note that since the continuation is allocated on store (\texttt{kstore}),
-a new continuation address \texttt{new_kaddr} is constructed, and a new frame
-\texttt{Frame(x, e, env)} paired with the current continuation address \texttt{kaddr}
+CESK machines. But note that since the continuation is allocated on store (@kstore@),
+a new continuation address @new_kaddr@ is constructed, and a new frame
+@Frame(x, e, env)@ paired with the current continuation address @kaddr@
 is joined into this address. Finally, a list of states is generated.
 
-2) An atomic expression \texttt{ae} sits on the control position of the state.
-The value of \texttt{ae} is being returned to its caller. 
-We dereference the continuation address \texttt{kaddr} and obtain a set of
-continuations \texttt{conts}.
+2) An atomic expression @ae@ sits on the control position of the state.
+The value of @ae@ is being returned to its caller. 
+We dereference the continuation address @kaddr@ and obtain a set of
+continuations @conts@.
 For each continuation in the set, we construct an environment based on the
-environment of the frame (\texttt{env_f}), and bind \texttt{x} to a newly created binding 
-address \texttt{baddr}; also update the store with \texttt{baddr} and 
-the values that \texttt{ae} represents. In every generated states,
-the control becomes the expression \texttt{e} in the frame,
-and as we can tell from the name, continuation address \texttt{f_kaddr} 
+environment of the frame (@env_f@), and bind @x@ to a newly created binding 
+address @baddr@; also update the store with @baddr@ and 
+the values that @ae@ represents. In every generated states,
+the control becomes the expression @e@ in the frame,
+and as we can tell from the name, continuation address @f_kaddr@
 also comes from the frame.
 
-\begin{verbatim}
+\begin{lstlisting}
 def step(s: State): List[State] = {
@@
 }
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+TODO
+
+\begin{lstlisting}
 def k: Int = 0
-def tick(s: State): Time = (s.e::s.time).take(k)
-\end{verbatim}
+def tick(s: State): Time = (s.e :: s.time).take(k)
+\end{lstlisting}
 
@@
 
-To run (analyze) a program, as CESK machine, we use \texttt{inject} to construct
+To run (analyze) a program, as CESK machine, we use @inject@ to construct
 the initial state given a program. 
 Note that the initial continuation store has a special mapping that maps continuation address
-\texttt{Halt} to empty set of continuations.
+@Halt@ to empty set of continuations.
 We also provide an empty program contour as initial time.
-\begin{verbatim}
+\begin{lstlisting}
 def inject(e: Expr): State = 
@@
         Halt, List())
-\end{verbatim}
+\end{lstlisting}
 
 However, comparing with concrete CESK machine, the
-\texttt{drive} function performs collecting semantics instead of valuation
-semantics. That is, for the purpose of analyzing programs, the function \texttt{drive}
+@drive@ function performs collecting semantics instead of valuation
+semantics. That is, for the purpose of analyzing programs, the function @drive@
 collects all the intermediate machine states as the program is abstractly executing.
-The following code shows a variant of work list algorithm that always applies \texttt{step}
-function to the first element \texttt{hd} of the work list \texttt{todo} if \texttt{hd}
-is unseen, then adds the result of \texttt{step} to the rest of work list and in 
-the meanwhile adds \texttt{hd} to the explored states set.
+The following code shows a variant of work list algorithm that always applies @step@
+function to the first element @hd@ of the work list @todo@ if @hd@
+is unseen, then adds the result of @step@ to the rest of work list and in 
+the meanwhile adds @hd@ to the explored states set.
 If the work list is empty, the function just returns the set of reachable states up to now.
 
-\begin{verbatim}
+\begin{lstlisting}
 def drive(todo: List[State], seen: Set[State]): Set[State] = todo match {
@@
 def analyze(e: Expr): Set[State] = drive(List(inject(e)), Set())
-\end{verbatim}
+\end{lstlisting}
 
-Finally, a user could use function \texttt{analyze} to obtain all reachable states 
+Finally, a user could use function @analyze@ to obtain all reachable states 
 for a given program.
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 (let ([id (lambda (z) z)])
@@
       x)))
-\end{verbatim}
+\end{lstlisting}
 
 In traditional $k$-CFA algorithm and the AAM we showed in section \ref{aam}, 
-the return-flow of invocation \texttt{(id 2)} 
-is going to both call-sites \texttt{(id 1)} and \texttt{(id 2)}. Therefore, the
-returned value \texttt{2} for variable \texttt{y} is also propagated to 
-variable \texttt{x} and then imprecise analysis result is arose.
+the return-flow of invocation @(id 2)@
+is going to both call-sites @(id 1)@ and @(id 1)@. Therefore, the
+returned value @2@ for variable @y@ is also propagated to 
+variable @x@ and then imprecise analysis result is arose.
 This return-flow merging is inevitable even we increase the context-sensitivity.
-If we use the monovariant analysis 0CFA, the analyzing result would be \texttt{x} and 
-\texttt{y} point to set \texttt{\{1, 2\}}, because it not distinguishes
-\texttt{z} at different call-sites.
-Under $1$-CFA, the algorithm is able to distinguish that variable \texttt{z} of function
-\texttt{id} has two different values at two call-sites, so variable \texttt{y} 
-would not be polluted by \texttt{1}, but it is still helplessness for distinguish
+If we use the monovariant analysis 0CFA, the analyzing result would be @x@ and 
+@y@ point to set @\{1, 2\}@, because it not distinguishes
+@z@ at different call-sites.
+Under $1$-CFA, the algorithm is able to distinguish that variable @z@ of function
+@id@ has two different values at two call-sites, so variable @y@ 
+would not be polluted by @1@, but it is still helplessness for distinguish
 return-flows.
@@
 
-In the definition of state \texttt{State}, continuation store disappears, and
-\texttt{konts} becomes a list of frames. Other components keep the same.
+In the definition of state @State@, continuation store disappears, and
+@konts@ becomes a list of frames. Other components keep the same.
 
-\begin{verbatim}
+\begin{lstlisting}
 case class State(e: Expr, env: Env, bstore: BStore, konts: List[Frame], time: Time)
-\end{verbatim}
+\end{lstlisting}
 
-The state transition function \texttt{step} is still non-determinism, but
-the only non-determinism happens when dereferencing the callee \texttt{f} 
-from the store at call-site \texttt{App(f, ae)}. 
-For each closure of \texttt{f} in the set, we do exactly
+The state transition function @step@ is still non-determinism, but
+the only non-determinism happens when dereferencing the callee @f@ 
+from the store at call-site @App(f, ae)@. 
+For each closure of @f@ in the set, we do exactly
 the same as what we did in concrete CESK machines: a new frame is constructed
 and pushed onto the stack.
-In the second case, since the continuation \texttt{konts} is a list of 
+In the second case, since the continuation @konts@ is a list of 
 frames, we can simply decompose it to differentiate an empty stack
@@
 is finished.
-Otherwise, we matched a frame \texttt{Frame(x, e, env_f)}. 
+Otherwise, we matched a frame @Frame(x, e, env_f)@. 
 This transition is deterministic, because we only have one frame, this frame
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 def step(s: State): List[State] = {
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 (letrec ([f1 (lambda (x) 
@@
     z))
-\end{verbatim}
+\end{lstlisting}
 
-\texttt{f1} and \texttt{f2} call each others, and the stack will be interleaved
-pushing \texttt{f1} and \texttt{f2} onto the top of current stack, but not existing 
-two stack components are the same in the state space.
+@f1@ and @f2@ call each others, and the stack will be interleaved
+pushing @f1@ and @f2@ onto the top of current stack, but not existing 
+two stack components are identical in the state space.
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 case class P4FContAddr(tgt: Expr, tgtEnv: Env, time: Time) extends KAddr
@@
   P4FContAddr(tgtExpr, tgtEnv, time)
-\end{verbatim}
+\end{lstlisting}
 
@@
 
+\begin{lstlisting}
+def drive_step(todo: List[State], seen: Set[State]): Set[State] = todo match {
+  case Nil => seen
+  case hd::tl if seen.contains(hd) => drive_step(tl, seen)
+  case hd::tl =>
+    val new_time = tick(hd)
+    val new_states = hd match {
+      case State(Let(x, App(f, ae), e), env, bstore, konts, time) 
+        if isAtomic(ae) =>
+        for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
+          val frame = Frame(x, e, env)
+          val baddr = allocBind(v, new_time)
+          val new_env = env_c + (v -> baddr)
+          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+          State(body, new_env, new_store, frame::konts, new_time)
+        }
+      case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
+        konts match {
+          case Nil => List()
+          case Frame(x, e, f_env)::konts =>
+            val baddr = allocBind(x, new_time)
+            val new_env = f_env + (x -> baddr)
+            val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+            List(State(e, new_env, new_store, konts, new_time))
+        }
+    }
+    drive_step(new_states ++ tl, seen + hd)
+}
+
+def analyze(e: Expr): Set[State] = drive_step(List(inject(e)), Set())
+\end{lstlisting}
+
 \subsection{Disentangling}
 
-for plain direct-style, the apply handles more cases.
+for plain direct-style, the continue function handles more cases.
+
+\begin{lstlisting}
+def drive_step(todo: List[State], seen: Set[State]): Set[State] = todo match {
+  case Nil => seen
+  case hd::tl if seen.contains(hd) => drive_step(tl, seen)
+  case hd::tl =>
+    val new_time = tick(hd)
+    hd match {
+      case State(Let(x, App(f, ae), e), env, bstore, konts, time)
+        if isAtomic(ae) =>
+        val closures = atomicEval(f, env, bstore).toList
+        val new_states = for (Clos(Lam(v, body), env_c) <- closures) yield {
+          val frame = Frame(x, e, env)
+          val baddr = allocBind(v, new_time)
+          val new_env = env_c + (v -> baddr)
+          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+          State(body, new_env, new_store, frame::konts, new_time)
+        }
+        drive_step(new_states ++ tl, seen + hd)
+      case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
+        continue(konts, atomicEval(ae, env, bstore), bstore, new_time, tl, seen + hd)
+    }
+}
+
+def continue(konts: List[Frame], ds: Set[Storable], bstore: BStore, time: Time, 
+             todo: List[State], seen: Set[State]): Set[State] = 
+  konts match {
+    case Nil => drive_step(todo, seen)
+    case Frame(x, e, f_env)::konts =>
+      val baddr = allocBind(x, time)
+      val new_env = f_env + (x -> baddr)
+      val new_store = bstore.update(baddr, ds)
+      val new_state = State(e, new_env, new_store, konts, time)
+      drive_step(new_state::todo, seen)
+  }
+\end{lstlisting}
 

### 2018-02-26 paper.tex
update

@@
 \usepackage[T1]{fontenc}
-\usepackage[scaled=0.9]{beramono}
+\usepackage[scaled=0.85]{beramono}
 \usepackage{amsmath}
@@
 the return-flow of invocation @(id 2)@
-is going to both call-sites @(id 1)@ and @(id 1)@. Therefore, the
+is going to both call-sites @(id 1)@ and @(id 2)@. Therefore, the
 returned value @2@ for variable @y@ is also propagated to 
@@
 This return-flow merging is inevitable even we increase the context-sensitivity.
-If we use the monovariant analysis 0CFA, the analyzing result would be @x@ and 
-@y@ point to set @\{1, 2\}@, because it not distinguishes
-@z@ at different call-sites.
+If we use the monovariant analysis (i.e. 0CFA), the analyzing result would be @x@ and 
+@y@ point to set @{1, 2}@, because the algorithm not distinguishes
+@z@ comes from different call-sites.
 Under $1$-CFA, the algorithm is able to distinguish that variable @z@ of function
 @id@ has two different values at two call-sites, so variable @y@ 
-would not be polluted by @1@, but it is still helplessness for distinguish
-return-flows.
+would not be polluted by @1@, but two call-sites still share the same continuation and
+it is still helplessness for distinguish return-flows.
 TODO: do need more details to explain this example? (as in P4F paper)
@@
 
-@f1@ and @f2@ call each others, and the stack will be interleaved
+@f1@ and @f2@ mutually call each others, and the stack will be alternatively
 pushing @f1@ and @f2@ onto the top of current stack, but not existing 
@@
 
-Recent years, there are great efforts \cite{vardoulakis2010cfa2, earl2012introspective, 
+In recent years, there are great efforts \cite{vardoulakis2010cfa2, earl2012introspective, 
 gilray2016pushdown, johnson2015abstracting} to achieve precise call/return 
@@
 computational inexpensive.
+The only change in code is when allocating a continuation address, not only
+keep track of the entry-point expression of callee, but also hold the target environment.
+The target environment is the environment extended from closure's environment
+(namely @val new_env = env_c + (v -> baddr)@). All the rest things remain the same.
 
@@
 
+TODO: explain the reason why it works.
+Notably, this change only causes constant-factor increase in time complexity of analysis.
+
 \textbf{Abstracting Definitional Interpreters.}
 
-On the other side, with the insight from Reynold that the defined language will inherit properties
+On the other side, with the insight from Reynold's seminal paper 
+\emph{Definitional interpreters for higher-order programming languages} \cite{reynolds1972definitional}
+that the defined language will inherit properties
 from the defining language, \citeauthor{darais2017abstracting} construct abstracting 
 definitional interpreters (ADI) that automatically inherit the pushdown control-flow 
-property from the defining language.
+property from its defining language.
+
 TODO: connection to this paper.

### 2018-02-27 paper.tex
update

@@
 \usepackage{flushend}
+\usepackage{tikz}
+
+\usetikzlibrary{arrows,chains,matrix,positioning,scopes}
+\makeatletter
+\tikzset{join/.code=\tikzset{after node path={%
+\ifx\tikzchainprevious\pgfutil@empty\else(\tikzchainprevious)%
+edge[every join]#1(\tikzchaincurrent)\fi}}}
+\makeatother
+
+\tikzset{>=stealth',every on chain/.append style={join},
+         every join/.style={->}}
+\tikzstyle{labeled}=[execute at begin node=$\scriptstyle,
+   execute at end node=$]
 
@@
 
-\subsection{Motivation}
-
 \subsection{Contributions}
@@
 
+The paper is organized as follows. 
+TODO: add a diagraph on transformations.
+Section \ref{background} reviews the background.
+
+\begin{tikzpicture}
+  \matrix (m) [matrix of math nodes, row sep=3em, column sep=3em]
+    { abstracting abstract machines & A  & B  & C  & 0 \\
+      abstract machines & A' & B' & C' & 0 \\ };
+  { [start chain] \chainin (m-1-1);
+    \chainin (m-1-2);
+    { [start branch=A] \chainin (m-2-2)
+        [join={node[right,labeled] {\eta_1}}];}
+    \chainin (m-1-3) [join={node[above,labeled] {\varphi}}];
+    { [start branch=B] \chainin (m-2-3)
+        [join={node[right,labeled] {\eta_2}}];}
+    \chainin (m-1-4) [join={node[above,labeled] {\psi}}];
+    { [start branch=C] \chainin (m-2-4)
+        [join={node[right,labeled] {\eta_3}}];}
+    \chainin (m-1-5); }
+  { [start chain] \chainin (m-2-1);
+    \chainin (m-2-2);
+    \chainin (m-2-3) [join={node[above,labeled] {\varphi'}}];
+    \chainin (m-2-4) [join={node[above,labeled] {\psi'}}];
+    \chainin (m-2-5); }
+\end{tikzpicture}
+
 \subsection{Style}
@@
 We use Scala language to demonstrate the idea and each step of transformations.
-There are two main reasons to do so:
+The reader should have moderate familiarity to Scala's syntax, such as
+case classes, pattern matching and for comprehension.
+
+There are two main reasons to use a real-world language:
 1) The code does not diminish the accuracy of the material
@@
 The Scala code in this paper can be easily back-translated into formal notations.
-2) The code is also executable with only few changes, which make it particularly fit for 
-presenting topic: syntactical transformations of programs.
-
-\section{Background}
+2) As a functional pearl, the code in this paper is executable with only few changes, 
+which make it particularly fit for presenting syntactical transformations on abstract
+interpreters.
+
+\section{Background} \label{background}
 
@@
 and continuations on the store, and then bound the addresses space to be finite.
-Thus the approximated machine-state space is finite, and computable.
+Since each component of state is finite, 
+thus the abstracted machine-state space is finite, and computable.
 
@@
 CESK machines, and also show that how to instantiate
-useful $k$-call-sensitive analysis.
+useful $k$-call-sensitive control-flow analysis.
 
@@
 
-Similar with CESK machines, the machine state of the AAM has a control expression,
+Similar with CESK machines, the machine state of AAM has a control expression,
 an environment, a store and continuation, and in addition with a time-stamp.
@@
 @Frame(x, e, env)@ paired with the current continuation address @kaddr@
-is joined into this address. Finally, a list of states is generated.
+is joined into @new_kaddr@. Finally, a list of successors states is generated.
 
@@
 
-To run (analyze) a program, as CESK machine, we use @inject@ to construct
+To run (analyze) a program, as CESK machine, we use function @inject@ to construct
 the initial state given a program. 
-Note that the initial continuation store has a special mapping that maps continuation address
+Note that the initial continuation store has a built-in mapping that maps continuation address
 @Halt@ to empty set of continuations.
@@
 A pushdown control-flow analysis provides as exact as runtime return-flow 
-when analyzing the program, but traditional control-flow analysis collapse 
-the state space using finite space abstraction thus cause imprecise stack modeling.
-
-To see the problem of spurious return-flows, we can consider the following example:
+when analyzing the program, but traditional control-flow analysis collapses 
+the state space into a finite space thus causes imprecise stack modeling.
+
+To see how traditional CFA suffers from spurious return-flows, 
+we can consider the following example:
 
@@
 @id@ has two different values at two call-sites, so variable @y@ 
-would not be polluted by @1@, but two call-sites still share the same continuation and
-it is still helplessness for distinguish return-flows.
+would not be polluted by @1@. 
+But two call-sites still share the same continuation and
+it is still helplessness for separating the return-flows.
 TODO: do need more details to explain this example? (as in P4F paper)
@@
 
-A naive solution to is to use precise call stack modeling as we did in concrete
+A naive solution to is to use precise call stack as we did in concrete
 CESK machine. Instead of allocating continuations on store and recording an 
-address of continuations, we leave the call stack unabstracted.
-
-In the definition of state @State@, continuation store disappears, and
-@konts@ becomes a list of frames. Other components keep the same.
+address of continuations in state, we leave the call stack unabstracted.
+
+In the definition of state @State@, the continuation store disappears, and
+component @konts@ becomes a list of frames. Other components keep the same.
 
@@
 
-The state transition function @step@ is still non-determinism, but
+The state transition function @step@ is still non-deterministic, but
 the only non-determinism happens when dereferencing the callee @f@ 
@@
 In the second case, since the continuation @konts@ is a list of 
-frames, we can simply decompose it to differentiate an empty stack
+frames, we can simply decompose it to differentiate an empty stack case
 or a frame exists on the top of stack. 
-If the stack is empty, then empty successors is returned, because the computation
-is finished.
-Otherwise, we matched a frame @Frame(x, e, env_f)@. 
-This transition is deterministic, because we only have one frame, this frame
-is then popped and the new state uses the rest of stack.
+If the stack is empty, then empty list of successors is returned, 
+because the computation is finished.
+Otherwise, we matched a frame @Frame(x, e, env_f)@ on the top of stack.
+The frame is then poped, and the new state uses the rest of stack.
+Since we can only access one frame at a time, this transition is deterministic.
 
@@
 
-In recent years, there are great efforts \cite{vardoulakis2010cfa2, earl2012introspective, 
+In recent years, there are significant efforts \cite{vardoulakis2010cfa2, earl2012introspective, 
 gilray2016pushdown, johnson2015abstracting} to achieve precise call/return 
@@
 \citeauthor{gilray2016pushdown} proposed a continuation address allocator for 
-small-step abstracting abstract machine that is both easy to implement and
+small-step abstracting abstract machines that is both simple to implement and
 computational inexpensive.
@@
 keep track of the entry-point expression of callee, but also hold the target environment.
-The target environment is the environment extended from closure's environment
+The target environment is the binding environment extended from closure's environment
 (namely @val new_env = env_c + (v -> baddr)@). All the rest things remain the same.
@@
 TODO: explain the reason why it works.
-Notably, this change only causes constant-factor increase in time complexity of analysis.
+
+Notably, this change only causes constant-factor increase in time complexity of analysis
+if the store is widened.
 
@@
 from the defining language, \citeauthor{darais2017abstracting} construct abstracting 
-definitional interpreters (ADI) that automatically inherit the pushdown control-flow 
+definitional interpreters that automatically inherit the pushdown control-flow 
 property from its defining language.

### 2018-02-27 paper.tex
update

@@
                                         %% contents suppressed with 'anonymous'
-\subtitle{Filling the Gap between Abstracting Abstract Machine and Abstracting Definitional Interpreter}        %% \subtitle is optional
+\subtitle{Filling the Gap between Abstracting Abstract Machines and Abstracting Definitional Interpreters}        %% \subtitle is optional
 \subtitlenote{with subtitle note}       %% \subtitlenote is optional;
@@
 
+boardly speaking, AAM and ADI they are both abstract interpreters.
+
 \subsection{Contributions}
@@
 The paper is organized as follows. 
-TODO: add a diagraph on transformations.
 Section \ref{background} reviews the background.
@@
 \begin{tikzpicture}
-  \matrix (m) [matrix of math nodes, row sep=3em, column sep=3em]
-    { abstracting abstract machines & A  & B  & C  & 0 \\
-      abstract machines & A' & B' & C' & 0 \\ };
-  { [start chain] \chainin (m-1-1);
-    \chainin (m-1-2);
-    { [start branch=A] \chainin (m-2-2)
-        [join={node[right,labeled] {\eta_1}}];}
-    \chainin (m-1-3) [join={node[above,labeled] {\varphi}}];
-    { [start branch=B] \chainin (m-2-3)
-        [join={node[right,labeled] {\eta_2}}];}
-    \chainin (m-1-4) [join={node[above,labeled] {\psi}}];
-    { [start branch=C] \chainin (m-2-4)
-        [join={node[right,labeled] {\eta_3}}];}
-    \chainin (m-1-5); }
-  { [start chain] \chainin (m-2-1);
-    \chainin (m-2-2);
-    \chainin (m-2-3) [join={node[above,labeled] {\varphi'}}];
-    \chainin (m-2-4) [join={node[above,labeled] {\psi'}}];
-    \chainin (m-2-5); }
+  \matrix (m) [matrix of math nodes, row sep=3em, column sep=3.7em]
+    { 
+      \begin{smallmatrix} \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
+      \begin{smallmatrix} \text{big-step} \\ \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  & 
+      \begin{smallmatrix} \text{disentangled}  \\ \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
+      \begin{smallmatrix} \text{refunctionalized}  \\ \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
+      \begin{smallmatrix} \text{abstracting} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
+      \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} & 
+        &
+      %\begin{smallmatrix} \text{big-step} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  & 
+        &
+      %\begin{smallmatrix} \text{disentangled} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
+        &
+      %\begin{smallmatrix} \text{refunctionalized} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
+      \begin{smallmatrix} \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\ 
+    };
+
+  { 
+    [start chain] \chainin (m-1-1);
+    \chainin (m-1-2) [join={node[above,labeled] {\text{fusion}}}];
+
+    \chainin (m-1-3) [join={node[above,labeled] {\text{disentangling}}}];
+
+    \chainin (m-1-4) [join={node[above,labeled] {\text{refunc.}}}];
+
+    \chainin (m-1-5) [join={node[above,labeled] {\text{monadify}}}];
+  }
+  { 
+    [start chain] 
+    \chainin (m-2-1);
+    { [start branch=A] \chainin (m-1-1)
+      [join={node[left, labeled] {\alpha}}];
+    }
+  }
+    %\chainin (m-2-2);
+    %\chainin (m-2-3);
+    %\chainin (m-2-4);
+  {
+    [start chain] 
+    \chainin (m-2-5); 
+    { [start branch=B] \chainin (m-1-5)
+      [join={node[left, labeled] {\alpha}}];
+    }
+  }
 \end{tikzpicture}
@@
 
-There are two main reasons to use a real-world language:
+There are two main reasons we use a real-world language:
 1) The code does not diminish the accuracy of the material
 than formal and mathematical notations, 
-which are heavily used in other papers in static analysis or semantics.
+which are heavily used in other static analysis or semantics papers.
 The Scala code in this paper can be easily back-translated into formal notations.

### 2018-02-27 paper.tex
update

@@
 We use Scala language to demonstrate the idea and each step of transformations.
-The reader should have moderate familiarity to Scala's syntax, such as
+We expect that readers have moderate familiarity to Scala's syntax, such as
 case classes, pattern matching and for comprehension.
@@
 \begin{lstlisting}
-def drive_step(todo: List[State], seen: Set[State]): Set[State] = todo match {
+def aval(state: State, todo: List[State], seen: Set[State]): Set[State] = { 
+  val newTime = tick(state)
+  state match {
+    case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
+      val newStates = for (Clos(Lam(v, body), c_env) <- aeval(f, env, bstore).toList) yield {
+        val frame = Frame(x, e, env)
+        val baddr = allocBind(v, newTime)
+        val newEnv = c_env + (v -> baddr)
+        val newStore = bstore.update(baddr, aeval(ae, env, bstore))
+        State(body, newEnv, newStore, frame::konts, newTime)
+      }   
+      collect(newStates ++ todo, seen + state)
+    case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
+      continue(konts, aeval(ae, env, bstore), bstore, newTime, todo, seen + state)
+  }   
+}
+
+def collect(todo: List[State], seen: Set[State]): Set[State] = { 
+  todo match {
     case Nil => seen
-  case hd::tl if seen.contains(hd) => drive_step(tl, seen)
-  case hd::tl =>
-    val new_time = tick(hd)
-    hd match {
-      case State(Let(x, App(f, ae), e), env, bstore, konts, time)
-        if isAtomic(ae) =>
-        val closures = atomicEval(f, env, bstore).toList
-        val new_states = for (Clos(Lam(v, body), env_c) <- closures) yield {
-          val frame = Frame(x, e, env)
-          val baddr = allocBind(v, new_time)
-          val new_env = env_c + (v -> baddr)
-          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-          State(body, new_env, new_store, frame::konts, new_time)
-        }
-        drive_step(new_states ++ tl, seen + hd)
-      case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
-        continue(konts, atomicEval(ae, env, bstore), bstore, new_time, tl, seen + hd)
+    case hd::tl if seen.contains(hd) => collect(tl, seen)
+    case hd::tl => aval(hd, tl, seen)
   }   
@@
 def continue(konts: List[Frame], ds: Set[Storable], bstore: BStore, time: Time,
-             todo: List[State], seen: Set[State]): Set[State] = 
-  konts match {
-    case Nil => drive_step(todo, seen)
+              todo: List[State], seen: Set[State]): Set[State] = konts match {
+  case Nil => collect(todo, seen)
   case Frame(x, e, f_env)::konts =>
     val baddr = allocBind(x, time)
-      val new_env = f_env + (x -> baddr)
-      val new_store = bstore.update(baddr, ds)
-      val new_state = State(e, new_env, new_store, konts, time)
-      drive_step(new_state::todo, seen)
-  }
+    val newEnv = f_env + (x -> baddr)
+    val newStore = bstore.update(baddr, ds) 
+    val newState = State(e, newEnv, newStore, konts, time)
+    collect(newState::todo, seen)
+}
+
+def analyze(e: Expr): Set[State] = aval(inject(e), List(), Set()) 
 \end{lstlisting}
@@
 allocating continuations of defined language on the heap of defining language.
+
+\begin{lstlisting}
+case class Config(e: Expr, env: Env, store: BStore, time: Time)
+case class ValStore(vs: Set[Storable], store: BStore, time: Time)
+type Cache = Map[Config, ValStore]
+
+type Continue = Function4[Set[Storable], BStore, Time, Cache, ValStore]
+
+def tick(s: Config): Time = (s.e::s.time).take(k)
+
+def aval(e: Expr, env: Env, bstore: BStore, time: Time, cache: Cache, continue: Continue): ValStore = {
+  val config = Config(e, env, bstore, time)
+  if (cache.contains(config)) {
+    val ValStore(vs, bstore, time) = cache(config)
+    return continue(vs, bstore, time, cache)
+  }
+  val newTime = tick(config)
+  e match {
+    case Let(x, App(f, ae), e) if isAtomic(ae) =>
+      val closures = aeval(f, env, bstore).toList
+
+      def nd(closures: List[Storable], accds: Set[Storable], accbstore: BStore, acctime: Time, cache: Cache): ValStore  = {
+        closures match {
+          case Nil =>
+            val baddr = allocBind(x, acctime)
+            val newEnv = env + (x -> baddr)
+            val newStore = accbstore.update(baddr, accds)
+            aval(e, newEnv, newStore, acctime, cache, (ds: Set[Storable], accbstore: BStore, acctime: Time, accCache: Cache) => {
+                val newCache = accCache + (Config(e, newEnv, newStore, acctime) -> ValStore(ds, accbstore, acctime))
+                continue(ds, accbstore, acctime, newCache)
+            })
+          case Clos(Lam(v, body), env_c)::rest =>
+            val baddr = allocBind(v, newTime)
+            val newEnv = env_c + (v -> baddr)
+            val newStore = accbstore.update(baddr, aeval(ae, env, bstore)) //TODO: which store?
+            aval(body, newEnv, newStore, newTime, cache, (ds: Set[Storable], accbstore: BStore, acctime: Time, accCache: Cache) => {
+                val newCache = accCache + (Config(body, newEnv, newStore, newTime) -> ValStore(ds, accbstore, acctime))
+                nd(rest, ds++accds, accbstore, acctime, newCache)
+            })
+        }
+      }
+
+      nd(closures, Set(), bstore, newTime, cache)
+    case ae if isAtomic(ae) =>
+      val ds = aeval(ae, env, bstore)
+      continue(ds, bstore, newTime, cache)
+  }
+}
+
+def done: Continue = (ds: Set[Storable], bstore: BStore, time: Time, cache: Cache) => {
+  ValStore(ds, bstore, time)
+}
+def analyze(e: Expr): ValStore = aval(e, Map(), Store[BAddr, Storable](Map()), List(), Map[Config, ValStore](), done)
+\end{lstlisting}
 

### 2018-02-27 paper.tex
update

@@
 1) @Halt@ which corresponds to the empty stack, and
-2) @ContAddr@ consists of the entry expression of callee and the time that
-creates this address, and they are also finite. 
+2) @ContAddr@ consists of the entry expressions of callee, which are also also
+finite.
+One may ask, why not also keep track of the time in continuation addresses?
 (TODO: why target expression, allocation polyvariance)
@@
 case object Halt extends KAddr
-case class ContAddr(tgt: Expr, time: Time) extends KAddr
+case class ContAddr(tgt: Expr) extends KAddr
 \end{lstlisting}
@@
 def allocBind(x: String, time: Time): BAddr = BAddr(x, time)
-def allocKont(tgtExpr: Expr, time: Time): KAddr = ContAddr(tgtExpr, time)
+def allocKont(tgtExpr: Expr): KAddr = ContAddr(tgtExpr)
 \end{lstlisting}
@@
         val new_bstore = bstore.update(baddr, atomicEval(ae, env, bstore))
-        val new_kaddr = allocKont(body, new_time)
+        val new_kaddr = allocKont(body)
         val new_kstore = kstore.update(new_kaddr, Cont(Frame(x, e, env), kaddr))
@@
 \begin{lstlisting}
-case class P4FContAddr(tgt: Expr, tgtEnv: Env, time: Time) extends KAddr
-def allocKont(tgtExpr: Expr, tgtEnv: Env, time: Time): KAddr = 
-  P4FContAddr(tgtExpr, tgtEnv, time)
+case class P4FContAddr(tgt: Expr, tgtEnv: Env) extends KAddr
+def allocKont(tgtExpr: Expr, tgtEnv: Env): KAddr = 
+  P4FContAddr(tgtExpr, tgtEnv)
 \end{lstlisting}
@@
 case class Config(e: Expr, env: Env, store: BStore, time: Time)
-case class ValStore(vs: Set[Storable], store: BStore, time: Time)
+case class ValStore(ds: Set[Storable], store: BStore, time: Time)
 type Cache = Map[Config, ValStore]
 
-type Continue = Function4[Set[Storable], BStore, Time, Cache, ValStore]
+type Continue = Function2[ValStore, Cache, ValStore]
 
@@
   if (cache.contains(config)) {
-    val ValStore(vs, bstore, time) = cache(config)
-    return continue(vs, bstore, time, cache)
+    return continue(cache(config), cache)
   }   
@@
 
-      def nd(closures: List[Storable], accds: Set[Storable], accbstore: BStore, acctime: Time, cache: Cache): ValStore  = {
+      def nd(closures: List[Storable], accds: Set[Storable],
+             accbstore: BStore, acctime: Time, cache: Cache): ValStore  = { 
         closures match {
@@
             val newStore = accbstore.update(baddr, accds)
-            aval(e, newEnv, newStore, acctime, cache, (ds: Set[Storable], accbstore: BStore, acctime: Time, accCache: Cache) => {
-                val newCache = accCache + (Config(e, newEnv, newStore, acctime) -> ValStore(ds, accbstore, acctime))
-                continue(ds, accbstore, acctime, newCache)
+            aval(e, newEnv, newStore, acctime, cache, (vs: ValStore, accCache: Cache) => {
+              val newCache = accCache + (Config(e, newEnv, newStore, acctime) -> vs) 
+              continue(vs, newCache)
             })  
@@
             val newStore = accbstore.update(baddr, aeval(ae, env, bstore)) //TODO: which store?
-            aval(body, newEnv, newStore, newTime, cache, (ds: Set[Storable], accbstore: BStore, acctime: Time, accCache: Cache) => {
-                val newCache = accCache + (Config(body, newEnv, newStore, newTime) -> ValStore(ds, accbstore, acctime))
-                nd(rest, ds++accds, accbstore, acctime, newCache)
+            aval(body, newEnv, newStore, newTime, cache, (vs: ValStore, accCache: Cache) => {
+              val newCache = accCache + (Config(body, newEnv, newStore, newTime) -> vs) 
+              nd(rest, vs.ds++accds, accbstore, acctime, newCache)
             })  
@@
       val ds = aeval(ae, env, bstore)
-      continue(ds, bstore, newTime, cache)
-  }
-}
-
-def done: Continue = (ds: Set[Storable], bstore: BStore, time: Time, cache: Cache) => {
-  ValStore(ds, bstore, time)
-}
-def analyze(e: Expr): ValStore = aval(e, Map(), Store[BAddr, Storable](Map()), List(), Map[Config, ValStore](), done)
+      continue(ValStore(ds, bstore, newTime), cache)
+  }   
+}
+
+def done: Continue = (vs: ValStore, cache: Cache) => { vs }
+def analyze(e: Expr): ValStore = aval(e, Map(), Store[BAddr, Storable](Map()),
+                                      List(), Map[Config, ValStore](), done)
 \end{lstlisting}

### 2018-02-27 paper.tex
update

@@
 As we mentioned in section \ref{anfsyntax}, the atomic expressions are either a variable, or
-a literal @lambda@ term. So the helper function @atomicEval@ handles these two
+a literal @lambda@ term. So the atomic expression evaluator @atomicEval@ handles these two
 cases and evaluates atomic expressions to closures in a straightforward way.
@@
 and its right-hand side is an function application.
-By calling function @atomicEval@, we obtain the closure that callee @f@ stands for.
+By calling evaluator @atomicEval@, we obtain the closure that callee @f@ stands for.
 The successor state's control then transfers to the @body@ expression of closure,
@@
 case class P4FContAddr(tgt: Expr, tgtEnv: Env) extends KAddr
-def allocKont(tgtExpr: Expr, tgtEnv: Env): KAddr = 
-  P4FContAddr(tgtExpr, tgtEnv)
+def allocKont(tgtExpr: Expr, tgtEnv: Env): KAddr = P4FContAddr(tgtExpr, tgtEnv)
 \end{lstlisting}
@@
 \textbf{Abstracting Definitional Interpreters.}
-
 On the other side, with the insight from Reynold's seminal paper 
@@
 We start from the AAM with unbounded stack.
+In fact, since later we will eliminate the call stack and transform to
+higher-order continuations, it does not matter what kind of AAM we start from.
 
@@
 allocating continuations of defined language on the heap of defining language.
+
+\subsubsection{First Try}
+
+simply trun evaluation contexts into higher-order continuations.
+We establish pushdown control-flow, but this is not computable.
 
@@
 
-\subsubsection{First Try}
-
-simply trun evaluation contexts into higher-order continuations
-
-\subsubsection{Computable Refunctionalization}
+\subsubsection{Caching Fixed-Points}
 

### 2018-02-28 paper.tex
update

@@
   Danvy shows the correspondence between concrete abstract machines and interpreters.
-  But the relations between small-step abstracting abstract machines and 
+  Yet, the relations between small-step abstracting abstract machines and 
   big-step abstracting definitional interpreters is not well studied.
@@
   abstract interpreter.
-  Remarkably, follow the idea that evaluation contexts are defunctionalized 
+  Remarkably, refunctionalization sequentializes the abstract evaluation.
+  And follow the idea that evaluation contexts are defunctionalized 
   continuations, we reveal how precise call/return match in control-flow 
@@
 sealed trait Expr
-case class Binding(x: String, e: Expr)
-
 case class Var(x: String) extends Expr
@@
 
+\begin{lstlisting}
+case class Config(e: Expr, env: Env)
+case class ValStore(ds: Set[Storable], store: BStore, time: Time)
+type Cache = Store[Config, ValStore]
+type RetType = (Set[ValStore], Cache)
+type Continue = Function5[Set[ValStore], BStore, Time, Cache, Cache, RetType]
+
+def tick(e: Expr, oldtime: Time): Time = (e::oldtime).take(k)
+
+def aval(e: Expr, env: Env, bstore: BStore, time: Time,
+         inCache: Cache, _outCache: Cache, continue: Continue): RetType = {
+  val config = Config(e, env)
+  if (_outCache.contains(config)) {
+    return continue(_outCache(config), bstore, List(), inCache, _outCache)
+  }
+  val outCache = _outCache.update(config, inCache.getOrElse(config, Set()))
+  val newTime = tick(e, time)
+  e match {
+    case Let(x, App(f, ae), e) if isAtomic(ae) =>
+      val closures = aeval(f, env, bstore).toList
+      def nd(closures: List[Storable], accds: Set[Storable], accbstore: BStore, acctime: Time,
+              accInCache: Cache, accOutCache: Cache): RetType  = closures match {
+        case Nil =>
+          // TODO: If we use `newTime` instead of `acctime`, is it m-CFA?
+          val baddr = allocBind(x, acctime)
+          val newEnv = env + (x -> baddr)
+          val newStore = accbstore.update(baddr, accds)
+          aval(e, newEnv, newStore, acctime, accInCache, accOutCache,
+            (vs: Set[ValStore], accStore: BStore, acctime: Time, accInCache: Cache, accOutCache: Cache) => {
+              val newOutCache = accOutCache.update(Config(e, newEnv), vs)
+              continue(vs, accStore, acctime, accInCache, newOutCache)
+            })
+        case Clos(Lam(v, body), env_c)::rest =>
+          val baddr = allocBind(v, newTime)
+          val newEnv = env_c + (v -> baddr)
+          val newStore = accbstore.update(baddr, aeval(ae, env, bstore)) //TODO: which store?
+          aval(body, newEnv, newStore, newTime, accInCache, accOutCache,
+            (vs: Set[ValStore], accStore: BStore, acctime: Time, accInCache: Cache, accOutCache) => {
+              val newOutCache = accOutCache.update(Config(body, newEnv), vs)
+              val vsds = vs.map(_.ds).foldLeft(Set[Storable]())(_ ++ _)
+              nd(rest, vsds++accds, accStore, acctime, accInCache, newOutCache)
+            })
+      }
+      nd(closures, Set(), bstore, newTime, inCache, outCache)
+    case ae if isAtomic(ae) =>
+      val ds = aeval(ae, env, bstore)
+      continue(Set(ValStore(ds, bstore, newTime)), bstore, newTime, inCache, outCache)
+    }
+}
+
+def done: Function2[Expr, Env, Continue] = (e: Expr, env: Env) => {
+  (vs: Set[ValStore], store: BStore, time: Time, inCache: Cache, outCache: Cache) => {
+    val newOutCache = outCache.update(Config(e, env), vs)
+    (vs, outCache)
+  }
+}
+
+def analyze1(e: Expr, cache: Cache): RetType = {
+  aval(e, Map(), Store[BAddr, Storable](Map()), List(), cache, cache, done(e, Map()))
+}
+
+def analyze(e: Expr): Set[ValStore] = {
+  def iter(cache: Cache): RetType = {
+    val (vs, newCache) = analyze1(e, cache)
+    if (cache == newCache) (vs, newCache)
+    else iter(newCache)
+  }
+  val (vs, cache) = iter(Store[Config, ValStore](Map()))
+  val summarized = cache.map.values.map(_.map(_.store).foldLeft(Store[BAddr, Storable](Map()))(_.join(_))).foldLeft(Store[BAddr, Storable](Map()))(_.join(_))
+  vs
+}
+\end{lstlisting}
+
 %probably Store Widening \& Abstract Garbage Collection
@@
 show the precision of stack
+
+On equivalence: since refunc interp does not have an explicit concept of $state$,
+so 1) we can compare the final states of P4F against with refunc interp,
+or 2) we can compare the summarized store of P4F against with summarized store of refunc interp.
 

### 2018-03-01 paper.tex
update

@@
 The paper is organized as follows.
-Section \ref{background} reviews the background.
+
+In section \ref{background}, we first describe the target language we will
+analyze and a concrete CESK machine as operational semantics for it.
+Then we review abstracting abstract machines (AAM) as an abstract interpretation
+counterpart of CESK machines.
 
@@
 
-In this section, we first describe the target language we will
-analyze and a concrete CESK machine as operational semantics for it.
-Then we review some recent works in control-flow analysis for
-higher-order languages:
-1) abstracting abstract machines (AAM) as an abstract interpretation
-counterpart of CESK machines, and
 2) the precise call/return match problem in control-flow analysis and its existing solutions.
-We will show in section 3 that the abstracting abstract machines form the basis of our transformations,
-and call/return match is obtained during transformation.
 
@@
 Although we just show the core calculus language, but it can be easily extended
-to support recursive bindings, conditionals, primitive types and operations on primitive types.
+to support recursive bindings (such as $letrec$), conditionals, primitive types and 
+operations on primitive types. These cases would be trival or simple, so we omit them
+in the paper.
 The transformations we will show in the rest of this paper

### 2018-03-06 paper.tex
reorganize

@@
   The transformations include fusing, disentangling, refunctionalizing 
-  and monadification (or un-CPS to direct-style). All of them properly handle the 
+  and monadification (or un-CPS to direct-style with delimited continuation). 
+  All of them properly handle the 
   collecting semantics and the non-determinism of abstract interpretation. 
@@
   between small-step AAM and big-step ADI by developing a series of
-  systematic transformations. The correspondence between abstract
+  systematic transformations. 
+  The correspondence between abstract
   machines and interpreters exists not only in concrete semantics artifact,
@@
 
-\item We present an abstract interpreter obtained from refunctionalizing
+\item A new understanding of pushdown control-flow analysis by 
+  refunctionalization.
+  We present an abstract interpreter obtained from refunctionalizing
   an small-step abstracting abstract machine. 
   This abstract interpreter naturally has pushdown control-flow property
-  since we represent the evaluation contexts of analyzed language using
-  continuations of the meta-language, and therefore sequentialize the
-  analyzing.
-  We also analyze its termination, soundness and complexity.
+  because the refunctionalization forces us to sequentialize the 
+  abstract exectuion.
+  We also analyze its termination and soundness.
 
@@
 \begin{tikzpicture}
-  \matrix (m) [matrix of math nodes, row sep=3em, column sep=3.7em]
+  \matrix (m) [matrix of math nodes, row sep=2.5em, column sep=3.7em]
     { 
+        &  &  &  & 
+      \begin{smallmatrix} \text{abstracting} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
       \begin{smallmatrix} \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
@@
       \begin{smallmatrix} \text{refunctionalized}  \\ \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
-      \begin{smallmatrix} \text{abstracting} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
+      \begin{smallmatrix} \text{monadic} \\ \text{abstracting} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
       \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} & 
@@
     };
-
   {
-    [start chain] \chainin (m-1-1);
-    \chainin (m-1-2) [join={node[above,labeled] {\text{fusion}}}];
-
-    \chainin (m-1-3) [join={node[above,labeled] {\text{disentangling}}}];
-
-    \chainin (m-1-4) [join={node[above,labeled] {\text{refunc.}}}];
-
-    \chainin (m-1-5) [join={node[above,labeled] {\text{monadify}}}];
+    [start chain] \chainin (m-2-4);
+    \chainin (m-1-5) [join={node[sloped,anchor=center,above,labeled] {\text{un-CPS}}}];
+    \chainin (m-2-5) [join={node[above,labeled] {\text{}}}];
+    \chainin (m-1-5) [join={node[above,labeled] {\text{}}}];
+  }
+  { 
+    [start chain] \chainin (m-2-1);
+    \chainin (m-2-2) [join={node[above,labeled] {\text{fusion}}}];
+    \chainin (m-2-3) [join={node[above,labeled] {\text{disentangling}}}];
+    \chainin (m-2-4) [join={node[above,labeled] {\text{refunc.}}}];
+    \chainin (m-2-5) [join={node[above,labeled] {\text{monadify}}}];
   }
@@
     [start chain] 
-    \chainin (m-2-1);
-    { [start branch=A] \chainin (m-1-1)
+    \chainin (m-3-1);
+    { [start branch=A] \chainin (m-2-1)
       [join={node[left, labeled] {\alpha}}];
@@
     [start chain] 
-    \chainin (m-2-5); 
-    { [start branch=B] \chainin (m-1-5)
-      [join={node[left, labeled] {\alpha}}];
+    \chainin (m-3-5); 
+    { [start branch=B] \chainin (m-2-5)
+      [join={node[right, labeled] {\alpha}}];
     }
@@
 \section{Background} \label{background}
-
-2) the precise call/return match problem in control-flow analysis and its existing solutions.
 
@@
 
-\subsection{Pushdown Control-Flow Analysis}
-
-Pushdown control-flow is a property in analysis that precisely models
-the runtime call-stack of the analyzed program. 
-A pushdown control-flow analysis provides as exact as runtime return-flow 
-when analyzing the program, but traditional control-flow analysis collapses 
-the state space into a finite space thus causes imprecise stack modeling.
-
-To see how traditional CFA suffers from spurious return-flows, 
-we can consider the following example:
-
-\begin{lstlisting}
-(let ([id (lambda (z) z)])
-  (let ([x (id 1)])
-    (let ([y (id 2)])
-      x)))
-\end{lstlisting}
-
-In traditional $k$-CFA algorithm and the AAM we showed in section \ref{aam}, 
-the return-flow of invocation @(id 2)@
-is going to both call-sites @(id 1)@ and @(id 2)@. Therefore, the
-returned value @2@ for variable @y@ is also propagated to 
-variable @x@ and then imprecise analysis result is arose.
-This return-flow merging is inevitable even we increase the context-sensitivity.
-If we use the monovariant analysis (i.e. 0CFA), the analyzing result would be @x@ and 
-@y@ point to set @{1, 2}@, because the algorithm not distinguishes
-@z@ comes from different call-sites.
-Under $1$-CFA, the algorithm is able to distinguish that variable @z@ of function
-@id@ has two different values at two call-sites, so variable @y@ 
-would not be polluted by @1@. 
-But two call-sites still share the same continuation and
-it is still helplessness for separating the return-flows.
-TODO: do need more details to explain this example? (as in P4F paper)
-
-In this section, we firstly show a simple but uncomputable solution to the
-problem; then briefly review several recent research works that enable
-computable and precise call/return match. Even though the first solution is
-uncomputable, but as we will see in section 3, it is the starting point of our
-transformation.
-
-\subsubsection{AAM with Unbound Stack}
-
-A naive solution to is to use precise call stack as we did in concrete
-CESK machine. Instead of allocating continuations on store and recording an 
-address of continuations in state, we leave the call stack unabstracted.
-
-In the definition of state @State@, the continuation store disappears, and
-component @konts@ becomes a list of frames. Other components keep the same.
-
-\begin{lstlisting}
-case class State(e: Expr, env: Env, bstore: BStore, konts: List[Frame], time: Time)
-\end{lstlisting}
-
-The state transition function @step@ is still non-deterministic, but
-the only non-determinism happens when dereferencing the callee @f@ 
-from the store at call-site @App(f, ae)@. 
-For each closure of @f@ in the set, we do exactly
-the same as what we did in concrete CESK machines: a new frame is constructed
-and pushed onto the stack.
-In the second case, since the continuation @konts@ is a list of 
-frames, we can simply decompose it to differentiate an empty stack case
-or a frame exists on the top of stack. 
-If the stack is empty, then empty list of successors is returned, 
-because the computation is finished.
-Otherwise, we matched a frame @Frame(x, e, env_f)@ on the top of stack.
-The frame is then poped, and the new state uses the rest of stack.
-Since we can only access one frame at a time, this transition is deterministic.
-
-\begin{lstlisting}
-def step(s: State): List[State] = {
-  val new_time = tick(s)
-  s match {
-    case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
-      for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
-        val frame = Frame(x, e, env)
-        val baddr = allocBind(v, new_time)
-        val new_env = env_c + (v -> baddr)
-        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-        State(body, new_env, new_store, frame::konts, new_time)
-      }
-    case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
-      konts match {
-        case Nil => List()
-        case Frame(x, e, env_f)::konts =>
-          val baddr = allocBind(x, new_time)
-          val new_env = env_f + (x -> baddr)
-          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-          List(State(e, new_env, new_store, konts, new_time))
-      }
-  }
-}
-\end{lstlisting}
-
-Unfortunately, even though other components is finite, this AAM with unbounded stack is uncomputable. 
-Because the unbounded stack can grow to arbitrary deep and this implies that the state space is possibly 
-infinite, therefore the analysis may not terminate for all programs if we simply enumerating 
-reachable states.
-To see this, consider a program that has two mutually recursived functions:
-
-\begin{lstlisting}
-(letrec ([f1 (lambda (x) 
-               (let ([x1 (f2 x)]) x1))]
-         [f2 (lambda (y)
-               (let ([y1 (f1 y)]) y1))])
-  (let ([z (f1 1)])
-    z))
-\end{lstlisting}
-
-@f1@ and @f2@ mutually call each others, and the stack will be alternatively
-pushing @f1@ and @f2@ onto the top of current stack, but not existing 
-two stack components are identical in the state space.
-
-\subsubsection{Computable Solutions}
-
-In recent years, there are significant efforts \cite{vardoulakis2010cfa2, earl2012introspective, 
-gilray2016pushdown, johnson2015abstracting} to achieve precise call/return 
-match based on small-step AAM which is what we described in section \ref{aam}.
-
-CFA2 is the first solution that solves the return-flows problem\cite{vardoulakis2010cfa2}.
-But CFA2 has several drawbacks: it works only on continuation-passing style programs, 
-and only allows monovariant analysis, in addition with an exponential time complexity.
-TODO: summarize PDCFA, AAC, P4F
-
-In the rest of this section, we take a closer look at the P4F which is the
-state-of-art techniques to accomplish perfect call-stack precision on small-step
-AAMs, and Abstracting Defnitional Interpreters (ADI) that adpots another 
-approach to pushdown control flow analysis.
-
-\textbf{Pushdown for Free.}
-\citeauthor{gilray2016pushdown} proposed a continuation address allocator for 
-small-step abstracting abstract machines that is both simple to implement and
-computational inexpensive.
-The only change in code is when allocating a continuation address, not only
-keep track of the entry-point expression of callee, but also hold the target environment.
-The target environment is the binding environment extended from closure's environment
-(namely @val new_env = env_c + (v -> baddr)@). All the rest things remain the same.
-
-\begin{lstlisting}
-case class P4FContAddr(tgt: Expr, tgtEnv: Env) extends KAddr
-def allocKont(tgtExpr: Expr, tgtEnv: Env): KAddr = P4FContAddr(tgtExpr, tgtEnv)
-\end{lstlisting}
-
-TODO: explain the reason why it works.
-
-Notably, this change only causes constant-factor increase in time complexity of analysis
-if the store is widened.
-
-\textbf{Abstracting Definitional Interpreters.}
-On the other side, with the insight from Reynold's seminal paper 
-\emph{Definitional interpreters for higher-order programming languages} \cite{reynolds1972definitional}
-that the defined language will inherit properties
-from the defining language, \citeauthor{darais2017abstracting} construct abstracting 
-definitional interpreters that automatically inherit the pushdown control-flow 
-property from its defining language.
-
-TODO: connection to this paper.
-TODO: although they establish pushdown control-flow analysis in different ways, 
-is there a correspondence between different methods?
-
-\section{From AAM to ADI}
+\section{Fusing}
 
@@
 higher-order continuations, it does not matter what kind of AAM we start from.
-
-\subsection{Fusing}
 
@@
 
-\subsection{Disentangling}
+\section{Disentangling}
 
@@
 
-\subsection{Refunctionalization}
+\section{Refunctionalization}
 
@@
 
+\subsection{Pushdown Control-Flow Analysis}
+
+Pushdown control-flow is a property in analysis that precisely models
+the runtime call-stack of the analyzed program. 
+A pushdown control-flow analysis provides as exact as runtime return-flow 
+when analyzing the program, but traditional control-flow analysis collapses 
+the state space into a finite space thus causes imprecise stack modeling.
+
+To see how traditional CFA suffers from spurious return-flows, 
+we can consider the following example:
+
+\begin{lstlisting}
+(let ([id (lambda (z) z)])
+  (let ([x (id 1)])
+    (let ([y (id 2)])
+      x)))
+\end{lstlisting}
+
+In traditional $k$-CFA algorithm and the AAM we showed in section \ref{aam}, 
+the return-flow of invocation @(id 2)@
+is going to both call-sites @(id 1)@ and @(id 2)@. Therefore, the
+returned value @2@ for variable @y@ is also propagated to 
+variable @x@ and then imprecise analysis result is arose.
+This return-flow merging is inevitable even we increase the context-sensitivity.
+If we use the monovariant analysis (i.e. 0CFA), the analyzing result would be @x@ and 
+@y@ point to set @{1, 2}@, because the algorithm not distinguishes
+@z@ comes from different call-sites.
+Under $1$-CFA, the algorithm is able to distinguish that variable @z@ of function
+@id@ has two different values at two call-sites, so variable @y@ 
+would not be polluted by @1@. 
+But two call-sites still share the same continuation and
+it is still helplessness for separating the return-flows.
+TODO: do need more details to explain this example? (as in P4F paper)
+
+In this section, we firstly show a simple but uncomputable solution to the
+problem; then briefly review several recent research works that enable
+computable and precise call/return match. Even though the first solution is
+uncomputable, but as we will see in section 3, it is the starting point of our
+transformation.
+
+\subsubsection{AAM with Unbound Stack}
+
+A naive solution to is to use precise call stack as we did in concrete
+CESK machine. Instead of allocating continuations on store and recording an 
+address of continuations in state, we leave the call stack unabstracted.
+
+In the definition of state @State@, the continuation store disappears, and
+component @konts@ becomes a list of frames. Other components keep the same.
+
+\begin{lstlisting}
+case class State(e: Expr, env: Env, bstore: BStore, konts: List[Frame], time: Time)
+\end{lstlisting}
+
+The state transition function @step@ is still non-deterministic, but
+the only non-determinism happens when dereferencing the callee @f@ 
+from the store at call-site @App(f, ae)@. 
+For each closure of @f@ in the set, we do exactly
+the same as what we did in concrete CESK machines: a new frame is constructed
+and pushed onto the stack.
+In the second case, since the continuation @konts@ is a list of 
+frames, we can simply decompose it to differentiate an empty stack case
+or a frame exists on the top of stack. 
+If the stack is empty, then empty list of successors is returned, 
+because the computation is finished.
+Otherwise, we matched a frame @Frame(x, e, env_f)@ on the top of stack.
+The frame is then poped, and the new state uses the rest of stack.
+Since we can only access one frame at a time, this transition is deterministic.
+
+\begin{lstlisting}
+def step(s: State): List[State] = {
+  val new_time = tick(s)
+  s match {
+    case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
+      for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
+        val frame = Frame(x, e, env)
+        val baddr = allocBind(v, new_time)
+        val new_env = env_c + (v -> baddr)
+        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+        State(body, new_env, new_store, frame::konts, new_time)
+      }
+    case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
+      konts match {
+        case Nil => List()
+        case Frame(x, e, env_f)::konts =>
+          val baddr = allocBind(x, new_time)
+          val new_env = env_f + (x -> baddr)
+          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+          List(State(e, new_env, new_store, konts, new_time))
+      }
+  }
+}
+\end{lstlisting}
+
+Unfortunately, even though other components is finite, this AAM with unbounded stack is uncomputable. 
+Because the unbounded stack can grow to arbitrary deep and this implies that the state space is possibly 
+infinite, therefore the analysis may not terminate for all programs if we simply enumerating 
+reachable states.
+To see this, consider a program that has two mutually recursived functions:
+
+\begin{lstlisting}
+(letrec ([f1 (lambda (x) 
+               (let ([x1 (f2 x)]) x1))]
+         [f2 (lambda (y)
+               (let ([y1 (f1 y)]) y1))])
+  (let ([z (f1 1)])
+    z))
+\end{lstlisting}
+
+@f1@ and @f2@ mutually call each others, and the stack will be alternatively
+pushing @f1@ and @f2@ onto the top of current stack, but not existing 
+two stack components are identical in the state space.
+
+\subsubsection{Computable Solutions}
+
+In recent years, there are significant efforts \cite{vardoulakis2010cfa2, earl2012introspective, 
+gilray2016pushdown, johnson2015abstracting} to achieve precise call/return 
+match based on small-step AAM which is what we described in section \ref{aam}.
+
+CFA2 is the first solution that solves the return-flows problem\cite{vardoulakis2010cfa2}.
+But CFA2 has several drawbacks: it works only on continuation-passing style programs, 
+and only allows monovariant analysis, in addition with an exponential time complexity.
+TODO: summarize PDCFA, AAC, P4F
+
+In the rest of this section, we take a closer look at the P4F which is the
+state-of-art techniques to accomplish perfect call-stack precision on small-step
+AAMs, and Abstracting Defnitional Interpreters (ADI) that adpots another 
+approach to pushdown control flow analysis.
+
 %probably Store Widening \& Abstract Garbage Collection
 %monadic
-\subsection{Monadification: to Abstracting Definitional Interpreter}
+\section{Monadification: to Abstracting Definitional Interpreter}
 
@@
 
-\subsection{Un-CPS: to Yet Another Abstracting Definitional Interpreter}
+\section{Un-CPS: to Yet Another Abstracting Definitional Interpreter}
 
@@
 \section{Refunctionalized AAM}
+
+May be insert to section refunc.
 
@@
 
-\section{Detour: Shiver's \textit{k}-CFA?}
+%\section{Detour: Shiver's \textit{k}-CFA?}
 
 \section{Related Work}
+
+\textbf{Pushdown for Free.}
+\citeauthor{gilray2016pushdown} proposed a continuation address allocator for 
+small-step abstracting abstract machines that is both simple to implement and
+computational inexpensive.
+The only change in code is when allocating a continuation address, not only
+keep track of the entry-point expression of callee, but also hold the target environment.
+The target environment is the binding environment extended from closure's environment
+(namely @val new_env = env_c + (v -> baddr)@). All the rest things remain the same.
+
+\begin{lstlisting}
+case class P4FContAddr(tgt: Expr, tgtEnv: Env) extends KAddr
+def allocKont(tgtExpr: Expr, tgtEnv: Env): KAddr = P4FContAddr(tgtExpr, tgtEnv)
+\end{lstlisting}
+
+TODO: explain the reason why it works.
+
+Notably, this change only causes constant-factor increase in time complexity of analysis
+if the store is widened.
+
+\textbf{Abstracting Definitional Interpreters.}
+On the other side, with the insight from Reynold's seminal paper 
+\emph{Definitional interpreters for higher-order programming languages} \cite{reynolds1972definitional}
+that the defined language will inherit properties
+from the defining language, \citeauthor{darais2017abstracting} construct abstracting 
+definitional interpreters that automatically inherit the pushdown control-flow 
+property from its defining language.
+
+TODO: connection to this paper.
+TODO: although they establish pushdown control-flow analysis in different ways, 
+is there a correspondence between different methods?
 

### 2018-03-07 paper.tex
reorganize

@@
                                         %% contents suppressed with 'anonymous'
-\subtitle{Filling the Gap between Abstracting Abstract Machines and Abstracting Definitional Interpreters}        %% \subtitle is optional
+\subtitle{Or Filling the Gap of Abstracting Abstract Machines to Abstracting Definitional Interpreters}        %% \subtitle is optional
 \subtitlenote{with subtitle note}       %% \subtitlenote is optional;
@@
   continuations, we reveal how precise call/return match in control-flow 
-  analysis can be obtained by refunctionalizing a small-step abstracting 
-  abstract machine,
+  analysis can be obtained by refunctionalizing a small-step abstract interpreter.
   as well as explain how it been lost by defunctionalizing the higher-order
@@
   an small-step abstracting abstract machine. 
-  This abstract interpreter naturally has pushdown control-flow property
+  This abstract interpreter naturally enjoys pushdown control-flow property
   because the refunctionalization forces us to sequentialize the 
   abstract exectuion.
-  We also analyze its termination and soundness.
+  We also analyze its correctness, soundness and termination.
 

### 2018-03-09 paper.tex
title page

@@
 \acmJournal{PACMPL}
-\acmVolume{1}
+\acmVolume{2}
 \acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
@@
 \acmYear{2018}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\acmMonth{9}
+\acmDOI{10.1145/nnnnnnn.nnnnnnn}
 \startPage{1}
@@
 %% Title information
-\title[Short Title]{Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)}         %% [Short Title] is optional;
+\title{Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)}         %% [Short Title] is optional;
                                         %% when present, will be used in
                                         %% header instead of Full Title.
-\titlenote{with title note}             %% \titlenote is optional;
+%\titlenote{with title note}             %% \titlenote is optional;
                                         %% can be repeated if necessary;
                                         %% contents suppressed with 'anonymous'
-\subtitle{Or Filling the Gap of Abstracting Abstract Machines to Abstracting Definitional Interpreters}        %% \subtitle is optional
-\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
+\subtitle{Filling the Gap Between Abstracting Abstract Machines and Abstracting Definitional Interpreters}        %% \subtitle is optional
+%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                         %% can be repeated if necessary;
@@
 %% comma separated list
-\keywords{pushdown analysis, abstract interpretation, static analysis}  %% \keywords are mandatory in final camera-ready submission
+%\keywords{pushdown analysis, abstract interpretation, static analysis}  %% \keywords are mandatory in final camera-ready submission
 

### 2018-03-09 paper.tex
reorganize

--- 2018-03-07 paper.tex
@@
   The transformations include fusing, disentangling, refunctionalizing 
-  and monadification (or un-CPS to direct-style with delimited continuation). 
+  and monadification (or un-CPS to direct-style with delimited continuations). 
   All of them properly handle the 
@@
   Given this insight, we reveal that why small-step AAM has no pushdown
-  control-flow property (if without extra effort), as well as explain
+  control-flow property (if without extra efforts), as well as explain
   that why ADI naturally extends pushdown control-flow property from its
@@
 Although we just show the core calculus language, but it can be easily extended
-to support recursive bindings (such as $letrec$), conditionals, primitive types and 
-operations on primitive types. These cases would be trival or simple, so we omit them
-in the paper.
+to support recursive bindings (such as @letrec@), conditionals, primitive types and 
+operations on primitive types. These cases would be trival or simple to implement, 
+so we omit them in the paper.
 The transformations we will show in the rest of this paper
-also work on abstract machines for plain, direct-style $\lambda$-calculus languages.
-TODO: add letrec?
+also work on abstract machines for plain direct-style $\lambda$-calculus languages.
 
@@
 e $\in$ Exp ::= ae
-          | (let ([x (f ae)]) e)
+          | (let ([x (ae ae)]) e)
 ae $\in$ AExp ::= x            
@@
 
-In ANF, all the function applications must be administrated in a @let@ expression,
-and then bounded to a variable name under current environment.
-We can use @letrec@ to create mutually recursive function bindings. But @letrec@
-only accepts atomic expressions that appear at the right-hand side of a binding.
+In ANF, an expression is an atomic expression or a @let@ expression.
+The restriction is all the function applications must be administrated in a @let@ expression,
+and then bounded to a variable name under the current environment.
+Both the operator and operand of function applications should be an atomic expression.
 The atomic expressions $ae$ are either a variable, or a literal @lambda@ term, which
@@
 The abstract syntax can be straightforwardly described in Scala as follows:
+TODO: update the abstract syntax.
 
@@
   case Var(x) => store(env(x))
-  case lam@Lam(x, body) => Clos(lam, env)
+  case lam @ Lam(x, body) => Clos(lam, env)
 }
@@
 
-To run (analyze) a program, as CESK machine, we use function @inject@ to construct
-the initial state given a program. 
+To run (analyze) a program, as CESK machines, we first use function @inject@ to construct
+the initial state given the program. 
 Note that the initial continuation store has a built-in mapping that maps continuation address
@@
 collects all the intermediate machine states as the program is abstractly executing.
-The following code shows a variant of work list algorithm that always applies @step@
-function to the first element @hd@ of the work list @todo@ if @hd@
-is unseen, then adds the result of @step@ to the rest of work list and in 
-the meanwhile adds @hd@ to the explored states set.
-If the work list is empty, the function just returns the set of reachable states up to now.
+The following code shows a variant of work list algorithm to find the fixed-point 
+of set of states.
+Function @drive@ always applies function @step@ to the head element @hd@ of 
+the work list @todo@ if @hd@ is unseen, then inserts the result of @step@ to 
+the rest of work list; and in the meanwhile adds @hd@ to the explored states set.
+If the work list is empty, the @drive@ just returns the set of reachable states 
+up to now.
 
@@
 Finally, a user could use function @analyze@ to obtain all reachable states 
-for a given program.
+as a finite set for a given program.
 
@@
 
+\subsection{Pushdown Machine}
+
+AAM with unbounded stack
+
 \section{Fusing}
 
-We start from the AAM with unbounded stack.
-In fact, since later we will eliminate the call stack and transform to
-higher-order continuations, it does not matter what kind of AAM we start from.
+In last section, we showed that by keeping an unabstracted stack in the state 
+space we can recover the precise call/return match. In this section, we start
+our first transformation: \textit{fusing} an AAM with unbounded stack.
+
+In fact, later we will eliminate the stack and transform it to continuations
+represented by higher-order functions.
+Because of transforming to higher-order functions forces us to sequentialize the execution order,
+it does not matter what kind of AAM we start from. 
+We choose to start from AAM with unbounded stack just because TODO.
 
@@
 
-In recent years, there are significant efforts \cite{vardoulakis2010cfa2, earl2012introspective, 
-gilray2016pushdown, johnson2015abstracting} to achieve precise call/return 
-match based on small-step AAM which is what we described in section \ref{aam}.
-
-CFA2 is the first solution that solves the return-flows problem\cite{vardoulakis2010cfa2}.
-But CFA2 has several drawbacks: it works only on continuation-passing style programs, 
-and only allows monovariant analysis, in addition with an exponential time complexity.
-TODO: summarize PDCFA, AAC, P4F
-
 In the rest of this section, we take a closer look at the P4F which is the
@@
 
-\textbf{Pushdown for Free.}
+\textbf{Pushdown Control-flow Analysis}
+In recent years, there are significant efforts \cite{vardoulakis2010cfa2, earl2012introspective, 
+gilray2016pushdown, johnson2015abstracting} to achieve precise call/return 
+match based on small-step AAM which is what we described in section \ref{aam}.
+
+CFA2 is the first solution that solves the return-flows problem\cite{vardoulakis2010cfa2}.
+But CFA2 has several drawbacks: it works on only continuation-passing style programs, 
+and not supports polyvariant analysis, in addition with an exponential time complexity.
+TODO: summarize PDCFA, AAC
+
 \citeauthor{gilray2016pushdown} proposed a continuation address allocator for 
@@
 computational inexpensive. 
-The only change in code is when allocating a continuation address, not only
+Based on the AAM we presented in Section \ref{aam}, the only change in code is when 
+allocating a continuation address, not only
 keep track of the entry-point expression of callee, but also hold the target environment.
@@
 (namely @val new_env = env_c + (v -> baddr)@). All the rest things remain the same.
+The following code shows the implementation:
 
@@
 
-Notably, this change only causes constant-factor increase in time complexity of analysis
-if the store is widened.
+Notably, this change only causes constant-factor increase in time complexity of 
+the analysis if the store is widened.
+
+TOOD: why only needs the target env. correspondence?
 
 \textbf{Abstracting Definitional Interpreters.}
-On the other side, with the insight from Reynold's seminal paper 
+On the other side of small-step AAM, with the insight from Reynold's seminal paper 
 \emph{Definitional interpreters for higher-order programming languages} \cite{reynolds1972definitional}
-that the defined language will inherit properties
-from the defining language, \citeauthor{darais2017abstracting} construct abstracting 
+that the defined language will inherit properties from the defining language, 
+\citeauthor{darais2017abstracting} construct abstracting 
 definitional interpreters that automatically inherit the pushdown control-flow 
@@
 
+\citeauthor{darais2017abstracting}'s abstracting definitional interpreters is written
+in monadic style.
+
 TODO: connection to this paper.
+TODO: delimited control vs monad. correspondence since the very beginning time.
 TODO: although they establish pushdown control-flow analysis in different ways, 
 is there a correspondence between different methods?
+
+\textbf{Defunctionalization and Refunctionalization.}
 

### 2018-03-10 paper.tex
update

@@
 both can be evaluated in a single step.
+We also assume that the program is globally alpha-renamed, i.e., all the
+variable names are unique.
 
@@
 
-TODO
+A histroy of last $k$ call sites is served as time information in
+$k$-call-sensitive analysis.
+The history is represented as a list of expressions and embedded
+in the allocated addresses.
+
+Every time when we transfer to next states, we use function @tick@ to
+fresh the time and use this new time when allocating addresses. The function
+@tick@ returns the $k$ front-most expressions given the current state and its
+time history.
 
@@
 
+If we instantiate $k$ to be $0$, the history degenerates to an empty list,
+then we obtain a monovariant analysis, which
+not differentiates values at different call sites. In this case, the address
+space collapses to the space of variable names.
+Note that regarding to the ambiguitiy in
+$k$-CFA\cite{Gilray:2016:ACP:2951913.2951936},
+here the code actually implements a call+return sensitivity.
+
 \subsubsection{Collecting Semantics}
 
-To run (analyze) a program, as CESK machines, we first use function @inject@ to construct
+To run (analyze) a program, as the CESK machines, we first use function @inject@ to construct
 the initial state given the program. 
@@
 
+Transform to Extended Continuation-Passing Style (ECPS), continuation and meta-continuation.
 simply trun evaluation contexts into higher-order continuations.
@@
 \begin{lstlisting}
-case class Config(e: Expr, env: Env, store: BStore, time: Time)
-case class ValStore(ds: Set[Storable], store: BStore, time: Time)
-type Cache = Map[Config, ValStore]
-
-type Continue = Function2[ValStore, Cache, ValStore]
-
-def tick(s: Config): Time = (s.e::s.time).take(k)
-
-def aval(e: Expr, env: Env, bstore: BStore, time: Time, cache: Cache, continue: Continue): ValStore = { 
-  val config = Config(e, env, bstore, time)
-  if (cache.contains(config)) {
-    return continue(cache(config), cache)
-  }   
-  val newTime = tick(config)
+case class VS(vals: Set[Storable], time: Time, store: BStore)
+type MCont = VS => VS
+type Cont  = MCont => (VS => VS)
+
+def aval(e: Expr, env: Env, store: BStore, time: Time, cont: Cont, rcont: MCont): VS = {
+  val new_time = (e::time).take(k)
+
   e match {
-    case Let(x, App(f, ae), e) if isAtomic(ae) =>
-      val closures = aeval(f, env, bstore).toList
-
-      def nd(closures: List[Storable], accds: Set[Storable],
-             accbstore: BStore, acctime: Time, cache: Cache): ValStore  = { 
+    case Let(x, App(f, ae), e) =>
+      val closures = atomicEval(f, env, store).toList.asInstanceOf[List[Clos]]
+
+      def ndCont(closures: List[Clos])(r: MCont)(vs: VS): VS = {
         closures match {
-          case Nil =>
-            val baddr = allocBind(x, acctime)
-            val newEnv = env + (x -> baddr)
-            val newStore = accbstore.update(baddr, accds)
-            aval(e, newEnv, newStore, acctime, cache, (vs: ValStore, accCache: Cache) => {
-              val newCache = accCache + (Config(e, newEnv, newStore, acctime) -> vs) 
-              continue(vs, newCache)
+          case Nil => r(vs)
+          case Clos(Lam(v, body), c_env)::rest =>
+            val baddr = allocBind(v, vs.time)
+            val new_env = c_env + (v -> baddr)
+            val new_store = vs.store.update(baddr, atomicEval(ae, env, vs.store))
+            aval(body, new_env, new_store, new_time, ndCont(rest), (vs1: VS) => {
+                    r(VS(vs.vals ++ vs1.vals, vs1.time, vs1.store))
                   })
-          case Clos(Lam(v, body), env_c)::rest =>
-            val baddr = allocBind(v, newTime)
-            val newEnv = env_c + (v -> baddr)
-            val newStore = accbstore.update(baddr, aeval(ae, env, bstore)) //TODO: which store?
-            aval(body, newEnv, newStore, newTime, cache, (vs: ValStore, accCache: Cache) => {
-              val newCache = accCache + (Config(body, newEnv, newStore, newTime) -> vs) 
-              nd(rest, vs.ds++accds, accbstore, acctime, newCache)
-            })  
-        }   
-      }
-
-      nd(closures, Set(), bstore, newTime, cache)
+        }
+      }
+      def mtCont(vs: VS): VS = {
+        val baddr = allocBind(x, vs.time)
+        val new_env = env + (x -> baddr)
+        val new_store = vs.store.update(baddr, vs.vals)
+        aval(e, new_env, new_store, vs.time, cont, rcont)
+      }
+
+      ndCont(closures)(mtCont)(VS(Set(), new_time, store))
+
     case ae if isAtomic(ae) =>
-      val ds = aeval(ae, env, bstore)
-      continue(ValStore(ds, bstore, newTime), cache)
-  }   
-}
-
-def done: Continue = (vs: ValStore, cache: Cache) => { vs }
-def analyze(e: Expr): ValStore = aval(e, Map(), Store[BAddr, Storable](Map()),
-                                      List(), Map[Config, ValStore](), done)
+      val ds = atomicEval(ae, env, store)
+      cont(rcont)(VS(ds, new_time, store))
+  }
+}
+
+def analyze(e: Expr) = aval(e, Map(),
+                            Store[BAddr, Storable](Map()),
+                            List(),
+                            (r: MCont) => (vs: VS) => vs,
+                            (vs: VS) => vs)
+
 \end{lstlisting}
@@
 \begin{lstlisting}
-case class Config(e: Expr, env: Env)
-case class ValStore(ds: Set[Storable], store: BStore, time: Time)
-type Cache = Store[Config, ValStore]
-type RetType = (Set[ValStore], Cache)
-type Continue = Function5[Set[ValStore], BStore, Time, Cache, Cache, RetType]
-
-def tick(e: Expr, oldtime: Time): Time = (e::oldtime).take(k)
-
-def aval(e: Expr, env: Env, bstore: BStore, time: Time,
-         inCache: Cache, _outCache: Cache, continue: Continue): RetType = {
-  val config = Config(e, env)
-  if (_outCache.contains(config)) {
-    return continue(_outCache(config), bstore, List(), inCache, _outCache)
-  }
-  val outCache = _outCache.update(config, inCache.getOrElse(config, Set()))
-  val newTime = tick(e, time)
-  e match {
-    case Let(x, App(f, ae), e) if isAtomic(ae) =>
-      val closures = aeval(f, env, bstore).toList
-      def nd(closures: List[Storable], accds: Set[Storable], accbstore: BStore, acctime: Time,
-              accInCache: Cache, accOutCache: Cache): RetType  = closures match {
-        case Nil =>
-          // TODO: If we use `newTime` instead of `acctime`, is it m-CFA?
-          val baddr = allocBind(x, acctime)
-          val newEnv = env + (x -> baddr)
-          val newStore = accbstore.update(baddr, accds)
-          aval(e, newEnv, newStore, acctime, accInCache, accOutCache,
-            (vs: Set[ValStore], accStore: BStore, acctime: Time, accInCache: Cache, accOutCache: Cache) => {
-              val newOutCache = accOutCache.update(Config(e, newEnv), vs)
-              continue(vs, accStore, acctime, accInCache, newOutCache)
-            })
-        case Clos(Lam(v, body), env_c)::rest =>
-          val baddr = allocBind(v, newTime)
-          val newEnv = env_c + (v -> baddr)
-          val newStore = accbstore.update(baddr, aeval(ae, env, bstore)) //TODO: which store?
-          aval(body, newEnv, newStore, newTime, accInCache, accOutCache,
-            (vs: Set[ValStore], accStore: BStore, acctime: Time, accInCache: Cache, accOutCache) => {
-              val newOutCache = accOutCache.update(Config(body, newEnv), vs)
-              val vsds = vs.map(_.ds).foldLeft(Set[Storable]())(_ ++ _)
-              nd(rest, vsds++accds, accStore, acctime, accInCache, newOutCache)
-            })
-      }
-      nd(closures, Set(), bstore, newTime, inCache, outCache)
-    case ae if isAtomic(ae) =>
-      val ds = aeval(ae, env, bstore)
-      continue(Set(ValStore(ds, bstore, newTime)), bstore, newTime, inCache, outCache)
-    }
-}
-
-def done: Function2[Expr, Env, Continue] = (e: Expr, env: Env) => {
-  (vs: Set[ValStore], store: BStore, time: Time, inCache: Cache, outCache: Cache) => {
-    val newOutCache = outCache.update(Config(e, env), vs)
-    (vs, outCache)
-  }
-}
-
-def analyze1(e: Expr, cache: Cache): RetType = {
-  aval(e, Map(), Store[BAddr, Storable](Map()), List(), cache, cache, done(e, Map()))
-}
-
-def analyze(e: Expr): Set[ValStore] = {
-  def iter(cache: Cache): RetType = {
-    val (vs, newCache) = analyze1(e, cache)
-    if (cache == newCache) (vs, newCache)
-    else iter(newCache)
-  }
-  val (vs, cache) = iter(Store[Config, ValStore](Map()))
-  val summarized = cache.map.values.map(_.map(_.store).foldLeft(Store[BAddr, Storable](Map()))(_.join(_))).foldLeft(Store[BAddr, Storable](Map()))(_.join(_))
-  vs
-}
 \end{lstlisting}
@@
 
-\subsubsection{Computable Solutions}
-
-In the rest of this section, we take a closer look at the P4F which is the
-state-of-art techniques to accomplish perfect call-stack precision on small-step
-AAMs, and Abstracting Defnitional Interpreters (ADI) that adpots another 
-approach to pushdown control flow analysis.
-
 %probably Store Widening \& Abstract Garbage Collection
@@
 computational inexpensive. 
-Based on the AAM we presented in Section \ref{aam}, the only change in code is when 
-allocating a continuation address, not only
-keep track of the entry-point expression of callee, but also hold the target environment.
+Based on the AAM we presented in Section \ref{aam}, the only change in code is
+not only keeping track of the entry-point expression of callee, but also holding the target environment
+when allocating a continuation address.
 The target environment is the binding environment extended from closure's environment
-(namely @val new_env = env_c + (v -> baddr)@). All the rest things remain the same.
+(namely @val new_env = env_c + (v -> baddr)@). All the other part of the code remain the same.
 The following code shows the implementation:

### 2018-03-11 paper.tex
update

@@
 and simplifies the analysis \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
-Here, we choose to use administrative normal form (ANF) which is a direct-style intermediate representation 
-as target for clarity but without losing simplicity and generality.
+Here, we choose to use $\lambda$-calculus in administrative normal form (ANF) 
+which is a direct-style intermediate representation as target for clarity 
+but without losing simplicity and generality.
 Although we just show the core calculus language, but it can be easily extended
@@
 \subsubsection{Allocating Addresses}
-Up to now, we did not describe much about the adresses in stores and the 
+Up to now, we did not describe much about allocating adresses in stores and the 
 time-stamp @Time@.
-But they are key ingredients to achieve analyses with different sensitivities,
-as well as a finite state space analysis.
+But in abstract interpretation, they are key ingredients to achieve 
+analyses with different sensitivities, as well as to a finite state space analysis\cite{Gilray:2016:ACP:2951913.2951936}.
 To effectively approximate the runtime behavior, a finite 
 program contour @time@ that encodes the program execution history is introduced.
-Here, we use a list of execution contexts (expressions) to reprent that, and
+Here, we use a list of execution contexts (expressions) to represent that, and
 as we will see in section \ref{kcfainst}, by using different @tick@
@@
 
-As we mentioned previously, the space of addresses is finite in AAM.
+As we mentioned previously, the space of addresses has to be finite in AAM.
 Binding addresses are parameterized by variable names and the time that creates this binding,
@@
 
-\section{Fusing}
+\section{Fusing} \label{fusing}
 
@@
 
-\section{Disentangling}
+\section{Disentangling} \label{disen}
 
@@
 
-\section{Refunctionalization}
+\section{Refunctionalization} \label{refunc}
 
@@
 %probably Store Widening \& Abstract Garbage Collection
-%monadic
-\section{Monadification: to Abstracting Definitional Interpreter}
-
-For one direction, we can transform the refunctionalized AAM to abstracting definitional interpreter
-by adding monads.
-
-\section{Un-CPS: to Yet Another Abstracting Definitional Interpreter}
-
-We can transform the refunctionalized AAM to direct style, but explicitly use 
-side effects such as assignment and mutation to handle collecting semantics and non-determinism.
+\section{To Direct-Style}
+
+In previous section, we presented a refunctionalized AAM that in extended continuation-passing style. 
+To obtain a definitional abstract interpreter, we need to transform it further to direct-style. 
+At this point, we have two choices: 
+1) By monadifying the continuations into monads, we obtain the \citeauthor{darais2017abstracting}'s 
+abstracting definitional interpreter in monadic style.
+2) By representing the extended continuation-passing style with delimited continuations,
+we derive a new form of abstract interpreters in direct-style.
+This coincidence should not be a surprise since there are extraordinarily rich literatures 
+show the correspondence between monads and continuations as of their very beginning era
+(TODO: cite Danvy/abs control/esce of fp).
+In this section, we present both of them.
+
+\subsection{Monadification} \label{monadify}
+
+In this section, we show how to transform the refunctionalized AAM to monadic-style
+abstracting definitional interpreters.
+
+\subsection{Un-CPS} \label{uncps}
+
+On another path, here, we introduce direct-style abstracting definitional interpreter 
+with delimited control operators @reset@ and @shift@.
+
+collecting semantics and non-determinism.
 
 TODO: use delimited continuation?
-Another form of abstracting definitional interpreter.
 
@@
 
-\citeauthor{gilray2016pushdown} proposed a continuation address allocator for 
-small-step abstracting abstract machines that is both simple to implement and
+\citeauthor{gilray2016pushdown} proposed a polyvariant address allocator for continuations
+under small-step abstracting abstract machines that is both simple to implement and
 computational inexpensive. 
@@
 
+composable.
 TODO: connection to this paper.
-TODO: delimited control vs monad. correspondence since the very beginning time.
 TODO: although they establish pushdown control-flow analysis in different ways, 

### 2018-03-11 paper.tex
update abstract

@@
   Abstracting abstract machines (AAM) is a systematic methodology 
-  for constructing abstract interpreters that derived from concrete
+  for constructing abstract interpreters that are derived from concrete
   small-step abstract machines. Recent progress applies the same idea 
-  on definitional interpreter and obtains a big-step 
-  abstracting definitional interpreter (ADI) written in monadic style.
+  on definitional interpreter and obtains big-step 
+  abstracting definitional interpreters (ADI) written in monadic style.
   Danvy shows the correspondence between concrete abstract machines and interpreters.
@@
   abstract interpreter.
+
   Remarkably, refunctionalization sequentializes the abstract evaluation.
-  And follow the idea that evaluation contexts are defunctionalized 
+  Following the idea that evaluation contexts are defunctionalized 
   continuations, we reveal how precise call/return match in control-flow 
-  analysis can be obtained by refunctionalizing a small-step abstract interpreter.
-  as well as explain how it been lost by defunctionalizing the higher-order
-  continuations into an firts-order data type.
+  analysis can be obtained by properly refunctionalizing a small-step abstract interpreter.
+  As well as we explain how it been lost by defunctionalizing the higher-order
+  continuations into an first-order data type.
 

### 2018-03-11 paper.tex
update code

@@
 case class VS(vals: Set[Storable], time: Time, store: BStore)
-type MCont = VS => VS
-type Cont  = MCont => (VS => VS)
-
-def aval(e: Expr, env: Env, store: BStore, time: Time, cont: Cont, rcont: MCont): VS = {
+
+type MCont = VS => Set[VS]
+type Cont  = MCont => (VS => Set[VS])
+
+def aval(e: Expr, env: Env, store: BStore, time: Time, cont: Cont, rcont: MCont): Set[VS] = {
   val new_time = (e::time).take(k)
-
   e match {
     case Let(x, App(f, ae), e) =>
-      val closures = atomicEval(f, env, store).toList.asInstanceOf[List[Clos]]
-
-      def ndCont(closures: List[Clos])(r: MCont)(vs: VS): VS = {
-        closures match {
-          case Nil => r(vs)
-          case Clos(Lam(v, body), c_env)::rest =>
-            val baddr = allocBind(v, vs.time)
+      val closures = aeval(f, env, store).toList.asInstanceOf[List[Clos]]
+
+      def nd(closures: List[Clos]): Set[VS] = {
+        val Clos(Lam(v, body), c_env)::rest = closures
+        val baddr = allocBind(v, new_time)
         val new_env = c_env + (v -> baddr)
-            val new_store = vs.store.update(baddr, atomicEval(ae, env, vs.store))
-            aval(body, new_env, new_store, new_time, ndCont(rest), (vs1: VS) => {
-                    r(VS(vs.vals ++ vs1.vals, vs1.time, vs1.store))
-                  })
-        }
-      }
-      def mtCont(vs: VS): VS = {
+        val new_store = store.update(baddr, aeval(ae, env, store))
+        val new_cont = (r: MCont) => (vs: VS) => {
           val baddr = allocBind(x, vs.time)
@@
           val new_store = vs.store.update(baddr, vs.vals)
-        aval(e, new_env, new_store, vs.time, cont, rcont)
-      }
-
-      ndCont(closures)(mtCont)(VS(Set(), new_time, store))
+          aval(e, new_env, new_store, vs.time, cont, r)
+        }
+        val new_rcont = if (rest.isEmpty) {
+          rcont 
+        } else (vs: VS) => {
+          nd(rest) + vs
+        }
+        aval(body, new_env, new_store, new_time, new_cont, new_rcont)
+      }
+
+      nd(closures)
 
     case ae if isAtomic(ae) =>
-      val ds = atomicEval(ae, env, store)
+      val ds = aeval(ae, env, store)
       cont(rcont)(VS(ds, new_time, store))
@@
 def analyze(e: Expr) = aval(e, Map(),
-                            Store[BAddr, Storable](Map()),
-                            List(),
-                            (r: MCont) => (vs: VS) => vs,
-                            (vs: VS) => vs)
-
+                            Store[BAddr, Storable](Map()), List(),
+                            (r: MCont) => (vs: VS) =>  r(vs),
+                            (vs: VS) => Set(vs))
 \end{lstlisting}

### 2018-03-11 paper.tex
update

@@
 
-  Remarkably, refunctionalization sequentializes the abstract evaluation.
   Following the idea that evaluation contexts are defunctionalized 
-  continuations, we reveal how precise call/return match in control-flow 
+  continuations, we also found refunctionalization sequentializes the 
+  non-deterministic abstract evaluation.
+  Remarkably, we reveal how precise call/return match in control-flow 
   analysis can be obtained by properly refunctionalizing a small-step abstract interpreter.
@@
 
-\subsection{Pushdown Machine}
-
-AAM with unbounded stack
+\subsection{One Step Back: Unabstracted Stack}
+
+In this section, we describe a variant of AAM that leaves stack to be 
+unabstracted. That is to to use a precise call stack as we did in concrete CESK machine. 
+Instead of allocating continuations on store and embedding
+addresses of continuations in states, we intended to use a list of frames
+to explicitly model the stack.
+By doing so, we recover the call stack as precise as runtime 
+(so called pushdown control-flow analysis), but since 
+the stack is unbounded thus the analysis is possibly not computable.
+For readers who do not familiarized pushdown analysis, we have a detailed 
+discussion in Section \ref{pdcfarevisit}.
+The reason we do so here is to establish an artifact with precise call/return match
+used for the next step of transformation.
+
+In the definition of state @State@, the continuation store disappears, and
+component @konts@ becomes a list of frames. Other components keep the same.
+
+\begin{lstlisting}
+case class State(e: Expr, env: Env, bstore: BStore, konts: List[Frame], time: Time)
+\end{lstlisting}
+
+The state transition function @step@ is still non-deterministic, but
+the only non-determinism happens when dereferencing the callee @f@ 
+from function applications @App(f, ae)@.
+
+\begin{lstlisting}
+def step(s: State): List[State] = {
+  val new_time = tick(s)
+  s match {
+    case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
+      for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
+        val frame = Frame(x, e, env)
+        val baddr = allocBind(v, new_time)
+        val new_env = env_c + (v -> baddr)
+        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+        State(body, new_env, new_store, frame::konts, new_time)
+      }
+    case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
+      val Frame(x, e, env_f)::konts = konts
+      val baddr = allocBind(x, new_time)
+      val new_env = env_f + (x -> baddr)
+      val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+      List(State(e, new_env, new_store, konts, new_time))
+  }
+}
+\end{lstlisting}
+
+In the first case of state matching, we may have multiple choices of closure for @f@.
+For each closure in the set, a new frame is constructed and pushed onto the stack.
+The code for handling the second case is the same as concrete CESK machines.
+The function @drive@ should be also updated to prevent we pass a state with
+empty frame list into function @step@.
+
+\textbf{Note on Computability.}
+Unfortunately, even though other components in states are finite, 
+this AAM with unbounded stack is still not computable. 
+Because the unbounded stack can grow to arbitrary deep which implies that the state space is possibly 
+infinite, therefore the analysis may not terminate for all programs if we simply enumerating 
+reachable states.
+To see this, consider a program that has two mutually recursived functions:
+
+\begin{lstlisting}
+(letrec ([f1 (lambda (x) 
+               (let ([x1 (f2 x)]) x1))]
+         [f2 (lambda (y)
+               (let ([y1 (f1 y)]) y1))])
+  (let ([z (f1 1)])
+    z))
+\end{lstlisting}
+
+@f1@ and @f2@ mutually call each others, and the stack will be alternatively
+pushing @f1@ and @f2@ onto the top of current stack, but not existing 
+two stack components are identical in the state space.
 
@@
 space we can recover the precise call/return match. In this section, we start
-our first transformation: \textit{fusing} an AAM with unbounded stack.
-
-In fact, later we will eliminate the stack and transform it to continuations
-represented by higher-order functions.
-Because of transforming to higher-order functions forces us to sequentialize the execution order,
-it does not matter what kind of AAM we start from. 
-We choose to start from AAM with unbounded stack just because TODO.
+our first transformation: \textit{fusing} an AAM with unbounded-stack.
+
+In fact, as later we will eliminate the stack and transform it to continuations
+that are represented by higher-order functions, it does not matter what kind of 
+AAM we start from. 
+Because of transforming to higher-order functions forces us to sequentialize 
+the order of abstract evaluation.
+We choose to start from AAM with unbounded-stack just because it has a 
+straightforward correspondence of pushdown semantics to refunctionalized AAM.
 
@@
   case Nil => seen
-  case hd::tl if seen.contains(hd) => drive_step(tl, seen)
   case hd::tl =>
+    if (seen.contains(hd)) return drive_step(tl, seen)
+
     val new_time = tick(hd)
@@
 def aval(state: State, todo: List[State], seen: Set[State]): Set[State] = { 
+  if (seen.contains(state)) return collect(todo, seen)
+
   val newTime = tick(state)
@@
     case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
-      val newStates = for (Clos(Lam(v, body), c_env) <- aeval(f, env, bstore).toList) yield {
+      val newStates = for (Clos(Lam(v, body), c_env) <- atomicEval(f, env, bstore).toList) yield {
         val frame = Frame(x, e, env)
@@
         val newEnv = c_env + (v -> baddr)
-        val newStore = bstore.update(baddr, aeval(ae, env, bstore))
+        val newStore = bstore.update(baddr, atomicEval(ae, env, bstore))
         State(body, newEnv, newStore, frame::konts, newTime)
@@
     case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
-      continue(konts, aeval(ae, env, bstore), bstore, newTime, todo, seen + state)
+      continue(konts, atomicEval(ae, env, bstore), bstore, newTime, todo, seen + state)
   }   
@@
     case Nil => seen
-    case hd::tl if seen.contains(hd) => collect(tl, seen)
     case hd::tl => aval(hd, tl, seen)
@@
 
+As we pointed earlier, refunctionalizing evaluation contexts to higher-order
+continuation functions forces us to somehow sequentialize the abstract evaluation.
+We identify that in order to sequentialize abstract evaluation, there are three
+fundamental differences with concrete evaluation: 
+1) The dereferencing store addresses is non-deterministic, therefore the returning
+may also be non-deterministic.
+2) We need sound collecting semantics rather than valuation semantics.
+3) The abstract evaluation has to be terminating.
+
 allocating continuations of defined language on the heap of defining language.
 
-\subsubsection{First Try}
+\subsection{First Try}
 
@@
     case Let(x, App(f, ae), e) =>
-      val closures = aeval(f, env, store).toList.asInstanceOf[List[Clos]]
+      val closures = atomicEval(f, env, store).toList.asInstanceOf[List[Clos]]
 
@@
         val new_env = c_env + (v -> baddr)
-        val new_store = store.update(baddr, aeval(ae, env, store))
+        val new_store = store.update(baddr, atomicEval(ae, env, store))
         val new_cont = (r: MCont) => (vs: VS) => {
@@
     case ae if isAtomic(ae) =>
-      val ds = aeval(ae, env, store)
+      val ds = atomicEval(ae, env, store)
       cont(rcont)(VS(ds, new_time, store))
@@
 
-\subsubsection{Caching Fixed-Points}
+\subsection{Caching Fixed-Points}
 
@@
 
-\subsection{Pushdown Control-Flow Analysis}
+\subsection{Pushdown Control-Flow Analysis, Revisited} \label{pdcfarevisit}
 
@@
 TODO: do need more details to explain this example? (as in P4F paper)
-
-In this section, we firstly show a simple but uncomputable solution to the
-problem; then briefly review several recent research works that enable
-computable and precise call/return match. Even though the first solution is
-uncomputable, but as we will see in section 3, it is the starting point of our
-transformation.
-
-\subsubsection{AAM with Unbound Stack}
-
-A naive solution to is to use precise call stack as we did in concrete
-CESK machine. Instead of allocating continuations on store and recording an 
-address of continuations in state, we leave the call stack unabstracted.
-
-In the definition of state @State@, the continuation store disappears, and
-component @konts@ becomes a list of frames. Other components keep the same.
-
-\begin{lstlisting}
-case class State(e: Expr, env: Env, bstore: BStore, konts: List[Frame], time: Time)
-\end{lstlisting}
-
-The state transition function @step@ is still non-deterministic, but
-the only non-determinism happens when dereferencing the callee @f@ 
-from the store at call-site @App(f, ae)@. 
-For each closure of @f@ in the set, we do exactly
-the same as what we did in concrete CESK machines: a new frame is constructed
-and pushed onto the stack.
-In the second case, since the continuation @konts@ is a list of 
-frames, we can simply decompose it to differentiate an empty stack case
-or a frame exists on the top of stack. 
-If the stack is empty, then empty list of successors is returned, 
-because the computation is finished.
-Otherwise, we matched a frame @Frame(x, e, env_f)@ on the top of stack.
-The frame is then poped, and the new state uses the rest of stack.
-Since we can only access one frame at a time, this transition is deterministic.
-
-\begin{lstlisting}
-def step(s: State): List[State] = {
-  val new_time = tick(s)
-  s match {
-    case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
-      for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
-        val frame = Frame(x, e, env)
-        val baddr = allocBind(v, new_time)
-        val new_env = env_c + (v -> baddr)
-        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-        State(body, new_env, new_store, frame::konts, new_time)
-      }
-    case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
-      konts match {
-        case Nil => List()
-        case Frame(x, e, env_f)::konts =>
-          val baddr = allocBind(x, new_time)
-          val new_env = env_f + (x -> baddr)
-          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-          List(State(e, new_env, new_store, konts, new_time))
-      }
-  }
-}
-\end{lstlisting}
-
-Unfortunately, even though other components is finite, this AAM with unbounded stack is uncomputable. 
-Because the unbounded stack can grow to arbitrary deep and this implies that the state space is possibly 
-infinite, therefore the analysis may not terminate for all programs if we simply enumerating 
-reachable states.
-To see this, consider a program that has two mutually recursived functions:
-
-\begin{lstlisting}
-(letrec ([f1 (lambda (x) 
-               (let ([x1 (f2 x)]) x1))]
-         [f2 (lambda (y)
-               (let ([y1 (f1 y)]) y1))])
-  (let ([z (f1 1)])
-    z))
-\end{lstlisting}
-
-@f1@ and @f2@ mutually call each others, and the stack will be alternatively
-pushing @f1@ and @f2@ onto the top of current stack, but not existing 
-two stack components are identical in the state space.
 

### 2018-03-12 paper.tex
update

@@
 Here, we use a list of execution contexts (expressions) to represent that, and
-as we will see in section \ref{kcfainst}, by using different @tick@
-function on time we may obtain a family of analyses.
+as we will see in section \ref{kcfainst}, by applying different @tick@
+functions to time we may obtain a family of analyses.
 
@@
 As we mentioned previously, the space of addresses has to be finite in AAM.
-Binding addresses are parameterized by variable names and the time that creates this binding,
+Binding addresses are parameterized by variable names and the time when creating this binding,
 which are both finite.
@@
 
-\subsubsection{$k$-call-sensitive Instantiation} \label{kcfainst}
+\subsubsection{$k$-Call-Sensitive Instantiation} \label{kcfainst}
 
@@
 
-The state transition function @step@ is still non-deterministic, but
-the only non-determinism happens when dereferencing the callee @f@ 
+The state transition function @step@ shows in below which is still 
+non-deterministic, but the only non-determinism happens when dereferencing the callee @f@ 
 from function applications @App(f, ae)@.
@@
 
-@f1@ and @f2@ mutually call each others, and the stack will be alternatively
-pushing @f1@ and @f2@ onto the top of current stack, but not existing 
-two stack components are identical in the state space.
+Function @f1@ and @f2@ mutually invokes each others, and the stack will be 
+alternatively pushing frame @f1@ and @f2@ onto the top of current stack, 
+but not existing two stack components are identical in the state space.
 
@@
 
-type MCont = VS => Set[VS]
-type Cont  = MCont => (VS => Set[VS])
-
-def aval(e: Expr, env: Env, store: BStore, time: Time, cont: Cont, rcont: MCont): Set[VS] = {
+type Ans = Set[VS]
+type NCont = VS => Ans
+type Cont = VS => (NCont => Ans)
+
+def aval(e: Expr, env: Env, store: BStore, time: Time, cont: Cont, ncont: NCont): Ans = {
   val new_time = (e::time).take(k)
@@
     case Let(x, App(f, ae), e) =>
-      val closures = atomicEval(f, env, store).toList.asInstanceOf[List[Clos]]
-
-      def nd(closures: List[Clos]): Set[VS] = {
+      val closures = aeval(f, env, store).toList.asInstanceOf[List[Clos]]
+
+      def nd(closures: List[Clos]): Ans = {
         val Clos(Lam(v, body), c_env)::rest = closures
@@
         val new_env = c_env + (v -> baddr)
-        val new_store = store.update(baddr, atomicEval(ae, env, store))
-        val new_cont = (r: MCont) => (vs: VS) => {
+        val new_store = store.update(baddr, aeval(ae, env, store))
+        val new_cont = (vs: VS) => (r: NCont) => {
           val baddr = allocBind(x, vs.time)
@@
         }
-        val new_rcont = if (rest.isEmpty) {
-          rcont 
+        val new_ncont = if (rest.isEmpty) {
+          ncont
         } else (vs: VS) => {
@@
         }
-        aval(body, new_env, new_store, new_time, new_cont, new_rcont)
+        // new_time is determined by the lexical scope
+        aval(body, new_env, new_store, new_time, new_cont, new_ncont)
       }
@@
       nd(closures)
-
     case ae if isAtomic(ae) =>
-      val ds = atomicEval(ae, env, store)
-      cont(rcont)(VS(ds, new_time, store))
-  }
-}
-
-def analyze(e: Expr) = aval(e, Map(),
-                            Store[BAddr, Storable](Map()), List(),
-                            (r: MCont) => (vs: VS) =>  r(vs),
+      val ds = aeval(ae, env, store)
+      cont(VS(ds, new_time, store))(ncont)
+  }
+}
+
+def mtEnv = Map[String, BAddr]()
+def mtStore = Store[BAddr, Storable](Map())
+def analyze(e: Expr) = aval(e, env, store, List(), 
+                            (vs: VS) => (r: NCont) => r(vs), 
                             (vs: VS) => Set(vs))
+
 \end{lstlisting}

### 2018-03-14 paper.tex
code

@@
 case class VS(vals: Set[Storable], time: Time, store: BStore)
-
-type Ans = Set[VS]
-type NCont = VS => Ans
-type Cont = VS => (NCont => Ans)
-
-def aval(e: Expr, env: Env, store: BStore, time: Time, cont: Cont, ncont: NCont): Ans = {
+type NCont = Set[VS] => Set[VS]
+type Cont  = NCont => (Set[VS] => Set[VS])
+
+def aval(e: Expr, env: Env, store: BStore, time: Time,
+          cont: Cont, ncont: NCont): Set[VS] = {
   val new_time = (e::time).take(k)
@@
 
-      def nd(closures: List[Clos]): Ans = {
+      def new_ncont(ret_vss: Set[VS])(result: Set[VS]): Set[VS] = {
+        val VS(vals, time, store) = ret_vss.head
+        val baddr = allocBind(x, time)
+        val new_env = env + (x -> baddr)
+        val new_store = store.update(baddr, vals)
+        aval(e, new_env, new_store, time, (r: NCont) => r, (vss1: Set[VS]) => {
+          if (ret_vss.tail.isEmpty) cont(ncont)(result ++ vss1)
+          else new_ncont(ret_vss.tail)(result ++ vss1)
+        })
+      }
+
+      def nd(closures: List[Clos], vss_acc: Set[VS]): Set[VS] = {
         val Clos(Lam(v, body), c_env)::rest = closures
@@
         val new_store = store.update(baddr, aeval(ae, env, store))
-        val new_cont = (vs: VS) => (r: NCont) => {
-          val baddr = allocBind(x, vs.time)
-          val new_env = env + (x -> baddr)
-          val new_store = vs.store.update(baddr, vs.vals)
-          aval(e, new_env, new_store, vs.time, cont, r)
-        }
-        val new_ncont = if (rest.isEmpty) {
-          ncont
-        } else (vs: VS) => {
-          nd(rest) + vs
-        }
-        // new_time is determined by the lexical scope
-        aval(body, new_env, new_store, new_time, new_cont, new_ncont)
-      }
-
-      nd(closures)
+        aval(body, new_env, new_store, new_time, (r: NCont) => r, (vss: Set[VS]) => {
+          if (rest.isEmpty) new_ncont(vss_acc ++ vss)(Set())
+          else nd(rest, vss_acc ++ vss)
+        })
+      }
+
+      nd(closures, Set())
+
     case ae if isAtomic(ae) =>
       val ds = aeval(ae, env, store)
-      cont(VS(ds, new_time, store))(ncont)
-  }
-}
-
+      cont(ncont)(Set(VS(ds, new_time, store)))
+  }
+}
+
+def mtTime = List()
 def mtEnv = Map[String, BAddr]()
 def mtStore = Store[BAddr, Storable](Map())
-def analyze(e: Expr) = aval(e, env, store, List(), 
-                            (vs: VS) => (r: NCont) => r(vs), 
-                            (vs: VS) => Set(vs))
-
+
+def analyze(e: Expr) = aval(e, mtEnv, mtStore, mtTime, 
+                            (r: NCont) => (vss: Set[VS]) =>  r(vss), 
+                            (vss: Set[VS]) => vss)
 \end{lstlisting}

### 2018-03-14 paper.tex
code

@@
 (TODO: cite Danvy/abs control/esce of fp).
-In this section, we present both of them.
+In fact, we can also just use the same caching algorithm but with side effect such as 
+assignments and mutations to achieve the same definitional interpreter with pushdown
+control flows.
+In this section, we present both of versions using monads and delimited controls.
 
@@
 
-collecting semantics and non-determinism.
-
-TODO: use delimited continuation?
+TODO: This version does not use delimited controls.
+
+\begin{lstlisting}
+def aval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): (Set[VS], Cache) = {
+  val config = Config(e, env, time)
+  if (cache.outContains(config)) return (cache.outGet(config), cache)
+
+  val new_cache = cache.outUpdate(config, cache.inGet(config))
+  val new_time = (e::time).take(k)
+
+  e match {
+    case Let(x, App(f, ae), e) =>
+      val closures = aeval(f, env, store).toList.asInstanceOf[List[Clos]]
+      val app_vss_cache = for (Clos(Lam(v, body), c_env) <- closures) yield {
+        val baddr = allocBind(v, new_time)
+        val new_env = c_env + (v -> baddr)
+        val new_store = store.update(baddr, aeval(ae, env, store))
+        aval(body, new_env, new_store, new_time, new_cache) //TODO: Do we have to use the latest cache?
+      }
+
+      val app_vss = app_vss_cache.map(_._1).foldLeft(Set[VS]())(_ ++ _)
+      val app_cache = app_vss_cache.map(_._2).foldLeft(mtCache)(_.outJoin(_))
+
+      val e_vss = for (VS(vals, time, store) <- app_vss.toList) yield {
+        val baddr = allocBind(x, time)
+        val new_env = env + (x -> baddr)
+        val new_store = store.update(baddr, vals)
+        aval(e, new_env, new_store, time, app_cache) //TODO: Do we have to use the latest cache?
+      }
+      val result = e_vss.map(_._1).foldLeft(Set[VS]())(_ ++ _)
+      (result, e_vss.map(_._2).foldLeft(mtCache)(_.outJoin(_)).outUpdate(config, result))
+
+    case ae if isAtomic(ae) =>
+      val ds = aeval(ae, env, store)
+      (Set(VS(ds, new_time, store)), new_cache.outUpdate(config, VS(ds, new_time, store)))
+  }
+}
+
+def analyze(e: Expr) = {
+  def iter(cache: Cache): (Set[VS], Cache) = {
+    val (vss, new_cache) = aval(e, mtEnv, mtStore, mtTime, cache)
+    if (new_cache.out == cache.out) (vss, new_cache)
+    else iter (Cache(new_cache.out, new_cache.out))
+  }
+  iter(mtCache)
+}
+\end{lstlisting}
 

### 2018-03-14 paper.tex
fix grammar mistakes

@@
   small-step abstract machines. Recent progress applies the same idea 
-  on definitional interpreter and obtains big-step 
-  abstracting definitional interpreters (ADI) written in monadic style.
+  on definitional interpreters and obtains big-step 
+  abstract definitional interpreters (ADI) written in monadic style.
   Danvy shows the correspondence between concrete abstract machines and interpreters.
@@
   In this paper, we show their correspondence and how to syntactically transform small-step 
-  abstracting abstract machines into big-step abstracting definitional 
+  abstract abstract machines into big-step abstract definitional 
   interpreters.
@@
   collecting semantics and the non-determinism of abstract interpretation. 
-  After each transformation, we also obtain an intermediary form of 
+  After each transformation, we also obtain an intermediate form of 
   abstract interpreter.
@@
   Following the idea that evaluation contexts are defunctionalized 
-  continuations, we also found refunctionalization sequentializes the 
+  continuations, we also found that refunctionalization sequentializes the 
   non-deterministic abstract evaluation.
-  Remarkably, we reveal how precise call/return match in control-flow 
+  Remarkably, we reveal how precise call/return matching in control-flow 
   analysis can be obtained by properly refunctionalizing a small-step abstract interpreter.
@@
 \begin{itemize}
-\item The main contribution of this paper is we fill the gap
+\item The main contribution of this paper is filling the gap
   between small-step AAM and big-step ADI by developing a series of
@@
         &  &  &  & 
-      \begin{smallmatrix} \text{abstracting} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
-      \begin{smallmatrix} \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
-      \begin{smallmatrix} \text{big-step} \\ \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  & 
-      \begin{smallmatrix} \text{disentangled}  \\ \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
-      \begin{smallmatrix} \text{refunctionalized}  \\ \text{abstracting} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
+      \begin{smallmatrix} \text{abstract} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
+      \begin{smallmatrix} \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
+      \begin{smallmatrix} \text{big-step} \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  & 
+      \begin{smallmatrix} \text{disentangled}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
+      \begin{smallmatrix} \text{refunctionalized}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} & 
       \begin{smallmatrix} \text{monadic} \\ \text{abstracting} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
@@
 
+\iffalse 
 \subsection{Style}
@@
 interpreters.
+\fi
 
@@
 Traditionally, continuation-passing style (CPS) is a popular intermediate representation
-for analyzing functional programs because of it exposes control transfer explicitly
+for analyzing functional programs because it exposes control transfer explicitly
 and simplifies the analysis \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
@@
 but without losing simplicity and generality.
-Although we just show the core calculus language, but it can be easily extended
+Although we just show the core calculus language, it can be easily extended
 to support recursive bindings (such as @letrec@), conditionals, primitive types and 
-operations on primitive types. These cases would be trival or simple to implement, 
+operations on primitive types. These cases would be trivial to implement, 
 so we omit them in the paper.
@@
 In ANF, an expression is an atomic expression or a @let@ expression.
-The restriction is all the function applications must be administrated in a @let@ expression,
-and then bounded to a variable name under the current environment.
-Both the operator and operand of function applications should be an atomic expression.
+The restriction is that all the function applications must be administrated in a @let@ expression,
+and then bound to a variable name under the current environment.
+Both the operator and operand of function applications are an atomic expressions.
 The atomic expressions $ae$ are either a variable, or a literal @lambda@ term, which
 both can be evaluated in a single step.
-We also assume that the program is globally alpha-renamed, i.e., all the
-variable names are unique.
+We also assume that all the variable names in the program are unique.
 
@@
 
-\textit{C}ontrol \textit{E}nvionment \textit{S}tore \textit{K}ontinuation (CESK) machine is an
+The CESK machine is an
 abstract machine for describing semantics of and evaluating $\lambda$-calculus \cite{felleisen1987calculus}.
-CESK machine models program execution as state transitions in a small-step fashion. As its name suggested,
-a machine state has four components: 1) Control is the expression currently being evaluated.
-2) Environment is the a map that contains the address of a variable that in the lexical
-scope of control.
-3) Store models the heap of a program, and it is also a map but maps from addresses to values.
-The addresses are infinite numbers starting from 0.
-In our toy language, the only category of value is closure, i.e., a function paired with an environment.
-4) Continuation represents the program stack.
+The CESK machine models program execution as state transitions in a small-step fashion. 
+As its name suggested, a machine state has four components: 
+1) \textit{Control} is the expression currently being evaluated.
+2) \textit{Environment} is a map that contains the address of a variable in the lexical scope.
+3) \textit{Store} models the heap of a program. It maps from addresses to values.
+The addresses space is numbers starting from 0.
+In our toy language, the only category of value is the closure, i.e., a function paired with an environment.
+4) \textit{Continuation} represents the program stack. \todo{explain more on stack}
 
@@
 It is worth noting that the continuation class @Kont@ is defined as a list of frames.
-The frame class @Frame@ stores the information of call-site, i.e., the information that
-can be used to resume the interrupted computation.
+The frame class @Frame@ stores the information of a single call-site, 
+i.e., the information that can be used to resume the interrupted computation.
 A @Frame@ constitutes of a variable name @x@ to be bind later, a control expression
@@
 \subsubsection{Single-step Transition}
-Before go into describing how the machine evaluates expressions, we firstly define several helper functions.
+Before going into describing how the machine evaluates expressions, we firstly define several helper functions.
 As we mentioned in section \ref{anfsyntax}, the atomic expressions are either a variable, or
@@
 cases and evaluates atomic expressions to closures in a straightforward way.
-The @alloc@ function generates a fresh address, and always allocates an unique integer under the domain
-of store.
-The @isAtomic@ function is used as a predicate of whether the expression is atomic.
+The @alloc@ function generates a fresh address, and always allocates unique integer 
+in the domain of store.
+The @isAtomic@ function is used as a predicate of the expression is atomic.
 
@@
 Now, we can faithfully describe the state transition function @step@,
-that given a machine state, function @step@ determines its successor state.
+that given a machine state, determines its successor state.
 
@@
 
-The first case is that the control of current state is a @Let@ expression,
+In the first case, the control of current state is a @Let@ expression,
 and its right-hand side is an function application.
@@
 \subsection{Abstracting Abstract Machine} \label{aam}
-Abstracting abstract machine (AAM) is a systematic methodology that derives sound 
+Abstracting abstract machines (AAM) is a systematic methodology that derives sound 
 abstract interpreters for higher-order functional languages from concrete 
@@
 Since the state space of concrete execution is possibly infinite,
-so to analyze programs, the key insight of AAM approach is to allocate both bindings
+to analyze programs, the key insight of AAM approach is to allocate both bindings
 and continuations on the store, and then bound the addresses space to be finite.
 Since each component of state is finite, 
-thus the abstracted machine-state space is finite, and computable.
+the abstracted machine-state space is finite, and computable.
 
 In this section, we derive the abstracting abstract machine from concrete
-CESK machines, and also show that how to instantiate
+CESK machines, and also show how to instantiate
 useful $k$-call-sensitive control-flow analysis.
@@
 
-Similar with CESK machines, the machine state of AAM has a control expression,
+Similar to CESK machines, the machine state of AAM has a control expression,
 an environment, a store and continuation, and in addition with a time-stamp.
-But there are several notable differences of store between AAMs and concrete CESK machines.
+But there are several notable differences between AAM's store and CESK machine's store.
 In AAM, the store maps addresses to sets of values, so it stores all possible values
-for a particular address. Thus the dereference an address becomes non-deterministic.
-Also, the store performs a join rather than overwriting when updating elements of the store.
+for a particular address. Thus dereferencing addresses becomes non-deterministic.
+Also, the store performs joining \todo{?} rather than overwriting when updating elements of the store.
 Third, the continuations are likewise allocated on store instead of formed a linked list,
@@
 
-For clarity, we divide the store into binding store @BStore@ and continuation store @KStore@. 
-The binding store maps binding addresses to sets of closure values, and continuation store maps
+For clarity, we divide the store into a binding store @BStore@ and a continuation store @KStore@. 
+The binding store maps binding addresses to sets of closure values, and the continuation store maps
 continuation addresses to sets of continuations.
-We define a generic class @Store[K,V]@ that performs join when updating elements
+We define a generic class @Store[K,V]@ that performs joining when updating elements
 in a store. By parameterizing @Store[K,V]@ with @[BAddr, Storable]@ and 
@@
 
-As a consequence, the components of a state is also changed: 
+As a consequence, the components of states are also changed: 
 the store is divided into binding store and continuation store;
 the continuation stack becomes an address.
-By dereferencing this address in continuation store, we can retrieve the 
+By dereferencing this address in a continuation store, we can retrieve the 
 actual transfer of controls. The definition of environment @Env@
@@
 \begin{lstlisting}
-case class State(e: Expr, env: Env, bstore: BStore, 
-                 kstore: KStore, k: KAddr, time: Time)
+case class State(e: Expr, env: Env, bstore: BStore, kstore: KStore, k: KAddr, time: Time)
 \end{lstlisting}
@@
 \subsubsection{Allocating Addresses}
-Up to now, we did not describe much about allocating adresses in stores and the 
-time-stamp @Time@.
+Up to now, we did not describe about allocating adresses in stores and the time stamp @Time@.
 But in abstract interpretation, they are key ingredients to achieve 
 analyses with different sensitivities, as well as to a finite state space analysis\cite{Gilray:2016:ACP:2951913.2951936}.
-To effectively approximate the runtime behavior, a finite 
-program contour @time@ that encodes the program execution history is introduced.
-Here, we use a list of execution contexts (expressions) to represent that, and
+To effectively approximate the runtime behavior, we introduce a finite 
+program contour @time@ that encodes the program execution history.
+Here, we use a list of execution contexts (expressions) to represent, and
 as we will see in section \ref{kcfainst}, by applying different @tick@
-functions to time we may obtain a family of analyses.
+functions on time stamp we may obtain a family of analyses.
 
@@
 As we mentioned previously, the space of addresses has to be finite in AAM.
-Binding addresses are parameterized by variable names and the time when creating this binding,
+Binding addresses are parameterized by variable names and the time of creating this binding,
 which are both finite.
@@
 1) @Halt@ which corresponds to the empty stack, and
-2) @ContAddr@ consists of the entry expressions of callee, which are also also
+2) @ContAddr@ consists of the entry expressions of callee, which are also
 finite.
@@
 
-We use two helper functions @allocBind@ and @allocKont@ that will be used later
+Two helper functions @allocBind@ and @allocKont@ that will be used later
 to allocate binding addresses and continuation addresses.
@@
 Since dereferencing an address becomes non-deterministic, our @atomicEval@
-function also truns into non-deterministic. Given an atomic expression @e@,
+function also truns to be non-deterministic. Given an atomic expression @e@,
 @atomicEval@ returns a set of storable values, i.e., closures, to the caller.
-If the expression is just a @lambda@ term, then the returned set is singleton.
+If the expression is just a @lambda@ term, then the returned set is a singleton.
 
@@
 
-The structure of function @step@ is similar with concrete CESK machines, 
-except non-determinism which makes @step@ returns a list of reachable 
-successor states.
-In the beginning of @step@, we call function @tick@ to update 
-the time.
-Then we have two cases to consider: 1) Current control expression is @Let@,
-and @App(f, ae)@ is being bind to @x@.
-Fristly we retrieve the set of closures that @f@ may represents.
-For each closure in the set, we do almost the same as what we did in concrete 
+The structure of function @step@ is similar to the concrete CESK machine, 
+except non-determinism which makes @step@ return a list of reachable successor states.
+In the beginning of @step@, we call function @tick@ to update the time.
+Then we have two cases to consider: 1) The current control expression is @Let@,
+and then the result of @App(f, ae)@ will be bound to variable @x@.
+Firstly we retrieve the set of closures that @f@ may represents.
+For each closure in the set, we do almost the same as what we did in the concrete 
 CESK machines. But note that since the continuation is allocated on store (@kstore@),
@@
 
-A histroy of last $k$ call sites is served as time information in
+A histroy of the last $k$ call sites is used as a finite program contour in
 $k$-call-sensitive analysis.
@@
 
-Every time when we transfer to next states, we use function @tick@ to
-fresh the time and use this new time when allocating addresses. The function
-@tick@ returns the $k$ front-most expressions given the current state and its
+Every time before transferring to successor states, we use function @tick@ to
+fresh the time stamp and use this new time stamp for successors when allocating addresses. 
+The function @tick@ returns the $k$ front-most expressions given the current state and its
 time history.
@@
 If we instantiate $k$ to be $0$, the history degenerates to an empty list,
-then we obtain a monovariant analysis, which
-not differentiates values at different call sites. In this case, the address
+and we obtain a monovariant analysis, which does
+not differentiate values at different call sites. In this case, the address
 space collapses to the space of variable names.
-Note that regarding to the ambiguitiy in
-$k$-CFA\cite{Gilray:2016:ACP:2951913.2951936},
-here the code actually implements a call+return sensitivity.
+Note that regarding the ambiguitiy in $k$-CFA\cite{Gilray:2016:ACP:2951913.2951936},
+the code here actually implements call+return sensitivity.
 
@@
 
-To run (analyze) a program, as the CESK machines, we first use function @inject@ to construct
-the initial state given the program. 
+To run (analyze) a program, as same as the CESK machines, 
+we first use function @inject@ to construct the initial state given the program. 
 Note that the initial continuation store has a built-in mapping that maps continuation address
-@Halt@ to empty set of continuations.
+@Halt@ to an empty set of continuations.
 We also provide an empty program contour as initial time.
@@
 
-However, comparing with concrete CESK machine, the
-@drive@ function performs collecting semantics instead of valuation
+However, comparing with the concrete CESK machine, the
+@drive@ function performs collecting semantics instead of the valuation
 semantics. That is, for the purpose of analyzing programs, the function @drive@
 collects all the intermediate machine states as the program is abstractly executing.
-The following code shows a variant of work list algorithm to find the fixed-point 
+The following code shows a variant of the work list algorithm to find the fixed-point 
 of set of states.
@@
 the work list @todo@ if @hd@ is unseen, then inserts the result of @step@ to 
-the rest of work list; and in the meanwhile adds @hd@ to the explored states set.
-If the work list is empty, the @drive@ just returns the set of reachable states 
+the rest of work list; and in the meantime adds @hd@ to the explored states set.
+If the work list is empty, then @drive@ just returns the set of reachable states 
 up to now.
@@
 
-Finally, a user could use function @analyze@ to obtain all reachable states 
-as a finite set for a given program.
+Finally, a user could use function @analyze@ to obtain all reachable states for a given program.
 
@@
 
-In this section, we describe a variant of AAM that leaves stack to be 
-unabstracted. That is to to use a precise call stack as we did in concrete CESK machine. 
-Instead of allocating continuations on store and embedding
-addresses of continuations in states, we intended to use a list of frames
+In this section, we describe a variant of AAM that leaves the stack to be 
+unbounded. That uses a precise call stack as we did in the concrete CESK machine. 
+Instead of allocating continuations in the store and embedding
+addresses of continuations in states, we intend to use a list of frames
 to explicitly model the stack.
-By doing so, we recover the call stack as precise as runtime 
+By doing so, we recover the call stack as precise as runtime's
 (so called pushdown control-flow analysis), but since 
-the stack is unbounded thus the analysis is possibly not computable.
-For readers who do not familiarized pushdown analysis, we have a detailed 
+the stack is unbounded, the analysis is potentially not computable.
+For readers who are not familiar with pushdown analysis, we have a detailed 
 discussion in Section \ref{pdcfarevisit}.
-The reason we do so here is to establish an artifact with precise call/return match
+The reason we show it here is to establish an artifact with precise call/return match
 used for the next step of transformation.
@@
 In the definition of state @State@, the continuation store disappears, and
-component @konts@ becomes a list of frames. Other components keep the same.
+component @konts@ becomes a list of frames. The other components keep the same.
 
@@
 
-The state transition function @step@ shows in below which is still 
+The state transition function @step@ shows below is still 
 non-deterministic, but the only non-determinism happens when dereferencing the callee @f@ 
@@
 
-In the first case of state matching, we may have multiple choices of closure for @f@.
+In the first case of pattern matching, we may have multiple choices of closure for callee @f@.
 For each closure in the set, a new frame is constructed and pushed onto the stack.
-The code for handling the second case is the same as concrete CESK machines.
-The function @drive@ should be also updated to prevent we pass a state with
+The code for handling the second case is the same as the concrete CESK machines.
+The function @drive@ is also updated (but not shown here) to prevent passing a state with
 empty frame list into function @step@.
@@
 \textbf{Note on Computability.}
-Unfortunately, even though other components in states are finite, 
+Unfortunately, even though, other components in states are finite, 
 this AAM with unbounded stack is still not computable. 
-Because the unbounded stack can grow to arbitrary deep which implies that the state space is possibly 
-infinite, therefore the analysis may not terminate for all programs if we simply enumerating 
-reachable states.
-To see this, consider a program that has two mutually recursived functions:
+Because the unbounded stack can grow to arbitrary depth which implies that the state space is possibly 
+infinite, therefore the analysis may not terminate for all programs if we simply enumerate reachable states.
+To see this, consider a program that has two mutually recursive functions:
 
@@
 
-Function @f1@ and @f2@ mutually invokes each others, and the stack will be 
+Function @f1@ and @f2@ mutually invokes each other, and the stack will be 
 alternatively pushing frame @f1@ and @f2@ onto the top of current stack, 
-but not existing two stack components are identical in the state space.
+but no existing two stack components are identical in the state space.
 
@@
 space we can recover the precise call/return match. In this section, we start
-our first transformation: \textit{fusing} an AAM with unbounded-stack.
+our first step of transformation: \textit{fusing} an AAM with unbounded-stack.
 
@@
 AAM we start from. 
-Because of transforming to higher-order functions forces us to sequentialize 
+Because transforming to higher-order functions forces us to sequentialize 
 the order of abstract evaluation.

### 2018-03-15 paper.tex
update

@@
 for plain direct-style, the continue function handles more cases.
+\todo{a defunctionalized object for non-determinism}
 
@@
 \begin{lstlisting}
-case class VS(vals: Set[Storable], time: Time, store: BStore)
-type NCont = Set[VS] => Set[VS]
-type Cont  = NCont => (Set[VS] => Set[VS])
-
-def aval(e: Expr, env: Env, store: BStore, time: Time,
-          cont: Cont, ncont: NCont): Set[VS] = {
+def nd[T,S](ts: Set[T], acc: S, f: (T, S, S=>S) => S, g: S => S): S = {
+  if (ts.isEmpty) g(acc)
+  else f(ts.head, acc, (vss: S) => nd(ts.tail, vss, f, g))
+}
+
+type Ans = Set[VS]
+type Cont = Ans => Ans
+
+def aval(e: Expr, env: Env, store: BStore, time: Time, cont: Cont): Ans = {
   val new_time = (e::time).take(k)
@@
     case Let(x, App(f, ae), e) =>
-      val closures = aeval(f, env, store).toList.asInstanceOf[List[Clos]]
-
-      def new_ncont(ret_vss: Set[VS])(result: Set[VS]): Set[VS] = {
-        val VS(vals, time, store) = ret_vss.head
+      val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
+      nd[Clos, Ans](closures, Set[VS](), { case (clos, acc, closnd) =>
+        val Clos(Lam(v, body), c_env) = clos
+        val baddr = allocBind(v, new_time)
+        val new_env = c_env + (v -> baddr)
+        val new_store = store.update(baddr, atomicEval(ae, env, store)) 
+        aval(body, new_env, new_store, new_time, (bodyvss: Set[VS]) => {
+          nd[VS, Ans](bodyvss, Set[VS](), { case (vs, acc_vss, bdnd) => 
+            val VS(vals, time, store) = vs
+            val new_env = env + (x -> baddr)
+            val new_store = store.update(baddr, vals)
+            aval(e, new_env, new_store, time, (evss: Ans) => bdnd(acc_vss ++ evss))
+          },
+          (evss: Ans) => closnd(evss ++ acc))
+        })
+      }, 
+      cont)
+    case ae if isAtomic(ae) => 
+      val ds = atomicEval(ae, env, store)
+      cont(Set(VS(ds, new_time, store))) 
+  } 
+}
+
+def mtTime = List()
+def mtEnv = Map[String, BAddr]()
+def mtStore = Store[BAddr, Storable](Map())
+
+def analyze(e: Expr) =
+  aval(e, mtEnv, mtStore, mtTime, (ans => ans))
+\end{lstlisting}
+
+\subsection{Caching Fixed-Points}
+
+fix point cache
+
+\begin{lstlisting}
+case class Config(e: Expr, env: Env, store: BStore, time: Time)
+case class Cache(in: Store[Config, VS], out: Store[Config, VS]) {
+  def inGet(config: Config): Set[VS] = in.getOrElse(config, Set())
+  def outGet(config: Config): Set[VS] = out.getOrElse(config, Set())
+  def outContains(config: Config): Boolean = out.contains(config)
+  def outUpdate(config: Config, vss: Set[VS]): Cache = Cache(in, out.update(config, vss))
+}
+\end{lstlisting}
+
+adding cache to eval
+
+The definition of Ans is changed to a case class that contains a set of @VS@ and
+a cache.
+\begin{lstlisting}
+case class Ans(vss: Set[VS], cache: Cache)
+\end{lstlisting}
+
+\begin{lstlisting}
+def aval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, cont: Cont): Ans = {
+  val config = Config(e, env, store, time)
+  if (cache.outContains(config)) return cont(Ans(cache.outGet(config), cache))
+
+  val new_cache = cache.outUpdate(config, cache.inGet(config))
+  val new_time = (e::time).take(k)
+
+  e match {
+    case Let(x, App(f, ae), e) =>
+      val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
+      nd[Clos, Ans](closures, Ans(Set[VS](), new_cache), { case (clos, Ans(acc, cache), closnd) =>
+        val Clos(Lam(v, body), c_env) = clos
+        val baddr = allocBind(v, new_time)
+        val new_env = c_env + (v -> baddr)
+        val new_store = store.update(baddr, atomicEval(ae, env, store))
+        aval(body, new_env, new_store, new_time, cache, { case Ans(bodyvss, bodycache) =>
+          nd[VS, Ans](bodyvss, Ans(Set[VS](), bodycache), { case (vs, Ans(acc_vss, cache), bdnd) =>
+            val VS(vals, time, store) = vs
             val baddr = allocBind(x, time)
@@
             val new_store = store.update(baddr, vals)
-        aval(e, new_env, new_store, time, (r: NCont) => r, (vss1: Set[VS]) => {
-          if (ret_vss.tail.isEmpty) cont(ncont)(result ++ vss1)
-          else new_ncont(ret_vss.tail)(result ++ vss1)
+            aval(e, new_env, new_store, time, cache, { case Ans(evss, ecache) => bdnd(Ans(acc_vss ++ evss, ecache)) })
+          },
+          { case Ans(evss, cache) => closnd(Ans(evss ++ acc, cache)) })
         })
-      }
-
-      def nd(closures: List[Clos], vss_acc: Set[VS]): Set[VS] = {
-        val Clos(Lam(v, body), c_env)::rest = closures
-        val baddr = allocBind(v, new_time)
-        val new_env = c_env + (v -> baddr)
-        val new_store = store.update(baddr, aeval(ae, env, store))
-        aval(body, new_env, new_store, new_time, (r: NCont) => r, (vss: Set[VS]) => {
-          if (rest.isEmpty) new_ncont(vss_acc ++ vss)(Set())
-          else nd(rest, vss_acc ++ vss)
+      },
+      { case Ans(vss, cache) => cont(Ans(vss, cache.outUpdate(config, vss))) })
+
+    case ae if isAtomic(ae) =>
+      val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
+      cont(Ans(vs, new_cache.outUpdate(config, vs)))
+  }
+}
+def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore) = {
+  def iter(cache: Cache): Ans = {
+    val Ans(vss, new_cache) = aval(e, env, store, mtTime, cache, { 
+      val initConfig = Config(e, env, store, mtTime)
+      case Ans(vss, cache) => Ans(vss, cache.outUpdate(initConfig, vss)) 
     })
-      }
-
-      nd(closures, Set())
-
-    case ae if isAtomic(ae) =>
-      val ds = aeval(ae, env, store)
-      cont(ncont)(Set(VS(ds, new_time, store)))
-  }
-}
-
-def mtTime = List()
-def mtEnv = Map[String, BAddr]()
-def mtStore = Store[BAddr, Storable](Map())
-
-def analyze(e: Expr) = aval(e, mtEnv, mtStore, mtTime, 
-                            (r: NCont) => (vss: Set[VS]) =>  r(vss), 
-                            (vss: Set[VS]) => vss)
-\end{lstlisting}
-
-\subsection{Caching Fixed-Points}
-
-fix point cache
-
-\begin{lstlisting}
+    if (new_cache.out == cache.out) { Ans(vss, new_cache) }
+    else { iter(Cache(new_cache.out, new_cache.out)) }
+  }    
+  iter(mtCache)
+}
 \end{lstlisting}
@@
 \begin{lstlisting}
-def aval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): (Set[VS], Cache) = {
-  val config = Config(e, env, time)
-  if (cache.outContains(config)) return (cache.outGet(config), cache)
-
-  val new_cache = cache.outUpdate(config, cache.inGet(config))
+def aval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans = {
+  val config = Config(e, env, store, time)
+  if (cache.outContains(config)) {
+    return Ans(cache.outGet(config), cache)
+  }
+
+  var new_cache = cache.outUpdate(config, cache.inGet(config))
   val new_time = (e::time).take(k)
-
   e match {
@@
       val closures = aeval(f, env, store).toList.asInstanceOf[List[Clos]]
-      val app_vss_cache = for (Clos(Lam(v, body), c_env) <- closures) yield {
+      val letvs = for (Clos(Lam(v, body), c_env) <- closures) yield {
         val baddr = allocBind(v, new_time)
@@
         val new_store = store.update(baddr, aeval(ae, env, store))
-        aval(body, new_env, new_store, new_time, new_cache) //TODO: Do we have to use the latest cache?
-      }
-
-      val app_vss = app_vss_cache.map(_._1).foldLeft(Set[VS]())(_ ++ _)
-      val app_cache = app_vss_cache.map(_._2).foldLeft(mtCache)(_.outJoin(_))
-
-      val e_vss = for (VS(vals, time, store) <- app_vss.toList) yield {
+        val Ans(bdv, bdcache) = aval(body, new_env, new_store, new_time, new_cache)
+        new_cache = bdcache
+        val evs = for (VS(vals, time, store) <- bdv) yield {
           val baddr = allocBind(x, time)
@@
           val new_store = store.update(baddr, vals)
-        aval(e, new_env, new_store, time, app_cache) //TODO: Do we have to use the latest cache?
-      }
-      val result = e_vss.map(_._1).foldLeft(Set[VS]())(_ ++ _)
-      (result, e_vss.map(_._2).foldLeft(mtCache)(_.outJoin(_)).outUpdate(config, result))
+          val Ans(ev, ecache) = aval(e, new_env, new_store, time, new_cache)
+          new_cache = ecache
+          ev
+        }
+        evs.foldLeft(Set[VS]())(_ ++ _)
+      }
+      val letv = letvs.foldLeft(Set[VS]())(_ ++ _)
+      Ans(letv, new_cache.outUpdate(config, letv))
 
     case ae if isAtomic(ae) =>
-      val ds = aeval(ae, env, store)
-      (Set(VS(ds, new_time, store)), new_cache.outUpdate(config, VS(ds, new_time, store)))
-  }
-}
-
-def analyze(e: Expr) = {
-  def iter(cache: Cache): (Set[VS], Cache) = {
-    val (vss, new_cache) = aval(e, mtEnv, mtStore, mtTime, cache)
-    if (new_cache.out == cache.out) (vss, new_cache)
-    else iter (Cache(new_cache.out, new_cache.out))
-  }
-  iter(mtCache)
+      val vs = Set(VS(aeval(ae, env, store), new_time, store))
+      Ans(vs, new_cache.outUpdate(config, vs))
+  }
 }

### 2018-03-15 paper.tex
misc

@@
 \begin{lstlisting}
-def nd[T,S](ts: Set[T], acc: S, f: (T, S, S=>S) => S, g: S => S): S = {
+def nd[T,S](ts: Set[T], acc: S, f: (T, S, S => S) => S, g: S => S): S = {
   if (ts.isEmpty) g(acc)

### 2018-03-15 paper.tex
writing

@@
 We choose to start from AAM with unbounded-stack just because it has a 
-straightforward correspondence of pushdown semantics to refunctionalized AAM.
+straightforward correspondence of pushdown semantics to the refunctionalized AAM.
 
@@
 
-fix point cache
+We will apply the fixed-point caching algorithm as described in
+\citeauthor{darais2017abstracting}'s abstract definitional interpreter.
 
@@
 
-\subsection{Higher-Order Controls?}
-
-TODO: analyze programs with call/cc, shift, reset.
-
 \subsection{Precise Call/Return Match}
@@
 
-\section{Perspective}
-
-\subsection{Defunctionalization}
-
-specify higher-order definitional interpreters using first-order means (cite)
-
-\subsection{Refunctionalization}
-
+%\section{Perspective}
+%\subsection{Defunctionalization}
+%specify higher-order definitional interpreters using first-order means (cite)
+%\subsection{Refunctionalization}
 %\section{Detour: Shiver's \textit{k}-CFA?}
@@
 
-\textbf{Pushdown Control-flow Analysis}
+\textbf{Abstracting Abstract Machines.}
+
+\textbf{Defunctionalization and Refunctionalization.}
+
+\textbf{Pushdown Control-flow Analysis.}
 In recent years, there are significant efforts \cite{vardoulakis2010cfa2, earl2012introspective, 
 gilray2016pushdown, johnson2015abstracting} to achieve precise call/return 
-match based on small-step AAM which is what we described in section \ref{aam}.
+match based on small-step abstracting abstract machines.
 
@@
 But CFA2 has several drawbacks: it works on only continuation-passing style programs, 
-and not supports polyvariant analysis, in addition with an exponential time complexity.
+and not supports polyvariant analysis, in addition with an exponential time
+complexity.
+
 TODO: summarize PDCFA, AAC
 
-\citeauthor{gilray2016pushdown} proposed a polyvariant address allocator for continuations
-under small-step abstracting abstract machines that is both simple to implement and
-computational inexpensive. 
-Based on the AAM we presented in Section \ref{aam}, the only change in code is
+\citeauthor{gilray2016pushdown} proposed a polyvariant continuation-addresses allocator
+for small-step AAM to achieve pushdown analysis \cite{gilray2016pushdown}.
+The method is both simple to implement and computational inexpensive.
+Based on the AAM we presented in Section \ref{aam}, the only change in the code is
 not only keeping track of the entry-point expression of callee, but also holding the target environment
-when allocating a continuation address.
-The target environment is the binding environment extended from closure's environment
-(namely @val new_env = env_c + (v -> baddr)@). All the other part of the code remain the same.
+when allocating a continuation addresses.
 The following code shows the implementation:
@@
 
-TODO: explain the reason why it works.
-
+The target environment is used for successor states and extended from closure's
+environment (namely @val new_env = env_c + (v -> baddr)@).
+All the other part of the code would not be changed.
 Notably, this change only causes constant-factor increase in time complexity of 
@@
 
-TOOD: why only needs the target env. correspondence?
-
 \textbf{Abstracting Definitional Interpreters.}
-On the other side of small-step AAM, with the insight from Reynold's seminal paper 
-\emph{Definitional interpreters for higher-order programming languages} \cite{reynolds1972definitional}
-that the defined language will inherit properties from the defining language, 
-\citeauthor{darais2017abstracting} construct abstracting 
-definitional interpreters that automatically inherit the pushdown control-flow 
-property from its defining language.
-
-\citeauthor{darais2017abstracting}'s abstracting definitional interpreters is written
-in monadic style.
-
-composable.
-TODO: connection to this paper.
-TODO: although they establish pushdown control-flow analysis in different ways, 
-is there a correspondence between different methods?
-
-\textbf{Defunctionalization and Refunctionalization.}
+Reynold's seminal paper \emph{Definitional interpreters for higher-order
+  programming languages} \cite{reynolds1972definitional} shows that in
+definitional interpreters the defined language can inherit properties from the defining language.
+With this insight, \citeauthor{darais2017abstracting} construct abstracting 
+definitional interpreters that automatically inherits the pushdown control-flow 
+property from its defining language
+because the defined language just uses the stack the meta-language \cite{darais2017abstracting}.
+\citeauthor{darais2017abstracting}'s abstracting definitional interpreters work
+on direct-style programs, and is written in monadic style. One of the advantages
+of monadic abstract interpreters is modular and composable, therefore deploying
+different sensitivities or features is just applying a different monad.
+Prior to that, \citeauthor{Sergey:2013:MAI:2491956.2491979} show a monadic
+abstract interpreter for small-step semantics.
+
+This paper is inspired a lot from \citeauthor{darais2017abstracting}'s work.
+Instead of just showing the final form of abstract definitional interpreters,
+We reveal that refunctionalization plays an important role for inheriting
+stack from the meta-language.
+The target of the series of transformations in this paper is abstract definitional
+interpreters, and we additionaly demonstrate abstract interpreters written in both
+monadic- and direct-style.
+The difference is ours work for A-Normal Form $\lambda$-calculus instead of plain
+$\lambda$-calculus, our work also can be easily adopted to plain
+$\lambda$-calculus.
 

### 2018-03-16 paper.tex
update

@@
     [start chain] \chainin (m-2-1);
-    \chainin (m-2-2) [join={node[above,labeled] {\text{fusion}}}];
+    \chainin (m-2-2) [join={node[above,labeled] {\begin{smallmatrix} \text{linearizing} \\ \text{+} \\ \text{fusing} \end{smallmatrix}}     }];
     \chainin (m-2-3) [join={node[above,labeled] {\text{disentangling}}}];
@@
 
-It is worth noting that the continuation class @Kont@ is defined as a list of frames.
+It is worth noting that the continuation class @Kont@ is defined as a list of
+frames.
+The frame can be considered as evaluation context in reduction semantics.
 The frame class @Frame@ stores the information of a single call-site, 
@@
 for a particular address. Thus dereferencing addresses becomes non-deterministic.
-Also, the store performs joining \todo{?} rather than overwriting when updating elements of the store.
+Also, the store performs joining rather than overwriting when updating elements
+of the store.
 Third, the continuations are likewise allocated on store instead of formed a linked list,
@@
 In the beginning of @step@, we call function @tick@ to update the time.
-Then we have two cases to consider: 1) The current control expression is @Let@,
+Then we have two cases to consider:
+
+1) The current control expression is @Let@,
 and then the result of @App(f, ae)@ will be bound to variable @x@.
-Firstly we retrieve the set of closures that @f@ may represents.
+Firstly we retrieve the set of closures that @f@ may represent.
 For each closure in the set, we do almost the same as what we did in the concrete 
@@
   s match {
-    case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) 
-      if isAtomic(ae) =>
+    case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) =>
       val closures = atomicEval(f, env, bstore).toList
@@
       }
-    case State(ae, env, bstore, kstore, kaddr, time) 
-      if isAtomic(ae) =>
+    case State(ae, env, bstore, kstore, kaddr, time) if isAtomic(ae) =>
       val conts = kstore(kaddr).toList
@@
     case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
-      val Frame(x, e, env_f)::konts = konts
+      konts match {
+        case Nil => List()
+        case Frame(x, e, env_f)::konts =>
           val baddr = allocBind(x, new_time)
@@
   }
+}
 \end{lstlisting}
@@
 The code for handling the second case is the same as the concrete CESK machines.
-The function @drive@ is also updated (but not shown here) to prevent passing a state with
-empty frame list into function @step@.
 
@@
 
+\section{Linearization} \label{linear}
+
+In last section, we showed that by keeping an unabstracted stack in the state 
+space we can recover the precise call/return match.
+From this section, we start describing the transformations step by step.
+Our base machine is the AAM with unbounded stack.
+In fact, as later we will transform the stack to higher-order functions
+represented continuations, it does not matter what kind of AAM we start from. 
+Because transforming to higher-order functions forces us to sequentialize 
+the order of abstract evaluation, and neither need to construct a frame of list
+nor to allocate continuations in store.
+We choose to start from AAM with unbounded stack just because it has a 
+equivalence of stack model to the final abstract definitional interpreters.
+
+\todo{Unbounded Stack AAM is PD, justify this}
+
+In Danvy's paper \textit{Defunctionalized Interpreters for Programming
+  Languages}, he metioned for deterministic languages, a reduction semantics
+is a structural operational semantics in continuation style, where the reduction
+context is a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}.
+Here we are facing a problem:
+the underlying semantics of AAM is fundamentally non-deterministic.
+But the evaluation context (the @Frame@ in our program) is not
+non-deterministic, and the work list @todo@ actually implicitly handles the non-determinism,
+if we simply refunctionalize the frames to functions, it would not helps toward
+abstract definitional interpreters.
+
+In order to address this problem, our first step of tramsformations
+is to linearize all the non-deterministic choices.
+This step turns the @step@ function to be deterministic.
+Likewise the @Frame@ saves the information of the caller in concrete executions,
+we introduce another layer of controls, and define a case class @NDCont@
+that saves the information at a fork point when we have multiple target closures.
+We also add a new field @ndk@ to the definition of state which is called
+@NDState@ now.
+
+\begin{lstlisting}
+case class NDCont(cls: List[Clos], argvs: Set[Storable], store: BStore, time: Time, frames: List[Frame])
+case class NDState(e: Expr, env: Env, bstore: BStore, konts: List[Frame], time: Time, ndk: List[NDCont]) {
+  def toState: State = State(e, env, bstore, konts, time)
+}
+\end{lstlisting}
+
+A @NDCont@ contains a list of closures that are possible functions to be invoked,
+a set of values that will be bind to the function's formal argument,
+and the store, time, and frames at the fork point.
+In the definition of @NDState@, an auxiliary function @toState@ is added that
+converts itself to @State@.
+
+Now the type of function @step@ becomes @NDState => NDState@.
+The following code shows the first case of matching an instance of @NDState@ in
+the @step@.
+
+\begin{lstlisting}
+case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
+  val closures = atomicEval(f, env, bstore).toList.asInstanceOf[List[Clos]]
+  val Clos(Lam(v, body), c_env) = closures.head
+  val frame = Frame(x, e, env)
+  val new_frames = frame::konts
+  val baddr = allocBind(v, new_time)
+  val new_env = c_env + (v -> baddr)
+  val argvs = atomicEval(ae, env, bstore)
+  val new_store = bstore.update(baddr, argvs)
+  val new_ndk = NDCont(closures.tail, argvs, bstore, new_time, new_frames)::ndk
+  NDState(body, new_env, new_store, new_frames, new_time, new_ndk)
+\end{lstlisting}
+
+By atomically evaluating @f@, we obtain a set of closures.
+The transition is deterministic to the first element of the closure set.
+So we just prepare new environment, new store, and new frame list only for the
+first closure in that set.
+A new non-deterministic continuation @new_ndk@ is also constructed,
+which contains the rest of closures, the values of argument @ae@, the store,
+the time and the new frame list.
+We use the store before updating because for different closures they may form
+different binding addresses @baddr@.
+We also use the new frames, because all the closures at this fork point share
+the same stack and return point.
+
+In the second case shown in below we will see how this @NDCont@ controls
+the non-determinism.
+
+\begin{lstlisting}
+case NDState(ae, env, bstore, konts, time, ndk) if isAtomic(ae) =>
+  konts match {
+    case Nil => ndk match {
+      case NDCont(Nil, _, _, _, _)::ndk => 
+        NDState(ae, env, bstore, konts, time, ndk) /* transfer to the next fork point */
+      case NDCont(cls, argvs, bstore, time, frames)::ndk => 
+        val Clos(Lam(v, body), c_env) = cls.head
+        val baddr = allocBind(v, time)
+        val new_env = c_env + (v -> baddr)
+        val new_store = bstore.update(baddr, argvs)
+        val new_ndk = NDCont(cls.tail, argvs, bstore, tile, frames)::ndk
+        /* resume the fork point with the next closure */
+        NDState(body, new_env, new_store, frames, time, new_ndk) 
+    }
+    case Frame(x, e, f_env)::konts =>
+      val baddr = allocBind(x, newTime)
+      val new_env = f_env + (x -> baddr)
+      val new_store = bstore.update(baddr, aeval(ae, env, bstore))
+      NDState(e, new_env, new_store, konts, newTime, ndk) /* normal return */
+  }
+\end{lstlisting}
+
+If the normal continuation @konts@ is an empty list, then it means we have reached
+the halt of computation; otherwise, we should return the values of @ae@ to its
+caller which is contained in the top frame of stack.
+
+But, since we have added non-determinism into state, an empty list of frames means
+we have the halt of \textit{one computation path}, and we should determine the next state
+by @NDCont@.
+So we have a pattern matching on @ndk@: 1) If the closure set is an empty set,
+then we had tried all the closures of this fork point and should move the the
+next fork point. 2) Otherwise, we pop a closure from the set, then build new
+environment and new store for the body expression of the closure;
+we use the same frame list and time that are copied from the fork point rather
+than current one.
+The non-determinism continuation @new_ndk@ is also updated by removal of that closure.
+
+\begin{lstlisting}
+def drive(nds: NDState, seen: Set[State]): Set[State] = {
+  nds match {
+    case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
+    case nds =>
+      val s = nds.toState
+      if (seen.contains(s)) drive(step(nds), seen)
+      else drive(step(nds), seen + s)
+  }
+}
+\end{lstlisting}
+
+The function @drive@ is also changed: there is no work list anymore, because all the
+potentially unexplored states are embedded in to the continuation for
+non-determinism.
+The termination of the analysis is when we reach a @NDState@ that the expression
+is atomic, and both the normal continuation and non-determinism continuation are
+empty list. This corresponds to the case that @todo@ list is empty.
+
+Now, we obtain a linearized abstract abstract machine. If we image that the classical
+AAM explores a graph of reachable states, then the linearized AAM flatten this
+graph.
+
 \section{Fusing} \label{fusing}
 
-In last section, we showed that by keeping an unabstracted stack in the state 
-space we can recover the precise call/return match. In this section, we start
-our first step of transformation: \textit{fusing} an AAM with unbounded-stack.
-
-In fact, as later we will eliminate the stack and transform it to continuations
-that are represented by higher-order functions, it does not matter what kind of 
-AAM we start from. 
-Because transforming to higher-order functions forces us to sequentialize 
-the order of abstract evaluation.
-We choose to start from AAM with unbounded-stack just because it has a 
-straightforward correspondence of pushdown semantics to the refunctionalized AAM.
-
-\begin{lstlisting}
-def drive_step(todo: List[State], seen: Set[State]): Set[State] = todo match {
-  case Nil => seen
-  case hd::tl =>
-    if (seen.contains(hd)) return drive_step(tl, seen)
-
-    val new_time = tick(hd)
-    val new_states = hd match {
-      case State(Let(x, App(f, ae), e), env, bstore, konts, time) 
-        if isAtomic(ae) =>
-        for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
-          val frame = Frame(x, e, env)
-          val baddr = allocBind(v, new_time)
-          val new_env = env_c + (v -> baddr)
-          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-          State(body, new_env, new_store, frame::konts, new_time)
-        }
-      case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
-        konts match {
-          case Nil => List()
-          case Frame(x, e, f_env)::konts =>
-            val baddr = allocBind(x, new_time)
-            val new_env = f_env + (x -> baddr)
-            val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-            List(State(e, new_env, new_store, konts, new_time))
-        }
-    }
-    drive_step(new_states ++ tl, seen + hd)
-}
-
-def analyze(e: Expr): Set[State] = drive_step(List(inject(e)), Set())
+Fusing is a tramsformation that combines function @step@ and @drive@ to one.
+The fused function @drive_step@ basically formed by merging function
+@step@ into function @drive@.
+@drive_step@ takes a @NDState@ and a set of explored states as argument,
+and returns a set of reachable states once terminates.
+
+\begin{lstlisting}
+def drive_step(nds: NDState, seen: Set[State]): Set[State] = {
+  nds match {
+    case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
+    case nds =>
+      val newTime = tick(nds)
+      val new_ndstate = nds match {
+        case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndkonts) =>
+          ......
+        case NDState(ae, env, bstore, konts, time, ndkonts) if isAtomic(ae) =>
+          ......
+      }
+      val s = nds.toState
+      if (seen.contains(s)) drive_step(new_ndstate, seen)
+      else drive_step(new_ndstate, seen + s)
+  }
+}
+\end{lstlisting}
+
+Now we have a single function to perform both abstract evaluation and intermediate
+states collecting. Given @inject(e)@ as the inital state and empty set as
+initial reached set, we can easily define the entrance function @analyze@.
+
+\begin{lstlisting}
+def analyze(e: Expr): Set[State] = drive_step(inject(e), Set())
 \end{lstlisting}
@@
 
-for plain direct-style, the continue function handles more cases.
-\todo{a defunctionalized object for non-determinism}
-
-\begin{lstlisting}
-def aval(state: State, todo: List[State], seen: Set[State]): Set[State] = { 
-  if (seen.contains(state)) return collect(todo, seen)
-
-  val newTime = tick(state)
-  state match {
-    case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
-      val newStates = for (Clos(Lam(v, body), c_env) <- atomicEval(f, env, bstore).toList) yield {
-        val frame = Frame(x, e, env)
-        val baddr = allocBind(v, newTime)
-        val newEnv = c_env + (v -> baddr)
-        val newStore = bstore.update(baddr, atomicEval(ae, env, bstore))
-        State(body, newEnv, newStore, frame::konts, newTime)
-      }   
-      collect(newStates ++ todo, seen + state)
-    case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
-      continue(konts, atomicEval(ae, env, bstore), bstore, newTime, todo, seen + state)
-  }   
-}
-
-def collect(todo: List[State], seen: Set[State]): Set[State] = { 
-  todo match {
-    case Nil => seen
-    case hd::tl => aval(hd, tl, seen)
-  } 
-}
-
-def continue(konts: List[Frame], ds: Set[Storable], bstore: BStore, time: Time,
-              todo: List[State], seen: Set[State]): Set[State] = konts match {
-  case Nil => collect(todo, seen)
-  case Frame(x, e, f_env)::konts =>
-    val baddr = allocBind(x, time)
-    val newEnv = f_env + (x -> baddr)
-    val newStore = bstore.update(baddr, ds) 
-    val newState = State(e, newEnv, newStore, konts, time)
-    collect(newState::todo, seen)
-}
-
-def analyze(e: Expr): Set[State] = aval(inject(e), List(), Set()) 
+In last section, the fused AAM combines all the functions into one and
+looks like a ``big-step'' interpreter, although it still has the machine
+state representation inside. In the disentangling transformation, we
+
+\begin{lstlisting}
 \end{lstlisting}
@@
 
+Defunctionalization and refunctionalization build connections between abstract
+machines and interpreters.
+
+For deterministic languages, a reduction semantics is a structural operational
+semantics in continuation style, where the reduction context is a
+defunctionalized continuation.
+Unfortunately, the abstract abstract machines is non-deterministic, where
+the non-determinism is handled by a work list.
+We can turn this explicit non-determinism into the reduction context of
+AAM.
+
 \textbf{Pushdown Control-flow Analysis.}
@@
 
-TODO: summarize PDCFA, AAC
+\todo{PDCFA, AAC}
 

### 2018-03-16 paper.tex
update

@@
 
-In last section, we showed that by keeping an unabstracted stack in the state 
+In the last section, we showed that by keeping an unabstracted stack in the state 
 space we can recover the precise call/return match.
@@
 the order of abstract evaluation, and neither need to construct a frame of list
-nor to allocate continuations in store.
+nor to allocate continuations in the store.
 We choose to start from AAM with unbounded stack just because it has a 
@@
 @NDState@ now.
+For the rest of the paper, to differentiate the two continuations, we would call
+the first one as normal continuation, and the second one as non-deterministic
+continuations.
 
@@
 
-In the second case shown in below we will see how this @NDCont@ controls
+In the second case shown in below we will see how @NDCont@ deal with
 the non-determinism.
@@
       case NDCont(Nil, _, _, _, _)::ndk => 
-        NDState(ae, env, bstore, konts, time, ndk) /* transfer to the next fork point */
+        NDState(ae, env, bstore, konts, time, ndk) /* transfer to the most recent fork point */
       case NDCont(cls, argvs, bstore, time, frames)::ndk => 
@@
       val new_env = f_env + (x -> baddr)
-      val new_store = bstore.update(baddr, aeval(ae, env, bstore))
+      val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
       NDState(e, new_env, new_store, konts, newTime, ndk) /* normal return */
@@
 But, since we have added non-determinism into state, an empty list of frames means
-we have the halt of \textit{one computation path}, and we should determine the next state
-by @NDCont@.
+we have the reached the halt of \textit{one computation path}, and we should
+determine the next state indicates by @NDCont@.
 So we have a pattern matching on @ndk@: 1) If the closure set is an empty set,
 then we had tried all the closures of this fork point and should move the the
-next fork point. 2) Otherwise, we pop a closure from the set, then build new
+next fork point. Since the @NDCont@ is represented by a list and we always
+append new element from its front, we are resuming to the most recent fork point by
+popping up the front-most element.
+2) Otherwise, we pop a closure from the set, then build new
 environment and new store for the body expression of the closure;
@@
     case nds =>
+      val s = nds.toState
+      val new_seen = if (seen.contains(s)) seen else seen+s
       val newTime = tick(nds)
@@
       }
+      drive_step(new_ndstate, new_seen)
+  }
+}
+\end{lstlisting}
+
+Now we have a single function to perform both abstract evaluation and intermediate
+states collecting. Given @inject(e)@ as the inital state and an empty set as
+initial reached set, we can easily define the entrance function @analyze@ as follow.
+
+\begin{lstlisting}
+def analyze(e: Expr): Set[State] = drive_step(inject(e), Set())
+\end{lstlisting}
+
+\section{Disentangling} \label{disen}
+
+In the last section, the fused AAM combines all the functions into one and
+looks like a ``big-step'' interpreter, although it still has the machine
+state representation inside.
+In the disentangling transformation, we
+identify the first-order data type represented evaluation contexts,
+and lift the code blocks that handles these evaluation contexts to be individual functions.
+
+Since there are two layers of continuations, as a result,
+we obtain three mutually recursive functions, @drive_step@,
+@continue@ and @ndcontinue@. The function @drive_step@ plays the same
+role as before; @continue@ handles the evaluation context (@Frame@) of the analyzed
+language, which the function is called from @drive_step@ when encountering an atomic
+expression; @ndcontinue@ dispatches non-deterministic continuations @NDCont@,
+and is invoked from @continue@ when the normal stack is empty.
+
+\begin{lstlisting}
+def drive_step(nds: NDState, seen: Set[State]): Set[State] = {
+  nds match {
+    case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
+    case nds =>
       val s = nds.toState
-      if (seen.contains(s)) drive_step(new_ndstate, seen)
-      else drive_step(new_ndstate, seen + s)
-  }
-}
-\end{lstlisting}
-
-Now we have a single function to perform both abstract evaluation and intermediate
-states collecting. Given @inject(e)@ as the inital state and empty set as
-initial reached set, we can easily define the entrance function @analyze@.
-
-\begin{lstlisting}
-def analyze(e: Expr): Set[State] = drive_step(inject(e), Set())
-\end{lstlisting}
-
-\section{Disentangling} \label{disen}
-
-In last section, the fused AAM combines all the functions into one and
-looks like a ``big-step'' interpreter, although it still has the machine
-state representation inside. In the disentangling transformation, we
-
-\begin{lstlisting}
-\end{lstlisting}
+      val new_seen = if (seen.contains(s)) seen else seen + s
+      val new_time = tick(nds)
+      nds match {
+        case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
+          ......
+          drive_step(NDState(body, new_env, new_store, new_frames, new_time, new_ndk), new_seen)
+
+        case NDState(ae, env, bstore, konts, time, ndk) if isAtomic(ae) =>
+          continue(nds, new_seen)
+      }
+  }
+}
+\end{lstlisting}
+
+The above code shows the skeleton of function @drive_step@.
+In the first case of pattern matching, the code for constructing new
+environment, store and frames is omitted.
+At the end of this case, we invoke @drive_step@ recursively.
+The code for the second case is replaced by a function call to @continue@.
+Now a vague shape of interpreer in continuation style is emerged.
+
+\begin{lstlisting}
+def continue(nds: NDState, seen: Set[State]): Set[State] = {
+  val NDState(ae, env, bstore, konts, time, ndk) = nds
+  val new_time = tick(nds)
+  konts match {
+    case Nil => ndcontinue(nds, seen)
+    case Frame(x, e, f_env)::konts =>
+      val baddr = allocBind(x, new_time)
+      val new_env = f_env + (x -> baddr)
+      val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+      drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), seen)
+  }
+}
+\end{lstlisting}
+
+The function @continue@ is largely what we have for the second case in
+@drive_step@, except a function call to @ndcontinue@ is placed when
+the the normal continuation is empty.
+Since we are analyzing a language in ANF, the first-order data representation of
+continuation is simplified to @List@.
+An empty list @Nil@ represents the halt of a computation path.
+A cons @::@ means the top frame of the list holds the binding variable, and
+evaluation context.
+An analogue to this in standard CEK machines would be an abstract class (e.g., @cont@) with three
+variants @halt@, @arg(value, cont)@ and @arg(exp, env, cont)@.
+
+\begin{lstlisting}
+def ndcontinue(nds: NDState, seen: Set[State]): Set[State] = {
+  val NDState(ae, env, bstore, konts, time, ndk) = nds
+  ndk match {
+    case NDCont(Nil, _, _, _, _)::ndk =>
+      drive_step(NDState(ae, env, bstore, konts, time, ndk), seen)
+    case NDCont(cls, argvs, bstore, time, frames)::ndk =>
+      val Clos(Lam(v, body), c_env) = cls.head
+      val baddr = allocBind(v, time)
+      val new_env = c_env + (v -> baddr)
+      val new_store = bstore.update(baddr, argvs)
+      drive_step(NDState(body, new_env, new_store, frames, time,
+                         NDCont(cls.tail, argvs, bstore, time, frames)::ndk),
+                 seen)
+  }
+}
+\end{lstlisting}
+
+The shape for function @ndcontinue@ is similar to function @continue@, after
+all, they both use @List@ as a representation of continuation.
 

### 2018-03-16 paper.tex
update

@@
 determine the next state indicates by @NDCont@.
-So we have a pattern matching on @ndk@: 1) If the closure set is an empty set,
+So we have a pattern matching on @ndk@:
+\begin{itemize}
+  \item If the closure set is an empty set,
 then we had tried all the closures of this fork point and should move the the
@@
 popping up the front-most element.
-2) Otherwise, we pop a closure from the set, then build new
+\item Otherwise, we pop a closure from the set, then build new
 environment and new store for the body expression of the closure;
@@
 The non-determinism continuation @new_ndk@ is also updated by removal of that closure.
+\end{itemize}
 
@@
 The shape for function @ndcontinue@ is similar to function @continue@, after
-all, they both use @List@ as a representation of continuation.
+all, they both use @List@ as a representation of continuations.
+When the closure set is empty, we recursively call @drive_step@ with the rest
+of @ndk@, since in this case the normal stack @konts@ is still empty, so
+the state will be dispatched to @ndcontinue@ again.
+Otherwise, we invoke @drive_step@ with the body expression of the next closure object.
 
@@
 
-As we pointed earlier, refunctionalizing evaluation contexts to higher-order
-continuation functions forces us to somehow sequentialize the abstract evaluation.
-We identify that in order to sequentialize abstract evaluation, there are three
-fundamental differences with concrete evaluation: 
-1) The dereferencing store addresses is non-deterministic, therefore the returning
-may also be non-deterministic.
-2) We need sound collecting semantics rather than valuation semantics.
-3) The abstract evaluation has to be terminating.
-
-allocating continuations of defined language on the heap of defining language.
+Refunctionalization transforms first-order data type represented
+evaluation contexts to higher-order functions.
+Besides sequentializing the order of abstract evaluation by functions,
+this transformation can be also regarded as allocating continuation functions of
+the defined language in the heap of the meta language.
+
+In this section, several other notable changes have been made:
+\begin{itemize}
+\item We do not have the concept of @state@ (or @NDState@) anymore.
+  The components of the states are lifted as arguments of the evaluation
+  function.
+\item Since our final target is abstract definitional interpreters while still
+non-deterministic, start from this section, our abstract semantic artifact
+returns a set of final values insteaf of collected states.
+\item As we are getting close to the
+abstract definitional interperters, the name of @drive_step@ is updated to
+@aeval@ (for \emph{abstract eval}).
+\end{itemize}
+
+For clearity, in this section, we first show how to refunctionalize the normal continuation and
+non-deterministic continuation without collecting semantics.
+Then adding an appropriate caching algorithm to make sure the analysis will
+terminate at some fixed-point is just changing the program to cache-passing style.
+Finally, we examine how computable pushdown control-flow analysis is established
+through refunctionalization and caching.
 
@@
 
-Transform to Extended Continuation-Passing Style (ECPS), continuation and meta-continuation.
-simply trun evaluation contexts into higher-order continuations.
-We establish pushdown control-flow, but this is not computable.
-
+To properly represented final values, we introduce the case class @VS@ which
+contains a set of storable values, a time stamp, and a store.
+\begin{lstlisting}
+case class VS(vals: Set[Storable], time: Time, store: BStore)
+\end{lstlisting}
+An instance of @VS@ represents the computational result of one path in
+the non-deterministic evaluation.
+The reason that we include a store is there might be a case two pathes have the
+same values but different accumulated side effects such as memory allocations.
+Since the whole computation is non-deterministic, the type of final result
+values is a collection @Set[VS]@. For convenience, we define it as answer type @Ans@:
+\begin{lstlisting}
+type Ans = Set[VS]
+\end{lstlisting}
+
+As we mentioned previously, the components of states are lifted as arguments to
+@aeval@:
+\begin{lstlisting}
+def aeval(e: Expr, env: Env, store: BStore, time: Time, continue: Cont): Ans = ......
+type Cont = Ans => Ans
+\end{lstlisting}
+
+An additional argument @continue@ which has type @Cont@ is also introduced.
+Recall that in the disentangled AAM, the second case in function @drive_step@
+makes a call to @continue@. Here, the reemerging @continue@ is a function
+with type @Ans => Ans@. The function @aeval@ does not return, instead, it calls
+@continue@ with the result values.
+
+\textbf{Non-Determinism Abstraction.}
+To handle the inevitable non-determinism, we defined a generic function @nd@ as follows:
 \begin{lstlisting}
@@
 }
-
-type Ans = Set[VS]
-type Cont = Ans => Ans
-
-def aval(e: Expr, env: Env, store: BStore, time: Time, cont: Cont): Ans = {
+\end{lstlisting}
+
+Given a set of element of type @T@, an initial accumulated value of type @S@, a function @f@
+that works on element type @T@ and accumulated type @S@, and a function @g@ that
+works on and returns value of type @S@, @nd@ applies @f@ to each element in the
+set iteratively, accumulates the value and finally applies @g@ on it. 
+Seeing this one may bring @fold@ to mind, indeed, it is an implementation of
+@fold@ in continuation style. But here we have two continuations, one for the
+next element in the set, and one for the final result.
+Sometime it is also called extended continuation-passing style (ECPS) that has a continuation and a
+meta-continuation \cite{Danvy:1990:AC:91556.91622}.
+\todo{correspondence to disentangled version, refunc the NDCont}
+
+\textbf{First Version of Evaluation Function.}
+Since our first version of @aeval@ just implements the non-deterministic
+evaluation, and given the defined @nd@, the evaluation function can be easily
+defined.
+The outer call to function @nd@ evaluates over a set of possible closures of
+@f@. For each closures, we go into its body with new environment and store.
+Remember that function @aeval@ returns a set of @VS@ to its continuation,
+so the @bodyvss@ is a set of @VS@ that represents values from possible
+computation pathes when evaluating @body@.
+
+Then for the inner call of function @nd@ evaluates over the set of body values.
+For each @VS@, we use the time and store
+The inner application of @aeval@ returns values of one computation path to its
+continuation, where the continuation will accumulate the result and transfer
+evaluation to the next body value.
+
+\begin{lstlisting}
+def aeval(e: Expr, env: Env, store: BStore, time: Time, continue: Cont): Ans = {
   val new_time = (e::time).take(k)
@@
       val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
-      nd[Clos, Ans](closures, Set[VS](), { case (clos, acc, closnd) =>
+      nd[Clos, Ans](closures, Set[VS](), { case (clos, accouter, closnd) =>
         val Clos(Lam(v, body), c_env) = clos
@@
         val new_store = store.update(baddr, atomicEval(ae, env, store)) 
-        aval(body, new_env, new_store, new_time, (bodyvss: Set[VS]) => {
-          nd[VS, Ans](bodyvss, Set[VS](), { case (vs, acc_vss, bdnd) => 
+        aeval(body, new_env, new_store, new_time, (bodyvss: Set[VS]) => {
+          nd[VS, Ans](bodyvss, Set[VS](), { case (vs, accinner, bdnd) => 
             val VS(vals, time, store) = vs
@@
             val new_store = store.update(baddr, vals)
-            aval(e, new_env, new_store, time, (evss: Ans) => bdnd(acc_vss ++ evss))
+            aeval(e, new_env, new_store, time, 
+                  /* accumulate the values of one path, and call next body value*/
+                  (evss: Ans) => bdnd(accinner ++ evss)) 
           },
-          (evss: Ans) => closnd(evss ++ acc))
+          /* accumulate the values of pathes to accouter, and call next closure */
+          (evss: Ans) => closnd(evss ++ accouter)) 
         })
       }, 
-      cont)
+      continue)
     case ae if isAtomic(ae) => 
       val ds = atomicEval(ae, env, store)
-      cont(Set(VS(ds, new_time, store))) 
-  } 
-}
-
-def mtTime = List()
-def mtEnv = Map[String, BAddr]()
-def mtStore = Store[BAddr, Storable](Map())
-
+      continue(Set(VS(ds, new_time, store))) 
+  } 
+}
+\end{lstlisting}
+
+At the end, the initial invocation to @aeval@ is passed with an indentify
+function as halting continuation.
+\begin{lstlisting}
 def analyze(e: Expr) =
-  aval(e, mtEnv, mtStore, mtTime, (ans => ans))
+  aeval(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List(), (ans => ans))
 \end{lstlisting}
@@
 \begin{lstlisting}
-def aval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, cont: Cont): Ans = {
+def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, cont: Cont): Ans = {
   val config = Config(e, env, store, time)
@@
         val new_store = store.update(baddr, atomicEval(ae, env, store))
-        aval(body, new_env, new_store, new_time, cache, { case Ans(bodyvss, bodycache) =>
+        aeval(body, new_env, new_store, new_time, cache, { case Ans(bodyvss, bodycache) =>
           nd[VS, Ans](bodyvss, Ans(Set[VS](), bodycache), { case (vs, Ans(acc_vss, cache), bdnd) =>
@@
             val new_store = store.update(baddr, vals)
-            aval(e, new_env, new_store, time, cache, { case Ans(evss, ecache) => bdnd(Ans(acc_vss ++ evss, ecache)) })
+            aeval(e, new_env, new_store, time, cache, { case Ans(evss, ecache) => bdnd(Ans(acc_vss ++ evss, ecache)) })
           },
@@
   def iter(cache: Cache): Ans = {
-    val Ans(vss, new_cache) = aval(e, env, store, mtTime, cache, { 
+    val Ans(vss, new_cache) = aeval(e, env, store, mtTime, cache, { 
       val initConfig = Config(e, env, store, mtTime)
@@
 \begin{lstlisting}
-def aval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans = {
+def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans = {
   val config = Config(e, env, store, time)
@@
         val new_store = store.update(baddr, aeval(ae, env, store))
-        val Ans(bdv, bdcache) = aval(body, new_env, new_store, new_time, new_cache)
+        val Ans(bdv, bdcache) = aeval(body, new_env, new_store, new_time, new_cache)
         new_cache = bdcache
@@
           val new_store = store.update(baddr, vals)
-          val Ans(ev, ecache) = aval(e, new_env, new_store, time, new_cache)
+          val Ans(ev, ecache) = aeval(e, new_env, new_store, time, new_cache)
           new_cache = ecache

### 2018-03-16 paper.tex
update

@@
 As we mentioned previously, the components of states are lifted as arguments to
-@aeval@:
-\begin{lstlisting}
-def aeval(e: Expr, env: Env, store: BStore, time: Time, continue: Cont): Ans = ......
+@aeval@. Therefore, the pattern matching on state also becomes just matching on
+expression @e@.
+\begin{lstlisting}
 type Cont = Ans => Ans
+def aeval(e: Expr, env: Env, store: BStore, time: Time, continue: Cont): Ans = {
+  val new_time = (e::time).take(k)
+  e match {
+    case Let(x, App(f, ae), e) =>
+      ......
+    case ae if isAtomic(ae) => 
+      ......
+  }
+}
 \end{lstlisting}
@@
 Since our first version of @aeval@ just implements the non-deterministic
-evaluation, and given the defined @nd@, the evaluation function can be easily
-defined.
-The outer call to function @nd@ evaluates over a set of possible closures of
-@f@. For each closures, we go into its body with new environment and store.
-Remember that function @aeval@ returns a set of @VS@ to its continuation,
-so the @bodyvss@ is a set of @VS@ that represents values from possible
-computation pathes when evaluating @body@.
-
-Then for the inner call of function @nd@ evaluates over the set of body values.
-For each @VS@, we use the time and store
-The inner application of @aeval@ returns values of one computation path to its
-continuation, where the continuation will accumulate the result and transfer
-evaluation to the next body value.
-
-\begin{lstlisting}
-def aeval(e: Expr, env: Env, store: BStore, time: Time, continue: Cont): Ans = {
-  val new_time = (e::time).take(k)
-  e match {
+evaluation, and given the defined @nd@, the evaluation function @aeval@ can be easily
+defined. For the case that @e@ is a @let@ expression,
+the overall idea is to perform a two step depth-first evluation. \todo{refine}
+
+\begin{lstlisting}
 case Let(x, App(f, ae), e) =>
@@
   continue)
+\end{lstlisting}
+
+The outer function call @nd@ evaluates over the set of possible closures of
+@f@. For each closures, we go into its body with new environment and new store.
+Remember that the function @aeval@ returns a set of @VS@ to its continuation,
+so the continuation argument @bodyvss@ is a set of @VS@ that represents values
+from multiple computation pathes when evaluating this single @body@ expression.
+
+Then for the inner call of function @nd@ evaluates over the set of body values @bodyvss@.
+For each @VS@, the time stamp and store in @VS@ will be instantiated to
+@aeval@.
+The inner application of @aeval@ returns set of values of one computation path to its
+continuation, where the continuation will accumulate the result and transfer
+evaluation to the next body value.
+
+\begin{lstlisting}
 case ae if isAtomic(ae) => 
@@
   continue(Set(VS(ds, new_time, store))) 
-  } 
-}
-\end{lstlisting}
+\end{lstlisting}
+
+If the expression is atomic, then we just need to construct a new instance of
+@VS@, return it to the @continue@ function.
 
@@
 
+In the last section, we just established a computable pushdown control-flow
+analysis through refunctionalization and caching. In this section,
+we examine how pushdown control-flow is obtained and what we have done to
+achieve it.
+
+\textbf{The Problem of Return-flows.}
 Pushdown control-flow is a property in analysis that precisely models
@@
 
-To see how traditional CFA suffers from spurious return-flows, 
+To see how traditional control-flow analysis suffers from spurious return-flows, 
 we can consider the following example:
@@
 
-In traditional $k$-CFA algorithm and the AAM we showed in section \ref{aam}, 
-the return-flow of invocation @(id 2)@
-is going to both call-sites @(id 1)@ and @(id 2)@. Therefore, the
-returned value @2@ for variable @y@ is also propagated to 
+In $k$-CFA algorithm or the abstract abstract machine we showed in section
+\ref{aam},
+The call sites @(id 2)@ and @(id 1)@ share the same return flow,
+so the of invocation @(id 2)@ returns to both call-sites @[x (id 1)]@ and @[y (id 2)]@.
+Therefore, the returned value @2@ for variable @y@ is also propagated to 
 variable @x@ and then imprecise analysis result is arose.
 This return-flow merging is inevitable even we increase the context-sensitivity.
-If we use the monovariant analysis (i.e. 0CFA), the analyzing result would be @x@ and 
-@y@ point to set @{1, 2}@, because the algorithm not distinguishes
+If we use the monovariant analysis (i.e., 0CFA), the analyzing result would be @x@ and 
+@y@ point to value set @{1, 2}@, because the algorithm not distinguishes
 @z@ comes from different call-sites.
@@
 would not be polluted by @1@.
-But two call-sites still share the same continuation and
+However, variable @x@ still points to value set @{1, 2}@.
+Because two call-sites still share the same continuation and
 it is still helplessness for separating the return-flows.
-TODO: do need more details to explain this example? (as in P4F paper)
+
+\textbf{Call/Return Matching through Refunctionalization and Caching.}
+\todo{how refunc and caching help}
 

### 2018-03-16 paper.tex
update

@@
 %% Title information
-\title{Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)}         %% [Short Title] is optional;
+%\title{Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)}         %% [Short Title] is optional;
+\title{Refunctionalization of Abstracting Abstract Machines (Functional Pearl)}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
   abstract definitional interpreters (ADI) written in monadic style.
-  Danvy shows the correspondence between concrete abstract machines and interpreters.
   Yet, the relations between small-step abstracting abstract machines and 
@@
   interpreters.
-  The transformations include fusing, disentangling, refunctionalizing 
-  and monadification (or un-CPS to direct-style with delimited continuations). 
-  All of them properly handle the 
-  collecting semantics and the non-determinism of abstract interpretation. 
-  After each transformation, we also obtain an intermediate form of 
-  abstract interpreter.
-
-  Following the idea that evaluation contexts are defunctionalized 
-  continuations, we also found that refunctionalization sequentializes the 
-  non-deterministic abstract evaluation.
+  The transformations include linearization, fusing, disentangling, refunctionalizing 
+  and monadification (or un-CPS to direct-style). Linearization expresses
+  non-deterministic choices into first-order data types, and then refunctionalization
+  sequentializes the evaluation order by higher-order functions.
+  All of transformations properly handle the collecting semantics and the
+  non-determinism of abstract interpretation. 
+  %After each transformation, we also obtain an intermediate form of abstract interpreter.
+
+  Following the idea that in deterministic languages,
+  evaluation contexts in reduction semantics are defunctionalized continuations,
+  we further show in non-deterministic languages, evaluation contexts are refunctionalized to
+  extended continuations style.
   Remarkably, we reveal how precise call/return matching in control-flow 
-  analysis can be obtained by properly refunctionalizing a small-step abstract interpreter.
-  As well as we explain how it been lost by defunctionalizing the higher-order
-  continuations into an first-order data type.
-
-  TODO: Maybe something Shiver's formulation of CFA in denotational style.
-
+  analysis is obtained by refunctionalizing a small-step abstract
+  abstract machine with properly caching.
+
+  %As well as we explain how it been lost by defunctionalizing the higher-order continuations into an first-order data type. 
 \end{abstract}

### 2018-03-16 paper.tex
update

@@
 In section \ref{background}, we first describe the target language we will
-analyze and a concrete CESK machine as operational semantics for it.
+analyze and a concrete CESK machine as small-step operational semantics for it.
 Then we review abstracting abstract machines (AAM) as an abstract interpretation
-counterpart of CESK machines.
+counterpart of CESK machines. In section \ref{linear}, we convert the
+non-deterministic choices of closures to another layer of continuations,
+therefore linearize the abstract execution.
+Section \ref{fusing} fuses the driver function and step function to one single
+function.
+In section \ref{disen}, we further identify the first-order data type representations of
+continuations, and disentangle the code blocks for dispatching data types to separate
+functions.
+Section \ref{refunc} represents the refunctionalized abstract abstract machines,
+In this part, the continuations represented by data type trun to higher-order
+functions.
+In Section \ref{directstyle}, we transform the refunctionalized AAM to two
+different direct style, one is implemented with monads, the other is using
+delimited controls.
+
+\todo{Add section numbers to the figure}
 
@@
 %probably Store Widening \& Abstract Garbage Collection
-\section{To Direct-Style}
+\section{To Direct-Style} \label{directstyle}
 

### 2018-03-16 paper.tex
Fixed spacing, typos

@@
 
-boardly speaking, AAM and ADI they are both abstract interpreters.
+broadly speaking, AAM and ADI they are both abstract interpreters.
 
@@
   because the refunctionalization forces us to sequentialize the
-  abstract exectuion.
+  abstract execution.
   We also analyze its correctness, soundness and termination.
@@
 Section \ref{refunc} represents the refunctionalized abstract abstract machines,
-In this part, the continuations represented by data type trun to higher-order
+In this part, the continuations represented by data type turn to higher-order
 functions.
@@
 \subsubsection{Allocating Addresses}
-Up to now, we did not describe about allocating adresses in stores and the time stamp @Time@.
+Up to now, we did not describe about allocating addresses in stores and the time stamp @Time@.
 But in abstract interpretation, they are key ingredients to achieve
@@
 Since dereferencing an address becomes non-deterministic, our @atomicEval@
-function also truns to be non-deterministic. Given an atomic expression @e@,
+function also turns to be nondeterministic. Given an atomic expression @e@,
 @atomicEval@ returns a set of storable values, i.e., closures, to the caller.
@@
 
-A histroy of the last $k$ call sites is used as a finite program contour in
+A history of the last $k$ call sites is used as a finite program contour in
 $k$-call-sensitive analysis.
@@
 space collapses to the space of variable names.
-Note that regarding the ambiguitiy in $k$-CFA\cite{Gilray:2016:ACP:2951913.2951936},
+Note that regarding the ambiguity in $k$-CFA\cite{Gilray:2016:ACP:2951913.2951936},
 the code here actually implements call+return sensitivity.
@@
 In Danvy's paper \textit{Defunctionalized Interpreters for Programming
-  Languages}, he metioned for deterministic languages, a reduction semantics
+  Languages}, he mentioned for deterministic languages, a reduction semantics
 is a structural operational semantics in continuation style, where the reduction
@@
 
-In order to address this problem, our first step of tramsformations
+In order to address this problem, our first step of transformations
 is to linearize all the non-deterministic choices.
@@
 
-Fusing is a tramsformation that combines function @step@ and @drive@ to one.
+Fusing is a transformation that combines function @step@ and @drive@ to one.
 The fused function @drive_step@ basically formed by merging function
@@
 Now we have a single function to perform both abstract evaluation and intermediate
-states collecting. Given @inject(e)@ as the inital state and an empty set as
+states collecting. Given @inject(e)@ as the initial state and an empty set as
 initial reached set, we can easily define the entrance function @analyze@ as follow.
@@
 The code for the second case is replaced by a function call to @continue@.
-Now a vague shape of interpreer in continuation style is emerged.
+Now a vague shape of interpreter in continuation style is emerged.
 
@@
 non-deterministic, start from this section, our abstract semantic artifact
-returns a set of final values insteaf of collected states.
+returns a set of final values instead of collected states.
 \item As we are getting close to the
-abstract definitional interperters, the name of @drive_step@ is updated to
+abstract definitional interpreters, the name of @drive_step@ is updated to
 @aeval@ (for \emph{abstract eval}).
@@
 
-For clearity, in this section, we first show how to refunctionalize the normal continuation and
+For clarity, in this section, we first show how to refunctionalize the normal continuation and
 non-deterministic continuation without collecting semantics.
@@
 defined. For the case that @e@ is a @let@ expression,
-the overall idea is to perform a two step depth-first evluation. \todo{refine}
+the overall idea is to perform a two step depth-first evaluation. \todo{refine}
 
@@
       },
-      /* accumulate the values of pathes to accouter, and call next closure */
+      /* accumulate the values of paths to accouter, and call next closure */
       (evss: Ans) => closnd(evss ++ accouter))
@@
 so the continuation argument @bodyvss@ is a set of @VS@ that represents values
-from multiple computation pathes when evaluating this single @body@ expression.
+from multiple computation pates when evaluating this single @body@ expression.
 
@@
 The target of the series of transformations in this paper is abstract definitional
-interpreters, and we additionaly demonstrate abstract interpreters written in both
+interpreters, and we additionally demonstrate abstract interpreters written in both
 monadic- and direct-style.

### 2018-03-16 paper.tex
Abstract grammar (teehee)

@@
   small-step abstract machines. Recent progress applies the same idea
-  on definitional interpreters and obtains big-step
+  on definitional interpreters, and obtains big-step
   abstract definitional interpreters (ADI) written in monadic style.
@@
   interpreters.
-  The transformations include linearization, fusing, disentangling, refunctionalizing
+  The transformations include linearization, fusing, disentangling, refunctionalizing,
   and monadification (or un-CPS to direct-style). Linearization expresses
-  non-deterministic choices into first-order data types, and then refunctionalization
+  non-deterministic choices into first-order data types, after which refunctionalization
   sequentializes the evaluation order by higher-order functions.
-  All of transformations properly handle the collecting semantics and the
-  non-determinism of abstract interpretation.
+  All transformations properly handle the collecting semantics and the
+  nondeterminism of abstract interpretation.
   %After each transformation, we also obtain an intermediate form of abstract interpreter.

### 2018-03-16 paper.tex
More abstract grammar

@@
 
-  Following the idea that in deterministic languages,
+  Following the idea that in deterministic languages
   evaluation contexts in reduction semantics are defunctionalized continuations,
-  we further show in non-deterministic languages, evaluation contexts are refunctionalized to
-  extended continuations style.
+  we further show that in nondeterministic languages, evaluation contexts are
+  refunctionalized to extended continuations style.
   Remarkably, we reveal how precise call/return matching in control-flow
   analysis is obtained by refunctionalizing a small-step abstract
-  abstract machine with properly caching.
+  abstract machine with proper caching.
 

### 2018-03-16 paper.tex
Contributions

@@
 
-broadly speaking, AAM and ADI they are both abstract interpreters.
+Abstracting abstract machines (AAMs) and abstract definitional interpreters (ADIs) are, broadly speaking, both examples of abstract interpreters.
+\note{JD: need more bullet points here and I can fill this in}
 
@@
 
-the essence of abstracting definitional interpreter.
+% the essence of abstracting definitional interpreter.
+
+We summarize our contributions as follows:
 
 \begin{itemize}
-\item The main contribution of this paper is filling the gap
+\item The main contribution of this paper is the filling of the gap
   between small-step AAM and big-step ADI by developing a series of
   systematic transformations.
-  The correspondence between abstract
+  We show that the correspondence between abstract
   machines and interpreters exists not only in concrete semantics artifact,
@@
   Galois connection?
+  \note{JD: Section number?}
 
@@
   We present an abstract interpreter obtained from refunctionalizing
-  an small-step abstracting abstract machine.
-  This abstract interpreter naturally enjoys pushdown control-flow property
-  because the refunctionalization forces us to sequentialize the
+  a small-step abstracting abstract machine.
+  This abstract interpreter naturally exhibits the pushdown control-flow property,
+  as the refunctionalization forces us to sequentialize the
   abstract execution.
-  We also analyze its correctness, soundness and termination.
+  We also analyze its correctness, soundness, and termination.
 
 \item Further, we show that defunctionalization and refunctionalization
-  of abstract interpreters play important roles for call stack of analyzed language.
-  Given this insight, we reveal that why small-step AAM has no pushdown
-  control-flow property (if without extra efforts), as well as explain
-  that why ADI naturally extends pushdown control-flow property from its
+  of abstract interpreters play important roles for the call stack of the analyzed language.
+  Given this insight, we reveal precisely \emph{why} small-step AAM has no pushdown
+  control-flow property (without extra efforts), as well as explain
+  why ADI naturally extends pushdown control-flow property from its
   defining language.

### 2018-03-16 paper.tex
Outline

@@
 
-The paper is organized as follows.
-
-In section \ref{background}, we first describe the target language we will
-analyze and a concrete CESK machine as small-step operational semantics for it.
-Then we review abstracting abstract machines (AAM) as an abstract interpretation
-counterpart of CESK machines. In section \ref{linear}, we convert the
-non-deterministic choices of closures to another layer of continuations,
-therefore linearize the abstract execution.
-Section \ref{fusing} fuses the driver function and step function to one single
+The remainder of this paper is organized as follows.
+We begin by describing the target language we will analyze in Section~\ref{background},
+including a concrete CESK machine as small-step operational semantics for it.
+We continue with a review of abstracting abstract machines (AAM) as an abstract interpretation
+counterpart of CESK machines. Following this, Section~\ref{linear} shows a conversion from the
+nondeterministic choices of closures to another layer of continuations,
+therefore linearizing the abstract execution.
+Section~\ref{fusing} then discusses fusing the driver and step functions to a single
 function.
-In section \ref{disen}, we further identify the first-order data type representations of
+We further identify in Section~\ref{disen} the first-order data type representations of
 continuations, and disentangle the code blocks for dispatching data types to separate
 functions.
-Section \ref{refunc} represents the refunctionalized abstract abstract machines,
-In this part, the continuations represented by data type turn to higher-order
+Section~\ref{refunc} then represents the refunctionalized abstract abstract machines.
+In this section, the continuations represented by data types become higher-order
 functions.
-In Section \ref{directstyle}, we transform the refunctionalized AAM to two
+In Section~\ref{directstyle}, we transform the refunctionalized AAM to two
 different direct style, one is implemented with monads, the other is using
 delimited controls.
+
+Finally, we discuss related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.
 
@@
 
-\section{Related Work}
+\section{Related Work}\label{sec:related}
 
@@
 
-\section{Conclusion}
+\section{Conclusion}\label{sec:conclusion}
 

### 2018-03-16 paper.tex
formatting

@@
 
+\note{TR: can this be merged with contributions? (they currently lack section refs)}
 The remainder of this paper is organized as follows.
@@
 
-\todo{Add section numbers to the figure}
+\todo{Add section numbers to the figure}\note{TR: put this into a proper figure, add known danvy sequence 
+for concrete evaluators at the bottom (in grey or dashed lines)}
 
@@
 \begin{lstlisting}
-e $\in$ Exp ::= ae
-          | (let ([x (ae ae)]) e)
-ae $\in$ AExp ::= x
-            | lam
-lam $\in$ Lam ::= (lambda (x) e)
-x ... are variable names
+  e $\in$ Exp ∂::= ae | (let ([x (ae ae)]) e)
+ ae $\in$ AExp ::= ∂x | lam
+lam $\in$ Lam ∂::= (lambda (x) e)
+  x $\in$ Var   (variable names)
 \end{lstlisting}
@@
 The abstract syntax can be straightforwardly described in Scala as follows:
-TODO: update the abstract syntax.
+\note{TR: grammar below isn't ANF. Either make it so, or clarify that 
+we assume ANF only as convention and don't need/want to enforce it in
+the term structure}
 
@@
 \section{Conclusion}\label{sec:conclusion}
+\note{todo}
 
@@
 %% Appendix
-\appendix
-\section{Appendix}
-
-Text of appendix \ldots
+% \appendix
+% \section{Appendix}
+
+% Text of appendix \ldots
 

### 2018-03-16 paper.tex
misc

@@
 %% Title information
-%\title{Pushdown Control-Flow Analysis via Refunctionalization (Functional Pearl)}         %% [Short Title] is optional;
+%\title{Pushdown Control Flow Analysis via Refunctionalization (Functional Pearl)}         %% [Short Title] is optional;
 \title{Refunctionalization of Abstracting Abstract Machines (Functional Pearl)}         %% [Short Title] is optional;
@@
   and monadification (or un-CPS to direct-style). Linearization expresses
-  non-deterministic choices into first-order data types, after which refunctionalization
+  non-deterministic choices by first-order data types, after which refunctionalization
   sequentializes the evaluation order by higher-order functions.
@@
 
-\subsection{Contributions}
+\note{rest of intro goes here}
+
+The rest of this paper is structured as follows: we begin by reviewing some background
+necessary for this work in Section~\ref{background}, as well as introducing some of the basic
+code structures used throughout the paper. We then address the main contribution of this
+paper, which is the filling of the gap between small-step AAM and big-step ADI by developing a
+series of systematic transformations. Those transformations are summarized here, with their
+associated section:
+
+\begin{itemize}
+\item We show how \textbf{linearization} (Section~\ref{linear}) can be used to...
+\item In Section~\ref{fusing}, we then present the \textbf{fusing} transformation, and present...
+\item Section~\ref{disen} discusses \textbf{disentangling}, which shows...
+\item \textbf{Refunctionalization} is shown in Section~\ref{refunc}, and brings a new understanding of pushdown control flow analysis.
+In this section, we present an abstract interpreter obtained from refunctionalizing a small-step abstracting abstract machine which naturally exhibits the pushdown control-flow property (as refunctionalization requires abstract execution to be sequentialized).
+\item The last transformation we show is that of transforming a definitional abstract interpreter to a \textbf{direct-style} interpreter (Section~\ref{directstyle}).
+
+\end{itemize}
+
+These transformations are used throughout the paper, with refunctionalization and
+defunctionalization of abstract interpreters playing important roles for the call stack of
+the analyzed language (i.e., Scala)\note{right?}. All of this insight allows the revelation of
+precisely \emph{why} small-step AAM does not have the pushdown control-flow property
+(without extra effort), as well as explains why ADI naturally extends pushdown control-flow
+property from its defining language.
+
+We complete this paper by discussing related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.
+
+% \subsection{Contributions}
 
@@
 
-We summarize our contributions as follows:
-
-\begin{itemize}
-\item The main contribution of this paper is the filling of the gap
-  between small-step AAM and big-step ADI by developing a series of
-  systematic transformations.
-  We show that the correspondence between abstract
-  machines and interpreters exists not only in concrete semantics artifact,
-  but also in abstract semantics artifact.
-  Galois connection?
-  \note{JD: Section number?}
-
-\item A new understanding of pushdown control-flow analysis by
-  refunctionalization.
-  We present an abstract interpreter obtained from refunctionalizing
-  a small-step abstracting abstract machine.
-  This abstract interpreter naturally exhibits the pushdown control-flow property,
-  as the refunctionalization forces us to sequentialize the
-  abstract execution.
-  We also analyze its correctness, soundness, and termination.
-
-\item Further, we show that defunctionalization and refunctionalization
-  of abstract interpreters play important roles for the call stack of the analyzed language.
-  Given this insight, we reveal precisely \emph{why} small-step AAM has no pushdown
-  control-flow property (without extra efforts), as well as explain
-  why ADI naturally extends pushdown control-flow property from its
-  defining language.
-
-\end{itemize}
-
-\subsection{Outline}
-
-\note{TR: can this be merged with contributions? (they currently lack section refs)}
-The remainder of this paper is organized as follows.
-We begin by describing the target language we will analyze in Section~\ref{background},
-including a concrete CESK machine as small-step operational semantics for it.
-We continue with a review of abstracting abstract machines (AAM) as an abstract interpretation
-counterpart of CESK machines. Following this, Section~\ref{linear} shows a conversion from the
-nondeterministic choices of closures to another layer of continuations,
-therefore linearizing the abstract execution.
-Section~\ref{fusing} then discusses fusing the driver and step functions to a single
-function.
-We further identify in Section~\ref{disen} the first-order data type representations of
-continuations, and disentangle the code blocks for dispatching data types to separate
-functions.
-Section~\ref{refunc} then represents the refunctionalized abstract abstract machines.
-In this section, the continuations represented by data types become higher-order
-functions.
-In Section~\ref{directstyle}, we transform the refunctionalized AAM to two
-different direct style, one is implemented with monads, the other is using
-delimited controls.
-
-Finally, we discuss related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.
+% We summarize our contributions as follows:
+
+% \begin{itemize}
+% \item The main contribution of this paper is the filling of the gap
+%   between small-step AAM and big-step ADI by developing a series of
+%   systematic transformations.
+%   We show that the correspondence between abstract
+%   machines and interpreters exists not only in concrete semantics artifacts,
+%   but also in abstract semantics artifacts.
+%   Galois connection?
+%   \note{JD: Section number?}
+
+% \item A new understanding of pushdown control flow analysis by
+%   refunctionalization.
+%   We present an abstract interpreter obtained from refunctionalizing
+%   a small-step abstracting abstract machine.
+%   This abstract interpreter naturally exhibits the pushdown control-flow property,
+%   as the refunctionalization forces us to sequentialize the
+%   abstract execution.
+%   We also analyze its correctness, soundness, and termination.
+
+% \item Further, we show that defunctionalization and refunctionalization
+%   of abstract interpreters play important roles for the call stack of the analyzed language.
+%   Given this insight, we reveal precisely \emph{why} small-step AAM has no pushdown
+%   control-flow property (without extra efforts), as well as explain
+%   why ADI naturally extends pushdown control-flow property from its
+%   defining language.
+
+% \end{itemize}
+
+% \subsection{Outline}
+
+% \note{TR: can this be merged with contributions? (they currently lack section refs)}
+% The remainder of this paper is organized as follows.
+% We begin by describing the target language we will analyze in Section~\ref{background},
+% including a concrete CESK machine as small-step operational semantics for it.
+% We continue with a review of abstracting abstract machines (AAM) as an abstract interpretation
+% counterpart of CESK machines. Following this, Section~\ref{linear} shows a conversion from the
+% nondeterministic choices of closures to another layer of continuations,
+% therefore linearizing the abstract execution.
+% Section~\ref{fusing} then discusses fusing the driver and step functions to a single
+% function.
+% We further identify in Section~\ref{disen} the first-order data type representations of
+% continuations, and disentangle the code blocks for dispatching data types to separate
+% functions.
+% Section~\ref{refunc} then represents the refunctionalized abstract abstract machines.
+% In this section, the continuations represented by data types become higher-order
+% functions.
+% In Section~\ref{directstyle}, we transform the refunctionalized AAM to two
+% different direct styles: one implemented with monads, the other using
+% delimited controls.
+
+% Finally, we discuss related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.
 
@@
 for concrete evaluators at the bottom (in grey or dashed lines)}
+\note{JD: should these be abstracting abstract machines?}
 
@@
 for analyzing functional programs because it exposes control transfer explicitly
-and simplifies the analysis \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
-Here, we choose to use $\lambda$-calculus in administrative normal form (ANF)
-which is a direct-style intermediate representation as target for clarity
+and simplifies analysis~\cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
+Here, we choose to use $\lambda$-calculus in administrative normal form (ANF),
+which is a direct-style intermediate representation, as our target for clarity,
 but without losing simplicity and generality.
-Although we just show the core calculus language, it can be easily extended
-to support recursive bindings (such as @letrec@), conditionals, primitive types and
+Although we only show the core calculus language, it can be easily extended
+to support recursive bindings (such as @letrec@), conditionals, primitive types, and
 operations on primitive types. These cases would be trivial to implement,
-so we omit them in the paper.
+so we elide them in this paper.
 The transformations we will show in the rest of this paper
@@
 To begin, we now present the concrete syntax of a call-by-value $\lambda$-calculus language
-in ANF \cite{flanagan1993essence}.
+in ANF~\cite{flanagan1993essence}.
 
@@
 
-In ANF, an expression is an atomic expression or a @let@ expression.
-The restriction is that all the function applications must be administrated in a @let@ expression,
-and then bound to a variable name under the current environment.
-Both the operator and operand of function applications are an atomic expressions.
-The atomic expressions $ae$ are either a variable, or a literal @lambda@ term, which
-both can be evaluated in a single step.
+In ANF, an expression is either an atomic expression or a @let@ expression.
+A restriction exists which states that all function applications must be administrated
+within a @let@ expression and then bound to a variable name under the current environment.
+Both the operator and operand of function applications are atomic expressions.
+An atomic expression $ae$ is either a variable or a literal @lambda@ term, either of which
+can be evaluated in a single step.
 We also assume that all the variable names in the program are unique.
@@
 The CESK machine models program execution as state transitions in a small-step fashion.
-As its name suggested, a machine state has four components:
+As its name suggests, a machine state has four components:
 1) \textit{Control} is the expression currently being evaluated.
 2) \textit{Environment} is a map that contains the address of a variable in the lexical scope.
-3) \textit{Store} models the heap of a program. It maps from addresses to values.
-The addresses space is numbers starting from 0.
-In our toy language, the only category of value is the closure, i.e., a function paired with an environment.
+3) \textit{Store} models the heap of a program as a map from addresses to values.
+The address space consists of numbers (0-indexed).
+In our toy language, the only category of value is a closure, i.e., a function paired with an environment.
 4) \textit{Continuation} represents the program stack. \todo{explain more on stack}
 
-We show the Scala representations for the components of CESK machine.
+The Scala representations for the components of CESK machine are as follows:
 
@@
 It is worth noting that the continuation class @Kont@ is defined as a list of
-frames.
-The frame can be considered as evaluation context in reduction semantics.
-The frame class @Frame@ stores the information of a single call-site,
-i.e., the information that can be used to resume the interrupted computation.
-A @Frame@ constitutes of a variable name @x@ to be bind later, a control expression
-that resumes to, and its environment. (TODO: mention that frame list vs KArg/KFun)
-
-\subsubsection{Single-step Transition}
-Before going into describing how the machine evaluates expressions, we firstly define several helper functions.
-As we mentioned in section \ref{anfsyntax}, the atomic expressions are either a variable, or
-a literal @lambda@ term. So the atomic expression evaluator @atomicEval@ handles these two
-cases and evaluates atomic expressions to closures in a straightforward way.
-The @alloc@ function generates a fresh address, and always allocates unique integer
-in the domain of store.
-The @isAtomic@ function is used as a predicate of the expression is atomic.
+frames, where a frame can be considered an evaluation context in reduction semantics.
+We represent frames using the @Frame@ class, which stores the information of a single
+call-site, i.e., the information that can be used to resume the interrupted computation.
+A @Frame@ constitutes a variable name @x@ to be bound later, a control expression
+to which the program may resume, and its environment. \note{JD: someone check this and make sure I got it right}
+\todo{mention that frame list vs KArg/KFun}
+
+\subsubsection{Single-Step Transition}
+Before describing how the machine evaluates expressions, we must first define several helper
+functions. As mentioned in Section~\ref{anfsyntax}, atomic expressions are either a variable
+or a literal @lambda@ term. As such, the atomic expression evaluator @atomicEval@ handles
+these two cases and evaluates atomic expressions to closures in a straightforward way.
+The @alloc@ function generates a fresh address, and always allocates a unique integer
+in the domain of @store@.
+The @isAtomic@ function is used as a predicate to determine if the expression is atomic.
 
@@
 
-Now, we can faithfully describe the state transition function @step@,
-that given a machine state, determines its successor state.
+We can now faithfully describe the state transition function @step@,
+which when given a machine state, determines its successor state.
 
@@
 
-In the first case, the control of current state is a @Let@ expression,
-and its right-hand side is an function application.
-By calling evaluator @atomicEval@, we obtain the closure that callee @f@ stands for.
-The successor state's control then transfers to the @body@ expression of closure,
-with an updated environment and an update store. The new environment is extended
-from closure's environment and mapped @v@ to a fresh address @addr@.
+As shown and previously discussed, we examine the only two cases which the state may be.
+
+\textbf{\texttt{Let} expression.} In the first @case@ statement shown in the previous code,
+the control of the current state matches as a @Let@ expression, with its right-hand side a
+function application.
+By calling the @atomicEval@ evaluator, we obtain the closure for which the callee @f@ stands.
+The successor state's control then transfers to the @body@ expression of the closure
+with an updated environment and store. The new environment is extended
+from the closure's environment and mapped from @v@ to a fresh address @addr@.
 The new store is extended with @addr@ mapping to the value of @ae@,
-which is evaluated from @atomicEval@.
-In the meanwhile, a new frame @frame@ is pushed onto the stack @k@.
-This @frame@ contains the variable name @x@ at left-hand side position of @Let@,
+which in turn is evaluated from @atomicEval@.
+Finally, a new frame @frame@ is pushed onto the stack @k@, where
+@frame@ contains the variable name @x@ at the left-hand position of the @Let@,
 the body expression of @Let@, and the lexical environment of the body expression.
 
-The second case for @step@ is the control is transferred to an atomic expression.
-We extract the top frame of continuations firstly.
-The control (that is an atomic expression) of current state is the evaluated term
-that being bind to the variable @x@ from the top frame.
+\textbf{Atomic expression.} If the state is not a @Let@ expression, then it must be
+an atomic expression, as seen in the above code. In this scenario,
+we begin by extracting the top frame of all available continuations.
+The control (i.e., an atomic expression) of the current state is the evaluated term
+that is being bound to the variable @x@ from the top frame.
 The environment and store are updated with @x@ mapping to the closure value of @ae@.
-Then the successor state is transferred to expression @e@ from the top frame,
+Finally, the successor state is transferred to expression @e@ from the top frame,
 which is the body of a @Let@ expression, with the updated environment, store, and
-the rest of stack @ks@.
+the rest of the stack @ks@.
 
@@
 
-To run the program, we first use function @inject@ to construct an initial machine
+To run the program, we first use the @inject@ function (below) to construct an initial machine
 state given an expression @e@. The initial state contains an empty environment,
-store and stack.
+store, and stack.
 \begin{lstlisting}
@@
 
-Then the @drive@ function is used to evaluate
-to a final state by iteratively apply @step@ on current state until we reach a state
-that the control is an atomic expression and the continuation stack is empty.
-And of course, we can extract the value from the final state at last.
+The @drive@ function is then used to evaluate
+to a final state by iteratively applying @step@ on the current state until a state is reached
+ion which the control is an atomic expression and the continuation stack is empty.
+Naturally, we can then extract the value from the final state at last.
 
@@
 
-\subsection{Abstracting Abstract Machine} \label{aam}
+\subsection{Abstracting Abstract Machines} \label{aam}
 Abstracting abstract machines (AAM) is a systematic methodology that derives sound
@@
 abstract machines \cite{van2012systematic, van2010abstracting}.
-An abstracting abstract machine implements computable abstract semantics that
+An abstracting abstract machine implements computable abstract semantics which
 approximates the runtime behaviors of programs.
 Since the state space of concrete execution is possibly infinite,
-to analyze programs, the key insight of AAM approach is to allocate both bindings
+the key insight of AAM approach when analyzing programs is to allocate both bindings
 and continuations on the store, and then bound the addresses space to be finite.
 Since each component of state is finite,
-the abstracted machine-state space is finite, and computable.
+the abstracted machine-state space is also finite, and therefore computable.
 
@@
 CESK machines, and also show how to instantiate
-useful $k$-call-sensitive control-flow analysis.
+useful $k$-call-sensitive control flow analysis.
 
@@
 Similar to CESK machines, the machine state of AAM has a control expression,
-an environment, a store and continuation, and in addition with a time-stamp.
-But there are several notable differences between AAM's store and CESK machine's store.
-In AAM, the store maps addresses to sets of values, so it stores all possible values
-for a particular address. Thus dereferencing addresses becomes non-deterministic.
-Also, the store performs joining rather than overwriting when updating elements
-of the store.
-Third, the continuations are likewise allocated on store instead of formed a linked list,
-and the continuation in a state then turns into a continuation address.
-
-For clarity, we divide the store into a binding store @BStore@ and a continuation store @KStore@.
-The binding store maps binding addresses to sets of closure values, and the continuation store maps
-continuation addresses to sets of continuations.
-We define a generic class @Store[K,V]@ that performs joining when updating elements
-in a store. By parameterizing @Store[K,V]@ with @[BAddr, Storable]@ and
-@[KAddr, Cont]@, we obtain @BStore@ and @KStore@, perspectively.
-
-TODO(rephrase): Note that the store is monotonic, the store is always growing and never becoming
-smaller. This property guarantees we eventually will find a fixed point.
+an environment, a store, and continuation, as well as a timestamp.
+However, there are several notable differences between AAM's store and CESK machine's store.
+In AAM, the store maps addresses to sets of values; it stores all possible values
+for a particular address. As such, dereferencing addresses becomes nondeterministic.
+Also, the store performs \emph{joining}, rather than overwriting, when updating elements.
+Furthermore, the continuations are likewise allocated on the store instead of formed into a
+linked list, and the continuations are in a state which then turns into a continuation
+address.
+
+For clarity, we divide the store into two separate stores: the binding stores @BStore@, and
+the continuation store @KStore@.
+The binding store maps binding addresses to sets of closure values, whereas the continuation
+store maps continuation addresses to sets of continuations.
+We then define a generic class @Store[K,V]@ that performs joining when updating elements
+in a store (below). By parameterizing @Store[K,V]@ with @[BAddr, Storable]@ and
+@[KAddr, Cont]@, we obtain @BStore@ and @KStore@, respectively.
+
+We note that the \note{store? update function? can a store be monotonic?} is monotonic; it continuously grows and never shrinks. This property
+guarantees that in analyzing the \note{update function? store? pick the right word}, a fixed point can always be reached.
 
@@
 
-The co-domain of binding stores @Storable@ is the same
-as we defined for CESK machines.
-The co-domain of continuation stores @Cont@ is comprised of
-a frame @Frame@ and a continuation address @KAddr@.
+The codomain of all binding stores @Storable@ is the same
+as previously defined for CESK machines.
+The codomain of continuation stores @Cont@, on the other hand, is comprised of
+a @Frame@ object and a continuation address @KAddr@.
 To mimic the runtime call stack, @KAddr@ plays the role of
 representing the remaining stack frames.
-TODO: talk about nondet of cont deref?
+\todo{talk about nondet of cont deref?}
 
@@
 As a consequence, the components of states are also changed:
-the store is divided into binding store and continuation store;
+the store is divided into binding stores and continuation stores;
 the continuation stack becomes an address.
@@
 
-\subsubsection{Single-step Transition}
-
-Since dereferencing an address becomes non-deterministic, our @atomicEval@
+\subsubsection{Single-Step Transition}
+
+Since dereferencing an address becomes nondeterministic, our @atomicEval@
 function also turns to be nondeterministic. Given an atomic expression @e@,
@@
 The structure of function @step@ is similar to the concrete CESK machine,
-except non-determinism which makes @step@ return a list of reachable successor states.
+except nondeterminism which makes @step@ return a list of reachable successor states.
 In the beginning of @step@, we call function @tick@ to update the time.
@@
 By doing so, we recover the call stack as precise as runtime's
-(so called pushdown control-flow analysis), but since
+(so called pushdown control flow analysis), but since
 the stack is unbounded, the analysis is potentially not computable.
@@
 The state transition function @step@ shows below is still
-non-deterministic, but the only non-determinism happens when dereferencing the callee @f@
+nondeterministic, but the only nondeterminism happens when dereferencing the callee @f@
 from function applications @App(f, ae)@.
@@
 Here we are facing a problem:
-the underlying semantics of AAM is fundamentally non-deterministic.
+the underlying semantics of AAM is fundamentally nondeterministic.
 But the evaluation context (the @Frame@ in our program) is not
-non-deterministic, and the work list @todo@ actually implicitly handles the non-determinism,
+nondeterministic, and the work list @todo@ actually implicitly handles the nondeterminism,
 if we simply refunctionalize the frames to functions, it would not helps toward
@@
 In order to address this problem, our first step of transformations
-is to linearize all the non-deterministic choices.
+is to linearize all the nondeterministic choices.
 This step turns the @step@ function to be deterministic.
@@
 For the rest of the paper, to differentiate the two continuations, we would call
-the first one as normal continuation, and the second one as non-deterministic
+the first one as normal continuation, and the second one as nondeterministic
 continuations.
@@
 first closure in that set.
-A new non-deterministic continuation @new_ndk@ is also constructed,
+A new nondeterministic continuation @new_ndk@ is also constructed,
 which contains the rest of closures, the values of argument @ae@, the store,
@@
 In the second case shown in below we will see how @NDCont@ deal with
-the non-determinism.
+the nondeterminism.
 
@@
 
-But, since we have added non-determinism into state, an empty list of frames means
+But, since we have added nondeterminism into state, an empty list of frames means
 we have the reached the halt of \textit{one computation path}, and we should
@@
 than current one.
-The non-determinism continuation @new_ndk@ is also updated by removal of that closure.
+The nondeterminism continuation @new_ndk@ is also updated by removal of that closure.
 \end{itemize}
@@
 potentially unexplored states are embedded in to the continuation for
-non-determinism.
+nondeterminism.
 The termination of the analysis is when we reach a @NDState@ that the expression
-is atomic, and both the normal continuation and non-determinism continuation are
+is atomic, and both the normal continuation and nondeterminism continuation are
 empty list. This corresponds to the case that @todo@ list is empty.
@@
 language, which the function is called from @drive_step@ when encountering an atomic
-expression; @ndcontinue@ dispatches non-deterministic continuations @NDCont@,
+expression; @ndcontinue@ dispatches nondeterministic continuations @NDCont@,
 and is invoked from @continue@ when the normal stack is empty.
@@
 \item Since our final target is abstract definitional interpreters while still
-non-deterministic, start from this section, our abstract semantic artifact
+nondeterministic, start from this section, our abstract semantic artifact
 returns a set of final values instead of collected states.
@@
 For clarity, in this section, we first show how to refunctionalize the normal continuation and
-non-deterministic continuation without collecting semantics.
+nondeterministic continuation without collecting semantics.
 Then adding an appropriate caching algorithm to make sure the analysis will
 terminate at some fixed-point is just changing the program to cache-passing style.
-Finally, we examine how computable pushdown control-flow analysis is established
+Finally, we examine how computable pushdown control flow analysis is established
 through refunctionalization and caching.
@@
 An instance of @VS@ represents the computational result of one path in
-the non-deterministic evaluation.
+the nondeterministic evaluation.
 The reason that we include a store is there might be a case two pathes have the
 same values but different accumulated side effects such as memory allocations.
-Since the whole computation is non-deterministic, the type of final result
+Since the whole computation is nondeterministic, the type of final result
 values is a collection @Set[VS]@. For convenience, we define it as answer type @Ans@:
@@
 
-\textbf{Non-Determinism Abstraction.}
-To handle the inevitable non-determinism, we defined a generic function @nd@ as follows:
+\textbf{NonDeterminism Abstraction.}
+To handle the inevitable nondeterminism, we defined a generic function @nd@ as follows:
 \begin{lstlisting}
@@
 \textbf{First Version of Evaluation Function.}
-Since our first version of @aeval@ just implements the non-deterministic
+Since our first version of @aeval@ just implements the nondeterministic
 evaluation, and given the defined @nd@, the evaluation function @aeval@ can be easily
@@
 
-\subsection{Pushdown Control-Flow Analysis, Revisited} \label{pdcfarevisit}
+\subsection{Pushdown Control Flow Analysis, Revisited} \label{pdcfarevisit}
 
@@
 the runtime call-stack of the analyzed program.
-A pushdown control-flow analysis provides as exact as runtime return-flow
-when analyzing the program, but traditional control-flow analysis collapses
+A pushdown control flow analysis provides as exact as runtime return-flow
+when analyzing the program, but traditional control flow analysis collapses
 the state space into a finite space thus causes imprecise stack modeling.
 
-To see how traditional control-flow analysis suffers from spurious return-flows,
+To see how traditional control flow analysis suffers from spurious return-flows,
 we can consider the following example:
@@
 defunctionalized continuation.
-Unfortunately, the abstract abstract machines is non-deterministic, where
-the non-determinism is handled by a work list.
-We can turn this explicit non-determinism into the reduction context of
+Unfortunately, the abstract abstract machines is nondeterministic, where
+the nondeterminism is handled by a work list.
+We can turn this explicit nondeterminism into the reduction context of
 AAM.
 
-\textbf{Pushdown Control-flow Analysis.}
+\textbf{Pushdown Control flow Analysis.}
 In recent years, there are significant efforts \cite{vardoulakis2010cfa2, earl2012introspective,

### 2018-03-16 paper.tex
Forgot to push :/

--- 2018-03-16 paper.tex
@@
 It is worth noting that the continuation class @Kont@ is defined as a list of
-frames.
-The frame can be considered as evaluation context in reduction semantics.
-The frame class @Frame@ stores the information of a single call-site,
-i.e., the information that can be used to resume the interrupted computation.
-A @Frame@ constitutes of a variable name @x@ to be bind later, a control expression
-that resumes to, and its environment. (TODO: mention that frame list vs KArg/KFun)
-
-\subsubsection{Single-step Transition}
-Before going into describing how the machine evaluates expressions, we firstly define several helper functions.
-As we mentioned in section \ref{anfsyntax}, the atomic expressions are either a variable, or
-a literal @lambda@ term. So the atomic expression evaluator @atomicEval@ handles these two
-cases and evaluates atomic expressions to closures in a straightforward way.
+frames, where a frame can be considered an evaluation context in reduction semantics.
+We represent frames using the @Frame@ class, which stores the information of a single
+call-site, i.e., the information that can be used to resume the interrupted computation.
+A @Frame@ constitutes a variable name @x@ to be bound later, a control expression
+to which the program may resume, and its environment. \note{JD: someone check this and make sure I got it right}
+\todo{mention that frame list vs KArg/KFun}
+
+\subsubsection{Single-Step Transition}
+Before describing how the machine evaluates expressions, we must first define several helper
+functions. As mentioned in Section~\ref{anfsyntax}, atomic expressions are either a variable
+or a literal @lambda@ term. As such, the atomic expression evaluator @atomicEval@ handles
+these two cases and evaluates atomic expressions to closures in a straightforward way.
 The @alloc@ function generates a fresh address, and always allocates unique integer
 in the domain of store.
-The @isAtomic@ function is used as a predicate of the expression is atomic.
+The @isAtomic@ function is used as a predicate to determine if the expression is atomic.
 
@@
 Now, we can faithfully describe the state transition function @step@,
-that given a machine state, determines its successor state.
+that, when given a machine state, determines its successor state.
 

### 2018-03-16 paper.tex
Reworking contributions+outline

@@
 
-\subsection{Contributions}
+\note{rest of intro goes here}
+
+The rest of this paper is structured as follows: we begin by reviewing some background
+necessary for this work in Section~\ref{background}, as well as introducing some of the basic
+code structures used throughout the paper. We then address the main contribution of this
+paper, which is the filling of the gap between small-step AAM and big-step ADI by developing a
+series of systematic transformations. Those transformations are summarized here, with their
+associated section:
+
+\begin{itemize}
+\item We show how \textbf{linearization} (Section~\ref{linear}) can be used to...
+\item In Section~\ref{fusing}, we then present the \textbf{fusing} transformation, and present...
+\item Section~\ref{disen} discusses \textbf{disentangling}, which shows...
+\item \textbf{Refunctionalization} is shown in Section~\ref{refunc}, and brings a new understanding of pushdown control-flow analysis.
+In this section, we present an abstract interpreter obtained from refunctionalizing a small-step abstracting abstract machine which naturally exhibits the pushdown control-flow property (as refunctionalization requires abstract execution to be sequentialized).
+\item The last transformation we show is that of transforming a definitional abstract interpreter to a \textbf{direct-style} interpreter (Section~\ref{directstyle}).
+
+\end{itemize}
+
+These transformations are used throughout the paper, with refunctionalization and
+defunctionalization of abstract interpreters playing important roles for the call stack of
+the analyzed language (i.e., Scala)\note{right?}. All of this insight allows the revelation of
+precisely \emph{why} small-step AAM does not have the pushdown control-flow property
+(without extra effort), as well as explains why ADI naturally extends pushdown control-flow
+property from its defining language.
+
+We complete this paper by discussing related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.
+
+% \subsection{Contributions}
 
@@
 
-We summarize our contributions as follows:
-
-\begin{itemize}
-\item The main contribution of this paper is the filling of the gap
-  between small-step AAM and big-step ADI by developing a series of
-  systematic transformations.
-  We show that the correspondence between abstract
-  machines and interpreters exists not only in concrete semantics artifact,
-  but also in abstract semantics artifact.
-  Galois connection?
-  \note{JD: Section number?}
-
-\item A new understanding of pushdown control-flow analysis by
-  refunctionalization.
-  We present an abstract interpreter obtained from refunctionalizing
-  a small-step abstracting abstract machine.
-  This abstract interpreter naturally exhibits the pushdown control-flow property,
-  as the refunctionalization forces us to sequentialize the
-  abstract execution.
-  We also analyze its correctness, soundness, and termination.
-
-\item Further, we show that defunctionalization and refunctionalization
-  of abstract interpreters play important roles for the call stack of the analyzed language.
-  Given this insight, we reveal precisely \emph{why} small-step AAM has no pushdown
-  control-flow property (without extra efforts), as well as explain
-  why ADI naturally extends pushdown control-flow property from its
-  defining language.
-
-\end{itemize}
-
-\subsection{Outline}
-
-\note{TR: can this be merged with contributions? (they currently lack section refs)}
-The remainder of this paper is organized as follows.
-We begin by describing the target language we will analyze in Section~\ref{background},
-including a concrete CESK machine as small-step operational semantics for it.
-We continue with a review of abstracting abstract machines (AAM) as an abstract interpretation
-counterpart of CESK machines. Following this, Section~\ref{linear} shows a conversion from the
-nondeterministic choices of closures to another layer of continuations,
-therefore linearizing the abstract execution.
-Section~\ref{fusing} then discusses fusing the driver and step functions to a single
-function.
-We further identify in Section~\ref{disen} the first-order data type representations of
-continuations, and disentangle the code blocks for dispatching data types to separate
-functions.
-Section~\ref{refunc} then represents the refunctionalized abstract abstract machines.
-In this section, the continuations represented by data types become higher-order
-functions.
-In Section~\ref{directstyle}, we transform the refunctionalized AAM to two
-different direct style, one is implemented with monads, the other is using
-delimited controls.
-
-Finally, we discuss related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.
+% We summarize our contributions as follows:
+
+% \begin{itemize}
+% \item The main contribution of this paper is the filling of the gap
+%   between small-step AAM and big-step ADI by developing a series of
+%   systematic transformations.
+%   We show that the correspondence between abstract
+%   machines and interpreters exists not only in concrete semantics artifacts,
+%   but also in abstract semantics artifacts.
+%   Galois connection?
+%   \note{JD: Section number?}
+
+% \item A new understanding of pushdown control-flow analysis by
+%   refunctionalization.
+%   We present an abstract interpreter obtained from refunctionalizing
+%   a small-step abstracting abstract machine.
+%   This abstract interpreter naturally exhibits the pushdown control-flow property,
+%   as the refunctionalization forces us to sequentialize the
+%   abstract execution.
+%   We also analyze its correctness, soundness, and termination.
+
+% \item Further, we show that defunctionalization and refunctionalization
+%   of abstract interpreters play important roles for the call stack of the analyzed language.
+%   Given this insight, we reveal precisely \emph{why} small-step AAM has no pushdown
+%   control-flow property (without extra efforts), as well as explain
+%   why ADI naturally extends pushdown control-flow property from its
+%   defining language.
+
+% \end{itemize}
+
+% \subsection{Outline}
+
+% \note{TR: can this be merged with contributions? (they currently lack section refs)}
+% The remainder of this paper is organized as follows.
+% We begin by describing the target language we will analyze in Section~\ref{background},
+% including a concrete CESK machine as small-step operational semantics for it.
+% We continue with a review of abstracting abstract machines (AAM) as an abstract interpretation
+% counterpart of CESK machines. Following this, Section~\ref{linear} shows a conversion from the
+% nondeterministic choices of closures to another layer of continuations,
+% therefore linearizing the abstract execution.
+% Section~\ref{fusing} then discusses fusing the driver and step functions to a single
+% function.
+% We further identify in Section~\ref{disen} the first-order data type representations of
+% continuations, and disentangle the code blocks for dispatching data types to separate
+% functions.
+% Section~\ref{refunc} then represents the refunctionalized abstract abstract machines.
+% In this section, the continuations represented by data types become higher-order
+% functions.
+% In Section~\ref{directstyle}, we transform the refunctionalized AAM to two
+% different direct styles: one implemented with monads, the other using
+% delimited controls.
+
+% Finally, we discuss related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.
 

### 2018-03-17 paper.tex
Typos

@@
 these two cases and evaluates atomic expressions to closures in a straightforward way.
-The @alloc@ function generates a fresh address, and always allocates unique integer
-in the domain of store.
+The @alloc@ function generates a fresh address, and always allocates a unique integer
+in the domain of @store@.
 The @isAtomic@ function is used as a predicate to determine if the expression is atomic.
@@
 
-Now, we can faithfully describe the state transition function @step@,
-that, when given a machine state, determines its successor state.
+We can now faithfully describe the state transition function @step@,
+which when given a machine state, determines its successor state.
 
@@
 In the first case, the control of current state is a @Let@ expression,
-and its right-hand side is an function application.
-By calling evaluator @atomicEval@, we obtain the closure that callee @f@ stands for.
+and its right-hand side is a function application.
+By calling the @atomicEval@ evaluator, we obtain the closure for which the callee @f@ stands.
 The successor state's control then transfers to the @body@ expression of closure,

### 2018-03-17 paper.tex
2.2.2

@@
 
-In the first case, the control of current state is a @Let@ expression,
-and its right-hand side is a function application.
+As shown and previously discussed, we examine the only two cases which the state may be.
+
+\textbf{\texttt{Let} expression.} In the first @case@ statement shown in the previous code,
+the control of the current state matches as a @Let@ expression, with its right-hand side a
+function application.
 By calling the @atomicEval@ evaluator, we obtain the closure for which the callee @f@ stands.
-The successor state's control then transfers to the @body@ expression of closure,
-with an updated environment and an update store. The new environment is extended
-from closure's environment and mapped @v@ to a fresh address @addr@.
+The successor state's control then transfers to the @body@ expression of the closure
+with an updated environment and store. The new environment is extended
+from the closure's environment and mapped from @v@ to a fresh address @addr@.
 The new store is extended with @addr@ mapping to the value of @ae@,
-which is evaluated from @atomicEval@.
-In the meanwhile, a new frame @frame@ is pushed onto the stack @k@.
-This @frame@ contains the variable name @x@ at left-hand side position of @Let@,
+which in turn is evaluated from @atomicEval@.
+Finally, a new frame @frame@ is pushed onto the stack @k@, where
+@frame@ contains the variable name @x@ at the left-hand position of the @Let@,
 the body expression of @Let@, and the lexical environment of the body expression.
 
-The second case for @step@ is the control is transferred to an atomic expression.
-We extract the top frame of continuations firstly.
-The control (that is an atomic expression) of current state is the evaluated term
-that being bind to the variable @x@ from the top frame.
+\textbf{Atomic expression.} If the state is not a @Let@ expression, then it must be
+an atomic expression, as seen in the above code. In this scenario,
+we begin by extracting the top frame of all available continuations.
+The control (i.e., an atomic expression) of the current state is the evaluated term
+that is being bound to the variable @x@ from the top frame.
 The environment and store are updated with @x@ mapping to the closure value of @ae@.
-Then the successor state is transferred to expression @e@ from the top frame,
+Finally, the successor state is transferred to expression @e@ from the top frame,
 which is the body of a @Let@ expression, with the updated environment, store, and
-the rest of stack @ks@.
+the rest of the stack @ks@.
 

### 2018-03-17 paper.tex
update intro

@@
 
-Abstracting abstract machines (AAMs) and abstract definitional interpreters (ADIs) are, broadly speaking, both examples of abstract interpreters.
-\note{JD: need more bullet points here and I can fill this in}
+Abstracting abstract machines (AAM) is a systematic methodology proposed by \citeauthor{van2012systematic}
+for constructing abstract interpreters of higher-order functional languages \cite{van2012systematic, van2010abstracting}. 
+Given a concrete small-step abstract machine, (e.g, the CESK machine, Krivine's machine or etc), 
+by allocating continuations in the store and bound both the value addresses and continuation
+addresses to be finite, we obtain an abstract interpreter with finite state space which 
+can be used for performing sound static analysis.
+One can further instantiate different polyvariant control-flow analysis by using different address allocators \cite{Gilray:2016:ACP:2951913.2951936}.
+
+Applying the same idea on big-step definitional interpreters, \citeauthor{darais2017abstracting} 
+build abstract definitional interpreters (ADI) that are written in monadic style \cite{darais2017abstracting}
+By changing the underlying
+monads, the definition of interpreter has no change, but we can recover different semantics 
+including the concrete semantics and various abstract semantics.
+
+Broadly speaking, abstract abstract machines and abstract definitional interpreters are both
+abstract interpreters. They are obtained by applying a combination of abstractions to their
+concrete counterparts, abstract machines and definitional interperters, perspectively.
+The relations between abstract machines, definitional interpreters, and monadic interpreters 
+is intensively studied by Danvy and his collaborator in the past years 
+\cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268}.
+
+But in terms of their abstract semantic artifacts, 
+the relations between small-step abstract abstract machines and big-step abstract 
+definitional interpreters, and the question that how can we interdrive each other is not
+well studied.
+In this functional peral, we practical answer this question by presenting a series of 
+syntactical transformations on the program from small-step abstract abstract machines 
+to big-step abstract definitional interpreters.
+
+\subsection{outline}
 
@@
 \item Section~\ref{disen} discusses \textbf{disentangling}, which shows...
-\item \textbf{Refunctionalization} is shown in Section~\ref{refunc}, and brings a new understanding of pushdown control-flow analysis.
-In this section, we present an abstract interpreter obtained from refunctionalizing a small-step abstracting abstract machine which naturally exhibits the pushdown control-flow property (as refunctionalization requires abstract execution to be sequentialized).
-\item The last transformation we show is that of transforming a definitional abstract interpreter to a \textbf{direct-style} interpreter (Section~\ref{directstyle}).
+\item \textbf{Refunctionalization} is shown in Section~\ref{refunc}, and brings a new understanding 
+  of pushdown control-flow analysis. In this section, we present an abstract interpreter obtained 
+  from refunctionalizing a small-step abstracting abstract machine which naturally exhibits the 
+  pushdown control-flow property (as refunctionalization requires abstract execution to be sequentialized).
+\item The last transformation we show is that of transforming a definitional abstract interpreter 
+  to a \textbf{direct-style} interpreter (Section~\ref{directstyle}).
 
@@
 defunctionalization of abstract interpreters playing important roles for the call stack of
-the analyzed language (i.e., Scala)\note{right?}. All of this insight allows the revelation of
+the analyzed language. All of this insight allows the revelation of
 precisely \emph{why} small-step AAM does not have the pushdown control-flow property
@@
 
-% \item A new understanding of pushdown control-flow analysis by
+% \item A new understanding of pushdown control flow analysis by
 %   refunctionalization.

### 2018-03-17 paper.tex
update intro

@@
 Applying the same idea on big-step definitional interpreters, \citeauthor{darais2017abstracting} 
-build abstract definitional interpreters (ADI) that are written in monadic style \cite{darais2017abstracting}
-By changing the underlying
-monads, the definition of interpreter has no change, but we can recover different semantics 
-including the concrete semantics and various abstract semantics.
+build abstract definitional interpreters (ADI) that are written in monadic style \cite{darais2017abstracting}.
+One of the advantages of monadic interpreter is modular and composable.
+By changing the underlying monads, the definition of interpreter has no change, 
+but we can recover different semantics including the concrete semantics and various 
+abstract semantics such as context-sensitivity and abstract garbage collection.
 
@@
 concrete counterparts, abstract machines and definitional interperters, perspectively.
-The relations between abstract machines, definitional interpreters, and monadic interpreters 
-is intensively studied by Danvy and his collaborator in the past years 
-\cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268}.
+In concrete world, the relations among reduction semantics, abstract machines, definitional interpreters,
+and monadic interpreters is intensively studied by Danvy and his collaborator in the past years 
+\cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}.
+The abstract machines implement structural operational semantics in continuation style, 
+where the reduction contexts are defunctionalized continuations. One can derive definitional
+interpreters by refunctionalizing the evaluation contexts of abstract machines; 
+and by defunctionalizing the higher-order functions one may obtain abstract machines in the 
+reversed direction.
 
@@
 well studied.
-In this functional peral, we practical answer this question by presenting a series of 
-syntactical transformations on the program from small-step abstract abstract machines 
-to big-step abstract definitional interpreters.
-
-\subsection{outline}
+The fundamental difference between concrete semantics artifacts and abstract semantics 
+artifacts is non-determinism. In addition to insure termination, abstract semantics 
+artifacts usually equipped with a cache of explored state that is guaranteed to reach 
+the least fixed-point eventually.
+In this functional peral, we practically answer the question and relate AAM and ADI 
+by presenting a series of syntactical transformations on the program from
+small-step abstract abstract machines to big-step abstract definitional interpreters.
+
+\subsection{Outline}
 

### 2018-03-17 paper.tex
2.3.1

--- 2018-03-17 paper.tex
@@
   and monadification (or un-CPS to direct-style). Linearization expresses
-  non-deterministic choices by first-order data types, after which refunctionalization
+  nondeterministic choices into first-order data types, after which refunctionalization
   sequentializes the evaluation order by higher-order functions.
@@
 
-Abstracting abstract machines (AAM) is a systematic methodology proposed by \citeauthor{van2012systematic}
-for constructing abstract interpreters of higher-order functional languages \cite{van2012systematic, van2010abstracting}. 
-Given a concrete small-step abstract machine, (e.g, the CESK machine, Krivine's machine or etc), 
-by allocating continuations in the store and bound both the value addresses and continuation
-addresses to be finite, we obtain an abstract interpreter with finite state space which 
-can be used for performing sound static analysis.
-One can further instantiate different polyvariant control-flow analysis by using different address allocators \cite{Gilray:2016:ACP:2951913.2951936}.
-
-Applying the same idea on big-step definitional interpreters, \citeauthor{darais2017abstracting} 
-build abstract definitional interpreters (ADI) that are written in monadic style \cite{darais2017abstracting}
-By changing the underlying
-monads, the definition of interpreter has no change, but we can recover different semantics 
-including the concrete semantics and various abstract semantics.
-
-Broadly speaking, abstract abstract machines and abstract definitional interpreters are both
-abstract interpreters. They are obtained by applying a combination of abstractions to their
-concrete counterparts, abstract machines and definitional interperters, perspectively.
-The relations between abstract machines, definitional interpreters, and monadic interpreters 
-is intensively studied by Danvy and his collaborator in the past years 
-\cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268}.
-
-But in terms of their abstract semantic artifacts, 
-the relations between small-step abstract abstract machines and big-step abstract 
-definitional interpreters, and the question that how can we interdrive each other is not
-well studied.
-In this functional peral, we practical answer this question by presenting a series of 
-syntactical transformations on the program from small-step abstract abstract machines 
-to big-step abstract definitional interpreters.
-
-\subsection{outline}
+Abstracting abstract machines (AAMs) and abstract definitional interpreters (ADIs) are, broadly speaking, both examples of abstract interpreters.
+\note{JD: need more bullet points here and I can fill this in}
 
@@
 \item Section~\ref{disen} discusses \textbf{disentangling}, which shows...
-\item \textbf{Refunctionalization} is shown in Section~\ref{refunc}, and brings a new understanding 
-  of pushdown control-flow analysis. In this section, we present an abstract interpreter obtained 
-  from refunctionalizing a small-step abstracting abstract machine which naturally exhibits the 
-  pushdown control-flow property (as refunctionalization requires abstract execution to be sequentialized).
-\item The last transformation we show is that of transforming a definitional abstract interpreter 
-  to a \textbf{direct-style} interpreter (Section~\ref{directstyle}).
+\item \textbf{Refunctionalization} is shown in Section~\ref{refunc}, and brings a new understanding of pushdown control flow analysis.
+In this section, we present an abstract interpreter obtained from refunctionalizing a small-step abstracting abstract machine which naturally exhibits the pushdown control-flow property (as refunctionalization requires abstract execution to be sequentialized).
+\item The last transformation we show is that of transforming a definitional abstract interpreter to a \textbf{direct-style} interpreter (Section~\ref{directstyle}).
 
@@
 defunctionalization of abstract interpreters playing important roles for the call stack of
-the analyzed language. All of this insight allows the revelation of
+the analyzed language (i.e., Scala)\note{right?}. All of this insight allows the revelation of
 precisely \emph{why} small-step AAM does not have the pushdown control-flow property

### 2018-03-17 paper.tex
intro

--- 2018-03-17 paper.tex
@@
 artifacts is non-determinism. In addition to insure termination, abstract semantics 
-artifacts usually equipped with a cache of explored state that is guaranteed to reach 
+artifacts are usually equipped with a cache of explored state that is guaranteed to reach 
 the least fixed-point eventually.
@@
 
+In addition, the abstract abstract machines with unbounded stack naturally corresponds 
+to abstract definitional interpreters. We show after refunctionalizing an AAM with unbounded
+stack and with proper caching algorithm, the pushdown control-flow analysis can be obtained.
+
 \subsection{Outline}
-
-\note{rest of intro goes here}
 
@@
 \begin{itemize}
-\item We show how \textbf{linearization} (Section~\ref{linear}) can be used to...
-\item In Section~\ref{fusing}, we then present the \textbf{fusing} transformation, and present...
-\item Section~\ref{disen} discusses \textbf{disentangling}, which shows...
-\item \textbf{Refunctionalization} is shown in Section~\ref{refunc}, and brings a new understanding 
-  of pushdown control-flow analysis. In this section, we present an abstract interpreter obtained 
-  from refunctionalizing a small-step abstracting abstract machine which naturally exhibits the 
-  pushdown control-flow property (as refunctionalization requires abstract execution to be sequentialized).
-\item The last transformation we show is that of transforming a definitional abstract interpreter 
-  to a \textbf{direct-style} interpreter (Section~\ref{directstyle}).
+\item We show \textbf{linearization} in Section~\ref{linear}. By expressing the 
+  non-deterministic choices as first-order data type, we linearize the execution 
+  of abstract abstract machines, therefore the transition of machine state becomes deterministic.
+    Notably, this introduces anther layer of controls that also will be refunctionalized later.
+
+\item In Section~\ref{fusing}, we then present the \textbf{fusing} transformation. 
+  The fusing transformation simply combines the single-step function @step@ and the driving
+  function into one, but keeps all the machine state representation.
+
+\item Section~\ref{disen} discusses \textbf{disentangling}, which disassemble the fused AAM to be
+  several individual functions, and each function handles one data type represented continuation.
+
+\item \textbf{Refunctionalization} is shown in Section~\ref{refunc} which sequentializes
+  the order of abstract execution by higher-order functions.
+  For clarity, we first present the refunctionalized AAM without caching. Then we adopt
+    a different caching algorithm to guarantee the termination of abstract interpretation.
+  In this section, we also review pushdown control-flow analysis and examine how computable 
+    and precise call/return match is obtained through these transformations so far.
+
+\item The last transformation we show is that of transforming the refunctionalized AAM
+  to a \textbf{direct-style} interpreter (Section~\ref{directstyle}) by using delimited
+    controls.
 
@@
 defunctionalization of abstract interpreters playing important roles for the call stack of
-the analyzed language. All of this insight allows the revelation of
-precisely \emph{why} small-step AAM does not have the pushdown control-flow property
-(without extra effort), as well as explains why ADI naturally extends pushdown control-flow
-property from its defining language.
+the analyzed language. By refunctionalization, the call stack of the analyzed language
+is blended into the call stack of the defining language. This provides another perspective to 
+explain why \citeauthor{darais2017abstracting}'s abstract definitional interpreters 
+is able to extend the the pushdown control-flow property from its defining language.
 
@@
 
-1) The current control expression is @Let@,
+\begin{itemize}
+  \item The current control expression is @Let@,
 and then the result of @App(f, ae)@ will be bound to variable @x@.
@@
 
-2) An atomic expression @ae@ sits on the control position of the state.
+  \item An atomic expression @ae@ sits on the control position of the state.
 The value of @ae@ is being returned to its caller.
@@
 also comes from the frame.
+\end{itemize}
 

### 2018-03-17 paper.tex
Fixing to itemize, etc.

@@
 
-\textbf{\texttt{Let} expression.} In the first @case@ statement shown in the previous code,
+\begin{itemize}
+
+\item In the first @case@ statement shown in the previous code,
 the control of the current state matches as a @Let@ expression, with its right-hand side a
@@
 
-\textbf{Atomic expression.} If the state is not a @Let@ expression, then it must be
+\item If the state is not a @Let@ expression, then it must be
 an atomic expression, as seen in the above code. In this scenario,
@@
 the rest of the stack @ks@.
+
+\end{itemize}
 
@@
 \subsubsection{Allocating Addresses}
-Up to now, we did not describe about allocating addresses in stores and the time stamp @Time@.
-But in abstract interpretation, they are key ingredients to achieve
-analyses with different sensitivities, as well as to a finite state space analysis\cite{Gilray:2016:ACP:2951913.2951936}.
+Up to this point, we have not described allocating addresses in stores,
+nor handling the time stamp @Time@.
+In abstract interpretation, however, these are key ingredients to achieve
+analyses with different sensitivities, as well as to perform a finite
+state space analysis\cite{Gilray:2016:ACP:2951913.2951936}.
 To effectively approximate the runtime behavior, we introduce a finite
-program contour @time@ that encodes the program execution history.
-Here, we use a list of execution contexts (expressions) to represent, and
-as we will see in section \ref{kcfainst}, by applying different @tick@
-functions on time stamp we may obtain a family of analyses.
+program counter @time@ that encodes the program execution history.
+We use a list of execution contexts (expressions) to represent this, and
+as we will see in Section~\ref{kcfainst}, by applying different @tick@
+functions on the timestamp, we are able to obtain a family of analyses.
 
@@
 
-As we mentioned previously, the space of addresses has to be finite in AAM.
-Binding addresses are parameterized by variable names and the time of creating this binding,
-which are both finite.
+As previously mentioned, the space of addresses must be finite in AAM.
+Binding addresses are parameterized by variable names and the creation time of the binding,
+both of which are finite.
 Continuation addresses @KAddr@ has two variants:
@@
 One may ask, why not also keep track of the time in continuation addresses?
-(TODO: why target expression, allocation polyvariance)
+\todo{why target expression, allocation polyvariance}
 
@@
 
-Two helper functions @allocBind@ and @allocKont@ that will be used later
-to allocate binding addresses and continuation addresses.
+We introduce two helper functions, @allocBind@ and @allocKont@, which will be
+used to allocate binding addresses and continuation addresses.
 
@@
 Since dereferencing an address becomes nondeterministic, our @atomicEval@
-function also turns to be nondeterministic. Given an atomic expression @e@,
-@atomicEval@ returns a set of storable values, i.e., closures, to the caller.
-If the expression is just a @lambda@ term, then the returned set is a singleton.
+function (below) is also nondeterministic. Given an atomic expression @e@,
+@atomicEval@ returns a set of storable values (i.e., closures) to the caller.
+If the expression is simply a @lambda@ term, the returned set is a singleton.
 
@@
 The structure of function @step@ is similar to the concrete CESK machine,
-except nondeterminism which makes @step@ return a list of reachable successor states.
-In the beginning of @step@, we call function @tick@ to update the time.
-Then we have two cases to consider:
+except the nondeterminism which makes @step@ return a list of reachable successor states.
+We have two cases to consider (code shown below):
 
 \begin{itemize}
-  \item The current control expression is @Let@,
-and then the result of @App(f, ae)@ will be bound to variable @x@.
-Firstly we retrieve the set of closures that @f@ may represent.
-For each closure in the set, we do almost the same as what we did in the concrete
-CESK machines. But note that since the continuation is allocated on store (@kstore@),
-a new continuation address @new_kaddr@ is constructed, and a new frame
-@Frame(x, e, env)@ paired with the current continuation address @kaddr@
-is joined into @new_kaddr@. Finally, a list of successors states is generated.
+  \item If the current control expression is a @Let@, then
+the result of @App(f, ae)@ will be bound to variable @x@. In this case, we retrieve the set
+of closures that @f@ may represent. For each closure in the set, we perform nearly the same
+operations as in the concrete CESK machines, with an important difference: the
+continuation is allocated on the store @kstore@, so a new continuation address @new_kaddr@
+must be constructed and a new frame @Frame(x, e, env)@ paired with the current continuation
+address @kaddr@ is merged into @new_kaddr@. Finally, a list of successor states is generated.
 

### 2018-03-17 paper.tex
tweak section 3
Editing intro

@@
 
-Abstracting abstract machines (AAM) is a systematic methodology proposed by \citeauthor{van2012systematic}
-for constructing abstract interpreters of higher-order functional languages \cite{van2012systematic, van2010abstracting}. 
-Given a concrete small-step abstract machine, (e.g, the CESK machine, Krivine's machine or etc), 
-by allocating continuations in the store and bound both the value addresses and continuation
-addresses to be finite, we obtain an abstract interpreter with finite state space which 
+Abstracting abstract machines (AAM) is a systematic methodology proposed by~\citeauthor{van2012systematic}
+for constructing abstract interpreters of higher-order functional languages~\cite{van2012systematic, van2010abstracting}.
+Given a concrete small-step abstract machine, (e.g, the CESK machine, Krivine's machine, etc.),
+by allocating continuations in the store and binding both the value addresses and continuation
+addresses to be finite, we obtain an abstract interpreter with a finite state space which
 can be used for performing sound static analysis.
-One can further instantiate different polyvariant control-flow analysis by using different address allocators \cite{Gilray:2016:ACP:2951913.2951936}.
-
-Applying the same idea on big-step definitional interpreters, \citeauthor{darais2017abstracting} 
-build abstract definitional interpreters (ADI) that are written in monadic style \cite{darais2017abstracting}.
-One of the advantages of monadic interpreter is modular and composable.
-By changing the underlying monads, the definition of interpreter has no change, 
-but we can recover different semantics including the concrete semantics and various 
+One can further instantiate different polyvariant control flow analyses by using different address allocators \cite{Gilray:2016:ACP:2951913.2951936}.
+
+Applying the same idea to big-step definitional interpreters, \citeauthor{darais2017abstracting}
+built abstract definitional interpreters (ADI) that are written in monadic style \cite{darais2017abstracting}.
+One of the advantages of a monadic interpreter is that it is modular and composable.
+By changing the underlying monads, the definition of the interpreter is not modified,
+but we can recover different semantics, including the concrete semantics and various
 abstract semantics such as context-sensitivity and abstract garbage collection.
@@
 abstract interpreters. They are obtained by applying a combination of abstractions to their
-concrete counterparts, abstract machines and definitional interperters, perspectively.
-In concrete world, the relations among reduction semantics, abstract machines, definitional interpreters,
-and monadic interpreters is intensively studied by Danvy and his collaborator in the past years 
+concrete counterparts, abstract machines and definitional interpreters, respectively.
+In the concrete world, the relations among reduction semantics, abstract machines,
+definitional interpreters, and monadic interpreters is intensively studied by Danvy and his
+collaborators in the past years
 \cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}.
-The abstract machines implement structural operational semantics in continuation style, 
+These abstract machines implement structural operational semantics in continuation style,
 where the reduction contexts are defunctionalized continuations. One can derive definitional
-interpreters by refunctionalizing the evaluation contexts of abstract machines; 
-and by defunctionalizing the higher-order functions one may obtain abstract machines in the 
-reversed direction.
-
-But in terms of their abstract semantic artifacts, 
+interpreters by refunctionalizing the evaluation contexts of abstract machines,
+and by defunctionalizing the higher-order functions, one may obtain abstract machines in the
+reverse direction.
+
+But in terms of their abstract semantics artifacts,
 the relations between small-step abstract abstract machines and big-step abstract
-definitional interpreters, and the question that how can we interdrive each other is not
-well studied.
+definitional interpreters, as well as the ability to derive one from the other is not
+well-studied.
 The fundamental difference between concrete semantics artifacts and abstract semantics
-artifacts is non-determinism. In addition to insure termination, abstract semantics 
+artifacts is nondeterminism. In addition to ensuring termination, abstract semantics
 artifacts are usually equipped with a cache of explored state that is guaranteed to reach
 the least fixed-point eventually.
-In this functional peral, we practically answer the question and relate AAM and ADI 
+In this functional pearl, we practically answer the question and relate AAM and ADI
 by presenting a series of syntactical transformations on the program from
@@
 
-In addition, the abstract abstract machines with unbounded stack naturally corresponds 
+In addition, the abstract abstract machines with unbounded stack naturally correspond
 to abstract definitional interpreters. We show after refunctionalizing an AAM with unbounded
-stack and with proper caching algorithm, the pushdown control-flow analysis can be obtained.
+stack, and with the proper caching algorithm, the pushdown control-flow analysis can be
+obtained.
 
@@
 \item We show \textbf{linearization} in Section~\ref{linear}. By expressing the
-  non-deterministic choices as first-order data type, we linearize the execution 
-  of abstract abstract machines, therefore the transition of machine state becomes deterministic.
-    Notably, this introduces anther layer of controls that also will be refunctionalized later.
+  nondeterministic choices as a first-order data type, we linearize the execution
+  of abstract abstract machines; the transition of machine states therefore becomes deterministic. Notably, this introduces another layer of controls that will also be refunctionalized later.
 
@@
   The fusing transformation simply combines the single-step function @step@ and the driving
-  function into one, but keeps all the machine state representation.
-
-\item Section~\ref{disen} discusses \textbf{disentangling}, which disassemble the fused AAM to be
-  several individual functions, and each function handles one data type represented continuation.
+  function into one, but keeps all the machine state representations.
+
+\item Section~\ref{disen} discusses \textbf{disentangling}, which disassembles the fused AAM to be several individual functions, with each function handling one data type represented by a continuation.
 
@@
   the order of abstract execution by higher-order functions.
-  For clarity, we first present the refunctionalized AAM without caching. Then we adopt
-    a different caching algorithm to guarantee the termination of abstract interpretation.
+  For clarity, we first present the refunctionalized AAM without caching. We then adopt a different caching algorithm to guarantee the termination of abstract interpretation.
   In this section, we also review pushdown control-flow analysis and examine how computable
-    and precise call/return match is obtained through these transformations so far.
+    and precise call/return matching is obtained through these transformations.
 
 \item The last transformation we show is that of transforming the refunctionalized AAM
-  to a \textbf{direct-style} interpreter (Section~\ref{directstyle}) by using delimited
-    controls.
+  to a \textbf{direct-style} interpreter (Section~\ref{directstyle}) by using delimited controls.
 
@@
 Now, we obtain a linearized abstract abstract machine. If we image that the classical
-AAM explores a graph of reachable states, then the linearized AAM flatten this
-graph.
+AAM explores a graph of reachable states, then the linearized AAM just flattens the
+graph to a linear sequence.
 

### 2018-03-17 paper.tex
2.3.3

@@
 
-  \item An atomic expression @ae@ sits on the control position of the state.
-The value of @ae@ is being returned to its caller.
-We dereference the continuation address @kaddr@ and obtain a set of
-continuations @conts@.
-For each continuation in the set, we construct an environment based on the
-environment of the frame (@env_f@), and bind @x@ to a newly created binding
-address @baddr@; also update the store with @baddr@ and
-the values that @ae@ represents. In every generated states,
-the control becomes the expression @e@ in the frame,
-and as we can tell from the name, continuation address @f_kaddr@
-also comes from the frame.
+  \item In the second case, an atomic expression @ae@ sits on the control position of the
+state. Here, the value of @ae@ is being returned to its caller.
+In order to accomplish this, we dereference the continuation address @kaddr@ and obtain a set
+of continuations @conts@. For each continuation in the set, we construct an environment based
+on the environment @env_f@ of the frame, and bind @x@ to a newly created binding address
+@baddr@. We must also update the store with @baddr@ and the values that @ae@ represents. In
+every generated state, the control becomes the expression @e@ in the frame, and as we can
+tell from the name, the continuation address @f_kaddr@ also comes from the frame.
 \end{itemize}

### 2018-03-17 paper.tex
update section 7

@@
 
-In previous section, we presented a refunctionalized AAM that in extended continuation-passing style.
-To obtain a definitional abstract interpreter, we need to transform it further to direct-style.
-At this point, we have two choices:
-1) By monadifying the continuations into monads, we obtain the \citeauthor{darais2017abstracting}'s
-abstracting definitional interpreter in monadic style.
-2) By representing the extended continuation-passing style with delimited continuations,
+In previous section, we presented a refunctionalized AAM with @nd@ operator 
+in extended continuation-passing style.
+To obtain definitional abstract interpreters, we now transform it further to direct-style.
+
+To do this at this point, we actually have multiple choices:
+\begin{itemize}
+  \item By monadifying the continuations into monads, we obtain the abstracting
+    definitional interpreters in monadic style which are what \citeauthor{darais2017abstracting}
+    described \cite{darais2017abstracting}.
+  \item By representing the extended continuation-passing style with delimited controls,
     we derive a new form of abstract interpreters in direct-style.
-This coincidence should not be a surprise since there are extraordinarily rich literatures
-show the correspondence between monads and continuations as of their very beginning era
-(TODO: cite Danvy/abs control/esce of fp).
-In fact, we can also just use the same caching algorithm but with side effect such as
-assignments and mutations to achieve the same definitional interpreter with pushdown
-control flows.
-In this section, we present both of versions using monads and delimited controls.
-
-\subsection{Monadification} \label{monadify}
-
-In this section, we show how to transform the refunctionalized AAM to monadic-style
-abstracting definitional interpreters.
-
-\subsection{Un-CPS} \label{uncps}
+  \item In fact, we may also just use the same caching algorithm but with side effect such as
+    assignments and mutations to update the cache and then achieve the same 
+    definitional interpreter with pushdown control flows. In this case, nondeterminism can be 
+    well handled by for comprehension.
+\end{itemize}
+
+These coincidence should not be a surprise since there are extraordinarily rich literatures
+show the correspondence between monads and continuation-passing style as well as delimited 
+controls since their very beginning era. \todo{cite}
+
+In this section, we present the version that using delimited controls.
+
+\subsection{Un-CPS by Delimited Controls} \label{uncps}
 

### 2018-03-17 paper.tex
Guannan rekt my bad English :D

@@
 Given a concrete small-step abstract machine, (e.g, the CESK machine, Krivine's machine, etc.),
-by allocating continuations in the store and binding both the value addresses and continuation
+by allocating continuations in the store and bounding both the value addresses and continuation
 addresses to be finite, we obtain an abstract interpreter with a finite state space which

### 2018-03-17 paper.tex
section 7

@@
 
-To begin, we now present the concrete syntax of a call-by-value $\lambda$-calculus language
+To begin, we present the concrete syntax of a call-by-value $\lambda$-calculus language
 in ANF~\cite{flanagan1993essence}.
@@
 
-A history of the last $k$ call sites is used as a finite program contour in
-$k$-call-sensitive analysis.
+In $k$-call-sensitive analysis, a history of the last $k$ call sites is
+used as a finite program counter.
 The history is represented as a list of expressions and embedded
@@
 
-Every time before transferring to successor states, we use function @tick@ to
-fresh the time stamp and use this new time stamp for successors when allocating addresses.
-The function @tick@ returns the $k$ front-most expressions given the current state and its
+Before transferring to successor states, we must use the @tick@ function to
+refresh the timestamp, and then use this new timestamp for successors when allocating addresses.
+The @tick@ function returns the $k$ front-most expressions given the current state and its
 time history.
@@
 If we instantiate $k$ to be $0$, the history degenerates to an empty list,
-and we obtain a monovariant analysis, which does
-not differentiate values at different call sites. In this case, the address
+and we obtain a monovariant analysis (i.e., it does
+not differentiate values at different call sites). In this case, the address
 space collapses to the space of variable names.
 Note that regarding the ambiguity in $k$-CFA\cite{Gilray:2016:ACP:2951913.2951936},
-the code here actually implements call+return sensitivity.
+the code here actually implements call+return \note{call+return, call/return, etc.? Check this} sensitivity.
 
@@
 
-To run (analyze) a program, as same as the CESK machines,
-we first use function @inject@ to construct the initial state given the program.
-Note that the initial continuation store has a built-in mapping that maps continuation address
-@Halt@ to an empty set of continuations.
-We also provide an empty program contour as initial time.
+Similar to the CESK machines, to run (analyze) a program
+we first use the @inject@ function to construct the initial state given to the program.
+Note that the initial continuation store has a built-in mapping that maps the continuation
+address for @Halt@ to an empty set of continuations.
+We also provide an empty program counter as our initial time.
 \begin{lstlisting}
@@
 
-However, comparing with the concrete CESK machine, the
+However, in contrast to the concrete CESK machine, the
 @drive@ function performs collecting semantics instead of the valuation
@@
 collects all the intermediate machine states as the program is abstractly executing.
-The following code shows a variant of the work list algorithm to find the fixed-point
-of set of states.
+The following code shows a variant of the worklist algorithm to find the fixed-point
+of the set of states.\note{JD: I changed all of this to worklist instead of work list in favor of the worklist algorithm. Change this back if it's wrong.}
 Function @drive@ always applies function @step@ to the head element @hd@ of
-the work list @todo@ if @hd@ is unseen, then inserts the result of @step@ to
-the rest of work list; and in the meantime adds @hd@ to the explored states set.
-If the work list is empty, then @drive@ just returns the set of reachable states
-up to now.
+the worklist @todo@ if @hd@ is unseen. It then inserts the result of @step@ to
+the rest of worklist, and in the meantime adds @hd@ to the explored states set.
+If the worklist is empty, @drive@ simply returns the set of reachable states
+up to the current execution point.
 
@@
 
-Finally, a user could use function @analyze@ to obtain all reachable states for a given program.
+Finally, a user may invoke the @analyze@ function to obtain all reachable states for a given
+program.
 
@@
 
-In this section, we describe a variant of AAM that leaves the stack to be
-unbounded. That uses a precise call stack as we did in the concrete CESK machine.
+In this section, we describe a variant of AAM that allows the stack to be
+unbounded which uses a precise call stack as we did in the concrete CESK machine.
 Instead of allocating continuations in the store and embedding
@@
 to explicitly model the stack.
-By doing so, we recover the call stack as precise as runtime's
+By doing so, we recover the call stack as precise as runtimes
 (so called pushdown control flow analysis), but since
@@
 For readers who are not familiar with pushdown analysis, we have a detailed
-discussion in Section \ref{pdcfarevisit}.
-The reason we show it here is to establish an artifact with precise call/return match
+discussion in Section~\ref{pdcfarevisit}.
+The reason we show it here is to establish an artifact with precise call/return matching
 used for the next step of transformation.
 
-In the definition of state @State@, the continuation store disappears, and
-component @konts@ becomes a list of frames. The other components keep the same.
+In the definition of @State@, the continuation store disappears, and
+component @konts@ becomes a list of frames. The other components remain unchanged.
 
@@
 
-The state transition function @step@ shows below is still
-nondeterministic, but the only nondeterminism happens when dereferencing the callee @f@
-from function applications @App(f, ae)@.
+The state transition function @step@ shown below is still
+nondeterministic, but the only nondeterminism happening is when dereferencing the callee @f@
+from the function application @App(f, ae)@.
 
@@
 For each closure in the set, a new frame is constructed and pushed onto the stack.
-The code for handling the second case is the same as the concrete CESK machines.
+The code for handling the second case (atomic expressions) is the same as the concrete CESK
+machines.
 
 \textbf{Note on Computability.}
-Unfortunately, even though, other components in states are finite,
-this AAM with unbounded stack is still not computable.
-Because the unbounded stack can grow to arbitrary depth which implies that the state space is possibly
-infinite, therefore the analysis may not terminate for all programs if we simply enumerate reachable states.
-To see this, consider a program that has two mutually recursive functions:
+Unfortunately, even though other components in the states are finite,
+this AAM with an unbounded stack is still not computable.
+This is because the unbounded stack can grow to arbitrary depth which implies that the state
+space is possibly infinite; the analysis may therefore not terminate for all programs if we
+simply enumerate reachable states. To see this, consider a program that has two mutually
+recursive functions:
 
@@
 
-Function @f1@ and @f2@ mutually invokes each other, and the stack will be
-alternatively pushing frame @f1@ and @f2@ onto the top of current stack,
-but no existing two stack components are identical in the state space.
+Function @f1@ and @f2@ mutually invoke each other, so the stack will
+alternate pushing frames @f1@ and @f2@ onto the top of current stack.
+However, no two existing stack components are identical in the state space.
 
@@
 
-In the last section, we showed that by keeping an unabstracted stack in the state
-space we can recover the precise call/return match.
-From this section, we start describing the transformations step by step.
-Our base machine is the AAM with unbounded stack.
-In fact, as later we will transform the stack to higher-order functions
-represented continuations, it does not matter what kind of AAM we start from.
-Because transforming to higher-order functions forces us to sequentialize
-the order of abstract evaluation, and neither need to construct a frame of list
-nor to allocate continuations in the store.
-We choose to start from AAM with unbounded stack just because it has a
-equivalence of stack model to the final abstract definitional interpreters.
+In the previous section, we show that by keeping an unabstracted stack in the state
+space, we can recover the precise call/return match.
+In this and following sections, we begin describing the transformations step by step.
+Our base machine for now is the AAM with unbounded stack, though
+as we will later transform the stack to higher-order functions
+represent \note{``representing'' or ``represented by''?} continuations, it does not matter what kind of AAM we start from.
+This is because transforming to higher-order functions forces us to sequentialize
+the order of abstract evaluation, and neither versions of AAM requires construction of a frame of lists \note{should this be frame of lists? sounds weird}, nor to allocate continuations in the store.
+Thus, our choice to start from an AAM with unbounded stack is motivated simply because it has an equivalent stack model to abstract definitional interpreters (our final target).
 
@@
 
-On another path, here, we introduce direct-style abstracting definitional interpreter
-with delimited control operators @reset@ and @shift@.
-
-TODO: This version does not use delimited controls.
-
-\begin{lstlisting}
-def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans = {
-  val config = Config(e, env, store, time)
-  if (cache.outContains(config)) {
-    return Ans(cache.outGet(config), cache)
-  }
-
-  var new_cache = cache.outUpdate(config, cache.inGet(config))
-  val new_time = (e::time).take(k)
-  e match {
-    case Let(x, App(f, ae), e) =>
-      val closures = aeval(f, env, store).toList.asInstanceOf[List[Clos]]
-      val letvs = for (Clos(Lam(v, body), c_env) <- closures) yield {
-        val baddr = allocBind(v, new_time)
-        val new_env = c_env + (v -> baddr)
-        val new_store = store.update(baddr, aeval(ae, env, store))
-        val Ans(bdv, bdcache) = aeval(body, new_env, new_store, new_time, new_cache)
-        new_cache = bdcache
-        val evs = for (VS(vals, time, store) <- bdv) yield {
-          val baddr = allocBind(x, time)
-          val new_env = env + (x -> baddr)
-          val new_store = store.update(baddr, vals)
-          val Ans(ev, ecache) = aeval(e, new_env, new_store, time, new_cache)
-          new_cache = ecache
-          ev
-        }
-        evs.foldLeft(Set[VS]())(_ ++ _)
-      }
-      val letv = letvs.foldLeft(Set[VS]())(_ ++ _)
-      Ans(letv, new_cache.outUpdate(config, letv))
-
-    case ae if isAtomic(ae) =>
-      val vs = Set(VS(aeval(ae, env, store), new_time, store))
-      Ans(vs, new_cache.outUpdate(config, vs))
-  }
-}
+As we identified before,
+
+\begin{lstlisting}
 \end{lstlisting}

### 2018-03-17 paper.tex
Finished section 2

@@
 
-In the previous section, we show that by keeping an unabstracted stack in the state
-space, we can recover the precise call/return match.
-In this and following sections, we begin describing the transformations step by step.
-Our base machine for now is the AAM with unbounded stack, though
-as we will later transform the stack to higher-order functions
-represent \note{``representing'' or ``represented by''?} continuations, it does not matter what kind of AAM we start from.
-This is because transforming to higher-order functions forces us to sequentialize
-the order of abstract evaluation, and neither versions of AAM requires construction of a frame of lists \note{should this be frame of lists? sounds weird}, nor to allocate continuations in the store.
-Thus, our choice to start from an AAM with unbounded stack is motivated simply because it has an equivalent stack model to abstract definitional interpreters (our final target).
+In the last section, we showed that by keeping an unabstracted stack in the state
+space we can recover the precise call/return match.
+From this section, we start describing the transformations step by step.
+Our base machine is the AAM with unbounded stack.
+In fact, as later we will transform the stack to higher-order functions
+represented continuations, it does not matter what kind of AAM we start from.
+Because transforming to higher-order functions forces us to sequentialize
+the order of abstract evaluation, and neither need to construct a frame of list
+nor to allocate continuations in the store.
+We choose to start from AAM with unbounded stack just because it has a
+equivalence of stack model to the final abstract definitional interpreters.
 
@@
 
-As we identified before,
-
-\begin{lstlisting}
+On another path, here, we introduce direct-style abstracting definitional interpreter
+with delimited control operators @reset@ and @shift@.
+
+TODO: This version does not use delimited controls.
+
+\begin{lstlisting}
+def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans = {
+  val config = Config(e, env, store, time)
+  if (cache.outContains(config)) {
+    return Ans(cache.outGet(config), cache)
+  }
+
+  var new_cache = cache.outUpdate(config, cache.inGet(config))
+  val new_time = (e::time).take(k)
+  e match {
+    case Let(x, App(f, ae), e) =>
+      val closures = aeval(f, env, store).toList.asInstanceOf[List[Clos]]
+      val letvs = for (Clos(Lam(v, body), c_env) <- closures) yield {
+        val baddr = allocBind(v, new_time)
+        val new_env = c_env + (v -> baddr)
+        val new_store = store.update(baddr, aeval(ae, env, store))
+        val Ans(bdv, bdcache) = aeval(body, new_env, new_store, new_time, new_cache)
+        new_cache = bdcache
+        val evs = for (VS(vals, time, store) <- bdv) yield {
+          val baddr = allocBind(x, time)
+          val new_env = env + (x -> baddr)
+          val new_store = store.update(baddr, vals)
+          val Ans(ev, ecache) = aeval(e, new_env, new_store, time, new_cache)
+          new_cache = ecache
+          ev
+        }
+        evs.foldLeft(Set[VS]())(_ ++ _)
+      }
+      val letv = letvs.foldLeft(Set[VS]())(_ ++ _)
+      Ans(letv, new_cache.outUpdate(config, letv))
+
+    case ae if isAtomic(ae) =>
+      val vs = Set(VS(aeval(ae, env, store), new_time, store))
+      Ans(vs, new_cache.outUpdate(config, vs))
+  }
+}
 \end{lstlisting}

### 2018-03-17 paper.tex
update intro

--- 2018-03-17 paper.tex
@@
   Yet, the relations between small-step abstracting abstract machines and
-  big-step abstracting definitional interpreters is not well studied.
+  big-step abstracting definitional interpreters are not well studied.
 
@@
 
-Abstracting abstract machines (AAM) is a systematic methodology proposed by~\citeauthor{van2012systematic}
-for constructing abstract interpreters of higher-order functional languages~\cite{van2012systematic, van2010abstracting}.
+Defining a language by building an interpreter for it can be traced to the very
+early time of programming languages research \cite{landin1966next, reynolds1972definitional}.
+Nowadays, even an undergraduate student in computer science is able to build toy languages
+through interpreters.
+But building a sound abstract interpreter remains as an esoteric and difficult task until recent years.
+
+\citeauthor{van2012systematic} proposed Abstracting Abstract Machines (AAM) methodology which can used
+for constructing abstract interpreters of higher-order functional languages from concrete abstract machines
+~\cite{van2012systematic, van2010abstracting}.
 Given a concrete small-step abstract machine, (e.g, the CESK machine, Krivine's machine, etc.),
@@
 but we can recover different semantics, including the concrete semantics and various
-abstract semantics such as context-sensitivity and abstract garbage collection.
-
-Broadly speaking, abstract abstract machines and abstract definitional interpreters are both
-abstract interpreters. They are obtained by applying a combination of abstractions to their
+abstract semantics such as context-sensitivity and abstract garbage collection \cite{Sergey:2013:MAI:2491956.2491979}.
+
+Broadly speaking, abstract abstract machines and abstract definitional 
+interpreters are both some kind of abstract interpreters.
+They are obtained by applying a combination of abstractions to their
 concrete counterparts, abstract machines and definitional interpreters, respectively.
+An interesting question would be how we can interderive these abstract semantics artifacts 
+from the other one.
+
 In the concrete world, the relations among reduction semantics, abstract machines,
@@
 \cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}.
-These abstract machines implement structural operational semantics in continuation style,
+The abstract machines implement structural operational semantics in continuation style,
 where the reduction contexts are defunctionalized continuations. One can derive definitional
@@
 
-But in terms of their abstract semantics artifacts,
+On the contrary, in terms of their abstract semantics artifacts,
 the relations between small-step abstract abstract machines and big-step abstract
-definitional interpreters, as well as the ability to derive one from the other is not
+definitional interpreters, as well as the question of deriving one from the other is not
 well-studied.

### 2018-03-17 paper.tex
update

@@
 %\title{Pushdown Control Flow Analysis via Refunctionalization (Functional Pearl)}         %% [Short Title] is optional;
-\title{Refunctionalization of Abstracting Abstract Machines (Functional Pearl)}         %% [Short Title] is optional;
+\title{Refunctionalization of Abstract Abstract Machines (Functional Pearl)}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
 
-\subsection{Outline}
-
-The rest of this paper is structured as follows: we begin by reviewing some background
+\subsection{Contribution}
+
+We begin by reviewing some background
 necessary for this work in Section~\ref{background}, as well as introducing some of the basic
@@
 
-\todo{Add section numbers to the figure}\note{TR: put this into a proper figure, add known danvy sequence
+\note{TR: put this into a proper figure, add known danvy sequence
 for concrete evaluators at the bottom (in grey or dashed lines)}
-\note{JD: should these be abstracting abstract machines?}
 
@@
 and simplifies analysis~\cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
-Here, we choose to use $\lambda$-calculus in administrative normal form (ANF),
+Here, we choose to use $\lambda$-calculus in administrative normal form (ANF) \cite{flanagan1993essence},
 which is a direct-style intermediate representation, as our target for clarity,
 but without losing simplicity and generality.
+The transformations we will show in the rest of this paper
+also work on abstract machines for plain direct-style $\lambda$-calculus languages.
 Although we only show the core calculus language, it can be easily extended
@@
 so we elide them in this paper.
-The transformations we will show in the rest of this paper
-also work on abstract machines for plain direct-style $\lambda$-calculus languages.
 
 To begin, we present the concrete syntax of a call-by-value $\lambda$-calculus language
-in ANF~\cite{flanagan1993essence}.
+in ANF.
 

### 2018-03-17 paper.tex
update

@@
 
-We note that the \note{store? update function? can a store be monotonic?} is monotonic; it continuously grows and never shrinks. This property
-guarantees that in analyzing the \note{update function? store? pick the right word}, a fixed point can always be reached.
+We note that both the value store and continuation store are monotonic; 
+it continuously grows and never shrinks. This property guarantees that 
+a fixed point of store can always be reached through Kleene's 
+iteration when analyzing the program
 
@@
 
-The codomain of all binding stores @Storable@ is the same
-as previously defined for CESK machines.
+The codomain of binding stores @Storable@ is the same as previously defined for CESK machines.
 The codomain of continuation stores @Cont@, on the other hand, is comprised of
@@
 representing the remaining stack frames. 
-\todo{talk about nondet of cont deref?}
+But since the continuation store may contain multiple continuations,
+the dereferencing of continuation addresses is also nondeterministic.
 
@@
 Note that regarding the ambiguity in $k$-CFA\cite{Gilray:2016:ACP:2951913.2951936},
-the code here actually implements call+return \note{call+return, call/return, etc.? Check this} sensitivity.
+the code here actually implements call+return sensitivity.
 
@@
 The following code shows a variant of the worklist algorithm to find the fixed-point
-of the set of states.\note{JD: I changed all of this to worklist instead of work list in favor of the worklist algorithm. Change this back if it's wrong.}
+of the set of states.
 Function @drive@ always applies function @step@ to the head element @hd@ of

### 2018-03-17 paper.tex
Finished Linearization

@@
 In Danvy's paper \textit{Defunctionalized Interpreters for Programming
-  Languages}, he mentioned for deterministic languages, a reduction semantics
+  Languages}, he mentions that for deterministic languages, a reduction semantics
 is a structural operational semantics in continuation style, where the reduction
 context is a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}.
-Here we are facing a problem:
-the underlying semantics of AAM is fundamentally nondeterministic.
-But the evaluation context (the @Frame@ in our program) is not
-nondeterministic, and the work list @todo@ actually implicitly handles the nondeterminism,
-if we simply refunctionalize the frames to functions, it would not helps toward
+This poses a problem, as the underlying semantics of AAM is fundamentally nondeterministic.
+But the evaluation context (i.e., the @Frame@ in our program) is \emph{not}
+nondeterministic, and the worklist @todo@ actually implicitly handles the nondeterminism.
+Thus, if we simply refunctionalize the frames to functions, it does not help us move toward
 abstract definitional interpreters.
@@
 In order to address this problem, our first step of transformations
-is to linearize all the nondeterministic choices.
-This step turns the @step@ function to be deterministic.
-Likewise the @Frame@ saves the information of the caller in concrete executions,
-we introduce another layer of controls, and define a case class @NDCont@
+is to linearize all nondeterministic choices.
+This step removes all nondeterminism from the @step@ function.
+Likewise, the @Frame@ saves the information of the caller in concrete executions.
+We then introduce another layer of controls, and define a case class @NDCont@
 that saves the information at a fork point when we have multiple target closures.
-We also add a new field @ndk@ to the definition of state which is called
-@NDState@ now.
-For the rest of the paper, to differentiate the two continuations, we would call
-the first one as normal continuation, and the second one as nondeterministic
-continuations.
+We also add a new field @ndk@ to the definition of state, which we now call @NDState@.
+For clarity of presentation in differentiating between the two continuations, we elect to call
+the first a \emph{normal continuation}, and the second as a \emph{nondeterministic
+continuation}.
 
@@
 
-A @NDCont@ contains a list of closures that are possible functions to be invoked,
+Each @NDCont@ object contains a list of closures that are possible functions to be invoked,
 a set of values that will be bind to the function's formal argument,
@@
 
-Now the type of function @step@ becomes @NDState => NDState@.
+Function @step@ now becomes of type @NDState => NDState@.
 The following code shows the first case of matching an instance of @NDState@ in
-the @step@.
+the @step@ function:
 
@@
 
-By atomically evaluating @f@, we obtain a set of closures.
-The transition is deterministic to the first element of the closure set.
-So we just prepare new environment, new store, and new frame list only for the
+By atomically evaluating @f@, we obtain a set of closures, with the transition being deterministic in regards to the first element of the closure set.
+As such, we prepare a new environment, a new store, and a new frame list only for the
 first closure in that set.
@@
 which contains the rest of closures, the values of argument @ae@, the store,
-the time and the new frame list.
-We use the store before updating because for different closures they may form
+the time, and the new frame list.
+We use the store before updating, because for different closures they may form
 different binding addresses @baddr@.
-We also use the new frames, because all the closures at this fork point share
+We also use the new frames, because all closures at this fork point share
 the same stack and return point.
 
-In the second case shown in below we will see how @NDCont@ deal with
-the nondeterminism.
+In the second case (shown below), we will see how @NDCont@ deals with
+nondeterminism.
 
@@
 If the normal continuation @konts@ is an empty list, then it means we have reached
-the halt of computation; otherwise, we should return the values of @ae@ to its
-caller which is contained in the top frame of stack.
-
-But, since we have added nondeterminism into state, an empty list of frames means
+the end of the computation (i.e., a halt); otherwise, we should return the values of @ae@ to its caller which is contained in the top frame of stack.
+
+However, since we have added nondeterminism into state, an empty list of frames means
 we have the reached the halt of \textit{one computation path}, and we should
-determine the next state indicates by @NDCont@.
-So we have a pattern matching on @ndk@:
+determine the next state indicated by @NDCont@.
+Thus, we have a pattern matching on @ndk@:
 \begin{itemize}
   \item If the closure set is an empty set,
-then we had tried all the closures of this fork point and should move the the
-next fork point. Since the @NDCont@ is represented by a list and we always
-append new element from its front, we are resuming to the most recent fork point by
+then we have tried all the closures of this fork point and should move to the
+next fork point. Since the @NDCont@ is represented by a list, and we always
+append new elements to its front, we are resuming to the most recent fork point by
 popping up the front-most element.
-\item Otherwise, we pop a closure from the set, then build new
-environment and new store for the body expression of the closure;
+  \item Otherwise, we pop a closure from the set, then build a new
+environment and a new store for the body expression of the closure;
 we use the same frame list and time that are copied from the fork point rather
 than current one.
-The nondeterminism continuation @new_ndk@ is also updated by removal of that closure.
+The nondeterministic continuation @new_ndk@ is also updated by the removal of that closure.
 \end{itemize}
@@
 
-The function @drive@ is also changed: there is no work list anymore, because all the
-potentially unexplored states are embedded in to the continuation for
+The @drive@ function is also changed: there is no worklist anymore, as all the
+potentially unexplored states are embedded into the continuation for
 nondeterminism.
-The termination of the analysis is when we reach a @NDState@ that the expression
-is atomic, and both the normal continuation and nondeterminism continuation are
-empty list. This corresponds to the case that @todo@ list is empty.
-
-Now, we obtain a linearized abstract abstract machine. If we image that the classical
-AAM explores a graph of reachable states, then the linearized AAM just flattens the
+The termination of the analysis occurs when we reach an @NDState@ in which the expression
+is atomic and both the normal continuation and nondeterminism continuation are
+empty lists. This corresponds to the case that @todo@ list is empty.
+
+At this point, we have obtain a linearized abstract abstract machine. If we imagine that the
+classical AAM explores a graph of reachable states, then the linearized AAM flattens the
 graph to a linear sequence.

### 2018-03-17 paper.tex
section 7

@@
 
-Fusing is a transformation that combines function @step@ and @drive@ to one.
-The fused function @drive_step@ basically formed by merging function
-@step@ into function @drive@.
-@drive_step@ takes a @NDState@ and a set of explored states as argument,
-and returns a set of reachable states once terminates.
+Fusing is a transformation that combines the @step@ and @drive@ functions into a single
+function.
+The fused function @drive_step@ is essentially formed by merging the functionality of
+@step@ into the @drive@ function.
+@drive_step@ takes an @NDState@ and a set of explored states as an argument
+and returns a set of reachable states once it terminates.
+\note{Do we really need this section?
+It seems unlikely that this requires its own section.
+Could we just append to the last section: For convenience, we combine these functions into one, called \texttt{drivestep}?}
 
@@
 
-Now we have a single function to perform both abstract evaluation and intermediate
-states collecting. Given @inject(e)@ as the initial state and an empty set as
-initial reached set, we can easily define the entrance function @analyze@ as follow.
+With this, have a single function to perform both abstract evaluation and the collection of
+intermediate states. Given @inject(e)@ as an initial state and an empty set as the
+initial set of states reached, we can easily define the entrance function @analyze@ as
+follows:
 
@@
 
-In the last section, the fused AAM combines all the functions into one and
-looks like a ``big-step'' interpreter, although it still has the machine
+With fusing complete, our AAM
+appears similar to a ``big-step'' interpreter, although it still has the machine
 state representation inside.
 In the disentangling transformation, we
-identify the first-order data type represented evaluation contexts,
-and lift the code blocks that handles these evaluation contexts to be individual functions.
-
-Since there are two layers of continuations, as a result,
-we obtain three mutually recursive functions, @drive_step@,
-@continue@ and @ndcontinue@. The function @drive_step@ plays the same
-role as before; @continue@ handles the evaluation context (@Frame@) of the analyzed
-language, which the function is called from @drive_step@ when encountering an atomic
-expression; @ndcontinue@ dispatches nondeterministic continuations @NDCont@,
-and is invoked from @continue@ when the normal stack is empty.
+identify the first-order data types which represent evaluation contexts
+and lift the code blocks that handle these evaluation contexts to be individual functions.
+
+Since there are two layers of continuations,
+we obtain three mutually recursive functions: @drive_step@, which plays the same
+role as before; @continue@, which is called from @drive_step@ when encountering an atomic expression, and handles the evaluation context (@Frame@) of the analyzed
+language; and @ndcontinue@, which dispatches nondeterministic continuations @NDCont@,
+and which is invoked from @continue@ when the normal stack is empty.
 
@@
 
-The above code shows the skeleton of function @drive_step@.
-In the first case of pattern matching, the code for constructing new
-environment, store and frames is omitted.
-At the end of this case, we invoke @drive_step@ recursively.
-The code for the second case is replaced by a function call to @continue@.
-Now a vague shape of interpreter in continuation style is emerged.
+The above code shows the skeleton of @drive_step@.
+At the end of the first case of pattern matching, we invoke @drive_step@ recursively\footnote{The code for constructing new environments, stores, and frames is elided for simplicity of presentation.}.
+The code for the second case is replaced entirely by a function call to @continue@.
+The astute reader will notice a vague shape of interpreter in continuation style has emerged.
 
@@
 
-The function @continue@ is largely what we have for the second case in
-@drive_step@, except a function call to @ndcontinue@ is placed when
-the the normal continuation is empty.
+@continue@ is mainly what we have seen for the second case in
+@drive_step@, except that @ndcontinue@ is called when
+the normal continuation is empty.
 Since we are analyzing a language in ANF, the first-order data representation of
-continuation is simplified to @List@.
-An empty list @Nil@ represents the halt of a computation path.
-A cons @::@ means the top frame of the list holds the binding variable, and
+continuations is simplified to a Scala @List@.
+An empty list @Nil@ represents the halt of a computation path, whereas
+a cons @::@ means the top frame of the list holds the binding variable and
 evaluation context.
-An analogue to this in standard CEK machines would be an abstract class (e.g., @cont@) with three
-variants @halt@, @arg(value, cont)@ and @arg(exp, env, cont)@.
+An analogy to this in standard CEK machines would be an abstract class (e.g., @cont@) with
+three variants: @halt@, @arg(value, cont)@, and @arg(exp, env, cont)@.
 
@@
 
-The shape for function @ndcontinue@ is similar to function @continue@, after
-all, they both use @List@ as a representation of continuations.
+The shape for @ndcontinue@ is similar to @continue@; after
+all, they both use a @List@ representation for continuations.
 When the closure set is empty, we recursively call @drive_step@ with the rest
@@
 
-Refunctionalization transforms first-order data type represented
+Refunctionalization transforms first-order data types represented as \note{``represented as'' or ``representing?'' It should be representing, right?}
 evaluation contexts to higher-order functions.
@@
 this transformation can be also regarded as allocating continuation functions of
-the defined language in the heap of the meta language.
+the defined language in the heap of the metalanguage.
 
@@
 function.
-\item Since our final target is abstract definitional interpreters while still
-nondeterministic, start from this section, our abstract semantic artifact
+\item Since our final target is an abstract definitional interpreter (though still
+nondeterministic), starting from this section our abstract semantic artifact
 returns a set of final values instead of collected states.
-\item As we are getting close to the
+\item As we are getting close to achieving
 abstract definitional interpreters, the name of @drive_step@ is updated to
@@
 
-For clarity, in this section, we first show how to refunctionalize the normal continuation and
+For clarity, we first show how to refunctionalize normal continuations and
 nondeterministic continuation without collecting semantics.
-Then adding an appropriate caching algorithm to make sure the analysis will
-terminate at some fixed-point is just changing the program to cache-passing style.
-Finally, we examine how computable pushdown control flow analysis is established
+Then, adding an appropriate caching algorithm to make sure the analysis will
+terminate at some fixed-point is a simple transformation of the program to cache-passing
+style. Finally, we examine how computable pushdown control flow analysis is established
 through refunctionalization and caching.
@@
 
-To properly represented final values, we introduce the case class @VS@ which
-contains a set of storable values, a time stamp, and a store.
+To properly represent final values, we introduce the case class @VS@ which
+contains a set of storable values, a timestamp, and a store.
 \begin{lstlisting}
@@
 
-We will apply the fixed-point caching algorithm as described in
-\citeauthor{darais2017abstracting}'s abstract definitional interpreter.
-
+Based on the refunctionalized AAM from the last section, we use the same @nd@ nondeterministic operator
+but further extend it to a fixed-point cached evaluation in this section. \todo{extend this}
+
+\textbf{Configuration.}
+Now we do not have the concept of explicit stack and state,
+but to represent the cache, we introduce the configuration @Config@ as 
+a state-like definition which borrows the components from state but without 
+continuations.
 \begin{lstlisting}
 case class Config(e: Expr, env: Env, store: BStore, time: Time)
+\end{lstlisting}
+
+\textbf{Cache.}
+We will apply the fixed-point caching algorithm as described from
+\citeauthor{darais2017abstracting}'s ADI paper.
+Note that the latent assumption of the worklist algorithm in classical AAM
+is if we have seen a state @s@, then it means we also have seen the successors
+of state @s@. 
+The caching algorithm we adopted here replay the same assumption but in a big-step
+manner: if we have seen the configuration @c@, then it means we also have seen
+the values that are evaluated by the configuration @c@.
+
+\begin{lstlisting}
 case class Cache(in: Store[Config, VS], out: Store[Config, VS]) {
@@
 
-adding cache to eval
-
+\textbf{Cache-Passing Style.}
 The definition of Ans is changed to a case class that contains a set of @VS@ and
@@
 \begin{lstlisting}
-def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, cont: Cont): Ans = {
+def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, continue: Cont): Ans = {
   val config = Config(e, env, store, time)
-  if (cache.outContains(config)) return cont(Ans(cache.outGet(config), cache))
+  if (cache.outContains(config)) return continue(Ans(cache.outGet(config), cache))
 
@@
       },
-      { case Ans(vss, cache) => cont(Ans(vss, cache.outUpdate(config, vss))) })
+      { case Ans(vss, cache) => continue(Ans(vss, cache.outUpdate(config, vss))) })
 
@@
       val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
-      cont(Ans(vs, new_cache.outUpdate(config, vs)))
-  }
-}
+      continue(Ans(vs, new_cache.outUpdate(config, vs)))
+  }
+}
+\end{lstlisting}
+
+\begin{lstlisting}
 def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore) = {

### 2018-03-17 paper.tex
Finished disentangling

--- 2018-03-17 paper.tex
@@
 
-The function @continue@ is largely what we have for the second case in
-@drive_step@, except a function call to @ndcontinue@ is placed when
-the the normal continuation is empty.
+@continue@ is largely what we have for the second case in
+@drive_step@, except that @ndcontinue@ is called when
+the normal continuation is empty.
 Since we are analyzing a language in ANF, the first-order data representation of
-continuation is simplified to @List@.
-An empty list @Nil@ represents the halt of a computation path.
-A cons @::@ means the top frame of the list holds the binding variable, and
+continuations is simplified to a Scala @List@.
+An empty list @Nil@ represents the halt of a computation path, whereas
+a cons @::@ means the top frame of the list holds the binding variable and
 evaluation context.
-An analogue to this in standard CEK machines would be an abstract class (e.g., @cont@) with three
-variants @halt@, @arg(value, cont)@ and @arg(exp, env, cont)@.
+An analogy to this in standard CEK machines would be an abstract class (e.g., @cont@) with
+three variants: @halt@, @arg(value, cont)@, and @arg(exp, env, cont)@.
 

### 2018-03-17 paper.tex
Typo
update

--- 2018-03-17 paper.tex
@@
 
+The case class @Cache@ is defined as follow:
+
 \begin{lstlisting}
@@
 
+The @Cache@ contains two maps @in@ and @out@ which are both mapping from 
+configurations to sets of @VS@.
+
+When we enter to function @aeval@
+
 \textbf{Cache-Passing Style.}
+
 The definition of Ans is changed to a case class that contains a set of @VS@ and
@@
 analysis through refunctionalization and caching. In this section,
-we examine how pushdown control-flow is obtained and what we have done to
+we revisit the pushdown control-flow problem and examine what we have done to
 achieve it.

### 2018-03-17 paper.tex
Working on refunc

--- 2018-03-17 paper.tex
@@
 
-@continue@ is largely what we have for the second case in
+@continue@ is mainly what we have seen for the second case in
 @drive_step@, except that @ndcontinue@ is called when

### 2018-03-17 paper.tex
section 6

--- 2018-03-17 paper.tex
@@
 
-\textbf{First Version of Evaluation Function.}
+\textbf{Evaluation without Caching.}
 Since our first version of @aeval@ just implements the nondeterministic
@@
 
-The case class @Cache@ is defined as follow:
+The case class @Cache@ and its operations are defined as follow:
 
@@
 The @Cache@ contains two maps @in@ and @out@ which are both mapping from 
-configurations to sets of @VS@.
-
-When we enter to function @aeval@
+configurations to sets of @VS@. The @in@ cache contains mappings from 
+last iteration of evaluation, and @out@ cache contains mappings after this
+iteration of evaluation. Once we have evaluated a term to some values, 
+we update the @out@ cache. In the next time of iteration, we will use @out@
+from last iteration as @in@ cache for this iteration.
 
 \textbf{Cache-Passing Style.}
-
-The definition of Ans is changed to a case class that contains a set of @VS@ and
+Now we can modify out interpreter to cache-passing style. 
+First the return type @Ans@ is changed to a case class that contains a set of @VS@ and
 a cache.
+
 \begin{lstlisting}
@@
 \end{lstlisting}
+
+The skeleton of function @aeval@ largely remain the same. 
+Every place has a value of @Ans@ is also replaced as direct pattern matching so 
+that we can use the fields in @Ans@ immediately.
+The cache is used in a monotonic way: each function call to @aeval@ or @nd@ is
+passed with the latest cache from the most recent continuation's result.
+
+When we enter to function @aeval@, we firstly lookup in the @out@ cache, 
+if there are some values for current configuration, then we use immediately.
+This corresponds to small-step AAM with a worklist when we have seen a state 
+then we can simply discard the state and continue working the rest of worklist @todo@.
+
+Otherwise, if we not hit the @out@ cache, we retrieve the values from @in@ cache, and update @out@ cache
+with values from @in@ cache.  The retrieving from @in@ cache may return an empty set of values if 
+there is no such mapping for the configuration from the previous iteration.
+It actually first conservatively assumes that all computation can be diverged,
+if not, we will update its values in the mapping after evaluation.
+In terms of partial orders and lattices, it is also the case we start 
+the Kleene iteration from the bottom of the lattice.
 
@@
   val config = Config(e, env, store, time)
+  /* lookups the out cache, uses the values if contains this config */
   if (cache.outContains(config)) return continue(Ans(cache.outGet(config), cache))
-
+  /* uses the values from in cache */
   val new_cache = cache.outUpdate(config, cache.inGet(config))
   val new_time = (e::time).take(k)
-
   e match {
@@
       nd[Clos, Ans](closures, Ans(Set[VS](), new_cache), { case (clos, Ans(acc, cache), closnd) =>
-        val Clos(Lam(v, body), c_env) = clos
-        val baddr = allocBind(v, new_time)
-        val new_env = c_env + (v -> baddr)
-        val new_store = store.update(baddr, atomicEval(ae, env, store))
+        ......
         aeval(body, new_env, new_store, new_time, cache, { case Ans(bodyvss, bodycache) =>
           nd[VS, Ans](bodyvss, Ans(Set[VS](), bodycache), { case (vs, Ans(acc_vss, cache), bdnd) =>
-            val VS(vals, time, store) = vs
-            val baddr = allocBind(x, time)
-            val new_env = env + (x -> baddr)
-            val new_store = store.update(baddr, vals)
-            aeval(e, new_env, new_store, time, cache, { case Ans(evss, ecache) => bdnd(Ans(acc_vss ++ evss, ecache)) })
+            ......
+            aeval(e, new_env, new_store, time, cache, { case Ans(evss, ecache) => 
+              bdnd(Ans(acc_vss ++ evss, ecache)) 
+            })
           },
@@
       },
+      /* updates the out cache after evaluation */
       { case Ans(vss, cache) => continue(Ans(vss, cache.outUpdate(config, vss))) })
@@
       val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
-      continue(Ans(vs, new_cache.outUpdate(config, vs)))
-  }
-}
-\end{lstlisting}
+      continue(Ans(vs, new_cache.outUpdate(config, vs))) /* updates the out cache after evaluation */
+  }
+}
+\end{lstlisting}
+
+After done the evaluation, we obtain a set of @VS@ and need to update the @out@ cache.
+In the first case of pattern matching, it happens inside of the last continuation of 
+outer @nd@. 
+For the second case, we update the @out@ cache before calling the continuation @continue@.
 

### 2018-03-17 paper.tex
update code

@@
       val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
-      nd[Clos, Ans](closures, Ans(Set[VS](), new_cache), { case (clos, Ans(acc, cache), closnd) =>
+      nd[Clos, Ans](closures, Ans(Set[VS](), new_cache), { case (clos, Ans(accouter, cache), closnd) =>
         ......
@@
             aeval(e, new_env, new_store, time, cache, { case Ans(evss, ecache) => 
-              bdnd(Ans(acc_vss ++ evss, ecache)) 
+              bdnd(Ans(accinner ++ evss, ecache)) 
             })
           },
-          { case Ans(evss, cache) => closnd(Ans(evss ++ acc, cache)) })
+          { case Ans(evss, cache) => closnd(Ans(evss ++ accouter, cache)) })
         })

### 2018-03-17 paper.tex
Adding figure

@@
 
-\note{TR: put this into a proper figure, add known danvy sequence
-for concrete evaluators at the bottom (in grey or dashed lines)}
-
-\begin{tikzpicture}
-  \matrix (m) [matrix of math nodes, row sep=2.5em, column sep=3.7em]
-    {
-        &  &  &  &
-      \begin{smallmatrix} \text{abstract} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
-      \begin{smallmatrix} \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
-      \begin{smallmatrix} \text{big-step} \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  &
-      \begin{smallmatrix} \text{disentangled}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
-      \begin{smallmatrix} \text{refunctionalized}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
-      \begin{smallmatrix} \text{monadic} \\ \text{abstracting} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
-      \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} &
-        &
-      %\begin{smallmatrix} \text{big-step} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  &
-        &
-      %\begin{smallmatrix} \text{disentangled} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
-        &
-      %\begin{smallmatrix} \text{refunctionalized} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
-      \begin{smallmatrix} \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
-    };
-  {
-    [start chain] \chainin (m-2-4);
-    \chainin (m-1-5) [join={node[sloped,anchor=center,above,labeled] {\text{un-CPS}}}];
-    \chainin (m-2-5) [join={node[above,labeled] {\text{}}}];
-    \chainin (m-1-5) [join={node[above,labeled] {\text{}}}];
-  }
-  {
-    [start chain] \chainin (m-2-1);
-    \chainin (m-2-2) [join={node[above,labeled] {\begin{smallmatrix} \text{linearizing} \\ \text{+} \\ \text{fusing} \end{smallmatrix}}     }];
-    \chainin (m-2-3) [join={node[above,labeled] {\text{disentangling}}}];
-    \chainin (m-2-4) [join={node[above,labeled] {\text{refunc.}}}];
-    \chainin (m-2-5) [join={node[above,labeled] {\text{monadify}}}];
-  }
-  {
-    [start chain]
-    \chainin (m-3-1);
-    { [start branch=A] \chainin (m-2-1)
-      [join={node[left, labeled] {\alpha}}];
-    }
-  }
-    %\chainin (m-2-2);
-    %\chainin (m-2-3);
-    %\chainin (m-2-4);
-  {
-    [start chain]
-    \chainin (m-3-5);
-    { [start branch=B] \chainin (m-2-5)
-      [join={node[right, labeled] {\alpha}}];
-    }
-  }
-\end{tikzpicture}
+% \note{TR: put this into a proper figure, add known danvy sequence
+% for concrete evaluators at the bottom (in grey or dashed lines)}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=100mm,scale=1.0]{figures/lines.png}
+\end{center}
+\end{figure}
+
+% \begin{tikzpicture}
+%   \matrix (m) [matrix of math nodes, row sep=2.5em, column sep=3.7em]
+%     {
+%         &  &  &  &
+%       \begin{smallmatrix} \text{abstract} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
+%       \begin{smallmatrix} \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
+%       \begin{smallmatrix} \text{big-step} \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  &
+%       \begin{smallmatrix} \text{disentangled}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
+%       \begin{smallmatrix} \text{refunctionalized}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
+%       \begin{smallmatrix} \text{monadic} \\ \text{abstracting} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
+%       \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} &
+%         &
+%       %\begin{smallmatrix} \text{big-step} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  &
+%         &
+%       %\begin{smallmatrix} \text{disentangled} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
+%         &
+%       %\begin{smallmatrix} \text{refunctionalized} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
+%       \begin{smallmatrix} \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
+%     };
+%   {
+%     [start chain] \chainin (m-2-4);
+%     \chainin (m-1-5) [join={node[sloped,anchor=center,above,labeled] {\text{un-CPS}}}];
+%     \chainin (m-2-5) [join={node[above,labeled] {\text{}}}];
+%     \chainin (m-1-5) [join={node[above,labeled] {\text{}}}];
+%   }
+%   {
+%     [start chain] \chainin (m-2-1);
+%     \chainin (m-2-2) [join={node[above,labeled] {\begin{smallmatrix} \text{linearizing} \\ \text{+} \\ \text{fusing} \end{smallmatrix}}     }];
+%     \chainin (m-2-3) [join={node[above,labeled] {\text{disentangling}}}];
+%     \chainin (m-2-4) [join={node[above,labeled] {\text{refunc.}}}];
+%     \chainin (m-2-5) [join={node[above,labeled] {\text{monadify}}}];
+%   }
+%   {
+%     [start chain]
+%     \chainin (m-3-1);
+%     { [start branch=A] \chainin (m-2-1)
+%       [join={node[left, labeled] {\alpha}}];
+%     }
+%   }
+%     %\chainin (m-2-2);
+%     %\chainin (m-2-3);
+%     %\chainin (m-2-4);
+%   {
+%     [start chain]
+%     \chainin (m-3-5);
+%     { [start branch=B] \chainin (m-2-5)
+%       [join={node[right, labeled] {\alpha}}];
+%     }
+%     {
+%       [start branch=D, color=gray] \chainin (m-3-1)
+%       [join={node[right, labeled] {\text{}}}];
+%     }
+%   }
+%   {
+%     [start chain, color=gray] \chainin (m-3-1);
+%     {
+%       [start branch=C] \chainin (m-3-5)
+%       [join={node[above, labeled] {\text{Known Danvy Sequence for Concrete Evaluators}}}];
+%     }
+%   }
+% \end{tikzpicture}
 

### 2018-03-17 paper.tex
update

@@
 \citeauthor{van2012systematic} proposed Abstracting Abstract Machines (AAM) methodology which can used
-for constructing abstract interpreters of higher-order functional languages from concrete abstract machines
+for constructing sound abstract interpreters of higher-order functional languages from concrete abstract machines
 ~\cite{van2012systematic, van2010abstracting}.
@@
 Based on the refunctionalized AAM from the last section, we use the same @nd@ nondeterministic operator
-but further extend it to a fixed-point cached evaluation in this section. \todo{extend this}
+but further extend it to a fixed-point cached evaluation in this section. 
+The fixed-point caching algorithm performs a sound over-approximation of all possible 
+concrete reachable pathes and values, and also prevents non-termination when analyzing diverged programs.
 
@@
 \textbf{Cache.}
-We will apply the fixed-point caching algorithm as described from
-\citeauthor{darais2017abstracting}'s ADI paper.
-Note that the latent assumption of the worklist algorithm in classical AAM
+Recall that the latent assumption of the worklist algorithm in classical AAM
 is if we have seen a state @s@, then it means we also have seen the successors
@@
 the values that are evaluated by the configuration @c@.
+Here we will apply the fixed-point caching algorithm as described from
+\citeauthor{darais2017abstracting}'s ADI paper \cite{darais2017abstracting}.
 
@@
 if not, we will update its values in the mapping after evaluation.
-In terms of partial orders and lattices, it is also the case we start 
+In terms of partial orders and lattices, it is the case that starts
 the Kleene iteration from the bottom of the lattice.
@@
 \begin{lstlisting}
-def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore) = {
+def analyze(e: Expr) = {
   def iter(cache: Cache): Ans = {
-    val Ans(vss, new_cache) = aeval(e, env, store, mtTime, cache, {
-      val initConfig = Config(e, env, store, mtTime)
+    val Ans(vss, new_cache) = aeval(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List(), cache, {
+      val initConfig = Config(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List())
       case Ans(vss, cache) => Ans(vss, cache.outUpdate(initConfig, vss))
@@
 \end{lstlisting}
+
+Finally, the @analyze@ function as the entrance of the analysis does a looping iteration
+to find the fixed-point on caches starting from an empty cache. 
+If the @out@ cache of this iteration is equivalent to the @out@ cache from the last iteration,
+then we have reached a fixed-point that over-approximates the concrete evaluations,
+and thus can be returned.
+Otherwise, a next iteration with the new @out@ cache will be activated.
+Note that the continuation of the initial call to @aeval@ sets up the values for the 
+initial configuration in the final cache.
 

### 2018-03-17 paper.tex
update fig

--- 2018-03-17 paper.tex
@@
 
-\subsection{Contribution}
-
-We begin by reviewing some background
-necessary for this work in Section~\ref{background}, as well as introducing some of the basic
-code structures used throughout the paper. We then address the main contribution of this
-paper, which is the filling of the gap between small-step AAM and big-step ADI by developing a
-series of systematic transformations. Those transformations are summarized here, with their
-associated section:
-
-\begin{itemize}
-\item We show \textbf{linearization} in Section~\ref{linear}. By expressing the
-  nondeterministic choices as a first-order data type, we linearize the execution
-  of abstract abstract machines; the transition of machine states therefore becomes deterministic. Notably, this introduces another layer of controls that will also be refunctionalized later.
-
-\item In Section~\ref{fusing}, we then present the \textbf{fusing} transformation.
-  The fusing transformation simply combines the single-step function @step@ and the driving
-  function into one, but keeps all the machine state representations.
-
-\item Section~\ref{disen} discusses \textbf{disentangling}, which disassembles the fused AAM to be several individual functions, with each function handling one data type represented by a continuation.
-
-\item \textbf{Refunctionalization} is shown in Section~\ref{refunc} which sequentializes
-  the order of abstract execution by higher-order functions.
-  For clarity, we first present the refunctionalized AAM without caching. We then adopt a different caching algorithm to guarantee the termination of abstract interpretation.
-  In this section, we also review pushdown control-flow analysis and examine how computable
-    and precise call/return matching is obtained through these transformations.
-
-\item The last transformation we show is that of transforming the refunctionalized AAM
-  to a \textbf{direct-style} interpreter (Section~\ref{directstyle}) by using delimited controls.
-
-\end{itemize}
-
-These transformations are used throughout the paper, with refunctionalization and
-defunctionalization of abstract interpreters playing important roles for the call stack of
-the analyzed language. By refunctionalization, the call stack of the analyzed language
-is blended into the call stack of the defining language. This provides another perspective to
-explain why \citeauthor{darais2017abstracting}'s abstract definitional interpreters
-is able to extend the the pushdown control-flow property from its defining language.
-
-We complete this paper by discussing related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.
-
-% \subsection{Contributions}
-
-% the essence of abstracting definitional interpreter.
-
-% We summarize our contributions as follows:
-
-% \begin{itemize}
-% \item The main contribution of this paper is the filling of the gap
-%   between small-step AAM and big-step ADI by developing a series of
-%   systematic transformations.
-%   We show that the correspondence between abstract
-%   machines and interpreters exists not only in concrete semantics artifacts,
-%   but also in abstract semantics artifacts.
-%   Galois connection?
-%   \note{JD: Section number?}
-
-% \item A new understanding of pushdown control flow analysis by
-%   refunctionalization.
-%   We present an abstract interpreter obtained from refunctionalizing
-%   a small-step abstracting abstract machine.
-%   This abstract interpreter naturally exhibits the pushdown control-flow property,
-%   as the refunctionalization forces us to sequentialize the
-%   abstract execution.
-%   We also analyze its correctness, soundness, and termination.
-
-% \item Further, we show that defunctionalization and refunctionalization
-%   of abstract interpreters play important roles for the call stack of the analyzed language.
-%   Given this insight, we reveal precisely \emph{why} small-step AAM has no pushdown
-%   control-flow property (without extra efforts), as well as explain
-%   why ADI naturally extends pushdown control-flow property from its
-%   defining language.
-
-% \end{itemize}
-
-% \subsection{Outline}
-
-% \note{TR: can this be merged with contributions? (they currently lack section refs)}
-% The remainder of this paper is organized as follows.
-% We begin by describing the target language we will analyze in Section~\ref{background},
-% including a concrete CESK machine as small-step operational semantics for it.
-% We continue with a review of abstracting abstract machines (AAM) as an abstract interpretation
-% counterpart of CESK machines. Following this, Section~\ref{linear} shows a conversion from the
-% nondeterministic choices of closures to another layer of continuations,
-% therefore linearizing the abstract execution.
-% Section~\ref{fusing} then discusses fusing the driver and step functions to a single
-% function.
-% We further identify in Section~\ref{disen} the first-order data type representations of
-% continuations, and disentangle the code blocks for dispatching data types to separate
-% functions.
-% Section~\ref{refunc} then represents the refunctionalized abstract abstract machines.
-% In this section, the continuations represented by data types become higher-order
-% functions.
-% In Section~\ref{directstyle}, we transform the refunctionalized AAM to two
-% different direct styles: one implemented with monads, the other using
-% delimited controls.
-
-% Finally, we discuss related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.
-
-\note{TR: put this into a proper figure, add known danvy sequence
-for concrete evaluators at the bottom (in grey or dashed lines)}
-
+\begin{figure}
+\begin{center}
  \begin{tikzpicture}
@@
      }
+     {
+       [start branch=D, color=gray] \chainin (m-3-1)
+       [join={node[right, labeled] {\text{}}}];
+     }
+   }
+   {
+     [start chain, color=gray] \chainin (m-3-1);
+     {
+       [start branch=C] \chainin (m-3-5)
+       [join={node[above, labeled] {\text{Known as Danvy's correspondence for concrete semantics artifacts }}}];
+     }
    }
  \end{tikzpicture}
+ \caption{Transformations from AAM to ADI} \label{fig:trans}
+\end{center}
+\end{figure}
+
+\subsection{Contribution}
+
+We begin by reviewing some background
+necessary for this work in Section~\ref{background}, as well as introducing some of the basic
+code structures used throughout the paper. We then address the main contribution of this
+paper, which is the filling of the gap between small-step AAM and big-step ADI by developing a
+series of systematic transformations. Figure \ref{fig:trans} shows the transformations.
+
+Those transformations are summarized here, with their associated section:
+
+\begin{itemize}
+\item We show \textbf{linearization} in Section~\ref{linear}. By expressing the
+  nondeterministic choices as a first-order data type, we linearize the execution
+  of abstract abstract machines; the transition of machine states therefore becomes deterministic. Notably, this introduces another layer of controls that will also be refunctionalized later.
+
+\item In Section~\ref{fusing}, we then present the \textbf{fusing} transformation.
+  The fusing transformation simply combines the single-step function @step@ and the driving
+  function into one, but keeps all the machine state representations.
+
+\item Section~\ref{disen} discusses \textbf{disentangling}, which disassembles the fused AAM to be several individual functions, with each function handling one data type represented by a continuation.
+
+\item \textbf{Refunctionalization} is shown in Section~\ref{refunc} which sequentializes
+  the order of abstract execution by higher-order functions.
+  For clarity, we first present the refunctionalized AAM without caching. We then adopt a different caching algorithm to guarantee the termination of abstract interpretation.
+  In this section, we also review pushdown control-flow analysis and examine how computable
+    and precise call/return matching is obtained through these transformations.
+
+\item The last transformation we show is that of transforming the refunctionalized AAM
+  to a \textbf{direct-style} interpreter (Section~\ref{directstyle}) by using delimited controls.
+
+\end{itemize}
+
+These transformations are used throughout the paper, with refunctionalization and
+defunctionalization of abstract interpreters playing important roles for the call stack of
+the analyzed language. By refunctionalization, the call stack of the analyzed language
+is blended into the call stack of the defining language. This provides another perspective to
+explain why \citeauthor{darais2017abstracting}'s abstract definitional interpreters
+is able to extend the the pushdown control-flow property from its defining language.
+
+We complete this paper by discussing related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.
+
+% \subsection{Contributions}
+
+% the essence of abstracting definitional interpreter.
+
+% We summarize our contributions as follows:
+
+% \begin{itemize}
+% \item The main contribution of this paper is the filling of the gap
+%   between small-step AAM and big-step ADI by developing a series of
+%   systematic transformations.
+%   We show that the correspondence between abstract
+%   machines and interpreters exists not only in concrete semantics artifacts,
+%   but also in abstract semantics artifacts.
+%   Galois connection?
+%   \note{JD: Section number?}
+
+% \item A new understanding of pushdown control flow analysis by
+%   refunctionalization.
+%   We present an abstract interpreter obtained from refunctionalizing
+%   a small-step abstracting abstract machine.
+%   This abstract interpreter naturally exhibits the pushdown control-flow property,
+%   as the refunctionalization forces us to sequentialize the
+%   abstract execution.
+%   We also analyze its correctness, soundness, and termination.
+
+% \item Further, we show that defunctionalization and refunctionalization
+%   of abstract interpreters play important roles for the call stack of the analyzed language.
+%   Given this insight, we reveal precisely \emph{why} small-step AAM has no pushdown
+%   control-flow property (without extra efforts), as well as explain
+%   why ADI naturally extends pushdown control-flow property from its
+%   defining language.
+
+% \end{itemize}
+
+% \subsection{Outline}
+
+% \note{TR: can this be merged with contributions? (they currently lack section refs)}
+% The remainder of this paper is organized as follows.
+% We begin by describing the target language we will analyze in Section~\ref{background},
+% including a concrete CESK machine as small-step operational semantics for it.
+% We continue with a review of abstracting abstract machines (AAM) as an abstract interpretation
+% counterpart of CESK machines. Following this, Section~\ref{linear} shows a conversion from the
+% nondeterministic choices of closures to another layer of continuations,
+% therefore linearizing the abstract execution.
+% Section~\ref{fusing} then discusses fusing the driver and step functions to a single
+% function.
+% We further identify in Section~\ref{disen} the first-order data type representations of
+% continuations, and disentangle the code blocks for dispatching data types to separate
+% functions.
+% Section~\ref{refunc} then represents the refunctionalized abstract abstract machines.
+% In this section, the continuations represented by data types become higher-order
+% functions.
+% In Section~\ref{directstyle}, we transform the refunctionalized AAM to two
+% different direct styles: one implemented with monads, the other using
+% delimited controls.
+
+% Finally, we discuss related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.
+
+% \note{TR: put this into a proper figure, add known danvy sequence
+% for concrete evaluators at the bottom (in grey or dashed lines)}
+
+%\includegraphics[width=100mm,scale=1.0]{figures/lines.png}
 

### 2018-03-17 paper.tex
Working on 6

@@
 the nondeterministic evaluation.
-The reason that we include a store is there might be a case two pathes have the
-same values but different accumulated side effects such as memory allocations.
-Since the whole computation is nondeterministic, the type of final result
-values is a collection @Set[VS]@. For convenience, we define it as answer type @Ans@:
+The reason that we include a store is that there might be a case in which two paths have the same values but different accumulated side effects (e.g., memory allocations).
+Since the whole computation is nondeterministic, the type of the final result
+values is a collection @Set[VS]@. For convenience, we define it as a type @Ans@:
 \begin{lstlisting}
@@
 
-As we mentioned previously, the components of states are lifted as arguments to
-@aeval@. Therefore, the pattern matching on state also becomes just matching on
+As previously mentioned, the components of states are lifted as arguments to
+@aeval@. Therefore, the pattern matching on the state becomes just a match on
 expression @e@.
@@
 An additional argument @continue@ which has type @Cont@ is also introduced.
-Recall that in the disentangled AAM, the second case in function @drive_step@
+Recall that in the disentangled AAM, the second case in @drive_step@
 makes a call to @continue@. Here, the reemerging @continue@ is a function
-with type @Ans => Ans@. The function @aeval@ does not return, instead, it calls
+with type @Ans => Ans@. The function @aeval@ does not return: instead, it calls
 @continue@ with the result values.
 
-\textbf{NonDeterminism Abstraction.}
-To handle the inevitable nondeterminism, we defined a generic function @nd@ as follows:
+\textbf{Nondeterminism Abstraction.}
+To handle the inevitable nondeterminism, we define a generic function @nd@ as follows:
 \begin{lstlisting}
@@
 
-Given a set of element of type @T@, an initial accumulated value of type @S@, a function @f@
+Given a set of elements of type @T@, an initial accumulated value of type @S@, a function @f@
 that works on element type @T@ and accumulated type @S@, and a function @g@ that
-works on and returns value of type @S@, @nd@ applies @f@ to each element in the
-set iteratively, accumulates the value and finally applies @g@ on it.
-Seeing this one may bring @fold@ to mind, indeed, it is an implementation of
-@fold@ in continuation style. But here we have two continuations, one for the
-next element in the set, and one for the final result.
-Sometime it is also called extended continuation-passing style (ECPS) that has a continuation and a
-meta-continuation \cite{Danvy:1990:AC:91556.91622}.
+works on and returns a value of type @S@, @nd@ applies @f@ to each element in the
+set iteratively, accumulates the value, and finally applies @g@ to it.
+Readers familiar with functional programming may think of @fold@ at this point; indeed, it is an implementation of @fold@ in continuation style.
+But here we have two continuations, one for the next element in the set, and one for the final result.
+This is sometimes referred to as an extended continuation-passing style (ECPS) that has a continuation and a meta-continuation \cite{Danvy:1990:AC:91556.91622}.
 \todo{correspondence to disentangled version, refunc the NDCont}
@@
 \textbf{Evaluation without Caching.}
-Since our first version of @aeval@ just implements the nondeterministic
-evaluation, and given the defined @nd@, the evaluation function @aeval@ can be easily
-defined. For the case that @e@ is a @let@ expression,
-the overall idea is to perform a two step depth-first evaluation. \todo{refine}
+Since our first version of @aeval@ implements the nondeterministic
+evaluation, given the defined @nd@, the evaluation function @aeval@ can be easily
+defined. In the case that @e@ is a @let@ expression,
+the overall idea is to perform a two-step, depth-first evaluation. \todo{refine}
 
@@
 The outer function call @nd@ evaluates over the set of possible closures of
-@f@. For each closures, we go into its body with new environment and new store.
-Remember that the function @aeval@ returns a set of @VS@ to its continuation,
-so the continuation argument @bodyvss@ is a set of @VS@ that represents values
-from multiple computation pates when evaluating this single @body@ expression.
-
-Then for the inner call of function @nd@ evaluates over the set of body values @bodyvss@.
-For each @VS@, the time stamp and store in @VS@ will be instantiated to
+@f@. For each such closure, we go into its body with a new environment and a new store.
+Remember that the function @aeval@ returns a set of @VS@ objects to its continuation,
+so the continuation argument @bodyvss@ is a set of @VS@ objects which represent values
+from multiple computation paths when evaluating this single @body@ expression.
+
+We next consider the inner call of function @nd@ which evaluates over the set of body values @bodyvss@.
+For each @VS@, the timestamp and store in @VS@ will be instantiated to
 @aeval@.
-The inner application of @aeval@ returns set of values of one computation path to its
+The inner application of @aeval@ returns a set of values of one computation path to its
 continuation, where the continuation will accumulate the result and transfer

### 2018-03-17 paper.tex
spacing

@@
 
-The outer function call @nd@ evaluates over the set of possible closures of
+The outer call to @nd@ evaluates over the set of possible closures of
 @f@. For each such closure, we go into its body with a new environment and a new store.
@@
 
-We next consider the inner call of function @nd@ which evaluates over the set of body values @bodyvss@.
+We next consider the inner call to @nd@, which evaluates over the set of body values @bodyvss@.
 For each @VS@, the timestamp and store in @VS@ will be instantiated to
@@
 
-If the expression is atomic, then we just need to construct a new instance of
-@VS@, return it to the @continue@ function.
+If the expression is atomic, then we simply need to construct a new instance of
+@VS@ and return it to the @continue@ function.
 

### 2018-03-17 paper.tex
Working on 6

@@
 
-If the expression is atomic, then we simply need to construct a new instance of
+If the expression is atomic, then we simply construct a new instance of
 @VS@ and return it to the @continue@ function.
 
-At the end, the initial invocation to @aeval@ is passed with an indentify
-function as halting continuation.
+At the end, the initial invocation to @aeval@ is passed with an identity
+function as a halting continuation:
 \begin{lstlisting}
@@
 
-Based on the refunctionalized AAM from the last section, we use the same @nd@ nondeterministic operator
-but further extend it to a fixed-point cached evaluation in this section.
+Based on the refunctionalized AAM from the last section, we use the same nondeterministic
+operator @nd@, but further extend it to a fixed-point cached evaluation in this section.
 The fixed-point caching algorithm performs a sound over-approximation of all possible
-concrete reachable pathes and values, and also prevents non-termination when analyzing diverged programs.
+concrete reachable paths and values, and also prevents non-termination when analyzing diverged programs.
 
 \textbf{Configuration.}
-Now we do not have the concept of explicit stack and state,
-but to represent the cache, we introduce the configuration @Config@ as
-a state-like definition which borrows the components from state but without
-continuations.
+We do not have the concept of an explicit stack or state,
+so to represent the cache we introduce a configuration @Config@ as
+a state-like definition which borrows the components from state, sans continuations.
 \begin{lstlisting}
@@
 Recall that the latent assumption of the worklist algorithm in classical AAM
-is if we have seen a state @s@, then it means we also have seen the successors
+is that if we have seen a state @s@, it means we also have seen the successors
 of state @s@.
-The caching algorithm we adopted here replay the same assumption but in a big-step
+The caching algorithm we adopt here replays the same assumption but in a big-step
 manner: if we have seen the configuration @c@, then it means we also have seen
@@
 
-The @Cache@ contains two maps @in@ and @out@ which are both mapping from
-configurations to sets of @VS@. The @in@ cache contains mappings from
-last iteration of evaluation, and @out@ cache contains mappings after this
+The @Cache@ contains two maps @in@ and @out@ which are both mappings from
+configurations to sets of @VS@. The @in@ cache contains mappings from the
+previous iteration of evaluation, and the @out@ cache contains mappings after the current
 iteration of evaluation. Once we have evaluated a term to some values,
-we update the @out@ cache. In the next time of iteration, we will use @out@
-from last iteration as @in@ cache for this iteration.
+we update the @out@ cache. In the next iteration, we will use the @out@
+from the previous iteration as the @in@ cache for the current iteration.
 
 \textbf{Cache-Passing Style.}
-Now we can modify out interpreter to cache-passing style.
-First the return type @Ans@ is changed to a case class that contains a set of @VS@ and
+We can now transform our interpreter into cache-passing style.
+We begin by changing the return type @Ans@ to a case class which contains a @VS@ set and
 a cache.
@@
 
-The skeleton of function @aeval@ largely remain the same.
-Every place has a value of @Ans@ is also replaced as direct pattern matching so
+The skeleton of function @aeval@ largely remains the same.
+Every place which has a value of @Ans@ is also replaced as direct pattern matching so
 that we can use the fields in @Ans@ immediately.
@@
 
-When we enter to function @aeval@, we firstly lookup in the @out@ cache,
-if there are some values for current configuration, then we use immediately.
-This corresponds to small-step AAM with a worklist when we have seen a state
-then we can simply discard the state and continue working the rest of worklist @todo@.
-
-Otherwise, if we not hit the @out@ cache, we retrieve the values from @in@ cache, and update @out@ cache
-with values from @in@ cache.  The retrieving from @in@ cache may return an empty set of values if
-there is no such mapping for the configuration from the previous iteration.
-It actually first conservatively assumes that all computation can be diverged,
-if not, we will update its values in the mapping after evaluation.
+Upon entering @aeval@, we first determine whether the @out@ cache
+contains some values for the current configuration and use them immediately (if present).
+This corresponds to small-step AAM with a worklist: when we have seen a state,
+we can simply discard the state and continue working through the rest of the worklist @todo@.
+
+If, however, we have not hit the @out@ cache, we retrieve the values from the @in@ cache, and update the @out@ cache
+with values from the @in@ cache. This retrieval from the @in@ cache may return an empty set of values if there is no such mapping for the configuration from the previous iteration.
+In reality, we first conservatively assume that all computations can be diverged, and
+if not, we update its values in the mapping after evaluation.
 In terms of partial orders and lattices, it is the case that starts

### 2018-03-17 paper.tex
Finished 6.2

@@
 
-After done the evaluation, we obtain a set of @VS@ and need to update the @out@ cache.
-In the first case of pattern matching, it happens inside of the last continuation of
-outer @nd@.
-For the second case, we update the @out@ cache before calling the continuation @continue@.
+After completing the evaluation, we obtain a set of @VS@ objects and must update the @out@ cache.
+In the first case of pattern matching, this happens inside the last continuation of the
+outer @nd@ call.
+For the second case, we update the @out@ cache before calling the @continue@ continuation.
 
@@
 
-Finally, the @analyze@ function as the entrance of the analysis does a looping iteration
+Finally, the @analyze@ function (as the entrance of the analysis) does a looping iteration
 to find the fixed-point on caches starting from an empty cache.
@@
 and thus can be returned.
-Otherwise, a next iteration with the new @out@ cache will be activated.
+Otherwise, the next iteration with the new @out@ cache will be activated.
 Note that the continuation of the initial call to @aeval@ sets up the values for the

### 2018-03-17 paper.tex
Finished 6

@@
 
-In the last section, we just established a computable pushdown control-flow
+In the previous section, we have established a computable pushdown control-flow
 analysis through refunctionalization and caching. In this section,
-we revisit the pushdown control-flow problem and examine what we have done to
-achieve it.
-
-\textbf{The Problem of Return-flows.}
-Pushdown control-flow is a property in analysis that precisely models
-the runtime call-stack of the analyzed program.
-A pushdown control flow analysis provides as exact as runtime return-flow
-when analyzing the program, but traditional control flow analysis collapses
-the state space into a finite space thus causes imprecise stack modeling.
+we revisit the pushdown control flow problem and examine what we have done to
+overcome it.
+
+\textbf{The Problem with Return-flows.}
+Pushdown control flow is a property in analysis that precisely models
+the runtime call stack of the analyzed program.
+A pushdown control flow analysis provides an as-exact-as-runtime return-flow
+when analyzing a program, but traditional control flow analysis collapses
+the state space into a finite space and thus causes imprecise stack modeling.
 
@@
 
-In $k$-CFA algorithm or the abstract abstract machine we showed in section
-\ref{aam},
-The call sites @(id 2)@ and @(id 1)@ share the same return flow,
-so the of invocation @(id 2)@ returns to both call-sites @[x (id 1)]@ and @[y (id 2)]@.
+In $k$-CFA algorithm or the abstract abstract machine shown in Section~\ref{aam},
+the call sites @(id 2)@ and @(id 1)@ share the same return flow,
+so the invocation of @(id 2)@ returns to both call-sites @[x (id 1)]@ and @[y (id 2)]@.
 Therefore, the returned value @2@ for variable @y@ is also propagated to
-variable @x@ and then imprecise analysis result is arose.
-This return-flow merging is inevitable even we increase the context-sensitivity.
-If we use the monovariant analysis (i.e., 0CFA), the analyzing result would be @x@ and
-@y@ point to value set @{1, 2}@, because the algorithm not distinguishes
-@z@ comes from different call-sites.
+variable @x@, causing imprecise analysis results to arise.
+This return-flow merging is inevitable, even when increasing the context-sensitivity.
+If we use the monovariant analysis (i.e., $0$-CFA), the analysis result would be such that @x@ and @y@ point to value set @{1, 2}@, because the algorithm does not distinguish that
+@z@ comes from a different call site.
 Under $1$-CFA, the algorithm is able to distinguish that variable @z@ of function
-@id@ has two different values at two call-sites, so variable @y@
+@id@ has two different values at two call sites, so variable @y@
 would not be polluted by @1@.
-However, variable @x@ still points to value set @{1, 2}@.
-Because two call-sites still share the same continuation and
-it is still helplessness for separating the return-flows.
+However, variable @x@ still points to value set @{1, 2}@, because two call-sites still share the same continuation $1$-CFA is still unable to separate the return-flows.
 

### 2018-03-17 paper.tex
update

@@
 
-The abstract syntax can be straightforwardly described in Scala as follows:
-\note{TR: grammar below isn't ANF. Either make it so, or clarify that
-we assume ANF only as convention and don't need/want to enforce it in
-the term structure}
+The abstract syntax in Scala is shown as follows. We assume the source 
+program conforms ANF's convention, and not force it in the term structure 
+of Scala constructs.
 
@@
 In our toy language, the only category of value is a closure, i.e., a function paired with an environment.
-4) \textit{Continuation} represents the program stack. \todo{explain more on stack}
-
-The Scala representations for the components of CESK machine are as follows:
+4) \textit{Continuation} represents the program stack. In this paper, 
+we instantiate the stack as a list of frames because the ANF simplifies the evaluation context.
+For direct style $\lambda$-calculus, the continuations of CEK/CESK machines can be presented by 
+variants of a data type.
+
+The Scala representations for the components of the CESK machine are as follows:
 
@@
 A @Frame@ constitutes a variable name @x@ to be bound later, a control expression
-to which the program may resume, and its environment. \note{JD: someone check this and make sure I got it right}
-\todo{mention that frame list vs KArg/KFun}
+to which the program may resume, and its environment.
 
@@
 1) @Halt@ which corresponds to the empty stack, and
-2) @ContAddr@ consists of the entry expressions of callee, which are also
+2) @ContAddr@ consists of the target expressions of callee, which are also
 finite.
-One may ask, why not also keep track of the time in continuation addresses?
-\todo{why target expression, allocation polyvariance}
+%One may ask, why not also keep track of the time in continuation addresses?
+%\todo{why target expression, allocation polyvariance}
 
@@
 as we will later transform the stack to higher-order functions
-represent \note{``representing'' or ``represented by''?} continuations, it does not matter what kind of AAM we start from.
+representing continuations, it does not matter what kind of AAM we start from.
 This is because transforming to higher-order functions forces us to sequentialize
@@
 and returns a set of reachable states once it terminates.
-\note{Do we really need this section?
-It seems unlikely that this requires its own section.
-Could we just append to the last section: For convenience, we combine these functions into one, called \texttt{drivestep}?}
 
@@
 
-Refunctionalization transforms first-order data types represented as \note{``represented as'' or ``representing?'' It should be representing, right?}
+Refunctionalization transforms first-order data types representing
 evaluation contexts to higher-order functions.
@@
 works on and returns a value of type @S@, @nd@ applies @f@ to each element in the
-set iteratively, accumulates the value, and finally applies @g@ to it.
-Readers familiar with functional programming may think of @fold@ at this point; indeed, it is an implementation of @fold@ in continuation style.
-But here we have two continuations, one for the next element in the set, and one for the final result.
-This is sometimes referred to as an extended continuation-passing style (ECPS) that has a continuation and a meta-continuation \cite{Danvy:1990:AC:91556.91622}.
-\todo{correspondence to disentangled version, refunc the NDCont}
+set iteratively, accumulates the value and finally applies @g@ on it.
+
+If we look at @continue@ and @ndcontinue@ in the disentangled AAM, we will find
+that @nd@ is an abstraction of those two functions. In function @continue@,
+there are two out calls to @ndcontinue@ and @drive_step@, this is the reason why 
+we have functions @f@ and @g@ in @nd@.
+
+Readers familiar with functional programming may think of @fold@ at this point; indeed, 
+it is an implementation of @fold@ in continuation style. But here we have two 
+continuations, one for the next element in the set, and one for the final result.
+This is sometimes referred to as an extended continuation-passing style (ECPS) 
+that has a continuation and a meta-continuation \cite{Danvy:1990:AC:91556.91622}.
 
@@
 
-This paper is inspired a lot from \citeauthor{darais2017abstracting}'s work.
+This paper is inspired greatly from \citeauthor{darais2017abstracting}'s work.
 Instead of just showing the final form of abstract definitional interpreters,

### 2018-03-17 paper.tex
Done with first pass

@@
 
-In previous section, we presented a refunctionalized AAM with @nd@ operator
+In the previous section, we presented a refunctionalized AAM utilizing the @nd@ operator
 in extended continuation-passing style.
@@
 
-To do this at this point, we actually have multiple choices:
+We have multiple choices of how to proceed:
 \begin{itemize}
-  \item By monadifying the continuations into monads, we obtain the abstracting
+  \item By monadifying the continuations, we obtain the abstracting
     definitional interpreters in monadic style which are what \citeauthor{darais2017abstracting}
-    described \cite{darais2017abstracting}.
+    describe \cite{darais2017abstracting}.
   \item By representing the extended continuation-passing style with delimited controls,
     we derive a new form of abstract interpreters in direct-style.
-  \item In fact, we may also just use the same caching algorithm but with side effect such as
-    assignments and mutations to update the cache and then achieve the same
+  \item In fact, we may also just use the same caching algorithm but with side effects such as
+    assignments and mutations to update the cache, and then achieve the same
     definitional interpreter with pushdown control flows. In this case, nondeterminism can be
-    well handled by for comprehension.
+    easily handled for comprehension.
 \end{itemize}
 
-These coincidence should not be a surprise since there are extraordinarily rich literatures
-show the correspondence between monads and continuation-passing style as well as delimited
-controls since their very beginning era. \todo{cite}
-
-In this section, we present the version that using delimited controls.
+These coincidences should not be a surprise since existing literature is extraordinarily rich
+in showing the correspondence between monads and continuation-passing style as well as
+delimited controls since their very beginning era. \todo{cite}
+
+In this section, we present the version that uses delimited controls.
 
@@
 
-As we identified before,
+As identified before, \note{JD: compelling...}
 
@@
 
-May be insert to section refunc.
+Maybe insert to section refunc.
 
@@
 defunctionalized continuation.
-Unfortunately, the abstract abstract machines is nondeterministic, where
-the nondeterminism is handled by a work list.
+Unfortunately, abstract abstract machines are nondeterministic, where
+the nondeterminism is handled by a worklist.
 We can turn this explicit nondeterminism into the reduction context of
@@
 
-\textbf{Pushdown Control flow Analysis.}
+\textbf{Pushdown Control Flow Analysis.}
 In recent years, there are significant efforts \cite{vardoulakis2010cfa2, earl2012introspective,
@@
 
-CFA2 is the first solution that solves the return-flows problem\cite{vardoulakis2010cfa2}.
-But CFA2 has several drawbacks: it works on only continuation-passing style programs,
-and not supports polyvariant analysis, in addition with an exponential time
+CFA2 is the first solution that solves the return-flows problem\cite{vardoulakis2010cfa2},
+but CFA2 has several drawbacks: it works only on continuation-passing style programs,
+and does not supports polyvariant analysis, in addition to an exponential time
 complexity.
@@
 for small-step AAM to achieve pushdown analysis \cite{gilray2016pushdown}.
-The method is both simple to implement and computational inexpensive.
-Based on the AAM we presented in Section \ref{aam}, the only change in the code is
-not only keeping track of the entry-point expression of callee, but also holding the target environment
-when allocating a continuation addresses.
+This method is both simple to implement and computationally inexpensive.
+Based on the AAM we presented in Section \ref{aam}, the only changes in the code required is
+not only keeping track of the entry-point expression of callee, but also holding the target
+environment when allocating a continuation addresses.
 The following code shows the implementation:
@@
 
-The target environment is used for successor states and extended from closure's
+The target environment is used for successor states and extended from the closure's
 environment (namely @val new_env = env_c + (v -> baddr)@).
-All the other part of the code would not be changed.
+No other pieces of code would need to be modified.
 Notably, this change only causes constant-factor increase in time complexity of
@@
   programming languages} \cite{reynolds1972definitional} shows that in
-definitional interpreters the defined language can inherit properties from the defining language.
+definitional interpreters, the defined language can inherit properties from the defining language.
 With this insight, \citeauthor{darais2017abstracting} construct abstracting
-definitional interpreters that automatically inherits the pushdown control-flow
+definitional interpreters which automatically inherit the pushdown control-flow
 property from its defining language
-because the defined language just uses the stack the meta-language \cite{darais2017abstracting}.
-\citeauthor{darais2017abstracting}'s abstracting definitional interpreters work
-on direct-style programs, and is written in monadic style. One of the advantages
-of monadic abstract interpreters is modular and composable, therefore deploying
-different sensitivities or features is just applying a different monad.
+because the defined language simply uses the stack the
+meta-language~\cite{darais2017abstracting}. \citeauthor{darais2017abstracting}'s abstracting
+definitional interpreters work on direct-style programs, and is written in monadic style. One
+of the advantages of monadic abstract interpreters is modular and composable, therefore
+deploying different sensitivities or features is just applying a different monad.
 Prior to that, \citeauthor{Sergey:2013:MAI:2491956.2491979} show a monadic
@@
 
-This paper is inspired greatly from \citeauthor{darais2017abstracting}'s work.
-Instead of just showing the final form of abstract definitional interpreters,
-We reveal that refunctionalization plays an important role for inheriting
-stack from the meta-language.
+This paper is greatly inspired by~\citeauthor{darais2017abstracting}'s work.
+Instead of simply showing the final form of abstract definitional interpreters,
+we reveal that refunctionalization plays an important role for inheriting the
+stack from the metalanguage.
+
 The target of the series of transformations in this paper is abstract definitional
@@
 monadic- and direct-style.
-The difference is ours work for A-Normal Form $\lambda$-calculus instead of plain
-$\lambda$-calculus, our work also can be easily adopted to plain
+The difference is that ours works for A-Normal Form $\lambda$-calculus, instead of plain
+$\lambda$-calculus, although our work can also be easily adopted to handle a plain
 $\lambda$-calculus.

### 2018-03-17 paper.tex
update

@@
        [start branch=C] \chainin (m-3-5)
-       [join={node[above, labeled] {\text{Known as Danvy's correspondence for concrete semantics artifacts }}}];
+       [join={node[above, labeled] {\text{Known as Danvy's correspondences for concrete semantics artifacts }}}];
      }
@@
 
-\subsection{One Step Back: Unabstracted Stack}
+\subsection{One Step Back: Unabstracted Stack} \label{unabs}
 
@@
 This is because transforming to higher-order functions forces us to sequentialize
-the order of abstract evaluation, and neither versions of AAM requires construction of a frame of lists \note{should this be frame of lists? sounds weird}, nor to allocate continuations in the store.
-Thus, our choice to start from an AAM with unbounded stack is motivated simply because it has an equivalent stack model to abstract definitional interpreters (our final target).
-
-\todo{Unbounded Stack AAM is PD, justify this}
+the order of abstract evaluation, and neither requiring construction of a frame on stack, 
+nor to allocate continuations in the store.
+Thus, our choice to start from an AAM with unbounded stack is motivated simply because 
+it has an equivalent stack model to abstract definitional interpreters (our final target).
 
@@
 \textbf{Pushdown Control Flow Analysis.}
-In recent years, there are significant efforts \cite{vardoulakis2010cfa2, earl2012introspective,
+In recent years, there are significant efforts~\cite{vardoulakis2010cfa2, earl2012introspective,
 gilray2016pushdown, johnson2015abstracting} to achieve precise call/return
@@
 
-CFA2 is the first solution that solves the return-flows problem\cite{vardoulakis2010cfa2},
+CFA2 is the first solution that solves the return-flows problem~\cite{vardoulakis2010cfa2},
 but CFA2 has several drawbacks: it works only on continuation-passing style programs,
@@
 
-\todo{PDCFA, AAC}
-
-\citeauthor{gilray2016pushdown} proposed a polyvariant continuation-addresses allocator
+As mentioned in Section~\ref{unabs}, maintaining the perfect precision provided by an unbounded-stack machine is essential when performing the proposed transformations.
+Pushdown control flow analysis (PDCFA) is a mechanism which maintains this precision through the use of a specialized state graph representing all possible stacks contained within the unbounded-stack machine.
+These state graphs are known as \emph{Dyke state graphs}~\cite{gilray2016pushdown}.
+
+Similar to PDCFA, abstracting abstract control (AAC) is another strategy for maintaining stack precision without any loss.
+As shown by~\citeauthor{gilray2016pushdown}, AAC functions similar to PDFCA, including the fact that neither method requires any context sensitivity.
+AAC functions by utilizing continuations which are specific to both the source and target states of a call-site transition, which guarantees that no merging will occur during returns~\cite{gilray2016pushdown}.
+
+\citeauthor{gilray2016pushdown} further proposed a polyvariant continuation-addresses allocator
 for small-step AAM to achieve pushdown analysis \cite{gilray2016pushdown}.

### 2018-03-17 paper.tex
Adding PDFCA and AAC in related work

@@
        [start branch=C] \chainin (m-3-5)
-       [join={node[above, labeled] {\text{Known as Danvy's correspondences for concrete semantics artifacts }}}];
+       [join={node[above, labeled] {\text{Known as Danvy's correspondence for concrete semantics artifacts }}}];
      }
@@
 This is because transforming to higher-order functions forces us to sequentialize
-the order of abstract evaluation, and neither requiring construction of a frame on stack, 
-nor to allocate continuations in the store.
-Thus, our choice to start from an AAM with unbounded stack is motivated simply because 
-it has an equivalent stack model to abstract definitional interpreters (our final target).
+the order of abstract evaluation, and neither versions of AAM requires construction of a frame of lists \note{should this be frame of lists? sounds weird}, nor to allocate continuations in the store.
+Thus, our choice to start from an AAM with unbounded stack is motivated simply because it has an equivalent stack model to abstract definitional interpreters (our final target).
+
+\todo{Unbounded Stack AAM is PD, justify this}
 

### 2018-03-17 paper.tex
related work

--- 2018-03-17 paper.tex
@@
 
-\textbf{Abstracting Abstract Machines.}
+\textbf{Abstract Interpretation and Control-flow Analysis.}
+\citeauthor{cousot1977abstract} invented abstract interpretation as a sound approach 
+to approximate the program's runtime behaviors\cite{cousot1977abstract}.
+Control-flow analysis is one of the instances of abstract interpretation on functional programs 
+that can be traced to Jones \cite{jones1981flow}. 
+\citeauthor{Shivers:1988:CFA:53990.54007} introduces $k$-CFA which uses $k$ recent calling contexts as program 
+contour that differentiates values from different context \cite{Shivers:1988:CFA:53990.54007, Shivers:1991:SSC:115865.115884}.
+The modern formulation of control-flow analysis is the abstracting abstract machines 
+methodology \cite{van2012systematic, van2010abstracting} which forms the starting point of this paper.
+Modern programming languages such as Java\cite{might2010resolving} and Racket \cite{Tobin-Hochstadt:2012:HSE:2384616.2384655}
+also can be modeled by abstracting abstract machine approach.
 
 \textbf{Defunctionalization and Refunctionalization.}
-
 Defunctionalization and refunctionalization build connections between abstract

### 2018-03-17 paper.tex
Proofreading related work

--- 2018-03-17 paper.tex
@@
 
-\textbf{Abstracting Abstract Machines.}
+\textbf{Abstract Interpretation and Control-flow Analysis.}
+\citeauthor{cousot1977abstract} invented abstract interpretation as a sound approach
+to approximate a program's runtime behavior\cite{cousot1977abstract}.
+Control flow analysis is one instance of abstract interpretation on functional programs
+that can be traced to Jones \cite{jones1981flow}.
+\citeauthor{Shivers:1988:CFA:53990.54007} introduces $k$-CFA which uses $k$ recent calling contexts as program contour that differentiates values from different contexts \cite{Shivers:1988:CFA:53990.54007, Shivers:1991:SSC:115865.115884}.
+The modern formulation of control flow analysis is the abstracting abstract machines
+methodology \cite{van2012systematic, van2010abstracting} which forms the starting point of this paper.
+Modern programming languages such as Java\cite{might2010resolving} and Racket \cite{Tobin-Hochstadt:2012:HSE:2384616.2384655}
+can also be modeled by the abstracting abstract machine approach.
 
 \textbf{Defunctionalization and Refunctionalization.}
-
 Defunctionalization and refunctionalization build connections between abstract

### 2018-03-17 paper.tex
Fixing my mistake

@@
 To effectively approximate the runtime behavior, we introduce a finite
-program counter @time@ that encodes the program execution history.
+program contour @time@ that encodes the program execution history.
 We use a list of execution contexts (expressions) to represent this, and
@@
 In $k$-call-sensitive analysis, a history of the last $k$ call sites is
-used as a finite program counter.
+used as a finite program contour.
 The history is represented as a list of expressions and embedded
@@
 address for @Halt@ to an empty set of continuations.
-We also provide an empty program counter as our initial time.
+We also provide an empty program contour as our initial time.
 \begin{lstlisting}

### 2018-03-17 paper.tex
update

@@
 Defining a language by building an interpreter for it can be traced to the very
-early time of programming languages research \cite{landin1966next, reynolds1972definitional}.
+early days of programming languages research \cite{landin1966next, reynolds1972definitional}.
 Nowadays, even an undergraduate student in computer science is able to build toy languages
 through interpreters.
-But building a sound abstract interpreter remains as an esoteric and difficult task until recent years.
-
-\citeauthor{van2012systematic} proposed Abstracting Abstract Machines (AAM) methodology which can used
-for constructing sound abstract interpreters of higher-order functional languages from concrete abstract machines
+But building a sound abstract interpreter remained an esoteric and difficult task until very recently.
+
+\citeauthor{van2012systematic} proposed the Abstracting Abstract Machines (AAM) methodology which provides a recipe
+for constructing sound abstract interpreters for higher-order functional languages from concrete abstract machines
 ~\cite{van2012systematic, van2010abstracting}.
@@
 Broadly speaking, abstract abstract machines and abstract definitional
-interpreters are both some kind of abstract interpreters.
+interpreters are different forms of abstract interpreters.
 They are obtained by applying a combination of abstractions to their
 concrete counterparts, abstract machines and definitional interpreters, respectively.
-An interesting question would be how we can interderive these abstract semantics artifacts
-from the other one.
+An interesting question, and the subject of this paper, is how we can interderive these abstract semantics artifacts
+from the respective other one.
 
 In the concrete world, the relations among reduction semantics, abstract machines,
-definitional interpreters, and monadic interpreters is intensively studied by Danvy and his
-collaborators in the past years
+definitional interpreters, and monadic interpreters have been intensively studied by Danvy and his
+collaborators
 \cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}.
-The abstract machines implement structural operational semantics in continuation style,
+The concrete abstract machines implement structural operational semantics in continuation-passing style,
 where the reduction contexts are defunctionalized continuations. One can derive definitional
@@
 
-On the contrary, in terms of their abstract semantics artifacts,
+In the domain of abstract semantics artifacts, by contrast,
 the relations between small-step abstract abstract machines and big-step abstract
-definitional interpreters, as well as the question of deriving one from the other is not
-well-studied.
+definitional interpreters, as well as the question of deriving one from the other, are not
+well studied.
 The fundamental difference between concrete semantics artifacts and abstract semantics
@@
 the least fixed-point eventually.
-In this functional pearl, we practically answer the question and relate AAM and ADI
+In this functional pearl, we provide a constructive answer to the question 
+of interderiving abstract semantics artifacts and relate AAM and ADI
 by presenting a series of syntactical transformations on the program from
@@
 In addition, the abstract abstract machines with unbounded stack naturally correspond
-to abstract definitional interpreters. We show after refunctionalizing an AAM with unbounded
+to abstract definitional interpreters. We show that after refunctionalizing an AAM with unbounded
 stack, and with the proper caching algorithm, the pushdown control-flow analysis can be
@@
 
-\subsection{Contribution}
+\subsection{Contributions}
 
@@
 explain why \citeauthor{darais2017abstracting}'s abstract definitional interpreters
-is able to extend the the pushdown control-flow property from its defining language.
+is able to inherit the pushdown control-flow property from its defining language.
 
@@
 
-As mentioned in Section~\ref{unabs}, maintaining the perfect precision provided by an unbounded-stack machine is essential when performing the proposed transformations.
-Pushdown control flow analysis (PDCFA) is a mechanism which maintains this precision through the use of a specialized state graph representing all possible stacks contained within the unbounded-stack machine.
-These state graphs are known as \emph{Dyke state graphs}~\cite{gilray2016pushdown}.
+Pushdown control flow analysis (PDCFA) is another mechanism which maintains this precision 
+through the use of a specialized state graph representing all possible stacks contained 
+within the unbounded-stack machine. These state graphs are known as Dyke state graphs.
 

### 2018-03-17 paper.tex
pass over intro

@@
 
-Pushdown control flow analysis (PDCFA) is another mechanism which maintains this precision 
-through the use of a specialized state graph representing all possible stacks contained 
-within the unbounded-stack machine. These state graphs are known as Dyke state graphs.
+As mentioned in Section~\ref{unabs}, maintaining the perfect precision provided by an unbounded-stack machine is essential when performing the proposed transformations.
+Pushdown control flow analysis (PDCFA) is a mechanism which maintains this precision through the use of a specialized state graph representing all possible stacks contained within the unbounded-stack machine.
+These state graphs are known as \emph{Dyke state graphs}~\cite{gilray2016pushdown}.
 

### 2018-03-17 paper.tex
update

--- 2018-03-17 paper.tex
@@
 
-\item \textbf{Refunctionalization} is shown in Section~\ref{refunc} which sequentializes
+\item \textbf{Refunctionalization} is shown in Section~\ref{sec:refunc} which sequentializes
   the order of abstract execution by higher-order functions.
@@
 
-\section{Refunctionalization} \label{refunc}
+\section{Refunctionalization} \label{sec:refunc}
 
@@
 
+\iffalse
 \section{Refunctionalized AAM}
-
 Maybe insert to section refunc.
-
 \subsection{Store Widening}
-
 \subsection{Precise Call/Return Match}
-
 show the precision of stack
-
 On equivalence: since refunc interp does not have an explicit concept of $state$,
@@
 or 2) we can compare the summarized store of P4F against with summarized store of refunc interp.
-
 \subsection{Termination}
-
 show that number of continuation is finite.
-
 \subsection{Soundness}
-
 \subsection{Complexity}
+\fi
 
@@
 machines and interpreters.
-
-For deterministic languages, a reduction semantics is a structural operational
-semantics in continuation style, where the reduction context is a
-defunctionalized continuation.
-Unfortunately, abstract abstract machines are nondeterministic, where
-the nondeterminism is handled by a worklist.
-We can turn this explicit nondeterminism into the reduction context of
-AAM.
+\citeauthor{reynolds1972definitional} first shows refunctionalization as a program transformation technique that 
+can be used to transform higher-order functions to first-order functions 
+when sepcifing a programming language by interpreter \cite{reynolds1972definitional}.
+Refunctionalization is the reversed transformation.
+
+Over the years, Danvy and his collaborators applied refunctionalization and defunctionalization to 
+many different abstract machines and semantics, including the CEK machines, the CLS machines, 
+the SECD machines, and etc
+\cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}.
+And other applications are also found to be applicable, such as Dyck words recognizer and Dijkstra's shunting-yard algorithm \cite{Danvy:2006:RW:2171265.2171268}.
+
+For deterministic languages, \citeauthor{Danvy:2008:DIP:1411204.1411206} shows 
+a reduction semantics is a structural operational semantics in continuation style, 
+where the reduction context is a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}.
+Our work further shows for nondeterministic languages, there is another layer of continuations
+that controls the nondeterministic choices. After identifing and explicitly exposing (section \ref{linear}) that layer
+of nondeterministic continuations, the whole program can be then refunctionalized to extended continuation-passing style.
+
+Fusion as one of the transformation in our paper is also studied by \citeauthor{Ohori:2007:LFF:1190216.1190241} \cite{Ohori:2007:LFF:1190216.1190241}.
 
@@
 complexity.
-
-Pushdown control flow analysis (PDCFA) is another mechanism which maintains this precision 
-through the use of a specialized state graph representing all possible stacks contained 
-within the unbounded-stack machine. These state graphs are known as Dyke state graphs.
-
-Similar to PDCFA, abstracting abstract control (AAC) is another strategy for maintaining stack precision without any loss.
-As shown by~\citeauthor{gilray2016pushdown}, AAC functions similar to PDFCA, including the fact that neither method requires any context sensitivity.
-AAC functions by utilizing continuations which are specific to both the source and target states of a call-site transition, which guarantees that no merging will occur during returns~\cite{gilray2016pushdown}.
+Pushdown control flow analysis (PDCFA) is a mechanism which maintains this precision 
+through the use of a Dyke state graph representing all possible stacks contained 
+within the unbounded-stack machine\cite{earl2012introspective, earl2010pushdown}.
+Similar to PDCFA, abstracting abstract control (AAC) is another strategy for 
+maintaining stack precision \cite{johnson2015abstracting}.
+AAC functions by utilizing continuations which are specific to both 
+the source and target states of a call-site transition, 
+which guarantees that no spurious merging will occur during returns.
 
@@
 for small-step AAM to achieve pushdown analysis \cite{gilray2016pushdown}.
-This method is both simple to implement and computationally inexpensive.
+This method is both simple to implement and computationally inexpensive and so called Pushdown for Free (P4F).
 Based on the AAM we presented in Section \ref{aam}, the only changes in the code required is
@@
 environment when allocating a continuation addresses.
-The following code shows the implementation:
-
-\begin{lstlisting}
-case class P4FContAddr(tgt: Expr, tgtEnv: Env) extends KAddr
-def allocKont(tgtExpr: Expr, tgtEnv: Env): KAddr = P4FContAddr(tgtExpr, tgtEnv)
-\end{lstlisting}
-
-The target environment is used for successor states and extended from the closure's
-environment (namely @val new_env = env_c + (v -> baddr)@).
 No other pieces of code would need to be modified.
@@
 the analysis if the store is widened.
+
+\iffalse
+The following code shows the implementation:
+\begin{lstlisting}
+case class P4FContAddr(tgt: Expr, tgtEnv: Env) extends KAddr
+def allocKont(tgtExpr: Expr, tgtEnv: Env): KAddr = P4FContAddr(tgtExpr, tgtEnv)
+\end{lstlisting}
+The target environment is used for successor states and extended from the closure's
+environment (namely @val new_env = env_c + (v -> baddr)@).
+\fi
 
@@
 Prior to that, \citeauthor{Sergey:2013:MAI:2491956.2491979} show a monadic
-abstract interpreter for small-step semantics.
+abstract interpreter for small-step semantics \cite{Sergey:2013:MAI:2491956.2491979}.
 
 This paper is greatly inspired by~\citeauthor{darais2017abstracting}'s work.
-Instead of simply showing the final form of abstract definitional interpreters,
+The difference on the surface is that ours works for A-Normal Form $\lambda$-calculus, 
+instead of plain $\lambda$-calculus, although our work can also be easily adopted to handle a plain
+$\lambda$-calculus.
+But except of simply showing the final form of abstract definitional interpreters,
 we reveal that refunctionalization plays an important role for inheriting the
 stack from the metalanguage.
-
 The target of the series of transformations in this paper is abstract definitional
-interpreters, and we additionally demonstrate abstract interpreters written in both
-monadic- and direct-style.
-The difference is that ours works for A-Normal Form $\lambda$-calculus, instead of plain
-$\lambda$-calculus, although our work can also be easily adopted to handle a plain
-$\lambda$-calculus.
+interpreters, and we additionally demonstrate abstract interpreters written direct-style with 
+delimited controls. The correspondence of delimited controls (as well as continuations) and monads
+are well-known \cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions}.
 

### 2018-03-17 paper.tex
Related work section

@@
     definitional interpreter with pushdown control flows. In this case, nondeterminism can be
-    easily handled for comprehension.
+    easily handled via @for@ comprehension.
 \end{itemize}
@@
 machines and interpreters.
-\citeauthor{reynolds1972definitional} first shows refunctionalization as a program transformation technique that 
+\citeauthor{reynolds1972definitional} first shows defunctionalization as a program transformation technique that
 can be used to transform higher-order functions to first-order functions
-when sepcifing a programming language by interpreter \cite{reynolds1972definitional}.
-Refunctionalization is the reversed transformation.
+when specifying a programming language by an interpreter \cite{reynolds1972definitional}.
+Refunctionalization is the reverse of this transformation.
 
@@
 many different abstract machines and semantics, including the CEK machines, the CLS machines,
-the SECD machines, and etc
-\cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}.
-And other applications are also found to be applicable, such as Dyck words recognizer and Dijkstra's shunting-yard algorithm \cite{Danvy:2006:RW:2171265.2171268}.
+the SECD machines, etc.
+\cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}
+Other applications are also found to be applicable, such as Dyck word recognizer, and Dijkstra's shunting-yard algorithm \cite{Danvy:2006:RW:2171265.2171268}.
 
@@
 where the reduction context is a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}.
-Our work further shows for nondeterministic languages, there is another layer of continuations
-that controls the nondeterministic choices. After identifing and explicitly exposing (section \ref{linear}) that layer
-of nondeterministic continuations, the whole program can be then refunctionalized to extended continuation-passing style.
+Our work further shows that for nondeterministic languages, there is another layer of continuations
+that controls the nondeterministic choices. After identifying and explicitly exposing (Section~\ref{linear}) that layer
+of nondeterministic continuations, the whole program can then be refunctionalized to extended continuation-passing style.
 

### 2018-03-17 paper.tex
section 7

@@
 in showing the correspondence between monads and continuation-passing style as well as
-delimited controls since their very beginning era. \todo{cite}
+delimited controls since their very beginning era \cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions}.
 
@@
 
+\todo{Update the code}
 \subsection{Un-CPS by Delimited Controls} \label{uncps}
 
-As identified before, \note{JD: compelling...}
-
-\begin{lstlisting}
-\end{lstlisting}
-
-\iffalse
-\section{Refunctionalized AAM}
-Maybe insert to section refunc.
-\subsection{Store Widening}
-\subsection{Precise Call/Return Match}
-show the precision of stack
-On equivalence: since refunc interp does not have an explicit concept of $state$,
-so 1) we can compare the final states of P4F against with refunc interp,
-or 2) we can compare the summarized store of P4F against with summarized store of refunc interp.
-\subsection{Termination}
-show that number of continuation is finite.
-\subsection{Soundness}
-\subsection{Complexity}
-\fi
+As identified before, there are two layers of continuations, one for the normal stack of 
+the analyzed language, and the other for the nondeterministic choices of closures.
+This corresponds to the extended continuation-passing style.
+But in our defining language Scala, the delimited controls rely on the continuation-passing
+transformation which is only one layer.
+To address this problem, base on the @nd@ operator we have before, we introduce 
+another operator @ndcps@ for handling non-deterministic choices.
+
+The @ndcps@ itself is implemented with delimited controls. 
+The function takes two arguments: a set of elements with type @T@ and an initial 
+accumulated value of type @Set[S]@. 
+Function @ndcps@ iteratively returns a value of type @T@ with annotated by
+@@cps[Set[T]]@ that denotes the final returned value of @ndcps@ is type @Set[T]@.
+The @f@ introduced by the @shift@ operator is the delimited continuation at the 
+call-site of @ndcps@.
+
+\begin{lstlisting}
+def nd[T,S](ts: Set[T], acc: Set[S], k: T => Set[S]): Set[S] = {
+  if (ts.isEmpty) acc
+  else {
+    val vss = k(ts.head)
+    nd(ts.tail, acc ++ vss, k)
+  }
+}
+
+def ndcps[T,S](ts: Set[T], acc: Set[S]): T @cps[Set[S]] = shift { f: (T => Set[S]) =>
+  nd(ts, acc, f)
+}
+\end{lstlisting}
+
+We add an annotation @@cps[Ans]@ on the return type of function @aeval@.
+Now for every non-deterministic choices in the abstract interpreter
+
+\begin{lstlisting}
+def aeval(e: Expr, env: Env, store: BStore, time: Time): Ans @cps[Ans] = {
+  val new_time = (e::time).take(k)
+  e match {
+    case Let(x, App(f, ae), e) =>
+      val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
+      val Clos(Lam(v, body), c_env) = ndcps[Clos, VS](closures, Set[VS]())
+      val vbaddr = allocBind(v, new_time)
+      val new_cenv = c_env + (v -> vbaddr)
+      val new_store = store.update(vbaddr, aeval(ae, env, store))
+      val bodyvss = aeval(body, new_cenv, new_store, new_time)
+      val VS(vals, time, store) = ndcps[VS, VS](bodyvss, Set[VS]())
+      val xbaddr = allocBind(x, time)
+      val new_env = env + (x -> xbaddr)
+      val new_vsstore = store.update(xbaddr, vals)
+      aeval(e, new_env, new_vsstore, time)
+
+    case ae if isAtomic(ae) =>
+      val ds = atomicEval(ae, env, store)
+      Set(VS(ds, new_time, store))
+  }
+}
+\end{lstlisting}
+
+\begin{lstlisting}
+def analyze(e: Expr) = {
+  def iter(cache: Cache): Ans = {
+    val Ans(vss, new_cache) = aeval(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List(), cache, {
+      val initConfig = Config(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List())
+      case Ans(vss, cache) => Ans(vss, cache.outUpdate(initConfig, vss))
+    })
+    if (new_cache.out == cache.out) { Ans(vss, new_cache) }
+    else { iter(Cache(new_cache.out, new_cache.out)) }
+  }
+  iter(mtCache)
+}
+\end{lstlisting}
 
@@
 the analysis if the store is widened.
-
-\iffalse
-The following code shows the implementation:
-\begin{lstlisting}
-case class P4FContAddr(tgt: Expr, tgtEnv: Env) extends KAddr
-def allocKont(tgtExpr: Expr, tgtEnv: Env): KAddr = P4FContAddr(tgtExpr, tgtEnv)
-\end{lstlisting}
-The target environment is used for successor states and extended from the closure's
-environment (namely @val new_env = env_c + (v -> baddr)@).
-\fi
 

### 2018-03-17 paper.tex
section 7

@@
 Function @ndcps@ iteratively returns a value of type @T@ with annotated by
-@@cps[Set[T]]@ that denotes the final returned value of @ndcps@ is type @Set[T]@.
+\texttt{\@ cps[Set[T]]} that denotes the final returned value of @ndcps@ is type @Set[T]@.
 The @f@ introduced by the @shift@ operator is the delimited continuation at the 
@@
 
-We add an annotation @@cps[Ans]@ on the return type of function @aeval@.
-Now for every non-deterministic choices in the abstract interpreter
+We add an annotation \texttt{ cps[Ans]} on the return type of function @aeval@.
+Now for every non-deterministic choices in the abstract interpreter, we can simply use
+@ndcps@ on them and write the program sequentially as a concrete interpreter.
+The cache is still accumulated to the next call to @ndcps@ or @aeval@ given the 
+fact that the abstract interpreter should always use the latest cache.
 
@@
 \end{lstlisting}
+
+The function @analyze@ is also changed by adding a @reset@ around the function call to @aeval@.
 

### 2018-03-17 paper.tex
Proofreading intro to 7, my last commit for the night

--- 2018-03-17 paper.tex
@@
 
-As identified before, there are two layers of continuations, one for the normal stack of 
-the analyzed language, and the other for the nondeterministic choices of closures.
+As previously identified, there are two layers of continuations: one for the normal stack of the analyzed language, and the other for the nondeterministic choices of closures.
 This corresponds to the extended continuation-passing style.
-But in our defining language Scala, the delimited controls rely on the continuation-passing
-transformation which is only one layer.
-To address this problem, base on the @nd@ operator we have before, we introduce 
-another operator @ndcps@ for handling non-deterministic choices.
+But in our defining language (Scala), the delimited controls rely on the continuation-passing transformation which is only one layer.
+To address this problem, based on the @nd@ operator we have before, we introduce another operator @ndcps@ for handling non-deterministic choices.
 
 The @ndcps@ itself is implemented with delimited controls.
-The function takes two arguments: a set of elements with type @T@ and an initial 
-accumulated value of type @Set[S]@. 
-Function @ndcps@ iteratively returns a value of type @T@ with annotated by
-@@cps[Set[T]]@ that denotes the final returned value of @ndcps@ is type @Set[T]@.
-The @f@ introduced by the @shift@ operator is the delimited continuation at the 
-call-site of @ndcps@.
+The function takes two arguments: a set of elements with type @T@, and an initial accumulated value of type @Set[S]@.
+The @ndcps@ function iteratively returns a value of type @T@ with annotated by
+\texttt{\@ cps[Set[T]]} \note{JD: Tiark, can you fix the at sign? I fixed it for the other paper, but my fix doesn't work here} that denotes the final returned value of @ndcps@ is type @Set[T]@.
+The @f@ introduced by the @shift@ operator is the delimited continuation at the call-site of @ndcps@.
 
@@
 
-We add an annotation @@cps[Ans]@ on the return type of function @aeval@.
-Now for every non-deterministic choices in the abstract interpreter
+We add an annotation \texttt{ cps[Ans]} on the return type of the @aeval@ function.
+Now, for every nondeterministic choice in the abstract interpreter, we can simply call @ndcps@ on them and write the program sequentially as a concrete interpreter.
+The cache is still accumulated by the next call to @ndcps@ or @aeval@, given the fact that the abstract interpreter should always use the latest cache.
 

### 2018-03-17 paper.tex
update

@@
                                         %% contents suppressed with 'anonymous'
-\subtitle{Filling the Gap Between Abstracting Abstract Machines and Abstracting Definitional Interpreters}        %% \subtitle is optional
+\subtitle{Filling the Gap Between Abstract Abstract Machines and Abstract Definitional Interpreters}        %% \subtitle is optional
 %\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
@@
   The transformations include linearization, fusing, disentangling, refunctionalizing,
-  and monadification (or un-CPS to direct-style). Linearization expresses
+  and un-CPS to direct-style with delimited controls. Linearization expresses
   non-deterministic choices by first-order data types, after which refunctionalization
@@
 
-\textbf{Nondeterminism Abstraction.}
+\textbf{Nondeterminism Abstractions.}
 To handle the inevitable nondeterminism, we define a generic function @nd@ as follows:
@@
 defined. In the case that @e@ is a @let@ expression,
-the overall idea is to perform a two-step, depth-first evaluation. \todo{refine}
+we invoke @nd@ for twice and perform a nesting two-step, depth-first evaluation
+over possible values of @App(f, ae)@.
 
@@
 \textbf{Call/Return Matching through Refunctionalization and Caching.}
-\todo{how refunc and caching help}
+The refunctionalized AAM we obtained has the perfect match on return-flows even though we do not have a stack model.
+The higher-order functions representing continuations already connect all the execution in order.
+In fact, the continuations (call stack) of the analyzed language is blended into the call stack of our defining language (Scala) through refunctionalization, then the call/return of the analyzed language is naturally matched.
+
+In fact, what we started is an AAM with an unbounded stack which already precisely matches the calls and returns.
+Refunctionalization makes the interpreter be a ``big-step'' semantics artifact, then the consequence is we have no places to save the context information. 
+But refunctionalization not only forces us to sequentialize the order non-deterministic evaluation through higher-order functions representing continuations, 
+but also drive us to use a new caching algorithm.
+
+Indeed, the caching algorithm plays an important role to make the analysis computable.
+
+An interesting implication of this is if we apply the caching algorithm with some necessary changes to small-step abstract machines with an unbounded stack, we are also able to establish a computable and precise call/return match.
 
@@
 But in our defining language (Scala), the delimited controls rely on the continuation-passing transformation which is only one layer.
-To address this problem, based on the @nd@ operator we have before, we introduce another operator @ndcps@ for handling non-deterministic choices.
-
-The @ndcps@ itself is implemented with delimited controls.
-The function takes two arguments: a set of elements with type @T@, and an initial accumulated value of type @Set[S]@.
+
+To address this problem, we introduce two operators @nd@ and @ndcps@ for handling non-deterministic choices.
+
+\textbf{Nondeterminism Abstractions, Again.}
+The first operator @nd@ is simplified from the previous one. 
+Now @nd@ only takes a function @f@ that applies on elements of type @T@, and does recursive call on itself over elements in set @ts@.
+
+\begin{lstlisting}
+def nd[T,S](ts: Set[T], acc: Set[S], f: T => Set[S]): Set[S] = {
+  if (ts.isEmpty) acc
+  else nd(ts.tail, acc ++ vss, f(ts.head))
+}
+\end{lstlisting}
+
+The second operator @ndcps@ is implemented with delimited controls but simply calls our first operator @nd@.
+\begin{lstlisting}
+def ndcps[T,S](ts: Set[T], acc: Set[S]): T @cps[Set[S]] = shift { f: (T => Set[S]) =>
+  nd(ts, acc, f)
+}
+\end{lstlisting}
+
+The function @ndcps@ takes two arguments: a set of elements with type @T@, and an initial accumulated value of type @Set[S]@.
 The @ndcps@ function iteratively returns a value of type @T@ with annotated by
@@
 The @f@ introduced by the @shift@ operator is the delimited continuation at the call-site of @ndcps@.
-
-\begin{lstlisting}
-def nd[T,S](ts: Set[T], acc: Set[S], k: T => Set[S]): Set[S] = {
-  if (ts.isEmpty) acc
-  else {
-    val vss = k(ts.head)
-    nd(ts.tail, acc ++ vss, k)
-  }
-}
-
-def ndcps[T,S](ts: Set[T], acc: Set[S]): T @cps[Set[S]] = shift { f: (T => Set[S]) =>
-  nd(ts, acc, f)
-}
-\end{lstlisting}
 
@@
       val new_cenv = c_env + (v -> vbaddr)
-      val new_store = store.update(vbaddr, aeval(ae, env, store))
-      val bodyvss = aeval(body, new_cenv, new_store, new_time)
-      val VS(vals, time, store) = ndcps[VS, VS](bodyvss, Set[VS]())
+      val new_cstore = store.update(vbaddr, aeval(ae, env, store))
+      val bodyvss = aeval(body, new_cenv, new_cstore, new_time)
+      val VS(vals, time, vsstore) = ndcps[VS, VS](bodyvss, Set[VS]())
       val xbaddr = allocBind(x, time)
       val new_env = env + (x -> xbaddr)
-      val new_vsstore = store.update(xbaddr, vals)
-      aeval(e, new_env, new_vsstore, time)
+      val new_store = vsstore.update(xbaddr, vals)
+      aeval(e, new_env, new_store, time)
 
@@
 The function @analyze@ is also changed by adding a @reset@ around the function call to @aeval@.
+The afterwards updating on cache becomes sequential.
 
@@
   def iter(cache: Cache): Ans = {
-    val Ans(vss, new_cache) = aeval(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List(), cache, {
+    reset {
+      val Ans(vss, anscache) = aval(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List(), cache)
       val initConfig = Config(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List())
-      case Ans(vss, cache) => Ans(vss, cache.outUpdate(initConfig, vss))
-    })
+      val new_cache = anscache.outUpdate(initConfig, vss)
       if (new_cache.out == cache.out) { Ans(vss, new_cache) }
@@
     }
+  }
   iter(mtCache)
@@
 \end{lstlisting}
+
+Now we have arrive the end of this series of transformations. Starting from an abstract abstract machine, 
+eventually we obtain an abstract interpreter with pushdown control-flows and written in purely direct-style.
 
@@
 interpreters, and we additionally demonstrate abstract interpreters written direct-style with
-delimited controls. The correspondence of delimited controls (as well as continuations) and monads
+delimited controls. 
+We use the delimited controls @shift@ and @reset@ that are implemented in Scala \cite{rompf2009implementing}.
+The correspondence of delimited controls (as well as continuations) and monads
 are well-known \cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions}.
@@
 \section{Conclusion}\label{sec:conclusion}
-\note{todo}
+
+In this functional pearl, we fill the gap between small-step abstract abstract machines
+and big-step abstract definitional interpreters by developing a series of syntactical transformations.
+Among these transformations, linearization turns worklist to another layer of continuations; 
+refunctionalization converts the first-order data types representing continuations to higher-order 
+functions; and finally un-CPS with delimited controls transform the abstract interpreter to a sequence of 
+expressions.
+
+The broader ...\todo{conc}
 

### 2018-03-17 paper.tex
fix at sign

@@
 The @ndcps@ function iteratively returns a value of type @T@ with annotated by
-\texttt{\@ cps[Set[T]]} \note{JD: Tiark, can you fix the at sign? I fixed it for the other paper, but my fix doesn't work here} that denotes the final returned value of @ndcps@ is type @Set[T]@.
+\verb|@cps[Set[T]]| that denotes the final returned value of @ndcps@ is type @Set[T]@.
 The @f@ introduced by the @shift@ operator is the delimited continuation at the call-site of @ndcps@.
 
-We add an annotation \texttt{ cps[Ans]} on the return type of the @aeval@ function.
+We add an annotation \verb|@cps[Ans]| on the return type of the @aeval@ function.
 Now, for every nondeterministic choice in the abstract interpreter, we can simply call @ndcps@ on them and write the program sequentially as a concrete interpreter.

### 2018-03-17 paper.tex
update

@@
 
-The abstract syntax in Scala is shown as follows. We assume the source
-program conforms ANF's convention, and not force it in the term structure
+The abstract syntax in Scala is shown as follows. We assume that the source
+program conforms to the ANF convention, and we do not enforce it in the term structure
 of Scala constructs.
@@
 
-In this section, we present the version that uses delimited controls.
-
-\todo{Update the code}
+In this section, we present the second version that uses delimited controls.
+
 \subsection{Un-CPS by Delimited Controls} \label{uncps}
@@
 
-\textbf{Nondeterminism Abstractions, Again.}
-The first operator @nd@ is simplified from the previous one. 
-Now @nd@ only takes a function @f@ that applies on elements of type @T@, and does recursive call on itself over elements in set @ts@.
-
-\begin{lstlisting}
-def nd[T,S](ts: Set[T], acc: Set[S], f: T => Set[S]): Set[S] = {
+\textbf{Nondeterminism Operators, Again.}
+The first operator @nd@ is simplified from the previous one, but specialized with type @Ans@ and @Cache@. 
+We may still keep the types to be generic, but the purpose of presentation clearly, exposing @Cache@ explictly
+reveals that cache should be monotonically accumulated to the rest of computation.
+Now @nd@ only takes a function @f@ that applies on type @(T, Cache)@ where value of type @T@ comes from the set,
+and @nd@ does recursive call on itself over elements in the set @ts@.
+We also move the accumulation operation (@acc++f(ts.head)@) inside of @nd@.
+
+\begin{lstlisting}
+def nd[T](ts: Set[T], acc: Ans, k: ((T, Cache)) => Ans): Ans = {
   if (ts.isEmpty) acc
-  else nd(ts.tail, acc ++ vss, f(ts.head))
-}
-\end{lstlisting}
-
-The second operator @ndcps@ is implemented with delimited controls but simply calls our first operator @nd@.
-\begin{lstlisting}
-def ndcps[T,S](ts: Set[T], acc: Set[S]): T @cps[Set[S]] = shift { f: (T => Set[S]) =>
+  else nd(ts.tail, acc ++ k(ts.head, acc.cache), k)
+}
+\end{lstlisting}
+
+The second operator @ndcps@ is implemented with delimited controls but simply calls our 
+first operator @nd@ in body.
+\begin{lstlisting}
+def ndcps[T](ts: Set[T], acc: Ans): (T, Cache) @cps[Ans] = shift { f: (((T, Cache)) => Ans) =>
   nd(ts, acc, f)
@@
 \end{lstlisting}
-
-The function @ndcps@ takes two arguments: a set of elements with type @T@, and an initial accumulated value of type @Set[S]@.
-The @ndcps@ function iteratively returns a value of type @T@ with annotated by
-\verb|@cps[Set[T]]| that denotes the final returned value of @ndcps@ is type @Set[T]@.
-The @f@ introduced by the @shift@ operator is the delimited continuation at the call-site of @ndcps@.
-
-We add an annotation \verb|@cps[Ans]| on the return type of the @aeval@ function.
-Now, for every nondeterministic choice in the abstract interpreter, we can simply call @ndcps@ on them and write the program sequentially as a concrete interpreter.
-The cache is still accumulated by the next call to @ndcps@ or @aeval@, given the fact that the abstract interpreter should always use the latest cache.
-
-\begin{lstlisting}
-def aeval(e: Expr, env: Env, store: BStore, time: Time): Ans @cps[Ans] = {
+The function @ndcps@ takes two arguments: a set of elements with type @T@, 
+and an initial accumulated value of type @Ans@.
+The @ndcps@ iteratively returns an element from the set which the element has type @T@ and annotated by
+\texttt{\@ cps[Ans]}. This annotation denotes the final returned value of @ndcps@ is type @Ans@.
+The @f@ introduced by the @shift@ operator is the delimited continuation from the call-site of @ndcps@.
+
+\textbf{}
+We add an annotation \texttt{ cps[Ans]} on the return type of the @aeval@ function.
+Now, for every nondeterministic choice in the abstract interpreter, we can simply call 
+@ndcps@ on them and write the program sequentially as a concrete interpreter.
+@ndcps@ also returns the latest cache to its left-hand side definition.
+The cache is still accumulated to the next call to @ndcps@ or @aeval@, 
+given the fact that the abstract interpreter should always use the latest cache.
+
+\begin{lstlisting}
+def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans @cps[Ans] = {
+  val config = Config(e, env, store, time)
+  if (cache.outContains(config)) Ans(cache.outGet(config), cache)
+  else {
     val new_time = (e::time).take(k)
+    val new_cache = cache.outUpdate(config, cache.inGet(config))
     e match {
@@
         val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
-      val Clos(Lam(v, body), c_env) = ndcps[Clos, VS](closures, Set[VS]())
+        val (Clos(Lam(v, body), c_env), clscache) = ndcps[Clos](closures, Ans(Set[VS](), new_cache))
         val vbaddr = allocBind(v, new_time)
@@
         val new_cstore = store.update(vbaddr, aeval(ae, env, store))
-      val bodyvss = aeval(body, new_cenv, new_cstore, new_time)
-      val VS(vals, time, vsstore) = ndcps[VS, VS](bodyvss, Set[VS]())
-      val xbaddr = allocBind(x, time)
-      val new_env = env + (x -> xbaddr)
-      val new_store = vsstore.update(xbaddr, vals)
-      aeval(e, new_env, new_store, time)
+        /* evaluates the function application App(f, ae) */
+        val Ans(bodyvss, bodycache) = aeval(body, new_cenv, new_cstore, new_time, clscache)
+        val (VS(vals, time, vsstore), vscache) = ndcps[VS](bodyvss, Ans(Set[VS](), bodycache))
+        val baddr = allocBind(x, time)
+        val new_env = env + (x -> baddr)
+        val new_store = vsstore.update(baddr, vals)
+        /* evaluates the body expression from the let */
+        val Ans(finval, fincache) = aeval(e, new_env, new_store, time, vscache)
+        Ans(finval, fincache.outUpdate(config, finval))
 
       case ae if isAtomic(ae) =>
-      val ds = atomicEval(ae, env, store)
-      Set(VS(ds, new_time, store))
+        val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
+        Ans(vs, cache.outUpdate(config, vs))
+    }
   }
@@
     reset {
-      val Ans(vss, anscache) = aval(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List(), cache)
+      val Ans(vss, anscache) = aeval(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List(), cache)
       val initConfig = Config(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List())
@@
 
-Now we have arrive the end of this series of transformations. Starting from an abstract abstract machine, 
-eventually we obtain an abstract interpreter with pushdown control-flows and written in purely direct-style.
+Now we have arrive the end of this series of transformations. Starting from an small-step abstract abstract machine, 
+eventually we obtain an big-step definitional interpreter with pushdown control-flows and written in direct-style.
 
@@
 the analysis if the store is widened.
+
+Our work establish the pushdown control-flow analysis through refunctionalization and a
+proper caching algorithm. The call/return is naturally matched in the stack of 
+the meta language.
 
@@
 
-In this functional pearl, we fill the gap between small-step abstract abstract machines
-and big-step abstract definitional interpreters by developing a series of syntactical transformations.
-Among these transformations, linearization turns worklist to another layer of continuations; 
-refunctionalization converts the first-order data types representing continuations to higher-order 
-functions; and finally un-CPS with delimited controls transform the abstract interpreter to a sequence of 
-expressions.
-
-The broader ...\todo{conc}
+In this functional pearl, we fill the gap between small-step abstract abstract machines and big-step abstract definitional interpreters by developing a series of syntactical transformations.
+Among these transformations, linearization turns a worklist into another layer of continuations; refunctionalization converts the first-order data types representing continuations to higher-order functions; and finally un-CPS with delimited control transforms the abstract interpreter into a sequence of expressions, which looks no different to a concrete interpreter.
+
+We show the correspondence not only exists between concrete semantics artifacts but also exists between abstract semantics artifacts.
+An interesting open question would be whether there also exists a correspondence of static analyses formalized in different denotational style and operational style.
 

### 2018-03-17 paper.tex

@@
 over possible values of @App(f, ae)@.
+The values of branches from on fork point will be accumulated, and then
+returned to the upper fork point.
 
@@
 
-In this section, we present the second version that uses delimited controls.
+In this section, we present the second version that uses delimited control operators.
 
@@
 
-As previously identified, there are two layers of continuations: one for the normal stack of the analyzed language, and the other for the nondeterministic choices of closures.
+As previously identified, there are two layers of continuations: one for the normal stack of the 
+analyzed language, and the other for the nondeterministic choices of closures.
 This corresponds to the extended continuation-passing style.
@@
 
-To address this problem, we introduce two operators @nd@ and @ndcps@ for handling non-deterministic choices.
+To address this problem, we reintroduce two operators @nd@ and @ndcps@ for handling non-deterministic choices
+in a way that mimics extended continuation-passing style.
 
@@
 and @nd@ does recursive call on itself over elements in the set @ts@.
-We also move the accumulation operation (@acc++f(ts.head)@) inside of @nd@.
+We also move the accumulation operation (@++@) inside of @nd@.
 
@@
 
-The second operator @ndcps@ is implemented with delimited controls but simply calls our 
+The second operator @ndcps@ is implemented with delimited control operator @shift@ but simply calls our 
 first operator @nd@ in body.
@@
 and an initial accumulated value of type @Ans@.
-The @ndcps@ iteratively returns an element from the set which the element has type @T@ and annotated by
-\texttt{\@ cps[Ans]}. This annotation denotes the final returned value of @ndcps@ is type @Ans@.
+The @ndcps@ iteratively returns an element from the set along with the latest cache 
+, and the return type is also annotated by \verb|@cps[Ans]|. 
+This annotation denotes the final returned value of @ndcps@ is type @Ans@.
 The @f@ introduced by the @shift@ operator is the delimited continuation from the call-site of @ndcps@.
 
-\textbf{}
-We add an annotation \texttt{ cps[Ans]} on the return type of the @aeval@ function.
-Now, for every nondeterministic choice in the abstract interpreter, we can simply call 
-@ndcps@ on them and write the program sequentially as a concrete interpreter.
+Now we can use these two operators to redefine @aeval@.
+We add an annotation \verb|@cps[Ans]| on the return type of the @aeval@ function.
+Then, for every nondeterministic choice in the abstract interpreter, we can simply call 
+@ndcps@ on the set of possible choices and write the program sequentially as a concrete interpreter.
 @ndcps@ also returns the latest cache to its left-hand side definition.
@@
 
-The function @analyze@ is also changed by adding a @reset@ around the function call to @aeval@.
+The function @analyze@ is also changed by adding a @reset@ operator around the function call to @aeval@.
 The afterwards updating on cache becomes sequential.
@@
 Our work establish the pushdown control-flow analysis through refunctionalization and a
-proper caching algorithm. The call/return is naturally matched in the stack of 
+proper caching algorithm. The call/return is naturally matched by the stack of 
 the meta language.

### 2018-04-15 paper/paper.tex
Update the code using delimited control

--- 2018-03-17 paper.tex
@@
 \begin{lstlisting}
+val mtEnv = Map[String, BAddr]()
+val mtStore = Store[BAddr, Storable](Map())
+val mtTime = List()
+
 def analyze(e: Expr) = {
   def iter(cache: Cache): Ans = {
-    reset {
-      val Ans(vss, anscache) = aeval(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List(), cache)
+    val Ans(vss, anscache) = reset { aeval(e, mtEnv, mtStore, mtTime, cache) }
     val initConfig = Config(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List())
@@
     else { iter(Cache(new_cache.out, new_cache.out)) }
-    }
   }

### 2018-04-15 code/src/main/scala/gnw/refunc/Test.scala
update tests

@@ -8,41 +8,28 @@ import ANFAAM._
 // TODO: reorganize the tests
 
 object RefuncTest {
-  def test_smallstep_time() {
-    val initenv = Map("f" -> BAddr("f", List()))
-    val initstore = Store[BAddr, Storable](Map(BAddr("f", List()) -> Set(Clos(Lam("x", Var("x")), Map()),
-                                                                         Clos(Lam("y", Var("y")), Map()))))
-    val ndprog = Let("a", App(Var("f"), Num(3)), Var("a"))
-    ANFAAM.k = 1
-    println(SmallStep.analyze(ndprog, initenv, initstore).mkString("\n"))
-    //println(RefuncNoCache.analyze(ndprog, initenv, initstore))
-    //println(RefuncNoCache3.analyze(ndprog, initenv, initstore))
-  }
-
+  val mtTime = List()
+  val mtEnv = Map[String, BAddr]()
+  val mtStore = Store[BAddr, Storable](Map())
+  
   def test_refunc_nd() {
     ANFAAM.k = 0
-    val ndprog = Let("a", App(Var("f"), Num(3)),
-                    Var("a"))
-                    //Num(10))
+    val ndprog = Let("a", App(Var("f"), Num(3)), Var("a"))
     val initenv = Map("f" -> BAddr("f", List()))
     val initstore = Store[BAddr, Storable](Map(BAddr("f", List()) -> Set(Clos(Lam("x", Var("x")), Map()),
                                                                          Clos(Lam("y", Num(2)), Map()),
                                                                          Clos(Lam("z", Num(1)), Map()))))
-    //println(SmallStepUBStack.analyze(ndprog,initenv, initstore))
-    //println(LinearSmallStepUBStack.analyze(ndprog,initenv, initstore))
 
     assert(SmallStepUBStack.analyze(ndprog,initenv, initstore) == 
-      LinearSmallStepUBStack.analyze(ndprog,initenv, initstore))
-
-    assert(FusedLinearSmallStepUBStack.analyze(ndprog,initenv, initstore) == 
-      LinearSmallStepUBStack.analyze(ndprog,initenv, initstore))
-
-    //assert(RefuncNoCache2.analyze(ndprog, initenv, initstore) ==
-    //         RefuncNoCache3.analyze(ndprog, initenv, initstore))
-    //assert(RefuncCPSNoCacheBF.analyze(ndprog, initenv, initstore) ==
[…]

### 2018-04-15 code/src/main/scala/gnw/refunc/DirectStyle.scala
Update trivial cases

@@ -28,6 +28,19 @@ object DirectStyleDC {
       val new_time = (e::time).take(k)
       val new_cache = cache.outUpdate(config, cache.inGet(config))
       e match {
+        case Let(x, ae, e) if isAtomic(ae) =>
+          val baddr = allocBind(x, new_time)
+          val new_env = env + (x -> baddr)
+          val new_store = store.update(baddr, aeval(ae, env, store))
+          val Ans(eval, ecache) = aval(e, new_env, new_store, new_time, new_cache)
+          Ans(eval, ecache.outUpdate(config, eval))
+
+        case Letrec(bds, body) => 
+          val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
+          val new_store = bds.foldLeft(store)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
+          val Ans(finval, fincache) = aval(body, new_env, new_store, new_time, new_cache)
+          Ans(finval, fincache.outUpdate(config, finval))
+
         case Let(x, App(f, ae), e) =>
           val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
           val (Clos(Lam(v, body), c_env), clscache) = ndcps[Clos](closures, Ans(Set[VS](), new_cache))
@@ -55,13 +68,11 @@ object DirectStyleDC {
 
   def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore) = {
     def iter(cache: Cache): Ans = {
-      reset {
-        val Ans(vss, anscache) = aval(e, env, store, mtTime, cache)
-        val initConfig = Config(e, env, store, mtTime)
-        val new_cache = anscache.outUpdate(initConfig, vss)
-        if (new_cache.out == cache.out) { Ans(vss, new_cache) }
-        else { iter(Cache(new_cache.out, new_cache.out)) }
-      }
+      val Ans(vss, anscache) = reset { aval(e, env, store, mtTime, cache) }
+      val initConfig = Config(e, env, store, mtTime)
+      val new_cache = anscache.outUpdate(initConfig, vss)
+      if (new_cache.out == cache.out) { Ans(vss, new_cache) }
+      else { iter(Cache(new_cache.out, new_cache.out)) }
     }
     iter(Cache.mtCache)
   }

### 2018-04-15 code/src/main/scala/gnw/refunc/RefuncCPS.scala

@@ -13,6 +13,8 @@ case class Cache(in: Store[Config, VS], out: Store[Config, VS]) {
   def outUpdate(config: Config, vss: Set[VS]): Cache = { Cache(in, out.update(config, vss)) }
   def outUpdate(config: Config, vs: VS): Cache = { Cache(in, out.update(config, vs)) }
   def outJoin(c: Cache): Cache = { Cache(in, out.join(c.out)) }
+
+  def outVS: Set[VS] = { out.map.values.foldLeft(Set[VS]())(_ ++ _) }
 }
 
 object Cache {
@@ -46,6 +48,21 @@ object RefuncCPS {
     val new_cache = cache.outUpdate(config, cache.inGet(config))
 
     e match {
+      case Let(x, ae, e) if isAtomic(ae) =>
+        val baddr = allocBind(x, new_time)
+        val new_env = env + (x -> baddr)
+        val new_store = store.update(baddr, aeval(ae, env, store))
+        aval(e, new_env, new_store, new_time, new_cache, { case Ans(evss, ecache) => 
+          cont(Ans(evss, ecache.outUpdate(config, evss)))
+        })
+
+      case Letrec(bds, body) => 
+        val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
+        val new_store = bds.foldLeft(store)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
+        aval(body, new_env, new_store, new_time, new_cache, { case Ans(evss, ecache) => 
+          cont(Ans(evss, ecache.outUpdate(config, evss)))
+        })
+
       case Let(x, App(f, ae), e) =>
         val closures = aeval(f, env, store).asInstanceOf[Set[Clos]]
         nd[Clos, Ans](closures, Ans(Set[VS](), new_cache), { case (clos, Ans(acc, cache), closnd) =>

### 2018-04-15 code/src/main/scala/gnw/refunc/Test.scala
clean up tests

@@ -5,51 +5,56 @@ import gnw.refunc.ast._
 
 import ANFAAM._
 
-// TODO: reorganize the tests
-
 object RefuncTest {
   val mtTime = List()
   val mtEnv = Map[String, BAddr]()
   val mtStore = Store[BAddr, Storable](Map())
+  import SmallStepUBStack._
+
+  def summarizeSS(ss: Set[State]): BStore = {
+    ss.map(_.bstore).foldLeft(mtStore)(_.join(_))
+  }
+
+  def summarizeVSS(vss: Set[VS]): BStore = {
+    vss.map(_.store).foldLeft(mtStore)(_.join(_))
+  }
   
+  def basic_test(k: Int, prog: Expr, initenv: Env, initstore: BStore) {
+    ANFAAM.k = k
+    
+    assert(SmallStepUBStack.analyze(prog,initenv, initstore) == 
+           LinearSmallStepUBStack.analyze(prog, initenv, initstore))
+
+    assert(LinearSmallStepUBStack.analyze(prog, initenv, initstore) ==
+           FusedLinearSmallStepUBStack.analyze(prog,initenv, initstore))
+
+    assert(FusedLinearSmallStepUBStack.analyze(prog, initenv, initstore) ==
+           DisLinearSmallStepUBStack.analyze(prog, initenv, initstore))
+
+    assert(SmallStepP4F.analyze(prog, initenv, initstore).map(_.bstore).foldLeft(mtStore)(_.join(_)) ==
+           summarizeVSS(RefuncCPS.analyze(prog, initenv, initstore).vss))
+
+    assert(summarizeSS(DisLinearSmallStepUBStack.analyze(prog, initenv, initstore)) ==
+           summarizeVSS(RefuncCPS.analyze(prog, initenv, initstore).vss))
+
+    assert(RefuncCPS.analyze(prog, initenv, initstore) ==
+           DirectStyleDC.analyze(prog, initenv, initstore))
+
+    assert(DirectStyleDC.analyze(prog, initenv, initstore) ==
+           DirectStyleSideEff.analyze(prog, initenv, initstore))
+  }
+
   def test_refunc_nd() {
-    ANFAAM.k = 0
     val ndprog = Let("a", App(Var("f"), Num(3)), Var("a"))
     val initenv = Map("f" -> BAddr("f", List()))
     val initstore = Store[BAddr, Storable](Map(BAddr("f", List()) -> Set(Clos(Lam("x", Var("x")), Map()),
[…]

### 2018-04-15 code/src/main/scala/gnw/refunc/DirectStyle.scala
Add supports for let/letrec

@@ -93,6 +93,13 @@ object DirectStyleSideEff {
     val new_time = (e::time).take(k)
 
     e match {
+      case Let(x, ae, e) if isAtomic(ae) =>
+        val baddr = allocBind(x, new_time)
+        val new_env = env + (x -> baddr)
+        val new_store = store.update(baddr, aeval(ae, env, store))
+        val Ans(eval, ecache) = aval(e, new_env, new_store, new_time, new_cache)
+        Ans(eval, ecache.outUpdate(config, eval))
+
       case Letrec(bds, body) =>
         val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => {
                                           accenv + (bd.x -> allocBind(bd.x, new_time))

### 2018-04-15 code/src/main/scala/gnw/refunc/DisentangledLinearSmallStepUBStack.scala

@@ -26,6 +26,17 @@ object DisLinearSmallStepUBStack {
         val new_seen = if (seen.contains(s)) seen else seen + s
         val new_time = tick(nds)
         nds match {
+          case NDState(Let(x, ae, e), env, bstore, konts, time, ndk) if isAtomic(ae) =>
+            val baddr = allocBind(x, new_time)
+            val new_env = env + (x -> baddr)
+            val new_store = bstore.update(baddr, aeval(ae, env, bstore))
+            drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), new_seen)
+
+          case NDState(Letrec(bds, body), env, bstore, konts, time, ndk) =>
+            val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
+            val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
+            drive_step(NDState(body, new_env, new_store, konts, new_time, ndk), new_seen)
+            
           case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
             val closures = atomicEval(f, env, bstore).toList.asInstanceOf[List[Clos]]
             val Clos(Lam(v, body), c_env) = closures.head

### 2018-04-15 code/src/main/scala/gnw/refunc/FusedLinearSmallStepUBStack.scala

@@ -24,6 +24,17 @@ object FusedLinearSmallStepUBStack {
       case nds =>
         val new_time = tick(nds)
         val new_ndstate = nds match {
+          case NDState(Let(x, ae, e), env, bstore, konts, time, ndk) if isAtomic(ae) =>
+            val baddr = allocBind(x, new_time)
+            val new_env = env + (x -> baddr)
+            val new_store = bstore.update(baddr, aeval(ae, env, bstore))
+            NDState(e, new_env, new_store, konts, new_time, ndk)
+
+          case NDState(Letrec(bds, body), env, bstore, konts, time, ndk) =>
+            val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
+            val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
+            NDState(body, new_env, new_store, konts, new_time, ndk)
+
           case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
             val closures = atomicEval(f, env, bstore).toList.asInstanceOf[List[Clos]]
             val Clos(Lam(v, body), c_env) = closures.head

### 2018-04-15 code/src/main/scala/gnw/refunc/LinearSmallStepUBStack.scala

@@ -21,6 +21,17 @@ object LinearSmallStepUBStack {
   def step(nds: NDState): NDState = {
     val new_time = tick(nds)
     nds match {
+      case NDState(Let(x, ae, e), env, bstore, konts, time, ndk) if isAtomic(ae) =>
+        val baddr = allocBind(x, new_time)
+        val new_env = env + (x -> baddr)
+        val new_store = bstore.update(baddr, aeval(ae, env, bstore))
+        NDState(e, new_env, new_store, konts, new_time, ndk)
+
+      case NDState(Letrec(bds, body), env, bstore, konts, time, ndk) =>
+        val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
+        val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
+        NDState(body, new_env, new_store, konts, new_time, ndk)
+
       case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
         val closures = atomicEval(f, env, bstore).toList.asInstanceOf[List[Clos]]
         val Clos(Lam(v, body), c_env) = closures.head

### 2018-05-02 paper/paper.tex
Tweaked a little bit on abstract

@@
 
-  Following the idea that in deterministic languages
-  evaluation contexts in reduction semantics are defunctionalized continuations,
+  Following the idea that in deterministic languages,
+  evaluation contexts of reduction semantics are defunctionalized continuations,
   we further show that in nondeterministic languages, evaluation contexts are

### 2018-05-02 paper/paper.tex
Fixed a typo

@@
 ~\cite{van2012systematic, van2010abstracting}.
-Given a concrete small-step abstract machine, (e.g, the CESK machine, Krivine's machine, etc.),
+Given a concrete small-step abstract machine (e.g. the CESK machine, Krivine's machine, etc.),
 by allocating continuations in the store and bounding both the value addresses and continuation

### 2018-05-21 paper/paper.tex
Update the format

@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 %% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
+\documentclass[acmsmall]{acmart}\settopmatter{}
 
@@
 %% use 'none' for review submission.
-\setcopyright{none}
+% TODO: Which one?
+%\setcopyright{none}
 %\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
+\setcopyright{acmlicensed}
 %\setcopyright{rightsretained}
@@
 
-\lstMakeShortInline[keywordstyle=,%
-                    flexiblecolumns=false,%
-                    %basewidth={0.56em, 0.52em},%
-                    mathescape=false,%
-                    basicstyle=\tt]@
-
 \begin{document}
@@
 %% Title information
-%\title{Pushdown Control Flow Analysis via Refunctionalization (Functional Pearl)}         %% [Short Title] is optional;
 \title{Refunctionalization of Abstract Abstract Machines (Functional Pearl)}         %% [Short Title] is optional;
@@
 %% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
+\author{Guannan Wei}
+%\authornote{with author1 note}          %% \authornote is optional;
                                         %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+% TODO: Register a OrcID
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
 \affiliation{
   \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
+  \department{Department of Computer Science}              %% \department is recommended
+  \institution{Purdue University}            %% \institution is required
+  \streetaddress{305 N. University Street}
+  \city{West Lafayette}
+  \state{IN}
+  \postcode{47907}
+  \country{USA}                    %% \country is recommended
+}
+\email{wei220@purdue.edu}          %% \email is recommended
+
+\author{James Decker}
+%\authornote{with author1 note}          %% \authornote is optional;
+                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+  \position{Position1}
+  \department{Department of Computer Science}              %% \department is recommended
+  \institution{Purdue University}            %% \institution is required
+  \streetaddress{305 N. University Street}
+  \city{West Lafayette}
+  \state{IN}
+  \postcode{47907}
+  \country{USA}                    %% \country is recommended
+}
+\email{decker31@purdue.edu}          %% \email is recommended
+
+\author{Tiark Rompf}
+%\authornote{with author1 note}          %% \authornote is optional;
+                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+  \position{Position1}
+  \department{Department of Computer Science}              %% \department is recommended
+  \institution{Purdue University}            %% \institution is required
+  \streetaddress{305 N. University Street}
+  \city{West Lafayette}
+  \state{IN}
+  \postcode{47907}
+  \country{USA}                    %% \country is recommended
+}
+\email{tiark@purdue.edu}          %% \email is recommended
+
+\iffalse
 %% Author with two affiliations and emails.
@@
 \email{first2.last2@inst2b.org}         %% \email is recommended
+\fi
+
+\lstMakeShortInline[keywordstyle=,%
+                    flexiblecolumns=false,%
+                    %basewidth={0.56em, 0.52em},%
+                    mathescape=false,%
+                    basicstyle=\tt]@
 
@@
 
+% TODO: Update CCS
 %% 2012 ACM Computing Classification System (CSS) concepts
@@
 \end{CCSXML}
-
 \ccsdesc[500]{Software and its engineering~General programming languages}
@@
 %% comma separated list
-%\keywords{pushdown analysis, abstract interpretation, static analysis}  %% \keywords are mandatory in final camera-ready submission
+\keywords{refunctionalization, abstract machines, abstract interpretation}  %% \keywords are mandatory in final camera-ready submission
 
@@
 first operator @nd@ in body.
+
 \begin{lstlisting}
@@
 \end{lstlisting}
+
 The function @ndcps@ takes two arguments: a set of elements with type @T@, 

### 2018-05-21 paper/paper.tex
Fix typos and mistakes

@@
 To run the program, we first use the @inject@ function (below) to construct an initial machine
-state given an expression @e@. The initial state contains an empty environment,
+state given a closed expression @e@. The initial state contains an empty environment,
 store, and stack.
@@
 to a final state by iteratively applying @step@ on the current state until a state is reached
-ion which the control is an atomic expression and the continuation stack is empty.
+which the control is an atomic expression and the continuation stack is empty.
 Naturally, we can then extract the value from the final state at last.
@@
 In the definition of @State@, the continuation store disappears, and
-component @konts@ becomes a list of frames. The other components remain unchanged.
+component @konts@ becomes a list of frames. 
+The other components remain unchanged as in the Section \ref{aam}.
 
@@
 nondeterministic, and the worklist @todo@ actually implicitly handles the nondeterminism.
-Thus, if we simply refunctionalize the frames to functions, it does not help us move toward
-abstract definitional interpreters.
+Thus, if we simply refunctionalize the frames to higher-order functions, 
+it does not help us to move toward abstract definitional interpreters.
 
@@
 For clarity of presentation in differentiating between the two continuations, we elect to call
-the first a \emph{normal continuation}, and the second as a \emph{nondeterministic
-continuation}.
+the first a \emph{normal continuation}, and the second a \emph{nondeterministic continuation}.
 
@@
 Each @NDCont@ object contains a list of closures that are possible functions to be invoked,
-a set of values that will be bind to the function's formal argument,
+a set of values that will be bound to the function's formal argument,
 and the store, time, and frames at the fork point.
@@
 However, since we have added nondeterminism into state, an empty list of frames means
-we have the reached the halt of \textit{one computation path}, and we should
+we have reached the halt of \textit{one computation path}, and we should
 determine the next state indicated by @NDCont@.
@@
 
-With this, have a single function to perform both abstract evaluation and the collection of
+With this, we have a single function to perform both abstract evaluation and the collection of
 intermediate states. Given @inject(e)@ as an initial state and an empty set as the
@@
 
-@continue@ is mainly what we have seen for the second case in
+Function @continue@ is mainly what we have seen for the second case in
 @drive_step@, except that @ndcontinue@ is called when
@@
 defined. In the case that @e@ is a @let@ expression,
-we invoke @nd@ for twice and perform a nesting two-step, depth-first evaluation
+we invoke @nd@ twice and perform a nesting two-step, depth-first evaluation
 over possible values of @App(f, ae)@.
-The values of branches from on fork point will be accumulated, and then
+The values of branches from a fork point will be accumulated, and then
 returned to the upper fork point.
@@
 The fixed-point caching algorithm performs a sound over-approximation of all possible
-concrete reachable paths and values, and also prevents non-termination when analyzing diverged programs.
+concrete reachable paths and values, and also prevents non-termination when analyzing 
+programs that may diverge.
 
@@
 But refunctionalization not only forces us to sequentialize the order non-deterministic evaluation through higher-order functions representing continuations, 
-but also drive us to use a new caching algorithm.
+but also drives us to use a new caching algorithm.
 
@@
 The first operator @nd@ is simplified from the previous one, but specialized with type @Ans@ and @Cache@. 
-We may still keep the types to be generic, but the purpose of presentation clearly, exposing @Cache@ explictly
+We may still keep the types generic, but the presentation clearly, exposing @Cache@ explictly
 reveals that cache should be monotonically accumulated to the rest of computation.
-Now @nd@ only takes a function @f@ that applies on type @(T, Cache)@ where value of type @T@ comes from the set,
-and @nd@ does recursive call on itself over elements in the set @ts@.
+Now @nd@ only takes a function @f@ that applies on type @(T, Cache)@ where the value of type @T@ comes from the set,
+and @nd@ performs recursive call on itself over elements in the set @ts@.
 We also move the accumulation operation (@++@) inside of @nd@.
@@
 and an initial accumulated value of type @Ans@.
-The @ndcps@ iteratively returns an element from the set along with the latest cache 
-, and the return type is also annotated by \verb|@cps[Ans]|. 
+The @ndcps@ iteratively returns an element from the set along with the latest cache, 
+and the return type is also annotated by \verb|@cps[Ans]|. 
 This annotation denotes the final returned value of @ndcps@ is type @Ans@.
@@
 
-Now we have arrive the end of this series of transformations. Starting from an small-step abstract abstract machine, 
+Now we have arrived the end of this series of transformations. Starting from an small-step abstract abstract machine, 
 eventually we obtain an big-step definitional interpreter with pushdown control-flows and written in direct-style.

### 2018-05-21 paper/paper.tex
tweak a little bit

@@
 definitional interpreters which automatically inherit the pushdown control-flow
-property from its defining language
-because the defined language simply uses the stack the
+property from its defining language because the defined language simply uses the stack of the
 meta-language~\cite{darais2017abstracting}. \citeauthor{darais2017abstracting}'s abstracting
@@
 $\lambda$-calculus.
-But except of simply showing the final form of abstract definitional interpreters,
+But instead of directly showing the final form of abstract definitional interpreters,
 we reveal that refunctionalization plays an important role for inheriting the
@@
 delimited controls. 
-We use the delimited controls @shift@ and @reset@ that are implemented in Scala \cite{rompf2009implementing}.
+We use the delimited control operators @shift@ and @reset@ that are implemented in Scala \cite{rompf2009implementing}.
 The correspondence of delimited controls (as well as continuations) and monads
@@
 
-We show the correspondence not only exists between concrete semantics artifacts but also exists between abstract semantics artifacts.
+We show the functional correspondence not only exists between concrete semantics artifacts but also exists between abstract semantics artifacts.
 An interesting open question would be whether there also exists a correspondence of static analyses formalized in different denotational style and operational style.

### 2018-05-30 code/src/main/scala/gnw/refunc/Test.scala
clean up & add notes

@@ -6,10 +6,8 @@ import gnw.refunc.ast._
 import ANFAAM._
 
 object RefuncTest {
-  val mtTime = List()
-  val mtEnv = Map[String, BAddr]()
-  val mtStore = Store[BAddr, Storable](Map())
   import SmallStepUBStack._
+  val bot = Set()
 
   def summarizeSS(ss: Set[State]): BStore = {
     ss.map(_.bstore).foldLeft(mtStore)(_.join(_))
@@ -134,7 +132,7 @@ object RefuncTest {
 
     //println(p4fstore) //TODO: how to get the store
     /* The final value should be empty */
-    assert(DirectStyleDC.analyze(mutrec).vss == Set())
+    assert(DirectStyleDC.analyze(mutrec).vss == bot)
     assert(RefuncCPS.analyze(mutrec) == DirectStyleDC.analyze(mutrec))
 
     /************************************************************************/
@@ -156,6 +154,11 @@ object RefuncTest {
   }
 
   def main(args: Array[String]) {
+    //TODO: README file, extended syntax 
+    //TODO: Terminating programs
+    //TODO: Non-terminating programs, expected bot
+    //TODO: Mixing terminating and non-terminating part
+    //TODO: Use scala-test
     test_non_term()
     test_refunc_nd()
   }

### 2018-05-30 paper/paper.tex
misc

@@
 
-  In this paper, we show their correspondence and how to syntactically transform small-step
-  abstract abstract machines into big-step abstract definitional
+  In this paper, we show their functional correspondence and how to syntactically 
+  transform small-step abstract abstract machines into big-step abstract definitional
   interpreters.

### 2018-05-30 src/main/scala/refunc/ANFAAMBasis.scala
reorganize

@@ -1,7 +1,7 @@
-package gnw.refunc
+package refunc
 
 import scala.util.continuations._
-import gnw.refunc.ast._
+import refunc.ast._
 
 object ANFAAM {
   def isAtomic(e: Expr): Boolean =

### 2018-05-30 src/main/scala/refunc/CESK.scala

@@ -1,6 +1,6 @@
-package gnw.refunc
+package refunc
 
-import gnw.refunc.ast._
+import refunc.ast._
 
 object CESK {
   type Addr = Int

### 2018-05-30 src/main/scala/refunc/DirectStyle.scala

@@ -1,7 +1,7 @@
-package gnw.refunc
+package refunc
 
 import scala.util.continuations._
-import gnw.refunc.ast._
+import refunc.ast._
 import ANFAAM._
 
 object DirectStyleDC {

### 2018-05-30 src/main/scala/refunc/DisentangledLinearSmallStepUBStack.scala

@@ -1,10 +1,10 @@
-package gnw.refunc
+package refunc
 
 import scala.util.continuations._
-import gnw.refunc.ast._
+import refunc.ast._
 import ANFAAM._
 
-object FusedLinearSmallStepUBStack {
+object DisLinearSmallStepUBStack {
   import SmallStepUBStack._
 
   case class NDCont(cls: List[Clos], argvs: Set[Storable], store: BStore, time: Time, frames: List[Frame])
@@ -22,19 +22,21 @@ object FusedLinearSmallStepUBStack {
     nds match {
       case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
       case nds =>
+        val s = nds.toState
+        val new_seen = if (seen.contains(s)) seen else seen + s
         val new_time = tick(nds)
-        val new_ndstate = nds match {
+        nds match {
           case NDState(Let(x, ae, e), env, bstore, konts, time, ndk) if isAtomic(ae) =>
             val baddr = allocBind(x, new_time)
             val new_env = env + (x -> baddr)
             val new_store = bstore.update(baddr, aeval(ae, env, bstore))
-            NDState(e, new_env, new_store, konts, new_time, ndk)
+            drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), new_seen)
 
           case NDState(Letrec(bds, body), env, bstore, konts, time, ndk) =>
             val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
             val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
-            NDState(body, new_env, new_store, konts, new_time, ndk)
-
+            drive_step(NDState(body, new_env, new_store, konts, new_time, ndk), new_seen)
+            
           case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
             val closures = atomicEval(f, env, bstore).toList.asInstanceOf[List[Clos]]
             val Clos(Lam(v, body), c_env) = closures.head
@@ -45,30 +47,40 @@ object FusedLinearSmallStepUBStack {
             val new_store = bstore.update(baddr, argvs)
             val new_frames = frame::konts
             val new_ndk = NDCont(closures.tail, argvs, bstore, new_time, new_frames)::ndk
[…]

### 2018-05-30 src/main/scala/refunc/FusedLinearSmallStepUBStack.scala

@@ -1,10 +1,10 @@
-package gnw.refunc
+package refunc
 
 import scala.util.continuations._
-import gnw.refunc.ast._
+import refunc.ast._
 import ANFAAM._
 
-object DisLinearSmallStepUBStack {
+object FusedLinearSmallStepUBStack {
   import SmallStepUBStack._
 
   case class NDCont(cls: List[Clos], argvs: Set[Storable], store: BStore, time: Time, frames: List[Frame])
@@ -22,21 +22,19 @@ object DisLinearSmallStepUBStack {
     nds match {
       case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
       case nds =>
-        val s = nds.toState
-        val new_seen = if (seen.contains(s)) seen else seen + s
         val new_time = tick(nds)
-        nds match {
+        val new_ndstate = nds match {
           case NDState(Let(x, ae, e), env, bstore, konts, time, ndk) if isAtomic(ae) =>
             val baddr = allocBind(x, new_time)
             val new_env = env + (x -> baddr)
             val new_store = bstore.update(baddr, aeval(ae, env, bstore))
-            drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), new_seen)
+            NDState(e, new_env, new_store, konts, new_time, ndk)
 
           case NDState(Letrec(bds, body), env, bstore, konts, time, ndk) =>
             val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
             val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
-            drive_step(NDState(body, new_env, new_store, konts, new_time, ndk), new_seen)
-            
+            NDState(body, new_env, new_store, konts, new_time, ndk)
+
           case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
             val closures = atomicEval(f, env, bstore).toList.asInstanceOf[List[Clos]]
             val Clos(Lam(v, body), c_env) = closures.head
@@ -47,40 +45,30 @@ object DisLinearSmallStepUBStack {
             val new_store = bstore.update(baddr, argvs)
             val new_frames = frame::konts
             val new_ndk = NDCont(closures.tail, argvs, bstore, new_time, new_frames)::ndk
[…]

### 2018-05-30 src/main/scala/refunc/LinearSmallStepUBStack.scala

@@ -1,7 +1,7 @@
-package gnw.refunc
+package refunc
 
 import scala.util.continuations._
-import gnw.refunc.ast._
+import refunc.ast._
 import ANFAAM._
 
 object LinearSmallStepUBStack {

### 2018-05-30 src/main/scala/refunc/RefuncCPS.scala

@@ -1,7 +1,7 @@
-package gnw.refunc
+package refunc
 
 import scala.util.continuations._
-import gnw.refunc.ast._
+import refunc.ast._
 import ANFAAM._
 
 object RefuncCPS {

### 2018-05-30 src/main/scala/refunc/RefuncCPSNoCache.scala

@@ -1,7 +1,7 @@
-package gnw.refunc
+package refunc
 
 import scala.util.continuations._
-import gnw.refunc.ast._
+import refunc.ast._
 
 import ANFAAM._
 

### 2018-05-30 src/main/scala/refunc/SmallStep.scala

@@ -1,7 +1,7 @@
-package gnw.refunc
+package refunc
 
 import scala.util.continuations._
-import gnw.refunc.ast._
+import refunc.ast._
 import ANFAAM._
 
 object SmallStep {

### 2018-05-30 src/main/scala/refunc/SmallStepUBStack.scala

@@ -1,7 +1,7 @@
-package gnw.refunc
+package refunc
 
 import scala.util.continuations._
-import gnw.refunc.ast._
+import refunc.ast._
 import ANFAAM._
 
 object SmallStepUBStack {

### 2018-05-30 src/main/scala/refunc/Test.scala

@@ -1,8 +1,7 @@
-package gnw.refunc
+package refunc
 
 import scala.util.continuations._
-import gnw.refunc.ast._
-
+import refunc.ast._
 import ANFAAM._
 
 object RefuncTest {

### 2018-05-30 src/main/scala/refunc/ast/Expr.scala

@@ -1,4 +1,4 @@
-package gnw.refunc.ast
+package refunc.ast
 
 sealed trait Expr
 sealed trait Value extends Expr

### 2018-05-30 src/test/scala/refunc/Test.scala

@@ -1,0 +1,7 @@
+package refunc
+
+import org.scalatest._
+
+class RefuncTest extends FlatSpec with Matchers {
+  println("Refunctionalization of Abstract Abstract Machines")
+}

### 2018-05-30 src/main/scala/refunc/Test.scala
update tests

@@ -5,17 +5,15 @@ import refunc.ast._
 import ANFAAM._
 
 object RefuncTest {
-  import SmallStepUBStack._
   val bot = Set()
 
-  def summarizeSS(ss: Set[State]): BStore = {
-    ss.map(_.bstore).foldLeft(mtStore)(_.join(_))
+  def summarize(ss: Set[BStore]): BStore = { ss.foldLeft(mtStore)(_.join(_)) }
+  def summarizeState(ss: Set[ANFAAM.State]): BStore = { summarize(ss.map(_.bstore)) }
+  def summarizeVSS(vss: Set[VS]): BStore = { summarize(vss.map(_.store)) }
+  def summarizeUBState(ss: Set[SmallStepUBStack.State]): BStore = { 
+    summarize(ss.map(_.bstore)) 
   }
 
-  def summarizeVSS(vss: Set[VS]): BStore = {
-    vss.map(_.store).foldLeft(mtStore)(_.join(_))
-  }
-  
   def basic_test(k: Int, prog: Expr, initenv: Env, initstore: BStore) {
     ANFAAM.k = k
     
@@ -28,10 +26,10 @@ object RefuncTest {
     assert(FusedLinearSmallStepUBStack.analyze(prog, initenv, initstore) ==
            DisLinearSmallStepUBStack.analyze(prog, initenv, initstore))
 
-    assert(SmallStepP4F.analyze(prog, initenv, initstore).map(_.bstore).foldLeft(mtStore)(_.join(_)) ==
+    assert(summarizeState(SmallStepP4F.analyze(prog, initenv, initstore)) ==
            summarizeVSS(RefuncCPS.analyze(prog, initenv, initstore).vss))
 
-    assert(summarizeSS(DisLinearSmallStepUBStack.analyze(prog, initenv, initstore)) ==
+    assert(summarizeUBState(DisLinearSmallStepUBStack.analyze(prog, initenv, initstore)) ==
            summarizeVSS(RefuncCPS.analyze(prog, initenv, initstore).vss))
 
     assert(RefuncCPS.analyze(prog, initenv, initstore) ==
@@ -47,9 +45,7 @@ object RefuncTest {
     val initstore = Store[BAddr, Storable](Map(BAddr("f", List()) -> Set(Clos(Lam("x", Var("x")), Map()),
                                                                          Clos(Lam("y", Num(2)), Map()),
                                                                          Clos(Lam("z", Num(1)), Map()))))
-    
[…]

### 2018-05-30 paper/paper.tex
Update the style of subtitles in section related work

@@
 
-\textbf{Abstract Interpretation and Control-flow Analysis.}
+\textit{Abstract Interpretation and Control-flow Analysis.}
 \citeauthor{cousot1977abstract} invented abstract interpretation as a sound approach
@@
 Modern programming languages such as Java\cite{might2010resolving} and Racket \cite{Tobin-Hochstadt:2012:HSE:2384616.2384655}
-can also be modeled by the abstracting abstract machine approach.
-
-\textbf{Defunctionalization and Refunctionalization.}
+can also be modeled by the abstracting abstract machine approach. \\
+
+\textit{Defunctionalization and Refunctionalization.}
 Defunctionalization and refunctionalization build connections between abstract
@@
 
-Fusion as one of the transformation in our paper is also studied by \citeauthor{Ohori:2007:LFF:1190216.1190241} \cite{Ohori:2007:LFF:1190216.1190241}.
-
-\textbf{Pushdown Control Flow Analysis.}
+Fusion as one of the transformation in our paper is also studied by 
+\citeauthor{Ohori:2007:LFF:1190216.1190241} \cite{Ohori:2007:LFF:1190216.1190241}. \\
+
+\textit{Pushdown Control Flow Analysis.}
 In recent years, there are significant efforts~\cite{vardoulakis2010cfa2, earl2012introspective,
@@
 proper caching algorithm. The call/return is naturally matched by the stack of 
-the meta language.
-
-\textbf{Abstracting Definitional Interpreters.}
+the meta language. \\
+
+\textit{Abstracting Definitional Interpreters.}
 Reynold's seminal paper \emph{Definitional interpreters for higher-order

### 2018-05-31 paper/paper.tex
update ref; fix format and code

@@
 Defining a language by building an interpreter for it can be traced to the very
-early days of programming languages research \cite{landin1966next, reynolds1972definitional}.
+early days of programming languages research \cite{landin1966next, Reynolds:72}.
 Nowadays, even an undergraduate student in computer science is able to build toy languages
@@
 can be used for performing sound static analysis.
-One can further instantiate different polyvariant control flow analyses by using different address allocators \cite{Gilray:2016:ACP:2951913.2951936}.
+One can further instantiate different polyvariant control-flow analyses by using different address allocators \cite{Gilray:2016:ACP:2951913.2951936}.
 
@@
 CESK machines, and also show how to instantiate
-useful $k$-call-sensitive control flow analysis.
+useful $k$-call-sensitive control-flow analysis.
 
@@
 By doing so, we recover the call stack as precise as runtimes
-(so called pushdown control flow analysis), but since
+(so called pushdown control-flow analysis), but since
 the stack is unbounded, the analysis is potentially not computable.
@@
 terminate at some fixed-point is a simple transformation of the program to cache-passing
-style. Finally, we examine how computable pushdown control flow analysis is established
+style. Finally, we examine how computable pushdown control-flow analysis is established
 through refunctionalization and caching.
@@
 \begin{lstlisting}
+val mtEnv = Map[String, BAddr](); val mtStore = Store[BAddr, Storable](Map()); val mtTime = List()
+val mtCache = Cache(Store[Config, VS](Map()), Store[Config, VS](Map()))
+
 def analyze(e: Expr) = {
   def iter(cache: Cache): Ans = {
-    val Ans(vss, new_cache) = aeval(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List(), cache, {
-      val initConfig = Config(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List())
-      case Ans(vss, cache) => Ans(vss, cache.outUpdate(initConfig, vss))
+    val Ans(vss, new_cache) = aeval(e, mtEnv, mtStore, mtTime, cache, {
+      case Ans(vss, cache) => 
+        val initConfig = Config(e, mtEnv, mtStore, mtTime) 
+        Ans(vss, cache.outUpdate(initConfig, vss))
     })
@@
 
-\subsection{Pushdown Control Flow Analysis, Revisited} \label{pdcfarevisit}
+\subsection{Pushdown Control-Flow Analysis, Revisited} \label{pdcfarevisit}
 
@@
 analysis through refunctionalization and caching. In this section,
-we revisit the pushdown control flow problem and examine what we have done to
+we revisit the pushdown control-flow problem and examine what we have done to
 overcome it.
@@
 \textbf{The Problem with Return-flows.}
-Pushdown control flow is a property in analysis that precisely models
+Pushdown control-flow is a property in analysis that precisely models
 the runtime call stack of the analyzed program.
-A pushdown control flow analysis provides an as-exact-as-runtime return-flow
-when analyzing a program, but traditional control flow analysis collapses
+A pushdown control-flow analysis provides an as-exact-as-runtime return-flow
+when analyzing a program, but traditional control-flow analysis collapses
 the state space into a finite space and thus causes imprecise stack modeling.
 
-To see how traditional control flow analysis suffers from spurious return-flows,
+To see how traditional control-flow analysis suffers from spurious return-flows,
 we can consider the following example:
@@
 This return-flow merging is inevitable, even when increasing the context-sensitivity.
-If we use the monovariant analysis (i.e., $0$-CFA), the analysis result would be such that @x@ and @y@ point to value set @{1, 2}@, because the algorithm does not distinguish that
-@z@ comes from a different call site.
+If we use the monovariant analysis (i.e., $0$-CFA), the analysis result would be such 
+that @x@ and @y@ point to value set @{1, 2}@, because the algorithm does not distinguish 
+that @z@ comes from a different call site.
 Under $1$-CFA, the algorithm is able to distinguish that variable @z@ of function
@@
 would not be polluted by @1@.
-However, variable @x@ still points to value set @{1, 2}@, because two call-sites still share the same continuation $1$-CFA is still unable to separate the return-flows.
+However, variable @x@ still points to value set @{1, 2}@, because two call-sites still 
+share the same continuation and $1$-CFA is still unable to separate the return-flows.
 
 \textbf{Call/Return Matching through Refunctionalization and Caching.}
-The refunctionalized AAM we obtained has the perfect match on return-flows even though we do not have a stack model.
-The higher-order functions representing continuations already connect all the execution in order.
-In fact, the continuations (call stack) of the analyzed language is blended into the call stack of our defining language (Scala) through refunctionalization, then the call/return of the analyzed language is naturally matched.
-
-In fact, what we started is an AAM with an unbounded stack which already precisely matches the calls and returns.
-Refunctionalization makes the interpreter be a ``big-step'' semantics artifact, then the consequence is we have no places to save the context information. 
-But refunctionalization not only forces us to sequentialize the order non-deterministic evaluation through higher-order functions representing continuations, 
+The refunctionalized AAM we obtained has the perfect match on return-flows even though 
+we do not have a stack model. The higher-order functions representing continuations 
+already connect all the execution in order. In fact, the continuations (call stack) 
+of the analyzed language is blended into the call stack of our defining language 
+(Scala) through refunctionalization, then the call/return of the analyzed language 
+is naturally matched.
+
+In fact, what we started is an AAM with an unbounded stack which already precisely 
+matches the calls and returns.
+Refunctionalization makes the interpreter be a ``big-step'' semantics artifact, 
+then the consequence is we have no places to save the context information. 
+But refunctionalization not only forces us to sequentialize the order non-deterministic 
+evaluation through higher-order functions representing continuations, 
 but also drives us to use a new caching algorithm.
@@
 
-An interesting implication of this is if we apply the caching algorithm with some necessary changes to small-step abstract machines with an unbounded stack, we are also able to establish a computable and precise call/return match.
+An interesting implication of this is if we apply the caching algorithm with some 
+necessary changes to small-step abstract machines with an unbounded stack, 
+we are also able to establish a computable and precise call/return match.
 
@@
     assignments and mutations to update the cache, and then achieve the same
-    definitional interpreter with pushdown control flows. In this case, nondeterminism can be
+    definitional interpreter with pushdown control-flows. In this case, nondeterminism can be
     easily handled via @for@ comprehension.
@@
 \begin{lstlisting}
-val mtEnv = Map[String, BAddr]()
-val mtStore = Store[BAddr, Storable](Map())
-val mtTime = List()
-
 def analyze(e: Expr) = {
@@
     val Ans(vss, anscache) = reset { aeval(e, mtEnv, mtStore, mtTime, cache) }
-    val initConfig = Config(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List())
+    val initConfig = Config(e, mtEnv, mtStore, mtTime) 
     val new_cache = anscache.outUpdate(initConfig, vss)
@@
 
-\textit{Abstract Interpretation and Control-flow Analysis.}
+\textit{Abstract Interpretation and Control-Flow Analysis.}
 \citeauthor{cousot1977abstract} invented abstract interpretation as a sound approach
 to approximate a program's runtime behavior\cite{cousot1977abstract}.
-Control flow analysis is one instance of abstract interpretation on functional programs
+Control-flow analysis is one instance of abstract interpretation on functional programs
 that can be traced to Jones \cite{jones1981flow}.
 \citeauthor{Shivers:1988:CFA:53990.54007} introduces $k$-CFA which uses $k$ recent calling contexts as program contour that differentiates values from different contexts \cite{Shivers:1988:CFA:53990.54007, Shivers:1991:SSC:115865.115884}.
-The modern formulation of control flow analysis is the abstracting abstract machines
+The modern formulation of control-flow analysis is the abstracting abstract machines
 methodology \cite{van2012systematic, van2010abstracting} which forms the starting point of this paper.
@@
 machines and interpreters.
-\citeauthor{reynolds1972definitional} first shows defunctionalization as a program transformation technique that
+\citeauthor{Reynolds:72} first shows defunctionalization as a program transformation technique that
 can be used to transform higher-order functions to first-order functions
-when specifying a programming language by an interpreter \cite{reynolds1972definitional}.
+when specifying a programming language by an interpreter \cite{Reynolds:72, Reynolds:HOSC98-revisited}.
 Refunctionalization is the reverse of this transformation.
@@
 
-\textit{Pushdown Control Flow Analysis.}
+\textit{Pushdown Control-Flow Analysis.}
 In recent years, there are significant efforts~\cite{vardoulakis2010cfa2, earl2012introspective,
@@
 complexity.
-Pushdown control flow analysis (PDCFA) is a mechanism which maintains this precision
+Pushdown control-flow analysis (PDCFA) is a mechanism which maintains this precision
 through the use of a Dyke state graph representing all possible stacks contained
@@
 \textit{Abstracting Definitional Interpreters.}
-Reynold's seminal paper \emph{Definitional interpreters for higher-order
-  programming languages} \cite{reynolds1972definitional} shows that in
+Reynold's seminal paper \emph{Definitional interpreters for higher-order programming languages} 
+\cite{Reynolds:72, Reynolds:HOSC98-revisited} shows that in
 definitional interpreters, the defined language can inherit properties from the defining language.

### 2018-05-31 src/main/scala/refunc/Main.scala
migrate to scalatest

@@ -1,0 +1,9 @@
+package refunc
+
+object RefuncAAM {
+  def main(args: Array[String]) {
+    println("Artifact of \"Refunctionalization of Abstract Abstract Machines\"")
+    println("To run test:")
+    println("  sbt test")
+  }
+}

### 2018-05-31 src/main/scala/refunc/Test.scala

@@ -1,153 +1,0 @@
-package refunc
-
-import scala.util.continuations._
-import refunc.ast._
-import ANFAAM._
-
-object RefuncTest {
-  val bot = Set()
-
-  def summarize(ss: Set[BStore]): BStore = { ss.foldLeft(mtStore)(_.join(_)) }
-  def summarizeState(ss: Set[ANFAAM.State]): BStore = { summarize(ss.map(_.bstore)) }
-  def summarizeVSS(vss: Set[VS]): BStore = { summarize(vss.map(_.store)) }
-  def summarizeUBState(ss: Set[SmallStepUBStack.State]): BStore = { 
-    summarize(ss.map(_.bstore)) 
-  }
-
-  def basic_test(k: Int, prog: Expr, initenv: Env, initstore: BStore) {
-    ANFAAM.k = k
-    
-    assert(SmallStepUBStack.analyze(prog,initenv, initstore) == 
-           LinearSmallStepUBStack.analyze(prog, initenv, initstore))
-
-    assert(LinearSmallStepUBStack.analyze(prog, initenv, initstore) ==
-           FusedLinearSmallStepUBStack.analyze(prog,initenv, initstore))
-
-    assert(FusedLinearSmallStepUBStack.analyze(prog, initenv, initstore) ==
-           DisLinearSmallStepUBStack.analyze(prog, initenv, initstore))
-
-    assert(summarizeState(SmallStepP4F.analyze(prog, initenv, initstore)) ==
-           summarizeVSS(RefuncCPS.analyze(prog, initenv, initstore).vss))
-
-    assert(summarizeUBState(DisLinearSmallStepUBStack.analyze(prog, initenv, initstore)) ==
-           summarizeVSS(RefuncCPS.analyze(prog, initenv, initstore).vss))
-
-    assert(RefuncCPS.analyze(prog, initenv, initstore) ==
-           DirectStyleDC.analyze(prog, initenv, initstore))
-
-    assert(DirectStyleDC.analyze(prog, initenv, initstore) ==
-           DirectStyleSideEff.analyze(prog, initenv, initstore))
-  }
-
-  def test_refunc_nd() {
-    val ndprog = Let("a", App(Var("f"), Num(3)), Var("a"))
-    val initenv = Map("f" -> BAddr("f", List()))
-    val initstore = Store[BAddr, Storable](Map(BAddr("f", List()) -> Set(Clos(Lam("x", Var("x")), Map()),
-                                                                         Clos(Lam("y", Num(2)), Map()),
[…]

### 2018-05-31 src/test/scala/refunc/Test.scala

@@ -1,7 +1,191 @@
 package refunc
 
 import org.scalatest._
+import refunc.ast._
+import ANFAAM._
 
-class RefuncTest extends FlatSpec with Matchers {
-  println("Refunctionalization of Abstract Abstract Machines")
+trait RefuncTest {
+  val bot = Set()
+
+  def summarize(ss: Set[BStore]): BStore = { ss.foldLeft(mtStore)(_.join(_)) }
+  def summarizeState(ss: Set[ANFAAM.State]): BStore = { summarize(ss.map(_.bstore)) }
+  def summarizeVSS(vss: Set[VS]): BStore = { summarize(vss.map(_.store)) }
+  def summarizeUBState(ss: Set[SmallStepUBStack.State]): BStore = { 
+    summarize(ss.map(_.bstore)) 
+  }
+
+  def basic_test(k: Int, prog: Expr, initenv: Env, initstore: BStore) {
+    ANFAAM.k = k
+    
+    assert(SmallStepUBStack.analyze(prog,initenv, initstore) == 
+           LinearSmallStepUBStack.analyze(prog, initenv, initstore))
+
+    assert(LinearSmallStepUBStack.analyze(prog, initenv, initstore) ==
+           FusedLinearSmallStepUBStack.analyze(prog,initenv, initstore))
+
+    assert(FusedLinearSmallStepUBStack.analyze(prog, initenv, initstore) ==
+           DisLinearSmallStepUBStack.analyze(prog, initenv, initstore))
+
+    assert(summarizeUBState(DisLinearSmallStepUBStack.analyze(prog, initenv, initstore)) ==
+           summarizeVSS(RefuncCPS.analyze(prog, initenv, initstore).vss))
+
+    assert(RefuncCPS.analyze(prog, initenv, initstore) ==
+           DirectStyleDC.analyze(prog, initenv, initstore))
+
+    assert(DirectStyleDC.analyze(prog, initenv, initstore) ==
+           DirectStyleSideEff.analyze(prog, initenv, initstore))
+
+    /* The refunctionalized abstract interpreter in CPS has the same 
+     * summarized store with small-step AAM with P4F allocator. */
+    assert(summarizeState(SmallStepP4F.analyze(prog, initenv, initstore)) ==
+           summarizeVSS(RefuncCPS.analyze(prog, initenv, initstore).vss))
+
+    /* The refunctionalized abstract interpreter using delimited continuations
[…]

### 2018-05-31 paper/paper.tex
reformat

@@
 \begin{abstract}
-  Abstracting abstract machines (AAM) is a systematic methodology
-  for constructing abstract interpreters that are derived from concrete
-  small-step abstract machines. Recent progress applies the same idea
-  on definitional interpreters, and obtains big-step
-  abstract definitional interpreters (ADI) written in monadic style.
-  Yet, the relations between small-step abstracting abstract machines and
-  big-step abstracting definitional interpreters are not well studied.
+  Abstracting abstract machines (AAM) is a systematic methodology for constructing 
+  abstract interpreters that are derived from concrete small-step abstract machines.
+  Recent progress applies the same idea on definitional interpreters, and obtains
+  big-step abstract definitional interpreters (ADI) written in monadic style.
+  Yet, the relations between small-step abstracting abstract machines and big-step 
+  abstracting definitional interpreters are not well studied.
 
@@
   sequentializes the evaluation order by higher-order functions.
-  All transformations properly handle the collecting semantics and the
-  nondeterminism of abstract interpretation.
-  %After each transformation, we also obtain an intermediate form of abstract interpreter.
-
-  Following the idea that in deterministic languages,
-  evaluation contexts of reduction semantics are defunctionalized continuations,
-  we further show that in nondeterministic languages, evaluation contexts are
-  refunctionalized to extended continuations style.
-  Remarkably, we reveal how precise call/return matching in control-flow
-  analysis is obtained by refunctionalizing a small-step abstract
-  abstract machine with proper caching.
-
-  %As well as we explain how it been lost by defunctionalizing the higher-order continuations into an first-order data type.
+  All transformations properly handle the collecting semantics and the nondeterminism 
+  of abstract interpretation.
+
+  Following the idea that in deterministic languages, evaluation contexts of reduction
+  semantics are defunctionalized continuations, we further show that in nondeterministic 
+  languages, evaluation contexts are refunctionalized to extended continuations style.
+  Remarkably, we reveal how precise call/return matching in control-flow analysis is 
+  obtained by refunctionalizing a small-step abstract abstract machine with proper caching.
 \end{abstract}
@@
 early days of programming languages research \cite{landin1966next, Reynolds:72}.
-Nowadays, even an undergraduate student in computer science is able to build toy languages
-through interpreters.
-But building a sound abstract interpreter remained an esoteric and difficult task until very recently.
-
-\citeauthor{van2012systematic} proposed the Abstracting Abstract Machines (AAM) methodology which provides a recipe
-for constructing sound abstract interpreters for higher-order functional languages from concrete abstract machines
+Nowadays, even an undergraduate student in computer science is able to build toy 
+languages through interpreters.
+But building a sound abstract interpreter remained an esoteric and difficult task 
+until very recently.
+
+\citeauthor{van2012systematic} proposed the Abstracting Abstract Machines (AAM) 
+methodology which provides a recipe for constructing sound abstract interpreters 
+for higher-order functional languages from concrete abstract machines 
 ~\cite{van2012systematic, van2010abstracting}.
-Given a concrete small-step abstract machine (e.g. the CESK machine, Krivine's machine, etc.),
-by allocating continuations in the store and bounding both the value addresses and continuation
-addresses to be finite, we obtain an abstract interpreter with a finite state space which
-can be used for performing sound static analysis.
-One can further instantiate different polyvariant control-flow analyses by using different address allocators \cite{Gilray:2016:ACP:2951913.2951936}.
+Given a concrete small-step abstract machine (e.g. the CESK machine, Krivine's 
+machine, etc.), by allocating continuations in the store and bounding both the 
+value addresses and continuation addresses to be finite, we obtain an abstract 
+interpreter with a finite state space which can be used for performing sound 
+static analysis.
+One can further instantiate different polyvariant control-flow analyses by using
+different address allocators \cite{Gilray:2016:ACP:2951913.2951936}.
 
 Applying the same idea to big-step definitional interpreters, \citeauthor{darais2017abstracting}
-built abstract definitional interpreters (ADI) that are written in monadic style \cite{darais2017abstracting}.
-One of the advantages of a monadic interpreter is that it is modular and composable.
-By changing the underlying monads, the definition of the interpreter is not modified,
-but we can recover different semantics, including the concrete semantics and various
-abstract semantics such as context-sensitivity and abstract garbage collection \cite{Sergey:2013:MAI:2491956.2491979}.
-
-Broadly speaking, abstract abstract machines and abstract definitional
-interpreters are different forms of abstract interpreters.
-They are obtained by applying a combination of abstractions to their
-concrete counterparts, abstract machines and definitional interpreters, respectively.
-An interesting question, and the subject of this paper, is how we can interderive these abstract semantics artifacts
-from the respective other one.
+built abstract definitional interpreters (ADI) that are written in monadic style 
+\cite{darais2017abstracting}. One of the advantages of a monadic interpreter is 
+that it is modular and composable. By changing the underlying monads, the definition
+of the interpreter is not modified, but we can recover different semantics, including
+the concrete semantics and various abstract semantics such as context-sensitivity and
+abstract garbage collection \cite{Sergey:2013:MAI:2491956.2491979}.
+
+Broadly speaking, abstract abstract machines and abstract definitional interpreters are
+different forms of abstract interpreters. They are obtained by applying a combination 
+of abstractions to their concrete counterparts, abstract machines and definitional 
+interpreters, respectively. An interesting question, and the subject of this paper, is 
+how we can interderive these abstract semantics artifacts from the respective other one.
 
 In the concrete world, the relations among reduction semantics, abstract machines,
-definitional interpreters, and monadic interpreters have been intensively studied by Danvy and his
-collaborators
-\cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}.
-The concrete abstract machines implement structural operational semantics in continuation-passing style,
-where the reduction contexts are defunctionalized continuations. One can derive definitional
-interpreters by refunctionalizing the evaluation contexts of abstract machines,
-and by defunctionalizing the higher-order functions, one may obtain abstract machines in the
-reverse direction.
-
-In the domain of abstract semantics artifacts, by contrast,
-the relations between small-step abstract abstract machines and big-step abstract
-definitional interpreters, as well as the question of deriving one from the other, are not
-well studied.
+definitional interpreters, and monadic interpreters have been intensively studied by
+Danvy and his collaborators \cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202,
+danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, 
+Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}.
+The concrete abstract machines implement structural operational semantics in 
+continuation-passing style, where the reduction contexts are defunctionalized continuations.
+One can derive definitional interpreters by refunctionalizing the evaluation contexts of 
+abstract machines, and by defunctionalizing the higher-order functions, one may obtain 
+abstract machines in the reverse direction.
+
+In the domain of abstract semantics artifacts, by contrast, the relations between small-step
+abstract abstract machines and big-step abstract definitional interpreters, as well as the 
+question of deriving one from the other, are not well studied.
 The fundamental difference between concrete semantics artifacts and abstract semantics
@@
 the least fixed-point eventually.
-In this functional pearl, we provide a constructive answer to the question
-of interderiving abstract semantics artifacts and relate AAM and ADI
-by presenting a series of syntactical transformations on the program from
-small-step abstract abstract machines to big-step abstract definitional interpreters.
+In this functional pearl, we provide a constructive answer to the question of interderiving
+abstract semantics artifacts and relate AAM and ADI by presenting a series of syntactical
+transformations on the program from small-step abstract abstract machines to big-step 
+abstract definitional interpreters.
 
 In addition, the abstract abstract machines with unbounded stack naturally correspond
-to abstract definitional interpreters. We show that after refunctionalizing an AAM with unbounded
-stack, and with the proper caching algorithm, the pushdown control-flow analysis can be
-obtained.
+to abstract definitional interpreters. We show that after refunctionalizing an AAM with 
+unbounded stack, and with the proper caching algorithm, the pushdown control-flow analysis 
+can be obtained.
 
@@
 
-We begin by reviewing some background
-necessary for this work in Section~\ref{background}, as well as introducing some of the basic
-code structures used throughout the paper. We then address the main contribution of this
-paper, which is the filling of the gap between small-step AAM and big-step ADI by developing a
-series of systematic transformations. Figure \ref{fig:trans} shows the transformations.
+We begin by reviewing some background necessary for this work in Section~\ref{background}, 
+as well as introducing some of the basic code structures used throughout the paper. 
+We then address the main contribution of this paper, which is the filling of the gap between
+small-step AAM and big-step ADI by developing a series of systematic transformations. 
+Figure \ref{fig:trans} shows the transformations.
 
@@
     nondeterministic choices as a first-order data type, we linearize the execution
-  of abstract abstract machines; the transition of machine states therefore becomes deterministic. Notably, this introduces another layer of controls that will also be refunctionalized later.
+    of abstract abstract machines; the transition of machine states therefore becomes 
+    deterministic. Notably, this introduces another layer of controls that will also 
+    be refunctionalized later.
 
   \item In Section~\ref{fusing}, we then present the \textbf{fusing} transformation.
-  The fusing transformation simply combines the single-step function @step@ and the driving
-  function into one, but keeps all the machine state representations.
-
-\item Section~\ref{disen} discusses \textbf{disentangling}, which disassembles the fused AAM to be several individual functions, with each function handling one data type represented by a continuation.
+    The fusing transformation simply combines the single-step function @step@ and the 
+    driving function into one, but keeps all the machine state representations.
+
+  \item Section~\ref{disen} discusses \textbf{disentangling}, which disassembles the 
+    fused AAM to be several individual functions, with each function handling one data 
+    type represented by a continuation.
 
   \item \textbf{Refunctionalization} is shown in Section~\ref{sec:refunc} which sequentializes
-  the order of abstract execution by higher-order functions.
-  For clarity, we first present the refunctionalized AAM without caching. We then adopt a different caching algorithm to guarantee the termination of abstract interpretation.
+    the order of abstract execution by higher-order functions. For clarity, we first present 
+    the refunctionalized AAM without caching. We then adopt a different caching algorithm to 
+    guarantee the termination of abstract interpretation.
     In this section, we also review pushdown control-flow analysis and examine how computable
@@
 
-\item The last transformation we show is that of transforming the refunctionalized AAM
-  to a \textbf{direct-style} interpreter (Section~\ref{directstyle}) by using delimited controls.
-
+  \item The last transformation we show is that of transforming the refunctionalized AAM to a
+    \textbf{direct-style} interpreter (Section~\ref{directstyle}) by using delimited controls.
 \end{itemize}
@@
 
-We complete this paper by discussing related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.
+We complete this paper by discussing related work in Section~\ref{sec:related}, followed by 
+concluding thoughts in Section~\ref{sec:conclusion}.
 
@@
 Traditionally, continuation-passing style (CPS) is a popular intermediate representation
-for analyzing functional programs because it exposes control transfer explicitly
-and simplifies analysis~\cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
-Here, we choose to use $\lambda$-calculus in administrative normal form (ANF) \cite{flanagan1993essence},
-which is a direct-style intermediate representation, as our target for clarity,
-but without losing simplicity and generality.
-The transformations we will show in the rest of this paper
-also work on abstract machines for plain direct-style $\lambda$-calculus languages.
-Although we only show the core calculus language, it can be easily extended
-to support recursive bindings (such as @letrec@), conditionals, primitive types, and
-operations on primitive types. These cases would be trivial to implement,
+for analyzing functional programs because it exposes control transfer explicitly and 
+simplifies analysis~\cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
+Here, we choose to use $\lambda$-calculus in administrative normal form (ANF)
+\cite{flanagan1993essence}, which is a direct-style intermediate representation, as our 
+target for clarity, but without losing simplicity and generality. The transformations 
+we will show in the rest of this paper also work on abstract machines for plain 
+direct-style $\lambda$-calculus languages. Although we only show the core calculus language,
+it can be easily extended to support recursive bindings (such as @letrec@), conditionals, 
+primitive types, and operations on primitive types. These cases would be trivial to implement,
 so we elide them in this paper.
@@
 An atomic expression $ae$ is either a variable or a literal @lambda@ term, either of which
-can be evaluated in a single step.
-We also assume that all the variable names in the program are unique.
-
-The abstract syntax in Scala is shown as follows. We assume that the source
-program conforms to the ANF convention, and we do not enforce it in the term structure
-of Scala constructs.
+can be evaluated in a single step. We also assume that all the variable names in the program
+ are unique.
+
+The abstract syntax in Scala is shown as follows. We assume that the source program conforms
+to the ANF convention, and we do not enforce it in the term structure of Scala constructs.
 
@@
 
-The CESK machine is an
-abstract machine for describing semantics of and evaluating $\lambda$-calculus \cite{felleisen1987calculus}.
-The CESK machine models program execution as state transitions in a small-step fashion.
-As its name suggests, a machine state has four components:
+The CESK machine is an abstract machine for describing semantics of and evaluating 
+$\lambda$-calculus \cite{felleisen1987calculus}. The CESK machine models program execution 
+as state transitions in a small-step fashion. As its name suggests, a machine state has 
+four components:
 1) \textit{Control} is the expression currently being evaluated.
@@
   The address space consists of numbers (0-indexed).
-In our toy language, the only category of value is a closure, i.e., a function paired with an environment.
-4) \textit{Continuation} represents the program stack. In this paper,
-we instantiate the stack as a list of frames because the ANF simplifies the evaluation context.
-For direct style $\lambda$-calculus, the continuations of CEK/CESK machines can be presented by
+  In our toy language, the only category of value is a closure, i.e., a function paired with
+  an environment.
+4) \textit{Continuation} represents the program stack. In this paper, we instantiate the 
+  stack as a list of frames because the ANF simplifies the evaluation context. For direct 
+  style $\lambda$-calculus, the continuations of CEK/CESK machines can be presented by
   variants of a data type.
@@
 
-\item In the first @case@ statement shown in the previous code,
-the control of the current state matches as a @Let@ expression, with its right-hand side a
-function application.
+\item In the first @case@ statement shown in the previous code, the control of the current 
+  state matches as a @Let@ expression, with its right-hand side a function application.
   By calling the @atomicEval@ evaluator, we obtain the closure for which the callee @f@ stands.
-The successor state's control then transfers to the @body@ expression of the closure
-with an updated environment and store. The new environment is extended
-from the closure's environment and mapped from @v@ to a fresh address @addr@.
-The new store is extended with @addr@ mapping to the value of @ae@,
-which in turn is evaluated from @atomicEval@.
-Finally, a new frame @frame@ is pushed onto the stack @k@, where
-@frame@ contains the variable name @x@ at the left-hand position of the @Let@,
-the body expression of @Let@, and the lexical environment of the body expression.
-
-\item If the state is not a @Let@ expression, then it must be
-an atomic expression, as seen in the above code. In this scenario,
-we begin by extracting the top frame of all available continuations.
-The control (i.e., an atomic expression) of the current state is the evaluated term
-that is being bound to the variable @x@ from the top frame.
+  The successor state's control then transfers to the @body@ expression of the closure with 
+  an updated environment and store. The new environment is extended from the closure's 
+  environment and mapped from @v@ to a fresh address @addr@. The new store is extended with
+  @addr@ mapping to the value of @ae@, which in turn is evaluated from @atomicEval@.
+  Finally, a new frame @frame@ is pushed onto the stack @k@, where the @frame@ contains 
+  the variable name @x@ at the left-hand position of the @Let@, the body expression of 
+  @Let@, and the lexical environment of the body expression.
+
+\item If the state is not a @Let@ expression, then it must be an atomic expression, as seen
+  in the above code. In this scenario, we begin by extracting the top frame of all available
+  continuations. The control (i.e., an atomic expression) of the current state is the 
+  evaluated term that is being bound to the variable @x@ from the top frame.
   The environment and store are updated with @x@ mapping to the closure value of @ae@.
@@
 
-To run the program, we first use the @inject@ function (below) to construct an initial machine
-state given a closed expression @e@. The initial state contains an empty environment,
+To run the program, we first use the @inject@ function (below) to construct an initial 
+machine state given a closed expression @e@. The initial state contains an empty environment,
 store, and stack.
@@
 
-The @drive@ function is then used to evaluate
-to a final state by iteratively applying @step@ on the current state until a state is reached
-which the control is an atomic expression and the continuation stack is empty.
-Naturally, we can then extract the value from the final state at last.
+The @drive@ function is then used to evaluate to a final state by iteratively applying 
+@step@ on the current state until a state is reached which the control is an atomic 
+expression and the continuation stack is empty. Naturally, we can then extract the value
+from the final state at last.
 
@@
 Abstracting abstract machines (AAM) is a systematic methodology that derives sound
-abstract interpreters for higher-order functional languages from concrete
-abstract machines \cite{van2012systematic, van2010abstracting}.
-An abstracting abstract machine implements computable abstract semantics which
-approximates the runtime behaviors of programs.
-Since the state space of concrete execution is possibly infinite,
-the key insight of AAM approach when analyzing programs is to allocate both bindings
-and continuations on the store, and then bound the addresses space to be finite.
-Since each component of state is finite,
-the abstracted machine-state space is also finite, and therefore computable.
+abstract interpreters for higher-order functional languages from concrete abstract 
+machines \cite{van2012systematic, van2010abstracting}. An abstracting abstract machine
+implements computable abstract semantics which approximates the runtime behaviors of 
+programs.
+Since the state space of concrete execution is possibly infinite, the key insight of 
+AAM approach when analyzing programs is to allocate both bindings and continuations 
+on the store, and then bound the addresses space to be finite. Since each component
+of state is finite, the abstracted machine-state space is also finite, and therefore
+computable.
 
 In this section, we derive the abstracting abstract machine from concrete
-CESK machines, and also show how to instantiate
-useful $k$-call-sensitive control-flow analysis.
+CESK machines, and also show how to instantiate useful $k$-call-sensitive 
+control-flow analysis.
 
@@
 
-The codomain of binding stores @Storable@ is the same as previously defined for CESK machines.
-The codomain of continuation stores @Cont@, on the other hand, is comprised of
-a @Frame@ object and a continuation address @KAddr@.
-To mimic the runtime call stack, @KAddr@ plays the role of
-representing the remaining stack frames.
-But since the continuation store may contain multiple continuations,
-the dereferencing of continuation addresses is also nondeterministic.
+The codomain of binding stores @Storable@ is the same as previously defined for CESK 
+machines. The codomain of continuation stores @Cont@, on the other hand, is comprised of
+a @Frame@ object and a continuation address @KAddr@. To mimic the runtime call stack, 
+@KAddr@ plays the role of representing the remaining stack frames.
+But since the continuation store may contain multiple continuations, the dereferencing 
+of continuation addresses is also nondeterministic.
 
@@
 
-As a consequence, the components of states are also changed:
-the store is divided into binding stores and continuation stores;
-the continuation stack becomes an address.
-By dereferencing this address in a continuation store, we can retrieve the
-actual transfer of controls. The definition of environment @Env@
-remains the same.
+As a consequence, the components of states are also changed: the store is divided
+into binding stores and continuation stores; the continuation stack becomes an address.
+By dereferencing this address in a continuation store, we can retrieve the actual 
+transfer of controls. The definition of environment @Env@ remains the same.
 
@@
 As previously mentioned, the space of addresses must be finite in AAM.
-Binding addresses are parameterized by variable names and the creation time of the binding,
-both of which are finite.
+Binding addresses are parameterized by variable names and the creation 
+time of the binding, both of which are finite.
 Continuation addresses @KAddr@ has two variants:
@@
 \begin{itemize}
-  \item If the current control expression is a @Let@, then
-the result of @App(f, ae)@ will be bound to variable @x@. In this case, we retrieve the set
-of closures that @f@ may represent. For each closure in the set, we perform nearly the same
-operations as in the concrete CESK machines, with an important difference: the
-continuation is allocated on the store @kstore@, so a new continuation address @new_kaddr@
-must be constructed and a new frame @Frame(x, e, env)@ paired with the current continuation
-address @kaddr@ is merged into @new_kaddr@. Finally, a list of successor states is generated.
+  \item If the current control expression is a @Let@, then the result of @App(f, ae)@ will 
+    be bound to variable @x@. In this case, we retrieve the set of closures that @f@ may 
+    represent. For each closure in the set, we perform nearly the same operations as in the
+    concrete CESK machines, with an important difference: the continuation is allocated on 
+    the store @kstore@, so a new continuation address @new_kaddr@ must be constructed and 
+    a new frame @Frame(x, e, env)@ paired with the current continuation address @kaddr@ is 
+    merged into @new_kaddr@. Finally, a list of successor states is generated.
 
@@
     state. Here, the value of @ae@ is being returned to its caller.
-In order to accomplish this, we dereference the continuation address @kaddr@ and obtain a set
-of continuations @conts@. For each continuation in the set, we construct an environment based
-on the environment @env_f@ of the frame, and bind @x@ to a newly created binding address
-@baddr@. We must also update the store with @baddr@ and the values that @ae@ represents. In
-every generated state, the control becomes the expression @e@ in the frame, and as we can
-tell from the name, the continuation address @f_kaddr@ also comes from the frame.
+    In order to accomplish this, we dereference the continuation address @kaddr@ and obtain 
+    a set of continuations @conts@. For each continuation in the set, we construct an 
+    environment based on the environment @env_f@ of the frame, and bind @x@ to a newly 
+    created binding address @baddr@. We must also update the store with @baddr@ and the 
+    values that @ae@ represents. In every generated state, the control becomes the expression 
+    @e@ in the frame, and as we can tell from the name, the continuation address @f_kaddr@ 
+    also comes from the frame.
 \end{itemize}
@@
 
-In $k$-call-sensitive analysis, a history of the last $k$ call sites is
-used as a finite program contour.
-The history is represented as a list of expressions and embedded
+In $k$-call-sensitive analysis, a history of the last $k$ call sites is used as a 
+finite program contour. The history is represented as a list of expressions and embedded
 in the allocated addresses.
 
-Before transferring to successor states, we must use the @tick@ function to
-refresh the timestamp, and then use this new timestamp for successors when allocating addresses.
+Before transferring to successor states, we must use the @tick@ function to refresh the 
+timestamp, and then use this new timestamp for successors when allocating addresses.
 The @tick@ function returns the $k$ front-most expressions given the current state and its
@@
 
-If we instantiate $k$ to be $0$, the history degenerates to an empty list,
-and we obtain a monovariant analysis (i.e., it does
-not differentiate values at different call sites). In this case, the address
-space collapses to the space of variable names.
+If we instantiate $k$ to be $0$, the history degenerates to an empty list, and we obtain
+a monovariant analysis (i.e., it does not differentiate values at different call sites).
+In this case, the address space collapses to the space of variable names.
 Note that regarding the ambiguity in $k$-CFA\cite{Gilray:2016:ACP:2951913.2951936},
@@
 
-Similar to the CESK machines, to run (analyze) a program
-we first use the @inject@ function to construct the initial state given to the program.
-Note that the initial continuation store has a built-in mapping that maps the continuation
-address for @Halt@ to an empty set of continuations.
-We also provide an empty program contour as our initial time.
+Similar to the CESK machines, to run (analyze) a program we first use the @inject@ function
+to construct the initial state given to the program. Note that the initial continuation store
+has a built-in mapping that maps the continuation address for @Halt@ to an empty set of 
+continuations. We also provide an empty program contour as our initial time.
+
 \begin{lstlisting}
@@
 
-However, in contrast to the concrete CESK machine, the
-@drive@ function performs collecting semantics instead of the valuation
-semantics. That is, for the purpose of analyzing programs, the function @drive@
-collects all the intermediate machine states as the program is abstractly executing.
-The following code shows a variant of the worklist algorithm to find the fixed-point
-of the set of states.
+However, in contrast to the concrete CESK machine, the @drive@ function performs 
+collecting semantics instead of the valuation semantics. That is, for the purpose of
+analyzing programs, the function @drive@ collects all the intermediate machine states 
+as the program is abstractly executing. The following code shows a variant of the 
+worklist algorithm to find the fixed-point of the set of states.
 Function @drive@ always applies function @step@ to the head element @hd@ of
@@
 
-Finally, a user may invoke the @analyze@ function to obtain all reachable states for a given
-program.
+Finally, a user may invoke the @analyze@ function to obtain all reachable states for 
+a given program.
 
@@
 discussion in Section~\ref{pdcfarevisit}.
-The reason we show it here is to establish an artifact with precise call/return matching
-used for the next step of transformation.
+The reason we show it here is to establish an artifact with precise call/return 
+matching used for the next step of transformation.
 
@@
 
-By atomically evaluating @f@, we obtain a set of closures, with the transition being deterministic in regards to the first element of the closure set.
+By atomically evaluating @f@, we obtain a set of closures, with the transition being 
+deterministic in regards to the first element of the closure set.
 As such, we prepare a new environment, a new store, and a new frame list only for the
@@
 If the normal continuation @konts@ is an empty list, then it means we have reached
-the end of the computation (i.e., a halt); otherwise, we should return the values of @ae@ to its caller which is contained in the top frame of stack.
+the end of the computation (i.e., a halt); otherwise, we should return the values of 
+@ae@ to its caller which is contained in the top frame of stack.
 
@@
 we obtain three mutually recursive functions: @drive_step@, which plays the same
-role as before; @continue@, which is called from @drive_step@ when encountering an atomic expression, and handles the evaluation context (@Frame@) of the analyzed
+role as before; @continue@, which is called from @drive_step@ when encountering an atomic 
+expression, and handles the evaluation context (@Frame@) of the analyzed
 language; and @ndcontinue@, which dispatches nondeterministic continuations @NDCont@,
@@
 The above code shows the skeleton of @drive_step@.
-At the end of the first case of pattern matching, we invoke @drive_step@ recursively\footnote{The code for constructing new environments, stores, and frames is elided for simplicity of presentation.}.
+At the end of the first case of pattern matching, we invoke @drive_step@ recursively 
+\footnote{The code for constructing new environments, stores, and frames is elided for 
+simplicity of presentation.}.
 The code for the second case is replaced entirely by a function call to @continue@.
@@
 the nondeterministic evaluation.
-The reason that we include a store is that there might be a case in which two paths have the same values but different accumulated side effects (e.g., memory allocations).
+The reason that we include a store is that there might be a case in which two paths
+have the same values but different accumulated side effects (e.g., memory allocations).
 Since the whole computation is nondeterministic, the type of the final result
@@
 
-If, however, we have not hit the @out@ cache, we retrieve the values from the @in@ cache, and update the @out@ cache
-with values from the @in@ cache. This retrieval from the @in@ cache may return an empty set of values if there is no such mapping for the configuration from the previous iteration.
+If, however, we have not hit the @out@ cache, we retrieve the values from the @in@ cache, 
+and update the @out@ cache with values from the @in@ cache. This retrieval from the @in@ 
+cache may return an empty set of values if there is no such mapping for the configuration 
+from the previous iteration.
 In reality, we first conservatively assume that all computations can be diverged, and
@@
 in showing the correspondence between monads and continuation-passing style as well as
-delimited controls since their very beginning era \cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions}.
+delimited controls since their very beginning era \cite{Danvy:1990:AC:91556.91622, wadler1992essence,
+danvy1992representing, moggi1991notions}.
 
@@
 
-As previously identified, there are two layers of continuations: one for the normal stack of the 
-analyzed language, and the other for the nondeterministic choices of closures.
+As previously identified, there are two layers of continuations: one for the normal stack 
+of the analyzed language, and the other for the nondeterministic choices of closures.
 This corresponds to the extended continuation-passing style.
-But in our defining language (Scala), the delimited controls rely on the continuation-passing transformation which is only one layer.
-
-To address this problem, we reintroduce two operators @nd@ and @ndcps@ for handling non-deterministic choices
-in a way that mimics extended continuation-passing style.
+But in our defining language (Scala), the delimited controls rely on the continuation-passing 
+transformation which is only one layer.
+
+To address this problem, we reintroduce two operators @nd@ and @ndcps@ for handling 
+non-deterministic choices in a way that mimics extended continuation-passing style.
 
 \textbf{Nondeterminism Operators, Again.}
-The first operator @nd@ is simplified from the previous one, but specialized with type @Ans@ and @Cache@. 
-We may still keep the types generic, but the presentation clearly, exposing @Cache@ explictly
-reveals that cache should be monotonically accumulated to the rest of computation.
-Now @nd@ only takes a function @f@ that applies on type @(T, Cache)@ where the value of type @T@ comes from the set,
-and @nd@ performs recursive call on itself over elements in the set @ts@.
-We also move the accumulation operation (@++@) inside of @nd@.
+The first operator @nd@ is simplified from the previous one, but specialized with type 
+@Ans@ and @Cache@. We may still keep the types generic, but the presentation clearly, 
+exposing @Cache@ explictly reveals that cache should be monotonically accumulated to 
+the rest of computation. Now @nd@ only takes a function @f@ that applies on type 
+@(T, Cache)@ where the value of type @T@ comes from the set, and @nd@ performs recursive 
+call on itself over elements in the set @ts@. We also move the accumulation operation 
+(@++@) inside of @nd@.
 
@@
 
-The second operator @ndcps@ is implemented with delimited control operator @shift@ but simply calls our 
-first operator @nd@ in body.
+The second operator @ndcps@ is implemented with delimited control operator 
+@shift@ but simply calls our first operator @nd@ in body.
 
@@
 This annotation denotes the final returned value of @ndcps@ is type @Ans@.
-The @f@ introduced by the @shift@ operator is the delimited continuation from the call-site of @ndcps@.
+The @f@ introduced by the @shift@ operator is the delimited continuation from the 
+call-site of @ndcps@.
 
@@
 Then, for every nondeterministic choice in the abstract interpreter, we can simply call 
-@ndcps@ on the set of possible choices and write the program sequentially as a concrete interpreter.
-@ndcps@ also returns the latest cache to its left-hand side definition.
+@ndcps@ on the set of possible choices and write the program sequentially as a concrete 
+interpreter. @ndcps@ also returns the latest cache to its left-hand side definition.
 The cache is still accumulated to the next call to @ndcps@ or @aeval@, 
@@
 
-The function @analyze@ is also changed by adding a @reset@ operator around the function call to @aeval@.
-The afterwards updating on cache becomes sequential.
+The function @analyze@ is also changed by adding a @reset@ operator around the function 
+call to @aeval@. The afterwards updating on cache becomes sequential.
 
@@
 
-Now we have arrived the end of this series of transformations. Starting from an small-step abstract abstract machine, 
-eventually we obtain an big-step definitional interpreter with pushdown control-flows and written in direct-style.
+Now we have arrived the end of this series of transformations. Starting from an 
+small-step abstract abstract machine,  eventually we obtain an big-step definitional
+interpreter with pushdown control-flows and written in direct-style.
 
@@
 that can be traced to Jones \cite{jones1981flow}.
-\citeauthor{Shivers:1988:CFA:53990.54007} introduces $k$-CFA which uses $k$ recent calling contexts as program contour that differentiates values from different contexts \cite{Shivers:1988:CFA:53990.54007, Shivers:1991:SSC:115865.115884}.
+\citeauthor{Shivers:1988:CFA:53990.54007} introduces $k$-CFA which uses $k$ recent calling
+contexts as program contour that differentiates values from different contexts 
+\cite{Shivers:1988:CFA:53990.54007, Shivers:1991:SSC:115865.115884}.
 The modern formulation of control-flow analysis is the abstracting abstract machines
-methodology \cite{van2012systematic, van2010abstracting} which forms the starting point of this paper.
-Modern programming languages such as Java\cite{might2010resolving} and Racket \cite{Tobin-Hochstadt:2012:HSE:2384616.2384655}
+methodology \cite{van2012systematic, van2010abstracting} which forms the starting point
+of this paper.
+Modern programming languages such as Java\cite{might2010resolving} and Racket 
+\cite{Tobin-Hochstadt:2012:HSE:2384616.2384655}
 can also be modeled by the abstracting abstract machine approach. \\
@@
 machines and interpreters.
-\citeauthor{Reynolds:72} first shows defunctionalization as a program transformation technique that
-can be used to transform higher-order functions to first-order functions
+\citeauthor{Reynolds:72} first shows defunctionalization as a program transformation 
+technique that can be used to transform higher-order functions to first-order functions
 when specifying a programming language by an interpreter \cite{Reynolds:72, Reynolds:HOSC98-revisited}.
@@
 
-Over the years, Danvy and his collaborators applied refunctionalization and defunctionalization to
-many different abstract machines and semantics, including the CEK machines, the CLS machines,
+Over the years, Danvy and his collaborators applied refunctionalization and defunctionalization 
+to many different abstract machines and semantics, including the CEK machines, the CLS machines,
 the SECD machines, etc.
-\cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}
-Other applications are also found to be applicable, such as Dyck word recognizer, and Dijkstra's shunting-yard algorithm \cite{Danvy:2006:RW:2171265.2171268}.
+\cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, 
+Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268,
+ danvy2009towards, biernacka2009towards}
+Other applications are also found to be applicable, such as Dyck word recognizer, and 
+Dijkstra's shunting-yard algorithm \cite{Danvy:2006:RW:2171265.2171268}.
 
@@
 Our work further shows that for nondeterministic languages, there is another layer of continuations
-that controls the nondeterministic choices. After identifying and explicitly exposing (Section~\ref{linear}) that layer
-of nondeterministic continuations, the whole program can then be refunctionalized to extended continuation-passing style.
+that controls the nondeterministic choices. After identifying and explicitly exposing 
+(Section~\ref{linear}) that layer of nondeterministic continuations, the whole program can 
+then be refunctionalized to extended continuation-passing style.
 
@@
 for small-step AAM to achieve pushdown analysis \cite{gilray2016pushdown}.
-This method is both simple to implement and computationally inexpensive and so called Pushdown for Free (P4F).
-Based on the AAM we presented in Section \ref{aam}, the only changes in the code required is
-not only keeping track of the entry-point expression of callee, but also holding the target
-environment when allocating a continuation addresses.
+This method is both simple to implement and computationally inexpensive and so called Pushdown 
+for Free (P4F). Based on the AAM we presented in Section \ref{aam}, the only changes in the 
+code required is not only keeping track of the entry-point expression of callee, but also 
+holding the target environment when allocating a continuation addresses.
 No other pieces of code would need to be modified.
@@
 delimited controls. 
-We use the delimited control operators @shift@ and @reset@ that are implemented in Scala \cite{rompf2009implementing}.
-The correspondence of delimited controls (as well as continuations) and monads
-are well-known \cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions}.
+We use the delimited control operators @shift@ and @reset@ that are implemented in Scala 
+\cite{rompf2009implementing}. The correspondence of delimited controls (as well as continuations) 
+and monads are well-known \cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions}.
 
@@
 
-In this functional pearl, we fill the gap between small-step abstract abstract machines and big-step abstract definitional interpreters by developing a series of syntactical transformations.
-Among these transformations, linearization turns a worklist into another layer of continuations; refunctionalization converts the first-order data types representing continuations to higher-order functions; and finally un-CPS with delimited control transforms the abstract interpreter into a sequence of expressions, which looks no different to a concrete interpreter.
-
-We show the functional correspondence not only exists between concrete semantics artifacts but also exists between abstract semantics artifacts.
-An interesting open question would be whether there also exists a correspondence of static analyses formalized in different denotational style and operational style.
+In this functional pearl, we fill the gap between small-step abstract abstract machines 
+and big-step abstract definitional interpreters by developing a series of syntactical 
+transformations. Among these transformations, linearization turns a worklist into another
+layer of continuations; refunctionalization converts the first-order data types representing
+continuations to higher-order functions; and finally un-CPS with delimited control transforms
+the abstract interpreter into a sequence of expressions, which looks no different to a
+concrete interpreter.
+
+We show the functional correspondence not only exists between concrete semantics artifacts
+but also exists between abstract semantics artifacts. An interesting open question would
+be whether there also exists a correspondence of static analyses formalized in different 
+denotational style and operational style.
 

### 2018-06-01 src/main/scala/refunc/ANFAAMBasis.scala
update

@@ -78,8 +78,9 @@ object ANFAAM {
     def outContains(config: Config): Boolean = out.contains(config)
     def outUpdate(config: Config, vss: Set[VS]): Cache = { Cache(in, out.update(config, vss)) }
     def outUpdate(config: Config, vs: VS): Cache = { Cache(in, out.update(config, vs)) }
-    def outJoin(c: Cache): Cache = { Cache(in, out.join(c.out)) }
-
+    def join(c: Cache): Cache = { Cache(in.join(c.in), out.join(c.out)) }
+    
+    def inVS: Set[VS] = { in.map.values.foldLeft(Set[VS]())(_ ++ _) }
     def outVS: Set[VS] = { out.map.values.foldLeft(Set[VS]())(_ ++ _) }
   }
 
@@ -89,7 +90,7 @@ object ANFAAM {
 
   case class Ans(vss: Set[VS], cache: Cache) {
     def ++(ans: Ans): Ans = {
-      Ans(vss ++ ans.vss, ans.cache.outJoin(cache))
+      Ans(vss ++ ans.vss, cache.join(ans.cache))
     }
   }
 

### 2018-06-01 src/main/scala/refunc/DirectStyle.scala

@@ -5,7 +5,6 @@ import refunc.ast._
 import ANFAAM._
 
 object DirectStyleDC {
-  /* Depth first evaluation */
   import SmallStepUBStack._
   
   type Cont = Ans => Ans
@@ -30,33 +29,34 @@ object DirectStyleDC {
         case Let(x, ae, e) if isAtomic(ae) =>
           val baddr = allocBind(x, new_time)
           val new_env = env + (x -> baddr)
-          val new_store = store.update(baddr, aeval(ae, env, store))
-          val Ans(eval, ecache) = aval(e, new_env, new_store, new_time, new_cache)
-          Ans(eval, ecache.outUpdate(config, eval))
+          val new_store = store.update(baddr, atomicEval(ae, env, store))
+          val Ans(evss, ecache) = aval(e, new_env, new_store, new_time, new_cache)
+          Ans(evss, ecache.outUpdate(config, evss))
 
         case Letrec(bds, body) => 
           val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
-          val new_store = bds.foldLeft(store)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
-          val Ans(finval, fincache) = aval(body, new_env, new_store, new_time, new_cache)
-          Ans(finval, fincache.outUpdate(config, finval))
+          val new_store = bds.foldLeft(store)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time),
+                                                                                         atomicEval(bd.e, new_env, accbst)) })
+          val Ans(bdss, bdcache) = aval(body, new_env, new_store, new_time, new_cache)
+          Ans(bdss, bdcache.outUpdate(config, bdss))
 
         case Let(x, App(f, ae), e) =>
           val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
           val (Clos(Lam(v, body), c_env), clscache) = ndcps[Clos](closures, Ans(Set[VS](), new_cache))
           val vbaddr = allocBind(v, new_time)
           val new_cenv = c_env + (v -> vbaddr)
-          val new_cstore = store.update(vbaddr, aeval(ae, env, store))
-          val Ans(bodyvss, bodycache) = aval(body, new_cenv, new_cstore, new_time, clscache)
[…]

### 2018-06-01 src/main/scala/refunc/DirectStyleNoCache.scala

@@ -5,6 +5,8 @@ import refunc.ast._
 
 import ANFAAM._
 
+/* Exprimental implementations */
+
 object DirectStyleDCNoCache2 {
   /* Depth first evaluation */
   import SmallStepUBStack._

### 2018-06-01 src/main/scala/refunc/Main.scala

@@ -1,9 +1,28 @@
 package refunc
 
+import ast._
+import ANFAAM._
+
 object RefuncAAM {
   def main(args: Array[String]) {
     println("Artifact of \"Refunctionalization of Abstract Abstract Machines\"")
     println("To run test:")
     println("  sbt test")
+
+  val prog = Let("a", App(Var("f"), Num(3)), Var("a"))
+  val initenv = Map("f" -> BAddr("f", List()))
+  val initstore = Store[BAddr, Storable](Map(BAddr("f", List()) ->
+                                               Set(Clos(Lam("x", Var("x")), Map()),
+                                                   Clos(Lam("x1", Let("t1", App(Var("f"), Var("x1")), Var("t1"))),
+                                                        Map("f" -> BAddr("f", List()))),
+                                                   Clos(Lam("y", Num(2)), Map()),
+                                                   Clos(Lam("z", Num(1)), Map()))))
+  ANFAAM.k = 0
+      println(DirectStyleDC.analyze(prog, initenv, initstore).cache.in.map.size)
+      println(RefuncCPS.analyze(prog, initenv, initstore).cache.in.map.size)
+        assert(RefuncCPS.analyze(prog, initenv, initstore).cache.in.map.size == DirectStyleDC.analyze(prog, initenv, initstore).cache.in.map.size)
+        assert(RefuncCPS.analyze(prog, initenv, initstore).cache.out.map.size == DirectStyleDC.analyze(prog, initenv, initstore).cache.out.map.size)
+        assert(RefuncCPS.analyze(prog, initenv, initstore).cache.inVS == DirectStyleDC.analyze(prog, initenv, initstore).cache.inVS)
+        assert(RefuncCPS.analyze(prog, initenv, initstore).cache.outVS == DirectStyleDC.analyze(prog, initenv, initstore).cache.outVS)
   }
 }

### 2018-06-01 src/main/scala/refunc/RefuncCPS.scala

@@ -5,7 +5,6 @@ import refunc.ast._
 import ANFAAM._
 
 object RefuncCPS {
-  /* Depth First Evaluation */
   import SmallStepUBStack._
 
   type Cont = Ans => Ans
@@ -17,52 +16,57 @@ object RefuncCPS {
   
   def aval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, cont: Cont): Ans = {
     val config = Config(e, env, store, time)
-    if (cache.outContains(config)) {
-      return cont(Ans(cache.outGet(config), cache))
-    }
+    if (cache.outContains(config)) cont(Ans(cache.outGet(config), cache))
+    else {
+      val new_time = (e::time).take(k)
+      val new_cache = cache.outUpdate(config, cache.inGet(config))
 
-    val new_time = (e::time).take(k)
-    val new_cache = cache.outUpdate(config, cache.inGet(config))
-
-    e match {
-      case Let(x, ae, e) if isAtomic(ae) =>
-        val baddr = allocBind(x, new_time)
-        val new_env = env + (x -> baddr)
-        val new_store = store.update(baddr, aeval(ae, env, store))
-        aval(e, new_env, new_store, new_time, new_cache, { case Ans(evss, ecache) => 
-          cont(Ans(evss, ecache.outUpdate(config, evss)))
-        })
-
-      case Letrec(bds, body) => 
-        val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
-        val new_store = bds.foldLeft(store)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
-        aval(body, new_env, new_store, new_time, new_cache, { case Ans(evss, ecache) => 
-          cont(Ans(evss, ecache.outUpdate(config, evss)))
-        })
-
-      case Let(x, App(f, ae), e) =>
-        val closures = aeval(f, env, store).asInstanceOf[Set[Clos]]
-        nd[Clos, Ans](closures, Ans(Set[VS](), new_cache), { case (clos, Ans(acc, cache), closnd) =>
-          val Clos(Lam(v, body), c_env) = clos
-          val baddr = allocBind(v, new_time)
-          val new_env = c_env + (v -> baddr)
-          val new_store = store.update(baddr, aeval(ae, env, store))
-          aval(body, new_env, new_store, new_time, cache, { case Ans(bodyvss, bodycache) =>
[…]

### 2018-06-01 src/main/scala/refunc/RefuncCPSNoCache.scala

@@ -6,7 +6,6 @@ import refunc.ast._
 import ANFAAM._
 
 object RefuncCPSNoCache {
-  /* Depth first evaluation */
   import SmallStepUBStack._
   
   type Ans = Set[VS]
@@ -51,13 +50,10 @@ object RefuncCPSNoCache {
     aval(e, env, store, mtTime, (vss => vss))
 }
 
+
+
+/* Experimental implementation with breath first evaluation */
 object RefuncCPSNoCacheBF {
-  /* Implements non-determinism by fold-with-continuations.
-   * The strategy it uses is to first try all closures, collect
-   * values of App(f, ae) - result_vss, and for each of them,
-   * go into let's body `e` seperately.
-   * Breath first evaluation.
-   */
   import SmallStepUBStack._
   
   type Cont = Set[VS] => Set[VS]

### 2018-06-01 src/test/scala/refunc/Test.scala

@@ -4,9 +4,23 @@ import org.scalatest._
 import refunc.ast._
 import ANFAAM._
 
-trait RefuncTest {
+/**
+ * Description:
+ * 
+ * class RefuncTest provides some auxiliary functions;
+ * class TDProgTest contains test cases that programs are terminated and deterministic;
+ * class TNDProgTest contains test cases that programs are terminated but non-deterministic;
+ * class NDProgTest contains test cases that programs are non-terminated;
+ * class NTTNDProgTest contains test cases that programs are non-deterministic, but some branches
+ * are terminated, while some other branches are not.
+ *
+ * For convenience, we test non-deterministic cases through passing an initial store that maps
+ * an address to multiple target closures.
+ */
+
+trait RefuncTest extends FunSuite {
+  final val K = 3 /* K-call-site context sensitivity */
   val bot = Set()
-  val K = 3 /* K-call-site context sensitivity */
 
   def summarize(ss: Set[BStore]): BStore = { ss.foldLeft(mtStore)(_.join(_)) }
   def summarizeState(ss: Set[ANFAAM.State]): BStore = { summarize(ss.map(_.bstore)) }
@@ -14,16 +28,14 @@ trait RefuncTest {
   def summarizeUBState(ss: Set[SmallStepUBStack.State]): BStore = {
     summarize(ss.map(_.bstore)) 
   }
-}
+  
+  val description = "None"
 
-/** 
- *  Test terminated non-deterministic programs.
- */
-class NDProgTest extends FunSuite with RefuncTest {
+  /* Assertions for TDProgTest, TNDProgTest and NTTNDProgTest. */
   def runTest(id: Int, prog: Expr, initenv: Env, initstore: BStore) {
     for (k <- 0 to K) {
       ANFAAM.k = k
-      test(s"Testing terminated non-deterministic program $id under k=$k") {
+      test(s"Testing $description case $id under k=$k") {
         /* The result analyzed by the small-step AAM with unbounded stack (SmallStepUBStack)
          * should have the same result analyzed by its linearized counterpart 
          * (LinearSmallStepUBStack). */
@@ -89,6 +101,56 @@ class NDProgTest extends FunSuite with RefuncTest {
       }
     }
   }
+  
+  /* Common expressions */
+  val id = Lam("t", Var("t"))
+}
+
+/** 
+ *  Test terminated deterministic programs.
+ */
[…]

### 2018-06-01 src/test/scala/refunc/Test.scala
update test cases

@@ -112,22 +112,8 @@ trait RefuncTest extends FunSuite {
 class TDProgTest extends RefuncTest {
   override val description = "terminated deterministic"
 
-  /** 
-   * test case 1
-   * exp: 
-   * init env: ∅
-   * init store: ∅ 
-   */
-
-  /** 
-   * test case 2
-   * exp:
-   * init env: ∅
-   * init store: ∅ 
-   */
-
   /**
-   * test case 3
+   * test case 1
    * exp: (let ([id (lambda (t) t)])
    *        (let ([x (id 1)])
    *          (let ([m (id 4)])
@@ -137,13 +123,13 @@ class TDProgTest extends RefuncTest {
    * init env: ∅
    * init store: ∅ 
    */
-  val exp3 = Let("id", id, 
+  val exp1 = Let("id", id, 
              Let("x", App(id, Num(1)), 
              Let("m", App(id, Num(4)), 
              Let("y", App(id, Num(2)), 
              Let("z", App(id, Num(3)), 
                Var("x"))))))
-  runTest(3, exp3, mtEnv, mtStore)
+  runTest(1, exp1, mtEnv, mtStore)
 }
 
 /** 
@@ -234,7 +220,7 @@ class NDProgTest extends RefuncTest {
          *
          * Other intermediate forms of abstract interpreters in our transformations (LinearSmallStepUBStack, 
          * FusedLinearSmallStepUBStack and DisLinearSmallStepUBStack) lack proper caching mechanisms, 
-         * running them on non-terminated programs would cause the abstract interpreter to diverge.
+         * running them with non-terminated programs would cause the abstract interpreter to diverge.
          * So we don't test them here. */
         assert(RefuncCPS.analyze(prog, initenv, initstore).vss == bot)
         assert(DirectStyleDC.analyze(prog, initenv, initstore).vss == bot)
@@ -277,6 +263,42 @@ class NDProgTest extends RefuncTest {
                     Let("res", App(Var("f2"), Num(1)),
                         Var("res")))
   runTest(2, exp2, mtEnv, mtStore)
+
+  /**
+   * test case 3
+   * exp: (let ([a (f 3)])
+   *        (let ([b (g 4)])
+   *          b))
+   * init env: exp3_env1
+   * init store: exp3_store1
+   */
+  val exp3 = Let("a", App(Var("f"), Num(3)), 
+                Let("b", App(Var("g"), Num(4)),
+                  Var("b")))
[…]

### 2018-06-01 src/main/scala/refunc/RefuncCPS.scala
fix typos

@@ -86,7 +86,7 @@ object RefuncCPS {
 
 
 
-/* Experimental implementation with breath first evaluation */
+/* Experimental implementation with breadth first evaluation */
 object RefuncCPSBF {
   import SmallStepUBStack._
   

### 2018-06-01 src/main/scala/refunc/RefuncCPSNoCache.scala

@@ -52,7 +52,7 @@ object RefuncCPSNoCache {
 
 
 
-/* Experimental implementation with breath first evaluation */
+/* Experimental implementation with breadth first evaluation */
 object RefuncCPSNoCacheBF {
   import SmallStepUBStack._
   

### 2018-06-01 src/main/scala/refunc/Main.scala
remove useless code

@@ -8,21 +8,5 @@ object RefuncAAM {
     println("Artifact of \"Refunctionalization of Abstract Abstract Machines\"")
     println("To run test:")
     println("  sbt test")
-
-  val prog = Let("a", App(Var("f"), Num(3)), Var("a"))
-  val initenv = Map("f" -> BAddr("f", List()))
-  val initstore = Store[BAddr, Storable](Map(BAddr("f", List()) ->
-                                               Set(Clos(Lam("x", Var("x")), Map()),
-                                                   Clos(Lam("x1", Let("t1", App(Var("f"), Var("x1")), Var("t1"))),
-                                                        Map("f" -> BAddr("f", List()))),
-                                                   Clos(Lam("y", Num(2)), Map()),
-                                                   Clos(Lam("z", Num(1)), Map()))))
-  ANFAAM.k = 0
-      println(DirectStyleDC.analyze(prog, initenv, initstore).cache.in.map.size)
-      println(RefuncCPS.analyze(prog, initenv, initstore).cache.in.map.size)
-        assert(RefuncCPS.analyze(prog, initenv, initstore).cache.in.map.size == DirectStyleDC.analyze(prog, initenv, initstore).cache.in.map.size)
-        assert(RefuncCPS.analyze(prog, initenv, initstore).cache.out.map.size == DirectStyleDC.analyze(prog, initenv, initstore).cache.out.map.size)
-        assert(RefuncCPS.analyze(prog, initenv, initstore).cache.inVS == DirectStyleDC.analyze(prog, initenv, initstore).cache.inVS)
-        assert(RefuncCPS.analyze(prog, initenv, initstore).cache.outVS == DirectStyleDC.analyze(prog, initenv, initstore).cache.outVS)
   }
 }

### 2018-06-01 src/main/scala/refunc/ANFAAMBasis.scala
refactor; eliminate warnings

@@ -63,6 +63,7 @@ object ANFAAM {
     case Num(i) => Set(NumV(i))
     case Var(x) => bstore(env(x)).toSet
     case lam@Lam(x, body) => Set(Clos(lam, env))
+    case _ => throw new NotImplementedError(e.toString)
   }
 
   val atomicEval = aeval _

### 2018-06-01 src/main/scala/refunc/CESK.scala

@@ -2,7 +2,7 @@ package refunc
 
 import refunc.ast._
 
-object CESK {
+object Concrete {
   type Addr = Int
   type Env = Map[String, Addr]
   type Store = Map[Addr, Storable]
@@ -11,17 +11,22 @@ object CESK {
   case class Clos(v: Lam, env: Env) extends Storable
   case class Num(i: Int) extends Storable
 
-  case class Frame(x: String, e: Expr, env: Env)
-  type Kont = List[Frame]
-
-  case class State(e: Expr, env: Env, store: Store, k: Kont)
-
   def atomicEval(e: Expr, env: Env, store: Store): Storable = e match {
     case Var(x) => store(env(x))
     case lam@Lam(x, body) => Clos(lam, env)
+    case _ => throw new NotImplementedError(e.toString)
   }
+
   def alloc(store: Store): Addr = store.keys.size + 1
+
   def isAtomic(e: Expr): Boolean = e.isInstanceOf[Var] || e.isInstanceOf[Lam]
+}
+
+object CESK {
+  import Concrete._
+  case class Frame(x: String, e: Expr, env: Env)
+  type Kont = List[Frame]
+  case class State(e: Expr, env: Env, store: Store, k: Kont)
 
   def step(s: State): State = s match {
     case State(Let(x, ae, e), env, store, k) if isAtomic(ae) =>
@@ -50,16 +55,11 @@ object CESK {
 }
 
 object RefuncCESK {
-  import CESK._
+  import Concrete._
 
   case class VS(v: Storable, store: Store)
   type Ans = VS
   type Cont = Ans => Ans
-
-  def atomicEval(e: Expr, env: Env, store: Store): Storable = e match {
-    case Var(x) => store(env(x))
-    case lam@Lam(x, body) => Clos(lam, env)
-  }
 
   def eval(e: Expr, env: Env, store: Store, k: Cont): VS = e match {
     case Let(x, ae, e) if isAtomic(ae) =>

### 2018-06-01 src/main/scala/refunc/DisentangledLinearSmallStepUB.scala

@@ -81,6 +81,7 @@ object DisLinearSmallStepUBStack {
         drive_step(NDState(body, new_env, new_store, frames, time, 
                            NDCont(cls.tail, argvs, bstore, time, frames)::ndk),
                    seen)
+      case _ => throw new RuntimeException("Invalid NDCont")
     }
   }
 

### 2018-06-01 src/main/scala/refunc/FusedLinearSmallStepUB.scala

@@ -57,7 +57,9 @@ object FusedLinearSmallStepUBStack {
                   val baddr = allocBind(v, time)
                   val new_env = c_env + (v -> baddr)
                   val new_store = bstore.update(baddr, argvs)
-                  NDState(body, new_env, new_store, frames, time, NDCont(cls.tail, argvs, bstore, time, frames)::ndk)
+                  NDState(body, new_env, new_store, frames, time, 
+                          NDCont(cls.tail, argvs, bstore, time, frames)::ndk)
+                case _ => throw new RuntimeException("Invalid NDCont")
               }
               case Frame(x, e, f_env)::konts =>
                 val baddr = allocBind(x, new_time)

### 2018-06-01 src/main/scala/refunc/LinearSmallStepUB.scala

@@ -54,7 +54,9 @@ object LinearSmallStepUBStack {
               val baddr = allocBind(v, time)
               val new_env = c_env + (v -> baddr)
               val new_store = bstore.update(baddr, argvs)
-              NDState(body, new_env, new_store, frames, time, NDCont(cls.tail, argvs, bstore, time, frames)::ndk)
+              NDState(body, new_env, new_store, frames, time, 
+                      NDCont(cls.tail, argvs, bstore, time, frames)::ndk)
+            case _ => throw new RuntimeException("Invalid NDCont")
           }
           case Frame(x, e, f_env)::konts =>
             val baddr = allocBind(x, new_time)

### 2018-06-01 src/main/scala/refunc/DirectStyle.scala
refactor

@@ -34,9 +34,10 @@ object DirectStyleDC {
           Ans(evss, ecache.outUpdate(config, evss))
 
         case Letrec(bds, body) => 
-          val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
-          val new_store = bds.foldLeft(store)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time),
-                                                                                         atomicEval(bd.e, new_env, accbst)) })
+          val new_env = bds.foldLeft(env)((accenv: Env, bd: B) =>
+            accenv + (bd.x -> allocBind(bd.x, new_time)))
+          val new_store = bds.foldLeft(store)((accbst: BStore, bd: B) =>
+            accbst.update(allocBind(bd.x, new_time), atomicEval(bd.e, new_env, accbst)))
           val Ans(bdss, bdcache) = aval(body, new_env, new_store, new_time, new_cache)
           Ans(bdss, bdcache.outUpdate(config, bdss))
 
@@ -45,13 +46,13 @@ object DirectStyleDC {
           val (Clos(Lam(v, body), c_env), clscache) = ndcps[Clos](closures, Ans(Set[VS](), new_cache))
           val vbaddr = allocBind(v, new_time)
           val new_cenv = c_env + (v -> vbaddr)
-          val new_cstore = store.update(vbaddr, atomicEval(ae, env, store))
-          val Ans(bdvss, bdcache) = aval(body, new_cenv, new_cstore, new_time, clscache)
+          val new_store = store.update(vbaddr, atomicEval(ae, env, store))
+          val Ans(bdvss, bdcache) = aval(body, new_cenv, new_store, new_time, clscache)
           val (VS(vals, time, vsstore), vscache) = ndcps[VS](bdvss, Ans(Set[VS](), bdcache))
           val baddr = allocBind(x, time)
           val new_env = env + (x -> baddr)
-          val new_store = vsstore.update(baddr, vals)
-          val Ans(evss, ecache) = aval(e, new_env, new_store, time, vscache)
+          val new_vsstore = vsstore.update(baddr, vals)
+          val Ans(evss, ecache) = aval(e, new_env, new_vsstore, time, vscache)
           Ans(evss, ecache.outUpdate(config, evss))
 
         case ae if isAtomic(ae) =>
[…]

### 2018-06-01 src/main/scala/refunc/DirectStyleNoCache.scala

@@ -34,13 +34,13 @@ object DirectStyleDCNoCache2 {
         val Clos(Lam(v, body), c_env) = ndcps[Clos, VS](closures, Set[VS]())
         val vbaddr = allocBind(v, new_time)
         val new_cenv = c_env + (v -> vbaddr)
-        val new_cstore = store.update(vbaddr, aeval(ae, env, store))
-        val bodyvss = aval(body, new_cenv, new_cstore, new_time)
-        val VS(vals, time, vsstore) = ndcps[VS, VS](bodyvss, Set[VS]())
+        val new_store = store.update(vbaddr, aeval(ae, env, store))
+        val bdvss = aval(body, new_cenv, new_store, new_time)
+        val VS(vals, time, vsstore) = ndcps[VS, VS](bdvss, Set[VS]())
         val baddr = allocBind(x, time)
         val new_env = env + (x -> baddr)
-        val new_store = vsstore.update(baddr, vals)
-        aval(e, new_env, new_store, time)
+        val new_vsstore = vsstore.update(baddr, vals)
+        aval(e, new_env, new_vsstore, time)
   
       case ae if isAtomic(ae) =>
         val ds = atomicEval(ae, env, store)
@@ -51,6 +51,8 @@ object DirectStyleDCNoCache2 {
   def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore) =
     reset { aval(e, env, store, mtTime) }
 }
+
+/* Exprimental implementations */
 
 object DirectStyleDCNoCache {
   /* Depth first evaluation */
@@ -80,9 +82,9 @@ object DirectStyleDCNoCache {
           val new_env = c_env + (v -> baddr)
           val new_store = store.update(baddr, aeval(ae, env, store))
           reset { 
-            val bodyvss = aval(body, new_env, new_store, new_time)
+            val bdvss = aval(body, new_env, new_store, new_time)
             reset {
-              val (vs, acc_vss, bdnd) = ndcps[VS, Ans](bodyvss, Set[VS](), (evss: Ans) => closnd(evss ++ acc))
+              val (vs, acc_vss, bdnd) = ndcps[VS, Ans](bdvss, Set[VS](), (evss: Ans) => closnd(evss ++ acc))
               val VS(vals, time, store) = vs
               val baddr = allocBind(x, time)
               val new_env = env + (x -> baddr)

### 2018-06-01 src/main/scala/refunc/DisentangledLinearSmallStepUB.scala

@@ -33,8 +33,10 @@ object DisLinearSmallStepUBStack {
             drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), new_seen)
 
           case NDState(Letrec(bds, body), env, bstore, konts, time, ndk) =>
-            val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
-            val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
+            val new_env = bds.foldLeft(env)((accenv: Env, bd: B) =>
+              accenv + (bd.x -> allocBind(bd.x, new_time)))
+            val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) =>
+              accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)))
             drive_step(NDState(body, new_env, new_store, konts, new_time, ndk), new_seen)
             
           case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>

### 2018-06-01 src/main/scala/refunc/FusedLinearSmallStepUB.scala

@@ -31,8 +31,10 @@ object FusedLinearSmallStepUBStack {
             NDState(e, new_env, new_store, konts, new_time, ndk)
 
           case NDState(Letrec(bds, body), env, bstore, konts, time, ndk) =>
-            val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
-            val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
+            val new_env = bds.foldLeft(env)((accenv: Env, bd: B) =>
+              accenv + (bd.x -> allocBind(bd.x, new_time)))
+            val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) =>
+              accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)))
             NDState(body, new_env, new_store, konts, new_time, ndk)
 
           case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>

### 2018-06-01 src/main/scala/refunc/LinearSmallStepUB.scala

@@ -28,8 +28,10 @@ object LinearSmallStepUBStack {
         NDState(e, new_env, new_store, konts, new_time, ndk)
 
       case NDState(Letrec(bds, body), env, bstore, konts, time, ndk) =>
-        val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
-        val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
+        val new_env = bds.foldLeft(env)((accenv: Env, bd: B) =>
+          accenv + (bd.x -> allocBind(bd.x, new_time)))
+        val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) =>
+          accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)))
         NDState(body, new_env, new_store, konts, new_time, ndk)
 
       case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>

### 2018-06-01 src/main/scala/refunc/RefuncCPS.scala

@@ -31,9 +31,10 @@ object RefuncCPS {
           })
 
         case Letrec(bds, body) => 
-          val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
-          val new_store = bds.foldLeft(store)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), 
-                                                                                         atomicEval(bd.e, new_env, accbst)) })
+          val new_env = bds.foldLeft(env)((accenv: Env, bd: B) =>
+            accenv + (bd.x -> allocBind(bd.x, new_time)))
+          val new_store = bds.foldLeft(store)((accbst: BStore, bd: B) => 
+            accbst.update(allocBind(bd.x, new_time), atomicEval(bd.e, new_env, accbst)))
           aval(body, new_env, new_store, new_time, new_cache, { 
             case Ans(bdss, bdcache) => cont(Ans(bdss, bdcache.outUpdate(config, bdss)))
           })
@@ -44,8 +45,8 @@ object RefuncCPS {
             case (Clos(Lam(v, body), c_env), clsans, clsnd) =>
               val vbaddr = allocBind(v, new_time)
               val new_cenv = c_env + (v -> vbaddr)
-              val new_cstore = store.update(vbaddr, atomicEval(ae, env, store))
-              aval(body, new_cenv, new_cstore, new_time, clsans.cache, { 
+              val new_store = store.update(vbaddr, atomicEval(ae, env, store))
+              aval(body, new_cenv, new_store, new_time, clsans.cache, { 
                 case Ans(bdvss, bdcache) =>
                   nd[VS, Ans](bdvss, Ans(Set[VS](), bdcache), { 
                     case (VS(vals, time, vssstore), bdans, bdnd) =>
@@ -59,7 +60,8 @@ object RefuncCPS {
                     case eans => clsnd(eans ++ clsans)
                   })
               })
-          }, { case ans => cont(ans) })
+          }, 
+          cont)
     
         case ae if isAtomic(ae) =>
           val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
@@ -112,14 +114,18 @@ object RefuncCPSBF {
           val baddr = allocBind(v, new_time)
           val new_env = c_env + (v -> baddr)
           val new_store = store.update(baddr, atomicEval(ae, env, store))
[…]

### 2018-06-01 src/main/scala/refunc/RefuncCPSNoCache.scala

@@ -25,17 +25,17 @@ object RefuncCPSNoCache {
         nd[Clos, Ans](closures, Set[VS](), { case (clos, acc, closnd) =>
           val Clos(Lam(v, body), c_env) = clos
           val baddr = allocBind(v, new_time)
-          val new_env = c_env + (v -> baddr)
+          val new_cenv = c_env + (v -> baddr)
           val new_store = store.update(baddr, aeval(ae, env, store))
-          aval(body, new_env, new_store, new_time, (bodyvss: Set[VS]) => {
+          aval(body, new_cenv, new_store, new_time, (bodyvss: Set[VS]) => {
             nd[VS, Ans](bodyvss, Set[VS](), { case (vs, acc_vss, bdnd) =>
               val VS(vals, time, store) = vs
               val baddr = allocBind(x, time)
               val new_env = env + (x -> baddr)
               val new_store = store.update(baddr, vals)
-              aval(e, new_env, new_store, time, (evss: Ans) => bdnd(acc_vss ++ evss))
+              aval(e, new_env, new_store, time, { case evss => bdnd(acc_vss ++ evss) })
             },
-            (evss: Ans) => closnd(evss ++ acc))
+            { case evss => closnd(evss ++ acc) })
           })
         },
         cont)

### 2018-06-01 src/main/scala/refunc/SmallStep.scala

@@ -15,8 +15,10 @@ object SmallStep {
         List(State(e, new_env, newBStore, kstore, kaddr, new_time))
 
       case State(Letrec(bds, body), env, bstore, kstore, kaddr, time) =>
-        val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
-        val newBStore = bds.foldLeft(bstore)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
+        val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => 
+          accenv + (bd.x -> allocBind(bd.x, new_time)))
+        val newBStore = bds.foldLeft(bstore)((accbst: BStore, bd: B) => 
+          accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)))
         List(State(body, new_env, newBStore, kstore, kaddr, time))
 
       case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) if isAtomic(ae)=>

### 2018-06-01 src/main/scala/refunc/SmallStepP4F.scala

@@ -18,8 +18,10 @@ object SmallStepP4F {
         List(State(e, new_env, newBStore, kstore, kaddr, new_time))
 
       case State(Letrec(bds, body), env, bstore, kstore, kaddr, time) =>
-        val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
-        val newBStore = bds.foldLeft(bstore)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
+        val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => 
+          accenv + (bd.x -> allocBind(bd.x, new_time)))
+        val newBStore = bds.foldLeft(bstore)((accbst: BStore, bd: B) =>
+          accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)))
         List(State(body, new_env, newBStore, kstore, kaddr, time))
 
       case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) =>

### 2018-06-01 src/main/scala/refunc/SmallStepUB.scala

@@ -20,10 +20,14 @@ object SmallStepUBStack {
         val new_env = env + (x -> baddr)
         val new_store = bstore.update(baddr, aeval(ae, env, bstore))
         List(State(e, new_env, new_store, konts, new_time))
+
       case State(Letrec(bds, body), env, bstore, konts, time) =>
-        val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => { accenv + (bd.x -> allocBind(bd.x, new_time)) })
-        val newBStore = bds.foldLeft(bstore)((accbst: BStore, bd: B) => { accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)) })
+        val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => 
+          accenv + (bd.x -> allocBind(bd.x, new_time)))
+        val newBStore = bds.foldLeft(bstore)((accbst: BStore, bd: B) => 
+          accbst.update(allocBind(bd.x, new_time), aeval(bd.e, new_env, accbst)))
         List(State(body, new_env, newBStore, konts, time))
+
       case State(Let(x, App(f, ae), e), env, bstore, konts, time) =>
         for (Clos(Lam(v, body), c_env) <- aeval(f, env, bstore).toList) yield {
           val frame = Frame(x, e, env)
@@ -32,6 +36,7 @@ object SmallStepUBStack {
           val new_store = bstore.update(baddr, aeval(ae, env, bstore))
           State(body, new_env, new_store, frame::konts, new_time)
         }
+
       case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
         konts match {
           case Nil => List()

### 2018-06-01 src/main/scala/refunc/RefuncCPSNoCache.scala
misc

@@ -21,7 +21,6 @@ object RefuncCPSNoCache {
     e match {
       case Let(x, App(f, ae), e) =>
         val closures = aeval(f, env, store).asInstanceOf[Set[Clos]]
-
         nd[Clos, Ans](closures, Set[VS](), { case (clos, acc, closnd) =>
           val Clos(Lam(v, body), c_env) = clos
           val baddr = allocBind(v, new_time)
@@ -34,8 +33,9 @@ object RefuncCPSNoCache {
               val new_env = env + (x -> baddr)
               val new_store = store.update(baddr, vals)
               aval(e, new_env, new_store, time, { case evss => bdnd(acc_vss ++ evss) })
-            },
-            { case evss => closnd(evss ++ acc) })
+            }, { 
+              case evss => closnd(evss ++ acc) 
+            })
           })
         },
         cont)

### 2018-06-01 src/main/scala/refunc/DirectStyle.scala
renaming

@@ -133,10 +133,10 @@ object DirectStyleSideEff {
 
   def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore) = {
     def iter(cache: Cache): Ans = {
-      val Ans(vss, new_cache_) = aval(e, env, store, mtTime, cache)
-      val new_cache = new_cache_.outUpdate(Config(e, env, store, mtTime), vss)
-      if (new_cache.out == cache.out) { Ans(vss, new_cache) }
-      else { iter(Cache(new_cache.out, new_cache.out)) }
+      val Ans(vss, new_cache) = aval(e, env, store, mtTime, cache)
+      val updated_cache = new_cache.outUpdate(Config(e, env, store, mtTime), vss)
+      if (updated_cache.out == cache.out) { Ans(vss, updated_cache) }
+      else { iter(Cache(updated_cache.out, updated_cache.out)) }
     }
     iter(Cache.mtCache)
   }

### 2018-06-02 paper/paper.tex
update
update the code in paper

@@
 case Let(x, App(f, ae), e) =>
-  val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
-  nd[Clos, Ans](closures, Set[VS](), { case (clos, accouter, closnd) =>
+  val closures = aeval(f, env, store).asInstanceOf[Set[Clos]]
+  nd[Clos, Ans](closures, Set[VS](), { case (clos, clsacc, clsnd) =>
     val Clos(Lam(v, body), c_env) = clos
     val baddr = allocBind(v, new_time)
-    val new_env = c_env + (v -> baddr)
-    val new_store = store.update(baddr, atomicEval(ae, env, store))
-    aeval(body, new_env, new_store, new_time, (bodyvss: Set[VS]) => {
-      nd[VS, Ans](bodyvss, Set[VS](), { case (vs, accinner, bdnd) =>
+    val new_cenv = c_env + (v -> baddr)
+    val new_store = store.update(baddr, aeval(ae, env, store))
+    aeval(body, new_cenv, new_store, new_time, (bdvss: Set[VS]) => {
+      nd[VS, Ans](bdvss, Set[VS](), { case (vs, bdacc, bdnd) =>
         val VS(vals, time, store) = vs
+        val baddr = allocBind(x, time)
         val new_env = env + (x -> baddr)
         val new_store = store.update(baddr, vals)
-        aeval(e, new_env, new_store, time,
-              /* accumulate the values of one path, and call next body value*/
-              (evss: Ans) => bdnd(accinner ++ evss))
-      },
-      /* accumulate the values of paths to accouter, and call next closure */
-      (evss: Ans) => closnd(evss ++ accouter))
+        aeval(e, new_env, new_store, time, { 
+          /* accumulate the values of one path, and call next body value */
+          case evss => bdnd(bdacc ++ evss) 
+        })
+      }, { 
+        /* accumulate the values of paths, and call next closure */
+        case evss => clsnd(evss ++ clsacc) 
+      })
     })
@@
 Remember that the function @aeval@ returns a set of @VS@ objects to its continuation,
-so the continuation argument @bodyvss@ is a set of @VS@ objects which represent values
+so the continuation argument @bdvss@ is a set of @VS@ objects which represent values
 from multiple computation paths when evaluating this single @body@ expression.
 
-We next consider the inner call to @nd@, which evaluates over the set of body values @bodyvss@.
+We next consider the inner call to @nd@, which evaluates over the set of body values @bdvss@.
 For each @VS@, the timestamp and store in @VS@ will be instantiated to
@@
   /* lookups the out cache, uses the values if contains this config */
-  if (cache.outContains(config)) return continue(Ans(cache.outGet(config), cache))
-  /* uses the values from in cache */
+  if (cache.outContains(config)) cont(Ans(cache.outGet(config), cache))
+  else {
     val new_cache = cache.outUpdate(config, cache.inGet(config))
@@
         val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
-      nd[Clos, Ans](closures, Ans(Set[VS](), new_cache), { case (clos, Ans(accouter, cache), closnd) =>
-        ......
-        aeval(body, new_env, new_store, new_time, cache, { case Ans(bodyvss, bodycache) =>
-          nd[VS, Ans](bodyvss, Ans(Set[VS](), bodycache), { case (vs, Ans(acc_vss, cache), bdnd) =>
-            ......
-            aeval(e, new_env, new_store, time, cache, { case Ans(evss, ecache) =>
-              bdnd(Ans(accinner ++ evss, ecache))
+        nd[Clos, Ans](closures, Ans(Set[VS](), new_cache), { 
+          case (Clos(Lam(v, body), c_env), clsans, clsnd) =>
+            ...... /* sets vairables, e.g., new_cenv and new_store */
+            aeval(body, new_cenv, new_store, new_time, clsans.cache, { case Ans(bdvss, bdcache) =>
+                nd[VS, Ans](bdvss, Ans(Set[VS](), bdcache), { 
+                  case (VS(vals, time, vssstore), bdans, bdnd) =>
+                    ...... /* sets variables, e.g., new_env and new_store */
+                    aeval(e, new_env, new_store, time, bdans.cache, { 
+                      /* updates the out cache after evaluated to the end of one path  */
+                      case Ans(evss, ecache) => bdnd(bdans ++ Ans(evss, ecache.outUpdate(config, evss)))
+                    })
+                }, { case eans => clsnd(eans ++ clsans) })
             })
         }, 
-          { case Ans(evss, cache) => closnd(Ans(evss ++ accouter, cache)) })
-        })
-      },
-      /* updates the out cache after evaluation */
-      { case Ans(vss, cache) => continue(Ans(vss, cache.outUpdate(config, vss))) })
+        cont)
 
@@
         val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
-      continue(Ans(vs, new_cache.outUpdate(config, vs))) /* updates the out cache after evaluation */
+        cont(Ans(vs, new_cache.outUpdate(config, vs)))
+    }
   }
@@
 After completing the evaluation, we obtain a set of @VS@ objects and must update the @out@ cache.
-In the first case of pattern matching, this happens inside the last continuation of the
-outer @nd@ call.
+In the first case of pattern matching, this happens at the innermost continuation of @aeval@,
+since it is where we reach the end of one computation path.
 For the second case, we update the @out@ cache before calling the @continue@ continuation.
@@
         val new_cenv = c_env + (v -> vbaddr)
-        val new_cstore = store.update(vbaddr, aeval(ae, env, store))
+        val new_store = store.update(vbaddr, atomicEval(ae, env, store))
         /* evaluates the function application App(f, ae) */
-        val Ans(bodyvss, bodycache) = aeval(body, new_cenv, new_cstore, new_time, clscache)
-        val (VS(vals, time, vsstore), vscache) = ndcps[VS](bodyvss, Ans(Set[VS](), bodycache))
+        val Ans(bdvss, bdcache) = aeval(body, new_cenv, new_store, new_time, clscache)
+        val (VS(vals, time, vsstore), vscache) = ndcps[VS](bdvss, Ans(Set[VS](), bdcache))
         val baddr = allocBind(x, time)
         val new_env = env + (x -> baddr)
-        val new_store = vsstore.update(baddr, vals)
+        val new_vsstore = vsstore.update(baddr, vals)
         /* evaluates the body expression from the let */
-        val Ans(finval, fincache) = aeval(e, new_env, new_store, time, vscache)
-        Ans(finval, fincache.outUpdate(config, finval))
+        val Ans(evss, ecache) = aeval(e, new_env, new_vsstore, time, vscache)
+        Ans(evss, ecache.outUpdate(config, evss))
 
@@
         val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
-        Ans(vs, cache.outUpdate(config, vs))
+        Ans(vs, new_cache.outUpdate(config, vs))
     }
@@
 
-%\section{Perspective}
-%\subsection{Defunctionalization}
 %specify higher-order definitional interpreters using first-order means (cite)
-%\subsection{Refunctionalization}
 %\section{Detour: Shiver's \textit{k}-CFA?}

### 2018-06-02 src/main/scala/refunc/RefuncCPSNoCache.scala

@@ -21,20 +21,20 @@ object RefuncCPSNoCache {
     e match {
       case Let(x, App(f, ae), e) =>
         val closures = aeval(f, env, store).asInstanceOf[Set[Clos]]
-        nd[Clos, Ans](closures, Set[VS](), { case (clos, acc, closnd) =>
+        nd[Clos, Ans](closures, Set[VS](), { case (clos, clsacc, clsnd) =>
           val Clos(Lam(v, body), c_env) = clos
           val baddr = allocBind(v, new_time)
           val new_cenv = c_env + (v -> baddr)
           val new_store = store.update(baddr, aeval(ae, env, store))
-          aval(body, new_cenv, new_store, new_time, (bodyvss: Set[VS]) => {
-            nd[VS, Ans](bodyvss, Set[VS](), { case (vs, acc_vss, bdnd) =>
+          aval(body, new_cenv, new_store, new_time, (bdvss: Set[VS]) => {
+            nd[VS, Ans](bdvss, Set[VS](), { case (vs, bdacc, bdnd) =>
               val VS(vals, time, store) = vs
               val baddr = allocBind(x, time)
               val new_env = env + (x -> baddr)
               val new_store = store.update(baddr, vals)
-              aval(e, new_env, new_store, time, { case evss => bdnd(acc_vss ++ evss) })
+              aval(e, new_env, new_store, time, { case evss => bdnd(bdacc ++ evss) })
             }, { 
-              case evss => closnd(evss ++ acc) 
+              case evss => clsnd(evss ++ clsacc) 
             })
           })
         },

### 2018-06-02 paper/paper.tex
update code in paper

@@
         aeval(e, new_env, new_store, time, { 
-          /* accumulate the values of one path, and call next body value */
+          /* accumulate the values of one path, and call next body expression of f */
           case evss => bdnd(bdacc ++ evss) 
@@
       }, { 
-        /* accumulate the values of paths, and call next closure */
+        /* accumulate the values of paths, and call next target closure */
         case evss => clsnd(evss ++ clsacc) 

### 2018-06-04 src/main/scala/refunc/CESK.scala
Add concrete CES interp; remove useless imports

@@ -22,8 +22,9 @@ object Concrete {
   def isAtomic(e: Expr): Boolean = e.isInstanceOf[Var] || e.isInstanceOf[Lam]
 }
 
+import Concrete._
+
 object CESK {
-  import Concrete._
   case class Frame(x: String, e: Expr, env: Env)
   type Kont = List[Frame]
   case class State(e: Expr, env: Env, store: Store, k: Kont)
@@ -55,8 +56,6 @@ object CESK {
 }
 
 object RefuncCESK {
-  import Concrete._
-
   case class VS(v: Storable, store: Store)
   type Ans = VS
   type Cont = Ans => Ans
@@ -79,4 +78,23 @@ object RefuncCESK {
   }
 
   def run(e: Expr) = eval(e, Map(), Map(), (vs: VS) => vs)
+}
+
+object DirectStyleCES {
+  import RefuncCESK._
+  def eval(e: Expr, env: Env, store: Store): VS = e match {
+    case Let(x, ae, e) if isAtomic(ae) =>
+      val addr = alloc(store)
+      eval(e, env+(x->addr), store+(addr->atomicEval(ae, env, store)))
+
+    case Let(x, App(f, ae), e) =>
+      val Clos(Lam(v, body), env_c) = atomicEval(f, env, store)
+      val vaddr = alloc(store)
+      val VS(appv, appstore) = eval(body, env_c+(v->vaddr), store+(vaddr->atomicEval(ae, env, store)))
+      val xaddr = alloc(appstore)
+      eval(e, env+(x->xaddr), appstore+(xaddr->appv))
+
+    case ae if isAtomic(ae) =>
+      VS(atomicEval(ae, env, store), store)
+  }
 }

### 2018-06-04 src/main/scala/refunc/DirectStyle.scala

@@ -5,8 +5,6 @@ import refunc.ast._
 import ANFAAM._
 
 object DirectStyleDC {
-  import SmallStepUBStack._
-  
   type Cont = Ans => Ans
   
   @deprecated
@@ -81,7 +79,6 @@ object DirectStyleDC {
 
 object DirectStyleSideEff {
   /* Using side effect to update new_cache */
-  import SmallStepUBStack._
   import RefuncCPS._
 
   def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans = {

### 2018-06-04 src/main/scala/refunc/DirectStyleNoCache.scala

@@ -9,8 +9,6 @@ import ANFAAM._
 
 object DirectStyleDCNoCache2 {
   /* Depth first evaluation */
-  import SmallStepUBStack._
-  
   type Ans = Set[VS]
   type Cont = Ans => Ans
 
@@ -56,8 +54,6 @@ object DirectStyleDCNoCache2 {
 
 object DirectStyleDCNoCache {
   /* Depth first evaluation */
-  import SmallStepUBStack._
-  
   type Ans = Set[VS]
   type Cont = Ans => Ans
 

### 2018-06-04 src/main/scala/refunc/RefuncCPS.scala

@@ -5,8 +5,6 @@ import refunc.ast._
 import ANFAAM._
 
 object RefuncCPS {
-  import SmallStepUBStack._
-
   type Cont = Ans => Ans
   
   @deprecated
@@ -94,8 +92,6 @@ object RefuncCPS {
 
 /* Experimental implementation with breadth first evaluation */
 object RefuncCPSBF {
-  import SmallStepUBStack._
-  
   type Cont = (Set[VS], Cache) => (Set[VS], Cache)
 
   def nd[T](ts: Set[T], acc: Set[VS], cache: Cache, 

### 2018-06-04 src/main/scala/refunc/RefuncCPSNoCache.scala

@@ -6,8 +6,6 @@ import refunc.ast._
 import ANFAAM._
 
 object RefuncCPSNoCache {
-  import SmallStepUBStack._
-  
   type Ans = Set[VS]
   type Cont = Ans => Ans
 
@@ -54,8 +52,6 @@ object RefuncCPSNoCache {
 
 /* Experimental implementation with breadth first evaluation */
 object RefuncCPSNoCacheBF {
-  import SmallStepUBStack._
-  
   type Cont = Set[VS] => Set[VS]
 
   def nd[T](cs: List[T], acc: Set[VS], f: (T, Set[VS], Cont) => Set[VS], g: Cont): Set[VS] = {

### 2018-06-04 src/main/scala/refunc/CESK.scala
add comments

@@ -24,6 +24,7 @@ object Concrete {
 
 import Concrete._
 
+/* Concrete interpreter in CESK abstract machine style. */
 object CESK {
   case class Frame(x: String, e: Expr, env: Env)
   type Kont = List[Frame]
@@ -55,6 +56,7 @@ object CESK {
   def eval(e: Expr): State = drive(inject(e))
 }
 
+/* Refunctionalized concrete interpreter in CPS style. */
 object RefuncCESK {
   case class VS(v: Storable, store: Store)
   type Ans = VS
@@ -69,9 +71,9 @@ object RefuncCESK {
       val Clos(Lam(v, body), env_c) = atomicEval(f, env, store)
       val addr = alloc(store)
       eval(body, env_c+(v->addr), store+(addr->atomicEval(ae, env, store)), (vs: VS) => {
-              val addr = alloc(vs.store)
-              eval(e, env+(x->addr), vs.store+(addr->vs.v), k)
-            })
+        val addr = alloc(vs.store)
+        eval(e, env+(x->addr), vs.store+(addr->vs.v), k)
+      })
 
     case ae if isAtomic(ae) =>
       k(VS(atomicEval(ae, env, store), store))
@@ -80,6 +82,7 @@ object RefuncCESK {
   def run(e: Expr) = eval(e, Map(), Map(), (vs: VS) => vs)
 }
 
+/* Cocnrete interpreter in direct-style. */
 object DirectStyleCES {
   import RefuncCESK._
   def eval(e: Expr, env: Env, store: Store): VS = e match {

### 2018-06-04 paper/paper.tex
misc
renaming variables

@@
           drive_step(NDState(body, new_env, new_store, new_frames, new_time, new_ndk), new_seen)
-
         case NDState(ae, env, bstore, konts, time, ndk) if isAtomic(ae) =>
@@
       val new_store = bstore.update(baddr, argvs)
-      drive_step(NDState(body, new_env, new_store, frames, time,
-                         NDCont(cls.tail, argvs, bstore, time, frames)::ndk),
-                 seen)
+      val new_ndk = NDCont(cls.tail, argvs, bstore, time, frames)::ndk
+      drive_step(NDState(body, new_env, new_store, frames, time, new_ndk, seen))
   }
@@
 case ae if isAtomic(ae) =>
-  val ds = atomicEval(ae, env, store)
-  continue(Set(VS(ds, new_time, store)))
+  val ds = atomicEval(ae, env, store); continue(Set(VS(ds, new_time, store)))
 \end{lstlisting}

### 2018-06-04 src/main/scala/refunc/DirectStyle.scala

@@ -5,8 +5,6 @@ import refunc.ast._
 import ANFAAM._
 
 object DirectStyleDC {
-  type Cont = Ans => Ans
-  
   @deprecated
   def ndcps[T](ts: Set[T], acc: Ans): (T, Cache) @cps[Ans] = shift { 
     f: (((T, Cache)) => Ans) => nd(ts, acc, f)

### 2018-06-04 src/main/scala/refunc/SmallStep.scala

@@ -11,24 +11,24 @@ object SmallStep {
       case State(Let(x, ae, e), env, bstore, kstore, kaddr, time) if isAtomic(ae) =>
         val baddr = allocBind(x, new_time)
         val new_env = env + (x -> baddr)
-        val newBStore = bstore.update(baddr, atomicEval(ae, env, bstore))
-        List(State(e, new_env, newBStore, kstore, kaddr, new_time))
+        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+        List(State(e, new_env, new_store, kstore, kaddr, new_time))
 
       case State(Letrec(bds, body), env, bstore, kstore, kaddr, time) =>
         val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => 
           accenv + (bd.x -> allocBind(bd.x, new_time)))
-        val newBStore = bds.foldLeft(bstore)((accbst: BStore, bd: B) => 
+        val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) => 
           accbst.update(allocBind(bd.x, new_time), atomicEval(bd.e, new_env, accbst)))
-        List(State(body, new_env, newBStore, kstore, kaddr, time))
+        List(State(body, new_env, new_store, kstore, kaddr, time))
 
       case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) if isAtomic(ae)=>
         for (Clos(Lam(v, body), c_env) <- atomicEval(f, env, bstore).toList) yield {
           val baddr = allocBind(v, new_time)
           val new_env = c_env + (v -> baddr)
-          val newBStore = bstore.update(baddr, atomicEval(ae, env, bstore))
-          val newKAddr = allocKont(body, c_env, newBStore, new_time)
-          val newKStore = kstore.update(newKAddr, Cont(Frame(x, e, env), kaddr))
-          State(body, new_env, newBStore, newKStore, newKAddr, new_time)
+          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+          val new_kaddr = allocKont(body, c_env, new_store, new_time)
+          val new_kstore = kstore.update(new_kaddr, Cont(Frame(x, e, env), kaddr))
+          State(body, new_env, new_store, new_kstore, new_kaddr, new_time)
         }
 
       case State(ae, env, bstore, kstore, kaddr, time) if isAtomic(ae)=>

### 2018-06-04 src/main/scala/refunc/SmallStepP4F.scala

@@ -14,24 +14,24 @@ object SmallStepP4F {
       case State(Let(x, ae, e), env, bstore, kstore, kaddr, time) if isAtomic(ae) =>
         val baddr = allocBind(x, new_time)
         val new_env = env + (x -> baddr)
-        val newBStore = bstore.update(baddr, atomicEval(ae, env, bstore))
-        List(State(e, new_env, newBStore, kstore, kaddr, new_time))
+        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+        List(State(e, new_env, new_store, kstore, kaddr, new_time))
 
       case State(Letrec(bds, body), env, bstore, kstore, kaddr, time) =>
         val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => 
           accenv + (bd.x -> allocBind(bd.x, new_time)))
-        val newBStore = bds.foldLeft(bstore)((accbst: BStore, bd: B) =>
+        val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) =>
           accbst.update(allocBind(bd.x, new_time), atomicEval(bd.e, new_env, accbst)))
-        List(State(body, new_env, newBStore, kstore, kaddr, time))
+        List(State(body, new_env, new_store, kstore, kaddr, time))
 
       case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) =>
         for (Clos(Lam(v, body), c_env) <- atomicEval(f, env, bstore).toList) yield {
           val baddr = allocBind(v, new_time)
           val new_env = c_env + (v -> baddr)
-          val newBStore = bstore.update(baddr, atomicEval(ae, env, bstore))
-          val newKAddr = allocKont(body, c_env, newBStore, new_time)
-          val newKStore = kstore.update(newKAddr, Cont(Frame(x, e, env), kaddr))
-          State(body, new_env, newBStore, newKStore, newKAddr, new_time)
+          val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+          val new_kaddr = allocKont(body, c_env, new_store, new_time)
+          val new_kstore = kstore.update(new_kaddr, Cont(Frame(x, e, env), kaddr))
+          State(body, new_env, new_store, new_kstore, new_kaddr, new_time)
         }
 
       case State(ae, env, bstore, kstore, kaddr, time) if isAtomic(ae)=>

### 2018-06-05 paper/paper.tex
refactor function tick

@@
 program contour @time@ that encodes the program execution history.
-We use a list of execution contexts (expressions) to represent this, and
-as we will see in Section~\ref{kcfainst}, by applying different @tick@
-functions on the timestamp, we are able to obtain a family of analyses.
+The function @tick@ is used to refresh the "time" and get a "new time".
+We use a list of execution contexts (expressions) to represent the finite 
+program contour, and as we will see in Section~\ref{kcfainst}, by applying 
+different @tick@ functions on the timestamp, we are able to obtain a 
+family of analyses.
 
@@
 def step(s: State): List[State] = {
-  val new_time = tick(s)
+  val new_time = s.tick
   s match {
@@
 The @tick@ function returns the $k$ front-most expressions given the current state and its
-time history.
+time history. We implement @tick@ as a public method of case class @State@.
 
@@
 def k: Int = 0
-def tick(s: State): Time = (s.e :: s.time).take(k)
+case class State(e: Expr, env: Env, bstore: BStore, kstore: KStore, kaddr: KAddr, time: Time) {
+  def tick: Time = (e :: time).take(k)
+}
 \end{lstlisting}
@@
       val new_seen = if (seen.contains(s)) seen else seen+s
-      val newTime = tick(nds)
+      val newTime = nds.tick
       val new_ndstate = nds match {
@@
       val new_seen = if (seen.contains(s)) seen else seen + s
-      val new_time = tick(nds)
+      val new_time = nds.tick
       nds match {
@@
   val NDState(ae, env, bstore, konts, time, ndk) = nds
-  val new_time = tick(nds)
+  val new_time = nds.tick
   konts match {
@@
 contains a set of storable values, a timestamp, and a store.
+
 \begin{lstlisting}
@@
 \end{lstlisting}
+
 An instance of @VS@ represents the computational result of one path in
@@
 values is a collection @Set[VS]@. For convenience, we define it as a type @Ans@:
+
 \begin{lstlisting}
@@
 @aeval@. Therefore, the pattern matching on the state becomes just a match on
-expression @e@.
+expression @e@. The shape of function @aeval@ looks like:
+
 \begin{lstlisting}
@@
 def aeval(e: Expr, env: Env, store: BStore, time: Time, continue: Cont): Ans = {
-  val new_time = (e::time).take(k)
+  val new_time = ......
   e match {
-    case Let(x, App(f, ae), e) =>
-      ......
-    case ae if isAtomic(ae) =>
-      ......
+    case Let(x, App(f, ae), e) => ......
+    case ae if isAtomic(ae) => ......
   }
@@
 a state-like definition which borrows the components from state, sans continuations.
+
 \begin{lstlisting}
@@
     val new_cache = cache.outUpdate(config, cache.inGet(config))
-    val new_time = (e::time).take(k)
+    val new_time = config.tick 
     e match {
@@
   else {
-    val new_time = (e::time).take(k)
+    val new_time = config.tick
     val new_cache = cache.outUpdate(config, cache.inGet(config))

### 2018-06-05 paper/paper.tex
revise the nd/choices operators

@@
 To handle the inevitable nondeterminism, we define a generic function @nd@ as follows:
-\begin{lstlisting}
-def nd[T,S](ts: Set[T], acc: S, f: (T, S, S => S) => S, g: S => S): S = {
-  if (ts.isEmpty) g(acc)
-  else f(ts.head, acc, (vss: S) => nd(ts.tail, vss, f, g))
-}
-\end{lstlisting}
-
-Given a set of elements of type @T@, an initial accumulated value of type @S@, a function @f@
-that works on element type @T@ and accumulated type @S@, and a function @g@ that
-works on and returns a value of type @S@, @nd@ applies @f@ to each element in the
-set iteratively, accumulates the value and finally applies @g@ on it.
+
+\begin{lstlisting}
+def nd[T](ts: Iterable[T], acc: Ans, k: (T, Ans, Ans => Ans) => Ans, m: Ans => Ans): Ans = {
+  if (ts.isEmpty) m(acc)
+  else k(ts.head, acc, (ans: Ans) => nd(ts.tail, ans, k, m))
+}
+\end{lstlisting}
+
+Given a collection of elements of type @T@, an initial accumlated value of type 
+@Ans@, a function @k@ that works on element's type @T@ and accumulated value's 
+type @Ans@ (note that @k@ also takes a continuation of type @Ans=>Ans@), 
+and a function @m@ that works on and returns a value of type @Ans@, 
+@nd@ chooses an element in the collection and applies @k@ to it, accumulates 
+the value and finally applies @m@ on the accumulated value when there is no element 
+in the collection.
 
@@
 there are two out calls to @ndcontinue@ and @drive_step@, this is the reason why
-we have functions @f@ and @g@ in @nd@.
+we have functions @k@ and @m@ in @nd@.
 
@@
 it is an implementation of @fold@ in continuation style. But here we have two
-continuations, one for the next element in the set, and one for the final result.
+continuations, one for the next element in the collection, and one for the final result.
 This is sometimes referred to as an extended continuation-passing style (ECPS)
-that has a continuation and a meta-continuation \cite{Danvy:1990:AC:91556.91622}.
+that function @k@ is a normal continuation and @m@ is a meta-continuation 
+\cite{Danvy:1990:AC:91556.91622}.
 
@@
 case Let(x, App(f, ae), e) =>
-  val closures = aeval(f, env, store).asInstanceOf[Set[Clos]]
-  nd[Clos, Ans](closures, Set[VS](), { case (clos, clsacc, clsnd) =>
+  val closures = atomicEval(f, env, store)
+  nd[Storable](closures, Set[VS](), { case (clos, clsacc, clsnd) =>
     val Clos(Lam(v, body), c_env) = clos
@@
     val new_cenv = c_env + (v -> baddr)
-    val new_store = store.update(baddr, aeval(ae, env, store))
+    val new_store = store.update(baddr, atomicEval(ae, env, store))
     aeval(body, new_cenv, new_store, new_time, (bdvss: Set[VS]) => {
-      nd[VS, Ans](bdvss, Set[VS](), { case (vs, bdacc, bdnd) =>
+      nd[VS](bdvss, Set[VS](), { case (vs, bdacc, bdnd) =>
         val VS(vals, time, store) = vs
@@
         val new_store = store.update(baddr, vals)
-        aeval(e, new_env, new_store, time, { 
-          /* accumulate the values of one path, and call next body expression of f */
-          case evss => bdnd(bdacc ++ evss) 
+        /* evaluates and accumluates the values of one path, then call bdnd */
+        aeval(e, new_env, new_store, time, { case evss => bdnd(bdacc++evss) }) 
+      }, { case evss => clsnd(evss++clsacc) }) /* accumulate the values of (multiple) paths, then call clsnd */
     })
-      }, { 
-        /* accumulate the values of paths, and call next target closure */
-        case evss => clsnd(evss ++ clsacc) 
-      })
-    })
-  },
-  continue)
+  }, continue)
 \end{lstlisting}
@@
       case Let(x, App(f, ae), e) =>
-        val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
-        nd[Clos, Ans](closures, Ans(Set[VS](), new_cache), { 
+        val closures = atomicEval(f, env, store)
+        nd[Storable](closures, Ans(Set[VS](), new_cache), {
           case (Clos(Lam(v, body), c_env), clsans, clsnd) =>
@@
             aeval(body, new_cenv, new_store, new_time, clsans.cache, { case Ans(bdvss, bdcache) =>
-              nd[VS, Ans](bdvss, Ans(Set[VS](), bdcache), { 
+              nd[VS](bdvss, Ans(Set[VS](), bdcache), {
                 case (VS(vals, time, vssstore), bdans, bdnd) =>
@@
         },
-        cont)
+        continue)
 
@@
         val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
-        cont(Ans(vs, new_cache.outUpdate(config, vs)))
+        continue(Ans(vs, new_cache.outUpdate(config, vs)))
     }
@@
 
-To address this problem, we reintroduce two operators @nd@ and @ndcps@ for handling 
-non-deterministic choices in a way that mimics extended continuation-passing style.
+To address this problem, we reintroduce two operators @nd@ and @choices@ for handling 
+nondeterministic choices in a way that is direct-style but mimics extended 
+continuation-passing style.
 
 \textbf{Nondeterminism Operators, Again.}
-The first operator @nd@ is simplified from the previous one, but specialized with type 
-@Ans@ and @Cache@. We may still keep the types generic, but the presentation clearly, 
-exposing @Cache@ explictly reveals that cache should be monotonically accumulated to 
-the rest of computation. Now @nd@ only takes a function @f@ that applies on type 
-@(T, Cache)@ where the value of type @T@ comes from the set, and @nd@ performs recursive 
-call on itself over elements in the set @ts@. We also move the accumulation operation 
+The first operator @nd@ is adapted from the previous one, but here @nd@ has only one 
+continuation @k@. Introducing @Cache@ explicitly as the part of arguments of 
+continuation @k@ reveals that cache should be monotonically accumulated to 
+the rest of computation. Now the only continuation @k@ applies on type @(T, Cache)@ where
+the value of type @T@ comes from the collection, and @nd@ performs recursive invocation
+on itself over elements in the collection @ts@. We also move the accumulation operation
 (@++@) inside of @nd@.
@@
 \begin{lstlisting}
-def nd[T](ts: Set[T], acc: Ans, k: ((T, Cache)) => Ans): Ans = {
+def nd[T](ts: Iterable[T], acc: Ans, k: ((T, Cache)) => Ans): Ans = {
   if (ts.isEmpty) acc
@@
 
-The second operator @ndcps@ is implemented with delimited control operator 
+The second operator @choices@ is implemented with delimited control operator 
 @shift@ but simply calls our first operator @nd@ in body.
@@
 \begin{lstlisting}
-def ndcps[T](ts: Set[T], acc: Ans): (T, Cache) @cps[Ans] = shift { f: (((T, Cache)) => Ans) =>
-  nd(ts, acc, f)
-}
-\end{lstlisting}
-
-The function @ndcps@ takes two arguments: a set of elements with type @T@, 
+def choices[T](ts: Iterable[T], cache: Cache): (T, Cache) @cps[Ans] = shift {
+  f: (((T, Cache)) => Ans) => nd(ts, Ans(Set[VS](), cache), f)
+}
+\end{lstlisting}
+
+The function @choices@ takes two arguments: a set of elements with type @T@, 
 and an initial accumulated value of type @Ans@.
-The @ndcps@ iteratively returns an element from the set along with the latest cache, 
+The @choices@ iteratively returns an element from the collection along with the latest cache, 
 and the return type is also annotated by \verb|@cps[Ans]|. 
-This annotation denotes the final returned value of @ndcps@ is type @Ans@.
+This annotation denotes the final returned value of @choices@ is type @Ans@.
 The @f@ introduced by the @shift@ operator is the delimited continuation from the 
-call-site of @ndcps@.
+call-site of @choices@.
 
@@
 Then, for every nondeterministic choice in the abstract interpreter, we can simply call 
-@ndcps@ on the set of possible choices and write the program sequentially as a concrete 
-interpreter. @ndcps@ also returns the latest cache to its left-hand side definition.
-The cache is still accumulated to the next call to @ndcps@ or @aeval@, 
+@choices@ on the set of possible choices and write the program sequentially as concrete 
+interpreters, and the subsequent statements and expressions become the delimited continuation
+@f@. @choices@ also returns the latest cache to its left-hand side definition.
+The cache is still accumulated to the next call to @choices@ or @aeval@, 
 given the fact that the abstract interpreter should always use the latest cache.
@@
       case Let(x, App(f, ae), e) =>
-        val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
-        val (Clos(Lam(v, body), c_env), clscache) = ndcps[Clos](closures, Ans(Set[VS](), new_cache))
+        val closures = atomicEval(f, env, store)
+        val (Clos(Lam(v, body), c_env), clscache) = choices[Storable](closures, new_cache)
         val vbaddr = allocBind(v, new_time)
@@
         val Ans(bdvss, bdcache) = aeval(body, new_cenv, new_store, new_time, clscache)
-        val (VS(vals, time, vsstore), vscache) = ndcps[VS](bdvss, Ans(Set[VS](), bdcache))
+        val (VS(vals, time, vsstore), vscache) = choices[VS](bdvss, bdcache)
         val baddr = allocBind(x, time)
@@
 The function @analyze@ is also changed by adding a @reset@ operator around the function 
-call to @aeval@. The afterwards updating on cache becomes sequential.
+call to @aeval@ to set the delimiter. The afterwards updating on cache becomes sequential.
 

### 2018-06-05 src/main/scala/refunc/RefuncCPS.scala

@@ -18,9 +18,9 @@ object RefuncCPS {
     else k(ts.head, acc, (ans: Ans) => nd(ts.tail, ans, k, m))
   }
   
-  def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, cont: Cont): Ans = {
+  def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, continue: Cont): Ans = {
     val config = Config(e, env, store, time)
-    if (cache.outContains(config)) cont(Ans(cache.outGet(config), cache))
+    if (cache.outContains(config)) continue(Ans(cache.outGet(config), cache))
     else {
       val new_time = config.tick
       val new_cache = cache.outUpdate(config, cache.inGet(config))
@@ -31,7 +31,7 @@ object RefuncCPS {
           val new_env = env + (x -> baddr)
           val new_store = store.update(baddr, atomicEval(ae, env, store))
           aeval(e, new_env, new_store, new_time, new_cache, { 
-            case Ans(evss, ecache) => cont(Ans(evss, ecache.outUpdate(config, evss)))
+            case Ans(evss, ecache) => continue(Ans(evss, ecache.outUpdate(config, evss)))
           })
 
         case Letrec(bds, body) => 
@@ -40,7 +40,7 @@ object RefuncCPS {
           val new_store = bds.foldLeft(store)((accbst: BStore, bd: B) => 
             accbst.update(allocBind(bd.x, new_time), atomicEval(bd.e, new_env, accbst)))
           aeval(body, new_env, new_store, new_time, new_cache, { 
-            case Ans(bdss, bdcache) => cont(Ans(bdss, bdcache.outUpdate(config, bdss)))
+            case Ans(bdss, bdcache) => continue(Ans(bdss, bdcache.outUpdate(config, bdss)))
           })
 
         case Let(x, App(f, ae), e) =>
@@ -65,11 +65,11 @@ object RefuncCPS {
                   })
               })
           }, 
-          cont)
+          continue)
     
         case ae if isAtomic(ae) =>
           val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
-          cont(Ans(vs, new_cache.outUpdate(config, vs)))
+          continue(Ans(vs, new_cache.outUpdate(config, vs)))
       }
     }
   }
@@ -100,10 +100,10 @@ object RefuncCPSBF {
[…]

### 2018-06-05 src/main/scala/refunc/RefuncCPSNoCache.scala

@@ -9,38 +9,36 @@ object RefuncCPSNoCache {
   type Ans = Set[VS]
   type Cont = Ans => Ans
 
-  def nd[T,S](ts: Set[T], acc: S, f: (T, S, S=>S) => S, g: S => S): S = {
-    if (ts.isEmpty) g(acc)
-    else f(ts.head, acc, (vss: S) => nd(ts.tail, vss, f, g))
+  def nd[T](ts: Iterable[T], acc: Ans, k: (T, Ans, Ans=>Ans) => Ans, m: Ans=>Ans): Ans = {
+    if (ts.isEmpty) m(acc)
+    else k(ts.head, acc, (ans: Ans) => nd(ts.tail, ans, k, m))
   }
 
-  def aeval(e: Expr, env: Env, store: BStore, time: Time, cont: Cont): Ans = {
+  def aeval(e: Expr, env: Env, store: BStore, time: Time, continue: Cont): Ans = {
     val new_time = (e::time).take(k)
     e match {
       case Let(x, App(f, ae), e) =>
-        val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
-        nd[Clos, Ans](closures, Set[VS](), { case (clos, clsacc, clsnd) =>
+        val closures = atomicEval(f, env, store)
+        nd[Storable](closures, Set[VS](), { case (clos, clsacc, clsnd) =>
           val Clos(Lam(v, body), c_env) = clos
           val baddr = allocBind(v, new_time)
           val new_cenv = c_env + (v -> baddr)
           val new_store = store.update(baddr, atomicEval(ae, env, store))
           aeval(body, new_cenv, new_store, new_time, (bdvss: Set[VS]) => {
-            nd[VS, Ans](bdvss, Set[VS](), { case (vs, bdacc, bdnd) =>
+            nd[VS](bdvss, Set[VS](), { case (vs, bdacc, bdnd) =>
               val VS(vals, time, store) = vs
               val baddr = allocBind(x, time)
               val new_env = env + (x -> baddr)
               val new_store = store.update(baddr, vals)
               aeval(e, new_env, new_store, time, { case evss => bdnd(bdacc ++ evss) })
-            }, { 
-              case evss => clsnd(evss ++ clsacc) 
-            })
+            }, { case evss => clsnd(evss ++ clsacc) })
           })
         },
-        cont)
+        continue)
   
       case ae if isAtomic(ae) =>
[…]

### 2018-06-05 paper/paper.tex
reformat the dots

@@
         case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndkonts) =>
-          ......
+          ...
         case NDState(ae, env, bstore, konts, time, ndkonts) if isAtomic(ae) =>
-          ......
+          ...
       }
@@
         case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
-          ......
+          ...
           drive_step(NDState(body, new_env, new_store, new_frames, new_time, new_ndk), new_seen)
@@
 def aeval(e: Expr, env: Env, store: BStore, time: Time, continue: Cont): Ans = {
-  val new_time = ......
+  val new_time = ...
   e match {
-    case Let(x, App(f, ae), e) => ......
-    case ae if isAtomic(ae) => ......
+    case Let(x, App(f, ae), e) => ...
+    case ae if isAtomic(ae) => ...
   }
@@
           case (Clos(Lam(v, body), c_env), clsans, clsnd) =>
-            ...... /* sets vairables, e.g., new_cenv and new_store */
+            ... /* sets vairables, e.g., new_cenv and new_store */
             aeval(body, new_cenv, new_store, new_time, clsans.cache, { case Ans(bdvss, bdcache) =>
@@
                 case (VS(vals, time, vssstore), bdans, bdnd) =>
-                  ...... /* sets variables, e.g., new_env and new_store */
+                  ... /* sets variables, e.g., new_env and new_store */
                   aeval(e, new_env, new_store, time, bdans.cache, {

### 2018-06-06 paper/paper.tex
revise section 6 and 7

@@
 \begin{lstlisting}
-def inject(e: Expr): State =
-  State(e, Map(),
-        Store[BAddr, Storable](Map()),
-        Store[KAddr, Cont](Map(Halt -> Set())),
-        Halt, List())
+val mtTime = List();                         val mtStore = Store[BAddr, Storable](Map())
+val mtEnv = Map[String, BAddr]();            val mtKStore = Store[KAddr, Cont](Map(Halt -> Set()))
+def inject(e: Expr): State = State(e, mtEnv, mtStore, mtKStore, Halt, mtTime)
 \end{lstlisting}
@@
 have the same values but different accumulated side effects (e.g., memory allocations).
-Since the whole computation is nondeterministic, the type of the final result
+Since the whole computation is potentially nondeterministic, the type of the final result
 values is a collection @Set[VS]@. For convenience, we define it as a type @Ans@:
@@
 makes a call to @continue@. Here, the reemerging @continue@ is a function
-with type @Ans => Ans@. The function @aeval@ does not return: instead, it calls
+of type @Ans => Ans@. The function @aeval@ does not return: instead, it calls
 @continue@ with the result values.
@@
 \begin{lstlisting}
-def nd[T](ts: Iterable[T], acc: Ans, k: (T, Ans, Ans => Ans) => Ans, m: Ans => Ans): Ans = {
+def nd[T](ts: Iterable[T], acc: Ans, k: (T, Ans => Ans) => Ans, m: Ans => Ans): Ans = {
   if (ts.isEmpty) m(acc)
-  else k(ts.head, acc, (ans: Ans) => nd(ts.tail, ans, k, m))
-}
-\end{lstlisting}
-
+  else k(ts.head, (ans: Ans) => nd(ts.tail, acc ++ ans, k, m))
+}
+\end{lstlisting}
+
+%TODO: better word to replace `accumulate`
 Given a collection of elements of type @T@, an initial accumlated value of type 
-@Ans@, a function @k@ that works on element's type @T@ and accumulated value's 
-type @Ans@ (note that @k@ also takes a continuation of type @Ans=>Ans@), 
-and a function @m@ that works on and returns a value of type @Ans@, 
-@nd@ chooses an element in the collection and applies @k@ to it, accumulates 
-the value and finally applies @m@ on the accumulated value when there is no element 
-in the collection.
-
+@Ans@, a function @k@ that works on element's type @T@ and takes a continuation 
+of type @Ans=>Ans@, and a function @m@ that works on and returns a value of type 
+@Ans@ (which is also type @Ans=>Ans@), @nd@ chooses an element in the collection 
+and applies @k@ to it, accumulates the value (@acc++ans@ where the @ans@ is the 
+result comes from @k@) and eventually applies function @m@ on the accumulated 
+value when there is no element in the collection.
+
+%TODO: consider rewrite this paragraph and explain the relation to `continue`.
 If we look at @continue@ and @ndcontinue@ in the disentangled AAM, we will find
 that @nd@ is an abstraction of those two functions. In function @continue@,
-there are two out calls to @ndcontinue@ and @drive_step@, this is the reason why
-we have functions @k@ and @m@ in @nd@.
+there are two out calls to @drive_step@ and @ndcontinue@, here the application
+of @k@ simulates @drive_step@, and @m@ simulates @ndcontinue@.
 
 Readers familiar with functional programming may think of @fold@ at this point; indeed,
-it is an implementation of @fold@ in continuation style. But here we have two
-continuations, one for the next element in the collection, and one for the final result.
+it is a variant implementation of @fold@ in continuation style. But here we have two
+continuations, one for the next element in the collection, and one for the eventual result.
 This is sometimes referred to as an extended continuation-passing style (ECPS)
@@
 \textbf{Evaluation without Caching.}
-Since our first version of @aeval@ implements the nondeterministic
-evaluation, given the defined @nd@, the evaluation function @aeval@ can be easily
-defined. In the case that @e@ is a @let@ expression,
-we invoke @nd@ twice and perform a nesting two-step, depth-first evaluation
-over possible values of @App(f, ae)@.
-The values of branches from a fork point will be accumulated, and then
-returned to the upper fork point.
+Since our first version of @aeval@ implements only the nondeterministic evaluation, 
+given the defined @nd@ operator, the evaluation function @aeval@ can be easily defined. 
+In the case that @e@ is a @let@ expression, we invoke @nd@ twice and perform a nesting 
+two-step, depth-first evaluation over possible values of @App(f, ae)@.
+The values of branches from a fork point will be accumulated inside of @nd@, and 
+then returned to the upper fork point.
 
@@
   val closures = atomicEval(f, env, store)
-  nd[Storable](closures, Set[VS](), { case (clos, clsacc, clsnd) =>
+  nd[Storable](closures, Set[VS](), { case (clos, clsnd) =>
     val Clos(Lam(v, body), c_env) = clos
@@
     aeval(body, new_cenv, new_store, new_time, (bdvss: Set[VS]) => {
-      nd[VS](bdvss, Set[VS](), { case (vs, bdacc, bdnd) =>
+      nd[VS](bdvss, Set[VS](), { case (vs, bdnd) =>
         val VS(vals, time, store) = vs
@@
         val new_store = store.update(baddr, vals)
-        /* evaluates and accumluates the values of one path, then call bdnd */
-        aeval(e, new_env, new_store, time, { case evss => bdnd(bdacc++evss) }) 
-      }, { case evss => clsnd(evss++clsacc) }) /* accumulate the values of (multiple) paths, then call clsnd */
+        /* evaluate e with new env and store, then transfer to next body's value */
+        aeval(e, new_env, new_store, time, bdnd)
+      }, clsnd) /* transfer to next target closure of f */
     })
-  }, continue)
-\end{lstlisting}
-
-The outer call to @nd@ evaluates over the set of possible closures of
+  },
+  continue)
+\end{lstlisting}
+
+The outer application of @nd@ evaluates over the set of possible target closures of
 @f@. For each such closure, we go into its body with a new environment and a new store.
@@
 
-We next consider the inner call to @nd@, which evaluates over the set of body values @bdvss@.
-For each @VS@, the timestamp and store in @VS@ will be instantiated to
-@aeval@.
-The inner application of @aeval@ returns a set of values of one computation path to its
-continuation, where the continuation will accumulate the result and transfer
-evaluation to the next body value.
-
-\begin{lstlisting}
-case ae if isAtomic(ae) =>
-  val ds = atomicEval(ae, env, store); continue(Set(VS(ds, new_time, store)))
-\end{lstlisting}
-
-If the expression is atomic, then we simply construct a new instance of
-@VS@ and return it to the @continue@ function.
-
-At the end, the initial invocation to @aeval@ is passed with an identity
-function as a halting continuation:
-\begin{lstlisting}
-def analyze(e: Expr) =
-  aeval(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List(), (ans => ans))
+We next consider the inner application of @nd@, which evaluates over the set of body's 
+value/stores @bdvss@. For each @VS@, the timestamp and store in @VS@ will be instantiated to
+the next application of @aeval@. The inner application of @aeval@ evaluates @e@ with the
+substituted environment and store, then returns a set of values of one computation path 
+to its continuation @bdnd@, where the continuation will accumulate the result and 
+transfer evaluation to the next body value.
+
+The meta-continuation of the inner @nd@ is @clsnd@, meaning that we need to transfer 
+to the next target closure of @f@. The meta-continuation of the outer @nd@ is just 
+@continue@, as what we normally would do for CPS interpreters.
+
+\begin{lstlisting}
+case ae if isAtomic(ae) => val ds = atomicEval(ae, env, store); continue(Set(VS(ds, new_time, store)))
+\end{lstlisting}
+
+If the argument expression @e@ is atomic, which is the second case of match, we simply 
+construct a new instance of @VS@ and return it to the @continue@ function.
+
+At the end, the initial invocation to @aeval@ is passed with initial environments,
+especially an identity function as the halting continuation:
+\begin{lstlisting}
+def analyze(e: Expr) = aeval(e, mtEnv, mtStore, mtTime, (ans => ans))
 \end{lstlisting}
@@
 Based on the refunctionalized AAM from the last section, we use the same nondeterministic
-operator @nd@, but further extend it to a fixed-point cached evaluation in this section.
+operator @nd@, but further extend it to a fixed-point cached version in this section.
 The fixed-point caching algorithm performs a sound over-approximation of all possible
@@
 \textbf{Configuration.}
-We do not have the concept of an explicit stack or state,
-so to represent the cache we introduce a configuration @Config@ as
-a state-like definition which borrows the components from state, sans continuations.
+We do not have the concept of an explicit stack or state, so to represent the cache 
+we introduce a configuration @Config@ as a state-like definition which borrows the 
+components from state, sans continuations.
 
@@
 \textbf{Cache.}
-Recall that the latent assumption of the worklist algorithm in classical AAM
-is that if we have seen a state @s@, it means we also have seen the successors
-of state @s@.
+Recall that the latent assumption of the worklist algorithm in classical AAM is that 
+if we have seen a state @s@, it means we also have seen all the successors of state @s@.
 The caching algorithm we adopt here replays the same assumption but in a big-step 
-manner: if we have seen the configuration @c@, then it means we also have seen
-the values that are evaluated by the configuration @c@.
+manner: if we have seen the configuration @c@, then it means we also have seen the 
+values that are evaluated from the configuration @c@.
 Here we will apply the fixed-point caching algorithm as described from
@@
 
-The @Cache@ contains two maps @in@ and @out@ which are both mappings from
-configurations to sets of @VS@. The @in@ cache contains mappings from the
-previous iteration of evaluation, and the @out@ cache contains mappings after the current
-iteration of evaluation. Once we have evaluated a term to some values,
-we update the @out@ cache. In the next iteration, we will use the @out@
-from the previous iteration as the @in@ cache for the current iteration.
+The @Cache@ contains two maps @in@ and @out@ which are both mappings from configurations 
+to sets of @VS@. The @in@ cache contains mappings from the previous iteration of evaluation, 
+and the @out@ cache contains mappings after the current iteration of evaluation. 
+Once we have evaluated a term to some values, we update the @out@ cache. In the next 
+iteration, we will use the @out@ from the previous iteration as the @in@ cache for the 
+current iteration.
 
 \textbf{Cache-Passing Style.}
-We can now transform our interpreter into cache-passing style.
-We begin by changing the return type @Ans@ to a case class which contains a @VS@ set and
-a cache.
-
-\begin{lstlisting}
-case class Ans(vss: Set[VS], cache: Cache)
+We can now transform our interpreter into cache-passing style. We begin by changing the 
+return type @Ans@ to a case class which contains a @VS@ set and a cache. @Ans@ also supports
+its own accumulating operation @++@ which defines as union on @Set[VS]@ and join on caches.
+
+\begin{lstlisting}
+case class Ans(vss: Set[VS], cache: Cache) {
+  def ++(ans: Ans): Ans = { Ans(vss ++ ans.vss, cache.join(ans.cache)) }
+}
+\end{lstlisting}
+
+The nondeterministic operator @nd@ is also slightly rewrite to cache-passing style:
+continuation @k@ takes an additional argument of type @Cache@; the current accumulated
+cache @acc.cache@ is exposed to @k@ when applies.
+
+\begin{lstlisting}
+def nd[T](ts: Iterable[T], acc: Ans, k: (T, Cache, Ans => Ans) => Ans, m: Ans => Ans): Ans = {
+  if (ts.isEmpty) m(acc)
+  else k(ts.head, acc.cache, (ans: Ans) => nd(ts.tail, acc ++ ans, k, m))
+}
 \end{lstlisting}
@@
 The skeleton of function @aeval@ largely remains the same.
-Every place which has a value of @Ans@ is also replaced as direct pattern matching so
-that we can use the fields in @Ans@ immediately.
-The cache is used in a monotonic way: each function call to @aeval@ or @nd@ is
-passed with the latest cache from the most recent continuation's result.
+
+\begin{lstlisting}
+def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, continue: Cont): Ans = {
+  val config = Config(e, env, store, time)
+  if (cache.outContains(config)) cont(Ans(cache.outGet(config), cache))
+  else {
+    val new_cache = cache.outUpdate(config, cache.inGet(config))
+    val new_time = config.tick 
+    e match { ...  }
+  }
+}
+\end{lstlisting}
 
@@
 from the previous iteration.
-In reality, we first conservatively assume that all computations can be diverged, and
-if not, we update its values in the mapping after evaluation.
-In terms of partial orders and lattices, it is the case that starts
-the Kleene iteration from the bottom of the lattice.
-
-\begin{lstlisting}
-def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, continue: Cont): Ans = {
-  val config = Config(e, env, store, time)
-  /* lookups the out cache, uses the values if contains this config */
-  if (cache.outContains(config)) cont(Ans(cache.outGet(config), cache))
-  else {
-    val new_cache = cache.outUpdate(config, cache.inGet(config))
-    val new_time = config.tick 
-    e match {
+In reality, we first conservatively assume that all computations can be diverged, and if 
+not, we update its values in the mapping after evaluation. In terms of partial orders and 
+lattices, it is the case that starts the Kleene iteration from the bottom of the lattice.
+
+The two cases of match are shown below. Note that the cache is used in a monotonic way: 
+each function call to @aeval@ or @nd@ is passed with the latest cache from the most recent 
+continuation of @aeval@ or @nd@.
+
+\begin{lstlisting}
 case Let(x, App(f, ae), e) =>
@@
   nd[Storable](closures, Ans(Set[VS](), new_cache), {
-          case (Clos(Lam(v, body), c_env), clsans, clsnd) =>
-            ... /* sets vairables, e.g., new_cenv and new_store */
-            aeval(body, new_cenv, new_store, new_time, clsans.cache, { case Ans(bdvss, bdcache) =>
+    case (Clos(Lam(v, body), c_env), clscache, clsnd) =>
+      val vbaddr = allocBind(v, new_time); val new_cenv = c_env + (v -> vbaddr)
+      val new_store = store.update(vbaddr, atomicEval(ae, env, store))
+      aeval(body, new_cenv, new_store, new_time, clscache, { case Ans(bdvss, bdcache) =>
         nd[VS](bdvss, Ans(Set[VS](), bdcache), {
-                case (VS(vals, time, vssstore), bdans, bdnd) =>
-                  ... /* sets variables, e.g., new_env and new_store */
-                  aeval(e, new_env, new_store, time, bdans.cache, {
-                    /* updates the out cache after evaluated to the end of one path  */
-                    case Ans(evss, ecache) => bdnd(bdans ++ Ans(evss, ecache.outUpdate(config, evss)))
+          case (VS(vals, time, vssstore), bdvsscache, bdnd) =>
+            val baddr = allocBind(x, time); val new_env = env + (x -> baddr)
+            val new_store = vssstore.update(baddr, vals)
+            aeval(e, new_env, new_store, time, bdvsscache, {
+              case Ans(evss, ecache) => bdnd(Ans(evss, ecache.outUpdate(config, evss)))
             })
-              }, { case eans => clsnd(eans ++ clsans) })
+        }, clsnd)
       })
@@
   continue)
-
 case ae if isAtomic(ae) =>
@@
   continue(Ans(vs, new_cache.outUpdate(config, vs)))
-    }
-  }
-}
 \end{lstlisting}
@@
 \begin{lstlisting}
-val mtEnv = Map[String, BAddr](); val mtStore = Store[BAddr, Storable](Map()); val mtTime = List()
 val mtCache = Cache(Store[Config, VS](Map()), Store[Config, VS](Map()))
-
 def analyze(e: Expr) = {
   def iter(cache: Cache): Ans = {
-    val Ans(vss, new_cache) = aeval(e, mtEnv, mtStore, mtTime, cache, {
-      case Ans(vss, cache) => 
-        val initConfig = Config(e, mtEnv, mtStore, mtTime) 
-        Ans(vss, cache.outUpdate(initConfig, vss))
+    val Ans(vss, new_cache) = aeval(e, mtEnv, mtStore, mtTime, cache, { case Ans(vss, cache) => 
+        Ans(vss, cache.outUpdate(Config(e, mtEnv, mtStore, mtTime), vss))
     })
@@
 \textbf{Nondeterminism Operators, Again.}
-The first operator @nd@ is adapted from the previous one, but here @nd@ has only one 
-continuation @k@. Introducing @Cache@ explicitly as the part of arguments of 
-continuation @k@ reveals that cache should be monotonically accumulated to 
-the rest of computation. Now the only continuation @k@ applies on type @(T, Cache)@ where
-the value of type @T@ comes from the collection, and @nd@ performs recursive invocation
-on itself over elements in the collection @ts@. We also move the accumulation operation
-(@++@) inside of @nd@.
+%Introducing @Cache@ explicitly as the part of arguments of continuation @k@ reveals 
+%that cache should be monotonically accumulated to the rest of computation. 
+The first operator @nd@ is adapted from the previous one, but the meta-continuation
+@m@ has been eliminated here. Now the only continuation @k@ applies on tuple type @(T, Cache)@ 
+where the value of type @T@ comes from the collection, and @nd@ performs recursive invocation
+on itself over elements in the collection @ts@.
 
@@
 and an initial accumulated value of type @Ans@.
-The @choices@ iteratively returns an element from the collection along with the latest cache, 
-and the return type is also annotated by \verb|@cps[Ans]|. 
+The @choices@ iteratively returns an element from the collection along with the 
+latest cache, and the return type is also annotated by \verb|@cps[Ans]|. 
 This annotation denotes the final returned value of @choices@ is type @Ans@.
-The @f@ introduced by the @shift@ operator is the delimited continuation from the 
-call-site of @choices@.
+The @f@ is the delimited continuation captured by the @shift@ operator, which
+is form by the statements after @choices@'s call-site.
 
 Now we can use these two operators to redefine @aeval@.
-We add an annotation \verb|@cps[Ans]| on the return type of the @aeval@ function.
+We add an annotation \verb|@cps[Ans]| on the return type of the @aeval@ function to enable
+applying delimited control operators.
 Then, for every nondeterministic choice in the abstract interpreter, we can simply call 
@@
   else {
-    val new_time = config.tick
-    val new_cache = cache.outUpdate(config, cache.inGet(config))
+    val new_time = config.tick; val new_cache = cache.outUpdate(config, cache.inGet(config))
     e match {
@@
         val (Clos(Lam(v, body), c_env), clscache) = choices[Storable](closures, new_cache)
-        val vbaddr = allocBind(v, new_time)
-        val new_cenv = c_env + (v -> vbaddr)
+        val vbaddr = allocBind(v, new_time); val new_cenv = c_env + (v -> vbaddr)
         val new_store = store.update(vbaddr, atomicEval(ae, env, store))
-        /* evaluates the function application App(f, ae) */
         val Ans(bdvss, bdcache) = aeval(body, new_cenv, new_store, new_time, clscache)
         val (VS(vals, time, vsstore), vscache) = choices[VS](bdvss, bdcache)
-        val baddr = allocBind(x, time)
-        val new_env = env + (x -> baddr)
+        val baddr = allocBind(x, time); val new_env = env + (x -> baddr)
         val new_vsstore = vsstore.update(baddr, vals)
-        /* evaluates the body expression from the let */
         val Ans(evss, ecache) = aeval(e, new_env, new_vsstore, time, vscache)
         Ans(evss, ecache.outUpdate(config, evss))
-
       case ae if isAtomic(ae) =>
-        val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
-        Ans(vs, new_cache.outUpdate(config, vs))
+        val vss = Set(VS(atomicEval(ae, env, store), new_time, store))
+        Ans(vss, new_cache.outUpdate(config, vss))
     }
@@
     val Ans(vss, anscache) = reset { aeval(e, mtEnv, mtStore, mtTime, cache) }
-    val initConfig = Config(e, mtEnv, mtStore, mtTime) 
-    val new_cache = anscache.outUpdate(initConfig, vss)
+    val new_cache = anscache.outUpdate(Config(e, mtEnv, mtStore, mtTime), vss)
     if (new_cache.out == cache.out) { Ans(vss, new_cache) }

### 2018-06-06 src/main/scala/refunc/ANFAAMBasis.scala

@@ -61,7 +61,7 @@ object ANFAAM {
 
   def atomicEval(e: Expr, env: Env, bstore: BStore): Set[Storable] = e match {
     case Num(i) => Set(NumV(i))
-    case Var(x) => bstore(env(x)).toSet
+    case Var(x) => bstore(env(x))
     case lam@Lam(x, body) => Set(Clos(lam, env))
     case _ => throw new NotImplementedError(e.toString)
   }

### 2018-06-06 paper/paper.tex
revise section 7

@@
 The second operator @choices@ is implemented with delimited control operator 
-@shift@ but simply calls our first operator @nd@ in body.
+@shift@ but simply calls our first operator @nd@ in its body.
 
@@
 
-Now we can use these two operators to redefine @aeval@.
-We add an annotation \verb|@cps[Ans]| on the return type of the @aeval@ function to enable
-applying delimited control operators.
+Now we can use these two operators to redefine @aeval@. The abstract evaluator @aeval@ is
+rewritten in direct-style, so that the last argument @cont@ disappears; we also add an 
+annotation \verb|@cps[Ans]| on the return type to enable using delimited control operators 
+in it.
 Then, for every nondeterministic choice in the abstract interpreter, we can simply call 

### 2018-06-06 paper/paper.tex
revise section 2.4

@@
 unbounded which uses a precise call stack as we did in the concrete CESK machine.
-Instead of allocating continuations in the store and embedding
-addresses of continuations in states, we intend to use a list of frames
-to explicitly model the stack.
-By doing so, we recover the call stack as precise as runtimes
-(so called pushdown control-flow analysis), but since
-the stack is unbounded, the analysis is potentially not computable.
-For readers who are not familiar with pushdown analysis, we have a detailed
-discussion in Section~\ref{pdcfarevisit}.
-The reason we show it here is to establish an artifact with precise call/return 
-matching used for the next step of transformation.
+Instead of allocating continuations in the store and embedding addresses of 
+continuations in states, we intend to use a list of frames to explicitly model 
+the stack. By doing so, we recover the call stack as precise as runtimes
+(so called pushdown control-flow analysis), but since the stack is unbounded, 
+the state space is potentially infinite.
+
+The pushdown AAM with unbounded stack is also described by \citeauthor{van2012systematic} 
+\cite{van2012systematic}. 
+The reason we show it here is to keep the consistency of abstract semantics during 
+transformations -- since the final abstract definitional interpreters simply inherit
+a precise control-flow from its defining language, we also would like to start from a
+pushdown AAM that is also able to precisely match call and returns.
+For readers who are not familiar with pushdown analysis, we have a detailed discussion 
+in Section~\ref{pdcfarevisit}.
 
@@
   \item In fact, we may also just use the same caching algorithm but with side effects such as
-    assignments and mutations to update the cache, and then achieve the same
-    definitional interpreter with pushdown control-flows. In this case, nondeterminism can be
-    easily handled via @for@ comprehension.
+    assignments and mutations to update the cache, and then achieve the same definitional 
+    interpreter with pushdown control-flows. In this case, nondeterminism can be easily 
+    handled via @for@ comprehension.
 \end{itemize}
@@
 
+% TODO: P4F v.s. direct-style + caching
+% TODO: dyck graph summarization v.s. direct-style + caching
+
 In this functional pearl, we fill the gap between small-step abstract abstract machines 

### 2018-06-06 paper/paper.tex
misc

@@
 a precise control-flow from its defining language, we also would like to start from a
-pushdown AAM that is also able to precisely match call and returns.
+pushdown AAM that is also able to precisely match calls and returns.
 For readers who are not familiar with pushdown analysis, we have a detailed discussion 

### 2018-06-06 paper/paper.tex
revise the code

@@
 \begin{lstlisting}
-type Addr = Int
-type Env = Map[String, Addr]
-type Store = Map[Addr, Storable]
-
+type Addr = Int;           type Env = Map[String, Addr];           type Store = Map[Addr, Storable]
 abstract class Storable
 case class Clos(v: Lam, env: Env) extends Storable
-
-case class Frame(x: String, e: Expr, env: Env)
-type Kont = List[Frame]
-
+case class Frame(x: String, e: Expr, env: Env);                    type Kont = List[Frame]
 case class State(e: Expr, env: Env, store: Store, k: Kont)
@@
 \begin{lstlisting}
-abstract class Storable
-case class Clos(v: Lam, env: Env) extends Storable
-
 case class Frame(x: String, e: Expr, env: Env)
@@
 
-In the definition of @State@, the continuation store disappears, and
-component @konts@ becomes a list of frames. 
-The other components remain unchanged as in the Section \ref{aam}.
+In the definition of @State@, the continuation store and continuation address disappear;
+instead, a list of frames represents the stack. An empty list denotes that we have reached 
+the halt. The other components remain unchanged as in the Section \ref{aam}.
 
@@
 
-The state transition function @step@ shown below is still
-nondeterministic, but the only nondeterminism happening is when dereferencing the callee @f@
-from the function application @App(f, ae)@.
+The state transition function @step@ shown below is still nondeterministic, but the only 
+nondeterminism happening is when dereferencing the callee @f@ from the function application 
+@App(f, ae)@.
 
@@
   def toState: State = State(e, env, bstore, konts, time)
+  def tick: Time = ...
 }
@@
         val Clos(Lam(v, body), c_env) = cls.head
-        val baddr = allocBind(v, time)
-        val new_env = c_env + (v -> baddr)
+        val baddr = allocBind(v, time); val new_env = c_env + (v -> baddr)
         val new_store = bstore.update(baddr, argvs)
@@
     case Frame(x, e, f_env)::konts =>
-      val baddr = allocBind(x, newTime)
-      val new_env = f_env + (x -> baddr)
+      val baddr = allocBind(x, newTime); val new_env = f_env + (x -> baddr)
       val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
@@
     case nds =>
-      val s = nds.toState
+      val s = nds.toState; val newTime = nds.tick
       val new_seen = if (seen.contains(s)) seen else seen+s
-      val newTime = nds.tick
       val new_ndstate = nds match {
-        case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndkonts) =>
-          ...
-        case NDState(ae, env, bstore, konts, time, ndkonts) if isAtomic(ae) =>
-          ...
+        case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndkonts) => ...
+        case NDState(ae, env, bstore, konts, time, ndkonts) if isAtomic(ae) => ...
       }
@@
     case nds =>
-      val s = nds.toState
+      val s = nds.toState; val new_time = nds.tick
       val new_seen = if (seen.contains(s)) seen else seen + s
-      val new_time = nds.tick
       nds match {
@@
 def continue(nds: NDState, seen: Set[State]): Set[State] = {
-  val NDState(ae, env, bstore, konts, time, ndk) = nds
-  val new_time = nds.tick
+  val NDState(ae, env, bstore, konts, time, ndk) = nds; val new_time = nds.tick
   konts match {
@@
     case Frame(x, e, f_env)::konts =>
-      val baddr = allocBind(x, new_time)
-      val new_env = f_env + (x -> baddr)
+      val baddr = allocBind(x, new_time); val new_env = f_env + (x -> baddr)
       val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
@@
       val Clos(Lam(v, body), c_env) = cls.head
-      val baddr = allocBind(v, time)
-      val new_env = c_env + (v -> baddr)
+      val baddr = allocBind(v, time); val new_env = c_env + (v -> baddr)
       val new_store = bstore.update(baddr, argvs)
@@
     val Clos(Lam(v, body), c_env) = clos
-    val baddr = allocBind(v, new_time)
-    val new_cenv = c_env + (v -> baddr)
+    val baddr = allocBind(v, new_time); val new_cenv = c_env + (v -> baddr)
     val new_store = store.update(baddr, atomicEval(ae, env, store))
@@
         val VS(vals, time, store) = vs
-        val baddr = allocBind(x, time)
-        val new_env = env + (x -> baddr)
+        val baddr = allocBind(x, time); val new_env = env + (x -> baddr)
         val new_store = store.update(baddr, vals)
@@
 \begin{lstlisting}
-case class Config(e: Expr, env: Env, store: BStore, time: Time)
+case class Config(e: Expr, env: Env, store: BStore, time: Time) { ... }
 \end{lstlisting}

### 2018-06-13 paper/paper.tex
misc
update code code; minor revision

@@
 The @f@ is the delimited continuation captured by the @shift@ operator, which
-is form by the statements after @choices@'s call-site.
+is formed by the statements after @choices@'s call-site.
 
@@
 Control-flow analysis is one instance of abstract interpretation on functional programs
-that can be traced to Jones \cite{jones1981flow}.
-\citeauthor{Shivers:1988:CFA:53990.54007} introduces $k$-CFA which uses $k$ recent calling
-contexts as program contour that differentiates values from different contexts 
-\cite{Shivers:1988:CFA:53990.54007, Shivers:1991:SSC:115865.115884}.
-The modern formulation of control-flow analysis is the abstracting abstract machines
+that can be traced to Jones \cite{jones1981flow}. \citeauthor{Shivers:1988:CFA:53990.54007} 
+introduces $k$-CFA which uses $k$ recent calling contexts as program contour that 
+differentiates values from different contexts \cite{Shivers:1988:CFA:53990.54007, 
+Shivers:1991:SSC:115865.115884}.
+The recent formulation of control-flow analysis is the abstracting abstract machines
 methodology \cite{van2012systematic, van2010abstracting} which forms the starting point
@@
 
-%% Acknowledgments
+%% TODO: Acknowledgments
 \begin{acks}                            %% acks environment is optional

### 2018-06-14 paper/paper.tex
centering the code

@@
 
-\begin{lstlisting}
+%------------------------------------------------------------------------------------------------------------
+\begin{lstlisting}
+                                    ae $\in$ AExp ::= ∂x | lam
+                                     x $\in$ Var   (variable names)
+                                   lam $\in$ Lam ∂::= (lambda (x) e)
                                      e $\in$ Exp ∂::= ae | (let ([x (ae ae)]) e)
-   ae $\in$ AExp ::= ∂x | lam
-  lam $\in$ Lam ∂::= (lambda (x) e)
-    x $\in$ Var   (variable names)
 \end{lstlisting}
@@
                           }
-type BStore = Store[BAddr, Storable]
-type KStore = Store[KAddr, Cont]
+               type BStore = Store[BAddr, Storable];            type KStore = Store[KAddr, Cont]
 \end{lstlisting}
@@
 \begin{lstlisting}
-case class Frame(x: String, e: Expr, env: Env)
-case class Cont(frame: Frame, kaddr: KAddr)
+     case class Frame(x: String, e: Expr, env: Env);      case class Cont(frame: Frame, kaddr: KAddr)
 \end{lstlisting}
@@
 \begin{lstlisting}
+                           def allocKont(tgtExpr: Expr): KAddr = ContAddr(tgtExpr)
                            def allocBind(x: String, time: Time): BAddr = BAddr(x, time)
-def allocKont(tgtExpr: Expr): KAddr = ContAddr(tgtExpr)
 \end{lstlisting}
@@
               }
-
               def analyze(e: Expr): Set[State] = drive(List(inject(e)), Set())
@@
 \begin{lstlisting}
-case ae if isAtomic(ae) => val ds = atomicEval(ae, env, store); continue(Set(VS(ds, new_time, store)))
+               case ae if isAtomic(ae) => 
+                 val ds = atomicEval(ae, env, store); continue(Set(VS(ds, new_time, store)))
 \end{lstlisting}
@@
 its own accumulating operation @++@ which defines as union on @Set[VS]@ and join on caches.
-
+% TODO join?
 \begin{lstlisting}

### 2018-06-14 paper/paper.tex
update color; minor revision

@@
 We can now transform our interpreter into cache-passing style. We begin by changing the 
-return type @Ans@ to a case class which contains a @VS@ set and a cache. @Ans@ also supports
+return type @Ans@ to a case class which contains a @VS@ set and a cache. @Ans@ also implements 
 its own accumulating operation @++@ which defines as union on @Set[VS]@ and join on caches.
-% TODO join?
+The join operation of caches is defined as joining its two store components.
+
 \begin{lstlisting}

### 2018-06-14 paper/paper.tex
update the style of subsubsection

@@
 For each closure in the set, a new frame is constructed and pushed onto the stack.
-The code for handling the second case (atomic expressions) is the same as the concrete CESK
-machines.
-
-\textbf{Note on Computability.}
+The code for handling the second case (atomic expressions) is the same as the concrete CESK machines. \\
+
+\textit{Note on Computability.}
 Unfortunately, even though other components in the states are finite,
@@
 of type @Ans => Ans@. The function @aeval@ does not return: instead, it calls
-@continue@ with the result values.
-
-\textbf{Nondeterminism Abstractions.}
+@continue@ with the result values. \\
+
+\textit{Nondeterminism Abstractions.}
 To handle the inevitable nondeterminism, we define a generic function @nd@ as follows:
@@
 that function @k@ is a normal continuation and @m@ is a meta-continuation 
-\cite{Danvy:1990:AC:91556.91622}.
-
-\textbf{Evaluation without Caching.}
+\cite{Danvy:1990:AC:91556.91622}. \\
+
+\textit{Evaluation without Caching.}
 Since our first version of @aeval@ implements only the nondeterministic evaluation, 
@@
 concrete reachable paths and values, and also prevents non-termination when analyzing 
-programs that may diverge.
-
-\textbf{Configuration.}
+programs that may diverge. \\
+
+\textit{Configuration.}
 We do not have the concept of an explicit stack or state, so to represent the cache 
@@
 
-\textbf{Cache.}
+\textit{Cache.}
 Recall that the latent assumption of the worklist algorithm in classical AAM is that 
@@
 iteration, we will use the @out@ from the previous iteration as the @in@ cache for the 
-current iteration.
-
-\textbf{Cache-Passing Style.}
+current iteration. \\
+
+\textit{Cache-Passing Style.}
 We can now transform our interpreter into cache-passing style. We begin by changing the 
@@
 we revisit the pushdown control-flow problem and examine what we have done to
-overcome it.
-
-\textbf{The Problem with Return-flows.}
+overcome it. \\
+
+\textit{The Problem with Return-flows.}
 Pushdown control-flow is a property in analysis that precisely models
@@
 However, variable @x@ still points to value set @{1, 2}@, because two call-sites still 
-share the same continuation and $1$-CFA is still unable to separate the return-flows.
-
-\textbf{Call/Return Matching through Refunctionalization and Caching.}
+share the same continuation and $1$-CFA is still unable to separate the return-flows. \\
+
+\textit{Call/Return Matching through Refunctionalization and Caching.}
 The refunctionalized AAM we obtained has the perfect match on return-flows even though 
@@
 nondeterministic choices in a way that is direct-style but mimics extended 
-continuation-passing style.
-
-\textbf{Nondeterminism Operators, Again.}
+continuation-passing style. \\
+
+\textit{Nondeterminism Operators, Again.}
 %Introducing @Cache@ explicitly as the part of arguments of continuation @k@ reveals 

### 2018-06-17 paper/paper.tex
reformat code
fix cache in code and paper

@@
 \begin{lstlisting}
+                                     x $\in$ Variables
                                     ae $\in$ AExp ::= ∂x | lam
-                                     x $\in$ Var   (variable names)
                                    lam $\in$ Lam ∂::= (lambda (x) e)
@@
     })
-           if (new_cache.out == cache.out) { Ans(vss, new_cache) }
-           else { iter(Cache(new_cache.out, new_cache.out)) }
+    if (new_cache.out == new_cache.in) Ans(vss, new_cache)
+    else iter(Cache(new_cache.out, Store[Config, VS](Map())))
   }
@@
 
+% TODO: explain more on how caches work
 Finally, the @analyze@ function (as the entrance of the analysis) does a looping iteration
 to find the fixed-point on caches starting from an empty cache.
-If the @out@ cache of this iteration is equivalent to the @out@ cache from the last iteration,
-then we have reached a fixed-point that over-approximates the concrete evaluations,
-and thus can be returned.
-Otherwise, the next iteration with the new @out@ cache will be activated.
+If the @out@ cache of this iteration is equivalent to its @in@ cache, which means no more 
+fact was discovered during this iteration, then we have reached a fixed-point that 
+over-approximates the concrete evaluations, and thus can be returned.
+Otherwise, the next iteration with the new @out@ cache installed at the @in@ cache position 
+will be activated; in the meanwhile, an empty cache will be used as the @out@ cache.
 Note that the continuation of the initial call to @aeval@ sets up the values for the
@@
     describe \cite{darais2017abstracting}.
-  \item By representing the extended continuation-passing style with delimited controls,
-    we derive a new form of abstract interpreters in direct-style.
+  \item By representing the extended continuation-passing style with delimited control operators
+    such as @shift@ and @reset@, we derive a new form of abstract interpreters in direct-style.
   \item In fact, we may also just use the same caching algorithm but with side effects such as
     assignments and mutations to update the cache, and then achieve the same definitional 
-    interpreter with pushdown control-flows. In this case, nondeterminism can be easily 
-    handled via @for@ comprehension.
+    interpreter with pushdown control-flows. \citeauthor{friedman-ai} in a tutorial on abstract 
+    interpreters use this style to model caches and updates \cite{friedman-ai}.
+    In this case, nondeterminism can be easily handled via @for@ comprehension. 
 \end{itemize}
@@
     val new_cache = anscache.outUpdate(Config(e, mtEnv, mtStore, mtTime), vss)
-                 if (new_cache.out == cache.out) { Ans(vss, new_cache) }
-                 else { iter(Cache(new_cache.out, new_cache.out)) }
+    if (new_cache.out == new_cache.in) Ans(vss, new_cache)
+    else iter(Cache(new_cache.out, Store[Config, VS](Map())))
   }

### 2018-06-17 src/main/scala/refunc/DirectStyle.scala

@@ -68,8 +68,8 @@ object DirectStyleDC {
       val Ans(vss, anscache) = reset { aeval(e, env, store, mtTime, cache) }
       val init_config = Config(e, env, store, mtTime)
       val new_cache = anscache.outUpdate(init_config, vss)
-      if (new_cache.out == cache.out) { Ans(vss, new_cache) }
-      else { iter(Cache(new_cache.out, new_cache.out)) }
+      if (new_cache.out == new_cache.in) { Ans(vss, new_cache) }
+      else { iter(Cache(new_cache.out, Store[Config, VS](Map()))) }
     }
     iter(Cache.mtCache)
   }
@@ -135,8 +135,8 @@ object DirectStyleSideEff {
     def iter(cache: Cache): Ans = {
       val Ans(vss, new_cache) = aeval(e, env, store, mtTime, cache)
       val updated_cache = new_cache.outUpdate(Config(e, env, store, mtTime), vss)
-      if (updated_cache.out == cache.out) { Ans(vss, updated_cache) }
-      else { iter(Cache(updated_cache.out, updated_cache.out)) }
+      if (updated_cache.out == new_cache.in) { Ans(vss, updated_cache) }
+      else { iter(Cache(updated_cache.out, Store[Config, VS](Map()))) }
     }
     iter(Cache.mtCache)
   }

### 2018-06-17 src/main/scala/refunc/RefuncCPS.scala

@@ -72,8 +72,8 @@ object RefuncCPS {
           val init_config = Config(e, env, store, mtTime)
           Ans(vss, cache.outUpdate(init_config, vss))
       })
-      if (new_cache.out == cache.out) { Ans(vss, new_cache) }
-      else { iter(Cache(new_cache.out, new_cache.out)) }
+      if (new_cache.out == new_cache.in) { Ans(vss, new_cache) }
+      else { iter(Cache(new_cache.out, Store[Config, VS](Map()))) }
     }
     iter(Cache.mtCache)
   }

### 2018-06-17 paper/paper.tex
reformat code; fix typo

@@
   case class BAddr(x: String, time: Time) 
-
   abstract class KAddr
@@
     val config = Config(e, env, store, time)
-  if (cache.outContains(config)) cont(Ans(cache.outGet(config), cache))
+    if (cache.outContains(config)) continue(Ans(cache.outGet(config), cache))
     else {

### 2018-06-23 paper/paper.tex
Use singular form of control

@@
 %% For final camera-ready submission, w/ required CCS and ACM Reference
-\documentclass[acmsmall]{acmart}\settopmatter{}
+\documentclass[acmsmall, review]{acmart}\settopmatter{}
 
@@
   The transformations include linearization, fusing, disentangling, refunctionalizing,
-  and un-CPS to direct-style with delimited controls. Linearization expresses
+  and un-CPS to direct-style with delimited control. Linearization expresses
   non-deterministic choices by first-order data types, after which refunctionalization
@@
     of abstract abstract machines; the transition of machine states therefore becomes 
-    deterministic. Notably, this introduces another layer of controls that will also 
+    deterministic. Notably, this introduces another layer of control that will also 
     be refunctionalized later.
@@
   \item The last transformation we show is that of transforming the refunctionalized AAM to a
-    \textbf{direct-style} interpreter (Section~\ref{directstyle}) by using delimited controls.
+    \textbf{direct-style} interpreter (Section~\ref{directstyle}) by using delimited control operators.
 \end{itemize}
@@
 % different direct styles: one implemented with monads, the other using
-% delimited controls.
+% delimited control operators.
 
@@
 By dereferencing this address in a continuation store, we can retrieve the actual 
-transfer of controls. The definition of environment @Env@ remains the same.
+transfer of control. The definition of environment @Env@ remains the same.
 
@@
 Likewise, the @Frame@ saves the information of the caller in concrete executions.
-We then introduce another layer of controls, and define a case class @NDCont@
+We then introduce another layer of control, and define a case class @NDCont@
 that saves the information at a fork point when we have multiple target closures.
@@
 in showing the correspondence between monads and continuation-passing style as well as
-delimited controls since their very beginning era \cite{Danvy:1990:AC:91556.91622, wadler1992essence,
+delimited control since their very beginning era \cite{Danvy:1990:AC:91556.91622, wadler1992essence,
 danvy1992representing, moggi1991notions}.
@@
 
-\subsection{Un-CPS by Delimited Controls} \label{uncps}
+\subsection{Back to Direct-Style with Delimited Control} \label{uncps}
 
@@
 This corresponds to the extended continuation-passing style.
-But in our defining language (Scala), the delimited controls rely on the continuation-passing 
+But in our defining language (Scala), the delimited control operators rely on the continuation-passing 
 transformation which is only one layer.
@@
 interpreters, and we additionally demonstrate abstract interpreters written direct-style with
-delimited controls. 
+delimited control operators. 
 We use the delimited control operators @shift@ and @reset@ that are implemented in Scala 
-\cite{rompf2009implementing}. The correspondence of delimited controls (as well as continuations) 
+\cite{rompf2009implementing}. The correspondence of delimited control (as well as continuations) 
 and monads are well-known \cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions}.

### 2018-06-23 paper/paper.tex
abstract revised

@@
   interpreters.
-  The transformations include linearization, fusing, disentangling, refunctionalizing,
-  and un-CPS to direct-style with delimited control. Linearization expresses
-  non-deterministic choices by first-order data types, after which refunctionalization
-  sequentializes the evaluation order by higher-order functions.
+  The transformations include linearization, lightweight fusion, disentanglement, 
+  refunctionalization, and finally to direct-style with delimited control. 
+  Linearization expresses nondeterministic choices by first-order data types, 
+  after which refunctionalization transforms the datatype representing continuations
+  into higher-order functions.
+  Following the idea that in deterministic languages, reduction contexts of reduction
+  semantics are defunctionalized continuations, we further apply this idea to AAM --
+  reduction contexts of it are refunctionalized to extended continuation-passing style.
+
   All transformations properly handle the collecting semantics and the nondeterminism 
-  of abstract interpretation.
-
-  Following the idea that in deterministic languages, evaluation contexts of reduction
-  semantics are defunctionalized continuations, we further show that in nondeterministic 
-  languages, evaluation contexts are refunctionalized to extended continuations style.
-  Remarkably, we reveal how precise call/return matching in control-flow analysis is 
-  obtained by refunctionalizing a small-step abstract abstract machine with proper caching.
+  of abstract interpretation. Remarkably, we reveal how precise call/return matching in 
+  control-flow analysis can be obtained by refunctionalizing a small-step abstract abstract 
+  machine with proper caching.
 \end{abstract}
@@
 then the consequence is we have no places to save the context information. 
-But refunctionalization not only forces us to sequentialize the order non-deterministic 
+But refunctionalization not only forces us to sequentialize the order nondeterministic 
 evaluation through higher-order functions representing continuations, 
@@
 when specifying a programming language by an interpreter \cite{Reynolds:72, Reynolds:HOSC98-revisited}.
-Refunctionalization is the reverse of this transformation.
+Refunctionalization is the left inverse of this transformation.
 

### 2018-06-23 paper/paper.tex
some revision on Danvy's comments

@@
   Following the idea that in deterministic languages, reduction contexts of reduction
-  semantics are defunctionalized continuations, we further apply this idea to AAM --
-  reduction contexts of it are refunctionalized to extended continuation-passing style.
+  semantics are defunctionalized continuations, we further apply this idea to nondeterministic
+  AAM -- reduction contexts of it are refunctionalized to extended continuation-passing style.
 
@@
 interpreters, respectively. An interesting question, and the subject of this paper, is 
-how we can interderive these abstract semantics artifacts from the respective other one.
+how we can interderive these abstract semantic artifacts from the respective other one.
 
@@
 
-In the domain of abstract semantics artifacts, by contrast, the relations between small-step
+In the domain of abstract semantic artifacts, by contrast, the relations between small-step
 abstract abstract machines and big-step abstract definitional interpreters, as well as the 
 question of deriving one from the other, are not well studied.
-The fundamental difference between concrete semantics artifacts and abstract semantics
-artifacts is nondeterminism. In addition to ensuring termination, abstract semantics
-artifacts are usually equipped with a cache of explored state that is guaranteed to reach
-the least fixed-point eventually.
+One of the fundamental differences between concrete semantic artifacts and abstract semantic
+artifacts is nondeterminism. In addition to ensuring termination, abstract semantic
+artifacts are usually equipped with a cache of reachable states, the cache is
+updated in a monotonic way and guaranteed to reach the least fixed-point eventually.
 In this functional pearl, we provide a constructive answer to the question of interderiving
-abstract semantics artifacts and relate AAM and ADI by presenting a series of syntactical
+abstract semantic artifacts and relate AAM and ADI by presenting a series of syntactical
 transformations on the program from small-step abstract abstract machines to big-step 
@@
        \begin{smallmatrix} \text{refunctionalized}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
-       \begin{smallmatrix} \text{monadic} \\ \text{abstracting} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
+       \begin{smallmatrix} \text{monadic} \\ \text{abstract} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
        \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} &
@@
        [start branch=C] \chainin (m-3-5)
-       [join={node[above, labeled] {\text{Known as Danvy's correspondences for concrete semantics artifacts }}}];
+       [join={node[above, labeled] {\text{Known as Danvy's correspondences for concrete semantic artifacts }}}];
      }
@@
 We then address the main contribution of this paper, which is the filling of the gap between
-small-step AAM and big-step ADI by developing a series of systematic transformations. 
+small-step AAM and big-step ADI by applying a series of well-known systematic transformations
+drawn from concrete semantic artifacts.
 Figure \ref{fig:trans} shows the transformations.
@@
 
-  \item In Section~\ref{fusing}, we then present the \textbf{fusing} transformation.
+  \item In Section~\ref{fusing}, we then present the \textbf{lightweight fusion} transformation.
     The fusing transformation simply combines the single-step function @step@ and the 
@@
 
-  \item Section~\ref{disen} discusses \textbf{disentangling}, which disassembles the 
+  \item Section~\ref{disen} discusses \textbf{disentanglement}, which disassembles the 
     fused AAM to be several individual functions, with each function handling one data 
@@
 %   We show that the correspondence between abstract
-%   machines and interpreters exists not only in concrete semantics artifacts,
-%   but also in abstract semantics artifacts.
+%   machines and interpreters exists not only in concrete semantic artifacts,
+%   but also in abstract semantic artifacts.
 %   Galois connection?
@@
 call-site, i.e., the information that can be used to resume the interrupted computation.
-A @Frame@ constitutes a variable name @x@ to be bound later, a control expression
+A @Frame@ constitutes a variable name @x@ to be bound later, and a control component
 to which the program may resume, and its environment.
@@
 
-\item If the state is not a @Let@ expression, then it must be an atomic expression, as seen
-  in the above code. In this scenario, we begin by extracting the top frame of all available
-  continuations. The control (i.e., an atomic expression) of the current state is the 
-  evaluated term that is being bound to the variable @x@ from the top frame.
-  The environment and store are updated with @x@ mapping to the closure value of @ae@.
+\item If the control component is not a @Let@ expression, then it must be an atomic expression,
+  as seen in the above code. In this scenario, we begin by extracting the top frame of
+  all available continuations. The variable @x@ from the top frame will be bound to
+  the result of evaluating the atomic expression @ae@ by updating the environment and
+  store.
   Finally, the successor state is transferred to expression @e@ from the top frame,
@@
 
-Similar to CESK machines, the machine state of AAM has a control expression,
-an environment, a store, and continuation, as well as a timestamp.
-However, there are several notable differences between AAM's store and CESK machine's store.
-In AAM, the store maps addresses to sets of values; it stores all possible values
-for a particular address. As such, dereferencing addresses becomes nondeterministic.
-Also, the store performs \emph{joining}, rather than overwriting, when updating elements.
-Furthermore, the continuations are likewise allocated on the store instead of formed into a
-linked list, and the continuations are in a state which then turns into a continuation
-address.
+Similar to CESK machines, the machine state of AAM has a control, an environment, a store,
+and continuation, as well as a timestamp. However, there are several notable differences
+between AAM's store and CESK machine's store. In AAM, the store maps addresses to sets of
+values; it stores all possible values for a particular address. As such, dereferencing
+addresses becomes nondeterministic. Also, the store performs \emph{joining}, rather than
+overwriting, when updating elements. Furthermore, the continuations are likewise allocated
+on the store instead of formed into a linked list, and the continuation component becomes
+an address that maps to a set of continuations in the store instead of directly embedded
+in the state.
 
@@
 To effectively approximate the runtime behavior, we introduce a finite
-program contour @time@ that encodes the program execution history.
+program contours @time@ that encodes the program execution history.
 The function @tick@ is used to refresh the "time" and get a "new time".
-We use a list of execution contexts (expressions) to represent the finite 
-program contour, and as we will see in Section~\ref{kcfainst}, by applying 
-different @tick@ functions on the timestamp, we are able to obtain a 
-family of analyses.
+We use a list of expression (which are drawn from the control component)
+to encode the calling contexts history, and as we will see in Section~
+\ref{kcfainst}, by applying different @tick@ functions on the timestamp,
+we are able to obtain a family of analyses.
 
@@
 \begin{itemize}
-  \item If the current control expression is a @Let@, then the result of @App(f, ae)@ will 
+  \item If the current control component is a @Let@, then the result of @App(f, ae)@ will 
     be bound to variable @x@. In this case, we retrieve the set of closures that @f@ may 
@@
 continuations in states, we intend to use a list of frames to explicitly model 
-the stack. By doing so, we recover the call stack as precise as runtimes
+the stack. By doing so, we recover the call stack as the same as runtimes
 (so called pushdown control-flow analysis), but since the stack is unbounded, 
@@
 In the previous section, we show that by keeping an unabstracted stack in the state
-space, we can recover the precise call/return match.
-In this and following sections, we begin describing the transformations step by step.
-Our base machine for now is the AAM with unbounded stack, though
-as we will later transform the stack to higher-order functions
+space, we can recover the precise call/return match. Now we begin describing the
+transformations step by step. Our base machine for now is the AAM with unbounded
+stack, though as we will later transform the stack to higher-order functions
 representing continuations, it does not matter what kind of AAM we start from.
@@
 
-At this point, we have obtain a linearized abstract abstract machine. If we imagine that the
+At this point, we have obtained a linearized abstract abstract machine. If we imagine that the
 classical AAM explores a graph of reachable states, then the linearized AAM flattens the
@@
 
-\section{Fusing} \label{fusing}
-
-Fusing is a transformation that combines the @step@ and @drive@ functions into a single
-function.
+\section{Lightweight Fusion} \label{fusing}
+
+Then we apply a lightweight fusion transformation that combines the @step@ and @drive@
+functions into a single function.
 The fused function @drive_step@ is essentially formed by merging the functionality of
@@
 
-\section{Disentangling} \label{disen}
+\section{Disentanglement} \label{disen}
 
@@
 The code for the second case is replaced entirely by a function call to @continue@.
-The astute reader will notice a vague shape of interpreter in continuation style has emerged.
+The shape of an interpreter in continuation-passing style has emerged.
+%The astute reader will notice a vague shape of interpreter in continuation style has emerged.
 
@@
 matches the calls and returns.
-Refunctionalization makes the interpreter be a ``big-step'' semantics artifact, 
+Refunctionalization makes the interpreter be a ``big-step'' semantic artifact, 
 then the consequence is we have no places to save the context information. 
@@
 that can be traced to Jones \cite{jones1981flow}. \citeauthor{Shivers:1988:CFA:53990.54007} 
-introduces $k$-CFA which uses $k$ recent calling contexts as program contour that 
-differentiates values from different contexts \cite{Shivers:1988:CFA:53990.54007, 
+introduced $k$-CFA which uses $k$ recent calling contexts as program contours that 
+differentiate values from different contexts \cite{Shivers:1988:CFA:53990.54007, 
 Shivers:1991:SSC:115865.115884}.
@@
 technique that can be used to transform higher-order functions to first-order functions
-when specifying a programming language by an interpreter \cite{Reynolds:72, Reynolds:HOSC98-revisited}.
-Refunctionalization is the left inverse of this transformation.
+when writing a definitional interpreter \cite{Reynolds:72, Reynolds:HOSC98-revisited}.
+Refunctionalization is a left inverse of this transformation.
 
@@
 where the reduction context is a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}.
+% Danvy also shows in nondeterministic languages ...
 Our work further shows that for nondeterministic languages, there is another layer of continuations
@@
 
-Fusion as one of the transformation in our paper is also studied by 
+Lightweight fusion as one of the transformation in our paper is also studied by 
 \citeauthor{Ohori:2007:LFF:1190216.1190241} \cite{Ohori:2007:LFF:1190216.1190241}. \\
@@
 but CFA2 has several drawbacks: it works only on continuation-passing style programs,
-and does not supports polyvariant analysis, in addition to an exponential time
+and does not supports polyvariant analysis, in addition to having an exponential time
 complexity.
@@
 In this functional pearl, we fill the gap between small-step abstract abstract machines 
-and big-step abstract definitional interpreters by developing a series of syntactical 
+and big-step abstract definitional interpreters by applying a series of syntactical 
 transformations. Among these transformations, linearization turns a worklist into another
@@
 
-We show the functional correspondence not only exists between concrete semantics artifacts
-but also exists between abstract semantics artifacts. An interesting open question would
+We show the functional correspondence not only exists between concrete semantic artifacts
+but also exists between abstract semantic artifacts. An interesting open question would
 be whether there also exists a correspondence of static analyses formalized in different 

### 2018-06-23 paper/paper.tex
fix misc mistakes

@@
   Languages}, he mentions that for deterministic languages, a reduction semantics
-is a structural operational semantics in continuation style, where the reduction
+is a structural operational semantics in continuation-passing style, where the reduction
 context is a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}.
@@
 The shape of an interpreter in continuation-passing style has emerged.
-%The astute reader will notice a vague shape of interpreter in continuation style has emerged.
 
@@
 Readers familiar with functional programming may think of @fold@ at this point; indeed,
-it is a variant implementation of @fold@ in continuation style. But here we have two
+it is a variant implementation of @fold@ in continuation-passing style. But here we have two
 continuations, one for the next element in the collection, and one for the eventual result.
@@
 
-The function @analyze@ is also changed by adding a @reset@ operator around the function 
-call to @aeval@ to set the delimiter. The afterwards updating on cache becomes sequential.
+The function @analyze@ is also changed by calling a @reset@ operator around the function 
+call of @aeval@ to set the delimiter. The afterwards updating on cache becomes sequential.
 
@@
 
-Now we have arrived the end of this series of transformations. Starting from an 
-small-step abstract abstract machine, eventually we obtain an big-step definitional
+Now we have arrived the end of this series of transformations. Starting from a
+small-step abstract abstract machine, eventually we obtain a big-step definitional
 interpreter with pushdown control-flows and written in direct-style.
@@
 For deterministic languages, \citeauthor{Danvy:2008:DIP:1411204.1411206} shows
-a reduction semantics is a structural operational semantics in continuation style,
+a reduction semantics is a structural operational semantics in continuation-passing style,
 where the reduction context is a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}.
@@
 for Free (P4F). Based on the AAM we presented in Section \ref{aam}, the only changes in the 
-code required is not only keeping track of the entry-point expression of callee, but also 
+code required is not only keeping track of the entry-point expression of the callee, but also 
 holding the target environment when allocating a continuation addresses.
@@
 The difference on the surface is that ours works for A-Normal Form $\lambda$-calculus,
-instead of plain $\lambda$-calculus, although our work can also be easily adopted to handle a plain
+instead of plain $\lambda$-calculus, although our work can also be easily adapted to handle a plain
 $\lambda$-calculus.

### 2018-06-24 paper/paper.tex
some revisions

@@
   Linearization expresses nondeterministic choices by first-order data types, 
-  after which refunctionalization transforms the datatype representing continuations
+  after which refunctionalization transforms the data type representing continuations
   into higher-order functions.
@@
      [start chain] \chainin (m-2-1);
-     \chainin (m-2-2) [join={node[above,labeled] {\begin{smallmatrix} \text{linearizing} \\ \text{+} \\ \text{fusing} \end{smallmatrix}}     }];
-     \chainin (m-2-3) [join={node[above,labeled] {\text{disentangling}}}];
+     \chainin (m-2-2) [join={node[above,labeled] {\begin{smallmatrix} \text{linearization} \\ \text{+} \\ \text{fusion} \end{smallmatrix}}     }];
+     \chainin (m-2-3) [join={node[above,labeled] {\text{disentang.}}}];
      \chainin (m-2-4) [join={node[above,labeled] {\text{refunc.}}}];
@@
 
-  \item In Section~\ref{fusing}, we then present the \textbf{lightweight fusion} transformation.
+  \item In Section~\ref{fusing}, we then present the \textbf{lightweight fusion}.
     The fusing transformation simply combines the single-step function @step@ and the 
@@
     fused AAM to be several individual functions, with each function handling one data 
-    type represented by a continuation.
+    type representing continuation.
 
@@
 
-% \subsection{Contributions}
-
-% the essence of abstracting definitional interpreter.
-
-% We summarize our contributions as follows:
-
-% \begin{itemize}
-% \item The main contribution of this paper is the filling of the gap
-%   between small-step AAM and big-step ADI by developing a series of
-%   systematic transformations.
-%   We show that the correspondence between abstract
-%   machines and interpreters exists not only in concrete semantic artifacts,
-%   but also in abstract semantic artifacts.
-%   Galois connection?
-%   \note{JD: Section number?}
-
-% \item A new understanding of pushdown control flow analysis by
-%   refunctionalization.
-%   We present an abstract interpreter obtained from refunctionalizing
-%   a small-step abstracting abstract machine.
-%   This abstract interpreter naturally exhibits the pushdown control-flow property,
-%   as the refunctionalization forces us to sequentialize the
-%   abstract execution.
-%   We also analyze its correctness, soundness, and termination.
-
-% \item Further, we show that defunctionalization and refunctionalization
-%   of abstract interpreters play important roles for the call stack of the analyzed language.
-%   Given this insight, we reveal precisely \emph{why} small-step AAM has no pushdown
-%   control-flow property (without extra efforts), as well as explain
-%   why ADI naturally extends pushdown control-flow property from its
-%   defining language.
-
-% \end{itemize}
-
-% \subsection{Outline}
-
-% \note{TR: can this be merged with contributions? (they currently lack section refs)}
-% The remainder of this paper is organized as follows.
-% We begin by describing the target language we will analyze in Section~\ref{background},
-% including a concrete CESK machine as small-step operational semantics for it.
-% We continue with a review of abstracting abstract machines (AAM) as an abstract interpretation
-% counterpart of CESK machines. Following this, Section~\ref{linear} shows a conversion from the
-% nondeterministic choices of closures to another layer of continuations,
-% therefore linearizing the abstract execution.
-% Section~\ref{fusing} then discusses fusing the driver and step functions to a single
-% function.
-% We further identify in Section~\ref{disen} the first-order data type representations of
-% continuations, and disentangle the code blocks for dispatching data types to separate
-% functions.
-% Section~\ref{refunc} then represents the refunctionalized abstract abstract machines.
-% In this section, the continuations represented by data types become higher-order
-% functions.
-% In Section~\ref{directstyle}, we transform the refunctionalized AAM to two
-% different direct styles: one implemented with monads, the other using
-% delimited control operators.
-
-% Finally, we discuss related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.
-
-% \note{TR: put this into a proper figure, add known danvy sequence
-% for concrete evaluators at the bottom (in grey or dashed lines)}
-
-%\includegraphics[width=100mm,scale=1.0]{figures/lines.png}
-
 \iffalse
@@
 There are two main reasons we use a real-world language:
-1) The code does not diminish the accuracy of the material
-than formal and mathematical notations,
-which are heavily used in other static analysis or semantics papers.
+1) The code does not diminish the accuracy of the material than formal and mathematical
+notations, which are heavily used in other static analysis or semantics papers.
 The Scala code in this paper can be easily back-translated into formal notations.
@@
 @[KAddr, Cont]@, we obtain @BStore@ and @KStore@, respectively.
-
-We note that both the value store and continuation store are monotonic;
-it continuously grows and never shrinks. This property guarantees that
-a fixed point of store can always be reached through Kleene's
-iteration when analyzing the program
+We note that both the value store and continuation store are uodated monotonically;
+it continuously grows and never shrinks.
 
@@
 As a consequence, the components of states are also changed: the store is divided
-into binding stores and continuation stores; the continuation stack becomes an address.
+into binding stores and continuation stores; the continuation becomes an address
+that maps to a set of continuations in @KStore@.
 By dereferencing this address in a continuation store, we can retrieve the actual 
-transfer of control. The definition of environment @Env@ remains the same.
+transfers of control. The definition of environment @Env@ remains the same.
 
@@
 \subsubsection{Allocating Addresses}
-Up to this point, we have not described allocating addresses in stores,
-nor handling the time stamp @Time@.
-In abstract interpretation, however, these are key ingredients to achieve
-analyses with different sensitivities, as well as to perform a finite
-state space analysis\cite{Gilray:2016:ACP:2951913.2951936}.
-To effectively approximate the runtime behavior, we introduce a finite
-program contours @time@ that encodes the program execution history.
-The function @tick@ is used to refresh the "time" and get a "new time".
-We use a list of expression (which are drawn from the control component)
-to encode the calling contexts history, and as we will see in Section~
-\ref{kcfainst}, by applying different @tick@ functions on the timestamp,
-we are able to obtain a family of analyses.
+Up to this point, we have not described allocating addresses in stores, nor handling
+the time stamp @Time@. In abstract interpretation, however, these are key ingredients
+to achieve analyses with different sensitivities, as well as to perform a finite
+state space analysis \cite{Gilray:2016:ACP:2951913.2951936}.
+To effectively approximate the runtime behavior, we introduce a finite program contours
+@time@ that encodes the program execution history. The function @tick@ is used to refresh
+the "time" and get a "new time". We use a list of expression (which are drawn from the
+control component) to encode the calling contexts history, and as we will see in Section~
+\ref{kcfainst}, by applying different @tick@ functions on the timestamp, we are able to
+obtain a family of analyses.
 
@@
 1) @Halt@ which corresponds to the empty stack, and
-2) @ContAddr@ consists of the target expressions of callee, which are also
-finite.
+2) @ContAddr@ consists of the target expressions of callee, which are also finite.
 %One may ask, why not also keep track of the time in continuation addresses?
@@
 \end{lstlisting}
+
+Given the space of addresses are finite, we can conclude that there are finite number
+of environments and stores because the number of variables and closures are also finite.
+This property guarantees a finite space of reachable states, and we can always have a terminated
+analysis through Kleene's fixed-point iteration.
 
@@
 The reason we show it here is to keep the consistency of abstract semantics during 
-transformations -- since the final abstract definitional interpreters simply inherit
+transformations -- since the eventual abstract definitional interpreters simply inherit
 a precise control-flow from its defining language, we also would like to start from a
@@
 
-\textit{Note on Computability.}
-Unfortunately, even though other components in the states are finite,
-this AAM with an unbounded stack is still not computable.
+\textit{Note on Termination.}
+Unfortunately, even though other components in the states are finite, this AAM with an unbounded
+stack still may diverge when analyzing some programs.
 This is because the unbounded stack can grow to arbitrary depth which implies that the state
@@
 
-Function @f1@ and @f2@ mutually invoke each other, so the stack will
-alternate pushing frames @f1@ and @f2@ onto the top of current stack.
-However, no two existing stack components are identical in the state space.
+Function @f1@ and @f2@ mutually invoke each other, so the stack will alternately pushing
+frames @f1@ and @f2@ onto the top of current stack. However, no two existing stack
+components are identical in the state space. One can decide the reachability of a state
+through Dyck state graph \cite{earl2010pushdown, earl2012introspective}, but simply
+caching the explored states (i.e. @seen@) would not always make the analyzer terminate.
 
@@
 
-With fusing complete, our AAM
-appears similar to a ``big-step'' interpreter, although it still has the machine
-state representation inside.
-In the disentangling transformation, we
-identify the first-order data types which represent evaluation contexts
+With fusing complete, our AAM appears similar to a ``big-step'' interpreter, although it
+still has the machine state representation inside. In the disentanglement transformation,
+we identify the first-order data types which represent evaluation contexts
 and lift the code blocks that handle these evaluation contexts to be individual functions.
@@
 
-Function @continue@ is mainly what we have seen for the second case in
-@drive_step@, except that @ndcontinue@ is called when
-the normal continuation is empty.
+Function @continue@ is mainly what we have seen for the second case in @drive_step@,
+except that @ndcontinue@ is called when the normal continuation is empty.
 Since we are analyzing a language in ANF, the first-order data representation of
-continuations is simplified to a Scala @List@.
-An empty list @Nil@ represents the halt of a computation path, whereas
-a cons @::@ means the top frame of the list holds the binding variable and
-evaluation context.
-An analogy to this in standard CEK machines would be an abstract class (e.g., @cont@) with
-three variants: @halt@, @arg(value, cont)@, and @arg(exp, env, cont)@.
+continuations is simplified to a Scala @List@. An empty list @Nil@ represents the
+halt of a computation path, whereas a cons @::@ means the top frame of the list
+holds the binding variable and evaluation context.
+An analogy to this in standard CEK machines would be an abstract class (e.g., @cont@)
+with three variants: @halt@, @arg(value, cont)@, and @arg(exp, env, cont)@.
 

### 2018-06-24 paper/paper.tex
add non-final cases description

@@
 which when given a machine state, determines its successor state.
+The function @step@ is a partial function that only handles non-final states,
+which must have a successor; the final case of a state is handled by funciton
+@drive@, which will be mentioned in the next section.
 
@@
 
-As shown and previously discussed, we examine the only two cases which the state may be.
+As shown and previously discussed, we examine the only two non-final cases which the state
+may be.
 

### 2018-06-24 paper/paper.tex
a small revision
fix code

@@
 it can be easily extended to support recursive bindings (such as @letrec@), conditionals, 
-primitive types, and operations on primitive types. These cases would be trivial to implement,
-so we elide them in this paper.
+primitive types, and operations on primitive types. These cases would be straightforward to
+implement without introducing new transformations as far as the concerns of this paper,
+so we elide them here.
+% TODO: appendix, direct-style, numbers and if0
 
@@
   case class Store[K,V](map: Map[K, Set[V]]) {
-    def apply(addr: K): List[V] = map(addr).toList
+    def apply(addr: K): Set[V] = map(addr)
     def update(addr: K, d: Set[V]): Store[K,V] = {

### 2018-06-25 paper/paper.tex
revised section 8

@@
 
-\item If the control component is not a @Let@ expression, then it must be an atomic expression,
-  as seen in the above code. In this scenario, we begin by extracting the top frame of
-  all available continuations. The variable @x@ from the top frame will be bound to
-  the result of evaluating the atomic expression @ae@ by updating the environment and
-  store.
+\item If the control component is not a @Let@ expression, then it must be an atomic
+  expression, as seen in the above code. In this scenario, we begin by extracting the
+  top frame of all available continuations. The variable @x@ from the top frame will
+  be bound to the result of evaluating the atomic expression @ae@ by updating the
+  environment and store.
   Finally, the successor state is transferred to expression @e@ from the top frame,
@@
 of environments and stores because the number of variables and closures are also finite.
-This property guarantees a finite space of reachable states, and we can always have a terminated
-analysis through Kleene's fixed-point iteration.
+This property guarantees a finite space of reachable states, and we can always have a
+terminated analysis through Kleene's fixed-point iteration.
 
@@
 %probably Store Widening \& Abstract Garbage Collection
-\section{To Direct-Style} \label{directstyle}
+\section{Back to Direct-Style} \label{directstyle}
 
@@
 
-\subsection{Back to Direct-Style with Delimited Control} \label{uncps}
+\subsection{Back to Direct-Style with Control Operators} \label{uncps}
 
@@
 \textit{Abstract Interpretation and Control-Flow Analysis.}
-\citeauthor{cousot1977abstract} invented abstract interpretation as a sound approach
-to approximate a program's runtime behavior\cite{cousot1977abstract}.
-Control-flow analysis is one instance of abstract interpretation on functional programs
-that can be traced to Jones \cite{jones1981flow}. \citeauthor{Shivers:1988:CFA:53990.54007} 
-introduced $k$-CFA which uses $k$ recent calling contexts as program contours that 
-differentiate values from different contexts \cite{Shivers:1988:CFA:53990.54007, 
-Shivers:1991:SSC:115865.115884}.
+\citeauthor{cousot1977abstract} invented abstract interpretation as a sound approach to
+approximate a program's runtime behavior \cite{cousot1977abstract}. Control-flow analysis
+is one instance of abstract interpretation on functional programs that can be traced to
+Jones \cite{jones1981flow}. \citeauthor{Shivers:1988:CFA:53990.54007} introduced $k$-CFA
+which uses $k$ recent calling contexts as program contours that differentiate values from
+different contexts \cite{Shivers:1988:CFA:53990.54007, Shivers:1991:SSC:115865.115884}.
 The recent formulation of control-flow analysis is the abstracting abstract machines
 methodology \cite{van2012systematic, van2010abstracting} which forms the starting point
-of this paper.
-Modern programming languages such as Java\cite{might2010resolving} and Racket 
-\cite{Tobin-Hochstadt:2012:HSE:2384616.2384655}
-can also be modeled by the abstracting abstract machine approach. \\
+of this paper. Modern programming languages such as Java \cite{might2010resolving} and
+Racket \cite{Tobin-Hochstadt:2012:HSE:2384616.2384655} can also be modeled by the abstracting
+abstract machine approach. \\
 
 \textit{Defunctionalization and Refunctionalization.}
-Defunctionalization and refunctionalization build connections between abstract
-machines and interpreters.
-\citeauthor{Reynolds:72} first shows defunctionalization as a program transformation 
-technique that can be used to transform higher-order functions to first-order functions
-when writing a definitional interpreter \cite{Reynolds:72, Reynolds:HOSC98-revisited}.
+Defunctionalization and refunctionalization build connections between abstract machines and
+interpreters. \citeauthor{Reynolds:72} first shows defunctionalization as a program transformation 
+technique that can be used to transform higher-order functions to first-order functions when
+writing a definitional interpreter \cite{Reynolds:72, Reynolds:HOSC98-revisited}.
 Refunctionalization is a left inverse of this transformation.
@@
 Over the years, Danvy and his collaborators applied refunctionalization and defunctionalization 
-to many different abstract machines and semantics, including the CEK machines, the CLS machines,
-the SECD machines, etc.
-\cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing, 
+to many different abstract machines and interpreters and build functional correspondences among
+them \cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing,
 Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268,
- danvy2009towards, biernacka2009towards}
+danvy2009towards, biernacka2009towards}. Such semantic artifacts include the CEK machines,
+the CLS machines, the SECD machines, call-by-need evaluators, monadic evaluators, etc.
 Other applications are also found to be applicable, such as Dyck word recognizer, and 
-Dijkstra's shunting-yard algorithm \cite{Danvy:2006:RW:2171265.2171268}.
-
-For deterministic languages, \citeauthor{Danvy:2008:DIP:1411204.1411206} shows
-a reduction semantics is a structural operational semantics in continuation-passing style,
-where the reduction context is a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}.
-% Danvy also shows in nondeterministic languages ...
-Our work further shows that for nondeterministic languages, there is another layer of continuations
-that controls the nondeterministic choices. After identifying and explicitly exposing 
-(Section~\ref{linear}) that layer of nondeterministic continuations, the whole program can 
-then be refunctionalized to extended continuation-passing style.
+Dijkstra's shunting-yard algorithm and even Quicksort algorithm \cite{Danvy:2006:RW:2171265.2171268}.
+
+For deterministic languages, \citeauthor{Danvy:2008:DIP:1411204.1411206} shows a reduction semantics
+is a structural operational semantics in continuation-passing style, where the reduction contexts
+are defunctionalized (data type representing) continuations \cite{Danvy:2008:DIP:1411204.1411206}.
+Not unexpectedly, the correspondence also exists for nondeterministic languages. For example,
+\citeauthor{Danvy:2001:DW:773184.773202} show that refunctionalizing a regualr expressions matcher yields
+its counterpart with two layer of continuations \cite{Danvy:2001:DW:773184.773202}. And the multi-layer
+continuation-passing style is the underlying semantics of control operatos like @reset@ and @shift@
+\cite{Danvy:1990:AC:91556.91622}.
+
+Our work applies these ideas to the nondeterministic AAM. The additional layer of continuation controls
+the nondeterministic choice over closures. After identifying and explicitly exposing (Section~\ref{linear})
+that layer of nondeterministic continuations, the whole program can be refunctionalized to extended
+continuation-passing style, and then back to direct-style by the use of delimited control operators.
 
@@
 In recent years, there are significant efforts~\cite{vardoulakis2010cfa2, earl2012introspective,
-gilray2016pushdown, johnson2015abstracting} to achieve precise call/return
-match based on small-step abstracting abstract machines.
+gilray2016pushdown, johnson2015abstracting} to achieve precise call/return match based on small-step
+abstracting abstract machines.
 
 CFA2 is the first solution that solves the return-flows problem~\cite{vardoulakis2010cfa2},
-but CFA2 has several drawbacks: it works only on continuation-passing style programs,
-and does not supports polyvariant analysis, in addition to having an exponential time
-complexity.
-Pushdown control-flow analysis (PDCFA) is a mechanism which maintains this precision
-through the use of a Dyke state graph representing all possible stacks contained
-within the unbounded-stack machine\cite{earl2012introspective, earl2010pushdown}.
-Similar to PDCFA, abstracting abstract control (AAC) is another strategy for
-maintaining stack precision \cite{johnson2015abstracting}.
-AAC functions by utilizing continuations which are specific to both
-the source and target states of a call-site transition,
-which guarantees that no spurious merging will occur during returns.
+but CFA2 has several drawbacks: it works only on continuation-passing style programs, and
+does not support polyvariant analysis, in addition to having an exponential time complexity.
+Pushdown control-flow analysis (PDCFA) is a mechanism which maintains this precision through
+the use of a Dyke state graph representing all possible stacks contained within the unbounded-stack
+machine \cite{earl2012introspective, earl2010pushdown}. Similar to PDCFA, abstracting abstract
+control (AAC) is another strategy for maintaining stack precision \cite{johnson2015abstracting}.
+AAC functions by utilizing continuations which are specific to both the source and target states
+of a call-site transition, which guarantees that no spurious merging will occur during returns.
 
 \citeauthor{gilray2016pushdown} further proposed a polyvariant continuation-addresses allocator
-for small-step AAM to achieve pushdown analysis \cite{gilray2016pushdown}.
-This method is both simple to implement and computationally inexpensive and so called Pushdown 
-for Free (P4F). Based on the AAM we presented in Section \ref{aam}, the only changes in the 
-code required is not only keeping track of the entry-point expression of the callee, but also 
-holding the target environment when allocating a continuation addresses.
-No other pieces of code would need to be modified.
-Notably, this change only causes constant-factor increase in time complexity of
+for small-step AAM to achieve pushdown analysis \cite{gilray2016pushdown}. This method is both
+simple to implement and computationally inexpensive and so called Pushdown for Free (P4F).
+Based on the AAM we presented in Section \ref{aam}, the only changes in the code required is not
+only keeping track of the entry-point expression of the callee, but also holding the target
+environment when allocating a continuation addresses. No other pieces of code would need to be
+modified. Notably, this change only causes constant-factor increase in time complexity of
 the analysis if the store is widened.
@@
 Our work establish the pushdown control-flow analysis through refunctionalization and a
-proper caching algorithm. The call/return is naturally matched by the stack of 
-the meta language. \\
+proper caching algorithm. The call/return is naturally matched by the call stack of the
+meta-language. \\
 
@@
 Reynold's seminal paper \emph{Definitional interpreters for higher-order programming languages} 
-\cite{Reynolds:72, Reynolds:HOSC98-revisited} shows that in
-definitional interpreters, the defined language can inherit properties from the defining language.
-With this insight, \citeauthor{darais2017abstracting} construct abstracting
-definitional interpreters which automatically inherit the pushdown control-flow
-property from its defining language because the defined language simply uses the stack of the
-meta-language~\cite{darais2017abstracting}. \citeauthor{darais2017abstracting}'s abstracting
-definitional interpreters work on direct-style programs, and is written in monadic style. One
-of the advantages of monadic abstract interpreters is modular and composable, therefore
-deploying different sensitivities or features is just applying a different monad.
-Prior to that, \citeauthor{Sergey:2013:MAI:2491956.2491979} show a monadic
-abstract interpreter for small-step semantics \cite{Sergey:2013:MAI:2491956.2491979}.
-
-This paper is greatly inspired by~\citeauthor{darais2017abstracting}'s work.
-The difference on the surface is that ours works for A-Normal Form $\lambda$-calculus,
-instead of plain $\lambda$-calculus, although our work can also be easily adapted to handle a plain
-$\lambda$-calculus.
-But instead of directly showing the final form of abstract definitional interpreters,
-we reveal that refunctionalization plays an important role for inheriting the
-stack from the metalanguage.
-The target of the series of transformations in this paper is abstract definitional
-interpreters, and we additionally demonstrate abstract interpreters written direct-style with
-delimited control operators. 
-We use the delimited control operators @shift@ and @reset@ that are implemented in Scala 
-\cite{rompf2009implementing}. The correspondence of delimited control (as well as continuations) 
-and monads are well-known \cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions}.
+\cite{Reynolds:72, Reynolds:HOSC98-revisited} shows that in some definitional interpreters, the
+defined language can inherit properties from the defining language. With this insight,
+\citeauthor{darais2017abstracting} construct abstracting definitional interpreters which
+automatically inherit the pushdown control-flow property from its defining language because
+the defined language simply uses the call stack of the meta-language~\cite{darais2017abstracting}.
+\citeauthor{darais2017abstracting}'s abstracting definitional interpreters work on direct-style
+$\lambda$-calculus, and itself is written in monadic style. One of the advantages of monadic
+abstract interpreters is modular and composable, therefore deploying different sensitivities
+or features is just applying a different monad. Prior to that, \citeauthor{Sergey:2013:MAI:2491956.2491979}
+show a monadic abstract interpreter for small-step semantics \cite{Sergey:2013:MAI:2491956.2491979}.
+
+This paper is greatly inspired by~\citeauthor{darais2017abstracting}'s work. The difference on the
+surface is that ours works for A-Normal Form $\lambda$-calculus, instead of plain $\lambda$-calculus,
+although our work can also be easily adapted to handle a plain $\lambda$-calculus. But besides directly
+presenting the final form of abstract definitional interpreters, we reveal that refunctionalization
+plays an important role for inheriting the call stack from the meta-language. The target of the series
+of transformations in this paper is abstract definitional interpreters, and we additionally demonstrate
+abstract interpreters written direct-style with delimited control operators. We use the delimited control
+operators @shift@ and @reset@ that are implemented in Scala \cite{rompf2009implementing}. The
+correspondence of delimited control (as well as continuations) and monads are well-known
+\cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions}.
 

### 2018-06-25 paper/paper.tex
more revisions

@@
   The transformations include linearization, lightweight fusion, disentanglement, 
-  refunctionalization, and finally to direct-style with delimited control. 
+  refunctionalization, and finally back to direct-style with delimited control operators. 
   Linearization expresses nondeterministic choices by first-order data types, 
@@
   semantics are defunctionalized continuations, we further apply this idea to nondeterministic
-  AAM -- reduction contexts of it are refunctionalized to extended continuation-passing style.
+  AAM -- its reduction contexts are refunctionalized to extended continuation-passing style.
 
@@
 continuation-passing style, where the reduction contexts are defunctionalized continuations.
-One can derive definitional interpreters by refunctionalizing the evaluation contexts of 
+One can derive definitional interpreters by refunctionalizing the reduction contexts of 
 abstract machines, and by defunctionalizing the higher-order functions, one may obtain 
@@
   an environment.
-4) \textit{Continuation} represents the program stack. In this paper, we instantiate the 
-  stack as a list of frames because the ANF simplifies the evaluation context. For direct 
+4) \textit{Continuation} represents the program's call stack. In this paper, we instantiate the 
+  stack as a list of frames because the ANF simplifies the evaluation context. For direct % TODO
   style $\lambda$-calculus, the continuations of CEK/CESK machines can be presented by
@@
 It is worth noting that the continuation class @Kont@ is defined as a list of
-frames, where a frame can be considered an evaluation context in reduction semantics.
+frames, where a frame can be considered a reduction context in reduction semantics.
 We represent frames using the @Frame@ class, which stores the information of a single
@@
 
-As previously mentioned, the space of addresses must be finite in AAM.
-Binding addresses are parameterized by variable names and the creation 
-time of the binding, both of which are finite.
-Continuation addresses @KAddr@ has two variants:
+As previously mentioned, the space of states is finite when the space of addresses is finite.
+To make this happen, binding addresses are parameterized by variable names and the creation
+time of the binding, both of which are finite. Continuation addresses @KAddr@ has two variants:
 1) @Halt@ which corresponds to the empty stack, and
@@
 
-Given the space of addresses are finite, we can conclude that there are finite number
-of environments and stores because the number of variables and closures are also finite.
+Given that the space of addresses is finite, we can conclude that there are finite number
+of environments and stores because the numbers of variables and closures are also finite.
 This property guarantees a finite space of reachable states, and we can always have a
@@
 This is because transforming to higher-order functions forces us to sequentialize
-the order of abstract evaluation, and neither requiring construction of a frame on stack,
-nor to allocate continuations in the store.
+the order of abstract evaluation, and neither requiring construction of a frame on
+stack, nor to allocate continuations in the store.
 Thus, our choice to start from an AAM with unbounded stack is motivated simply because
@@
 
-In Danvy's paper \textit{Defunctionalized Interpreters for Programming
-  Languages}, he mentions that for deterministic languages, a reduction semantics
-is a structural operational semantics in continuation-passing style, where the reduction
-context is a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}.
-This poses a problem, as the underlying semantics of AAM is fundamentally nondeterministic.
-But the evaluation context (i.e., the @Frame@ in our program) is \emph{not}
-nondeterministic, and the worklist @todo@ actually implicitly handles the nondeterminism.
-Thus, if we simply refunctionalize the frames to higher-order functions, 
-it does not help us to move toward abstract definitional interpreters.
-
-In order to address this problem, our first step of transformations
-is to linearize all nondeterministic choices.
-This step removes all nondeterminism from the @step@ function.
+The underlying semantics of AAM is fundamentally nondeterministic: there are possibly
+multiple target closures when dereferencing an address, and we have to explore both branches
+if the conditionals exist in the language. But the reduction context (i.e., the @Frame@ in
+our program) is \emph{not} nondeterministic, and the worklist @todo@ actually implicitly
+handles the nondeterminism. Thus, if we simply refunctionalize the frames to higher-order
+functions, it does not help us to move toward abstract definitional interpreters.
+
+In Danvy's paper \textit{Defunctionalized Interpreters for Programming Languages},
+he mentions that for deterministic languages, a reduction semantics is a structural
+operational semantics in continuation-passing style, where the reduction context is
+a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}. Therefore
+refunctionalizing a deterministic language yields an interpreter in continuation-passing
+style; nevertheless, doing so for nondeterministic languages yields an interpreter
+with multiple layers of continuation \cite{Danvy:2006:RW:2171265.2171268}.
+
+We apply this idea to AAM and our first step of transformations is to linearize all
+nondeterministic choices. This step removes all nondeterminism from the @step@ function.
 Likewise, the @Frame@ saves the information of the caller in concrete executions.
@@
 For clarity of presentation in differentiating between the two continuations, we elect to call
-the first a \emph{normal continuation}, and the second a \emph{nondeterministic continuation}.
+the first a \emph{normal continuation}, and the second a \emph{meta-continuation}
+\footnote{Traditionally, they are also called \emph{success continuation} and \emph{failure
+continuation} \cite{Danvy:1990:AC:91556.91622}.}.
 
@@
 
+Here we also use first-order data types to explicitly represent the meta-continuation which
+controls the nondeterminism.
 Each @NDCont@ object contains a list of closures that are possible functions to be invoked,
-a set of values that will be bound to the function's formal argument,
-and the store, time, and frames at the fork point.
-In the definition of @NDState@, an auxiliary function @toState@ is added that
-converts itself to @State@.
+a set of values that will be bound to the function's formal argument, and the store, time,
+and frames at the fork point. In the definition of @NDState@, an auxiliary function @toState@
+is added that converts itself to @State@.
 
@@
 As such, we prepare a new environment, a new store, and a new frame list only for the
-first closure in that set.
-A new nondeterministic continuation @new_ndk@ is also constructed,
-which contains the rest of closures, the values of argument @ae@, the store,
-the time, and the new frame list.
-We use the store before updating, because for different closures they may form
-different binding addresses @baddr@.
-We also use the new frames, because all closures at this fork point share
-the same stack and return point.
-
-In the second case (shown below), we will see how @NDCont@ deals with
-nondeterminism.
+first closure in that set. A new meta-continuation @new_ndk@ is also constructed,
+which contains the rest of closures, the values of argument @ae@, the store, the time,
+and the new frame list. We use the store before updating, because for different closures
+they may form different binding addresses @baddr@. We also use the new frames, because
+all closures at this fork point share the same stack and return point.
+
+In the second case (shown below), we will see how @NDCont@ deals with nondeterminism.
 
@@
 However, since we have added nondeterminism into state, an empty list of frames means
-we have reached the halt of \textit{one computation path}, and we should
-determine the next state indicated by @NDCont@.
-Thus, we have a pattern matching on @ndk@:
+we have reached the halt of \textit{one computation path}, and we should determine the
+next state indicated by @NDCont@. Thus, we have a pattern matching on @ndk@:
+
 \begin{itemize}
@@
 than current one.
-The nondeterministic continuation @new_ndk@ is also updated by the removal of that closure.
+The meta-continuation @new_ndk@ is also updated by the removal of that closure.
 \end{itemize}
@@
 still has the machine state representation inside. In the disentanglement transformation,
-we identify the first-order data types which represent evaluation contexts
-and lift the code blocks that handle these evaluation contexts to be individual functions.
-
-Since there are two layers of continuations,
-we obtain three mutually recursive functions: @drive_step@, which plays the same
-role as before; @continue@, which is called from @drive_step@ when encountering an atomic 
-expression, and handles the evaluation context (@Frame@) of the analyzed
-language; and @ndcontinue@, which dispatches nondeterministic continuations @NDCont@,
-and which is invoked from @continue@ when the normal stack is empty.
+we identify the first-order data types which represent reduction contexts
+and lift the code blocks that handle these reduction contexts to be individual functions.
+
+Since there are two layers of continuations, we obtain three mutually recursive functions:
+@drive_step@, which plays the same role as before; @continue@, which is called from
+@drive_step@ when encountering an atomic expression, and handles the reduction context
+(@Frame@) of the analyzed language; and @ndcontinue@, which dispatches meta-continuations
+@NDCont@, and which is invoked from @continue@ when the normal stack is empty.
 
@@
 Since we are analyzing a language in ANF, the first-order data representation of
-continuations is simplified to a Scala @List@. An empty list @Nil@ represents the
+continuations is simplified to a Scala @List@. An empty list @Nil@ represents the %TODO
 halt of a computation path, whereas a cons @::@ means the top frame of the list
-holds the binding variable and evaluation context.
+holds the binding variable and context.
 An analogy to this in standard CEK machines would be an abstract class (e.g., @cont@)
@@
 For clarity, we first show how to refunctionalize normal continuations and
-nondeterministic continuation without collecting semantics.
+meta-continuation without collecting semantics.
 Then, adding an appropriate caching algorithm to make sure the analysis will
@@
 
-%TODO: better word to replace `accumulate`
-Given a collection of elements of type @T@, an initial accumlated value of type 
+Given a collection of elements of type @T@, an initial accumlator of type 
 @Ans@, a function @k@ that works on element's type @T@ and takes a continuation 
@@
 fact was discovered during this iteration, then we have reached a fixed-point that 
-over-approximates the concrete evaluations, and thus can be returned.
-Otherwise, the next iteration with the new @out@ cache installed at the @in@ cache position 
-will be activated; in the meanwhile, an empty cache will be used as the @out@ cache.
-Note that the continuation of the initial call to @aeval@ sets up the values for the
-initial configuration in the final cache.
+over-approximates the concrete evaluations, and thus the result can be returned.
+Otherwise, the next iteration with the new @out@ cache installed at the @in@ cache
+position will be activated; in the meanwhile, an empty cache will be used as the @out@
+cache. Note that the continuation of the initial call to @aeval@ sets up the values for
+the initial configuration in the final cache.
 
@@
 then the consequence is we have no places to save the context information. 
-But refunctionalization not only forces us to sequentialize the order nondeterministic 
-evaluation through higher-order functions representing continuations, 
-but also drives us to use a new caching algorithm.
+But refunctionalization not only forces us to sequentialize the order of nondeterministic 
+evaluation through higher-order functions representing continuations, but also drives us
+to use a new caching algorithm.
 
@@
 interpreters, and the subsequent statements and expressions become the delimited continuation
-@f@. @choices@ also returns the latest cache to its left-hand side definition.
-The cache is still accumulated to the next call to @choices@ or @aeval@, 
+@f@. The function @choices@ also returns the latest cache to its left-hand side when invoking.
+The cache will be still accumulated to the next call to @choices@ or @aeval@, 
 given the fact that the abstract interpreter should always use the latest cache.
@@
 Dijkstra's shunting-yard algorithm and even Quicksort algorithm \cite{Danvy:2006:RW:2171265.2171268}.
-
 For deterministic languages, \citeauthor{Danvy:2008:DIP:1411204.1411206} shows a reduction semantics
@@
 \citeauthor{Danvy:2001:DW:773184.773202} show that refunctionalizing a regualr expressions matcher yields
-its counterpart with two layer of continuations \cite{Danvy:2001:DW:773184.773202}. And the multi-layer
-continuation-passing style is the underlying semantics of control operatos like @reset@ and @shift@
-\cite{Danvy:1990:AC:91556.91622}.
+its counterpart with two layer of continuations \cite{Danvy:2001:DW:773184.773202}.
+And the multi-layer continuation-passing style is the underlying semantics of control operatos
+like @reset@ and @shift@ \cite{Danvy:1990:AC:91556.91622}.
 
@@
 the nondeterministic choice over closures. After identifying and explicitly exposing (Section~\ref{linear})
-that layer of nondeterministic continuations, the whole program can be refunctionalized to extended
+that layer of meta-continuations, the whole program can be refunctionalized to extended
 continuation-passing style, and then back to direct-style by the use of delimited control operators.

### 2018-06-25 paper/paper.tex
more revisions

@@
 early days of programming languages research \cite{landin1966next, Reynolds:72}.
-Nowadays, even an undergraduate student in computer science is able to build toy 
-languages through interpreters.
+Nowadays, even an undergraduate student in computer science is able to make his own
+programming language by building an interpreter.
 But building a sound abstract interpreter remained an esoteric and difficult task 
@@
 4) \textit{Continuation} represents the program's call stack. In this paper, we instantiate the 
-  stack as a list of frames because the ANF simplifies the evaluation context. For direct % TODO
-  style $\lambda$-calculus, the continuations of CEK/CESK machines can be presented by
-  variants of a data type.
-
+stack as a list of frames.
+
+%For direct style $\lambda$-calculus, the continuations of CEK/CESK machines can be presented by variants of a data type.
 The Scala representations for the components of the CESK machine are as follows:
@@
 frames, where a frame can be considered a reduction context in reduction semantics.
+An empty list represents empty call stack; otherwise, the head frame in the list
+represents the innermost context, and the reduction result of this frame will be used
+to fill the ``hole'' of its following element in the list.
 We represent frames using the @Frame@ class, which stores the information of a single
@@
 except that @ndcontinue@ is called when the normal continuation is empty.
-Since we are analyzing a language in ANF, the first-order data representation of
-continuations is simplified to a Scala @List@. An empty list @Nil@ represents the %TODO
+It works on the list representation of context: An empty list @Nil@ represents the
 halt of a computation path, whereas a cons @::@ means the top frame of the list

### 2018-06-25 paper/paper.tex
fix interderiving

@@
 updated in a monotonic way and guaranteed to reach the least fixed-point eventually.
-In this functional pearl, we provide a constructive answer to the question of interderiving
-abstract semantic artifacts and relate AAM and ADI by presenting a series of syntactical
-transformations on the program from small-step abstract abstract machines to big-step 
-abstract definitional interpreters.
+In this functional pearl, we make a first step to answering the question of interderiving
+abstract semantics artifacts, and provide a constructive answer to relate AAM and ADI 
+by presenting a series of syntactical transformations on the program from small-step 
+abstract abstract machines to big-step abstract definitional interpreters.
 
@@
 It is worth noting that the continuation class @Kont@ is defined as a list of
-frames, where a frame can be considered a reduction context in reduction semantics.
+frames, where a frame can be considered a representation of reduction context 
+in reduction semantics.
 An empty list represents empty call stack; otherwise, the head frame in the list

### 2018-06-25 paper/paper.tex
revise section 6.3

@@
 In the previous section, we have established a computable pushdown control-flow
-analysis through refunctionalization and caching. In this section,
-we revisit the pushdown control-flow problem and examine what we have done to
-overcome it. \\
+analysis through refunctionalization and caching. In this section, we revisit 
+the pushdown control-flow problem and examine what we have done to overcome it. \\
 
 \textit{The Problem with Return-flows.}
-Pushdown control-flow is a property in analysis that precisely models
-the runtime call stack of the analyzed program.
-A pushdown control-flow analysis provides an as-exact-as-runtime return-flow
-when analyzing a program, but traditional control-flow analysis collapses
-the state space into a finite space and thus causes imprecise stack modeling.
+Pushdown control-flow is a property in analysis that precisely models the runtime
+call stack of the analyzed program. A pushdown control-flow analysis provides an 
+as-exact-as-runtime return-flow when analyzing a program, but traditional 
+control-flow analysis collapses the state space into a finite space and thus 
+causes imprecise stack modeling.
 
@@
 In $k$-CFA algorithm or the abstract abstract machine shown in Section~\ref{aam},
-the call sites @(id 2)@ and @(id 1)@ share the same return flow,
-so the invocation of @(id 2)@ returns to both call-sites @[x (id 1)]@ and @[y (id 2)]@.
-Therefore, the returned value @2@ for variable @y@ is also propagated to
-variable @x@, causing imprecise analysis results to arise.
-This return-flow merging is inevitable, even when increasing the context-sensitivity.
-If we use the monovariant analysis (i.e., $0$-CFA), the analysis result would be such 
-that @x@ and @y@ point to value set @{1, 2}@, because the algorithm does not distinguish 
-that @z@ comes from a different call site.
-Under $1$-CFA, the algorithm is able to distinguish that variable @z@ of function
-@id@ has two different values at two call sites, so variable @y@
-would not be polluted by @1@.
-However, variable @x@ still points to value set @{1, 2}@, because two call-sites still 
+the call sites @(id 2)@ and @(id 1)@ share the same return flow, so the invocation
+of @(id 2)@ returns to both call-sites @[x (id 1)]@ and @[y (id 2)]@.  Therefore,
+the returned value @2@ for variable @y@ is also propagated to variable @x@, causing
+imprecise analysis results to arise. This return-flow merging is inevitable, even when
+increasing the context-sensitivity. Even worse, if we use the monovariant analysis (i.e.,
+$0$-CFA), the analysis result would be such that @x@ and @y@ point to value set @{1, 2}@,
+because the algorithm does not distinguish that @z@ (the argument of @id@) comes from a
+different call site. Under $1$-CFA, the algorithm is able to distinguish that variable
+@z@ has two different values at two call sites, so variable @y@ would not be polluted by
+@1@. However, variable @x@ still points to value set @{1, 2}@, because two call-sites still
 share the same continuation and $1$-CFA is still unable to separate the return-flows. \\
@@
 we do not have a stack model. The higher-order functions representing continuations
-already connect all the execution in order. In fact, the continuations (call stack) 
+already connect all the executions in order. In fact, the continuations (call stack)
 of the analyzed language is blended into the call stack of our defining language
@@
 In fact, what we started is an AAM with an unbounded stack which already precisely
-matches the calls and returns.
-Refunctionalization makes the interpreter be a ``big-step'' semantic artifact, 
-then the consequence is we have no places to save the context information. 
-But refunctionalization not only forces us to sequentialize the order of nondeterministic 
-evaluation through higher-order functions representing continuations, but also drives us
-to use a new caching algorithm.
-
-Indeed, the caching algorithm plays an important role to make the analysis computable.
-
+matches the calls and returns. Refunctionalization makes the interpreter be a ``big-step''
+semantic artifact, then the consequence is we have no places to explicitly save the context
+information. But refunctionalization not only forces us to sequentialize the order of
+nondeterministic evaluation through higher-order functions representing continuations,
+but also drives us to use a new caching algorithm.
+
+Indeed, the caching algorithm plays an important role to guarantee the analysis will
+terminate eventually.
 An interesting implication of this is if we apply the caching algorithm with some 
-necessary changes to small-step abstract machines with an unbounded stack, 
-we are also able to establish a computable and precise call/return match.
+necessary changes to small-step abstract machines with an unbounded stack, we are 
+also able to establish a computable and precise call/return match.
 

### 2018-06-26 paper/paper.tex
normal continuation -> continuation

@@
 For clarity of presentation in differentiating between the two continuations, we elect to call
-the first a \emph{normal continuation}, and the second a \emph{meta-continuation}
+the first a \emph{continuation}, and the second a \emph{meta-continuation}
 \footnote{Traditionally, they are also called \emph{success continuation} and \emph{failure
@@
 
-If the normal continuation @konts@ is an empty list, then it means we have reached
+If the continuation @konts@ is an empty list, then it means we have reached
 the end of the computation (i.e., a halt); otherwise, we should return the values of 
@@
 The termination of the analysis occurs when we reach an @NDState@ in which the expression
-is atomic and both the normal continuation and nondeterminism continuation are
-empty lists. This corresponds to the case that @todo@ list is empty.
+is atomic and both the continuation and meta-continuation are empty lists. This corresponds 
+to the case that @todo@ list is empty.
 
@@
 Function @continue@ is mainly what we have seen for the second case in @drive_step@,
-except that @ndcontinue@ is called when the normal continuation is empty.
+except that @ndcontinue@ is called when the continuation @konts@ is empty.
 It works on the list representation of context: An empty list @Nil@ represents the
@@
 
-For clarity, we first show how to refunctionalize normal continuations and
+For clarity, we first show how to refunctionalize continuations and
 meta-continuation without collecting semantics.
@@
 This is sometimes referred to as an extended continuation-passing style (ECPS)
-that function @k@ is a normal continuation and @m@ is a meta-continuation 
+that function @k@ is a continuation and @m@ is a meta-continuation 
 \cite{Danvy:1990:AC:91556.91622}. \\

### 2018-06-26 src/main/scala/refunc/ANFAAMBasis.scala
work on direct-style abs interp

@@ -4,6 +4,8 @@ import scala.util.continuations._
 import refunc.ast._
 
 object ANFAAM {
+  def isValue(e: Expr): Boolean = e.isInstanceOf[Value]
+
   def isAtomic(e: Expr): Boolean =
     e.isInstanceOf[Var] ||
     e.isInstanceOf[Lam] ||

### 2018-06-26 src/main/scala/refunc/ast/Expr.scala

@@ -9,7 +9,9 @@ case class  Num(i: Int) extends Expr
 case class  Var(x: String) extends Expr
 case class  App(e1: Expr, e2: Expr) extends Expr
 case class  Letrec(bds: List[B], body: Expr) extends Expr
-case class  Let(x: String, e: Expr, body: Expr) extends Expr
 case class  Lam(x: String, body: Expr) extends Value
 case class  If0(cnd: Expr, thn: Expr, els: Expr) extends Expr
+
+// not used for now
 case class  PrimOp(op: Symbol, e1: Expr, e2: Expr) extends Expr
+case class  Let(x: String, e: Expr, body: Expr) extends Expr

### 2018-06-26 src/main/scala/refunc/plainlambda/ADIPlainLam.scala

@@ -6,9 +6,9 @@ import refunc.ANFAAM._
 
 object DSAAM {
   trait Cont
-  //case object Mt extends Cont
   case class Ar(e: Expr, env: Env) extends Cont
   case class Fn(lam: Expr, env: Env) extends Cont
+  case class Br(thn: Expr, els: Expr, env: Env) extends Cont
 
   case class State(e: Expr, env: Env, store: BStore, konts: List[Cont], time: Time) {
     def tick: Time = (e :: time).take(k)
@@ -22,18 +22,24 @@ object DSAAM {
 object AAMUBPlainLam {
   import DSAAM._
 
-  def isValue(e: Expr): Boolean = e.isInstanceOf[Value]
-
   def step(s: State): List[State] = {
     val new_time = s.tick
     s match {
+      case State(Num(_), env, store, konts, time) =>
+        List(State(NumV, env, store, konts, new_time))
+
       case State(App(e1, e2), env, store, konts, time) =>
         List(State(e1, env, store, Ar(e2, env)::konts, new_time))
+
       case State(Var(x), env, store, konts, time) =>
         for (v <- atomicEval(Var(x), env, store).toList) yield v match {
-          case Clos(lam, c_env) => State(lam, c_env, store, konts, new_time)
           case NumV => State(NumV, env, store, konts, new_time)
+          case Clos(lam, c_env) => State(lam, c_env, store, konts, new_time)
         }
+
+      case State(If0(cnd, thn, els), env, store, konts, time) =>
+        List(State(cnd, env, store, Br(thn, els, env)::konts, new_time))
+
       case State(v, env, store, konts, time) if isValue(v)=>
         konts match {
           case Fn(Lam(x, body), c_env)::ks =>
@@ -43,7 +49,10 @@ object AAMUBPlainLam {
             List(State(body, new_env, new_store, ks, new_time))
           case Ar(arg, arg_env)::ks if v.isInstanceOf[Lam] =>
             List(State(arg, arg_env, store, Fn(v, env)::ks, new_time))
-          case _ => List() //including cases of type error, or spurious flows
+          case Br(thn, els, br_env)::ks if v == NumV =>
+            List(State(thn, br_env, store, ks, new_time),
+                 State(els, br_env, store, ks, new_time))
+          case _ => List() // including cases of type error, or spurious flows
         }
     }
   }
[…]

### 2018-06-27 paper/paper.tex
need to figure out arrow placement in figure

@@
   abstract interpreters that are derived from concrete small-step abstract machines.
-  Recent progress applies the same idea on definitional interpreters, and obtains
-  big-step abstract definitional interpreters (ADI) written in monadic style.
+  \citeauthor{darais2017abstracting} apply the same idea on monadic definitional interpreters, 
+  and obtain monadic abstract definitional interpreters (ADI) in big-step style.
   Yet, the relations between small-step abstracting abstract machines and big-step 
@@
   Linearization expresses nondeterministic choices by first-order data types, 
-  after which refunctionalization transforms the data type representing continuations
+  after which refunctionalization transforms the first-order datatypes representing continuations
   into higher-order functions. 
-  Following the idea that in deterministic languages, reduction contexts of reduction
-  semantics are defunctionalized continuations, we further apply this idea to nondeterministic
-  AAM -- its reduction contexts are refunctionalized to extended continuation-passing style.
+  The refunctionalized AAM is an abstract interpreter written in continuation-passing style
+  with two layers of continuation.
+
+  %Following the idea that in deterministic languages, reduction contexts of reduction
+  %semantics are defunctionalized continuations, we further apply this idea to nondeterministic
+  %AAM -- its reduction contexts are refunctionalized to continuation-passing style.
 
@@
 Defining a language by building an interpreter for it can be traced to the very
-early days of programming languages research \cite{landin1966next, Reynolds:72}.
-Nowadays, even an undergraduate student in computer science is able to make his own
+early days of programming languages research \cite{landin1966next, Reynolds:72, 
+McCarthy:1960:RFS:367177.367199}.
+Nowadays, even undergraduate students in computer science are able to make their own
 programming language by building an interpreter.
@@
 interpreters, respectively. An interesting question, and the subject of this paper, is 
-how we can interderive these abstract semantic artifacts from the respective other one.
+how we can derive an abstract semantic artifact from the respective other one.
 
@@
 One of the fundamental differences between concrete semantic artifacts and abstract semantic
-artifacts is nondeterminism. In addition to ensuring termination, abstract semantic
+artifacts is a sound latter one must tell you facts about all possible execution paths, and 
+thus nondeterminism is introduced. In addition to ensuring termination, abstract semantic
 artifacts are usually equipped with a cache of reachable states, the cache is
@@
      [start chain] \chainin (m-2-4);
-     \chainin (m-1-5) [join={node[sloped,anchor=center,above,labeled] {\text{un-CPS}}}];
+     \chainin (m-1-5) [double,join={node[sloped,anchor=center,above,labeled] {\text{CPS left-inv.}}}];
      \chainin (m-2-5) [join={node[above,labeled] {\text{}}}];
@@
    }
+   %{ [start chain] \chainin (m-1-5);
+   % \chainin (m-2-4) [join={node[sloped,anchor=center,above,labeled] {\text{CPS trans.}}}]; }
+   %
+   % TODO: find out how to make the arrows visually separate
+   % {
+   %   [start chain] \chainin (m-1-5);
+   %   \chainin (m-2-4) [bend left=5, join={node[sloped,anchor=center,below,labeled] {\text{CPS}}}];
+   % }
    {
@@
      \chainin (m-2-4) [join={node[above,labeled] {\text{refunc.}}}];
-     \chainin (m-2-5) [join={node[above,labeled] {\text{monadify}}}];
+     %\chainin (m-2-5) [join={node[above,labeled] {\text{monadify}}}];
    }
    {
-     [start chain]
+     [start chain, color=gray]
      \chainin (m-3-1);
@@
    {
-     [start chain]
+     [start chain, color=gray]
      \chainin (m-3-5);
@@
     of abstract abstract machines; the transition of machine states therefore becomes 
-    deterministic. Notably, this introduces another layer of control that will also 
-    be refunctionalized later.
-
-  \item In Section~\ref{fusing}, we then present the \textbf{lightweight fusion}.
+    deterministic. Notably, this explicitly expresses another layer of control that 
+    will also be refunctionalized later.
+
+  \item In Section~\ref{fusing}, we then apply a \textbf{lightweight fusion} 
+    \cite{DANVY2008100, Ohori:2007:LFF:1190216.1190241} to the linearized AAM. 
     The fusing transformation simply combines the single-step function @step@ and the 
@@
 
-  \item Section~\ref{disen} discusses \textbf{disentanglement}, which disassembles the 
-    fused AAM to be several individual functions, with each function handling one data 
-    type representing continuation.
-
-  \item \textbf{Refunctionalization} is shown in Section~\ref{sec:refunc} which sequentializes
-    the order of abstract execution by higher-order functions. For clarity, we first present 
-    the refunctionalized AAM without caching. We then adopt a different caching algorithm to 
-    guarantee the termination of abstract interpretation.
-    In this section, we also review pushdown control-flow analysis and examine how computable
-    and precise call/return matching is obtained through these transformations.
-
-  \item The last transformation we show is that of transforming the refunctionalized AAM to a
+  \item Section~\ref{disen} discusses \textbf{disentanglement}. In this part, we identify 
+    different first-order datatypes which represent different layers of continuation, and
+    disassemble their handlers to be separate higher-order functions.
+
+  \item Applying \textbf{refunctionalization} \cite{DANVY2009534, Danvy:2006:RW:2171265.2171268} 
+    to disentangled AAM is shown in Section~\ref{sec:refunc}. 
+    %as the left inverse of Reynolds' defunctionalization \cite{Reynold:72,Reynolds:HOSC98-revisited} 
+    Refunctionalization replaces the first-order datatypes representing continuations with
+    higher-order functions, and associates dispatching logic with proper higher-order functions. 
+    We then derive an abstract interpreter written in continuation-passing style with two
+    layers of continuations due to the nondeterminism.
+
+    For clarity, we first present the refunctionalized AAM without caching. 
+    We then adopt a caching algorithm \cite{darais2017abstracting} to guarantee the termination 
+    of abstract interpretation.
+    In the end of this section, we review the pushdown control-flow analysis and examine how 
+    computable and precise call/return matching is obtained through these transformations.
+
+  \item The last transformation we show is that of transforming the refunctionalized AAM back to a
     \textbf{direct-style} interpreter (Section~\ref{directstyle}) by using delimited control operators.
@@
 These transformations are used throughout the paper, with refunctionalization and
-defunctionalization of abstract interpreters playing important roles for the call stack of
-the analyzed language. By refunctionalization, the call stack of the analyzed language
-is blended into the call stack of the defining language. This provides another perspective to
+defunctionalization of abstract interpreters playing important roles for the stack model of
+the analyzed language. By refunctionalization, the call structure of the analyzed language
+is blended into the call structure of the defining language. This provides another perspective to
 explain why \citeauthor{darais2017abstracting}'s abstract definitional interpreters
@@
 
-In this section, we describe a variant of AAM that allows the stack to be
-unbounded which uses a precise call stack as we did in the concrete CESK machine.
+In this section, we describe a variant of AAM that allows the stack to be unbounded 
+which uses a precise model of call stack as we did in the concrete CESK machine.
 Instead of allocating continuations in the store and embedding addresses of 
 continuations in states, we intend to use a list of frames to explicitly model 
-the stack. By doing so, we recover the call stack as the same as runtimes
+the stack. By doing so, we recover the call structure as the same as runtimes
 (so called pushdown control-flow analysis), but since the stack is unbounded, 
@@
 transformations step by step. Our base machine for now is the AAM with unbounded
-stack, though as we will later transform the stack to higher-order functions
+stack, though as we will later transform the stack structure to higher-order functions
 representing continuations, it does not matter what kind of AAM we start from.
-This is because transforming to higher-order functions forces us to sequentialize
-the order of abstract evaluation, and neither requiring construction of a frame on
-stack, nor to allocate continuations in the store.
+This is because it would not be possible to construct a new frame on the stack, nor
+to allocate continuations in the store.
 Thus, our choice to start from an AAM with unbounded stack is motivated simply because
@@
 refunctionalizing a deterministic language yields an interpreter in continuation-passing
-style; nevertheless, doing so for nondeterministic languages yields an interpreter
-with multiple layers of continuation \cite{Danvy:2006:RW:2171265.2171268}.
+style. Nevertheless, doing so for nondeterministic languages yields an interpreter
+with multiple layers of continuation \cite{Danvy:2006:RW:2171265.2171268, DANVY2009534},
+where the additional layer of continuation represents the nondeterministic choices.
 
@@
 
-Refunctionalization transforms first-order data types representing
-evaluation contexts to higher-order functions.
-Besides sequentializing the order of abstract evaluation by functions,
-this transformation can be also regarded as allocating continuation functions of
-the defined language in the heap of the metalanguage.
+Refunctionalization transforms first-order data types representing evaluation contexts 
+to higher-order functions, and associates the dispatching logic of continuations with
+proper higher-order functions.
+This transformation can be also regarded as making the control flows of the defined 
+language be the part of control flows of the meta-language.
+%allocating continuation functions of the defined language in the heap of the meta-language.
 
@@
 Pushdown control-flow is a property in analysis that precisely models the runtime
-call stack of the analyzed program. A pushdown control-flow analysis provides an 
+call structure of the analyzed program. A pushdown control-flow analysis provides an 
 as-exact-as-runtime return-flow when analyzing a program, but traditional 
@@
 semantic artifact, then the consequence is we have no places to explicitly save the context
-information. But refunctionalization not only forces us to sequentialize the order of
-nondeterministic evaluation through higher-order functions representing continuations,
-but also drives us to use a new caching algorithm.
+information in store or stack. 
+But refunctionalization not only forces the nondeterministic control flow to be higher-order,
+but also drives us to use a different caching algorithm.
 
@@
 Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268,
-danvy2009towards, biernacka2009towards}. Such semantic artifacts include the CEK machines,
+DANVY2009534, danvy2009towards, biernacka2009towards}. Such semantic artifacts include the CEK machines,
 the CLS machines, the SECD machines, call-by-need evaluators, monadic evaluators, etc.
 Other applications are also found to be applicable, such as Dyck word recognizer, and 
-Dijkstra's shunting-yard algorithm and even Quicksort algorithm \cite{Danvy:2006:RW:2171265.2171268}.
+Dijkstra's shunting-yard algorithm and even Quicksort algorithm \cite{Danvy:2006:RW:2171265.2171268,
+DANVY2009534}.
 For deterministic languages, \citeauthor{Danvy:2008:DIP:1411204.1411206} shows a reduction semantics
@@
 Lightweight fusion as one of the transformation in our paper is also studied by 
-\citeauthor{Ohori:2007:LFF:1190216.1190241} \cite{Ohori:2007:LFF:1190216.1190241}. \\
+\citeauthor{Ohori:2007:LFF:1190216.1190241} \cite{Ohori:2007:LFF:1190216.1190241}.
+As shown by \citeauthor{DANVY2008100}, lightweight fusion can be used to show the 
+equivalence between small-step and big-step abstract machines \cite{DANVY2008100}.\\
 
@@
 automatically inherit the pushdown control-flow property from its defining language because
-the defined language simply uses the call stack of the meta-language~\cite{darais2017abstracting}.
+the defined language simply uses the call-stack model of the meta-language~\cite{darais2017abstracting}.
 \citeauthor{darais2017abstracting}'s abstracting definitional interpreters work on direct-style
@@
 
-This paper is greatly inspired by~\citeauthor{darais2017abstracting}'s work. The difference on the
-surface is that ours works for A-Normal Form $\lambda$-calculus, instead of plain $\lambda$-calculus,
-although our work can also be easily adapted to handle a plain $\lambda$-calculus. But besides directly
-presenting the final form of abstract definitional interpreters, we reveal that refunctionalization
-plays an important role for inheriting the call stack from the meta-language. The target of the series
-of transformations in this paper is abstract definitional interpreters, and we additionally demonstrate
-abstract interpreters written direct-style with delimited control operators. We use the delimited control
-operators @shift@ and @reset@ that are implemented in Scala \cite{rompf2009implementing}. The
-correspondence of delimited control (as well as continuations) and monads are well-known
+This paper is greatly inspired by~\citeauthor{darais2017abstracting}'s work \cite{darais2017abstracting}.
+The difference on the surface is that ours works for A-Normal Form $\lambda$-calculus, instead of plain 
+$\lambda$-calculus, although our work can also be easily adapted to handle a plain $\lambda$-calculus. 
+But besides directly presenting the final form of abstract definitional interpreters, we reveal that 
+refunctionalization plays an important role for inheriting the model of the stack from the meta-language. 
+The target of these transformations in this paper is abstract definitional interpreters, and we 
+additionally demonstrate abstract interpreters written direct-style with delimited control operators. 
+We use the delimited control operators @shift@ and @reset@ that are implemented in Scala 
+\cite{rompf2009implementing}. The correspondence of delimited continuations and monads are well-known
 \cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions}.
@@
 layer of continuations; refunctionalization converts the first-order data types representing
-continuations to higher-order functions; and finally un-CPS with delimited control transforms
-the abstract interpreter into a sequence of expressions, which looks no different to a
-concrete interpreter.
+continuations to higher-order functions; and finally a left-inverse CPS transformation turns
+the abstract interpreter into a sequence of expressions with delimited control operators,
+which looks almost no different to a concrete interpreter.
 
@@
 but also exists between abstract semantic artifacts. An interesting open question would
-be whether there also exists a correspondence of static analyses formalized in different 
-denotational style and operational style.
+be whether there also exists correspondences between static analyses that are formalized 
+in different denotational and operational styles.
 

### 2018-06-27 paper/paper.tex
fixes throughout

@@
   The transformations include linearization, lightweight fusion, disentanglement, 
-  refunctionalization, and finally back to direct-style with delimited control operators. 
+  refunctionalization, and finally transformation from CPS back to direct-style 
+  with delimited control operators. 
   Linearization expresses nondeterministic choices by first-order data types, 
@@
 Defining a language by building an interpreter for it can be traced to the very
-early days of programming languages research \cite{landin1966next, Reynolds:72, 
-McCarthy:1960:RFS:367177.367199}.
-Nowadays, even undergraduate students in computer science are able to make their own
-programming language by building an interpreter.
+early days of programming languages research \cite{McCarthy:1960:RFS:367177.367199,landin1966next, Reynolds:72}.
+Nowadays, even undergraduate students in computer science learn how to make their own
+programming languages by building an interpreter.
 But building a sound abstract interpreter remained an esoteric and difficult task 
@@
 abstract semantics artifacts, and provide a constructive answer to relate AAM and ADI 
-by presenting a series of syntactical transformations on the program from small-step 
+by presenting a series of syntactic transformations on the program from small-step 
 abstract abstract machines to big-step abstract definitional interpreters.
@@
      [start chain] \chainin (m-2-1);
-     \chainin (m-2-2) [join={node[above,labeled] {\begin{smallmatrix} \text{linearization} \\ \text{+} \\ \text{fusion} \end{smallmatrix}}     }];
+     \chainin (m-2-2) [join={node[above,labeled] {\begin{smallmatrix} \text{linearization} \\ \text{+} \\ \text{lightw. fusion} \end{smallmatrix}}     }];
      \chainin (m-2-3) [join={node[above,labeled] {\text{disentang.}}}];
@@
 as well as introducing some of the basic code structures used throughout the paper. 
-We then address the main contribution of this paper, which is the filling of the gap between
+We then address the main contribution of this paper: bridging the gap between
 small-step AAM and big-step ADI by applying a series of well-known systematic transformations
 drawn from concrete semantic artifacts.
-Figure \ref{fig:trans} shows the transformations.
-
-Those transformations are summarized here, with their associated section:
+Those transformations are shown in Figure~\ref{fig:trans} and summarized here, with their 
+respective associated sections:
 
@@
   \item Section~\ref{disen} discusses \textbf{disentanglement}. In this part, we identify 
-    different first-order datatypes which represent different layers of continuation, and
-    disassemble their handlers to be separate higher-order functions.
+    different first-order datatypes which represent different layers of continuations, and
+    disassemble their handlers into separate higher-order functions.
 
@@
 2) As a functional pearl, the code in this paper is executable with only few changes,
-which make it particularly fit for presenting syntactical transformations on abstract
+which make it particularly fit for presenting syntactic transformations on abstract
 interpreters.
@@
 In this functional pearl, we fill the gap between small-step abstract abstract machines
-and big-step abstract definitional interpreters by applying a series of syntactical
+and big-step abstract definitional interpreters by applying a series of syntactic 
 transformations. Among these transformations, linearization turns a worklist into another
@@
   %% extraction tools.
+  The authors are grateful to the anonymous reviewers for their 
+  helpful comments, and they are further indebted to Olivier Danvy 
+  for numerous invaluable comments on the draft.
+
   This material is based upon work supported by the
@@
   conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
+  of the authors and do not necessarily reflect the views of the
+  National Science Foundation. \note{TODO: grants}
 \end{acks}

### 2018-06-27 paper/paper.tex
A lot of revisions; still working on the figure text/layout

--- 2018-06-27 paper/paper.tex
@@
    % \chainin (m-2-4) [join={node[sloped,anchor=center,above,labeled] {\text{CPS trans.}}}]; }
-   %
-   % TODO: find out how to make the arrows visually separate
-   % {
-   %   [start chain] \chainin (m-1-5);
-   %   \chainin (m-2-4) [bend left=5, join={node[sloped,anchor=center,below,labeled] {\text{CPS}}}];
-   % }
    {
@@
 
-  \item Applying \textbf{refunctionalization} \cite{DANVY2009534, Danvy:2006:RW:2171265.2171268} 
+  \item Applying \textbf{Refunctionalization} \cite{DANVY2009534, Danvy:2006:RW:2171265.2171268} 
     to disentangled AAM is shown in Section~\ref{sec:refunc}. 

### 2018-06-28 paper/paper.tex
fix a typo
more fixes

--- 2018-06-27 paper/paper.tex
@@
 early days of programming languages research \cite{McCarthy:1960:RFS:367177.367199,landin1966next, Reynolds:72}.
-Nowadays, even undergraduate students in computer science learn how to make their own
-programming languages by building an interpreter.
+Nowadays, even undergraduate students in computer science learn how to define 
+their own programming languages by building an interpreter.
 But building a sound abstract interpreter remained an esoteric and difficult task 
@@
 drawn from concrete semantic artifacts.
-Those transformations are shown in Figure~\ref{fig:trans} and summarized here, with their 
-respective associated sections:
+Those transformations are shown in Figure~\ref{fig:trans} and summarized here, with full
+descriptions in their respective associated sections:
 
@@
 
-We complete this paper by discussing related work in Section~\ref{sec:related}, followed by 
+We discuss related work in Section~\ref{sec:related}, followed by 
 concluding thoughts in Section~\ref{sec:conclusion}.
@@
 writing a definitional interpreter \cite{Reynolds:72, Reynolds:HOSC98-revisited}.
-Refunctionalization is a left inverse of this transformation.
+Refunctionalization~\cite{TODOrefunc} is the left inverse of this transformation.
 
@@
 
-We show the functional correspondence not only exists between concrete semantic artifacts
-but also exists between abstract semantic artifacts. An interesting open question would
-be whether there also exists correspondences between static analyses that are formalized 
+We show that a functional correspondence exists not only between concrete semantic artifacts
+but also between abstract semantic artifacts. An interesting open question would
+be whether there also exist correspondences between static analyses that are formalized 
 in different denotational and operational styles.

### 2018-06-28 paper/paper.tex
Fixing typo
But -> However

@@
   The refunctionalized AAM is an abstract interpreter written in continuation-passing style
-  with two layers of continuation.
+  with two layers of continuations.
 
@@
 their own programming languages by building an interpreter.
-But building a sound abstract interpreter remained an esoteric and difficult task 
+However, building a sound abstract interpreter remained an esoteric and difficult task
 until very recently.

### 2018-06-28 paper/paper.tex
Clarification
Fixing intro

@@
 interpreters, respectively. An interesting question, and the subject of this paper, is
-how we can derive an abstract semantic artifact from the respective other one.
+how one can derive an abstract semantic artifact given the counterpart.
 
@@
 Danvy:2006:RW:2171265.2171268, danvy2009towards, biernacka2009towards}.
-The concrete abstract machines implement structural operational semantics in 
+These concrete abstract machines implement structural operational semantics in 
 continuation-passing style, where the reduction contexts are defunctionalized continuations.
@@
 One of the fundamental differences between concrete semantic artifacts and abstract semantic
-artifacts is a sound latter one must tell you facts about all possible execution paths, and 
-thus nondeterminism is introduced. In addition to ensuring termination, abstract semantic
-artifacts are usually equipped with a cache of reachable states, the cache is
+artifacts is that a sound instance of the latter must tell you facts about all possible execution paths, and
+thus introduce nondeterminism. In addition to ensuring termination, abstract semantic
+artifacts are usually equipped with a cache of reachable states; the cache is
 updated in a monotonic way and guaranteed to reach the least fixed-point eventually.
-In this functional pearl, we make a first step to answering the question of interderiving
-abstract semantics artifacts, and provide a constructive answer to relate AAM and ADI 
+In this functional pearl, we make a first step in answering the question of interderiving
+abstract semantics artifacts, and provide a constructive answer to relate AAMs and ADIs
 by presenting a series of syntactic transformations on the program from small-step
@@
 
-In addition, the abstract abstract machines with unbounded stack naturally correspond
-to abstract definitional interpreters. We show that after refunctionalizing an AAM with 
-unbounded stack, and with the proper caching algorithm, the pushdown control-flow analysis 
+In addition, those abstract abstract machines with unbounded stacks naturally correspond
+to abstract definitional interpreters. We show that after refunctionalizing an AAM with an
+unbounded stack (and utilizing the proper caching algorithm), the pushdown control-flow analysis
 can be obtained.

### 2018-06-28 paper/paper.tex
Minor grammer, whitespace

@@
 In this functional pearl, we make a first step in answering the question of interderiving
-abstract semantics artifacts, and provide a constructive answer to relate AAMs and ADIs
+abstract semantics artifacts, and provide a constructive answer to relate AAM and ADI
 by presenting a series of syntactic transformations on the program from small-step
@@
 We begin by reviewing some background necessary for this work in Section~\ref{background},
-as well as introducing some of the basic code structures used throughout the paper. 
+as well as by introducing some of the basic code structures used throughout the paper.
 We then address the main contribution of this paper: bridging the gap between
@@
 
-These transformations are used throughout the paper, with refunctionalization and
+These transformations are used throughout the paper, with the refunctionalization and
 defunctionalization of abstract interpreters playing important roles for the stack model of
@@
 is blended into the call structure of the defining language. This provides another perspective to
-explain why \citeauthor{darais2017abstracting}'s abstract definitional interpreters
+explain why \citeauthor{darais2017abstracting}'s abstract definitional interpreter
 is able to inherit the pushdown control-flow property from its defining language.

### 2018-06-28 paper/paper.tex
Minor grammar

@@
 primitive types, and operations on primitive types. These cases would be straightforward to
-implement without introducing new transformations as far as the concerns of this paper,
+implement without introducing new transformations related to the concerns of this paper,
 so we elide them here.
@@
 In ANF, an expression is either an atomic expression or a @let@ expression.
-A restriction exists which states that all function applications must be administrated
+A restriction exists which states that all function applications must be administered
 within a @let@ expression and then bound to a variable name under the current environment.
@@
 The abstract syntax in Scala is shown as follows. We assume that the source program conforms
-to the ANF convention, and we do not enforce it in the term structure of Scala constructs.
+to the ANF convention, and as such, do not enforce it in the term structure of Scala constructs.
 

### 2018-06-28 paper/paper.tex
Grammar

@@
 It is worth noting that the continuation class @Kont@ is defined as a list of
-frames, where a frame can be considered a representation of reduction context 
+frames, where a frame can be considered a representation of a reduction context
 in reduction semantics.
-An empty list represents empty call stack; otherwise, the head frame in the list
+An empty list represents an empty call stack; otherwise, the head frame in the list
 represents the innermost context, and the reduction result of this frame will be used
@@
 A @Frame@ constitutes a variable name @x@ to be bound later, and a control component
-to which the program may resume, and its environment.
+to which the program may resume, as well as its environment.
 

### 2018-06-28 paper/paper.tex
fix typo

@@
     different first-order datatypes which represent different layers of continuations, and
-    disassemble their handlers into separate higher-order functions.
+    disassemble their handlers into separate functions.
 

### 2018-06-28 paper/paper.tex
fix grammar, add ack

@@
 
-In addition, those abstract abstract machines with unbounded stacks naturally correspond
+In addition, those abstract abstract machines with an unbounded stack naturally correspond
 to abstract definitional interpreters. We show that after refunctionalizing an AAM with an
@@
 
-The function @analyze@ is also changed by calling a @reset@ operator around the function 
+The function @analyze@ is also changed by calling the @reset@ operator around the function 
 call of @aeval@ to set the delimiter. The afterwards updating on cache becomes sequential.
@@
   helpful comments, and they are further indebted to Olivier Danvy 
-  for numerous invaluable comments on the draft.
+  for numerous invaluable comments on the draft. We also thank
+  Kimball Germane and David Sherratt for their feedback, and
+  Thomas Gilray for the discussion on pushdown control-flow analysis.
 

### 2018-06-28 paper/paper.tex
update figure 1

@@
    {
-     [start chain] \chainin (m-2-4);
-     \chainin (m-1-5) [double,join={node[sloped,anchor=center,above,labeled] {\text{CPS left-inv.}}}];
+     [start chain, transform canvas={yshift=0.5ex,xshift=-0.5ex}] \chainin (m-2-4);
+     \chainin (m-1-5) [join={node[sloped,anchor=center,above,labeled] {\text{CPS left-inv.}}}];
+   }
+   {
+     [start chain, color=gray, transform canvas={yshift=-0.5ex,xshift=0.5ex}] \chainin (m-1-5);
+     \chainin (m-2-4) [join={node[sloped,anchor=center,below,labeled] {\text{CPS trans.}}}];
+   }
+   {
+     [start chain, color=gray] \chainin (m-2-5);
+     \chainin (m-1-5) [join={node[above,labeled] {\text{}}}];
      \chainin (m-2-5) [join={node[above,labeled] {\text{}}}];
-     \chainin (m-1-5) [join={node[above,labeled] {\text{}}}];
-   }
-   %{ [start chain] \chainin (m-1-5);
-   % \chainin (m-2-4) [join={node[sloped,anchor=center,above,labeled] {\text{CPS trans.}}}]; }
-   %
-   % TODO: find out how to make the arrows visually separate
-   % {
-   %   [start chain] \chainin (m-1-5);
-   %   \chainin (m-2-4) [bend left=5, join={node[sloped,anchor=center,below,labeled] {\text{CPS}}}];
-   % }
+     \chainin (m-2-4) [join={node[sloped,anchor=center,below,labeled] {\text{inline monad}}}];
+   }
    {
@@
    {
-     [start chain, color=gray]
+     [start chain, color=gray, dashed]
      \chainin (m-3-1);
@@
    {
-     [start chain, color=gray]
+     [start chain, color=gray, dashed]
      \chainin (m-3-5);

### 2018-06-28 paper/paper.tex
adjust figure 1

@@
          &  &  &  &
-       \begin{smallmatrix} \text{abstract} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
+       \begin{smallmatrix} \text{monadic} \\ \text{abstract} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
        \begin{smallmatrix} \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
@@
        \begin{smallmatrix} \text{refunctionalized}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
-       \begin{smallmatrix} \text{monadic} \\ \text{abstract} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
+       \begin{smallmatrix} \text{abstract} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
        \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} &
@@
    {
-     [start chain, transform canvas={yshift=0.5ex,xshift=-0.5ex}] \chainin (m-2-4);
-     \chainin (m-1-5) [join={node[sloped,anchor=center,above,labeled] {\text{CPS left-inv.}}}];
+     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-4);
+     \chainin (m-2-5) [join={node[sloped,anchor=center,above,labeled] {\text{CPS left-inv.}}}];
    }
    {
-     [start chain, color=gray, transform canvas={yshift=-0.5ex,xshift=0.5ex}] \chainin (m-1-5);
+     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-5);
      \chainin (m-2-4) [join={node[sloped,anchor=center,below,labeled] {\text{CPS trans.}}}];
@@
    {
-     [start chain, color=gray] \chainin (m-2-5);
+     [start chain, color=gray] \chainin (m-1-5);
+     \chainin (m-2-5) [join={node[above,labeled] {\text{}}}];
      \chainin (m-1-5) [join={node[above,labeled] {\text{}}}];
-     \chainin (m-2-5) [join={node[above,labeled] {\text{}}}];
-     \chainin (m-2-4) [join={node[sloped,anchor=center,below,labeled] {\text{inline monad}}}];
+     \chainin (m-2-4) [join={node[sloped,anchor=center,above,labeled] {\text{inline monads}}}];
    }

### 2018-06-28 paper/paper.tex
misc
Add a dashed rectangle

@@
        \begin{smallmatrix} \text{disentangled}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
-       \begin{smallmatrix} \text{refunctionalized}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
+       \begin{smallmatrix} \text{refunc.}  \\ \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
        \begin{smallmatrix} \text{abstract} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
@@
      \chainin (m-2-4) [join={node[above,labeled] {\text{refunc.}}}];
-     %\chainin (m-2-5) [join={node[above,labeled] {\text{monadify}}}];
    }
@@
    }
-     %\chainin (m-2-2);
-     %\chainin (m-2-3);
-     %\chainin (m-2-4);
    {
@@
    }
+   \draw[gray,dashed,transform canvas={xshift=0.5ex}] (m-1-5.north west) -- (m-2-5.south west);
+   \draw[gray,dashed,transform canvas={xshift=3.7ex, xscale=0.90}] (m-1-5.north west) -- (m-1-5.north east);
+   \draw[gray,dashed,transform canvas={xshift=-0.2ex}] (m-1-5.north east) -- (m-2-5.south east);
+   \draw[gray,dashed,transform canvas={xshift=3.7ex, xscale=0.90}] (m-2-5.south west) -- (m-2-5.south east);
  \end{tikzpicture}

### 2018-06-29 paper/paper.tex
add defunc to the figure

@@
      [start chain] \chainin (m-2-1);
-     \chainin (m-2-2) [join={node[above,labeled] {\begin{smallmatrix} \text{linearization} \\ \text{+} \\ \text{lightw. fusion} \end{smallmatrix}}     }];
+     \chainin (m-2-2) [join={node[above,labeled] {
+       \begin{smallmatrix} \text{linearization} \\ \text{+} \\ \text{lightw. fusion} \end{smallmatrix}
+     }}];
      \chainin (m-2-3) [join={node[above,labeled] {\text{disentang.}}}];
+   }
+   {
+     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-3);
      \chainin (m-2-4) [join={node[above,labeled] {\text{refunc.}}}];
@@
    {
-     [start chain, color=gray, dashed]
-     \chainin (m-3-1);
-     { [start branch=A] \chainin (m-2-1)
-       [join={node[left, labeled] {\alpha}}];
-     }
+     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-4);
+     \chainin (m-2-3) [join={node[below,labeled] {\text{defunc.}}}];
    }
    {
-     [start chain, color=gray, dashed]
-     \chainin (m-3-5);
-     { [start branch=B] \chainin (m-2-5)
+     [start chain, color=gray, dashed] \chainin (m-3-1);
+     { [start branch=A] \chainin (m-2-1) [join={node[left, labeled] {\alpha}}]; }
+   }
+   {
+     [start chain, color=gray, dashed] \chainin (m-3-5);
+     { 
+       [start branch=B] \chainin (m-2-5)
        [join={node[right, labeled] {\alpha}}];

### 2018-06-29 paper/paper.tex
revise frame/context

@@
 It is worth noting that the continuation class @Kont@ is defined as a list of
-frames, where a frame can be considered a representation of a reduction context
-in reduction semantics.
-An empty list represents an empty call stack; otherwise, the head frame in the list
-represents the innermost context, and the reduction result of this frame will be used
-to fill the ``hole'' of its following element in the list.
+frames to represent a reduction context.
+An empty list represents an empty context, and corresponds to halt. Otherwise, the head 
+frame in the list represents the innermost context, and the reduction result of this 
+frame will be used to fill the ``hole'' of its following frame in the list.
 We represent frames using the @Frame@ class, which stores the information of a single

### 2018-06-30 paper/paper.tex
misc
fix typos

@@
        \begin{smallmatrix} \text{monadic} \\ \text{abstract} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
+
        \begin{smallmatrix} \text{abstract} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
@@
        \begin{smallmatrix} \text{abstract} \\ \text{definitional} \\ \text{interpreters} \end{smallmatrix}  \\
-       \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} &
-         &
-       %\begin{smallmatrix} \text{big-step} \\ \text{abstract} \\ \text{machines} \end{smallmatrix}  &
-         &
-       %\begin{smallmatrix} \text{disentangled} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
-         &
-       %\begin{smallmatrix} \text{refunctionalized} \\ \text{abstract} \\ \text{machines} \end{smallmatrix} &
+
+       \begin{smallmatrix} \text{abstract} \\ \text{machines} \end{smallmatrix} & & & &
        \begin{smallmatrix} \text{definitional} \\ \text{interpreters} \end{smallmatrix} \\
@@
 Not unexpectedly, the correspondence also exists for nondeterministic languages. For example,
-\citeauthor{Danvy:2001:DW:773184.773202} show that refunctionalizing a regualr expressions matcher yields
+\citeauthor{Danvy:2001:DW:773184.773202} show that refunctionalizing a regular expressions matcher yields
 its counterpart with two layer of continuations \cite{Danvy:2001:DW:773184.773202}.
-And the multi-layer continuation-passing style is the underlying semantics of control operatos
+And the multi-layer continuation-passing style is the underlying semantics of control operators
 like @reset@ and @shift@ \cite{Danvy:1990:AC:91556.91622}.

### 2018-06-30 paper/paper.tex
replace abstracting with abstract

@@
   and obtain monadic abstract definitional interpreters (ADI) in big-step style.
-  Yet, the relations between small-step abstracting abstract machines and big-step 
-  abstracting definitional interpreters are not well studied.
+  Yet, the relations between small-step abstract abstract machines and big-step 
+  abstract definitional interpreters are not well studied.
 
@@
 abstract interpreters for higher-order functional languages from concrete abstract 
-machines \cite{van2012systematic, van2010abstracting}. An abstracting abstract machine
+machines \cite{van2012systematic, van2010abstracting}. An abstract abstract machine
 implements computable abstract semantics which approximates the runtime behaviors of 
@@
 
-In this section, we derive the abstracting abstract machine from concrete
+In this section, we derive the abstract abstract machine from concrete
 CESK machines, and also show how to instantiate useful $k$-call-sensitive 
@@
 \begin{itemize}
-  \item By monadifying the continuations, we obtain the abstracting
-    definitional interpreters in monadic style which are what \citeauthor{darais2017abstracting}
-    describe \cite{darais2017abstracting}.
+  \item By monadifying the continuations, we obtain the abstract definitional interpreters 
+    in monadic style which are what \citeauthor{darais2017abstracting} describe \cite{darais2017abstracting}.
   \item By representing the extended continuation-passing style with delimited control operators
@@
 gilray2016pushdown, johnson2015abstracting} to achieve precise call/return match based on small-step
-abstracting abstract machines.
+abstract abstract machines.
 
@@
 defined language can inherit properties from the defining language. With this insight,
-\citeauthor{darais2017abstracting} construct abstracting definitional interpreters which
+\citeauthor{darais2017abstracting} construct abstract definitional interpreters which
 automatically inherit the pushdown control-flow property from its defining language because
 the defined language simply uses the call-stack model of the meta-language~\cite{darais2017abstracting}.
-\citeauthor{darais2017abstracting}'s abstracting definitional interpreters work on direct-style
+\citeauthor{darais2017abstracting}'s abstract definitional interpreters work on direct-style
 $\lambda$-calculus, and itself is written in monadic style. One of the advantages of monadic

### 2018-06-30 paper/paper.tex
more revisions

@@
                                         %% contents suppressed with 'anonymous'
-\subtitle{Filling the Gap Between Abstract Abstract Machines and Abstract Definitional Interpreters}        %% \subtitle is optional
+\subtitle{Bridging the Gap Between Abstract Abstract Machines and Abstract Definitional Interpreters}        %% \subtitle is optional
 %\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
@@
   The transformations include linearization, lightweight fusion, disentanglement, 
-  refunctionalization, and finally transformation from CPS back to direct-style 
-  with delimited control operators. 
+  refunctionalization, and finally from CPS back to direct-style with delimited 
+  control operators. 
   Linearization expresses nondeterministic choices by first-order data types, 
-  after which refunctionalization transforms the first-order datatypes representing continuations
+  after which refunctionalization transforms the first-order data types representing continuations
   into higher-order functions. 
@@
 
-Applying the same idea to big-step definitional interpreters, \citeauthor{darais2017abstracting}
-built abstract definitional interpreters (ADI) that are written in monadic style
+Applying the same idea to monadic definitional interpreters, \citeauthor{darais2017abstracting}
+built monadic abstract definitional interpreters (ADI) in big-step style
 \cite{darais2017abstracting}. One of the advantages of a monadic interpreter is
@@
 One of the fundamental differences between concrete semantic artifacts and abstract semantic
-artifacts is that a sound instance of the latter must tell you facts about all possible execution paths, and
-thus introduce nondeterminism. In addition to ensuring termination, abstract semantic
+artifacts is that a sound instance of the latter must tell you facts about all possible execution 
+paths, and thus introduce nondeterminism. In addition to ensuring termination, abstract semantic
 artifacts are usually equipped with a cache of reachable states; the cache is
-updated in a monotonic way and guaranteed to reach the least fixed-point eventually.
+iteratively updated in a monotonic way and guaranteed to reach the least fixed-point eventually.
 In this functional pearl, we make a first step in answering the question of interderiving
@@
   \item Section~\ref{disen} discusses \textbf{disentanglement}. In this part, we identify 
-    different first-order datatypes which represent different layers of continuations, and
+    different first-order data types which represent different layers of continuations, and
     disassemble their handlers into separate functions.
@@
     %as the left inverse of Reynolds' defunctionalization \cite{Reynold:72,Reynolds:HOSC98-revisited} 
-    Refunctionalization replaces the first-order datatypes representing continuations with
+    Refunctionalization replaces the first-order data types representing continuations with
     higher-order functions, and associates dispatching logic with proper higher-order functions. 
@@
 The function @step@ is a partial function that only handles non-final states,
-which must have a successor; the final case of a state is handled by funciton
+which must have a successor; the final case of a state is handled by function
 @drive@, which will be mentioned in the next section.
@@
 machines \cite{van2012systematic, van2010abstracting}. An abstract abstract machine
-implements computable abstract semantics which approximates the runtime behaviors of 
+implements computable abstract semantics which approximates the run-time behaviors of 
 programs.
@@
 @[KAddr, Cont]@, we obtain @BStore@ and @KStore@, respectively.
-We note that both the value store and continuation store are uodated monotonically;
+We note that both the value store and continuation store are updated monotonically;
 it continuously grows and never shrinks.
@@
 
-The codomain of binding stores @Storable@ is the same as previously defined for CESK 
-machines. The codomain of continuation stores @Cont@, on the other hand, is comprised of
-a @Frame@ object and a continuation address @KAddr@. To mimic the runtime call stack, 
+The co-domain of binding stores @Storable@ is the same as previously defined for CESK 
+machines. The conman of continuation stores @Cont@, on the other hand, is comprised of
+a @Frame@ object and a continuation address @KAddr@. To mimic the run-time call stack, 
 @KAddr@ plays the role of representing the remaining stack frames.
@@
 state space analysis \cite{Gilray:2016:ACP:2951913.2951936}.
-To effectively approximate the runtime behavior, we introduce a finite program contours
+To effectively approximate the run-time behavior, we introduce a finite program contours
 @time@ that encodes the program execution history. The function @tick@ is used to refresh
@@
 as the program is abstractly executing. The following code shows a variant of the 
-worklist algorithm to find the fixed-point of the set of states.
+work-list algorithm to find the fixed-point of the set of states.
 Function @drive@ always applies function @step@ to the head element @hd@ of
-the worklist @todo@ if @hd@ is unseen. It then inserts the result of @step@ to
-the rest of worklist, and in the meantime adds @hd@ to the explored states set.
-If the worklist is empty, @drive@ simply returns the set of reachable states
+the work-list @todo@ if @hd@ is unseen. It then inserts the result of @step@ to
+the rest of work-list, and in the meantime adds @hd@ to the explored states set.
+If the work-list is empty, @drive@ simply returns the set of reachable states
 up to the current execution point.
@@
 continuations in states, we intend to use a list of frames to explicitly model 
-the stack. By doing so, we recover the call structure as the same as runtimes
+the stack. By doing so, we recover the call structure as the same as run-times
 (so called pushdown control-flow analysis), but since the stack is unbounded, 
@@
 
-The pushdown AAM with unbounded stack is also described by \citeauthor{van2012systematic} 
+The pushdown AAM with an unbounded stack is also described by \citeauthor{van2012systematic} 
 \cite{van2012systematic}. 
@@
 space, we can recover the precise call/return match. Now we begin describing the
-transformations step by step. Our base machine for now is the AAM with unbounded
-stack, though as we will later transform the stack structure to higher-order functions
+transformations step by step. Our base machine is the AAM with an unbounded
+stack, as we will later transform the stack structure to higher-order functions
 representing continuations, it does not matter what kind of AAM we start from.
 This is because it would not be possible to construct a new frame on the stack, nor
-to allocate continuations in the store.
-Thus, our choice to start from an AAM with unbounded stack is motivated simply because
+to allocate continuations in the store after Refunctionalization.
+Thus, our choice to start from an unbounded-stack AAM is motivated simply because
 it has an equivalent stack model to abstract definitional interpreters (our final target).
@@
 multiple target closures when dereferencing an address, and we have to explore both branches
-if the conditionals exist in the language. But the reduction context (i.e., the @Frame@ in
-our program) is \emph{not} nondeterministic, and the worklist @todo@ actually implicitly
+if conditionals exist in the analyzed language. But the reduction context (i.e., the @Frame@ in
+our program) is \emph{not} nondeterministic, and actually the work-list @todo@ implicitly
 handles the nondeterminism. Thus, if we simply refunctionalize the frames to higher-order
@@
 We then introduce another layer of control, and define a case class @NDCont@
-that saves the information at a fork point when we have multiple target closures.
-We also add a new field @ndk@ to the definition of state, which we now call @NDState@.
-For clarity of presentation in differentiating between the two continuations, we elect to call
-the first a \emph{continuation}, and the second a \emph{meta-continuation}
+that saves the information at a fork point when we have nondeterministic target closures.
+We also add a new field @ndk@ of type @NDCont@ to the definition of state, which we now 
+call @NDState@. For clarity of presentation in differentiating between the two continuations, 
+we elect to call the first a \emph{continuation}, and the second a \emph{meta-continuation}
 \footnote{Traditionally, they are also called \emph{success continuation} and \emph{failure
@@
 Each @NDCont@ object contains a list of closures that are possible functions to be invoked,
-a set of values that will be bound to the function's formal argument, and the store, time,
+a set of values that will be bound to the target function's formal argument, and the store, time,
 and frames at the fork point. In the definition of @NDState@, an auxiliary function @toState@
-is added that converts itself to @State@.
+is added that converts itself to type @State@.
 
@@
 they may form different binding addresses @baddr@. We also use the new frames, because
-all closures at this fork point share the same stack and return point.
+all closures at this fork point share the same stack structure and return point.
 
@@
 
-If the continuation @konts@ is an empty list, then it means we have reached
+Normally, if the continuation @konts@ is an empty list, then it means we have reached
 the end of the computation (i.e., a halt); otherwise, we should return the values of 
@@
 we have reached the halt of \textit{one computation path}, and we should determine the
-next state indicated by @NDCont@. Thus, we have a pattern matching on @ndk@:
+next state indicated by @NDCont@. Thus, we have a inner pattern matching on @ndk@:
 
 \begin{itemize}
-  \item If the closure set is an empty set,
+  \item If the closure set is an empty set (@Nil@ in the code since we list to represent that set),
 then we have tried all the closures of this fork point and should move to the
-next fork point. Since the @NDCont@ is represented by a list, and we always
+next fork point. Since the meta-continuation @nkd@ is represented by a list, and we always
 append new elements to its front, we are resuming to the most recent fork point by
-popping up the front-most element.
+popping up the front-most element and using the tail of that list.
   \item Otherwise, we pop a closure from the set, then build a new
@@
 than current one.
-The meta-continuation @new_ndk@ is also updated by the removal of that closure.
+The meta-continuation @new_ndk@ is also updated by the removal of that closure, which makes
+sure we will eventually shrink to the empty set case.
 \end{itemize}
@@
 
-The @drive@ function is also changed: there is no worklist anymore, as all the
+The @drive@ function is also changed: there is no work-list anymore, as all the
 potentially unexplored states are embedded into the continuation for
@@
 is atomic and both the continuation and meta-continuation are empty lists. This corresponds 
-to the case that @todo@ list is empty.
+to the case that @todo@ list is empty. The set of explored states is preserved as before.
 
@@
 classical AAM explores a graph of reachable states, then the linearized AAM flattens the
-graph to a linear sequence.
+graph to a linear sequence of states.
 
@@
 and lift the code blocks that handle these reduction contexts to be individual functions.
-
-Since there are two layers of continuations, we obtain three mutually recursive functions:
+After disentanglement, we will obtain an abstract abstract machine that in defunctionalized form.
+
+Since there are two layers of continuations, we disentangle them into three mutually recursive 
+functions:
 @drive_step@, which plays the same role as before; @continue@, which is called from
@@
 proper higher-order functions.
+After this transformation, we will obtain 
 This transformation can be also regarded as making the control flows of the defined 
@@
 \item We do not have the concept of @state@ (or @NDState@) anymore.
-The components of the states are lifted as arguments of the evaluation
-function.
+  The components of the states are lifted as arguments of the evaluation function.
 \item Since our final target is an abstract definitional interpreter (though still
@@
   returns a set of final values instead of collected states.
-\item As we are getting close to achieving
-abstract definitional interpreters, the name of @drive_step@ is updated to
-@aeval@ (for \emph{abstract eval}).
+\item As we are getting close to achieving abstract definitional interpreters, 
+  the name of @drive_step@ is updated to @aeval@ (for \emph{abstract eval}).
 \end{itemize}
 
-For clarity, we first show how to refunctionalize continuations and
-meta-continuation without collecting semantics.
-Then, adding an appropriate caching algorithm to make sure the analysis will
+For clarity, we first show how to refunctionalize continuations and meta-continuation 
+without collecting semantics.
+Then, adding an appropriate caching algorithm to ensure the analysis will
 terminate at some fixed-point is a simple transformation of the program to cache-passing
@@
 
-Given a collection of elements of type @T@, an initial accumlator of type 
+Given a collection of elements of type @T@, an initial accumulator of type 
 @Ans@, a function @k@ that works on element's type @T@ and takes a continuation 
@@
 \textit{Cache.}
-Recall that the latent assumption of the worklist algorithm in classical AAM is that 
+Recall that the latent assumption of the work-list algorithm in classical AAM is that 
 if we have seen a state @s@, it means we also have seen all the successors of state @s@.
@@
 contains some values for the current configuration and use them immediately (if present).
-This corresponds to small-step AAM with a worklist: when we have seen a state,
-we can simply discard the state and continue working through the rest of the worklist @todo@.
+This corresponds to small-step AAM with a work-list: when we have seen a state,
+we can simply discard the state and continue working through the rest of the work-list @todo@.
 
@@
 \textit{The Problem with Return-flows.}
-Pushdown control-flow is a property in analysis that precisely models the runtime
+Pushdown control-flow is a property in analysis that precisely models the run-time
 call structure of the analyzed program. A pushdown control-flow analysis provides an 
-as-exact-as-runtime return-flow when analyzing a program, but traditional 
+as-exact-as-run-time return-flow when analyzing a program, but traditional 
 control-flow analysis collapses the state space into a finite space and thus 
@@
 \begin{itemize}
-  \item By monadifying the continuations, we obtain the abstract definitional interpreters 
-    in monadic style which are what \citeauthor{darais2017abstracting} describe \cite{darais2017abstracting}.
+  \item By representing the continuations with monads\cite{filinski1994representing}, 
+    we obtain the abstract definitional interpreters in monadic style which are similar to what 
+    \citeauthor{darais2017abstracting} describe \cite{darais2017abstracting}.
   \item By representing the extended continuation-passing style with delimited control operators
@@
 
+\textit{Defunctionalization and Refunctionalization.}
+\citeauthor{Reynolds:72} first shows defunctionalization as a program transformation 
+technique that can be used to transform higher-order functions to first-order data types and 
+its first-order dispatching function when writing definitional interpreters
+\cite{Reynolds:72, Reynolds:HOSC98-revisited}. As generalized closure conversion,
+defunctionalization is also used in program analysis, transformation, and partial evaluation. %TODO cite
+Refunctionalization~\cite{Danvy:2006:RW:2171265.2171268, Danvy:2001:DW:773184.773202} is the 
+left inverse of defunctionalization, firstly considered by \citeauthor{Danvy:2001:DW:773184.773202}.
+
+\citeauthor{Ager:2003:FCE:888251.888254} revealed that defunctionalization and refunctionalization
+can be used to show the functional correspondence between interpreters and abstract machines
+\cite{Ager:2003:FCE:888251.888254}. A number of independently designed small-step semantic artifacts can 
+be obtained by closure-converting, CPS-transforming, and then defunctionalizing their big-step counterparts, 
+and vice versa by refunctionalizing \cite{DANVY2009534}. 
+Such semantic artifacts include SECD machines \cite{10.1007/11431664_4, Ager:2003:FCE:888251.888254, DanvySECDJ},
+CEK machines \cite{Ager:2003:FCE:888251.888254}, CLS machines  \cite{Ager:2003:FCE:888251.888254},
+Categorical Abstract Machines \cite{Ager:2003:FCE:888251.888254}, lazy abstract machines \cite{AGER2004223}, 
+monadic evaluators \cite{ager2005functional}, as well as languages with richer constructs
+\cite{Danvy:2008:DIP:1411204.1411206, danvy2009towards, biernacka2009towards}.
+
+Besides, Danvy and his collaborators observed the correspondence between evaluation contexts and continuations.
+For deterministic languages, \citeauthor{Danvy:2008:DIP:1411204.1411206} showed that reduction contexts
+are defunctionalized (data type representing) continuations and reduction semantics is a structural 
+operational semantics in continuation-passing style \cite{Danvy:2008:DIP:1411204.1411206}.
+Not unexpectedly, the correspondence also exists for nondeterministic languages. For example,
+\citeauthor{Danvy:2001:DW:773184.773202} showed that refunctionalizing a regular expression matcher yields
+its counterpart with two layer of continuations \cite{Danvy:2001:DW:773184.773202}.
+And the operational semantics of control operators (e.g., @reset@ and @shift@) can be given by an 
+abstract machine with multi-layer continuations \cite{Danvy:1990:AC:91556.91622, lmcs:2269}.
+
+More generally, other kinds of programs, instead of merely semantic artifacts, are also found to be applicable.
+As \citeauthor{Danvy:2001:DW:773184.773202} pointed out, defunctionalization is dual to Church encoding.
+Examples include backtracking algorithms (e.g., regular expression matcher), Dyck word recognizer, and 
+Dijkstra's shunting-yard algorithm and even Quicksort algorithm \cite{Danvy:2006:RW:2171265.2171268, 
+DANVY2009534, Danvy:2001:DW:773184.773202}. \\
+
+%Our work applies these ideas to the nondeterministic AAM. The additional layer of continuation controls
+%the nondeterministic choice over closures. After identifying and explicitly exposing (Section~\ref{linear})
+%that layer of meta-continuations, the whole program can be refunctionalized to extended
+%continuation-passing style, and then back to direct-style by the use of delimited control operators. \\
+
+\textit{Lightweight Fusion.}
+Lightweight fusion as one of the transformation in our paper is also studied by 
+\citeauthor{Ohori:2007:LFF:1190216.1190241} \cite{Ohori:2007:LFF:1190216.1190241}.
+As shown by \citeauthor{DANVY2008100}, lightweight fusion can be used to show the 
+equivalence between small-step and big-step abstract machines \cite{DANVY2008100}.\\
+
 \textit{Abstract Interpretation and Control-Flow Analysis.}
 \citeauthor{cousot1977abstract} invented abstract interpretation as a sound approach to
-approximate a program's runtime behavior \cite{cousot1977abstract}. Control-flow analysis
+approximate a program's run-time behavior \cite{cousot1977abstract}. Control-flow analysis
 is one instance of abstract interpretation on functional programs that can be traced to
@@
 
-\textit{Defunctionalization and Refunctionalization.}
-Defunctionalization and refunctionalization build connections between abstract machines and
-interpreters. \citeauthor{Reynolds:72} first shows defunctionalization as a program transformation 
-technique that can be used to transform higher-order functions to first-order functions when
-writing a definitional interpreter \cite{Reynolds:72, Reynolds:HOSC98-revisited}.
-Refunctionalization~\cite{TODOrefunc} is the left inverse of this transformation.
-
-Over the years, Danvy and his collaborators applied refunctionalization and defunctionalization 
-to many different abstract machines and interpreters and build functional correspondences among
-them \cite{Ager:2003:FCE:888251.888254, Danvy:2001:DW:773184.773202, danvy2004refocusing,
-Danvy:2008:DIP:1411204.1411206, AGER2004223, ager2005functional, Danvy:2006:RW:2171265.2171268,
-DANVY2009534, danvy2009towards, biernacka2009towards}. Such semantic artifacts include the CEK machines,
-the CLS machines, the SECD machines, call-by-need evaluators, monadic evaluators, etc.
-Other applications are also found to be applicable, such as Dyck word recognizer, and 
-Dijkstra's shunting-yard algorithm and even Quicksort algorithm \cite{Danvy:2006:RW:2171265.2171268,
-DANVY2009534}.
-For deterministic languages, \citeauthor{Danvy:2008:DIP:1411204.1411206} shows a reduction semantics
-is a structural operational semantics in continuation-passing style, where the reduction contexts
-are defunctionalized (data type representing) continuations \cite{Danvy:2008:DIP:1411204.1411206}.
-Not unexpectedly, the correspondence also exists for nondeterministic languages. For example,
-\citeauthor{Danvy:2001:DW:773184.773202} show that refunctionalizing a regular expressions matcher yields
-its counterpart with two layer of continuations \cite{Danvy:2001:DW:773184.773202}.
-And the multi-layer continuation-passing style is the underlying semantics of control operators
-like @reset@ and @shift@ \cite{Danvy:1990:AC:91556.91622}.
-
-Our work applies these ideas to the nondeterministic AAM. The additional layer of continuation controls
-the nondeterministic choice over closures. After identifying and explicitly exposing (Section~\ref{linear})
-that layer of meta-continuations, the whole program can be refunctionalized to extended
-continuation-passing style, and then back to direct-style by the use of delimited control operators.
-
-Lightweight fusion as one of the transformation in our paper is also studied by 
-\citeauthor{Ohori:2007:LFF:1190216.1190241} \cite{Ohori:2007:LFF:1190216.1190241}.
-As shown by \citeauthor{DANVY2008100}, lightweight fusion can be used to show the 
-equivalence between small-step and big-step abstract machines \cite{DANVY2008100}.\\
-
 \textit{Pushdown Control-Flow Analysis.}
@@
 \textit{Abstracting Definitional Interpreters.}
-Reynold's seminal paper \emph{Definitional interpreters for higher-order programming languages} 
+Reynolds' seminal paper \emph{Definitional interpreters for higher-order programming languages} 
 \cite{Reynolds:72, Reynolds:HOSC98-revisited} shows that in some definitional interpreters, the
@@
 The target of these transformations in this paper is abstract definitional interpreters, and we 
-additionally demonstrate abstract interpreters written direct-style with delimited control operators. 
-We use the delimited control operators @shift@ and @reset@ that are implemented in Scala 
+additionally demonstrate abstract interpreters written in direct-style with delimited control operators. 
+We use the delimited control operators @reset@ and @shift@ that are implemented in Scala 
 \cite{rompf2009implementing}. The correspondence of delimited continuations and monads are well-known
-\cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions}.
+\cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions, filinski1994representing}.
 
@@
 
-In this functional pearl, we fill the gap between small-step abstract abstract machines
+In this functional pearl, we bridge the gap between small-step abstract abstract machines
 and big-step abstract definitional interpreters by applying a series of syntactic 
-transformations. Among these transformations, linearization turns a worklist into another
+transformations. Among these transformations, linearization turns a work-list into another
 layer of continuations; refunctionalization converts the first-order data types representing
-continuations to higher-order functions; and finally a left-inverse CPS transformation turns
+continuations to higher-order functions; and finally a left-inverse CPS transformation converts
 the abstract interpreter into a sequence of expressions with delimited control operators,
-which looks almost no different to a concrete interpreter.
+which looks almost no different from a concrete interpreter.
 

### 2018-06-30 paper/paper.tex
fix typos

@@
 The co-domain of binding stores @Storable@ is the same as previously defined for CESK 
-machines. The conman of continuation stores @Cont@, on the other hand, is comprised of
+machines. The c-domain of continuation stores @Cont@, on the other hand, is comprised of
 a @Frame@ object and a continuation address @KAddr@. To mimic the run-time call stack, 
@@
 we have reached the halt of \textit{one computation path}, and we should determine the
-next state indicated by @NDCont@. Thus, we have a inner pattern matching on @ndk@:
+next state indicated by @NDCont@. Thus, we have an inner pattern matching on @ndk@:
 

### 2018-06-30 paper/paper.tex
add ref

@@
 \cite{Reynolds:72, Reynolds:HOSC98-revisited}. As generalized closure conversion,
-defunctionalization is also used in program analysis, transformation, and partial evaluation. %TODO cite
-Refunctionalization~\cite{Danvy:2006:RW:2171265.2171268, Danvy:2001:DW:773184.773202} is the 
-left inverse of defunctionalization, firstly considered by \citeauthor{Danvy:2001:DW:773184.773202}.
+defunctionalization is widely used in compilation and analysis \cite{pottier2006polymorphic, 
+Eisenberg:2014:PFT:2633357.2633361, fourtounis2014modular, 10.1007/3-540-46425-5_4, consel1993tour}.
+Refunctionalization~\cite{Danvy:2006:RW:2171265.2171268, Danvy:2001:DW:773184.773202} as a 
+left inverse of defunctionalization, was firstly considered by \citeauthor{Danvy:2001:DW:773184.773202}.
 

### 2018-06-30 paper/paper.tex
revise on section 3

@@
 
-At this point, we have obtained a linearized abstract abstract machine. If we imagine that the
-classical AAM explores a graph of reachable states, then the linearized AAM flattens the
-graph to a linear sequence of states.
+At this point, we have obtained a linearized abstract abstract machine. 
+As \citeauthor{10.1007/3-540-15975-4_33} and \citeauthor{danvy2006a} 
+showed, ``\textit{one programmer’s lazy-list constructors is another programmer’s success 
+and failure continuations}'' \cite{10.1007/3-540-15975-4_33, danvy2006a}; the @step@ function
+in AAM is the lazy-list constructor, and here we turn it into an explicit continuation.
+Intuitively, if we imagine that the classical AAM explores a graph of reachable states, 
+then the linearized AAM flattens the graph to a linear sequence of states.
 
@@
 \begin{itemize}
-  \item By representing the continuations with monads\cite{filinski1994representing}, 
+  \item By replacing the continuations with monads \cite{filinski1994representing}, 
     we obtain the abstract definitional interpreters in monadic style which are similar to what 
@@
 \textit{Defunctionalization and Refunctionalization.}
-\citeauthor{Reynolds:72} first shows defunctionalization as a program transformation 
+\citeauthor{Reynolds:72} first showed defunctionalization as a program transformation 
 technique that can be used to transform higher-order functions to first-order data types and 
@@
 monadic evaluators \cite{ager2005functional}, as well as languages with richer constructs
-\cite{Danvy:2008:DIP:1411204.1411206, danvy2009towards, biernacka2009towards}.
+\cite{Danvy:2008:DIP:1411204.1411206, danvy2009towards, biernacka2009towards}, etc.
 

### 2018-06-30 paper/paper.tex
misc
fix typo

@@
 to the next target closure of @f@. The meta-continuation of the outer @nd@ is just 
-@continue@, as what we normally would do for CPS interpreters.
+@continue@, as what we normally would do when writing CPS interpreters.
 
@@
 Racket \cite{Tobin-Hochstadt:2012:HSE:2384616.2384655} can also be modeled by the abstracting
-abstract machine approach. \\
+abstract machines approach. \\
 

### 2018-06-30 paper/paper.tex
revise a bit on section 6.1

@@
 Readers familiar with functional programming may think of @fold@ at this point; indeed,
-it is a variant implementation of @fold@ in continuation-passing style. But here we have two
-continuations, one for the next element in the collection, and one for the eventual result.
-This is sometimes referred to as an extended continuation-passing style (ECPS)
-that function @k@ is a continuation and @m@ is a meta-continuation 
-\cite{Danvy:1990:AC:91556.91622}. \\
+it is a variant implementation of @fold@ in continuation-passing style. But as we mentioned 
+in Section \ref{linear}, here we have two continuations, one continuation @k@ for the next 
+element in the collection, and one meta-continuation @m@ for the eventual accumulator.
+This is sometimes referred to as an extended continuation-passing style (ECPS) \cite{Danvy:1990:AC:91556.91622}. \\
 

### 2018-07-01 paper/paper.tex
front and back edits

@@
 \begin{abstract}
-  Abstracting abstract machines (AAM) is a systematic methodology for constructing 
-  abstract interpreters that are derived from concrete small-step abstract machines.
-  \citeauthor{darais2017abstracting} apply the same idea on monadic definitional interpreters, 
-  and obtain monadic abstract definitional interpreters (ADI) in big-step style.
-  Yet, the relations between small-step abstract abstract machines and big-step 
-  abstract definitional interpreters are not well studied.
-
-  In this paper, we show their functional correspondence and how to syntactically 
-  transform small-step abstract abstract machines into big-step abstract definitional
-  interpreters.
-  The transformations include linearization, lightweight fusion, disentanglement, 
-  refunctionalization, and finally from CPS back to direct-style with delimited 
-  control operators. 
-  Linearization expresses nondeterministic choices by first-order data types, 
-  after which refunctionalization transforms the first-order data types representing continuations
+  Abstracting abstract machines is a systematic methodology for constructing 
+  sound static analyses for higher-order languages, by deriving small-step \emph{abstract} 
+  abstract machines (AAM) that perform abstract interpretation from abstract machines that 
+  perform concrete evaluation.
+  \citeauthor{darais2017abstracting} apply the same underlying idea to monadic definitional interpreters, 
+  and obtain monadic \emph{abstract} definitional interpreters (ADI) that perform
+  abstract interpretation in big-step style using monads.
+  Yet, the relation between small-step abstract abstract machines and big-step 
+  abstract definitional interpreters is not well studied.
+
+  In this paper, we explain their functional correspondence and demonstrate how to 
+  syntactically transform small-step abstract abstract machines into big-step abstract 
+  definitional interpreters.
+  Building on known semantic interderivation techniques from the concrete evaluation
+  setting, the transformations include linearization, lightweight fusion, disentanglement, 
+  refunctionalization, and the left inverse of the CPS transform.
+  Linearization expresses nondeterministic choice through first-order data types, 
+  after which refunctionalization transforms the first-order data types that represent continuations
   into higher-order functions. 
-  The refunctionalized AAM is an abstract interpreter written in continuation-passing style
-  with two layers of continuations.
+  The refunctionalized AAM is an abstract interpreter written in continuation-passing style (CPS)
+  with two layers of continuations, which can be converted back to direct style with delimited 
+  control operators. Based on the known correspondence between delimited control and monads, 
+  we demonstrate that the explicit use of monads in abstract definitional interpreters
+  is optional.
 
@@
 
-Defining a language by building an interpreter for it can be traced to the very
-early days of programming languages research \cite{McCarthy:1960:RFS:367177.367199,landin1966next, Reynolds:72}.
-Nowadays, even undergraduate students in computer science learn how to define
+Implementing, and in some cases defining, a programming language by building an 
+interpreter for it can be traced to the very early days of programming languages 
+research~\cite{McCarthy:1960:RFS:367177.367199,landin1966next,Reynolds:72}.
+Nowadays, even undergraduate students in computer science learn how to implement
 their own programming languages by building an interpreter.
-However, building a sound abstract interpreter remained an esoteric and difficult task
-until very recently.
-
-\citeauthor{van2012systematic} proposed the Abstracting Abstract Machines (AAM)
+However, implementing a sound static analysis by building an
+abstract interpreter~\cite{cousot1977abstract} remained an esoteric and difficult 
+task until very recently.
+
+\Citet{van2012systematic,van2010abstracting} proposed the Abstracting Abstract Machines (AAM)
 methodology which provides a recipe for constructing sound abstract interpreters
-for higher-order functional languages from concrete abstract machines
-~\cite{van2012systematic, van2010abstracting}.
-Given a concrete small-step abstract machine (e.g. the CESK machine, Krivine's
-machine, etc.), by allocating continuations in the store and bounding both the
-value addresses and continuation addresses to be finite, we obtain an abstract
-interpreter with a finite state space which can be used for performing sound
-static analysis.
-One can further instantiate different polyvariant control-flow analyses by using
-different address allocators \cite{Gilray:2016:ACP:2951913.2951936}.
-
-Applying the same idea to monadic definitional interpreters, \citeauthor{darais2017abstracting}
-built monadic abstract definitional interpreters (ADI) in big-step style
-\cite{darais2017abstracting}. One of the advantages of a monadic interpreter is
+for higher-order functional languages from concrete abstract machines.
+%~\cite{van2012systematic, van2010abstracting}.
+Given a concrete small-step abstract machine that models a store 
+(e.g. the CESK machine, Krivine's machine, etc.), if we allocate continuations 
+in the store and draw all allocated addresses (both for regular values and 
+for continuations) from a finite set, then
+we obtain an abstract interpreter with a finite state space, which can be used for 
+performing sound static analysis and which is guaranteed to terminate.
+By using different address allocators \cite{Gilray:2016:ACP:2951913.2951936},
+one can further instantiate different polyvariant control-flow analyses.
+
+Applying the same idea to monadic definitional interpreters, \citet{darais2017abstracting}
+built monadic abstract definitional interpreters (ADI) in big-step style. 
+One of the advantages of a monadic interpreter is
 that it is modular and composable. By changing the underlying monads, the definition
@@
 interpreters, respectively. An interesting question, and the subject of this paper, is
-how one can derive an abstract semantic artifact given the counterpart.
+how one can derive one abstract semantic artifact given its counterpart.
+The main contribution of this paper is to explain the functional correspondence
+between these abstract semantic artifacts, and to demonstrate the concrete
+syntactic transformations necessary to transform an abstract abstract machine 
+into its corresponding abstract definitional interpreter.
 
@@
 
-\subsection{Contributions}
-
-We begin by reviewing some background necessary for this work in Section~\ref{background},
+\paragraph{Contributions and Outline}
+
+We begin by reviewing necessary background in Section~\ref{background},
 as well as by introducing some of the basic code structures used throughout the paper.
@@
 
-In this functional pearl, we bridge the gap between small-step abstract abstract machines
+In this Functional Pearl, we bridge the gap between small-step abstract abstract machines
 and big-step abstract definitional interpreters by applying a series of syntactic 
-transformations. Among these transformations, linearization turns a work-list into another
+transformations to transform the first into the latter. Among these transformations, linearization turns a work-list into an additional
 layer of continuations; refunctionalization converts the first-order data types representing
-continuations to higher-order functions; and finally a left-inverse CPS transformation converts
-the abstract interpreter into a sequence of expressions with delimited control operators,
-which looks almost no different from a concrete interpreter.
-
-We show that a functional correspondence exists not only between concrete semantic artifacts
-but also between abstract semantic artifacts. An interesting open question would
-be whether there also exist correspondences between static analyses that are formalized 
+continuations to higher-order functions; and finally the left-inverse of the CPS transform converts
+the CPS abstract interpreter into a direct-style abstract interpreter with delimited control operators, which looks almost identical to the corresponding concrete interpreter.
+
+This sequence of transformation demonstrates that a functional correspondence exists 
+not only between concrete semantic artifacts but also between abstract semantic artifacts. 
+An interesting open question would be whether there also exist correspondences between 
+static analyses that are formalized 
 in different denotational and operational styles.
@@
 
-  This material is based upon work supported by the
+  This work was supported in part by the
   \grantsponsor{GS100000001}{National Science
@@
   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the authors and do not necessarily reflect the views of the
-  National Science Foundation. \note{TODO: grants}
+  No.~\grantnum{GS100000001}{mmmmmmm} and by the
+  \grantsponsor{GS100000001}{DOE}{XXX} under Grant
+  No.~\grantnum{GS100000001}{nnnnnnn}.  \note{TODO: grants}
 \end{acks}

### 2018-07-01 paper/paper.tex
more edits front and back

@@
 between these abstract semantic artifacts, and to demonstrate the concrete
-syntactic transformations necessary to transform an abstract abstract machine 
-into its corresponding abstract definitional interpreter.
+syntactic transformation steps necessary to transform an abstract abstract machine 
+into its corresponding abstract definitional interpreter. The reverse direction
+is comparatively easier, and can be obtained by following the respective inverse 
+steps in the opposite order. This contribution fills an intellectual gap in our 
+understanding of abstract semantic artifacts, and it also opens up further practical 
+avenues for constructing static analysis tools based on principled and well-understand
+techniques.
 
@@
 iteratively updated in a monotonic way and guaranteed to reach the least fixed-point eventually.
-In this functional pearl, we make a first step in answering the question of interderiving
-abstract semantics artifacts, and provide a constructive answer to relate AAM and ADI
-by presenting a series of syntactic transformations on the program from small-step
-abstract abstract machines to big-step abstract definitional interpreters.
+% In this Functional Pearl, we make a first step in answering the question of interderiving
+% abstract semantics artifacts, and provide a constructive answer to relate AAM and ADI
+% by presenting a series of syntactic transformations on the program from small-step
+% abstract abstract machines to big-step abstract definitional interpreters.
 
 In addition, those abstract abstract machines with an unbounded stack naturally correspond
-to abstract definitional interpreters. We show that after refunctionalizing an AAM with an
-unbounded stack (and utilizing the proper caching algorithm), the pushdown control-flow analysis
-can be obtained.
+to abstract definitional interpreters. We show that refunctionalizing an AAM with an
+unbounded stack (and utilizing the proper caching algorithm) leads to a pushdown control-flow analysis.
 
@@
 In this section, we describe a variant of AAM that allows the stack to be unbounded 
-which uses a precise model of call stack as we did in the concrete CESK machine.
+which uses a precise model of the call stack as we did in the concrete CESK machine.
 Instead of allocating continuations in the store and embedding addresses of 
@@
 \textit{Defunctionalization and Refunctionalization.}
-\citeauthor{Reynolds:72} first showed defunctionalization as a program transformation 
-technique that can be used to transform higher-order functions to first-order data types and 
-its first-order dispatching function when writing definitional interpreters
-\cite{Reynolds:72, Reynolds:HOSC98-revisited}. As generalized closure conversion,
+\citet{Reynolds:72} proposed defunctionalization as a program transformation 
+technique that can be used to transform higher-order functions to first-order data types. 
+While the technique is very general, it was first presented as a key step in
+transforming higher-order definitional interpreters into their first-order 
+counterparts.
+As a form of generalized closure conversion,
 defunctionalization is widely used in compilation and analysis \cite{pottier2006polymorphic, 
 Eisenberg:2014:PFT:2633357.2633361, fourtounis2014modular, 10.1007/3-540-46425-5_4, consel1993tour}.
-Refunctionalization~\cite{Danvy:2006:RW:2171265.2171268, Danvy:2001:DW:773184.773202} as a 
-left inverse of defunctionalization, was firstly considered by \citeauthor{Danvy:2001:DW:773184.773202}.
-
-\citeauthor{Ager:2003:FCE:888251.888254} revealed that defunctionalization and refunctionalization
+Refunctionalization~\cite{Danvy:2006:RW:2171265.2171268} as a 
+left inverse of defunctionalization, was introduced 
+by \citeauthor{Danvy:2006:RW:2171265.2171268,DANVY2009534}.
+
+\Citet{Ager:2003:FCE:888251.888254} observed that defunctionalization and refunctionalization
 can be used to show the functional correspondence between interpreters and abstract machines
@@
 and vice versa by refunctionalizing \cite{DANVY2009534}. 
-Such semantic artifacts include SECD machines \cite{10.1007/11431664_4, Ager:2003:FCE:888251.888254, DanvySECDJ},
-CEK machines \cite{Ager:2003:FCE:888251.888254}, CLS machines  \cite{Ager:2003:FCE:888251.888254},
+Such semantic artifacts include the SECD machine \cite{10.1007/11431664_4, Ager:2003:FCE:888251.888254, DanvySECDJ},
+the CEK machine \cite{Ager:2003:FCE:888251.888254}, the CLS machine  \cite{Ager:2003:FCE:888251.888254},
 Categorical Abstract Machines \cite{Ager:2003:FCE:888251.888254}, lazy abstract machines \cite{AGER2004223}, 
@@
 Besides, Danvy and his collaborators observed the correspondence between evaluation contexts and continuations.
-For deterministic languages, \citeauthor{Danvy:2008:DIP:1411204.1411206} showed that reduction contexts
+For deterministic languages, \citet{Danvy:2008:DIP:1411204.1411206} showed that reduction contexts
 are defunctionalized (data type representing) continuations and reduction semantics is a structural 
-operational semantics in continuation-passing style \cite{Danvy:2008:DIP:1411204.1411206}.
+operational semantics in continuation-passing style.
 Not unexpectedly, the correspondence also exists for nondeterministic languages. For example,
-\citeauthor{Danvy:2001:DW:773184.773202} showed that refunctionalizing a regular expression matcher yields
-its counterpart with two layer of continuations \cite{Danvy:2001:DW:773184.773202}.
-And the operational semantics of control operators (e.g., @reset@ and @shift@) can be given by an 
+\citet{Danvy:2001:DW:773184.773202} showed that refunctionalizing a regular expression matcher yields
+its counterpart with two layer of continuations.
+In addition, the operational semantics of control operators (e.g., @shift@ and @reset@) can be given by an 
 abstract machine with multi-layer continuations \cite{Danvy:1990:AC:91556.91622, lmcs:2269}.
 
-More generally, other kinds of programs, instead of merely semantic artifacts, are also found to be applicable.
-As \citeauthor{Danvy:2001:DW:773184.773202} pointed out, defunctionalization is dual to Church encoding.
-Examples include backtracking algorithms (e.g., regular expression matcher), Dyck word recognizer, and 
-Dijkstra's shunting-yard algorithm and even Quicksort algorithm \cite{Danvy:2006:RW:2171265.2171268, 
+But as mentioned above, defunctionalization and refunctionalization are more general concepts, with a wide range of uses.
+For example, \citet{Danvy:2001:DW:773184.773202} showed that defunctionalization is dual to Church encoding, and further case studies where known
+higher-order and first-order implementations can be related through defunctionalization
+and refunctionalization include backtracking algorithms (e.g., regular expression matcher), Dyck word recognizers, Dijkstra's shunting-yard algorithm, and even Quicksort \cite{Danvy:2006:RW:2171265.2171268,
 DANVY2009534, Danvy:2001:DW:773184.773202}. \\
@@
 \textit{Lightweight Fusion.}
-Lightweight fusion as one of the transformation in our paper is also studied by 
+Lightweight fusion as one of the transformation in our paper was presented by 
 \citeauthor{Ohori:2007:LFF:1190216.1190241} \cite{Ohori:2007:LFF:1190216.1190241}.
@@
 environment when allocating a continuation addresses. No other pieces of code would need to be
-modified. Notably, this change only causes constant-factor increase in time complexity of
+modified. Notably, this change only causes a constant-factor increase in time complexity of
 the analysis if the store is widened.
 
-Our work establish the pushdown control-flow analysis through refunctionalization and a
-proper caching algorithm. The call/return is naturally matched by the call stack of the
-meta-language. \\
+Our work establishes the pushdown control-flow analysis through refunctionalization and a
+proper caching algorithm. The call/return flow is naturally matched by the call structure 
+of the meta-language. \\
 
@@
 Reynolds' seminal paper \emph{Definitional interpreters for higher-order programming languages} 
-\cite{Reynolds:72, Reynolds:HOSC98-revisited} shows that in some definitional interpreters, the
-defined language can inherit properties from the defining language. With this insight,
-\citeauthor{darais2017abstracting} construct abstract definitional interpreters which
+\citet{Reynolds:72, Reynolds:HOSC98-revisited} showed that in definitional interpreters, the
+defined language inherits properties such as the order of evaluation from the defining
+language, unless these properties are made explicit in the 
+interpreter. With this insight,
+\citet{darais2017abstracting} constructed abstract definitional interpreters which
 automatically inherit the pushdown control-flow property from its defining language because
-the defined language simply uses the call-stack model of the meta-language~\cite{darais2017abstracting}.
+the defined language simply uses the call-stack model of the meta-language.
 \citeauthor{darais2017abstracting}'s abstract definitional interpreters work on direct-style
-$\lambda$-calculus, and itself is written in monadic style. One of the advantages of monadic
-abstract interpreters is modular and composable, therefore deploying different sensitivities
-or features is just applying a different monad. Prior to that, \citeauthor{Sergey:2013:MAI:2491956.2491979}
-show a monadic abstract interpreter for small-step semantics \cite{Sergey:2013:MAI:2491956.2491979}.
-
-This paper is greatly inspired by~\citeauthor{darais2017abstracting}'s work \cite{darais2017abstracting}.
-The difference on the surface is that ours works for A-Normal Form $\lambda$-calculus, instead of plain 
-$\lambda$-calculus, although our work can also be easily adapted to handle a plain $\lambda$-calculus. 
-But besides directly presenting the final form of abstract definitional interpreters, we reveal that 
-refunctionalization plays an important role for inheriting the model of the stack from the meta-language. 
-The target of these transformations in this paper is abstract definitional interpreters, and we 
-additionally demonstrate abstract interpreters written in direct-style with delimited control operators. 
-We use the delimited control operators @reset@ and @shift@ that are implemented in Scala 
-\cite{rompf2009implementing}. The correspondence of delimited continuations and monads are well-known
+$\lambda$-calculus, and are themselves written in monadic style. One of the advantages of monadic
+abstract interpreters is modularity. Hence, deploying different sensitivities
+or features can be achieved by just applying a different monad. Prior to that, \citet{Sergey:2013:MAI:2491956.2491979}
+presented a monadic abstract interpreter for small-step semantics.
+
+This paper is greatly inspired by the work of~\citet{darais2017abstracting}.
+A first surface-level difference is that our presentation uses A-Normal Form $\lambda$-calculus, instead of plain 
+$\lambda$-calculus, although our work could be easily adapted to handle plain $\lambda$-calculus as well. 
+Going through the transformation steps from AAMs to ADIs instead of only considering the final ADI result provides additional insights. In particular, it reveals that 
+refunctionalization plays an important role for inheriting the call structure (i.e., the model of the stack) from the meta-language. 
+In addition, we demonstrate that the explicit use of monads in ADIs is optional, by translating to direct-style ADIs with delimited control operators in addition to the monadic version of \citet{darais2017abstracting}.
+We use the delimited control operators @shift@ and @reset@ \cite{Danvy:1990:AC:91556.91622} as implemented in Scala 
+\cite{rompf2009implementing}. The correspondence of delimited continuations and monads is well known
 \cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions, filinski1994representing}.

### 2018-07-01 paper/paper.tex
figure placeholder

@@
    {
-     [start chain] \chainin (m-2-1);
+     [start chain, transform canvas={yshift=0.5ex}] \chainin (m-2-1);
      \chainin (m-2-2) [join={node[above,labeled] {
@@
      \chainin (m-2-3) [join={node[above,labeled] {\text{disentang.}}}];
+   }
+   {
+     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-2);
+     \chainin (m-2-1) [join={node[below,labeled] {\text{List(Success)??}}}];
+   }
+   {
+     [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-3);
+     \chainin (m-2-2) [join={node[below,labeled] {\text{entang.}}}];
    }
@@
  \end{tikzpicture}
- \caption{Transformations from AAM to ADI} \label{fig:trans}
+ \caption{Functional correspondence between and transformations from AAM to ADI} \label{fig:trans}
 \end{center}

### 2018-07-01 paper/paper.tex
figure placeholder

@@
      [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-2);
-     \chainin (m-2-1) [join={node[below,labeled] {\text{List(Success)??}}}];
+     \chainin (m-2-1) [join={node[below,labeled] {\begin{smallmatrix} \text{splitting} \\ \text{+} \\ \text{List(Success)??} \end{smallmatrix}}}];
    }

### 2018-07-01 paper/paper.tex
contrib list

@@
 \section{Introduction}
+\vspace{-0.7ex}
 
@@
 we obtain an abstract interpreter with a finite state space, which can be used for 
-performing sound static analysis and which is guaranteed to terminate.
+performing sound static analysis and is guaranteed to terminate.
 By using different address allocators \cite{Gilray:2016:ACP:2951913.2951936},
@@
 question of deriving one from the other, are not well studied.
-One of the fundamental differences between concrete semantic artifacts and abstract semantic
-artifacts is that a sound instance of the latter must tell you facts about all possible execution 
-paths, and thus introduce nondeterminism. In addition to ensuring termination, abstract semantic
+One of the fundamental differences between concrete and abstract semantic
+artifacts is that a sound instance of the latter must consider facts about all possible execution 
+paths, and thus introduce a layer of nondeterminism. In addition to ensuring termination, abstract semantic
 artifacts are usually equipped with a cache of reachable states; the cache is
-iteratively updated in a monotonic way and guaranteed to reach the least fixed-point eventually.
+iteratively updated in a monotonic way, guaranteed to reach the least fixed-point eventually.
 % In this Functional Pearl, we make a first step in answering the question of interderiving
@@
 \begin{itemize}
-  \item We show \textbf{linearization} in Section~\ref{linear}. By expressing the
+  \item \textbf{Linearization}: By expressing the
     nondeterministic choices as a first-order data type, we linearize the execution
     of abstract abstract machines; the transition of machine states therefore becomes
-    deterministic. Notably, this explicitly expresses another layer of control that
-    will also be refunctionalized later.
-
-  \item In Section~\ref{fusing}, we then apply a \textbf{lightweight fusion} 
-    \cite{DANVY2008100, Ohori:2007:LFF:1190216.1190241} to the linearized AAM. 
-    The fusing transformation simply combines the single-step function @step@ and the 
-    driving function into one, but keeps all the machine state representations.
-
-  \item Section~\ref{disen} discusses \textbf{disentanglement}. In this part, we identify 
+    deterministic. Notably, this makes explicit another layer of control (Section~\ref{linear}). 
+    %that will also be refunctionalized later.
+
+  \item \textbf{Lightweight fusion}~\cite{DANVY2008100,Ohori:2007:LFF:1190216.1190241}:
+    We apply a fusion transformation to the linearized AAM, which
+    combines the single-step function @step@ and the 
+    driving function into one, but keeps all the machine state representations intact
+    (Section~\ref{fusing}).
+
+  \item \textbf{Disentanglement}: We identify 
     different first-order data types which represent different layers of continuations, and
-    disassemble their handlers into separate functions.
-
-  \item Applying \textbf{refunctionalization} \cite{DANVY2009534, Danvy:2006:RW:2171265.2171268} 
-    to disentangled AAM is shown in Section~\ref{sec:refunc}. 
+    disassemble their handlers into separate functions 
+    (Section~\ref{disen}).
+
+  \item \textbf{Refunctionalization}~\cite{DANVY2009534,Danvy:2006:RW:2171265.2171268}:
+    We apply refunctionalization to the disentangled AAM, which
     %as the left inverse of Reynolds' defunctionalization \cite{Reynold:72,Reynolds:HOSC98-revisited} 
-    Refunctionalization replaces the first-order data types representing continuations with
+    replaces the first-order data types representing continuations with
     higher-order functions, and associates dispatching logic with proper higher-order functions. 
-    We then derive an abstract interpreter written in continuation-passing style with two
+    We obtain an abstract interpreter written in continuation-passing style with two
     layers of continuations due to the nondeterminism.
-
+%
     For clarity, we first present the refunctionalized AAM without caching. 
@@
     In the end of this section, we review the pushdown control-flow analysis and examine how 
-    computable and precise call/return matching is obtained through these transformations.
-
-  \item The last transformation we show is that of transforming the refunctionalized AAM back to a
-    \textbf{direct-style} interpreter (Section~\ref{directstyle}) by using delimited control operators.
+    computable and precise call/return matching is obtained through these transformations
+    (Section~\ref{sec:refunc}).
+
+  \item \textbf{Direct-style transform}~\cite{DBLP:journals/scp/Danvy94}:
+    We finally transform the refunctionalized AAM back to a
+    direct-style interpreter by using delimited control operators
+    (Section~\ref{directstyle}).
 \end{itemize}

### 2018-07-01 paper/paper.tex
more edits througout

@@
     higher-order functions, and associates dispatching logic with proper higher-order functions. 
-    We obtain an abstract interpreter written in continuation-passing style with two
-    layers of continuations due to the nondeterminism.
+    We obtain an abstract interpreter written in continuation-passing style with an
+    additional layer of continuations due to nondeterminism.
 %
@@
 simplifies analysis~\cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
-Here, we choose to use $\lambda$-calculus in administrative normal form (ANF)
-\cite{flanagan1993essence}, which is a direct-style intermediate representation, as our 
-target for clarity, but without losing simplicity and generality. The transformations 
+Here, we choose a direct-style $\lambda$-calculus as our target where all ``serious'' 
+expressions (i.e., function calls) are let-bound. This style is variously known as 
+administrative normal form (ANF) \cite{flanagan1993essence}, or monadic normal 
+form \cite{moggi1991notions,DBLP:conf/cc/Danvy03}. The transformations 
 we will show in the rest of this paper also work on abstract machines for plain 
@@
 1) \textit{Control} is the expression currently being evaluated.
-2) \textit{Environment} is a map that contains the address of a variable in the lexical scope.
+2) \textit{Environment} is a map that contains the addresses of all variables in the lexical scope.
 3) \textit{Store} models the heap of a program as a map from addresses to values.
@@
   an environment.
-4) \textit{Continuation} represents the program's call stack. In this paper, we instantiate the 
-stack as a list of frames.
+4) \textit{Continuation} represents the program's execution context. In this paper, we instantiate the 
+execution context as a call stack consisting of a list of frames.
 
@@
 
-It is worth noting that the continuation class @Kont@ is defined as a list of
-frames to represent a reduction context.
+It is worth noting that the continuation class @Kont@ corresponds to a reduction 
+context in a reduction-based formulation of the semantics.
 An empty list represents an empty context, and corresponds to halt. Otherwise, the head 
@@
 which must have a successor; the final case of a state is handled by function
-@drive@, which will be mentioned in the next section.
+@drive@, which will be mentioned in Section~\ref{sec:drive}.
 
@@
   environment and store.
-  Finally, the successor state is transferred to expression @e@ from the top frame,
+  Finally, the successor state is expression @e@ from the top frame,
   which is the body of a @Let@ expression, with the updated environment, store, and
@@
 
-\subsubsection{Valuation}
+\subsubsection{Valuation}\label{sec:drive}
 
@@
 expression and the continuation stack is empty. Naturally, we can then extract the value
-from the final state at last.
+from the final state.
 
@@
 Since the state space of concrete execution is possibly infinite, the key insight of 
-AAM approach when analyzing programs is to allocate both bindings and continuations 
+AAM approach when analyzing programs is to allocate both variable bindings and continuations 
 on the store, and then bound the addresses space to be finite. Since each component
-of state is finite, the abstracted machine-state space is also finite, and therefore
-computable.
-
-In this section, we derive the abstract abstract machine from concrete
-CESK machines, and also show how to instantiate useful $k$-call-sensitive 
+of the machine state is finite, the abstracted machine-state space is also finite, 
+and therefore computable.
+
+In this section, we derive the abstract abstract machine from the concrete
+CESK machine, and also show how to instantiate useful $k$-call-sensitive 
 control-flow analysis.
@@
 
-For clarity, we divide the store into two separate stores: the binding stores @BStore@, and
+For clarity, we divide the store into two separate stores: the binding store @BStore@, and
 the continuation store @KStore@.
@@
 @[KAddr, Cont]@, we obtain @BStore@ and @KStore@, respectively.
-We note that both the value store and continuation store are updated monotonically;
-it continuously grows and never shrinks.
+We note that both the value store and the continuation store are updated monotonically;
+they grow continuously and never shrink.
 
@@
 
-The co-domain of binding stores @Storable@ is the same as previously defined for CESK 
-machines. The c-domain of continuation stores @Cont@, on the other hand, is comprised of
+The co-domain of the binding store @Storable@ is the same as previously defined for the CESK 
+machine. The co-domain of the continuation store @Cont@, on the other hand, is comprised of
 a @Frame@ object and a continuation address @KAddr@. To mimic the run-time call stack, 
 @KAddr@ plays the role of representing the remaining stack frames.
-But since the continuation store may contain multiple continuations, the dereferencing 
+But since the continuation store may contain multiple continuations, dereferencing 
 of continuation addresses is also nondeterministic.
@@
 
-As a consequence, the components of states are also changed: the store is divided
-into binding stores and continuation stores; the continuation becomes an address
+As a consequence, the components of machine states are also changed: the store is divided
+into a binding store and a continuation store; the continuation becomes an address
 that maps to a set of continuations in @KStore@.
 By dereferencing this address in a continuation store, we can retrieve the actual 
-transfers of control. The definition of environment @Env@ remains the same.
+control-transfer destination. The definition of environment @Env@ remains the same.
 
@@
 \subsubsection{Allocating Addresses}
-Up to this point, we have not described allocating addresses in stores, nor handling
-the time stamp @Time@. In abstract interpretation, however, these are key ingredients
-to achieve analyses with different sensitivities, as well as to perform a finite
+Up to this point, we have described neither allocating of addresses in stores, nor handling
+of time stamps @Time@. In abstract interpretation, however, these are key ingredients
+to realize analyses with different sensitivities, as well as to perform a finite
 state space analysis \cite{Gilray:2016:ACP:2951913.2951936}.
-To effectively approximate the run-time behavior, we introduce a finite program contours
+To effectively approximate the run-time behavior, we introduce a finite program \emph{contour}
+\note{define}
 @time@ that encodes the program execution history. The function @tick@ is used to refresh
-the "time" and get a "new time". We use a list of expression (which are drawn from the
-control component) to encode the calling contexts history, and as we will see in Section~
+the ``time'' and get a ''new time''. We use a list of expressions (which are drawn from the
+control component) to encode the calling context history, and as we will see in Section~
 \ref{kcfainst}, by applying different @tick@ functions on the timestamp, we are able to
@@
 As previously mentioned, the space of states is finite when the space of addresses is finite.
-To make this happen, binding addresses are parameterized by variable names and the creation
-time of the binding, both of which are finite. Continuation addresses @KAddr@ has two variants:
+To make this happen, addresses of variable bindings are parameterized by variable names and the creation
+time of the binding, both of which are finite. Continuation addresses @KAddr@ have two variants:
 1) @Halt@ which corresponds to the empty stack, and
-2) @ContAddr@ consists of the target expressions of callee, which are also finite.
+2) @ContAddr@ which consists of the call target expressions, also a finite set.
 %One may ask, why not also keep track of the time in continuation addresses?
@@
 
-Given that the space of addresses is finite, we can conclude that there are finite number
+Given that the space of addresses is finite, we can conclude that there are only finite numbers
 of environments and stores because the numbers of variables and closures are also finite.
-This property guarantees a finite space of reachable states, and we can always have a
-terminated analysis through Kleene's fixed-point iteration.
+This property guarantees a finite space of reachable states, and we can always realize a
+terminating analysis through Kleene's fixed-point iteration.
 
@@
 continuations in states, we intend to use a list of frames to explicitly model 
-the stack. By doing so, we recover the call structure as the same as run-times
+the stack. By doing so, we recover the call structure as the same as at run-time
 (so called pushdown control-flow analysis), but since the stack is unbounded, 
@@
 
-In the first case of pattern matching, we may have multiple choices of closure for callee @f@.
+In the first case of pattern matching, we may have multiple possible closures for callee @f@.
 For each closure in the set, a new frame is constructed and pushed onto the stack.
@@
 if conditionals exist in the analyzed language. But the reduction context (i.e., the @Frame@ in
-our program) is \emph{not} nondeterministic, and actually the work-list @todo@ implicitly
+our program) is \emph{not} nondeterministic, and the work-list implicitly
 handles the nondeterminism. Thus, if we simply refunctionalize the frames to higher-order
@@
 
-In Danvy's paper \textit{Defunctionalized Interpreters for Programming Languages},
-he mentions that for deterministic languages, a reduction semantics is a structural
-operational semantics in continuation-passing style, where the reduction context is
+In his paper \textit{Defunctionalized Interpreters for Programming Languages},
+Danvy observes that for deterministic languages, a reduction semantics can be seen as 
+a structural operational semantics in continuation-passing style, where the reduction context is
 a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}. Therefore
-refunctionalizing a deterministic language yields an interpreter in continuation-passing
-style. Nevertheless, doing so for nondeterministic languages yields an interpreter
+refunctionalizing such a reduction semantics for a deterministic language yields 
+an interpreter in continuation-passing style. 
+Nevertheless, doing so for nondeterministic languages yields an interpreter
 with multiple layers of continuation \cite{Danvy:2006:RW:2171265.2171268, DANVY2009534},
@@
 we elect to call the first a \emph{continuation}, and the second a \emph{meta-continuation}
-\footnote{Traditionally, they are also called \emph{success continuation} and \emph{failure
+\footnote{These are also called \emph{success continuation} and \emph{failure
 continuation} \cite{Danvy:1990:AC:91556.91622}.}.
@@
 At this point, we have obtained a linearized abstract abstract machine. 
-As \citeauthor{10.1007/3-540-15975-4_33} and \citeauthor{danvy2006a} 
-showed, ``\textit{one programmer’s lazy-list constructors is another programmer’s success 
-and failure continuations}'' \cite{10.1007/3-540-15975-4_33, danvy2006a}; the @step@ function
-in AAM is the lazy-list constructor, and here we turn it into an explicit continuation.
+It has been observed that ``\textit{failure can be replaced by a list of successes}''
+\cite{10.1007/3-540-15975-4_33} and that ``\textit{one programmer’s lazy-list 
+constructor is another programmer’s success and failure continuations}'' \cite{danvy2006a}; 
+the @step@ function in AAM is the lazy-list constructor, and here we turn it into an 
+explicit continuation.
 Intuitively, if we imagine that the classical AAM explores a graph of reachable states, 
@@
 of the analyzed language is blended into the call stack of our defining language
-(Scala) through refunctionalization, then the call/return of the analyzed language
+(Scala) through refunctionalization, then the call/return flow of the analyzed language
 is naturally matched.
 
-In fact, what we started is an AAM with an unbounded stack which already precisely
+In fact, what we started with is an AAM with an unbounded stack which already precisely
 matches the calls and returns. Refunctionalization makes the interpreter be a ``big-step''
@@
 in extended continuation-passing style.
-To obtain definitional abstract interpreters, we now transform it further to direct-style.
-
-We have multiple choices of how to proceed:
+To obtain a definitional abstract interpreter in direct style, 
+we have multiple choices of how to proceed:
 \begin{itemize}
@@
     we obtain the abstract definitional interpreters in monadic style which are similar to what 
-    \citeauthor{darais2017abstracting} describe \cite{darais2017abstracting}.
+    \citet{darais2017abstracting} describe.
   \item By representing the extended continuation-passing style with delimited control operators
     such as @shift@ and @reset@, we derive a new form of abstract interpreters in direct-style.
+    This transformation is the left-inverse of the CPS transform \cite{DBLP:journals/scp/Danvy94,DBLP:conf/lfp/DanvyL92}.
   \item In fact, we may also just use the same caching algorithm but with side effects such as
     assignments and mutations to update the cache, and then achieve the same definitional 
-    interpreter with pushdown control-flows. \citeauthor{friedman-ai} in a tutorial on abstract 
-    interpreters use this style to model caches and updates \cite{friedman-ai}.
-    In this case, nondeterminism can be easily handled via @for@ comprehension. 
+    interpreter with pushdown control-flows. \Citet{friedman-ai} in a tutorial on abstract 
+    interpreters use this style to model caches and updates.
+    In this case, nondeterminism can be easily handled via @for@ comprehensions. 
 \end{itemize}
 
-These coincidences should not be a surprise since existing literature is extraordinarily rich
+These coincidences should not be a surprise since the literature is rich
 in showing the correspondence between monads and continuation-passing style as well as
-delimited control since their very beginning era \cite{Danvy:1990:AC:91556.91622, wadler1992essence,
+delimited control \cite{Danvy:1990:AC:91556.91622, wadler1992essence,
 danvy1992representing, moggi1991notions}.
@@
 
-Now we have arrived the end of this series of transformations. Starting from a
+Now we have arrived at the end of this series of transformations. Starting from a
 small-step abstract abstract machine, eventually we obtain a big-step definitional
-interpreter with pushdown control-flows and written in direct-style.
+interpreter with pushdown control-flow, written in direct-style.
 
@@
 \textit{Abstract Interpretation and Control-Flow Analysis.}
-\citeauthor{cousot1977abstract} invented abstract interpretation as a sound approach to
+\citeauthor{cousot1977abstract} discovered abstract interpretation as a sound approach to
 approximate a program's run-time behavior \cite{cousot1977abstract}. Control-flow analysis
-is one instance of abstract interpretation on functional programs that can be traced to
-Jones \cite{jones1981flow}. \citeauthor{Shivers:1988:CFA:53990.54007} introduced $k$-CFA
-which uses $k$ recent calling contexts as program contours that differentiate values from
-different contexts \cite{Shivers:1988:CFA:53990.54007, Shivers:1991:SSC:115865.115884}.
-The recent formulation of control-flow analysis is the abstracting abstract machines
+is one instance of abstract interpretation on higher-order functional programs that can be traced to
+Jones \cite{jones1981flow}. \Citet{Shivers:1988:CFA:53990.54007,Shivers:1991:SSC:115865.115884} 
+introduced $k$-CFA which uses $k$ recent calling contexts as program contours that 
+differentiate values from different contexts.
+A recent formulation of control-flow analysis is the abstracting abstract machines
 methodology \cite{van2012systematic, van2010abstracting} which forms the starting point
@@
 \textit{Pushdown Control-Flow Analysis.}
-In recent years, there are significant efforts~\cite{vardoulakis2010cfa2, earl2012introspective,
-gilray2016pushdown, johnson2015abstracting} to achieve precise call/return match based on small-step
-abstract abstract machines.
-
-CFA2 is the first solution that solves the return-flows problem~\cite{vardoulakis2010cfa2},
-but CFA2 has several drawbacks: it works only on continuation-passing style programs, and
+There have been significant efforts to achieve precise call/return matching 
+based on small-step abstract abstract machines~\cite{vardoulakis2010cfa2, earl2012introspective,
+gilray2016pushdown, johnson2015abstracting}.
+
+CFA2 is the first solution that solves the return-flow problem~\cite{vardoulakis2010cfa2},
+but CFA2 has several limitation: it only works on continuation-passing style programs, and
 does not support polyvariant analysis, in addition to having an exponential time complexity.
@@
 
-\citeauthor{gilray2016pushdown} further proposed a polyvariant continuation-addresses allocator
-for small-step AAM to achieve pushdown analysis \cite{gilray2016pushdown}. This method is both
+\Citet{gilray2016pushdown} further proposed a polyvariant continuation-addresses allocator
+for small-step AAM to achieve pushdown analysis. This method is both
 simple to implement and computationally inexpensive and so called Pushdown for Free (P4F).
@@
 Reynolds' seminal paper \emph{Definitional interpreters for higher-order programming languages} 
-\citet{Reynolds:72, Reynolds:HOSC98-revisited} showed that in definitional interpreters, the
+\cite{Reynolds:72, Reynolds:HOSC98-revisited} showed that in definitional interpreters, the
 defined language inherits properties such as the order of evaluation from the defining

### 2018-07-01 paper/paper.tex
typos

@@
 already connect all the executions in order. In fact, the continuations (call stack)
-of the analyzed language is blended into the call stack of our defining language
-(Scala) through refunctionalization, then the call/return flow of the analyzed language
-is naturally matched.
+of the analyzed language are blended into the call stack of our defining language
+(Scala) through refunctionalization, and hence the call/return flow of the analyzed 
+language is naturally matched.
 
@@
 interpreters, and the subsequent statements and expressions become the delimited continuation
-@f@. The function @choices@ also returns the latest cache to its left-hand side when invoking.
+@f@. The function @choices@ also returns the latest cache to its left-hand side when invoked.
 The cache will be still accumulated to the next call to @choices@ or @aeval@, 

### 2018-07-01 paper/paper.tex
grammar

@@
 In fact, what we started with is an AAM with an unbounded stack which already precisely
-matches the calls and returns. Refunctionalization makes the interpreter be a ``big-step''
-semantic artifact, then the consequence is we have no places to explicitly save the context
-information in store or stack. 
+matches the calls and returns. Refunctionalization makes the interpreter a ``big-step''
+semantic artifact; the consequence is that we have no place to explicitly save the context
+information in the store or on the stack. 
 But refunctionalization not only forces the nondeterministic control flow to be higher-order,

### 2018-07-01 paper/paper.tex
more revisions

@@
 
-%For direct style $\lambda$-calculus, the continuations of CEK/CESK machines can be presented by variants of a data type.
 The Scala representations for the components of the CESK machine are as follows:
@@
 
-Similar to CESK machines, the machine state of AAM has a control, an environment, a store,
+Similar to the CESK machine, the machine state of AAM has a control, an environment, a store,
 and continuation, as well as a timestamp. However, there are several notable differences
@@
 state space analysis \cite{Gilray:2016:ACP:2951913.2951936}.
-To effectively approximate the run-time behavior, we introduce a finite program \emph{contour}
-\note{define}
-@time@ that encodes the program execution history. The function @tick@ is used to refresh
-the ``time'' and get a ''new time''. We use a list of expressions (which are drawn from the
-control component) to encode the calling context history, and as we will see in Section~
-\ref{kcfainst}, by applying different @tick@ functions on the timestamp, we are able to
-obtain a family of analyses.
+To effectively approximate the run-time behavior, we introduce program \emph{contours} @time@ 
+which are finite history of funciton calls till up to the current state \cite{Shivers:1991:SSC:115865.115884}.
+The function @tick@ is used to refresh the ``time'' and get a ''new time''. 
+We use a finite list of expressions (which are drawn from the control component) to encode the 
+calling context history, and as we will see in Section~\ref{kcfainst}, by applying different 
+@tick@ functions on the timestamp, we are able to obtain a family of analyses.
 
@@
     represent. For each closure in the set, we perform nearly the same operations as in the
-    concrete CESK machines, with an important difference: the continuation is allocated on 
+    concrete CESK machine, with an important difference: the continuation is allocated on 
     the store @kstore@, so a new continuation address @new_kaddr@ must be constructed and 
@@
 
-Similar to the CESK machines, to run (analyze) a program we first use the @inject@ function
+Similar to the CESK machine, to run (analyze) a program we first use the @inject@ function
 to construct the initial state given to the program. Note that the initial continuation store
@@
 For each closure in the set, a new frame is constructed and pushed onto the stack.
-The code for handling the second case (atomic expressions) is the same as the concrete CESK machines. \\
+The code for handling the second case (atomic expressions) is the same as the concrete CESK machine. \\
 
@@
 At this point, we have obtained a linearized abstract abstract machine. 
+Intuitively, if we imagine that the classical AAM explores a graph of reachable states, 
+then the linearized AAM flattens the graph to a linear sequence of states.
 It has been observed that ``\textit{failure can be replaced by a list of successes}''
@@
 constructor is another programmer’s success and failure continuations}'' \cite{danvy2006a}; 
-the @step@ function in AAM is the lazy-list constructor, and here we turn it into an 
-explicit continuation.
-Intuitively, if we imagine that the classical AAM explores a graph of reachable states, 
-then the linearized AAM flattens the graph to a linear sequence of states.
+the @step@ function in AAM is the list constructor used in a lazy way, and here we turn it 
+into an explicit data types representing continuation.
+The inverse transformation of this step is to apply \citeauthor{10.1007/3-540-15975-4_33}'s 
+technique that converts the data types representing meta-continuations (failure continuations) 
+back to list representation, where the elements in the list are plain machine states.
 
@@
 @step@ into the @drive@ function.
-@drive_step@ takes an @NDState@ and a set of explored states as an argument
+@drive_step@ takes an @NDState@ and a set of explored states @seen@ as arguments
 and returns a set of reachable states once it terminates.
@@
 
+With fusing complete, our AAM appears similar to a ``big-step'' interpreter, although it
+still has the machine state representation inside. The inverse transformation of this step
+would be extracting the work-list algorithm part to an individual function.
+
 \section{Disentanglement} \label{disen}
 
-With fusing complete, our AAM appears similar to a ``big-step'' interpreter, although it
-still has the machine state representation inside. In the disentanglement transformation,
-we identify the first-order data types which represent reduction contexts
-and lift the code blocks that handle these reduction contexts to be individual functions.
+In the disentanglement transformation, we identify the first-order data types which represent 
+reduction contexts and lift the code blocks that dispatch these reduction contexts to be individual 
+functions.
 After disentanglement, we will obtain an abstract abstract machine that in defunctionalized form.
@@
 At the end of the first case of pattern matching, we invoke @drive_step@ recursively 
+on the new @NDState@.
 \footnote{The code for constructing new environments, stores, and frames is elided for 
@@
 The code for the second case is replaced entirely by a function call to @continue@.
-The shape of an interpreter in continuation-passing style has emerged.
+The shape of an interpreter in continuation-passing style has emerged, even though
+the @continue@ function is not higher-order for the moment.
 
@@
 Function @continue@ is mainly what we have seen for the second case in @drive_step@,
-except that @ndcontinue@ is called when the continuation @konts@ is empty.
-It works on the list representation of context: An empty list @Nil@ represents the
+except that the dispatching logic for empty @konts@ is lifted to a separate function 
+@ndcontinue@.
+It matches on the list representation of context: An empty list @Nil@ represents the
 halt of a computation path, whereas a cons @::@ means the top frame of the list
-holds the binding variable and context.
-An analogy to this in standard CEK machines would be an abstract class (e.g., @cont@)
-with three variants: @halt@, @arg(value, cont)@, and @arg(exp, env, cont)@.
+holds the binding variable and the innermost context.
+%An analogy to this in standard CEK machines would be an abstract class (e.g., @cont@)
+%with three variants: @halt@, @arg(value, cont)@, and @arg(exp, env, cont)@.
 
@@
 
-The shape for @ndcontinue@ is similar to @continue@; after
-all, they both use a @List@ representation for continuations.
-When the closure set is empty, we recursively call @drive_step@ with the rest
-of @ndk@, since in this case the normal stack @konts@ is still empty, so
-the state will be dispatched to @ndcontinue@ again.
-Otherwise, we invoke @drive_step@ with the body expression of the next closure object.
+The shape for @ndcontinue@ is similar to @continue@; after all, they both use a 
+@List@ representation for continuations. When the closure set is empty, we recursively 
+call @drive_step@ with the rest of @ndk@, since in this case the continuation 
+@konts@ is still empty, so the state will be dispatched to @ndcontinue@ again.
+Otherwise, we invoke @drive_step@ with the a new @NDState@ where the control 
+is transferred to the body expression of the next closure object.
+
+The purpose of disentanglement is to reveal the defunctionalized form of AAM.
+Functions @continue@ and @ndcontinue@ that dispatches different data types representing 
+continuations are disentangled from @drive_step@. 
+The inverse transformation of this step would be simply inlining @ndcontinue@ into @continue@,
+and then inlining @continue@ into @drive_step@.
 
@@
     such as @shift@ and @reset@, we derive a new form of abstract interpreters in direct-style.
-    This transformation is the left-inverse of the CPS transform \cite{DBLP:journals/scp/Danvy94,DBLP:conf/lfp/DanvyL92}.
+    This transformation is the left-inverse of the CPS transform \cite{DBLP:journals/scp/Danvy94, DBLP:conf/lfp/DanvyL92}.
   \item In fact, we may also just use the same caching algorithm but with side effects such as
@@
 
-But as mentioned above, defunctionalization and refunctionalization are more general concepts, with a wide range of uses.
-For example, \citet{Danvy:2001:DW:773184.773202} showed that defunctionalization is dual to Church encoding, and further case studies where known
+But as mentioned above, defunctionalization and refunctionalization are more general concepts, 
+with a wide range of uses. For example, \citet{Danvy:2001:DW:773184.773202} showed that 
+defunctionalization is dual to Church encoding, and further case studies where known
 higher-order and first-order implementations can be related through defunctionalization
-and refunctionalization include backtracking algorithms (e.g., regular expression matcher), Dyck word recognizers, Dijkstra's shunting-yard algorithm, and even Quicksort \cite{Danvy:2006:RW:2171265.2171268,
-DANVY2009534, Danvy:2001:DW:773184.773202}. \\
+and refunctionalization include backtracking algorithms (e.g., regular expression matcher), 
+Dyck word recognizers, Dijkstra's shunting-yard algorithm, and even Quicksort 
+\cite{Danvy:2006:RW:2171265.2171268, DANVY2009534, Danvy:2001:DW:773184.773202}. \\
 
@@
 This paper is greatly inspired by the work of~\citet{darais2017abstracting}.
-A first surface-level difference is that our presentation uses A-Normal Form $\lambda$-calculus, instead of plain 
-$\lambda$-calculus, although our work could be easily adapted to handle plain $\lambda$-calculus as well. 
-Going through the transformation steps from AAMs to ADIs instead of only considering the final ADI result provides additional insights. In particular, it reveals that 
-refunctionalization plays an important role for inheriting the call structure (i.e., the model of the stack) from the meta-language. 
-In addition, we demonstrate that the explicit use of monads in ADIs is optional, by translating to direct-style ADIs with delimited control operators in addition to the monadic version of \citet{darais2017abstracting}.
+A first surface-level difference is that our presentation uses A-Normal Form $\lambda$-calculus, 
+instead of plain $\lambda$-calculus, although our work could be easily adapted to handle plain 
+$\lambda$-calculus as well. Going through the transformation steps from AAMs to ADIs instead of 
+only considering the final ADI result provides additional insights. In particular, it reveals that 
+refunctionalization plays an important role for inheriting the call structure (i.e., the model of 
+the stack) from the meta-language. 
+In addition, we demonstrate that the explicit use of monads in ADIs is optional, by translating to
+direct-style ADIs with delimited control operators in addition to the monadic version of \citet{darais2017abstracting}. 
 We use the delimited control operators @shift@ and @reset@ \cite{Danvy:1990:AC:91556.91622} as implemented in Scala 
@@
 and big-step abstract definitional interpreters by applying a series of syntactic 
-transformations to transform the first into the latter. Among these transformations, linearization turns a work-list into an additional
-layer of continuations; refunctionalization converts the first-order data types representing
-continuations to higher-order functions; and finally the left-inverse of the CPS transform converts
-the CPS abstract interpreter into a direct-style abstract interpreter with delimited control operators, which looks almost identical to the corresponding concrete interpreter.
+transformations to transform the first into the latter. Among these transformations, 
+linearization turns a work-list into an additional layer of continuations; refunctionalization 
+converts the first-order data types representing continuations to higher-order functions; 
+and finally the left-inverse of the CPS transformation converts the CPS abstract interpreter 
+into a direct-style abstract interpreter with delimited control operators, which looks almost 
+identical to the corresponding concrete interpreter.
 

### 2018-07-01 paper/paper.tex
revised some text

@@
 still has the machine state representation inside. The inverse transformation of this step
-would be extracting the work-list algorithm part to an individual function.
+would be splitting the work-list algorithm apart to an individual function.
 
@@
 In the first case of pattern matching, this happens at the innermost continuation of @aeval@,
-since it is where we reach the end of one computation path.
-For the second case, we update the @out@ cache before calling the @continue@ continuation.
+where we reach the end of one computation path.
+For the second case, we update the @out@ cache before calling the @continue@ function.
 
@@
 Finally, the @analyze@ function (as the entrance of the analysis) does a looping iteration
-to find the fixed-point on caches starting from an empty cache.
+to find the fixed-point of cache starting from an empty cache.
 If the @out@ cache of this iteration is equivalent to its @in@ cache, which means no more 
-fact was discovered during this iteration, then we have reached a fixed-point that 
-over-approximates the concrete evaluations, and thus the result can be returned.
-Otherwise, the next iteration with the new @out@ cache installed at the @in@ cache
-position will be activated; in the meanwhile, an empty cache will be used as the @out@
-cache. Note that the continuation of the initial call to @aeval@ sets up the values for
-the initial configuration in the final cache.
+fact was discovered during this iteration, then we have reached a fixed-point of cache that 
+has over-approximated the concrete evaluations, and thus the result can be returned.
+Otherwise, we install the latest @out@ cache to the @in@ cache position when initializing 
+the next iteration; in the meanwhile, an empty cache will be used as the @out@ cache. 
+Note that the initial continuation of calling @aeval@ updates the values for the initial 
+configuration in the final cache.
+
+% TODO: briefly explain how to defunc
 
@@
 
-In $k$-CFA algorithm or the abstract abstract machine shown in Section~\ref{aam},
+In $k$-CFA algorithm or using the abstract abstract machine shown in Section~\ref{aam},
 the call sites @(id 2)@ and @(id 1)@ share the same return flow, so the invocation
@@
 increasing the context-sensitivity. Even worse, if we use the monovariant analysis (i.e.,
-$0$-CFA), the analysis result would be such that @x@ and @y@ point to value set @{1, 2}@,
-because the algorithm does not distinguish that @z@ (the argument of @id@) comes from a
-different call site. Under $1$-CFA, the algorithm is able to distinguish that variable
+$0$-CFA), the analysis result would be such that @x@ and @y@ point to the value set @{1, 2}@.
+Because the algorithm does not distinguish incoming values from different call sites 
+for @z@ (the argument of @id@). Under $1$-CFA, the algorithm is able to distinguish that variable
 @z@ has two different values at two call sites, so variable @y@ would not be polluted by
-@1@. However, variable @x@ still points to value set @{1, 2}@, because two call-sites still
-share the same continuation and $1$-CFA is still unable to separate the return-flows. \\
+@1@. However, variable @x@ still points to the value set @{1, 2}@, because the two call-sites 
+still share the same continuation and $1$-CFA does not increase the ability to separate the 
+return-flows. \\
 
@@
 
-In fact, what we started with is an AAM with an unbounded stack which already precisely
+In fact, what we started with is an unbounded-stack AAM which already precisely
 matches the calls and returns. Refunctionalization makes the interpreter a ``big-step''
@@
 in extended continuation-passing style.
-To obtain a definitional abstract interpreter in direct style, 
-we have multiple choices of how to proceed:
+To obtain a definitional abstract interpreter in direct style, we have multiple choices 
+of how to proceed:
 \begin{itemize}
@@
 The first operator @nd@ is adapted from the previous one, but the meta-continuation
-@m@ has been eliminated here. Now the only continuation @k@ applies on tuple type @(T, Cache)@ 
-where the value of type @T@ comes from the collection, and @nd@ performs recursive invocation
-on itself over elements in the collection @ts@.
+@m@ has been eliminated here. Now the only explicit continuation @k@ applies on tuple 
+type @(T, Cache)@ where the value of type @T@ comes from the collection, and @nd@ performs 
+recursive invocation on itself over elements in the collection @ts@.
 
@@
 and an initial accumulated value of type @Ans@.
-The @choices@ iteratively returns an element from the collection along with the 
-latest cache, and the return type is also annotated by \verb|@cps[Ans]|. 
-This annotation denotes the final returned value of @choices@ is type @Ans@.
-The @f@ is the delimited continuation captured by the @shift@ operator, which
-is formed by the statements after @choices@'s call-site.
+The @choices@ iteratively returns an element of type @T@ from the collection 
+along with the latest cache. The return type of @choices@ is annotated with \verb|@cps[Ans]|,
+which denotes that the final returned value of @choices@ is type @Ans@.
+The function @f@ is the delimited continuation captured by the @shift@ operator, which
+is formed by the computation will be executed after the @choices@'s call-site.
 
@@
 annotation \verb|@cps[Ans]| on the return type to enable using delimited control operators 
-in it.
-Then, for every nondeterministic choice in the abstract interpreter, we can simply call 
+inside of it.
+Then, for every nondeterministic choice in the abstract interpretation, we can simply call 
 @choices@ on the set of possible choices and write the program sequentially as concrete 

### 2018-07-01 paper/paper.tex
update the figure

@@
      [start chain, color=gray, transform canvas={yshift=-0.5ex}] \chainin (m-2-2);
-     \chainin (m-2-1) [join={node[below,labeled] {\begin{smallmatrix} \text{splitting} \\ \text{+} \\ \text{List(Success)??} \end{smallmatrix}}}];
+     \chainin (m-2-1) [join={node[below,labeled] {\begin{smallmatrix} \text{splitting} \\ \text{+} \\ 
+     \text{replacing cont.} \\ \text{by list} \end{smallmatrix}}}];
    }
@@
 technique that converts the data types representing meta-continuations (failure continuations) 
-back to list representation, where the elements in the list are plain machine states.
+back to lazy list representation, where the elements in the list are plain machine states.
 

### 2018-07-01 paper/paper.tex
revised section 7

@@
 Intuitively, if we imagine that the classical AAM explores a graph of reachable states, 
-then the linearized AAM flattens the graph to a linear sequence of states.
+then the linearized AAM flattens the graph in depth-first order to a linear sequence of states.
 It has been observed that ``\textit{failure can be replaced by a list of successes}''
@@
 Finally, the @analyze@ function (as the entrance of the analysis) does a looping iteration
-to find the fixed-point of cache starting from an empty cache.
+to find the least fixed-point of cache starting from an empty cache.
 If the @out@ cache of this iteration is equivalent to its @in@ cache, which means no more 
@@
 
-The function @choices@ takes two arguments: a set of elements with type @T@, 
-and an initial accumulated value of type @Ans@.
-The @choices@ iteratively returns an element of type @T@ from the collection 
-along with the latest cache. The return type of @choices@ is annotated with \verb|@cps[Ans]|,
-which denotes that the final returned value of @choices@ is type @Ans@.
+The function @choices@ takes two arguments: a set of elements of type @T@, and an initial cache.
+The return type of @choices@ is @(T, Cache)@ with an annotation \verb|@cps[Ans]|, which 
+denotes that @choices@ will iteratively return an element of type @T@ from the collection
+along with the latest cache, but the final returned value of the continuation is type @Ans@.
 The function @f@ is the delimited continuation captured by the @shift@ operator, which
@@
 interpreters, and the subsequent statements and expressions become the delimited continuation
-@f@. The function @choices@ also returns the latest cache to its left-hand side when invoked.
+@f@. We have two invocations of @choices@, which correspond to the two invocations of @nd@ in
+the refunctionalized form.
+The function @choices@ also returns the latest cache to its left-hand side when invoked.
 The cache will be still accumulated to the next call to @choices@ or @aeval@, 
@@
 
-Now we have arrived at the end of this series of transformations. Starting from a
-small-step abstract abstract machine, eventually we obtain a big-step definitional
-interpreter with pushdown control-flow, written in direct-style.
-
-%specify higher-order definitional interpreters using first-order means (cite)
+Now we have arrived at the end of this series of transformations. 
+The inverse transformation of the last step is basically CPS transforming the function @aeval@ and @nd@.
+We may first explicitly expose the continuation at the level of @aeval@, and then convert @nd@ 
+to extended continuation-passing style, where a meta-continuation argument is added both for
+@nd@ and @nd@'s existing continuation @k@. Besides, these continuations have to be manifestly embedded
+at the call-site of @nd@, instead of implicitly provided by the @shift@ operator.
+
+To conclude, starting from a small-step abstract abstract machine, eventually we obtain a 
+big-step definitional interpreter with pushdown control-flow, written in direct-style.
+
 %\section{Detour: Shiver's \textit{k}-CFA?}

### 2018-07-02 paper/paper.tex
revision

@@
 the end of the computation (i.e., a halt); otherwise, we should return the values of 
-@ae@ to its caller which is contained in the top frame of stack.
-
-However, since we have added nondeterminism into state, an empty list of frames means
-we have reached the halt of \textit{one computation path}, and we should determine the
+@ae@ to its caller which is contained in the top frame of the stack.
+However, since we have added a meta-continuation into the state, an empty list of frames 
+means we have reached the halt of \textit{one computation path}, and we should determine the
 next state indicated by @NDCont@. Thus, we have an inner pattern matching on @ndk@:
@@
         val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-        drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), seen)
+        drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), seen) /* normal return */
     }
@@
       case NDCont(Nil, _, _, _, _)::ndk =>
+        /* transfer to the most recent fork point */
         drive_step(NDState(ae, env, bstore, konts, time, ndk), seen) 
@@
         val new_ndk = NDCont(cls.tail, argvs, bstore, time, frames)::ndk
+        /* resume the fork point with the next closure */
         drive_step(NDState(body, new_env, new_store, frames, time, new_ndk, seen))
@@
           val new_store = store.update(baddr, vals)
-          /* evaluate e with new env and store, then transfer to next body's value */
+          /* evaluate e with new env and store, then transfer to the next body's value */
           aeval(e, new_env, new_store, time, bdnd)
-        }, clsnd) /* transfer to next target closure of f */
+        }, clsnd) /* transfer to the next target closure of f */
       })
@@
 Not unexpectedly, the correspondence also exists for nondeterministic languages. For example,
-\citet{Danvy:2001:DW:773184.773202} showed that refunctionalizing a regular expression matcher yields
-its counterpart with two layer of continuations.
+\citet{Danvy:2001:DW:773184.773202} observed that refunctionalizing a regular expression matcher 
+with two stacks yields its counterpart with two layer of continuations.
 In addition, the operational semantics of control operators (e.g., @shift@ and @reset@) can be given by an 

### 2018-07-02 paper/paper.tex
fix typos

@@
 
-The nondeterministic operator @nd@ is also slightly rewrite to cache-passing style:
-continuation @k@ takes an additional argument of type @Cache@; the current accumulated
-cache @acc.cache@ is exposed to @k@ when applies.
+The nondeterministic operator @nd@ is also slightly rewritten to cache-passing style:
+the continuation @k@ takes an additional argument of type @Cache@; the current accumulated
+cache @acc.cache@ is exposed to @k@ when applied.
 

### 2018-07-02 paper/paper.tex
misc revision

@@
 \begin{itemize}
-  \item If the closure set is an empty set (@Nil@ in the code since we list to represent that set),
+  \item If the closure set is an empty set (@Nil@ in the code since we use list to represent that set),
 then we have tried all the closures of this fork point and should move to the
@@
   \item Otherwise, we pop a closure from the set, then build a new
-environment and a new store for the body expression of the closure;
-we use the same frame list and time that are copied from the fork point rather
+environment and a new store for the body expression of that closure.
+We use the same frame list and time which are copied from the fork point rather
 than current one.
-The meta-continuation @new_ndk@ is also updated by the removal of that closure, which makes
-sure we will eventually shrink to the empty set case.
+The meta-continuation @new_ndk@ is also updated by the removal of that popped closure, which makes
+sure the closure set will eventually shrink to empty.
 \end{itemize}

### 2018-07-02 src/main/scala/refunc/DisentangledLinearSmallStepUB.scala
update code

@@ -18,41 +18,37 @@ object DisLinearSmallStepUBStack {
     NDState(e, env, bstore, List(), List(), List())
 
   def drive_step(nds: NDState, seen: Set[State]): Set[State] = {
+    val s = nds.toState
+    val new_time = nds.tick
+    val new_seen = if (seen.contains(s)) seen else seen + s
     nds match {
-      case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
-      case nds =>
-        val s = nds.toState
-        val new_seen = if (seen.contains(s)) seen else seen + s
-        val new_time = nds.tick
-        nds match {
-          case NDState(Let(x, ae, e), env, bstore, konts, time, ndk) if isAtomic(ae) =>
-            val baddr = allocBind(x, new_time)
-            val new_env = env + (x -> baddr)
-            val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-            drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), new_seen)
+      case NDState(Let(x, ae, e), env, bstore, konts, time, ndk) if isAtomic(ae) =>
+        val baddr = allocBind(x, new_time)
+        val new_env = env + (x -> baddr)
+        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+        drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), new_seen)
 
-          case NDState(Letrec(bds, body), env, bstore, konts, time, ndk) =>
-            val new_env = bds.foldLeft(env)((accenv: Env, bd: B) =>
-              accenv + (bd.x -> allocBind(bd.x, new_time)))
-            val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) =>
-              accbst.update(allocBind(bd.x, new_time), atomicEval(bd.e, new_env, accbst)))
-            drive_step(NDState(body, new_env, new_store, konts, new_time, ndk), new_seen)
-            
-          case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
-            val closures = atomicEval(f, env, bstore).toList.asInstanceOf[List[Clos]]
-            val Clos(Lam(v, body), c_env) = closures.head
-            val frame = Frame(x, e, env)
-            val baddr = allocBind(v, new_time)
-            val new_env = c_env + (v -> baddr)
[…]

### 2018-07-02 src/main/scala/refunc/FusedLinearSmallStepUB.scala

@@ -18,60 +18,57 @@ object FusedLinearSmallStepUBStack {
     NDState(e, env, bstore, List(), List(), List())
 
   def drive_step(nds: NDState, seen: Set[State]): Set[State] = {
+    val s = nds.toState
+    val new_time = nds.tick
+    val new_seen = if (seen.contains(s)) seen else seen + s
     nds match {
-      case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
-      case nds =>
-        val new_time = nds.tick
-        val new_ndstate = nds match {
-          case NDState(Let(x, ae, e), env, bstore, konts, time, ndk) if isAtomic(ae) =>
+      case NDState(Let(x, ae, e), env, bstore, konts, time, ndk) if isAtomic(ae) =>
+        val baddr = allocBind(x, new_time)
+        val new_env = env + (x -> baddr)
+        val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
+        drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), new_seen)
+
+      case NDState(Letrec(bds, body), env, bstore, konts, time, ndk) =>
+        val new_env = bds.foldLeft(env)((accenv: Env, bd: B) =>
+          accenv + (bd.x -> allocBind(bd.x, new_time)))
+        val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) =>
+          accbst.update(allocBind(bd.x, new_time), atomicEval(bd.e, new_env, accbst)))
+        drive_step(NDState(body, new_env, new_store, konts, new_time, ndk), new_seen)
+
+      case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
+        val closures = atomicEval(f, env, bstore).toList.asInstanceOf[List[Clos]]
+        val Clos(Lam(v, body), c_env) = closures.head
+        val frame = Frame(x, e, env)
+        val baddr = allocBind(v, new_time)
+        val new_env = c_env + (v -> baddr)
+        val argvs = atomicEval(ae, env, bstore)
+        val new_store = bstore.update(baddr, argvs)
+        val new_frames = frame::konts
+        val new_ndk = NDCont(closures.tail, argvs, bstore, new_time, new_frames)::ndk
+        drive_step(NDState(body, new_env, new_store, new_frames, new_time, new_ndk), new_seen)
+
[…]

### 2018-07-02 src/main/scala/refunc/LinearSmallStepUB.scala

@@ -17,21 +17,21 @@ object LinearSmallStepUBStack {
   def inject(e: Expr, env: Env = Map(), bstore: Store[BAddr, Storable] = Store[BAddr, Storable](Map())): NDState =
     NDState(e, env, bstore, List(), List(), List())
 
-  def step(nds: NDState): NDState = {
+  def step(nds: NDState): Option[NDState] = {
     val new_time = nds.tick
     nds match {
       case NDState(Let(x, ae, e), env, bstore, konts, time, ndk) if isAtomic(ae) =>
         val baddr = allocBind(x, new_time)
         val new_env = env + (x -> baddr)
         val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-        NDState(e, new_env, new_store, konts, new_time, ndk)
+        Some(NDState(e, new_env, new_store, konts, new_time, ndk))
 
       case NDState(Letrec(bds, body), env, bstore, konts, time, ndk) =>
         val new_env = bds.foldLeft(env)((accenv: Env, bd: B) =>
           accenv + (bd.x -> allocBind(bd.x, new_time)))
         val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) =>
           accbst.update(allocBind(bd.x, new_time), atomicEval(bd.e, new_env, accbst)))
-        NDState(body, new_env, new_store, konts, new_time, ndk)
+        Some(NDState(body, new_env, new_store, konts, new_time, ndk))
 
       case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
         val closures = atomicEval(f, env, bstore).toList.asInstanceOf[List[Clos]]
@@ -43,38 +43,39 @@ object LinearSmallStepUBStack {
         val new_store = bstore.update(baddr, argvs)
         val new_frames = frame::konts
         val new_ndk = NDCont(closures.tail, argvs, bstore, new_time, new_frames)::ndk
-        NDState(body, new_env, new_store, new_frames, new_time, new_ndk)
+        Some(NDState(body, new_env, new_store, new_frames, new_time, new_ndk))
 
       case NDState(ae, env, bstore, konts, time, ndk) if isAtomic(ae) =>
         konts match {
           case Nil => ndk match {
+            case Nil => None
             case NDCont(Nil, _, _, _, _)::ndk => 
-              NDState(ae, env, bstore, konts, time, ndk) //NOTE: konts is Nil
+              Some(NDState(ae, env, bstore, konts, time, ndk)) //NOTE: konts is Nil
[…]

### 2018-07-03 paper/paper.tex
largely refactor section 6

@@
           val frame = Frame(x, e, env)
-          val baddr = allocBind(v, new_time)
-          val new_env = env_c + (v -> baddr)
+          val baddr = allocBind(v, new_time);  val new_env = env_c + (v -> baddr)
           val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
@@
           case Frame(x, e, env_f)::konts =>
-            val baddr = allocBind(x, new_time)
-            val new_env = env_f + (x -> baddr)
+            val baddr = allocBind(x, new_time);  val new_env = env_f + (x -> baddr)
             val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
@@
 and frames at the fork point. In the definition of @NDState@, an auxiliary function @toState@
-is added that converts itself to type @State@.
-
-Function @step@ now becomes of type @NDState => NDState@.
+is added to convert itself to type @State@.
+
+Function @step@ now becomes of type @NDState => Option[NDState]@. Returning a @None@ value means
+reaching the halt.
+
+\begin{lstlisting}
+  def step(nds: NDState): Option[NDState] = {
+    val new_time = nds.tick;  nds match { ... }
+  }
+\end{lstlisting}
+
 The following code shows the first case of matching an instance of @NDState@ in
@@
     val Clos(Lam(v, body), c_env) = closures.head
-    val frame = Frame(x, e, env)
-    val new_frames = frame::konts
-    val baddr = allocBind(v, new_time)
-    val new_env = c_env + (v -> baddr)
-    val argvs = atomicEval(ae, env, bstore)
+    val frame = Frame(x, e, env);        val new_frames = frame::konts
+    val baddr = allocBind(v, new_time);  val new_env = c_env + (v -> baddr)
+    val argvs = atomicEval(ae, env, bstore); 
     val new_store = bstore.update(baddr, argvs)
     val new_ndk = NDCont(closures.tail, argvs, bstore, new_time, new_frames)::ndk
-    NDState(body, new_env, new_store, new_frames, new_time, new_ndk)
+    Some(NDState(body, new_env, new_store, new_frames, new_time, new_ndk))
 \end{lstlisting}
@@
 they may form different binding addresses @baddr@. We also use the new frames, because
-all closures at this fork point share the same stack structure and return point.
+all the closures at this fork point share the same stack structure and return point.
 
@@
       case Nil => ndk match {
+        case Nil => None /* Halt */
         case NDCont(Nil, _, _, _, _)::ndk =>
-          NDState(ae, env, bstore, konts, time, ndk) /* transfer to the most recent fork point */
+          Some(NDState(ae, env, bstore, konts, time, ndk)) /* transfer to the most recent fork point */
         case NDCont(cls, argvs, bstore, time, frames)::ndk =>
@@
           /* resume the fork point with the next closure */
-          NDState(body, new_env, new_store, frames, time, new_ndk)
+          Some(NDState(body, new_env, new_store, frames, time, new_ndk))
       }
       case Frame(x, e, f_env)::konts =>
-        val baddr = allocBind(x, newTime); val new_env = f_env + (x -> baddr)
+        val baddr = allocBind(x, new_time);  val new_env = f_env + (x -> baddr)
         val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-        NDState(e, new_env, new_store, konts, newTime, ndk) /* normal return */
-    }
-\end{lstlisting}
-
-Normally, if the continuation @konts@ is an empty list, then it means we have reached
-the end of the computation (i.e., a halt); otherwise, we should return the values of 
-@ae@ to its caller which is contained in the top frame of the stack.
-However, since we have added a meta-continuation into the state, an empty list of frames 
-means we have reached the halt of \textit{one computation path}, and we should determine the
-next state indicated by @NDCont@. Thus, we have an inner pattern matching on @ndk@:
+        Some(NDState(e, new_env, new_store, konts, new_time, ndk)) /* normal return */
+    }
+\end{lstlisting}
+
+If the continuation @konts@ is empty, then it means we have reached an end of the computation;
+otherwise, we should return the values of @ae@ to its caller which is contained in the top
+frame of the stack. However, since a meta-continuation @ndk@ is introduced, we need to dispatch
+more cases when @konts@ is empty:
 
 \begin{itemize}
-  \item If the closure set is an empty set (@Nil@ in the code since we use list to represent that set),
-then we have tried all the closures of this fork point and should move to the
-next fork point. Since the meta-continuation @nkd@ is represented by a list, and we always
-append new elements to its front, we are resuming to the most recent fork point by
+  \item If @ndk@ is also empty, then we have reached the end of all computations. Besides,
+    the control component @ae@ now is an atomic value, therefore @None@ is returned.
+  \item If the target closure set of the front-most @NDCont@ is empty (@Nil@ in the code),
+    then we have computed all the closures at this fork point and should move to the next fork 
+    point. Since the meta-continuation @nkd@ is represented by a list, and we always
+    append new elements to its front, we resume to the most recent fork point by
     popping up the front-most element and using the tail of that list.
-  \item Otherwise, we pop a closure from the set, then build a new
-environment and a new store for the body expression of that closure.
-We use the same frame list and time which are copied from the fork point rather
-than current one.
-The meta-continuation @new_ndk@ is also updated by the removal of that popped closure, which makes
+  \item Otherwise, we pop a closure from the set, then build a new environment and store for 
+    the body expression of that closure. We use the same frame list and time which are copied 
+    from the fork point rather than current one.
+    The meta-continuation @new_ndk@ is updated by the removal of that popped closure, which makes
     sure the closure set will eventually shrink to empty.
@@
   def drive(nds: NDState, seen: Set[State]): Set[State] = {
-    nds match {
-      case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
-      case nds =>
-        val s = nds.toState
-        if (seen.contains(s)) drive(step(nds), seen)
-        else drive(step(nds), seen + s)
-    }
-  }
-\end{lstlisting}
-
-The @drive@ function is also changed: there is no work-list anymore, as all the
-potentially unexplored states are embedded into the continuation for
-nondeterminism.
-The termination of the analysis occurs when we reach an @NDState@ in which the expression
-is atomic and both the continuation and meta-continuation are empty lists. This corresponds 
-to the case that @todo@ list is empty. The set of explored states is preserved as before.
+    val s = nds.toState;  val new_seen = if (seen.contains(s)) seen else seen + s
+    step(nds) match {
+      case None => new_seen
+      case Some(nnds) => drive(nnds, new_seen)
+    }
+  }
+\end{lstlisting}
+
+The @drive@ function is also changed: there is no work-list anymore, as all the potentially 
+unexplored states are embedded into the meta-continuation.
+The termination of the analysis occurs when the value @None@ is returned from calling @step@.
+Otherwise, we extract the next state @nnds@ and recur on function @drive@.
+The set of explored states is preserved as before.
 
@@
 functions into a single function.
-The fused function @drive_step@ is essentially formed by merging the functionality of
-@step@ into the @drive@ function.
-@drive_step@ takes an @NDState@ and a set of explored states @seen@ as arguments
-and returns a set of reachable states once it terminates.
+The fused function @drive_step@ takes an @NDState@ and a set of explored states @seen@ 
+as arguments and returns a set of reachable states once it terminates.
+It is is essentially formed by merging the functionality of @step@ into the @drive@ function.
+Since each case in function @step@ are statically known, we inline the pattern match on the
+@Option@ type and replace it with @new_seen@ or a recursive call to @drive_step@ for @None@
+or @Some@, respectively.
 
@@
   def drive_step(nds: NDState, seen: Set[State]): Set[State] = {
+    val s = nds.toState;  val new_time = nds.tick
+    val new_seen = if (seen.contains(s)) seen else seen + s
     nds match {
-      case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
-      case nds =>
-        val s = nds.toState; val newTime = nds.tick
-        val new_seen = if (seen.contains(s)) seen else seen+s
-        val new_ndstate = nds match {
       case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndkonts) => ...
+        drive_step(NDState(body, new_env, new_store, new_frames, new_time, new_ndk), new_seen)
       case NDState(ae, env, bstore, konts, time, ndkonts) if isAtomic(ae) => ...
-        }
-        drive_step(new_ndstate, new_seen)
     }
@@
 intermediate states. Given @inject(e)@ as an initial state and an empty set as the
-initial set of states reached, we can easily define the entrance function @analyze@ as
-follows:
+initial set of states reached, we can define the entrance function @analyze@ as follows:
 
@@
 functions.
-After disentanglement, we will obtain an abstract abstract machine that in defunctionalized form.
+After disentanglement, we will obtain an abstract abstract machine in defunctionalized form.
 
@@
 functions:
-@drive_step@, which plays the same role as before; @continue@, which is called from
-@drive_step@ when encountering an atomic expression, and handles the reduction context
-(@Frame@) of the analyzed language; and @ndcontinue@, which dispatches meta-continuations
-@NDCont@, and which is invoked from @continue@ when the normal stack is empty.
+@drive_step@, which plays the same role as before; 
+@continue@, which is invoked from @drive_step@ when encountering an atomic expression, 
+  and dispatches the reduction context of the analyzed language; and 
+@ndcontinue@, which is invoked from @continue@ when the normal stack is empty,
+  and dispatches the meta-continuations @NDCont@.
 
@@
   def drive_step(nds: NDState, seen: Set[State]): Set[State] = {
+    ... 
     nds match {
-      case NDState(ae, _, _, Nil, _, Nil) if isAtomic(ae) => seen
-      case nds =>
-        val s = nds.toState; val new_time = nds.tick
-        val new_seen = if (seen.contains(s)) seen else seen + s
-        nds match {
-          case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) =>
-            ...
+      case NDState(Let(x, App(f, ae), e), env, bstore, konts, time, ndk) => ...
         drive_step(NDState(body, new_env, new_store, new_frames, new_time, new_ndk), new_seen)
-          case NDState(ae, env, bstore, konts, time, ndk) if isAtomic(ae) =>
-            continue(nds, new_seen)
-        }
+      case NDState(ae, env, bstore, konts, time, ndk) if isAtomic(ae) => continue(nds, new_seen)
     }
@@
         val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-        drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), seen) /* normal return */
-    }
-  }
-\end{lstlisting}
-
-Function @continue@ is mainly what we have seen for the second case in @drive_step@,
+        drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), seen)
+    }
+  }
+\end{lstlisting}
+
+Function @continue@ is identical to what we have seen for the second case in @drive_step@,
 except that the dispatching logic for empty @konts@ is lifted to a separate function 
@@
 halt of a computation path, whereas a cons @::@ means the top frame of the list
-holds the binding variable and the innermost context.
-%An analogy to this in standard CEK machines would be an abstract class (e.g., @cont@)
-%with three variants: @halt@, @arg(value, cont)@, and @arg(exp, env, cont)@.
+holds the innermost context.
 
@@
     ndk match {
+      case Nil => seen
       case NDCont(Nil, _, _, _, _)::ndk =>
-        /* transfer to the most recent fork point */
         drive_step(NDState(ae, env, bstore, konts, time, ndk), seen) 
@@
         val new_ndk = NDCont(cls.tail, argvs, bstore, time, frames)::ndk
-        /* resume the fork point with the next closure */
         drive_step(NDState(body, new_env, new_store, frames, time, new_ndk, seen))
@@
 
+Function @ndcontinue@ is identical to the dispatching logic in function @continue@ when @konts@ is empty.
+When @ndk@ is empty, the current explored states @seen@ is returned; the analysis then terminates.
+\iffalse
 The shape for @ndcontinue@ is similar to @continue@; after all, they both use a 
@@
 is transferred to the body expression of the next closure object.
+\fi
 
@@
 proper higher-order functions.
-After this transformation, we will obtain 
+After this transformation, we will obtain an abstract interpreter written in extended
+continuation-passing style\cite{Danvy:1990:AC:91556.91622}.
 This transformation can be also regarded as making the control flows of the defined 
-language be the part of control flows of the meta-language.
-%allocating continuation functions of the defined language in the heap of the meta-language.
-
-In this section, several other notable changes have been made:
+language be part of the control flows of the meta-language.
+
+\subsection{First Try}
+
+First of all, since there is no first-order data types representing model of stacks,
+we introduce a configuration class @Config@ as a state-like definition by eliminating
+@konts@ and @ndk@ from the definition of state. The type of return value of the abstract
+interpreter accordingly becomes @Set[Config]@.
+As we are getting close to achieving abstract definitional interpreters, 
+the name of @drive_step@ is updated to @aeval@ (for \emph{abstract eval}).
+
+\begin{lstlisting}
+  case class Config(e: Expr, env: Env, store: BStore, time: Time) { ... }
+\end{lstlisting}
+
+By looking at the definition of the first-order function @continue@ from the last section,
+we may find it calls the meta-continuation @ndcontinue@ inside of its body. 
+This observation leads us to define the types of higher-order continuation and meta-continuation:
+
+\begin{lstlisting}
+  type Cont = (Config, Set[Config], MCont) => Set[Config]
+  type MCont = (Config, Set[Config]) => Set[Config]
+\end{lstlisting}
+
+The continuation @Cont@ as a higher-order function takes a configuration, a set of explored
+configurations, as well as a meta-continuation as its arguments. While the meta-continuation
+only takes a configuration and a set of explored configurations as arguments.
+Now, the skeleton of function @aeval@ looks as follows:
+
+\begin{lstlisting}
+  def aeval(config: Config, seen: Set[Config], continue: Cont, mcontinue: MCont): Set[Config] = {
+    val Config(e, env, store, time) = config;  val new_time = config.tick
+    val new_seen = if (seen.contains(config)) seen else seen + config
+    e match {
+      case Let(x, App(f, ae), e) =>
+        val closures = atomicEval(f, env, store).toList.asInstanceOf[List[Clos]]
+        val Clos(Lam(v, body), c_env) = closures.head
+        val baddr = allocBind(v, new_time);      val new_env = c_env + (v -> baddr)
+        val argvs = atomicEval(ae, env, store);  val new_store = store.update(baddr, argvs)
+        val new_cont: Cont = ...
+        val new_mcont: MCont = ...
+        aeval(Config(body, new_env, new_store, new_time), new_seen, new_cont, new_mcont)
+      case ae if isAtomic(ae) => continue(config, new_seen, mcontinue)
+    }
+  }
+\end{lstlisting}
+
+Before jumping into the details of refunctionalizing the continuations and constructing
+@new_cont@ and @new_mcont@, we have two important observations in the disentangled AAM
+where the continuations were represented by data types:
+1) When constructing the @new_ndk@, we observe that it explicitly uses @new_frames@ as one
+argument of its top @NDCont@ object. The higer-order version of these continuations follows
+the same pattern.
+2) In the meta-continuation dispatching function @ndcontinue@, the size of the closures set is
+decreasing every time while the other parts of meta-continuation are kept. This means in 
+the higher-order version, we also need an ``ability'' to update the top frame of meta-continuation
+and keep the rest as the same.
+
+Let us first look into @new_cont@. The body of @new_cont@ is similar to the first-order dispatching
+function @continue@. The invocation of @aeval@ uses the latest meta-continuation @m@ instead of the outer
+one, @mcontinue@. Note that the new environment @new_env@ is updated based on the outer environment;
+while the atomic evaluation uses the latest environment @env_@.
+
+\begin{lstlisting}
+  val new_cont: Cont = { case (config@Config(ae, env_, store, time), seen, m) =>
+    val new_time = config.tick
+    val baddr = allocBind(x, new_time);  val new_env = env + (x -> baddr)
+    val new_store = store.update(baddr, atomicEval(ae, env_, store))
+    aeval(Config(e, new_env, new_store, new_time), seen, continue, m)
+  }
+\end{lstlisting}
+
+Then to effectively construct the meta-continuation @new_mcont@, we define an auxiliary function
+@makeMCont@, which takes a list of closures as arguments and returns the meta-continuation. 
+Function @makeMCont@ is a recursive function where the recursive invocation happens when 
+constructing the meta-continuation based on the rest of closures set. 
+Now we can dynamically generate the meta-continuation inside of a meta-continuation.
+Meanwhile, the @new_cont@ we defined above is used here, simply because different closures 
+actually share the same continuation.
+The case where the empty closures set occurs is also interesting. At this case,
+the continuation simply applies its meta-continuation @m@, and the meta-continuation is the
+outer meta-continuation @mcontinue@. It would be clear if we inductively think the base case 
+where the set contains only one closure, then @makeMCont@ is invoked with an empty set, 
+therefore the continuation should do nothing. %TODO: rewrite this part.
+
+\begin{lstlisting}
+  def makeMCont(cls: List[Clos]): MCont = {
+    val mcont: MCont = { case (config, seen) =>
+      if (cls.isEmpty) aeval(config, seen, (c, seen, m) => m(c, seen), mcontinue)
+      else {
+        val Clos(Lam(v, body), c_env) = cls.head
+        val baddr = allocBind(v, new_time);  val new_env = c_env + (v -> baddr)
+        val new_store = store.update(baddr, argvs)
+        aeval(Config(body, new_env, new_store, new_time), seen, new_cont, makeMCont(cls.tail))
+      }
+    } mcont }
+  val new_mcont: MCont = makeMCont(closures.tail)
+\end{lstlisting}
+
+One last very important thing is about the initial continuation and meta-continuation.
+The initial continuation is invoked when one computation path reaches its end,
+so we invoke the meta-continuation @m@ to continue exploring other paths.
+The initial meta-continuation is invoked when all computation paths reach their ends,
+so we terminate with the explored configurations @seen@. They correspond to the 
+cases where @konts@ is empty and @konts@ and @ndk@ are both empty, respectively.
+
+\begin{lstlisting}
+  def analyze(e: Expr): Set[Config] =
+    aeval(inject(e), Set(), { (c, seen, m) => m(c, seen) }, { (c, seen) => seen })
+\end{lstlisting}
+
+Now we have completely refunctionalized an abstract abstract machine; the two layers
+of continuations are represented by higher-order functions. The inverse transformation 
+of this step is known as \textit{de}functionalization, that converts the higher-order
+representing continuations back to first-order data types representation 
+\cite{Danvy:2001:DW:773184.773202, Danvy:2008:DIP:1411204.1411206}.
+
+\subsection{Simplification and Caching Fixed-Points}
+
+In this section, based on the refunctionalized AAM, we present several other transformations 
+toward a big-step abstract definitional interpreter:
+
 \begin{itemize}
-\item We do not have the concept of @state@ (or @NDState@) anymore.
-  The components of the states are lifted as arguments of the evaluation function.
+  \item The components of the configurations are lifted as arguments of the evaluation
+    function.
   \item Since our final target is an abstract definitional interpreter (though still
     nondeterministic), starting from this section our abstract semantic artifact
-  returns a set of final values instead of collected states.
-\item As we are getting close to achieving abstract definitional interpreters, 
-  the name of @drive_step@ is updated to @aeval@ (for \emph{abstract eval}).
+    returns a set of final values instead of collected configurations.
+  \item Configurations still play an important role in caching, where a cache
+    is a mapping from configurations to sets of final values. The caching algorithm
+    ensures termination of the analysis, as well as performs a sound over-approximation
+    of all possible concrete reachable paths and values.
+  \item We simplify the continuations with an nondeterministic operator @nd@ such 
+    that the function @aeval@ only uses one layer of continuation.
 \end{itemize}
 
-For clarity, we first show how to refunctionalize continuations and meta-continuation 
-without collecting semantics.
-Then, adding an appropriate caching algorithm to ensure the analysis will
-terminate at some fixed-point is a simple transformation of the program to cache-passing
-style. Finally, we examine how computable pushdown control-flow analysis is established
-through refunctionalization and caching.
-
-\subsection{First Try}
-
+\textit{Values.}
 To properly represent final values, we introduce the case class @VS@ which
@@
 
-An instance of @VS@ represents the computational result of one path in
-the nondeterministic evaluation. 
-The reason that we include a store is that there might be a case in which two paths
-have the same values but different accumulated side effects (e.g., memory allocations).
+An instance of @VS@ represents the computational result of one path in the nondeterministic 
+evaluation. The reason that we include a store is that there might be a case in which two 
+paths have the same values but different accumulated side effects (e.g., memory allocations).
 Since the whole computation is potentially nondeterministic, the type of the final result
-values is a collection @Set[VS]@. For convenience, we define it as a type @Ans@:
-
-\begin{lstlisting}
-  type Ans = Set[VS]
-\end{lstlisting}
-
-As previously mentioned, the components of states are lifted as arguments to
-@aeval@. Therefore, the pattern matching on the state becomes just a match on
-expression @e@. The shape of function @aeval@ looks like:
-
-\begin{lstlisting}
-  type Cont = Ans => Ans
-  def aeval(e: Expr, env: Env, store: BStore, time: Time, continue: Cont): Ans = {
-    val new_time = ...
-    e match {
-      case Let(x, App(f, ae), e) => ...
-      case ae if isAtomic(ae) => ...
-    }
-  }
-\end{lstlisting}
-
-An additional argument @continue@ which has type @Cont@ is also introduced.
-Recall that in the disentangled AAM, the second case in @drive_step@
-makes a call to @continue@. Here, the reemerging @continue@ is a function
-of type @Ans => Ans@. The function @aeval@ does not return: instead, it calls
-@continue@ with the result values. \\
-
-\textit{Nondeterminism Abstractions.}
-To handle the inevitable nondeterminism, we define a generic function @nd@ as follows:
-
-\begin{lstlisting}
-  def nd[T](ts: Iterable[T], acc: Ans, k: (T, Ans => Ans) => Ans, m: Ans => Ans): Ans = {
-    if (ts.isEmpty) m(acc)
-    else k(ts.head, (ans: Ans) => nd(ts.tail, acc ++ ans, k, m))
-  }
-\end{lstlisting}
-
-Given a collection of elements of type @T@, an initial accumulator of type 
-@Ans@, a function @k@ that works on element's type @T@ and takes a continuation 
-of type @Ans=>Ans@, and a function @m@ that works on and returns a value of type 
-@Ans@ (which is also type @Ans=>Ans@), @nd@ chooses an element in the collection 
-and applies @k@ to it, accumulates the value (@acc++ans@ where the @ans@ is the 
-result comes from @k@) and eventually applies function @m@ on the accumulated 
-value when there is no element in the collection.
-
-%TODO: consider rewrite this paragraph and explain the relation to `continue`.
-If we look at @continue@ and @ndcontinue@ in the disentangled AAM, we will find
-that @nd@ is an abstraction of those two functions. In function @continue@,
-there are two out calls to @drive_step@ and @ndcontinue@, here the application
-of @k@ simulates @drive_step@, and @m@ simulates @ndcontinue@.
-
-Readers familiar with functional programming may think of @fold@ at this point; indeed,
-it is a variant implementation of @fold@ in continuation-passing style. But as we mentioned 
-in Section \ref{linear}, here we have two continuations, one continuation @k@ for the next 
-element in the collection, and one meta-continuation @m@ for the eventual accumulator.
-This is sometimes referred to as an extended continuation-passing style (ECPS) \cite{Danvy:1990:AC:91556.91622}. \\
-
-\textit{Evaluation without Caching.}
-Since our first version of @aeval@ implements only the nondeterministic evaluation, 
-given the defined @nd@ operator, the evaluation function @aeval@ can be easily defined. 
-In the case that @e@ is a @let@ expression, we invoke @nd@ twice and perform a nesting 
-two-step, depth-first evaluation over possible values of @App(f, ae)@.
-The values of branches from a fork point will be accumulated inside of @nd@, and 
-then returned to the upper fork point.
-
-\begin{lstlisting}
-  case Let(x, App(f, ae), e) =>
-    val closures = atomicEval(f, env, store)
-    nd[Storable](closures, Set[VS](), { case (clos, clsnd) =>
-      val Clos(Lam(v, body), c_env) = clos
-      val baddr = allocBind(v, new_time); val new_cenv = c_env + (v -> baddr)
-      val new_store = store.update(baddr, atomicEval(ae, env, store))
-      aeval(body, new_cenv, new_store, new_time, (bdvss: Set[VS]) => {
-        nd[VS](bdvss, Set[VS](), { case (vs, bdnd) =>
-          val VS(vals, time, store) = vs
-          val baddr = allocBind(x, time); val new_env = env + (x -> baddr)
-          val new_store = store.update(baddr, vals)
-          /* evaluate e with new env and store, then transfer to the next body's value */
-          aeval(e, new_env, new_store, time, bdnd)
-        }, clsnd) /* transfer to the next target closure of f */
-      })
-    },
-    continue)
-\end{lstlisting}
-
-The outer application of @nd@ evaluates over the set of possible target closures of
-@f@. For each such closure, we go into its body with a new environment and a new store.
-Remember that the function @aeval@ returns a set of @VS@ objects to its continuation,
-so the continuation argument @bdvss@ is a set of @VS@ objects which represent values
-from multiple computation paths when evaluating this single @body@ expression.
-
-We next consider the inner application of @nd@, which evaluates over the set of body's 
-value/stores @bdvss@. For each @VS@, the timestamp and store in @VS@ will be instantiated to
-the next application of @aeval@. The inner application of @aeval@ evaluates @e@ with the
-substituted environment and store, then returns a set of values of one computation path 
-to its continuation @bdnd@, where the continuation will accumulate the result and 
-transfer evaluation to the next body value.
-
-The meta-continuation of the inner @nd@ is @clsnd@, meaning that we need to transfer 
-to the next target closure of @f@. The meta-continuation of the outer @nd@ is just 
-@continue@, as what we normally would do when writing CPS interpreters.
-
-\begin{lstlisting}
-  case ae if isAtomic(ae) => 
-    val ds = atomicEval(ae, env, store); continue(Set(VS(ds, new_time, store)))
-\end{lstlisting}
-
-If the argument expression @e@ is atomic, which is the second case of match, we simply 
-construct a new instance of @VS@ and return it to the @continue@ function.
-
-At the end, the initial invocation to @aeval@ is passed with initial environments,
-especially an identity function as the halting continuation:
-\begin{lstlisting}
-  def analyze(e: Expr) = aeval(e, mtEnv, mtStore, mtTime, (ans => ans))
-\end{lstlisting}
-
-\subsection{Caching Fixed-Points}
-
-Based on the refunctionalized AAM from the last section, we use the same nondeterministic
-operator @nd@, but further extend it to a fixed-point cached version in this section.
-The fixed-point caching algorithm performs a sound over-approximation of all possible
-concrete reachable paths and values, and also prevents non-termination when analyzing 
-programs that may diverge. \\
-
-\textit{Configuration.}
-We do not have the concept of an explicit stack or state, so to represent the cache 
-we introduce a configuration @Config@ as a state-like definition which borrows the 
-components from state, sans continuations.
-
-\begin{lstlisting}
-  case class Config(e: Expr, env: Env, store: BStore, time: Time) { ... }
-\end{lstlisting}
-
-\textit{Cache.}
+values is a collection @Set[VS]@. \\
+
+\textit{Caches.}
+% why cache? problem of set of states.
 Recall that the latent assumption of the work-list algorithm in classical AAM is that 
@@
 \citeauthor{darais2017abstracting}'s ADI paper \cite{darais2017abstracting}.
-
 The case class @Cache@ and its operations are defined as follow:
@@
 \textit{Cache-Passing Style.}
-We can now transform our interpreter into cache-passing style. We begin by changing the 
-return type @Ans@ to a case class which contains a @VS@ set and a cache. @Ans@ also implements 
+We can now transform our interpreter into cache-passing style. We begin by defining the 
+return type @Ans@ as a case class which contains a @VS@ set and a cache. @Ans@ also implements 
 its own accumulating operation @++@ which defines as union on @Set[VS]@ and join on caches.
@@
   case class Ans(vss: Set[VS], cache: Cache) {
-    def ++(ans: Ans): Ans = { Ans(vss ++ ans.vss, cache.join(ans.cache)) }
-  }
-\end{lstlisting}
-
-The nondeterministic operator @nd@ is also slightly rewritten to cache-passing style:
-the continuation @k@ takes an additional argument of type @Cache@; the current accumulated
-cache @acc.cache@ is exposed to @k@ when applied.
-
-\begin{lstlisting}
-  def nd[T](ts: Iterable[T], acc: Ans, k: (T, Cache, Ans => Ans) => Ans, m: Ans => Ans): Ans = {
-    if (ts.isEmpty) m(acc)
-    else k(ts.head, acc.cache, (ans: Ans) => nd(ts.tail, acc ++ ans, k, m))
-  }
-\end{lstlisting}
-
-The skeleton of function @aeval@ largely remains the same.
-
-\begin{lstlisting}
+    def ++(ans: Ans): Ans = { Ans(vss ++ ans.vss, cache.join(ans.cache)) } }
+\end{lstlisting}
+
+As previously mentioned, the components of configurations are lifted as arguments to
+@aeval@. Given the cache and simplified one-layer continuation (as we will see later), 
+the shape of function @aeval@ now looks like:
+
+\begin{lstlisting}
+  type Cont = Ans => Ans
   def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, continue: Cont): Ans = {
-    val config = Config(e, env, store, time)
+    val config = Config(e, env, store, time);  val new_time = config.tick
     if (cache.outContains(config)) continue(Ans(cache.outGet(config), cache))
-    else {
-      val new_cache = cache.outUpdate(config, cache.inGet(config))
-      val new_time = config.tick 
-      e match { ...  }
-    }
-  }
-\end{lstlisting}
-
-Upon entering @aeval@, we first determine whether the @out@ cache
-contains some values for the current configuration and use them immediately (if present).
+    else { val new_cache = cache.outUpdate(config, cache.inGet(config))
+           e match { ... } } }
+\end{lstlisting}
+
+Upon entering @aeval@, we first determine whether the @out@ cache contains some values 
+for the current configuration and use them immediately (if present).
 This corresponds to small-step AAM with a work-list: when we have seen a state,
@@
 In reality, we first conservatively assume that all computations can be diverged, and if 
-not, we update its values in the mapping after evaluation. In terms of partial orders and 
-lattices, it is the case that starts the Kleene iteration from the bottom of the lattice.
-
-The two cases of match are shown below. Note that the cache is used in a monotonic way: 
-each function call to @aeval@ or @nd@ is passed with the latest cache from the most recent 
-continuation of @aeval@ or @nd@.
+not, we update its values in the cache after evaluation. In terms of partial orders and 
+lattices, it is the case that starts the Kleene iteration from the bottom of the lattice.\\
+
+\textit{Nondeterminism Abstraction.}
+We observe that in the refunctionalized AAM, @makeMCont@ actually repeats the prologue
+part when matching a @let@ expression: they are both preparing a new environment and store 
+for the body expression of a closure. %TODO motivation
+We present a nondeterministic operator that abstracts this pattern:
+
+\begin{lstlisting}
+  def nd[T](ts: Iterable[T], acc: Ans, f: (T, Cache, Ans => Ans) => Ans, m: Ans => Ans): Ans = {
+    if (ts.isEmpty) m(acc)
+    else f(ts.head, acc.cache, (ans: Ans) => nd(ts.tail, acc ++ ans, f, m))
+  }
+\end{lstlisting}
+
+Given a collection of elements of type @T@, an initial accumulator of type 
+@Ans@, a function @f@ that works on element's type @T@ and takes a continuation 
+of type @Ans=>Ans@, and a function @m@ that is also type @Ans=>Ans@, @nd@ chooses
+an element in the collection and applies @f@ to it, accumulates the value 
+(@acc++ans@ where the @ans@ is the result comes from @f@) and eventually applies 
+function @m@ on the accumulated value when there is no element in the collection.
+This function is also written in cache-passing style, and the current accumulated 
+cache @acc.cache@ is exposed to @f@ when applied.
+
+Our function @nd@ is a variant implementation of @fold@ in continuation-passing style,
+and @f@ also has to be passed as a function that takes a continuation. \\
+
+\textit{Put Them Together.}
+Given the skeleton of @aeval@ in cache-and-continuation-passing style, we may now
+use the @nd@ operator to fill the details. The first case of match is that @e@
+is a @let@ expression:
 
@@
     val closures = atomicEval(f, env, store)
-    nd[Storable](closures, Ans(Set[VS](), new_cache), {
-      case (Clos(Lam(v, body), c_env), clscache, clsnd) =>
+    nd[Storable](closures, Ans(Set[VS](), new_cache), { case (cls, clscache, clsnd) =>
+        val Clos(Lam(v, body), c_env) = cls
         val vbaddr = allocBind(v, new_time);  val new_cenv = c_env + (v -> vbaddr)
@@
         aeval(body, new_cenv, new_store, new_time, clscache, { case Ans(bdvss, bdcache) =>
-          nd[VS](bdvss, Ans(Set[VS](), bdcache), {
-            case (VS(vals, time, vssstore), bdvsscache, bdnd) =>
+          nd[VS](bdvss, Ans(Set[VS](), bdcache), { case (VS(vals, time, vssstore), bdvsscache, bdnd) =>
               val baddr = allocBind(x, time);  val new_env = env + (x -> baddr)
@@
               aeval(e, new_env, new_store, time, bdvsscache, {
+                /* evaluate e with new env and store, then transfer to the next body's value */
                 case Ans(evss, ecache) => bdnd(Ans(evss, ecache.outUpdate(config, evss)))
               })
-          }, clsnd)
+          }, clsnd) /* transfer to the next target closure of f */
         })
-    },
-    continue)
+    }, continue)
+\end{lstlisting}
+
+We invoke @nd@ twice and perform a nesting two-step, depth-first evaluation over 
+possible values of @App(f, ae)@. The values of branches from a fork point will be
+accumulated inside of @nd@, and then returned to the upper fork point.
+
+The outer application of @nd@ evaluates over the set of possible target closures of
+@f@. For each such closure, we go into its body with a new environment and a new store.
+Remember that the function @aeval@ returns a set of @VS@ objects to its continuation,
+so the continuation argument @bdvss@ is a set of @VS@ objects which represent values
+from multiple computation paths when evaluating this single @body@ expression.
+
+We next consider the inner application of @nd@, which evaluates over the set of body's 
+value/stores @bdvss@. For each @VS@, the timestamp and store in @VS@ will be instantiated to
+the next application of @aeval@. The inner application of @aeval@ evaluates @e@ with the
+substituted environment and store, then returns a set of values of one computation path 
+to its continuation @bdnd@, where the continuation will accumulate the result and 
+transfer the evaluation to the next body value.
+The meta-continuation of the inner @nd@ is @clsnd@, meaning that we need to transfer 
+to the next target closure of @f@. The meta-continuation of the outer @nd@ is just 
+@continue@, as what we normally would do when writing CPS interpreters.
+
+Note that the cache is used in a monotonic way: each function call to @aeval@ or @nd@ 
+is passed with the latest cache from the most recent continuation of @aeval@ or @nd@.
+
+\begin{lstlisting}
   case ae if isAtomic(ae) =>
@@
 
-After completing the evaluation, we obtain a set of @VS@ objects and must update the @out@ cache.
+If the argument expression @e@ is atomic, which is the second case of match, we simply 
+construct a new instance of @VS@ and pass it to the @continue@ function.
+
+Once the evaluation is completed at each case, we obtain a set of @VS@ objects and must update the @out@ cache.
 In the first case of pattern matching, this happens at the innermost continuation of @aeval@,
@@
 configuration in the final cache.
-
-% TODO: briefly explain how to defunc
 
@@
 We may first explicitly expose the continuation at the level of @aeval@, and then convert @nd@ 
-to extended continuation-passing style, where a meta-continuation argument is added both for
-@nd@ and @nd@'s existing continuation @k@. Besides, these continuations have to be manifestly embedded
+to continuation-passing style. Besides, these continuations have to be manifestly embedded
 at the call-site of @nd@, instead of implicitly provided by the @shift@ operator.

### 2018-07-03 src/main/scala/refunc/LinearSmallStepUB.scala

@@ -48,7 +48,7 @@ object LinearSmallStepUBStack {
       case NDState(ae, env, bstore, konts, time, ndk) if isAtomic(ae) =>
         konts match {
           case Nil => ndk match {
-            case Nil => None
+            case Nil => None // Halt
             case NDCont(Nil, _, _, _, _)::ndk => 
               Some(NDState(ae, env, bstore, konts, time, ndk)) //NOTE: konts is Nil
             case NDCont(cls, argvs, bstore, time, frames)::ndk => 
@@ -71,11 +71,10 @@ object LinearSmallStepUBStack {
 
   def drive(nds: NDState, seen: Set[State]): Set[State] = {
     val s = nds.toState
+    val new_seen = if (seen.contains(s)) seen else seen + s
     step(nds) match {
-      case None => seen
-      case Some(nnds) => 
-        if (seen.contains(s)) drive(nnds, seen) 
-        else drive(nnds, seen + s)
+      case None => new_seen
+      case Some(nnds) => drive(nnds, new_seen)
     }
   }
 

### 2018-07-03 src/main/scala/refunc/RefuncECPS.scala

@@ -1,0 +1,75 @@
+package refunc
+
+import scala.util.continuations._
+import refunc.ast._
+import ANFAAM._
+
+object RefuncECPS {
+  import SmallStepUBStack._
+
+  def inject(e: Expr, env: Env = Map(), 
+             store: Store[BAddr, Storable] = Store[BAddr, Storable](Map())): Config =
+    Config(e, env, store, List())
+
+  type Cont = (Config, Set[Config], MCont) => Set[Config]
+  type MCont = (Config, Set[Config]) => Set[Config]
+
+  def aeval(config: Config, seen: Set[Config], continue: Cont, mcontinue: MCont): Set[Config] = {
+    val Config(e, env, store, time) = config
+    val new_time = config.tick
+    val new_seen = if (seen.contains(config)) seen else seen + config
+    e match {
+      case Let(x, ae, e) if isAtomic(ae) =>
+        val baddr = allocBind(x, new_time)
+        val new_env = env + (x -> baddr)
+        val new_store = store.update(baddr, atomicEval(ae, env, store))
+        aeval(Config(e, new_env, new_store, new_time), new_seen, continue, mcontinue)
+
+      case Letrec(bds, body) => 
+        val new_env = bds.foldLeft(env)((accenv: Env, bd: B) =>
+          accenv + (bd.x -> allocBind(bd.x, new_time)))
+        val new_store = bds.foldLeft(store)((accbst: BStore, bd: B) => 
+          accbst.update(allocBind(bd.x, new_time), atomicEval(bd.e, new_env, accbst)))
+        aeval(Config(body, new_env, new_store, new_time), new_seen, continue, mcontinue)
+
+      case Let(x, App(f, ae), e) =>
+        val closures = atomicEval(f, env, store).toList.asInstanceOf[List[Clos]]
+        val Clos(Lam(v, body), c_env) = closures.head
+        val baddr = allocBind(v, new_time)
+        val new_env = c_env + (v -> baddr)
+        val argvs = atomicEval(ae, env, store)
+        val new_store = store.update(baddr, argvs)
+        val new_cont: Cont = { case (config@Config(ae, env_, store, time), seen, m) => 
+          val new_time = config.tick
+          val baddr = allocBind(x, new_time)
+          val new_env = env + (x -> baddr)
+          val new_store = store.update(baddr, atomicEval(ae, env_, store))
[…]

### 2018-07-03 src/test/scala/refunc/Test.scala

@@ -23,8 +23,9 @@ trait RefuncTest extends FunSuite {
   val bot = Set()
 
   def summarize(ss: Set[BStore]): BStore = { ss.foldLeft(mtStore)(_.join(_)) }
-  def summarizeState(ss: Set[ANFAAM.State]): BStore = { summarize(ss.map(_.bstore)) }
   def summarizeVSS(vss: Set[VS]): BStore = { summarize(vss.map(_.store)) }
+  def summarizeState(ss: Set[ANFAAM.State]): BStore = { summarize(ss.map(_.bstore)) }
+  def summarizeConfig(cs: Set[Config]): BStore = { summarize(cs.map(_.store)) }
   def summarizeUBState(ss: Set[SmallStepUBStack.State]): BStore = {
     summarize(ss.map(_.bstore)) 
   }
@@ -68,6 +69,9 @@ trait RefuncTest extends FunSuite {
          */
         assert(summarizeUBState(DisLinearSmallStepUBStack.analyze(prog, initenv, initstore)) ==
                summarizeVSS(RefuncCPS.analyze(prog, initenv, initstore).vss))
+
+        assert(summarizeUBState(DisLinearSmallStepUBStack.analyze(prog, initenv, initstore)) ==
+               summarizeConfig(RefuncECPS.analyze(prog, initenv, initstore)))
       
         /* The result analyzed by the refunctionalized abstract interpreter written in CPS
          * (RefuncCPS) should have the same result analyzed by the direct-style abstract 

### 2018-07-03 paper/paper.tex
small tweaks; work in progress

@@
 \begin{lstlisting}
-  val new_cont: Cont = { case (config@Config(ae, env_, store, time), seen, m) =>
+  val new_cont: Cont = { case (config @ Config(ae, env_, store, time), seen, m) =>
     val new_time = config.tick
@@
     else { val new_cache = cache.outUpdate(config, cache.inGet(config))
-           e match { ... } } }
+           e match { ... } }
+  }
 \end{lstlisting}

### 2018-07-03 paper/paper.tex
tweaks & comments

@@
 As previously mentioned, the components of configurations are lifted as arguments to
-@aeval@. Given the cache and simplified one-layer continuation (as we will see later), 
+@aeval@. Given the cache and the simplified one-layer continuation (as we will see later), 
 the shape of function @aeval@ now looks like:
@@
 
+%TODO: co-inductive
 Upon entering @aeval@, we first determine whether the @out@ cache contains some values 
@@
               val new_store = vssstore.update(baddr, vals)
-              aeval(e, new_env, new_store, time, bdvsscache, {
+              aeval(e, new_env, new_store, time, bdvsscache, { case Ans(evss, ecache) =>
                 /* evaluate e with new env and store, then transfer to the next body's value */
-                case Ans(evss, ecache) => bdnd(Ans(evss, ecache.outUpdate(config, evss)))
+                bdnd(Ans(evss, ecache.outUpdate(config, evss)))
               })
@@
       else iter(Cache(new_cache.out, Store[Config, VS](Map())))
-    }
-    iter(mtCache)
-  }
+    } iter(mtCache) }
 \end{lstlisting}
@@
 configuration in the final cache.
+% TODO: summarize this section
 

### 2018-07-03 paper/paper.tex
continue revising section 6

@@
 
-First of all, since there is no first-order data types representing model of stacks,
+First of all, as we are getting close to achieving abstract definitional interpreters, 
+the name of @drive_step@ is updated to @aeval@ (for \emph{abstract eval}).
+Since there will be no first-order data types representing model of stacks,
 we introduce a configuration class @Config@ as a state-like definition by eliminating
@@
 interpreter accordingly becomes @Set[Config]@.
-As we are getting close to achieving abstract definitional interpreters, 
-the name of @drive_step@ is updated to @aeval@ (for \emph{abstract eval}).
 
@@
 configurations, as well as a meta-continuation as its arguments. While the meta-continuation
-only takes a configuration and a set of explored configurations as arguments.
+@MCont@ only takes a configuration and a set of explored configurations as arguments.
 Now, the skeleton of function @aeval@ looks as follows:
@@
 
-Before jumping into the details of refunctionalizing the continuations and constructing
+The continuation and meta-continuation are added as arguments to @aeval@.
+Before jumping into the details of refunctionalizing the stack structure and constructing
 @new_cont@ and @new_mcont@, we have two important observations in the disentangled AAM
@@
 The case where the empty closures set occurs is also interesting. At this case,
-the continuation simply applies its meta-continuation @m@, and the meta-continuation is the
-outer meta-continuation @mcontinue@. It would be clear if we inductively think the base case 
+the continuation simply applies its meta-continuation @m@, and the meta-continuation of @aeval@
+is the outer meta-continuation @mcontinue@. It would be clear if we inductively think the base case 
 where the set contains only one closure, then @makeMCont@ is invoked with an empty set, 
@@
 One last very important thing is about the initial continuation and meta-continuation.
-The initial continuation is invoked when one computation path reaches its end,
+The initial continuation will be invoked when one computation path reaches its end,
 so we invoke the meta-continuation @m@ to continue exploring other paths.
-The initial meta-continuation is invoked when all computation paths reach their ends,
+The initial meta-continuation will be invoked when all computation paths reach their ends,
 so we terminate with the explored configurations @seen@. They correspond to the 
@@
 
-In this section, based on the refunctionalized AAM, we present several other transformations 
-toward a big-step abstract definitional interpreter:
+In this section, based on the refunctionalized AAM, we further present several other 
+transformations toward a big-step abstract definitional interpreter:
 
@@
 
-\textit{Values.}
+\subsubsection{Values}
 To properly represent final values, we introduce the case class @VS@ which
@@
 Since the whole computation is potentially nondeterministic, the type of the final result
-values is a collection @Set[VS]@. \\
-
-\textit{Caches.}
+values is a collection @Set[VS]@.
+
+\subsubsection{Caches}
 % why cache? problem of set of states.
@@
 Once we have evaluated a term to some values, we update the @out@ cache. In the next 
-iteration, we will use the @out@ from the previous iteration as the @in@ cache for the 
-current iteration. \\
-
-\textit{Cache-Passing Style.}
+iteration, we will use the @out@ cache from the previous iteration as the @in@ cache.
+
+\subsubsection{Cache-Passing Style}
+%motivation
 We can now transform our interpreter into cache-passing style. We begin by defining the 
 return type @Ans@ as a case class which contains a @VS@ set and a cache. @Ans@ also implements 
-its own accumulating operation @++@ which defines as union on @Set[VS]@ and join on caches.
+its own accumulating operation @++@ which defines as a union of @Set[VS]@s and a join of caches.
 The join operation of caches is defined as joining its two store components.
@@
 Upon entering @aeval@, we first determine whether the @out@ cache contains some values 
-for the current configuration and use them immediately (if present).
+for the current configuration and use them immediately (if present) by calling @continue@.
 This corresponds to small-step AAM with a work-list: when we have seen a state,
-we can simply discard the state and continue working through the rest of the work-list @todo@.
+we can simply discard the state and continue working through the rest of the work-list.
 
@@
 cache may return an empty set of values if there is no such mapping for the configuration 
-from the previous iteration.
+from the last iteration.
 In reality, we first conservatively assume that all computations can be diverged, and if 
-not, we update its values in the cache after evaluation. In terms of partial orders and 
-lattices, it is the case that starts the Kleene iteration from the bottom of the lattice.\\
-
-\textit{Nondeterminism Abstraction.}
+not, we update its values in the cache after the evaluation. In terms of partial orders and 
+lattices, it is the case that starts the Kleene iteration from the bottom of the lattice.
+
+\subsubsection{Nondeterminism Abstraction}
 We observe that in the refunctionalized AAM, @makeMCont@ actually repeats the prologue
@@
 Our function @nd@ is a variant implementation of @fold@ in continuation-passing style,
-and @f@ also has to be passed as a function that takes a continuation. \\
-
-\textit{Put Them Together.}
+and @f@ also has to be passed as a function that takes a continuation.
+
+\subsubsection{Put Them Together}
 Given the skeleton of @aeval@ in cache-and-continuation-passing style, we may now
@@
 We next consider the inner application of @nd@, which evaluates over the set of body's 
-value/stores @bdvss@. For each @VS@, the timestamp and store in @VS@ will be instantiated to
+value/stores @bdvss@. For each @VS@, the time-stamp and store in @VS@ will be instantiated to
 the next application of @aeval@. The inner application of @aeval@ evaluates @e@ with the
@@
 configuration in the final cache.
+
 % TODO: summarize this section

### 2018-07-03 paper/paper.tex
simplify code

@@
     val new_store = store.update(baddr, atomicEval(ae, env_, store))
-    aeval(Config(e, new_env, new_store, new_time), seen, continue, m)
-  }
+    aeval(Config(e, new_env, new_store, new_time), seen, continue, m) }
 \end{lstlisting}
@@
 The case where the empty closures set occurs is also interesting. At this case,
-the continuation simply applies its meta-continuation @m@, and the meta-continuation of @aeval@
-is the outer meta-continuation @mcontinue@. It would be clear if we inductively think the base case 
-where the set contains only one closure, then @makeMCont@ is invoked with an empty set, 
-therefore the continuation should do nothing. %TODO: rewrite this part.
-
-\begin{lstlisting}
-  def makeMCont(cls: List[Clos]): MCont = {
-    val mcont: MCont = { case (config, seen) =>
-      if (cls.isEmpty) aeval(config, seen, (c, seen, m) => m(c, seen), mcontinue)
-      else {
+the outer meta-continuation @mcontinue@ is returned.
+It would be clear if we inductively think the base case where the closures set contains only one 
+closure, then @makeMCont@ is invoked with an empty set, therefore should be just the 
+current meta-continuation @mcontinue@.
+
+\begin{lstlisting}
+  def makeMCont(cls: List[Clos]): MCont =
+    if (cls.isEmpty) mcontinue else { (config, seen) =>
       val Clos(Lam(v, body), c_env) = cls.head
@@
       val new_store = store.update(baddr, argvs)
-        aeval(Config(body, new_env, new_store, new_time), seen, new_cont, makeMCont(cls.tail))
-      }
-    } mcont }
+      aeval(Config(body, new_env, new_store, new_time), seen, new_cont, makeMCont(cls.tail)) }
   val new_mcont: MCont = makeMCont(closures.tail)
@@
   def analyze(e: Expr): Set[Config] =
-    aeval(inject(e), Set(), { (c, seen, m) => m(c, seen) }, { (c, seen) => seen })
+    aeval(inject(e), Set(), (c, seen, m) => m(c, seen), (c, seen) => seen)
 \end{lstlisting}
@@
   case class Ans(vss: Set[VS], cache: Cache) {
-    def ++(ans: Ans): Ans = { Ans(vss ++ ans.vss, cache.join(ans.cache)) } }
+    def ++(ans: Ans): Ans = Ans(vss ++ ans.vss, cache.join(ans.cache)) }
 \end{lstlisting}
@@
 \begin{lstlisting}
-  def nd[T](ts: Iterable[T], acc: Ans, f: (T, Cache, Ans => Ans) => Ans, m: Ans => Ans): Ans = {
+  def nd[T](ts: Iterable[T], acc: Ans, f: (T, Cache, Ans => Ans) => Ans, m: Ans => Ans): Ans =
     if (ts.isEmpty) m(acc)
     else f(ts.head, acc.cache, (ans: Ans) => nd(ts.tail, acc ++ ans, f, m))
-  }
 \end{lstlisting}

### 2018-07-03 src/main/scala/refunc/RefuncECPS.scala

@@ -46,19 +46,15 @@ object RefuncECPS {
           val new_store = store.update(baddr, atomicEval(ae, env_, store))
           aeval(Config(e, new_env, new_store, new_time), seen, continue, m)
         }
-        def makeMCont(cls: List[Clos]): MCont = {
-          val mcont: MCont = { case (config, seen) =>
-            if (cls.isEmpty) aeval(config, seen, (c, seen, m) => m(c, seen), mcontinue)
-            else {
-              val Clos(Lam(v, body), c_env) = cls.head
-              val baddr = allocBind(v, new_time)
-              val new_env = c_env + (v -> baddr)
-              val new_store = store.update(baddr, argvs)
-              aeval(Config(body, new_env, new_store, new_time), seen, new_cont, makeMCont(cls.tail))
-            }
+        def makeMCont(cls: List[Clos]): MCont =
+          if (cls.isEmpty) mcontinue 
+          else { (config, seen) =>
+            val Clos(Lam(v, body), c_env) = cls.head
+            val baddr = allocBind(v, new_time)
+            val new_env = c_env + (v -> baddr)
+            val new_store = store.update(baddr, argvs)
+            aeval(Config(body, new_env, new_store, new_time), seen, new_cont, makeMCont(cls.tail))
           }
-          mcont
-        }
         val new_mcont: MCont = makeMCont(closures.tail)
         aeval(Config(body, new_env, new_store, new_time), new_seen, new_cont, new_mcont)
 

### 2018-07-03 paper/paper.tex
misc revision

@@
     additional layer of continuations due to nondeterminism.
-%
-    For clarity, we first present the refunctionalized AAM without caching. 
+
+    For clarity, we first present a vanilla version of refunctionalized AAM which simply
+    converts the stack structures to higher-order functions but keeps other parts unchanged.
     We then adopt a caching algorithm \cite{darais2017abstracting} to guarantee the termination 
@@
 \begin{lstlisting}
-                                  (letrec ([f1 (lambda (x)
-                                                 (let ([x1 (f2 x)]) x1))]
-                                           [f2 (lambda (y)
-                                                 (let ([y1 (f1 y)]) y1))])
-                                    (let ([z (f1 1)])
-                                      z))
+                            (letrec ([f1 (lambda (x) (let ([x1 (f2 x)]) x1))]
+                                     [f2 (lambda (y) (let ([y1 (f1 y)]) y1))])
+                               (let ([z (f1 1)]) z))
 \end{lstlisting}
@@
     step(nds) match {
-      case None => new_seen
-      case Some(nnds) => drive(nnds, new_seen)
-    }
-  }
+      case None => new_seen;  case Some(nnds) => drive(nnds, new_seen)
+    } }
 \end{lstlisting}
@@
 The @drive@ function is also changed: there is no work-list anymore, as all the potentially 
-unexplored states are embedded into the meta-continuation.
+unexplored states are embedded into the continuations.
 The termination of the analysis occurs when the value @None@ is returned from calling @step@.
@@
         drive_step(NDState(e, new_env, new_store, konts, new_time, ndk), seen)
-    }
-  }
+    } }
 \end{lstlisting}
@@
         drive_step(NDState(body, new_env, new_store, frames, time, new_ndk, seen))
-    }
-  }
+    } }
 \end{lstlisting}
@@
     function.
-  \item Since our final target is an abstract definitional interpreter (though still
-    nondeterministic), starting from this section our abstract semantic artifact
-    returns a set of final values instead of collected configurations.
+  \item The refunctionalized AAM collects all possible configurations, but our final target is an 
+    abstract definitional interpreter that cares about final evaluation results.
+    So starting from this section our abstract semantic artifact returns a set of final 
+    values instead of collected configurations. 
   \item Configurations still play an important role in caching, where a cache
@@
 As previously identified, there are two layers of continuations: one for the normal stack 
-of the analyzed language, and the other for the nondeterministic choices of closures.
+structure of the analyzed language, and the other for the nondeterministic choices of closures.
 This corresponds to the extended continuation-passing style.
@@
 We may first explicitly expose the continuation at the level of @aeval@, and then convert @nd@ 
-to continuation-passing style. Besides, these continuations have to be manifestly embedded
+to extended continuation-passing style. Besides, these continuations must be manifestly embedded
 at the call-site of @nd@, instead of implicitly provided by the @shift@ operator.
@@
 To conclude, starting from a small-step abstract abstract machine, eventually we obtain a 
-big-step definitional interpreter with pushdown control-flow, written in direct-style.
+big-step abstract definitional interpreter with pushdown control-flow, written in direct-style.
 

### 2018-07-03 paper/paper.tex
misc

@@
 Now we can dynamically generate the meta-continuation inside of a meta-continuation.
-Meanwhile, the @new_cont@ we defined above is used here, simply because different closures 
+Meanwhile, the @new_cont@ we defined above is used here, simply because of different closures 
 actually share the same continuation.
@@
 @z@ has two different values at two call sites, so variable @y@ would not be polluted by
-@1@. However, variable @x@ still points to the value set @{1, 2}@, because the two call-sites 
+@1@. However, variable @x@ still points to the value set @{1, 2}@, as the two call-sites 
 still share the same continuation and $1$-CFA does not increase the ability to separate the 

### 2018-07-04 paper/paper.tex
revision and update code

@@
 \subsubsection{Nondeterminism Abstraction}
-We observe that in the refunctionalized AAM, @makeMCont@ actually repeats the prologue
-part when matching a @let@ expression: they are both preparing a new environment and store 
-for the body expression of a closure. %TODO motivation
-We present a nondeterministic operator that abstracts this pattern:
+We observe that in the refunctionalized AAM, @makeMCont@ is invoked with the tail part of
+the closure set, and it actually repeats the code that creates the new environment and 
+store for the body expression of a closure.
+Based on observation, we present a nondeterministic operator that abstracts this pattern:
 

### 2018-07-04 src/main/scala/refunc/ANFAAMBasis.scala

@@ -37,6 +37,7 @@ object ANFAAM {
 
   abstract class Storable
   case class  Clos(v: Lam, env: Env) extends Storable
+  case class  CNumV(i: Int) extends Storable
   case object NumV extends Storable with Value
   type BStore = Store[BAddr, Storable]
 
@@ -62,7 +63,8 @@ object ANFAAM {
     State(e, env, bstore, Store[KAddr, Cont](Map(Halt -> Set())), Halt, List())
 
   def atomicEval(e: Expr, env: Env, bstore: BStore): Set[Storable] = e match {
-    case Num(_) | NumV => Set(NumV)
+    case Num(i) => Set(CNumV(i))
+    case NumV => Set(NumV)
     case Var(x) => bstore(env(x))
     case lam@Lam(x, body) => Set(Clos(lam, env))
     case _ => throw new NotImplementedError(e.toString)

### 2018-07-04 src/main/scala/refunc/DisentangledLinearSmallStepUB.scala

@@ -70,7 +70,7 @@ object DisLinearSmallStepUBStack {
     ndk match {
       case Nil => seen
       case NDCont(Nil, _, _, _, _)::ndk => 
-        drive_step(NDState(ae, env, bstore, konts, time, ndk), seen)
+        continue(NDState(ae, env, bstore, konts, time, ndk), seen)
       case NDCont(cls, argvs, bstore, time, frames)::ndk => 
         val Clos(Lam(v, body), c_env) = cls.head
         val baddr = allocBind(v, time)

### 2018-07-04 paper/paper.tex
revision

@@
 and update the @out@ cache with values from the @in@ cache. This retrieval from the @in@ 
-cache may return an empty set of values if there is no such mapping for the configuration 
-from the last iteration.
-In reality, we first conservatively assume that all computations can be diverged, and if 
-not, we update its values in the cache after the evaluation. In terms of partial orders and 
+cache may return an empty set of values if there is no such mapping for this configuration 
+from the previous iteration.
+In the initial iteration, we first conservatively assume that all computations can be diverged, 
+and if not, we update its values in the cache after the evaluation. In terms of partial orders and 
 lattices, it is the case that starts the Kleene iteration from the bottom of the lattice.
+The @out@ cache of this iteration will be used as @in@ cache for the next iteration.
 
@@
 
-% TODO: summarize this section
+Now, what we have obtained is an abstract interpreter written in continuation-passing style.
+Though the top-level function @aeval@ has only one continuation, but internally it uses
+a @nd@ operator which is written with two layers of continuations. Further more, we added
+a cache along with the interpreter and @nd@ operator.
 

### 2018-07-04 paper/paper.tex
fix typos

@@
 To effectively approximate the run-time behavior, we introduce program \emph{contours} @time@ 
-which are finite history of funciton calls till up to the current state \cite{Shivers:1991:SSC:115865.115884}.
+which are finite history of function calls till up to the current state \cite{Shivers:1991:SSC:115865.115884}.
 The function @tick@ is used to refresh the ``time'' and get a ''new time''. 
@@
 1) When constructing the @new_ndk@, we observe that it explicitly uses @new_frames@ as one
-argument of its top @NDCont@ object. The higer-order version of these continuations follows
+argument of its top @NDCont@ object. The higher-order version of these continuations follows
 the same pattern.
@@
 and refunctionalization include backtracking algorithms (e.g., regular expression matcher), 
-Dyck word recognizers, Dijkstra's shunting-yard algorithm, and even Quicksort 
+Dyck word recognizer, Dijkstra's shunting-yard algorithm, and even Quicksort 
 \cite{Danvy:2006:RW:2171265.2171268, DANVY2009534, Danvy:2001:DW:773184.773202}. \\
@@
 abstract interpreters is modularity. Hence, deploying different sensitivities
-or features can be achieved by just applying a different monad. Prior to that, \citet{Sergey:2013:MAI:2491956.2491979}
+or features can be achieved by just applying different monads. Prior to that, \citet{Sergey:2013:MAI:2491956.2491979}
 presented a monadic abstract interpreter for small-step semantics.

### 2018-07-04 paper/paper.tex
revision

@@
 is added to convert itself to type @State@.  
-
 Function @step@ now becomes of type @NDState => Option[NDState]@. Returning a @None@ value means
@@
 As such, we prepare a new environment, a new store, and a new frame list only for the
-first closure in that set. A new meta-continuation @new_ndk@ is also constructed,
-which contains the rest of closures, the values of argument @ae@, the store, the time,
-and the new frame list. We use the store before updating, because for different closures
-they may form different binding addresses @baddr@. We also use the new frames, because
-all the closures at this fork point share the same stack structure and return point.
+first closure in that set. A new meta-continuation @new_ndk@ is constructed,
+which contains the rest of closures, the values of evaluating @ae@, the store, the time,
+and the new frame list. We use the store before updating, because different 
+closures may form different binding addresses @baddr@. We also use the \textit{new} frames,
+because all the closures at this fork point share the same stack structure and return point.
 
@@
 
-If the continuation @konts@ is empty, then it means we have reached an end of the computation;
+If the continuation @konts@ is empty, then the machine has reached an end of the computation;
 otherwise, we should return the values of @ae@ to its caller which is contained in the top
@@
 \begin{itemize}
-  \item If @ndk@ is also empty, then we have reached the end of all computations. Besides,
-    the control component @ae@ now is an atomic value, therefore @None@ is returned.
-  \item If the target closure set of the front-most @NDCont@ is empty (@Nil@ in the code),
+  \item If the meta-continuation @ndk@ is also empty, then the machine has reached the end of \textit{all} 
+    computations. Besides, the control component @ae@ now is an atomic value, therefore @None@ is returned.
+  \item If the target closures set of the front-most @NDCont@ is empty (@Nil@ in the code),
     then we have computed all the closures at this fork point and should move to the next fork 
-    point. Since the meta-continuation @nkd@ is represented by a list, and we always
-    append new elements to its front, we resume to the most recent fork point by
-    popping up the front-most element and using the tail of that list.
-  \item Otherwise, we pop a closure from the set, then build a new environment and store for 
-    the body expression of that closure. We use the same frame list and time which are copied 
-    from the fork point rather than current one.
+    point. Since the meta-continuation @nkd@ is represented by a list, and we always append new 
+    elements to its front, we resume to the most recent fork point by popping up the front-most 
+    element and using the tail of that list.
+  \item Otherwise, we pop a closure from the set, and build a new environment and store for 
+    evaluating the body expression of that closure. We use the same frame list and time which 
+    are copied from the fork point rather than current one.
     The meta-continuation @new_ndk@ is updated by the removal of that popped closure, which makes
@@
 as arguments and returns a set of reachable states once it terminates.
-It is is essentially formed by merging the functionality of @step@ into the @drive@ function.
-Since each case in function @step@ are statically known, we inline the pattern match on the
-@Option@ type and replace it with @new_seen@ or a recursive call to @drive_step@ for @None@
-or @Some@, respectively.
+It is essentially formed by merging the functionality of @step@ into the @drive@ function.
+Since each case in function @step@ are statically known, we can inline the pattern match on the
+@Option@ type and replace it with returning @new_seen@ when @None@ is matched, or calling 
+@drive_step@ recursively when @Some@ is matched.
 
@@
 functions:
-@drive_step@, which plays the same role as before; 
-@continue@, which is invoked from @drive_step@ when encountering an atomic expression, 
+1) @drive_step@, which plays the same role as before; 
+2) @continue@, which is invoked from @drive_step@ when encountering an atomic expression, 
   and dispatches the reduction context of the analyzed language; and 
-@ndcontinue@, which is invoked from @continue@ when the normal stack is empty,
-  and dispatches the meta-continuations @NDCont@.
+3) @ndcontinue@, which is invoked from @continue@ when the normal stack is empty,
+  and dispatches the meta-continuation @NDCont@.
 
@@
 When @ndk@ is empty, the current explored states @seen@ is returned; the analysis then terminates.
-\iffalse
-The shape for @ndcontinue@ is similar to @continue@; after all, they both use a 
-@List@ representation for continuations. When the closure set is empty, we recursively 
-call @drive_step@ with the rest of @ndk@, since in this case the continuation 
-@konts@ is still empty, so the state will be dispatched to @ndcontinue@ again.
-Otherwise, we invoke @drive_step@ with the a new @NDState@ where the control 
-is transferred to the body expression of the next closure object.
-\fi
+Otherwise, a new @NDState@ is constructed depending on the status of the closures set.
+% first call to drive_step can be replaced with continue. ae is atomic, konts is empty.
 
@@
 language be part of the control flows of the meta-language.
+In this section, we first present a direct refunctionalization of the continuations.
+Next, we adopt a caching algorithm to obtain a computable analysis\cite{darais2017abstracting},
+and further prepare the interpreter for tranforming to direct-style later.
 
@@
 we introduce a configuration class @Config@ as a state-like definition by eliminating
-@konts@ and @ndk@ from the definition of state. The type of return value of the abstract
-interpreter accordingly becomes @Set[Config]@.
+@konts@ and @ndk@ from the definition of state. Accordingly, the abstract interpreter
+now returns @Set[Config]@.
 
@@
 we may find it calls the meta-continuation @ndcontinue@ inside of its body.
-This observation leads us to define the types of higher-order continuation and meta-continuation:
+This observation leads us to define the types of higher-order functions representing 
+continuation and meta-continuation:
 

### 2018-07-04 paper/paper.tex
fix grammar issues; update code

@@
     of abstract interpretation.
-    In the end of this section, we review the pushdown control-flow analysis and examine how 
+    At the end of this section, we review the pushdown control-flow analysis and examine how 
     computable and precise call/return matching is obtained through these transformations
@@
 
-The CESK machine is an abstract machine for describing semantics of and evaluating
+The CESK machine is an abstract machine for describing the semantics of and evaluating
 $\lambda$-calculus \cite{felleisen1987calculus}. The CESK machine models program execution
@@
 @step@ on the current state until a state is reached which the control is an atomic 
-expression and the continuation stack is empty. Naturally, we can then extract the value
+expression and the stack structure is empty. Naturally, we can then extract the value
 from the final state.
@@
 the stack. By doing so, we recover the call structure as the same as at run-time
-(so called pushdown control-flow analysis), but since the stack is unbounded, 
+(so-called pushdown control-flow analysis), but since the stack is unbounded, 
 the state space is potentially infinite.
@@
 stack still may diverge when analyzing some programs.
-This is because the unbounded stack can grow to arbitrary depth which implies that the state
+This is because the unbounded stack can grow to an arbitrary depth which implies that the state
 space is possibly infinite; the analysis may therefore not terminate for all programs if we
@@
 Function @f1@ and @f2@ mutually invoke each other, so the stack will alternately pushing
-frames @f1@ and @f2@ onto the top of current stack. However, no two existing stack
+frames @f1@ and @f2@ onto the top of the current stack. However, no two existing stack
 components are identical in the state space. One can decide the reachability of a state
@@
 It is essentially formed by merging the functionality of @step@ into the @drive@ function.
-Since each case in function @step@ are statically known, we can inline the pattern match on the
+Since each case in function @step@ is statically known, we can inline the pattern match on the
 @Option@ type and replace it with returning @new_seen@ when @None@ is matched, or calling 
@@
 With this, we have a single function to perform both abstract evaluation and the collection of
-intermediate states. Given @inject(e)@ as an initial state and an empty set as the
+intermediate states. Given @inject(e)@ as the initial state and an empty set as the
 initial set of states reached, we can define the entrance function @analyze@ as follows:
@@
 Next, we adopt a caching algorithm to obtain a computable analysis\cite{darais2017abstracting},
-and further prepare the interpreter for tranforming to direct-style later.
+and further prepare the interpreter for transforming to direct-style later.
 
@@
 we introduce a configuration class @Config@ as a state-like definition by eliminating
-@konts@ and @ndk@ from the definition of state. Accordingly, the abstract interpreter
+@konts@ and @ndk@ from the definition of states. Accordingly, the abstract interpreter
 now returns @Set[Config]@.
@@
     of all possible concrete reachable paths and values.
-  \item We simplify the continuations with an nondeterministic operator @nd@ such 
+  \item We simplify the continuations with a nondeterministic operator @nd@ such 
     that the function @aeval@ only uses one layer of continuation.
@@
 from the previous iteration.
-In the initial iteration, we first conservatively assume that all computations can be diverged, 
+In the initial iteration, we first conservatively assume that all computations can diverge, 
 and if not, we update its values in the cache after the evaluation. In terms of partial orders and 
 lattices, it is the case that starts the Kleene iteration from the bottom of the lattice.
-The @out@ cache of this iteration will be used as @in@ cache for the next iteration.
+The @out@ cache of this iteration will be used as the @in@ cache for the next iteration.
 
@@
 Given the skeleton of @aeval@ in cache-and-continuation-passing style, we may now
-use the @nd@ operator to fill the details. The first case of match is that @e@
+use the @nd@ operator to fill in the details. The first case of match is that @e@
 is a @let@ expression:
@@
     nd[Storable](closures, Ans(Set[VS](), new_cache), { case (cls, clscache, clsnd) =>
-        val Clos(Lam(v, body), c_env) = cls
-        val vbaddr = allocBind(v, new_time);  val new_cenv = c_env + (v -> vbaddr)
-        val new_store = store.update(vbaddr, atomicEval(ae, env, store))
+      val Clos(Lam(v, body), c_env) = cls;   val vbaddr = allocBind(v, new_time)
+      val new_cenv = c_env + (v -> vbaddr); val new_store = store.update(vbaddr, atomicEval(ae, env, store))
       aeval(body, new_cenv, new_store, new_time, clscache, { case Ans(bdvss, bdcache) =>
-          nd[VS](bdvss, Ans(Set[VS](), bdcache), { case (VS(vals, time, vssstore), bdvsscache, bdnd) =>
-              val baddr = allocBind(x, time);  val new_env = env + (x -> baddr)
-              val new_store = vssstore.update(baddr, vals)
+        nd[VS](bdvss, Ans(Set[VS](), bdcache), { case (vs, bdvsscache, bdnd) =>
+          val Val(vals, time, vsstore) = vs;  val baddr = allocBind(x, time)
+          val new_env = env + (x -> baddr);  val new_store = vssstore.update(baddr, vals)
           aeval(e, new_env, new_store, time, bdvsscache, { case Ans(evss, ecache) =>
@@
 Though the top-level function @aeval@ has only one continuation, but internally it uses
-a @nd@ operator which is written with two layers of continuations. Further more, we added
+a @nd@ operator which is written with two layers of continuations. Furthermore, we added
 a cache along with the interpreter and @nd@ operator.
@@
 \textit{The Problem with Return-flows.}
-Pushdown control-flow is a property in analysis that precisely models the run-time
+Pushdown control-flow is a property in the analysis that precisely models the run-time
 call structure of the analyzed program. A pushdown control-flow analysis provides an 
@@
 in extended continuation-passing style.
-To obtain a definitional abstract interpreter in direct style, we have multiple choices 
+To obtain a definitional abstract interpreter in direct-style, we have multiple choices 
 of how to proceed:
@@
 
-The second operator @choices@ is implemented with delimited control operator 
+The second operator @choices@ is implemented with the delimited control operator 
 @shift@ but simply calls our first operator @nd@ in its body.
@@
 The function @analyze@ is also changed by calling the @reset@ operator around the function 
-call of @aeval@ to set the delimiter. The afterwards updating on cache becomes sequential.
+call of @aeval@ to set the delimiter. The afterward updating on cache becomes sequential.
 
@@
 \citet{Danvy:2001:DW:773184.773202} observed that refunctionalizing a regular expression matcher 
-with two stacks yields its counterpart with two layer of continuations.
+with two stacks yields its counterpart with two layers of continuations.
 In addition, the operational semantics of control operators (e.g., @shift@ and @reset@) can be given by an 
@@
 CFA2 is the first solution that solves the return-flow problem~\cite{vardoulakis2010cfa2},
-but CFA2 has several limitation: it only works on continuation-passing style programs, and
+but CFA2 has several limitations: it only works on continuation-passing style programs, and
 does not support polyvariant analysis, in addition to having an exponential time complexity.

### 2018-07-05 paper/paper.tex
revise 6/7 and code

@@
 \begin{itemize}
-  \item The components of the configurations are lifted as arguments of the evaluation
+  \item The components of the configuration are lifted as arguments of the evaluation
     function.
@@
     ensures termination of the analysis, as well as performs a sound over-approximation
-    of all possible concrete reachable paths and values.
+    of all possible values and reachable paths.
   \item We simplify the continuations with a nondeterministic operator @nd@ such 
@@
 
-\subsubsection{Caches}
+\subsubsection{Cache-Passing Style}
 % why cache? problem of set of states.
@@
 iteration, we will use the @out@ cache from the previous iteration as the @in@ cache.
-
-\subsubsection{Cache-Passing Style}
-%motivation
-We can now transform our interpreter into cache-passing style. We begin by defining the 
+The iteration terminates when the @out@ cache is identical to the @in@ cache.
+
+We can now transform our abstract interpreter into cache-passing style. We begin by defining the 
 return type @Ans@ as a case class which contains a @VS@ set and a cache. @Ans@ also implements 
@@
     if (cache.outContains(config)) continue(Ans(cache.outGet(config), cache))
-    else { val new_cache = cache.outUpdate(config, cache.inGet(config))
-           e match { ... } }
-  }
+    else { 
+      val new_cache = cache.outUpdate(config, cache.inGet(config))
+      e match { ... } } }
 \end{lstlisting}
@@
 If, however, we have not hit the @out@ cache, we retrieve the values from the @in@ cache, 
-and update the @out@ cache with values from the @in@ cache. This retrieval from the @in@ 
+and update the @out@ cache with the values from the @in@ cache. This retrieval from the @in@ 
 cache may return an empty set of values if there is no such mapping for this configuration 
@@
 \subsubsection{Nondeterminism Abstraction}
+
 We observe that in the refunctionalized AAM, @makeMCont@ is invoked with the tail part of
@@
 \begin{lstlisting}
-  def nd[T](ts: Iterable[T], acc: Ans, f: (T, Cache, Ans => Ans) => Ans, m: Ans => Ans): Ans =
-    if (ts.isEmpty) m(acc)
-    else f(ts.head, acc.cache, (ans: Ans) => nd(ts.tail, acc ++ ans, f, m))
-\end{lstlisting}
-
-Given a collection of elements of type @T@, an initial accumulator of type 
-@Ans@, a function @f@ that works on element's type @T@ and takes a continuation 
-of type @Ans=>Ans@, and a function @m@ that is also type @Ans=>Ans@, @nd@ chooses
-an element in the collection and applies @f@ to it, accumulates the value 
-(@acc++ans@ where the @ans@ is the result comes from @f@) and eventually applies 
-function @m@ on the accumulated value when there is no element in the collection.
-This function is also written in cache-passing style, and the current accumulated 
-cache @acc.cache@ is exposed to @f@ when applied.
-
-Our function @nd@ is a variant implementation of @fold@ in continuation-passing style,
-and @f@ also has to be passed as a function that takes a continuation.
+  def nd[T](ts: Iterable[T], acc: Ans, k: (T, Cache) => Ans): Ans =
+    if (ts.isEmpty) acc else nd(ts.tail, acc ++ k(ts.head, acc.cache), k)
+\end{lstlisting}
+
+Given a collection of elements of type @T@, an initial accumulator of type @Ans@,
+a function @k@ that works on elements of type @T@, @nd@ chooses an element in
+the collection and applies @k@ to it iteratively, and accumulates the value that
+returned from @k@ when recuring on the rest of the collection. Eventually, @nd@
+returns the accumulated value @acc@ when there is no element in the collection.
+The latest accumulated cache @acc.cache@ is exposed to @k@ when applied.
 
 \subsubsection{Put Them Together}
-Given the skeleton of @aeval@ in cache-and-continuation-passing style, we may now
+
+With the skeleton of cache-and-continuation-passing style @aeval@ in mind, we may now
 use the @nd@ operator to fill in the details. The first case of match is that @e@
@@
     val closures = atomicEval(f, env, store)
-    nd[Storable](closures, Ans(Set[VS](), new_cache), { case (cls, clscache, clsnd) =>
+    nd[Storable](closures, Ans(Set[VS](), new_cache), { case (cls, clscache) =>
       val Clos(Lam(v, body), c_env) = cls;   val vbaddr = allocBind(v, new_time)
@@
       aeval(body, new_cenv, new_store, new_time, clscache, { case Ans(bdvss, bdcache) =>
-        nd[VS](bdvss, Ans(Set[VS](), bdcache), { case (vs, bdvsscache, bdnd) =>
+        nd[VS](bdvss, Ans(Set[VS](), bdcache), { case (vs, bdvsscache) =>
           val Val(vals, time, vsstore) = vs;  val baddr = allocBind(x, time)
@@
           aeval(e, new_env, new_store, time, bdvsscache, { case Ans(evss, ecache) => 
-            /* evaluate e with new env and store, then transfer to the next body's value */
-            bdnd(Ans(evss, ecache.outUpdate(config, evss)))
-          })
-        }, clsnd) /* transfer to the next target closure of f */
-      })
-    }, continue)
+            continue(Ans(evss, ecache.outUpdate(config, evss)))
+          }) }) }) })
 \end{lstlisting}
@@
 We invoke @nd@ twice and perform a nesting two-step, depth-first evaluation over 
-possible values of @App(f, ae)@. The values of branches from a fork point will be
-accumulated inside of @nd@, and then returned to the upper fork point.
-
+possible values of @App(f, ae)@. 
 The outer application of @nd@ evaluates over the set of possible target closures of
@@
 Remember that the function @aeval@ returns a set of @VS@ objects to its continuation,
-so the continuation argument @bdvss@ is a set of @VS@ objects which represent values
-from multiple computation paths when evaluating this single @body@ expression.
+so the continuation argument @bdvss@ represents a set of values from multiple computation 
+paths when evaluating this single @body@ expression.
 
 We next consider the inner application of @nd@, which evaluates over the set of body's 
-value/stores @bdvss@. For each @VS@, the time-stamp and store in @VS@ will be instantiated to
-the next application of @aeval@. The inner application of @aeval@ evaluates @e@ with the
-substituted environment and store, then returns a set of values of one computation path 
-to its continuation @bdnd@, where the continuation will accumulate the result and 
-transfer the evaluation to the next body value.
-The meta-continuation of the inner @nd@ is @clsnd@, meaning that we need to transfer 
-to the next target closure of @f@. The meta-continuation of the outer @nd@ is just 
-@continue@, as what we normally would do when writing CPS interpreters.
-
+value/stores @bdvss@. For each @VS@, the time-stamp and the store in @VS@ will be 
+instantiated to the next application of @aeval@. The environment @new_env@ is built on
+the outer environment variable @env@, which is the environment of this @let@ expression.
+The inner application of @aeval@ evaluates @e@ with the substituted environment and store,
+then returns a set of values of one computation path to its continuation, where we
+apply the @continue@ function with these values and the updated cache. 
 Note that the cache is used in a monotonic way: each function call to @aeval@ or @nd@ 
@@
 
-Once the evaluation is completed at each case, we obtain a set of @VS@ objects and must update the @out@ cache.
-In the first case of pattern matching, this happens at the innermost continuation of @aeval@,
-where we reach the end of one computation path.
+Once the evaluation is completed at each case, we obtain a set of @VS@ objects and must 
+update the @out@ cache. In the first case of pattern matching, this happens at the 
+innermost continuation of @aeval@, where we reach the end of one computation path.
 For the second case, we update the @out@ cache before calling the @continue@ function.
@@
     def iter(cache: Cache): Ans = { 
-      val Ans(vss, new_cache) = aeval(e, mtEnv, mtStore, mtTime, cache, { case Ans(vss, cache) => 
-          Ans(vss, cache.outUpdate(Config(e, mtEnv, mtStore, mtTime), vss))
-      })
-      if (new_cache.out == new_cache.in) Ans(vss, new_cache)
-      else iter(Cache(new_cache.out, Store[Config, VS](Map())))
-    } iter(mtCache) }
-\end{lstlisting}
-
-% TODO: explain more on how caches work
+      val Ans(vss, anscache) = aeval(e, mtEnv, mtStore, mtTime, cache, ans => ans)
+      if (anscache.out == anscache.in) Ans(vss, anscache)
+      else iter(Cache(anscache.out, Store[Config, VS](Map())))
+    } iter(Cache.mtCache) }
+\end{lstlisting}
+
 Finally, the @analyze@ function (as the entrance of the analysis) does a looping iteration
@@
 the next iteration; in the meanwhile, an empty cache will be used as the @out@ cache.
-Note that the initial continuation of calling @aeval@ updates the values for the initial 
-configuration in the final cache.
-
-Now, what we have obtained is an abstract interpreter written in continuation-passing style.
-Though the top-level function @aeval@ has only one continuation, but internally it uses
-a @nd@ operator which is written with two layers of continuations. Furthermore, we added
-a cache along with the interpreter and @nd@ operator.
+
+Now, as we can see, by using the @nd@ operator which abstracts over the meta-continuation,
+the evaluation function can be simplified to one layer continuation-passing style.
+The use two layers of continuations are unchained to two functions that each of them has one
+continuation.
+Now the continuation of @aeval@ plays the same role as the one in a concrete CPS interpreter.
+Furthermore, we added a cache along with the abstract interpreter and @nd@ operator.
 
@@
 
-In the previous section, we presented a refunctionalized AAM utilizing the @nd@ operator
-in extended continuation-passing style.
-To obtain a definitional abstract interpreter in direct-style, we have multiple choices 
-of how to proceed:
+In the previous section, we presented a refunctionalized AAM in extended continuation-passing style,
+and then simplified it by utilizing the @nd@ operator.
+To obtain a definitional abstract interpreter in direct-style, we have multiple choices of how to proceed:
+
 \begin{itemize}
@@
     \citet{darais2017abstracting} describe.
-  \item By representing the extended continuation-passing style with delimited control operators
+  \item By representing the continuation-passing style with delimited control operators
     such as @shift@ and @reset@, we derive a new form of abstract interpreters in direct-style.
@@
 
-As previously identified, there are two layers of continuations: one for the normal stack 
-structure of the analyzed language, and the other for the nondeterministic choices of closures.
-This corresponds to the extended continuation-passing style.
-But in our defining language (Scala), the delimited control operators rely on the continuation-passing 
-transformation which is only one layer.
-
-To address this problem, we reintroduce two operators @nd@ and @choices@ for handling 
-nondeterministic choices in a way that is direct-style but mimics extended 
-continuation-passing style. \\
-
-\textit{Nondeterminism Operators, Again.}
-%Introducing @Cache@ explicitly as the part of arguments of continuation @k@ reveals 
-%that cache should be monotonically accumulated to the rest of computation. 
-The first operator @nd@ is adapted from the previous one, but the meta-continuation
-@m@ has been eliminated here. Now the only explicit continuation @k@ applies on tuple 
-type @(T, Cache)@ where the value of type @T@ comes from the collection, and @nd@ performs 
-recursive invocation on itself over elements in the collection @ts@.
-
-\begin{lstlisting}
-  def nd[T](ts: Iterable[T], acc: Ans, k: ((T, Cache)) => Ans): Ans = {
-    if (ts.isEmpty) acc
-    else nd(ts.tail, acc ++ k(ts.head, acc.cache), k)
-  }
-\end{lstlisting}
-
-The second operator @choices@ is implemented with the delimited control operator 
-@shift@ but simply calls our first operator @nd@ in its body.
-
-\begin{lstlisting}
-  def choices[T](ts: Iterable[T], cache: Cache): (T, Cache) @cps[Ans] = shift {
-    f: (((T, Cache)) => Ans) => nd(ts, Ans(Set[VS](), cache), f)
-  }
-\end{lstlisting}
-
-The function @choices@ takes two arguments: a set of elements of type @T@, and an initial cache.
-The return type of @choices@ is @(T, Cache)@ with an annotation \verb|@cps[Ans]|, which 
-denotes that @choices@ will iteratively return an element of type @T@ from the collection
-along with the latest cache, but the final returned value of the continuation is type @Ans@.
-The function @f@ is the delimited continuation captured by the @shift@ operator, which
-is formed by the computation will be executed after the @choices@'s call-site.
-
-Now we can use these two operators to redefine @aeval@. The abstract evaluator @aeval@ is
-rewritten in direct-style, so that the last argument @cont@ disappears; we also add an 
-annotation \verb|@cps[Ans]| on the return type to enable using delimited control operators 
-inside of it.
-Then, for every nondeterministic choice in the abstract interpretation, we can simply call 
-@choices@ on the set of possible choices and write the program sequentially as concrete 
-interpreters, and the subsequent statements and expressions become the delimited continuation
-@f@. We have two invocations of @choices@, which correspond to the two invocations of @nd@ in
-the refunctionalized form.
-The function @choices@ also returns the latest cache to its left-hand side when invoked.
-The cache will be still accumulated to the next call to @choices@ or @aeval@, 
-given the fact that the abstract interpreter should always use the latest cache.
+We first transform the function @aeval@ to direct-style by removing the additional 
+continuation argument @continue@ and sequentializing the expressions in @aeval@.
+To enable using delimited control operators inside of @aeval@ we also add an 
+annotation \verb|@cps[Ans]| on the return type, which tells the compiler to CPS
+transform this function.
+Back to direct-style on the case where @e@ appears as an atomic expression is straightforward: 
+we simply remove the application of @continue@. Now the abstract evaluator @aeval@ looks as follows:
 
@@
       e match {
+        case Let(x, App(f, ae), e) => ...
+        case ae if isAtomic(ae) =>
+          val vss = Set(VS(atomicEval(ae, env, store), new_time, store))
+          Ans(vss, new_cache.outUpdate(config, vss))
+        } } }
+\end{lstlisting}
+
+The case where @e@ is a @let@ expression is not that obvious. As previously identified, 
+function @nd@ works on a unknown number of nondeterministic choices, and it takes a 
+function @k@ that works for every element in the collection. Here we present another
+function @choices@ that uses delimited control operator @shift@ to capture that ``@k@''
+when the program is direct-style and pass it to @nd@.
+
+\begin{lstlisting}
+  def nd[T](ts: Iterable[T], acc: Ans, k: ((T, Cache)) => Ans): Ans = {
+    if (ts.isEmpty) acc else nd(ts.tail, acc ++ k(ts.head, acc.cache), k)
+  }
+  def choices[T](ts: Iterable[T], cache: Cache): (T, Cache) @cps[Ans] = shift {
+    f: (((T, Cache)) => Ans) => nd(ts, Ans(Set[VS](), cache), f)
+  }
+\end{lstlisting}
+
+The function @choices@ takes two arguments: a set of elements of type @T@, and an initial cache.
+The return type of @choices@ is @(T, Cache)@ with an annotation \verb|@cps[Ans]|, which 
+denotes that @choices@ will iteratively return an element of type @T@ from the collection
+along with the latest cache once @f@ is invoked, but the final returned value of the 
+function is type @Ans@.
+The function @f@ is the delimited continuation captured by the @shift@ operator, which
+is formed by the computation will be executed after the @choices@'s call-site.
+Once the function @f@ (which is the function @k@ in @nd@) is invoked and returned, 
+it returns to its call-site where @nd@ will be invoked on the rest elements of the collection.
+Now we can use this function to sequentialize the @let@ case.
+
+\begin{lstlisting}
   case Let(x, App(f, ae), e) =>
@@
     Ans(evss, ecache.outUpdate(config, evss))
-        case ae if isAtomic(ae) =>
-          val vss = Set(VS(atomicEval(ae, env, store), new_time, store))
-          Ans(vss, new_cache.outUpdate(config, vss))
-      }
-    }
-  }
-\end{lstlisting}
+\end{lstlisting}
+
+For every nondeterministic choice in the abstract interpretation, we can simply call 
+@choices@ on the set of possible choices and write the program sequentially as concrete 
+interpreters, and the subsequent statements and expressions after the call-site become 
+the delimited continuation @f@. We have two invocations of @choices@, which correspond 
+to the two invocations of @nd@ in the refunctionalized form in the last section.
+The function @choices@ also returns the latest cache to its left-hand side when invoked.
+The cache will be still accumulated to the next call to @choices@ or @aeval@, 
+given the fact that the abstract interpreter should always use the latest cache.
 
 The function @analyze@ is also changed by calling the @reset@ operator around the function 
-call of @aeval@ to set the delimiter. The afterward updating on cache becomes sequential.
+call of @aeval@ to set the delimiter.
 
@@
       val Ans(vss, anscache) = reset { aeval(e, mtEnv, mtStore, mtTime, cache) }
-      val new_cache = anscache.outUpdate(Config(e, mtEnv, mtStore, mtTime), vss)
-      if (new_cache.out == new_cache.in) Ans(vss, new_cache)
-      else iter(Cache(new_cache.out, Store[Config, VS](Map())))
-    } iter(mtCache) }
+      if (anscache.out == anscache.in) Ans(vss, anscache)
+      else iter(Cache(anscache.out, Store[Config, VS](Map())))
+    } iter(Cache.mtCache) }
 \end{lstlisting}
@@
 Now we have arrived at the end of this series of transformations. 
-The inverse transformation of the last step is basically CPS transforming the function @aeval@ and @nd@.
-We may first explicitly expose the continuation at the level of @aeval@, and then convert @nd@ 
-to extended continuation-passing style. Besides, these continuations must be manifestly embedded
-at the call-site of @nd@, instead of implicitly provided by the @shift@ operator.
+The inverse transformation of the last step is basically CPS transforming the function @aeval@ and
+explicitly using function @nd@ instead of function @choices@. Then, these continuations must be 
+manifestly embedded at the call-site of @nd@, instead of implicitly provided by the @shift@ operator.
 

### 2018-07-05 src/main/scala/refunc/DirectStyle.scala

@@ -66,10 +66,9 @@ object DirectStyleDC {
   def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore) = {
     def iter(cache: Cache): Ans = {
       val Ans(vss, anscache) = reset { aeval(e, env, store, mtTime, cache) }
-      val init_config = Config(e, env, store, mtTime)
-      val new_cache = anscache.outUpdate(init_config, vss)
-      if (new_cache.out == new_cache.in) { Ans(vss, new_cache) }
-      else { iter(Cache(new_cache.out, Store[Config, VS](Map()))) }
+      assert(anscache.outContains(Config(e, env, store, mtTime)))
+      if (anscache.out == anscache.in) { Ans(vss, anscache) }
+      else { iter(Cache(anscache.out, Store[Config, VS](Map()))) }
     }
     iter(Cache.mtCache)
   }
@@ -133,10 +132,10 @@ object DirectStyleSideEff {
 
   def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore) = {
     def iter(cache: Cache): Ans = {
-      val Ans(vss, new_cache) = aeval(e, env, store, mtTime, cache)
-      val updated_cache = new_cache.outUpdate(Config(e, env, store, mtTime), vss)
-      if (updated_cache.out == new_cache.in) { Ans(vss, updated_cache) }
-      else { iter(Cache(updated_cache.out, Store[Config, VS](Map()))) }
+      val Ans(vss, anscache) = aeval(e, env, store, mtTime, cache)
+      assert(anscache.outContains(Config(e, env, store, mtTime)))
+      if (anscache.out == anscache.in) { Ans(vss, anscache) }
+      else { iter(Cache(anscache.out, Store[Config, VS](Map()))) }
     }
     iter(Cache.mtCache)
   }

### 2018-07-05 src/main/scala/refunc/RefuncCPS.scala

@@ -7,9 +7,9 @@ import ANFAAM._
 object RefuncCPS {
   type Cont = Ans => Ans
 
-  def nd[T](ts: Iterable[T], acc: Ans, k: (T, Cache, Ans => Ans) => Ans, m: Ans => Ans): Ans = {
-    if (ts.isEmpty) m(acc)
-    else k(ts.head, acc.cache, (ans: Ans) => nd(ts.tail, acc ++ ans, k, m))
+  def nd[T](ts: Iterable[T], acc: Ans, k: (T, Cache) => Ans): Ans = {
+    if (ts.isEmpty) acc
+    else nd(ts.tail, acc ++ k(ts.head, acc.cache), k)
   }
   
   def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, continue: Cont): Ans = {
@@ -40,23 +40,23 @@ object RefuncCPS {
         case Let(x, App(f, ae), e) =>
           val closures = atomicEval(f, env, store)
           nd[Storable](closures, Ans(Set[VS](), new_cache), { 
-            case (Clos(Lam(v, body), c_env), clscache, clsnd) =>
+            case (Clos(Lam(v, body), c_env), clscache) =>
               val vbaddr = allocBind(v, new_time)
               val new_cenv = c_env + (v -> vbaddr)
               val new_store = store.update(vbaddr, atomicEval(ae, env, store))
-              aeval(body, new_cenv, new_store, new_time, clscache, { case Ans(bdvss, bdcache) =>
-                nd[VS](bdvss, Ans(Set[VS](), bdcache), { 
-                  case (VS(vals, time, vssstore), bdvsscache, bdnd) =>
-                    val baddr = allocBind(x, time)
-                    val new_env = env + (x -> baddr)
-                    val new_store = vssstore.update(baddr, vals)
-                    aeval(e, new_env, new_store, time, bdvsscache, { 
-                      case Ans(evss, ecache) => bdnd(Ans(evss, ecache.outUpdate(config, evss)))
-                    })
-                }, clsnd)
+              aeval(body, new_cenv, new_store, new_time, clscache, { 
+                case Ans(bdvss, bdcache) =>
+                  nd[VS](bdvss, Ans(Set[VS](), bdcache), { 
+                    case (VS(vals, time, vssstore), bdvsscache) =>
+                      val baddr = allocBind(x, time)
+                      val new_env = env + (x -> baddr)
+                      val new_store = vssstore.update(baddr, vals)
[…]

### 2018-07-05 paper/paper.tex
minor revisions

@@
 these two cases and evaluates atomic expressions to closures in a straightforward way.
-The @alloc@ function generates a fresh address, and always allocates a unique integer
+The @alloc@ function generates a fresh address, and always allocates an unique integer
 in the domain of @store@.
@@
 
-The above code shows the skeleton of @drive_step@.
+The above code shows the skeleton of the @drive_step@.
 At the end of the first case of pattern matching, we invoke @drive_step@ recursively 
@@
 return type @Ans@ as a case class which contains a @VS@ set and a cache. @Ans@ also implements 
-its own accumulating operation @++@ which defines as a union of @Set[VS]@s and a join of caches.
+its own accumulating operation @++@ which defines as an union of @Set[VS]@s and a join of caches.
 The join operation of caches is defined as joining its two store components.
@@
 the collection and applies @k@ to it iteratively, and accumulates the value that
-returned from @k@ when recuring on the rest of the collection. Eventually, @nd@
+returned from @k@ when recurring on the rest of the collection. Eventually, @nd@
 returns the accumulated value @acc@ when there is no element in the collection.
@@
 the evaluation function can be simplified to one layer continuation-passing style.
-The use two layers of continuations are unchained to two functions that each of them has one
+The use of two layers of continuations are unchained to two functions that each of them has one
 continuation.
-Now the continuation of @aeval@ plays the same role as the one in a concrete CPS interpreter.
-Furthermore, we added a cache along with the abstract interpreter and @nd@ operator.
+The only continuation of @aeval@ plays the same role as the one in a concrete CPS interpreter.
+Furthermore, we added a cache along with the abstract interpreter and the @nd@ operator.
 
@@
 The case where @e@ is a @let@ expression is not that obvious. As previously identified, 
-function @nd@ works on a unknown number of nondeterministic choices, and it takes a 
+function @nd@ works on an unknown number of nondeterministic choices, and it takes a 
 function @k@ that works for every element in the collection. Here we present another
@@
 
-For every nondeterministic choice in the abstract interpretation, we can simply call 
+For every nondeterministic fork-point in the abstract interpretation, we can simply call 
 @choices@ on the set of possible choices and write the program sequentially as concrete 

### 2018-07-05 paper/paper.tex
minor revisions

@@
     values instead of collected configurations. 
-  \item Configurations still play an important role in caching, where a cache
+  \item The configurations still play an important role in caching, where a cache
     is a mapping from configurations to sets of final values. The caching algorithm
@@
     of all possible values and reachable paths.
-  \item We simplify the continuations with a nondeterministic operator @nd@ such 
-    that the function @aeval@ only uses one layer of continuation.
+  \item The function @aeval@ is simplified by introducing and using a nondeterministic 
+    operator @nd@ which makes @aeval@ only uses one layer of continuation.
 \end{itemize}
+
+We will begin by discussing these changes separately, and finally see how they work together.
 
@@
 continuation argument @continue@ and sequentializing the expressions in @aeval@.
-To enable using delimited control operators inside of @aeval@ we also add an 
+To enable using delimited control operators inside of @aeval@, we also add an 
 annotation \verb|@cps[Ans]| on the return type, which tells the compiler to CPS
@@
 Back to direct-style on the case where @e@ appears as an atomic expression is straightforward: 
-we simply remove the application of @continue@. Now the abstract evaluator @aeval@ looks as follows:
+we simply unwrap the application of @continue@. Now the abstract evaluator @aeval@ looks as follows:
 
@@
 function @choices@ that uses delimited control operator @shift@ to capture that ``@k@''
-when the program is direct-style and pass it to @nd@.
+when the program is written in direct-style and pass it to @nd@.
 
@@
 Once the function @f@ (which is the function @k@ in @nd@) is invoked and returned, 
-it returns to its call-site where @nd@ will be invoked on the rest elements of the collection.
-Now we can use this function to sequentialize the @let@ case.
+it returns to its call-site where @nd@ will be invoked on the remaining elements of the collection.
+Now we can use this function to sequentialize expressions in the @let@ case.
 
@@
 interpreters, and the subsequent statements and expressions after the call-site become 
-the delimited continuation @f@. We have two invocations of @choices@, which correspond 
-to the two invocations of @nd@ in the refunctionalized form in the last section.
+the delimited continuation @f@. We have two applications of @choices@, which correspond 
+to the two applications of @nd@ in the refunctionalized form in the last section.
 The function @choices@ also returns the latest cache to its left-hand side when invoked.
-The cache will be still accumulated to the next call to @choices@ or @aeval@, 
+The cache will be still accumulated to the following calls of @choices@ or @aeval@, 
 given the fact that the abstract interpreter should always use the latest cache.
@@
 The function @analyze@ is also changed by calling the @reset@ operator around the function 
-call of @aeval@ to set the delimiter.
+application of @aeval@ to set the delimiter.
 

### 2018-07-05 src/main/scala/refunc/DirectStyle.scala
add traces eq test

@@ -18,8 +18,11 @@ object DirectStyleDC {
   def choices[T](ts: Iterable[T], cache: Cache): (T, Cache) @cps[Ans] = shift { 
     f: (((T, Cache)) => Ans) => nd(ts, Ans(Set[VS](), cache), f)
   }
-  
+
+  var trace: List[Expr] = List()
+
   def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans @cps[Ans] = {
+    trace = e::trace
     val config = Config(e, env, store, time)
     if (cache.outContains(config)) Ans(cache.outGet(config), cache)
     else {
@@ -65,6 +68,7 @@ object DirectStyleDC {
 
   def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore) = {
     def iter(cache: Cache): Ans = {
+      trace = List()
       val Ans(vss, anscache) = reset { aeval(e, env, store, mtTime, cache) }
       assert(anscache.outContains(Config(e, env, store, mtTime)))
       if (anscache.out == anscache.in) { Ans(vss, anscache) }
@@ -78,7 +82,9 @@ object DirectStyleSideEff {
   /* Using side effect to update new_cache */
   import RefuncCPS._
 
+  var trace: List[Expr] = List()
   def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans = {
+    trace = e::trace
     val config = Config(e, env, store, time)
     if (cache.outContains(config)) {
       return Ans(cache.outGet(config), cache)
@@ -132,6 +138,7 @@ object DirectStyleSideEff {
 
   def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore) = {
     def iter(cache: Cache): Ans = {
+      trace = List()
       val Ans(vss, anscache) = aeval(e, env, store, mtTime, cache)
       assert(anscache.outContains(Config(e, env, store, mtTime)))
       if (anscache.out == anscache.in) { Ans(vss, anscache) }

### 2018-07-05 src/main/scala/refunc/DisentangledLinearSmallStepUB.scala

@@ -17,7 +17,10 @@ object DisLinearSmallStepUBStack {
   def inject(e: Expr, env: Env = Map(), bstore: Store[BAddr, Storable] = Store[BAddr, Storable](Map())): NDState =
     NDState(e, env, bstore, List(), List(), List())
 
+  var trace: List[Expr] = List()
+
   def drive_step(nds: NDState, seen: Set[State]): Set[State] = {
+    trace = nds.e::trace
     val s = nds.toState
     val new_time = nds.tick
     val new_seen = if (seen.contains(s)) seen else seen + s
@@ -83,7 +86,13 @@ object DisLinearSmallStepUBStack {
     }
   }
 
-  def analyze(e: Expr): Set[State] = drive_step(inject(e), Set())
+  def analyze(e: Expr): Set[State] = {
+    trace = List()
+    drive_step(inject(e), Set())
+  }
 
-  def analyze(e: Expr, env: Env, bstore: BStore): Set[State] = drive_step(inject(e, env, bstore), Set())
+  def analyze(e: Expr, env: Env, bstore: BStore): Set[State] = {
+    trace = List()
+    drive_step(inject(e, env, bstore), Set())
+  }
 }

### 2018-07-05 src/main/scala/refunc/LinearSmallStepUB.scala

@@ -17,7 +17,10 @@ object LinearSmallStepUBStack {
   def inject(e: Expr, env: Env = Map(), bstore: Store[BAddr, Storable] = Store[BAddr, Storable](Map())): NDState =
     NDState(e, env, bstore, List(), List(), List())
 
+  var trace: List[Expr] = List()
+
   def step(nds: NDState): Option[NDState] = {
+    trace = nds.e::trace
     val new_time = nds.tick
     nds match {
       case NDState(Let(x, ae, e), env, bstore, konts, time, ndk) if isAtomic(ae) =>
@@ -78,8 +81,14 @@ object LinearSmallStepUBStack {
     }
   }
 
-  def analyze(e: Expr): Set[State] = drive(inject(e), Set())
+  def analyze(e: Expr): Set[State] = {
+    trace = List()
+    drive(inject(e), Set())
+  }
 
-  def analyze(e: Expr, env: Env, bstore: BStore): Set[State] = drive(inject(e, env, bstore), Set())
+  def analyze(e: Expr, env: Env, bstore: BStore): Set[State] = {
+    trace = List()
+    drive(inject(e, env, bstore), Set())
+  }
 }
 

### 2018-07-05 src/main/scala/refunc/RefuncCPS.scala

@@ -12,7 +12,10 @@ object RefuncCPS {
     else nd(ts.tail, acc ++ k(ts.head, acc.cache), k)
   }
   
+  var trace: List[Expr] = List()
+
   def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, continue: Cont): Ans = {
+    trace = e::trace
     val config = Config(e, env, store, time)
     if (cache.outContains(config)) continue(Ans(cache.outGet(config), cache))
     else {
@@ -67,6 +70,7 @@ object RefuncCPS {
 
   def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore) = {
     def iter(cache: Cache): Ans = {
+      trace = List()
       val Ans(vss, anscache) = aeval(e, env, store, mtTime, cache, ans => ans)
       assert(anscache.outContains(Config(e, env, store, mtTime)))
       if (anscache.out == anscache.in) { Ans(vss, anscache) }

### 2018-07-05 src/main/scala/refunc/RefuncECPS.scala

@@ -14,7 +14,10 @@ object RefuncECPS {
   type Cont = (Config, Set[Config], MCont) => Set[Config]
   type MCont = (Config, Set[Config]) => Set[Config]
 
+  var trace: List[Expr] = List()
+
   def aeval(config: Config, seen: Set[Config], continue: Cont, mcontinue: MCont): Set[Config] = {
+    trace = config.e::trace
     val Config(e, env, store, time) = config
     val new_time = config.tick
     val new_seen = if (seen.contains(config)) seen else seen + config
@@ -63,9 +66,13 @@ object RefuncECPS {
     }
   }
 
-  def analyze(e: Expr): Set[Config] = 
+  def analyze(e: Expr): Set[Config] = {
+    trace = List()
     aeval(inject(e), Set(), { (c, seen, m) => m(c, seen) }, { (c, seen) => seen })
+  }
 
-  def analyze(e: Expr, env: Env, bstore: BStore): Set[Config] = 
+  def analyze(e: Expr, env: Env, bstore: BStore): Set[Config] = {
+    trace = List()
     aeval(inject(e, env, bstore), Set(), { (c, seen, m) => m(c, seen) }, { (c, seen) => seen })
+  }
 }

### 2018-07-05 src/test/scala/refunc/Test.scala

@@ -8,11 +8,11 @@ import ANFAAM._
  * Description:
  * 
  * class RefuncTest provides some auxiliary functions;
- * class TDProgTest contains test cases that programs are terminated and deterministic;
- * class TNDProgTest contains test cases that programs are terminated but non-deterministic;
- * class NDProgTest contains test cases that programs are non-terminated;
+ * class TDProgTest contains test cases that programs are terminating and deterministic;
+ * class TNDProgTest contains test cases that programs are terminating but non-deterministic;
+ * class NDProgTest contains test cases that programs are non-terminating;
  * class NTTNDProgTest contains test cases that programs are non-deterministic, but some branches
- * are terminated, while some other branches are not.
+ * are terminating, while some other branches are not.
  *
  * For convenience, we test non-deterministic cases through passing an initial store that maps
  * an address to multiple target closures.
@@ -54,6 +54,7 @@ trait RefuncTest extends FunSuite {
          * disentangled counterpart (DisLinearSmallStepUBStack). */
         assert(FusedLinearSmallStepUBStack.analyze(prog, initenv, initstore) ==
                DisLinearSmallStepUBStack.analyze(prog, initenv, initstore))
+        //assert(LinearSmallStepUBStack.trace == DisLinearSmallStepUBStack.trace)
         
         /* Since the disentangled linearized small-step AAM with unbounded stack 
          * (DisLinearSmallStepUBStack) returns a set of state, while the refunctionalized 
@@ -69,21 +70,25 @@ trait RefuncTest extends FunSuite {
          */
         assert(summarizeUBState(DisLinearSmallStepUBStack.analyze(prog, initenv, initstore)) ==
                summarizeVSS(RefuncCPS.analyze(prog, initenv, initstore).vss))
+        assert(DisLinearSmallStepUBStack.trace == RefuncCPS.trace)
 
         assert(summarizeUBState(DisLinearSmallStepUBStack.analyze(prog, initenv, initstore)) ==
                summarizeConfig(RefuncECPS.analyze(prog, initenv, initstore)))
+        assert(DisLinearSmallStepUBStack.trace == RefuncECPS.trace)
       
         /* The result analyzed by the refunctionalized abstract interpreter written in CPS
          * (RefuncCPS) should have the same result analyzed by the direct-style abstract 
          * interpreter using delimited continuations (DirectStyleDC). */
         assert(RefuncCPS.analyze(prog, initenv, initstore) ==
[…]

### 2018-07-05 paper/paper.tex
typos

@@
   \item In the second case, an atomic expression @ae@ sits on the control position of the
-    state. Here, the value of @ae@ is being returned to its caller.
+    state. Here, the values of @ae@ is being returned to its caller.
     In order to accomplish this, we dereference the continuation address @kaddr@ and obtain 
@@
 a function @k@ that works on elements of type @T@, @nd@ chooses an element in
-the collection and applies @k@ to it iteratively, and accumulates the value that
+the collection and applies @k@ to it iteratively, and accumulates the values that
 returned from @k@ when recurring on the rest of the collection. Eventually, @nd@

### 2018-07-05 paper/paper.tex
misc revision

@@
 
-Then to effectively construct the meta-continuation @new_mcont@, we define an auxiliary function
-@makeMCont@, which takes a list of closures as arguments and returns the meta-continuation. 
+Then to address the circularity of the meta-continuations, we define an auxiliary function
+@makeMCont@ to construct @new_mcont@, which takes a list of closures as the argument and returns 
+the meta-continuation. 
 Function @makeMCont@ is a recursive function where the recursive invocation happens when 

### 2018-07-06 paper/paper.tex
AAM -> A2M where appropriate

@@
   sound static analyses for higher-order languages, by deriving small-step \emph{abstract} 
-  abstract machines (AAM) that perform abstract interpretation from abstract machines that 
+  abstract machines (A\textsuperscript{2}Ms) that perform abstract interpretation from abstract machines that 
   perform concrete evaluation.
   \citeauthor{darais2017abstracting} apply the same underlying idea to monadic definitional interpreters, 
-  and obtain monadic \emph{abstract} definitional interpreters (ADI) that perform
+  and obtain monadic \emph{abstract} definitional interpreters (ADIs) that perform
   abstract interpretation in big-step style using monads.
@@
   into higher-order functions. 
-  The refunctionalized AAM is an abstract interpreter written in continuation-passing style (CPS)
+  The refunctionalized A\textsuperscript{2}M is an abstract interpreter written in continuation-passing style (CPS)
   with two layers of continuations, which can be converted back to direct style with delimited 
@@
 
-  All transformations properly handle the collecting semantics and the nondeterminism 
+  All transformations properly handle the collecting semantics and nondeterminism
   of abstract interpretation. Remarkably, we reveal how precise call/return matching in
@@
 Applying the same idea to monadic definitional interpreters, \citet{darais2017abstracting}
-built monadic abstract definitional interpreters (ADI) in big-step style. 
+built monadic abstract definitional interpreters (ADIs) in big-step style. 
 One of the advantages of a monadic interpreter is
@@
 different forms of abstract interpreters. They are obtained by applying a combination
-of abstractions to their concrete counterparts, abstract machines and definitional
+of abstractions to their concrete counterparts: abstract machines and definitional
 interpreters, respectively. An interesting question, and the subject of this paper, is
@@
 understanding of abstract semantic artifacts, and it also opens up further practical 
-avenues for constructing static analysis tools based on principled and well-understand
+avenues for constructing static analysis tools based on principled and well-understood
 techniques.
 
-In the concrete world, the relations among reduction semantics, abstract machines,
+In the concrete world, the relationships between reduction semantics, abstract machines,
 definitional interpreters, and monadic interpreters have been intensively studied by
@@
 
-In the domain of abstract semantic artifacts, by contrast, the relations between small-step
+In the domain of abstract semantic artifacts, by contrast, the relationship between small-step
 abstract abstract machines and big-step abstract definitional interpreters, as well as the
@@
 In addition, those abstract abstract machines with an unbounded stack naturally correspond
-to abstract definitional interpreters. We show that refunctionalizing an AAM with an
+to abstract definitional interpreters. We show that refunctionalizing an A\textsuperscript{2}M with an
 unbounded stack (and utilizing the proper caching algorithm) leads to a pushdown control-flow analysis.
@@
  \end{tikzpicture}
- \caption{Functional correspondence between and transformations from AAM to ADI} \label{fig:trans}
+ \caption{Functional correspondence between and transformations from A\textsuperscript{2}Ms to ADIs} \label{fig:trans}
 \end{center}
@@
 We then address the main contribution of this paper: bridging the gap between
-small-step AAM and big-step ADI by applying a series of well-known systematic transformations
+small-step A\textsuperscript{2}Ms and big-step ADIs by applying a series of well-known systematic transformations
 drawn from concrete semantic artifacts.
@@
     nondeterministic choices as a first-order data type, we linearize the execution
-    of abstract abstract machines; the transition of machine states therefore becomes
+    of abstract abstract machines; the transition of machine states thus becomes
     deterministic. Notably, this makes explicit another layer of control (Section~\ref{linear}). 
@@
   \item \textbf{Lightweight fusion}~\cite{DANVY2008100,Ohori:2007:LFF:1190216.1190241}:
-    We apply a fusion transformation to the linearized AAM, which
+    We apply a fusion transformation to the linearized A\textsuperscript{2}M, which
     combines the single-step function @step@ and the 
@@
   \item \textbf{Refunctionalization}~\cite{DANVY2009534,Danvy:2006:RW:2171265.2171268}:
-    We apply refunctionalization to the disentangled AAM, which
+    We apply refunctionalization to the disentangled A\textsuperscript{2}M, which
     %as the left inverse of Reynolds' defunctionalization \cite{Reynold:72,Reynolds:HOSC98-revisited} 
@@
 
-    For clarity, we first present a vanilla version of refunctionalized AAM which simply
+    For clarity, we first present a vanilla version of a refunctionalized A\textsuperscript{2}M which simply
     converts the stack structures to higher-order functions but keeps other parts unchanged.
@@
 
-  \item \textbf{Direct-style transform}~\cite{DBLP:journals/scp/Danvy94}:
-    We finally transform the refunctionalized AAM back to a
+  \item \textbf{Direct-style transformation}~\cite{DBLP:journals/scp/Danvy94}:
+    We finally transform the refunctionalized A\textsuperscript{2}M back to a
     direct-style interpreter by using delimited control operators
@@
 programs.
-Since the state space of concrete execution is possibly infinite, the key insight of 
+Since the state space of concrete execution is possibly infinite, the key insight of the
 AAM approach when analyzing programs is to allocate both variable bindings and continuations 
@@
 
-The pushdown AAM with an unbounded stack is also described by \citeauthor{van2012systematic} 
+The pushdown A\textsuperscript{2}M with an unbounded stack is also described by \citeauthor{van2012systematic} 
 \cite{van2012systematic}. 
@@
 a precise control-flow from its defining language, we also would like to start from a
-pushdown AAM that is also able to precisely match calls and returns.
+pushdown A\textsuperscript{2}M that is also able to precisely match calls and returns.
 For readers who are not familiar with pushdown analysis, we have a detailed discussion 
@@
 \textit{Note on Termination.}
-Unfortunately, even though other components in the states are finite, this AAM with an unbounded
+Unfortunately, even though other components in the states are finite, this A\textsuperscript{2}M with an unbounded
 stack still may diverge when analyzing some programs.
@@
 space, we can recover the precise call/return match. Now we begin describing the
-transformations step by step. Our base machine is the AAM with an unbounded
+transformations step by step. Our base machine is the A\textsuperscript{2}M with an unbounded
 stack, as we will later transform the stack structure to higher-order functions
-representing continuations, it does not matter what kind of AAM we start from.
+representing continuations, it does not matter what kind of A\textsuperscript{2}M we start from.
 This is because it would not be possible to construct a new frame on the stack, nor
 to allocate continuations in the store after Refunctionalization.
-Thus, our choice to start from an unbounded-stack AAM is motivated simply because
+Thus, our choice to start from an unbounded-stack A\textsuperscript{2}M is motivated simply because
 it has an equivalent stack model to abstract definitional interpreters (our final target).
@@
 Intuitively, if we imagine that the classical AAM explores a graph of reachable states, 
-then the linearized AAM flattens the graph in depth-first order to a linear sequence of states.
+then the linearized A\textsuperscript{2}M flattens the graph in depth-first order to a linear sequence of states.
 It has been observed that ``\textit{failure can be replaced by a list of successes}''
@@
 
-With fusing complete, our AAM appears similar to a ``big-step'' interpreter, although it
+With fusing complete, our A\textsuperscript{2}M appears similar to a ``big-step'' interpreter, although it
 still has the machine state representation inside. The inverse transformation of this step
@@
 Before jumping into the details of refunctionalizing the stack structure and constructing
-@new_cont@ and @new_mcont@, we have two important observations in the disentangled AAM
+@new_cont@ and @new_mcont@, we have two important observations in the disentangled A\textsuperscript{2}M
 where the continuations were represented by data types:
@@
 
-In this section, based on the refunctionalized AAM, we further present several other 
+In this section, based on the refunctionalized A\textsuperscript{2}M, we further present several other 
 transformations toward a big-step abstract definitional interpreter:
@@
     function.
-  \item The refunctionalized AAM collects all possible configurations, but our final target is an 
+  \item The refunctionalized A\textsuperscript{2}M collects all possible configurations, but our final target is an 
     abstract definitional interpreter that cares about final evaluation results.
@@
 
-We observe that in the refunctionalized AAM, @makeMCont@ is invoked with the tail part of
+We observe that in the refunctionalized A\textsuperscript{2}M, @makeMCont@ is invoked with the tail part of
 the closure set, and it actually repeats the code that creates the new environment and 
@@
 \textit{Call/Return Matching through Refunctionalization and Caching.}
-The refunctionalized AAM we obtained has the perfect match on return-flows even though
+The refunctionalized A\textsuperscript{2}M we obtained has the perfect match on return-flows even though
 we do not have a stack model. The higher-order functions representing continuations
@@
 
-In fact, what we started with is an unbounded-stack AAM which already precisely
+In fact, what we started with is an unbounded-stack A\textsuperscript{2}M which already precisely
 matches the calls and returns. Refunctionalization makes the interpreter a ``big-step''
@@
 
-In the previous section, we presented a refunctionalized AAM in extended continuation-passing style,
+In the previous section, we presented a refunctionalized A\textsuperscript{2}M in extended continuation-passing style,
 and then simplified it by utilizing the @nd@ operator.
@@
 \Citet{gilray2016pushdown} further proposed a polyvariant continuation-addresses allocator
-for small-step AAM to achieve pushdown analysis. This method is both
+for small-step A\textsuperscript{2}M to achieve pushdown analysis. This method is both
 simple to implement and computationally inexpensive and so called Pushdown for Free (P4F).
-Based on the AAM we presented in Section \ref{aam}, the only changes in the code required is not
+Based on the A\textsuperscript{2}M we presented in Section \ref{aam}, the only changes in the code required is not
 only keeping track of the entry-point expression of the callee, but also holding the target
@@
 instead of plain $\lambda$-calculus, although our work could be easily adapted to handle plain 
-$\lambda$-calculus as well. Going through the transformation steps from AAMs to ADIs instead of 
+$\lambda$-calculus as well. Going through the transformation steps from A\textsuperscript{2}Ms to ADIs instead of 
 only considering the final ADI result provides additional insights. In particular, it reveals that 

### 2018-07-06 paper/paper.tex
Proofreading

@@
 
-\Citet{van2012systematic,van2010abstracting} proposed the Abstracting Abstract Machines (AAM)
+\Citet{van2012systematic,van2010abstracting} proposed the Abstracting abstract machines (AAM)
 methodology which provides a recipe for constructing sound abstract interpreters
@@
 The CESK machine is an abstract machine for describing the semantics of and evaluating
-$\lambda$-calculus \cite{felleisen1987calculus}. The CESK machine models program execution
+a $\lambda$-calculus \cite{felleisen1987calculus}. The CESK machine models program execution
 as state transitions in a small-step fashion. As its name suggests, a machine state has
@@
 these two cases and evaluates atomic expressions to closures in a straightforward way.
-The @alloc@ function generates a fresh address, and always allocates an unique integer
+The @alloc@ function generates a fresh address, and always allocates a unique integer
 in the domain of @store@.
@@
 which must have a successor; the final case of a state is handled by function
-@drive@, which will be mentioned in Section~\ref{sec:drive}.
+@drive@, which will be explained in Section~\ref{sec:drive}.
 
@@
 As shown and previously discussed, we examine the only two non-final cases which the state
-may be.
+may be:
 
@@
 
-In this section, we derive the abstract abstract machine from the concrete
+In this section, we derive the respective abstract abstract machine from the concrete
 CESK machine, and also show how to instantiate useful $k$-call-sensitive 
@@
 
-Similar to the CESK machine, the machine state of AAM has a control, an environment, a store,
+Similar to the CESK machine, the machine state of an A\textsuperscript{2}M has a control, an environment, a store,
 and continuation, as well as a timestamp. However, there are several notable differences
-between AAM's store and CESK machine's store. In AAM, the store maps addresses to sets of
+between an A\textsuperscript{2}M's store and the CESK machine's store. In AAM, the store maps addresses to sets of
 values; it stores all possible values for a particular address. As such, dereferencing
@@
 
-Similar to the CESK machine, to run (analyze) a program we first use the @inject@ function
+Similar to the CESK machine, to run (analyze) a program, we first use the @inject@ function
 to construct the initial state given to the program. Note that the initial continuation store
@@
 However, in contrast to the concrete CESK machine, the @drive@ function performs 
-collecting semantics instead of the valuation semantics. That is, for the purpose of
+collecting semantics instead of valuation semantics. That is, for the purpose of
 analyzing programs, the function @drive@ collects all the intermediate machine states 
@@
 
-In this section, we describe a variant of AAM that allows the stack to be unbounded 
-which uses a precise model of the call stack as we did in the concrete CESK machine.
+In this section, we describe a variant of AAM that allows the stack to be unbounded,
+which uses a precise model of the call stack as in the concrete CESK machine.
 Instead of allocating continuations in the store and embedding addresses of 
@@
 
-The pushdown A\textsuperscript{2}M with an unbounded stack is also described by \citeauthor{van2012systematic} 
+The pushdown AAM with an unbounded stack is also described by \citeauthor{van2012systematic} 
 \cite{van2012systematic}. 
-The reason we show it here is to keep the consistency of abstract semantics during 
+We show it here is to maintain consistency of abstract semantics during
 transformations -- since the eventual abstract definitional interpreters simply inherit
-a precise control-flow from its defining language, we also would like to start from a
+a precise control-flow from their defining language, we also would like to start from a
 pushdown A\textsuperscript{2}M that is also able to precisely match calls and returns.
-For readers who are not familiar with pushdown analysis, we have a detailed discussion 
+For readers who are not familiar with pushdown analysis, we provide a detailed discussion
 in Section~\ref{pdcfarevisit}.
@@
 
-The state transition function @step@ shown below is still nondeterministic, but the only 
-nondeterminism happening is when dereferencing the callee @f@ from the function application 
-@App(f, ae)@.
+The state transition function @step@ shown below is still
+nondeterministic, but only when dereferencing the callee @f@ from the
+function application @App(f, ae)@.
 
@@
 \textit{Note on Termination.}
-Unfortunately, even though other components in the states are finite, this A\textsuperscript{2}M with an unbounded
-stack still may diverge when analyzing some programs.
+Unfortunately, even though other components in the states are finite, this version of AAM with an unbounded
+stack may still diverge when analyzing some programs.
 This is because the unbounded stack can grow to an arbitrary depth which implies that the state
@@
 
-Function @f1@ and @f2@ mutually invoke each other, so the stack will alternately pushing
+Function @f1@ and @f2@ mutually invoke each other, so the stack will alternately push
 frames @f1@ and @f2@ onto the top of the current stack. However, no two existing stack
@@
 transformations step by step. Our base machine is the A\textsuperscript{2}M with an unbounded
-stack, as we will later transform the stack structure to higher-order functions
+stack; as we will later transform the stack structure to higher-order functions
 representing continuations, it does not matter what kind of A\textsuperscript{2}M we start from.
 This is because it would not be possible to construct a new frame on the stack, nor
-to allocate continuations in the store after Refunctionalization.
+to allocate continuations in the store after refunctionalization.
 Thus, our choice to start from an unbounded-stack A\textsuperscript{2}M is motivated simply because
@@
 a structural operational semantics in continuation-passing style, where the reduction context is
-a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}. Therefore
+a defunctionalized continuation \cite{Danvy:2008:DIP:1411204.1411206}. Therefore,
 refunctionalizing such a reduction semantics for a deterministic language yields 
@@
 
-We apply this idea to AAM and our first step of transformations is to linearize all
+We apply this idea to AAM, and thus our first transformation step is to linearize all
 nondeterministic choices. This step removes all nondeterminism from the @step@ function.
@@
 is added to convert itself to type @State@.  
-Function @step@ now becomes of type @NDState => Option[NDState]@. Returning a @None@ value means
+@step@ now becomes of type @NDState => Option[NDState]@. Returning a @None@ value means
 reaching the halt.
@@
 
-If the continuation @konts@ is empty, then the machine has reached an end of the computation;
-otherwise, we should return the values of @ae@ to its caller which is contained in the top
+If the continuation @konts@ is empty, then the machine has reached an end of the computation.
+Otherwise, we should return the values of @ae@ to its caller which is contained in the top
 frame of the stack. However, since a meta-continuation @ndk@ is introduced, we need to dispatch
@@
   \item If the meta-continuation @ndk@ is also empty, then the machine has reached the end of \textit{all} 
-    computations. Besides, the control component @ae@ now is an atomic value, therefore @None@ is returned.
+    computations. Besides, the control component @ae@ now is an atomic value: @None@ is returned.
   \item If the target closures set of the front-most @NDCont@ is empty (@Nil@ in the code),
@@
     point. Since the meta-continuation @nkd@ is represented by a list, and we always append new 
-    elements to its front, we resume to the most recent fork point by popping up the front-most 
+    elements to its front, we resume to the most recent fork point by popping the front-most 
     element and using the tail of that list.
@@
     are copied from the fork point rather than current one.
-    The meta-continuation @new_ndk@ is updated by the removal of that popped closure, which makes
-    sure the closure set will eventually shrink to empty.
+    The meta-continuation @new_ndk@ is updated by the removal of that popped closure, which guarantees
+    the closure set will eventually shrink to empty.
 \end{itemize}
@@
 The termination of the analysis occurs when the value @None@ is returned from calling @step@.
-Otherwise, we extract the next state @nnds@ and recur on function @drive@.
+Otherwise, we extract the next state @nnds@ and recurse on function @drive@.
 The set of explored states is preserved as before.
@@
 Intuitively, if we imagine that the classical AAM explores a graph of reachable states, 
-then the linearized A\textsuperscript{2}M flattens the graph in depth-first order to a linear sequence of states.
+then the linearized version of AAM flattens the graph in depth-first order to a linear sequence of states.
 It has been observed that ``\textit{failure can be replaced by a list of successes}''
@@
 
-Then we apply a lightweight fusion transformation that combines the @step@ and @drive@
+We next apply a lightweight fusion transformation that combines the @step@ and @drive@
 functions into a single function.
@@
 It is essentially formed by merging the functionality of @step@ into the @drive@ function.
-Since each case in function @step@ is statically known, we can inline the pattern match on the
-@Option@ type and replace it with returning @new_seen@ when @None@ is matched, or calling 
+Since each case in function @step@ is statically known, we can inline the pattern matching on the
+@Option@ type and replace it with returning @new_seen@ when @None@ is matched, or by calling 
 @drive_step@ recursively when @Some@ is matched.
@@
 @ndcontinue@.
-It matches on the list representation of context: An empty list @Nil@ represents the
+It matches on the list representation of context: an empty list @Nil@ represents the
 halt of a computation path, whereas a cons @::@ means the top frame of the list
@@
 
-Function @ndcontinue@ is identical to the dispatching logic in function @continue@ when @konts@ is empty.
-When @ndk@ is empty, the current explored states @seen@ is returned; the analysis then terminates.
+Function @ndcontinue@ is identical to the dispatching logic in @continue@ when @konts@ is empty.
+When @ndk@ is empty, the current set of explored states @seen@ is returned, after which the analysis terminates.
 Otherwise, a new @NDState@ is constructed depending on the status of the closures set.
@@
 
-The purpose of disentanglement is to reveal the defunctionalized form of AAM.
-Functions @continue@ and @ndcontinue@ that dispatches different data types representing 
+The purpose of disentanglement is to reveal the defunctionalized form of our A\textsuperscript{2}M.
+Functions @continue@ and @ndcontinue@ that dispatch different data types representing 
 continuations are disentangled from @drive_step@. 
@@
 After this transformation, we will obtain an abstract interpreter written in extended
-continuation-passing style\cite{Danvy:1990:AC:91556.91622}.
+continuation-passing style~\cite{Danvy:1990:AC:91556.91622}.
 This transformation can be also regarded as making the control flows of the defined 
@@
 In this section, we first present a direct refunctionalization of the continuations.
-Next, we adopt a caching algorithm to obtain a computable analysis\cite{darais2017abstracting},
-and further prepare the interpreter for transforming to direct-style later.
+Next, we adopt a caching algorithm to obtain a computable analysis~\cite{darais2017abstracting},
+and further prepare the interpreter for transformation to direct-style at a later point.
 
@@
 This observation leads us to define the types of higher-order functions representing
-continuation and meta-continuation:
+continuations and meta-continuations:
 
@@
 
-The continuation @Cont@ as a higher-order function takes a configuration, a set of explored
-configurations, as well as a meta-continuation as its arguments. While the meta-continuation
+The continuation @Cont@ is a higher-order function which takes a configuration, a set of explored
+configurations, as well as a meta-continuation as its arguments, while the meta-continuation
 @MCont@ only takes a configuration and a set of explored configurations as arguments.
@@
 decreasing every time while the other parts of meta-continuation are kept. This means in 
-the higher-order version, we also need an ``ability'' to update the top frame of meta-continuation
+the higher-order version, we also need an ability to update the top frame of meta-continuation
 and keep the rest as the same.
@@
 Let us first look into @new_cont@. The body of @new_cont@ is similar to the first-order dispatching
-function @continue@. The invocation of @aeval@ uses the latest meta-continuation @m@ instead of the outer
-one, @mcontinue@. Note that the new environment @new_env@ is updated based on the outer environment;
-while the atomic evaluation uses the latest environment @env_@.
+function @continue@. The invocation of @aeval@ uses the latest
+meta-continuation @m@ instead of the outer @mcontinue@. Note that the
+new environment @new_env@ is updated based on the outer environment;
+the atomic evaluation uses the latest environment @env_@.
 
@@
 
-Then to address the circularity of the meta-continuations, we define an auxiliary function
-@makeMCont@ to construct @new_mcont@, which takes a list of closures as the argument and returns 
-the meta-continuation. 
+To address the circularity of the meta-continuations, we define an auxiliary function
+@makeMCont@ to construct @new_mcont@, which takes a list of closures as the argument and returns the meta-continuation.
 Function @makeMCont@ is a recursive function where the recursive invocation happens when 
@@
 the outer meta-continuation @mcontinue@ is returned.
-It would be clear if we inductively think the base case where the closures set contains only one 
-closure, then @makeMCont@ is invoked with an empty set, therefore should be just the 
+This becomes clear thinking inductively; our base case is such that the closures set contains only one 
+closure. @makeMCont@ is thus invoked with an empty set, and should therefore be just the
 current meta-continuation @mcontinue@.
@@
 
-One last very important thing is about the initial continuation and meta-continuation.
+We present one final, though very important, note regarding the initial continuation and meta-continuation.
 The initial continuation will be invoked when one computation path reaches its end,
@@
 
-Now we have completely refunctionalized an abstract abstract machine; the two layers
+We have now completely refunctionalized an abstract abstract machine; the two layers
 of continuations are represented by higher-order functions. The inverse transformation
-of this step is known as \textit{de}functionalization, that converts the higher-order
-representing continuations back to first-order data types representation 
-\cite{Danvy:2001:DW:773184.773202, Danvy:2008:DIP:1411204.1411206}.
+of this step is known as \textit{de}functionalization, which converts the higher-order functions
+representing continuations back to first-order data
+types~\cite{Danvy:2001:DW:773184.773202, Danvy:2008:DIP:1411204.1411206}.
 
@@
 
-In this section, based on the refunctionalized A\textsuperscript{2}M, we further present several other 
-transformations toward a big-step abstract definitional interpreter:
+In this section, based on the refunctionalized A\textsuperscript{2}M, we
+further present several other transformations toward a big-step abstract
+definitional interpreter:
 
@@
     function.
-  \item The refunctionalized A\textsuperscript{2}M collects all possible configurations, but our final target is an 
+  \item The refunctionalized AAM collects all possible configurations, but our final target is an 
     abstract definitional interpreter that cares about final evaluation results.
-    So starting from this section our abstract semantic artifact returns a set of final 
+    As such, starting from this section, our abstract semantic artifact returns a set of final 
     values instead of collected configurations. 
@@
 
-We will begin by discussing these changes separately, and finally see how they work together.
+We begin discussing these changes separately, and finally see how they work together.
 
@@
 if we have seen a state @s@, it means we also have seen all the successors of state @s@.
-The caching algorithm we adopt here replays the same assumption but in a big-step 
+The caching algorithm we adopt here replays the same assumption, but in a big-step 
 manner: if we have seen the configuration @c@, then it means we also have seen the 
@@
 @aeval@. Given the cache and the simplified one-layer continuation (as we will see later),
-the shape of function @aeval@ now looks like:
+the shape of function @aeval@ now looks as follows:
 
@@
 Upon entering @aeval@, we first determine whether the @out@ cache contains some values 
-for the current configuration and use them immediately (if present) by calling @continue@.
+for the current configuration, and (if present) use them immediately by
+calling @continue@.
 This corresponds to small-step AAM with a work-list: when we have seen a state,
@@
 
-We observe that in the refunctionalized A\textsuperscript{2}M, @makeMCont@ is invoked with the tail part of
-the closure set, and it actually repeats the code that creates the new environment and 
+We observe that in the refunctionalized AAM, @makeMCont@ is invoked with the tail part of
+the closure set, and actually repeats the code that creates the new environment and
 store for the body expression of a closure.
-Based on observation, we present a nondeterministic operator that abstracts this pattern:
+Based on this observation, we present a nondeterministic operator that abstracts this pattern:
 
@@
 
-Given a collection of elements of type @T@, an initial accumulator of type @Ans@,
-a function @k@ that works on elements of type @T@, @nd@ chooses an element in
+Given a collection of elements of type @T@, an initial accumulator of
+type @Ans@, and a function @k@ that works on elements of type @T@, @nd@
+chooses an element in
 the collection and applies @k@ to it iteratively, and accumulates the values that
@@
 
-\subsubsection{Put Them Together}
+\subsubsection{Putting Them Together}
 
 With the skeleton of cache-and-continuation-passing style @aeval@ in mind, we may now
-use the @nd@ operator to fill in the details. The first case of match is that @e@
-is a @let@ expression:
+use the @nd@ operator to fill in the details. The first case of our
+match is that @e@ is a @let@ expression:
 
@@
 
-We invoke @nd@ twice and perform a nesting two-step, depth-first evaluation over 
+We invoke @nd@ twice and perform a nesting, two-step, depth-first evaluation over 
 possible values of @App(f, ae)@. 
@@
 @f@. For each such closure, we go into its body with a new environment and a new store.
-Remember that the function @aeval@ returns a set of @VS@ objects to its continuation,
+Recall that the function @aeval@ returns a set of @VS@ objects to its continuation,
 so the continuation argument @bdvss@ represents a set of values from multiple computation
@@
 
-We next consider the inner application of @nd@, which evaluates over the set of body's 
-value/stores @bdvss@. For each @VS@, the time-stamp and the store in @VS@ will be 
+We next consider the inner application of @nd@, which evaluates over the set of the function body's 
+values/stores @bdvss@. For each @VS@, the timestamp and the store in @VS@ will be 
 instantiated to the next application of @aeval@. The environment @new_env@ is built on
@@
 Finally, the @analyze@ function (as the entrance of the analysis) does a looping iteration
-to find the least fixed-point of cache starting from an empty cache.
-If the @out@ cache of this iteration is equivalent to its @in@ cache, which means no more 
-fact was discovered during this iteration, then we have reached a fixed-point of cache that 
+to find the least fixed-point of the cache, starting from an empty cache.
+If the @out@ cache of this iteration is equivalent to its @in@ cache, which means no further
+information was discovered during this iteration, then we have reached a fixed-point of cache that 
 has over-approximated the concrete evaluations, and thus the result can be returned.
 Otherwise, we install the latest @out@ cache to the @in@ cache position when initializing 
-the next iteration; in the meanwhile, an empty cache will be used as the @out@ cache.
+the next iteration; in the meantime, an empty cache will be used as the @out@ cache.
 
 Now, as we can see, by using the @nd@ operator which abstracts over the meta-continuation,
-the evaluation function can be simplified to one layer continuation-passing style.
-The use of two layers of continuations are unchained to two functions that each of them has one
-continuation.
-The only continuation of @aeval@ plays the same role as the one in a concrete CPS interpreter.
-Furthermore, we added a cache along with the abstract interpreter and the @nd@ operator.
+the evaluation function can be simplified to a single-layer continuation-passing style.
+The use of two layers of continuations are unchained to two functions,
+in that each of them has one continuation.
+The only continuation of @aeval@ plays the same role as the one in a
+concrete CPS interpreter.
+Furthermore, we add a cache along with the abstract interpreter and the @nd@ operator.
 

### 2018-07-06 paper/paper.tex
Finished proofreading

@@
 
-In the previous section, we have established a computable pushdown control-flow
+In the previous section, we established a computable pushdown control-flow
 analysis through refunctionalization and caching. In this section, we revisit
@@
 call structure of the analyzed program. A pushdown control-flow analysis provides an
-as-exact-as-run-time return-flow when analyzing a program, but traditional 
-control-flow analysis collapses the state space into a finite space and thus 
-causes imprecise stack modeling.
+as-exact-as-runtime return-flow when analyzing a program, but traditional
+control-flow analysis collapses the state space into a finite space and
+thus causes imprecise stack modeling.
 
@@
 
-In $k$-CFA algorithm or using the abstract abstract machine shown in Section~\ref{aam},
+In the $k$-CFA algorithm, or in using the abstract abstract machine shown in Section~\ref{aam},
 the call sites @(id 2)@ and @(id 1)@ share the same return flow, so the invocation
@@
 
-In fact, what we started with is an unbounded-stack A\textsuperscript{2}M which already precisely
+In fact, we started with an unbounded-stack A\textsuperscript{2}M which already precisely
 matches the calls and returns. Refunctionalization makes the interpreter a ``big-step''
@@
 information in the store or on the stack.
-But refunctionalization not only forces the nondeterministic control flow to be higher-order,
+However, refunctionalization not only forces the nondeterministic control flow to be higher-order,
 but also drives us to use a different caching algorithm.
 
-Indeed, the caching algorithm plays an important role to guarantee the analysis will
-terminate eventually.
+Indeed, the caching algorithm plays an important role in guaranteeing
+the analysis will eventually terminate.
 An interesting implication of this is if we apply the caching algorithm with some
@@
     assignments and mutations to update the cache, and then achieve the same definitional
-    interpreter with pushdown control-flows. \Citet{friedman-ai} in a tutorial on abstract 
-    interpreters use this style to model caches and updates.
+    interpreter with pushdown control-flows. \Citet{friedman-ai}, in a tutorial on abstract
+    interpreters, use this style to model caches and updates.
     In this case, nondeterminism can be easily handled via @for@ comprehensions.
@@
 
-These coincidences should not be a surprise since the literature is rich
+These coincidences should not be a surprise, since the literature is rich
 in showing the correspondence between monads and continuation-passing style as well as
@@
 transform this function.
-Back to direct-style on the case where @e@ appears as an atomic expression is straightforward: 
-we simply unwrap the application of @continue@. Now the abstract evaluator @aeval@ looks as follows:
+Returning to direct-style in the case where @e@ appears as an atomic expression is straightforward:
+we simply unwrap the application of @continue@. The abstract evaluator @aeval@ now looks as follows:
 
@@
 
-The case where @e@ is a @let@ expression is not that obvious. As previously identified, 
-function @nd@ works on an unknown number of nondeterministic choices, and it takes a 
+The case where @e@ is a @let@ expression is less straightforward. As
+previously identified, @nd@ works on an unknown number of
+nondeterministic choices, and it takes a
 function @k@ that works for every element in the collection. Here we present another
-function @choices@ that uses delimited control operator @shift@ to capture that ``@k@''
+function @choices@ that uses delimited control operator @shift@ to capture that @k@
 when the program is written in direct-style and pass it to @nd@.
@@
 function is type @Ans@.
-The function @f@ is the delimited continuation captured by the @shift@ operator, which
-is formed by the computation will be executed after the @choices@'s call-site.
+The function @f@ is the delimited continuation captured by the @shift@ operator (formed by the computation), and will be executed after the @choices@'s call-site.
 Once the function @f@ (which is the function @k@ in @nd@) is invoked and returned,
-it returns to its call-site where @nd@ will be invoked on the remaining elements of the collection.
-Now we can use this function to sequentialize expressions in the @let@ case.
+it returns to its call site where @nd@ will be invoked on the remaining elements of the collection.
+Now we can use this function to sequentialize expressions in the @let@ case:
 
@@
 @choices@ on the set of possible choices and write the program sequentially as concrete
-interpreters, and the subsequent statements and expressions after the call-site become 
+interpreters, and the subsequent statements and expressions after the call site become
 the delimited continuation @f@. We have two applications of @choices@, which correspond
-to the two applications of @nd@ in the refunctionalized form in the last section.
-The function @choices@ also returns the latest cache to its left-hand side when invoked.
-The cache will be still accumulated to the following calls of @choices@ or @aeval@, 
+to the two applications of @nd@ in the refunctionalized form in the previous section.
+@choices@ also returns the latest cache to its left-hand side when invoked.
+The cache will continue to accumulate for the following calls of @choices@ or @aeval@,
 given the fact that the abstract interpreter should always use the latest cache.
 
-The function @analyze@ is also changed by calling the @reset@ operator around the function 
-application of @aeval@ to set the delimiter.
+@analyze@ is also changed by calling the @reset@ operator around the
+function application of @aeval@ to set the delimiter.
 
@@
 
-Now we have arrived at the end of this series of transformations. 
-The inverse transformation of the last step is basically CPS transforming the function @aeval@ and
+We have finally arrived at the end of this series of transformations.
+The inverse transformation of the last step is essentially that of performing a CPS transformation on the function @aeval@ and
 explicitly using function @nd@ instead of function @choices@. Then, these continuations must be
-manifestly embedded at the call-site of @nd@, instead of implicitly provided by the @shift@ operator.
+manifestly embedded at the call site of @nd@, instead of implicitly provided by the @shift@ operator.
 
@@
 and big-step abstract definitional interpreters by applying a series of syntactic
-transformations to transform the first into the latter. Among these transformations, 
-linearization turns a work-list into an additional layer of continuations; refunctionalization 
-converts the first-order data types representing continuations to higher-order functions; 
-and finally the left-inverse of the CPS transformation converts the CPS abstract interpreter 
+transformations to transform the former into the latter. Among these transformations,
+linearization turns a work-list into an additional layer of continuations, refunctionalization
+converts the first-order data types representing continuations to higher-order functions,
+and finally, the left-inverse of the CPS transformation converts the CPS abstract interpreter
 into a direct-style abstract interpreter with delimited control operators, which looks almost
@@
 This sequence of transformation demonstrates that a functional correspondence exists
-not only between concrete semantic artifacts but also between abstract semantic artifacts. 
-An interesting open question would be whether there also exist correspondences between 
-static analyses that are formalized 
-in different denotational and operational styles.
+not only between concrete semantic artifacts, but also between abstract semantic artifacts.
+An interesting open question remains of whether there also exist
+correspondences between static analyses that are formalized in different
+denotational and operational styles.
 

### 2018-07-06 paper/paper.tex
misc changes

@@
 
-Function @f1@ and @f2@ mutually invoke each other, so the stack will alternately push
-frames @f1@ and @f2@ onto the top of the current stack. However, no two existing stack
+Function @f1@ and @f2@ mutually invoke each other, so the frames of @f1@ and @f@ will be
+alternately pushed onto the top of the current stack. However, no two existing stack
 components are identical in the state space. One can decide the reachability of a state
@@
 function is type @Ans@.
-The function @f@ is the delimited continuation captured by the @shift@ operator (formed by the computation), and will be executed after the @choices@'s call-site.
+The function @f@ is the delimited continuation captured by the @shift@ operator.
+Whenever we call @choices@, the subsequent computations that will be executed after its call-site
+constitute the delimited continuation @f@.
 Once the function @f@ (which is the function @k@ in @nd@) is invoked and returned,

### 2018-07-07 paper/paper.tex
update code & revise bib

@@
 %% For final camera-ready submission, w/ required CCS and ACM Reference
-\documentclass[acmsmall, review]{acmart}\settopmatter{}
+\documentclass[acmsmall, screen]{acmart}\settopmatter{}
 
@@
 Given a concrete small-step abstract machine that models a store
-(e.g. the CESK machine, Krivine's machine, etc.), if we allocate continuations
+(e.g., the CESK machine, Krivine's machine, etc.), if we allocate continuations
 in the store and draw all allocated addresses (both for regular values and
@@
        [start branch=C] \chainin (m-3-5)
-       [join={node[above, labeled] {\text{Known as Danvy's correspondences for concrete semantic artifacts }}}];
+       [join={node[above, labeled] {\text{Functional correspondences for concrete semantic artifacts}}}];
      }
@@
 through Dyck state graph \cite{earl2010pushdown, earl2012introspective}, but simply
-caching the explored states (i.e. @seen@) would not always make the analyzer terminate.
+caching the explored states (i.e., @seen@) would not always make the analyzer terminate.
 
@@
         case NDCont(Nil, _, _, _, _)::ndk =>
-          Some(NDState(ae, env, bstore, konts, time, ndk)) /* transfer to the most recent fork point */
+          Some(NDState(ae, env, bstore, konts, time, ndk)) /* transfer to the front-most fork point */
         case NDCont(cls, argvs, bstore, time, frames)::ndk =>
@@
     point. Since the meta-continuation @nkd@ is represented by a list, and we always append new
-    elements to its front, we resume to the most recent fork point by popping the front-most
+    elements to its front, we resume to the next fork point by popping the front-most
     element and using the tail of that list.

### 2018-07-07 src/main/scala/refunc/RefuncCPS.scala

@@ -80,64 +80,63 @@ object RefuncCPS {
   }
 }
 
+object RefuncCPSNoCache {
+  type Ans = Set[Config]
+  type Cont = (Config, Ans) => Ans
 
+  def inject(e: Expr, env: Env = Map(), 
+             store: Store[BAddr, Storable] = Store[BAddr, Storable](Map())): Config =
+    Config(e, env, store, List())
 
-/* Experimental implementation with breadth first evaluation */
-object RefuncCPSBF {
-  type Cont = (Set[VS], Cache) => (Set[VS], Cache)
-
-  def nd[T](ts: Set[T], acc: Set[VS], cache: Cache, 
-            f: (T, Set[VS], Cache, Cont) => (Set[VS], Cache), g: Cont): (Set[VS], Cache) = {
-    if (ts.isEmpty) g(acc, cache)
-    else f(ts.head, acc, cache, (vss: Set[VS], cache: Cache) => nd(ts.tail, vss, cache, f, g))
+  def nd[T](ts: Iterable[T], acc: Ans, k: (T, Ans) => Ans): Ans = {
+    if (ts.isEmpty) acc
+    else nd(ts.tail, acc ++ k(ts.head, acc), k)
   }
-
-  def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, continue: Cont): (Set[VS], Cache) = {
-    val config = Config(e, env, store, time)
-    if (cache.outContains(config)) {
-      return continue(cache.outGet(config), cache)
-    }
-
-    val new_cache = cache.outUpdate(config, cache.inGet(config))
+  
+  def aeval(config: Config, seen: Ans, continue: Cont): Ans = {
+    trace = config.e::trace
+    val Config(e, env, store, time) = config
     val new_time = config.tick
+    val new_seen = if (seen.contains(config)) seen else seen + config
     e match {
+      case Let(x, ae, e) if isAtomic(ae) =>
+        val baddr = allocBind(x, new_time)
+        val new_env = env + (x -> baddr)
+        val new_store = store.update(baddr, atomicEval(ae, env, store))
+        aeval(Config(e, new_env, new_store, new_time), new_seen, continue)
+
+      case Letrec(bds, body) => 
+        val new_env = bds.foldLeft(env)((accenv: Env, bd: B) =>
+          accenv + (bd.x -> allocBind(bd.x, new_time)))
[…]

### 2018-07-07 src/main/scala/refunc/experimental/RefuncCPSBF.scala

@@ -1,0 +1,65 @@
+package refunc.experimental
+
+import refunc.ast._
+import refunc.ANFAAM._
+import scala.util.continuations._
+
+/* Experimental implementation with breadth first evaluation */
+object RefuncCPSBF {
+  type Cont = (Set[VS], Cache) => (Set[VS], Cache)
+
+  def nd[T](ts: Set[T], acc: Set[VS], cache: Cache, 
+            f: (T, Set[VS], Cache, Cont) => (Set[VS], Cache), g: Cont): (Set[VS], Cache) = {
+    if (ts.isEmpty) g(acc, cache)
+    else f(ts.head, acc, cache, (vss: Set[VS], cache: Cache) => nd(ts.tail, vss, cache, f, g))
+  }
+
+  def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, continue: Cont): (Set[VS], Cache) = {
+    val config = Config(e, env, store, time)
+    if (cache.outContains(config)) {
+      return continue(cache.outGet(config), cache)
+    }
+
+    val new_cache = cache.outUpdate(config, cache.inGet(config))
+    val new_time = config.tick
+    e match {
+      case Let(x, App(f, ae), e) =>
+        val closures = atomicEval(f, env, store).asInstanceOf[Set[Clos]]
+        nd(closures, Set[VS](), new_cache, (clos: Clos, acc: Set[VS], cache: Cache, ndk: Cont) => {
+          val Clos(Lam(v, body), c_env) = clos
+          val baddr = allocBind(v, new_time)
+          val new_env = c_env + (v -> baddr)
+          val new_store = store.update(baddr, atomicEval(ae, env, store))
+          aeval(body, new_env, new_store, new_time, cache, (bodyvss: Set[VS], cache: Cache) => { 
+            ndk(bodyvss++acc, cache) 
+          })
+        }, (result_vss: Set[VS], cache: Cache) => {
+          nd(result_vss, Set[VS](), cache, (vs: VS, acc: Set[VS], cache: Cache, ndk: Cont) => {
+            val VS(vals, time, store) = vs
+            val baddr = allocBind(x, time)
+            val new_env = env + (x -> baddr)
+            val new_store = store.update(baddr, vals)
+            aeval(e, new_env, new_store, time, cache, (evss: Set[VS], cache: Cache) => { 
+              ndk(evss++acc, cache) 
+            })
+          },
[…]

### 2018-07-07 src/main/scala/refunc/experimental/RefuncCPSNoCache.scala

@@ -5,7 +5,7 @@ import refunc.ast._
 
 import refunc.ANFAAM._
 
-object RefuncCPSNoCache {
+object RefuncCPSVSNoCache {
   type Ans = Set[VS]
   type Cont = Ans => Ans
 
@@ -58,8 +58,6 @@ object RefuncCPSNoCache {
   def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore) =
     aeval(e, env, store, mtTime, (vss => vss))
 }
-
-
 
 /* Experimental implementation with breadth first evaluation */
 object RefuncCPSNoCacheBF {

### 2018-07-07 src/test/scala/refunc/Test.scala

@@ -76,6 +76,10 @@ trait RefuncTest extends FunSuite {
         assert(summarizeUBState(DisLinearSmallStepUBStack.analyze(prog, initenv, initstore)) ==
                summarizeConfig(RefuncECPS.analyze(prog, initenv, initstore)))
         assert(DisLinearSmallStepUBStack.trace == RefuncECPS.trace)
+
+        assert(summarizeConfig(RefuncECPS.analyze(prog, initenv, initstore)) ==
+               summarizeConfig(RefuncCPSNoCache.analyze(prog, initenv, initstore)))
+        assert(RefuncCPSNoCache.trace == RefuncECPS.trace)
       
         /* The result analyzed by the refunctionalized abstract interpreter written in CPS
          * (RefuncCPS) should have the same result analyzed by the direct-style abstract 
@@ -112,7 +116,7 @@ trait RefuncTest extends FunSuite {
         /* Since here we test terminating programs, so RefuncCPS without caching should have
          * the same result against to RefuncCPS with caching. */
         assert(RefuncCPS.analyze(prog, initenv, initstore).vss ==
-               RefuncCPSNoCache.analyze(prog, initenv, initstore))
+               RefuncCPSVSNoCache.analyze(prog, initenv, initstore))
       }
     }
   }

### 2018-07-07 paper/paper.tex
Journal/CCS info

@@
 %% use defaults for review submission.
+\setcopyright{acmlicensed} %TODO
+\acmPrice{}
+\acmDOI{10.1145/3236800}
+\acmYear{2018}
+\copyrightyear{2018}
 \acmJournal{PACMPL}
 \acmVolume{2}
-\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2018}
+\acmNumber{ICFP}
+\acmArticle{105}
 \acmMonth{9}
-\acmDOI{10.1145/nnnnnnn.nnnnnnn}
+
 \startPage{1}
-
 %% Copyright information
@@
 %% use 'none' for review submission.
-% TODO: Which one?
 %\setcopyright{none}
 %\setcopyright{acmcopyright}
-\setcopyright{acmlicensed}
+%\setcopyright{acmlicensed} 
 %\setcopyright{rightsretained}
@@
   abstract machines (A\textsuperscript{2}Ms) that perform abstract interpretation from abstract machines that
-  perform concrete evaluation.
-  \citeauthor{darais2017abstracting} apply the same underlying idea to monadic definitional interpreters,
+  perform concrete evaluation. Darais et al. apply the same underlying idea to monadic definitional interpreters,
   and obtain monadic \emph{abstract} definitional interpreters (ADIs) that perform
@@
 
-% TODO: Update CCS
 %% 2012 ACM Computing Classification System (CSS) concepts
@@
 <concept>
-    <concept_id>10011007.10011006.10011008</concept_id>
-    <concept_desc>Software and its engineering~General programming languages</concept_desc>
+<concept_id>10003752.10003753.10010622</concept_id>
+<concept_desc>Theory of computation~Abstract machines</concept_desc>
 <concept_significance>500</concept_significance>
@@
 <concept>
-    <concept_id>10003456.10003457.10003521.10003525</concept_id>
-    <concept_desc>Social and professional topics~History of programming languages</concept_desc>
-    <concept_significance>300</concept_significance>
+<concept_id>10011007.10011006.10011008.10011009.10011012</concept_id>
+<concept_desc>Software and its engineering~Functional languages</concept_desc>
+<concept_significance>500</concept_significance>
 </concept>
@@
 \end{CCSXML}
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Social and professional topics~History of programming languages}
+
+\ccsdesc[500]{Theory of computation~Abstract machines}
+\ccsdesc[500]{Software and its engineering~Functional languages}
 %% End of generated code
@@
 %% comma separated list
-\keywords{refunctionalization, abstract machines, abstract interpretation}  %% \keywords are mandatory in final camera-ready submission
+\keywords{refunctionalization, abstract machines, control-flow analysis, Scala}  %% \keywords are mandatory in final camera-ready submission
 

### 2018-07-07 paper/paper.tex
replace A^2M to AAM

@@
   sound static analyses for higher-order languages, by deriving small-step \emph{abstract}
-  abstract machines (A\textsuperscript{2}Ms) that perform abstract interpretation from abstract machines that
+  abstract machines (AAMs) that perform abstract interpretation from abstract machines that
   perform concrete evaluation. Darais et al. apply the same underlying idea to monadic definitional interpreters,
@@
   into higher-order functions.
-  The refunctionalized A\textsuperscript{2}M is an abstract interpreter written in continuation-passing style (CPS)
+  The refunctionalized AAM is an abstract interpreter written in continuation-passing style (CPS)
   with two layers of continuations, which can be converted back to direct style with delimited
@@
 In addition, those abstract abstract machines with an unbounded stack naturally correspond
-to abstract definitional interpreters. We show that refunctionalizing an A\textsuperscript{2}M with an
+to abstract definitional interpreters. We show that refunctionalizing an AAM with an
 unbounded stack (and utilizing the proper caching algorithm) leads to a pushdown control-flow analysis.
@@
  \end{tikzpicture}
- \caption{Functional correspondence between and transformations from A\textsuperscript{2}Ms to ADIs} \label{fig:trans}
+ \caption{Functional correspondence between and transformations from AAMs to ADIs} \label{fig:trans}
 \end{center}
@@
 We then address the main contribution of this paper: bridging the gap between
-small-step A\textsuperscript{2}Ms and big-step ADIs by applying a series of well-known systematic transformations
+small-step AAMs and big-step ADIs by applying a series of well-known systematic transformations
 drawn from concrete semantic artifacts.
@@
   \item \textbf{Lightweight fusion}~\cite{DANVY2008100,Ohori:2007:LFF:1190216.1190241}:
-    We apply a fusion transformation to the linearized A\textsuperscript{2}M, which
+    We apply a fusion transformation to the linearized AAM, which
     combines the single-step function @step@ and the
@@
   \item \textbf{Refunctionalization}~\cite{DANVY2009534,Danvy:2006:RW:2171265.2171268}:
-    We apply refunctionalization to the disentangled A\textsuperscript{2}M, which
+    We apply refunctionalization to the disentangled AAM, which
     %as the left inverse of Reynolds' defunctionalization \cite{Reynold:72,Reynolds:HOSC98-revisited}
@@
 
-    For clarity, we first present a vanilla version of a refunctionalized A\textsuperscript{2}M which simply
+    For clarity, we first present a vanilla version of a refunctionalized AAM which simply
     converts the stack structures to higher-order functions but keeps other parts unchanged.
@@
   \item \textbf{Direct-style transformation}~\cite{DBLP:journals/scp/Danvy94}:
-    We finally transform the refunctionalized A\textsuperscript{2}M back to a
+    We finally transform the refunctionalized AAM back to a
     direct-style interpreter by using delimited control operators
@@
 
-Similar to the CESK machine, the machine state of an A\textsuperscript{2}M has a control, an environment, a store,
+Similar to the CESK machine, the machine state of an AAM has a control, an environment, a store,
 and continuation, as well as a timestamp. However, there are several notable differences
-between an A\textsuperscript{2}M's store and the CESK machine's store. In AAM, the store maps addresses to sets of
+between an AAM's store and the CESK machine's store. In AAM, the store maps addresses to sets of
 values; it stores all possible values for a particular address. As such, dereferencing
@@
 a precise control-flow from their defining language, we also would like to start from a
-pushdown A\textsuperscript{2}M that is also able to precisely match calls and returns.
+pushdown AAM that is also able to precisely match calls and returns.
 For readers who are not familiar with pushdown analysis, we provide a detailed discussion
@@
 space, we can recover the precise call/return match. Now we begin describing the
-transformations step by step. Our base machine is the A\textsuperscript{2}M with an unbounded
+transformations step by step. Our base machine is the AAM with an unbounded
 stack; as we will later transform the stack structure to higher-order functions
-representing continuations, it does not matter what kind of A\textsuperscript{2}M we start from.
+representing continuations, it does not matter what kind of AAM we start from.
 This is because it would not be possible to construct a new frame on the stack, nor
 to allocate continuations in the store after refunctionalization.
-Thus, our choice to start from an unbounded-stack A\textsuperscript{2}M is motivated simply because
+Thus, our choice to start from an unbounded-stack AAM is motivated simply because
 it has an equivalent stack model to abstract definitional interpreters (our final target).
@@
 
-With fusing complete, our A\textsuperscript{2}M appears similar to a ``big-step'' interpreter, although it
+With fusing complete, our AAM appears similar to a ``big-step'' interpreter, although it
 still has the machine state representation inside. The inverse transformation of this step
@@
 
-The purpose of disentanglement is to reveal the defunctionalized form of our A\textsuperscript{2}M.
+The purpose of disentanglement is to reveal the defunctionalized form of our AAM.
 Functions @continue@ and @ndcontinue@ that dispatch different data types representing
@@
 Before jumping into the details of refunctionalizing the stack structure and constructing
-@new_cont@ and @new_mcont@, we have two important observations in the disentangled A\textsuperscript{2}M
+@new_cont@ and @new_mcont@, we have two important observations in the disentangled AAM
 where the continuations were represented by data types:
@@
 
-In this section, based on the refunctionalized A\textsuperscript{2}M, we
+In this section, based on the refunctionalized AAM, we
 further present several other transformations toward a big-step abstract
@@
 \textit{Call/Return Matching through Refunctionalization and Caching.}
-The refunctionalized A\textsuperscript{2}M we obtained has the perfect match on return-flows even though
+The refunctionalized AAM we obtained has the perfect match on return-flows even though
 we do not have a stack model. The higher-order functions representing continuations
@@
 
-In fact, we started with an unbounded-stack A\textsuperscript{2}M which already precisely
+In fact, we started with an unbounded-stack AAM which already precisely
 matches the calls and returns. Refunctionalization makes the interpreter a ``big-step''
@@
 information in the store or on the stack.
-However, refunctionalization not only forces the nondeterministic control flow to be higher-order,
+However, refunctionalization not only forces the nondeterministic control flows to be higher-order,
 but also drives us to use a different caching algorithm.
@@
 
-%probably Store Widening \& Abstract Garbage Collection
 \section{Back to Direct-Style} \label{directstyle}
 
-In the previous section, we presented a refunctionalized A\textsuperscript{2}M in extended continuation-passing style,
+In the previous section, we presented a refunctionalized AAM in extended continuation-passing style,
 and then simplified it by utilizing the @nd@ operator.
@@
 \Citet{gilray2016pushdown} further proposed a polyvariant continuation-addresses allocator
-for small-step A\textsuperscript{2}M to achieve pushdown analysis. This method is both
+for small-step AAM to achieve pushdown analysis. This method is both
 simple to implement and computationally inexpensive and so called Pushdown for Free (P4F).
-Based on the A\textsuperscript{2}M we presented in Section \ref{aam}, the only changes in the code required is not
+Based on the AAM we presented in Section \ref{aam}, the only changes in the code required is not
 only keeping track of the entry-point expression of the callee, but also holding the target
@@
 defined language inherits properties such as the order of evaluation from the defining
-language, unless these properties are made explicit in the
-interpreter. With this insight,
+language, unless these properties are made explicit in the interpreter. With this insight,
 \citet{darais2017abstracting} constructed abstract definitional interpreters which
@@
 instead of plain $\lambda$-calculus, although our work could be easily adapted to handle plain
-$\lambda$-calculus as well. Going through the transformation steps from A\textsuperscript{2}Ms to ADIs instead of
+$\lambda$-calculus as well. Going through the transformation steps from AAMs to ADIs instead of
 only considering the final ADI result provides additional insights. In particular, it reveals that

### 2018-07-07 src/test/scala/refunc/Test.scala
Add comments

@@ -67,16 +67,26 @@ trait RefuncTest extends FunSuite {
          * interpreter against with AAM with P4F allocator.
          * 
          * The summarized store analyzed by DisLinearSmallStepUBStack machine should be
-         * the same with the one analyzed by RefuncCPS abstract interpreter.
+         * the same with the one analyzed by RefuncCPS abstract interpreter, as well as their
+         * traces.
          */
         assert(summarizeUBState(DisLinearSmallStepUBStack.analyze(prog, initenv, initstore)) ==
                summarizeVSS(RefuncCPS.analyze(prog, initenv, initstore).vss))
         assert(DisLinearSmallStepUBStack.trace == RefuncCPS.trace)
 
+        /* The summarized store analyzed by DisLinearSmallStepUBStack machine should be
+         * the same with the one analyzed by RefuncECPS abstract interpreter, as well as their
+         * traces.
+         */
         assert(summarizeUBState(DisLinearSmallStepUBStack.analyze(prog, initenv, initstore)) ==
                summarizeConfig(RefuncECPS.analyze(prog, initenv, initstore)))
         assert(DisLinearSmallStepUBStack.trace == RefuncECPS.trace)
 
+        /* The summarized store analyzed by RefuncECPS abstract interpreter should be
+         * the same with the one analyzed by RefuncCPSNoCache abstract interpreter, as well as their
+         * traces. Since in this part, we only test terminating programs, so RefuncCPS without caches
+         * works fine.
+         */
         assert(summarizeConfig(RefuncECPS.analyze(prog, initenv, initstore)) ==
                summarizeConfig(RefuncCPSNoCache.analyze(prog, initenv, initstore)))
         assert(RefuncCPSNoCache.trace == RefuncECPS.trace)

### 2018-07-07 paper/paper.tex
Update reference

@@
 language, unless these properties are made explicit in the interpreter. With this insight,
-\citet{darais2017abstracting} constructed abstract definitional interpreters which
-automatically inherit the pushdown control-flow property from its defining language because
-the defined language simply uses the call-stack model of the meta-language.
+\citeauthor{darais2017abstracting} constructed abstract definitional interpreters
+~\cite{darais2017abstracting, darais2017thesis} which automatically inherit the pushdown 
+control-flow property from its defining language because the defined language simply uses 
+the call-stack model of the meta-language.
 \citeauthor{darais2017abstracting}'s abstract definitional interpreters work on direct-style

### 2018-07-07 paper/paper.tex
revise section 6

@@
   type Addr = Int;     type Env = Map[String, Addr];    type Store = Map[Addr, Storable]
-  abstract class Storable
-  case class Clos(v: Lam, env: Env) extends Storable
+  abstract class Storable;            case class Clos(v: Lam, env: Env) extends Storable
   case class Frame(x: String, e: Expr, env: Env);       type Kont = List[Frame]
@@
 \textit{Call/Return Matching through Refunctionalization and Caching.}
-The refunctionalized AAM we obtained has the perfect match on return-flows even though
+The refunctionalized AAM with caching has the perfect match on return-flows even though
 we do not have a stack model. The higher-order functions representing continuations
-already connect all the executions in order. In fact, the continuations (call stack)
-of the analyzed language are blended into the call stack of our defining language
-(Scala) through refunctionalization, and hence the call/return flow of the analyzed
-language is naturally matched.
-
-In fact, we started with an unbounded-stack AAM which already precisely
-matches the calls and returns. Refunctionalization makes the interpreter a ``big-step''
-semantic artifact; the consequence is that we have no place to explicitly save the context
+already connect all the executions in order.
+
+In fact, we started with an unbounded-stack AAM which already precisely matches the calls and returns
+by explicitly using the data types to model the stack.
+However, in the refunctionalized AAM, the higher-order functions representing continuations of the analyzed 
+language are blended into the call stack of our defining language (Scala) through 
+refunctionalization, and hence the call/return flow of the analyzed language is naturally matched.
+This is why we say the pushdown property of the analysis is inherited from the defining language \cite{darais2017abstracting}.
+Another consequence of refunctionalization is that we have no place to explicitly save the context
 information in the store or on the stack.
-However, refunctionalization not only forces the nondeterministic control flows to be higher-order,
+
+Meanwhile, refunctionalization not only forces the nondeterministic control flows to be higher-order,
 but also drives us to use a different caching algorithm.
-
-Indeed, the caching algorithm plays an important role in guaranteeing
-the analysis will eventually terminate.
+Indeed, the caching algorithm plays an important role in guaranteeing the analysis will eventually terminate.
 An interesting implication of this is if we apply the caching algorithm with some

### 2018-07-07 paper/paper.tex
restructure section 6.2

@@
 <concept_desc>Software and its engineering~Functional languages</concept_desc>
-<concept_significance>500</concept_significance>
+<concept_significance>300</concept_significance>
+</concept>
+<concept>
+<concept_id>10011007.10011006.10011041.10010943</concept_id>
+<concept_desc>Software and its engineering~Interpreters</concept_desc>
+<concept_significance>300</concept_significance>
 </concept>
@@
 \ccsdesc[500]{Theory of computation~Abstract machines}
-\ccsdesc[500]{Software and its engineering~Functional languages}
+\ccsdesc[300]{Software and its engineering~Functional languages}
+\ccsdesc[300]{Software and its engineering~Interpreters}
 %% End of generated code
@@
           State(e, new_env, new_store, kstore, f_kaddr, new_time)
-        }
-    }
-  }
+        } } }
 \end{lstlisting}
@@
             List(State(e, new_env, new_store, konts, new_time))
-        }
-    }
-  }
+        } } }
 \end{lstlisting}
@@
       case NDState(ae, env, bstore, konts, time, ndkonts) if isAtomic(ae) => ...
-    }
-  }
+    } }
 \end{lstlisting}
@@
       case NDState(ae, env, bstore, konts, time, ndk) if isAtomic(ae) => continue(nds, new_seen)
-    }
-  }
+    } }
 \end{lstlisting}
@@
       case NDCont(Nil, _, _, _, _)::ndk =>
-        drive_step(NDState(ae, env, bstore, konts, time, ndk), seen)
+        /* drive_step() is equivalently replaced by ndcontinue() */
+        ndcontinue(NDState(ae, env, bstore, konts, time, ndk), seen)
       case NDCont(cls, argvs, bstore, time, frames)::ndk =>
-        val Clos(Lam(v, body), c_env) = cls.head
-        val baddr = allocBind(v, time);  val new_env = c_env + (v -> baddr)
-        val new_store = bstore.update(baddr, argvs)
+        val Clos(Lam(v, body), c_env) = cls.head;  val baddr = allocBind(v, time)
+        val new_env = c_env + (v -> baddr);  val new_store = bstore.update(baddr, argvs)
         val new_ndk = NDCont(cls.tail, argvs, bstore, time, frames)::ndk
@@
 
-Function @ndcontinue@ is identical to the dispatching logic in @continue@ when @konts@ is empty.
+Function @ndcontinue@ is identical to the dispatching logic in @continue@ when @konts@ is empty,
+except that the first application of @drive_step@ is reduced to @ndcontinue@ (we can safely do this because 
+@ae@ is atomic and @knots@ is empty).
 When @ndk@ is empty, the current set of explored states @seen@ is returned, after which the analysis terminates.
 Otherwise, a new @NDState@ is constructed depending on the status of the closures set.
-% first call to drive_step can be replaced with continue. ae is atomic, konts is empty.
 
@@
     if (cls.isEmpty) mcontinue else { (config, seen) =>
-      val Clos(Lam(v, body), c_env) = cls.head
-      val baddr = allocBind(v, new_time);  val new_env = c_env + (v -> baddr)
-      val new_store = store.update(baddr, argvs)
+      val Clos(Lam(v, body), c_env) = cls.head;  val baddr = allocBind(v, new_time)
+      val new_env = c_env + (v -> baddr);  val new_store = store.update(baddr, argvs)
       aeval(Config(body, new_env, new_store, new_time), seen, new_cont, makeMCont(cls.tail)) }
@@
 \begin{itemize}
+  \item The function @aeval@ is simplified by introducing and using a nondeterministic
+    operator @nd@ which makes @aeval@ only uses one layer of continuation.
   \item The components of the configuration are lifted as arguments of the evaluation
@@
     of all possible values and reachable paths.
-  \item The function @aeval@ is simplified by introducing and using a nondeterministic
-    operator @nd@ which makes @aeval@ only uses one layer of continuation.
 \end{itemize}
@@
 
+\subsubsection{Nondeterminism Abstraction}
+
+We observe that in the refunctionalized AAM, @makeMCont@ is invoked with the tail part of
+the closure set, and actually repeats the code that creates the new environment and
+store for the body expression of a closure.
+Based on this observation, we present a nondeterministic operator that abstracts this pattern:
+
+\begin{lstlisting}
+  type Ans = Set[Config]
+  def nd[T](ts: Iterable[T], acc: Ans, k: (T, Ans) => Ans): Ans =
+    if (ts.isEmpty) acc else nd(ts.tail, acc ++ k(ts.head, acc), k)
+\end{lstlisting}
+
+As a matter of convenience, we name the return type of the abstract interpreter to be @Ans@,
+which is @Set[Config]@ for the moment.
+Given a collection of elements of type @T@, an initial accumulator of type @Ans@, and a function
+@k@ that works on elements of type @T@, @nd@ chooses an element in the collection and applies @k@
+to it iteratively, and accumulates the values that returned from @k@ when recurring on the rest 
+of the collection. Eventually, @nd@ returns the accumulated value @acc@ when there is no element
+in the collection. The latest accumulated value @acc@ is exposed to @k@ when applied.
+
+Now we may refactor the function @aeval@ to one layer of continuation by using 
+the @nd@ operator.
+The use of two layers of continuations will be unchained to two functions, in that each of them 
+has one continuation.
+
+\begin{lstlisting}
+  type Cont = (Config, Ans) => Ans
+  def aeval(config: Config, seen: Ans, continue: Cont): Ans = { ...  e match {
+    case Let(x, App(f, ae), e) =>
+      val closures = atomicEval(f, env, store).toList.asInstanceOf[List[Clos]]
+      nd[Storable](closures, new_seen, { case (Clos(Lam(v, body), c_env), seen) =>
+        val baddr = allocBind(v, new_time);      val new_env = c_env + (v -> baddr)
+        val argvs = atomicEval(ae, env, store);  val new_store = store.update(baddr, argvs)
+        aeval(Config(body, new_env, new_store, new_time), seen, { case (config, seen) =>
+          val Config(ae, env_, store, time) = config;  val new_time = config.tick
+          val baddr = allocBind(x, new_time);  val new_env = env + (x -> baddr)
+          val new_store = store.update(baddr, atomicEval(ae, env_, store))
+          aeval(Config(e, new_env, new_store, new_time), seen, continue) }) })
+    case ae if isAtomic(ae) => continue(config, new_seen) } }
+\end{lstlisting}
+
+The circular meta-continuation is abstracted by the @nd@ operator, and only one continuation
+remains at the level of @aeval@. 
+When encountering a set of possible closures, we apply @nd@ on the set with the latest explored
+configurations and an anonymous function that works on each closure. The beginning part 
+(before calling @aeval@) of this anonymous function is essentially abstracted from @makeMCont@, 
+which constructs new environment and store for the closure being evaluated.
+Then we apply @aeval@ on the body expression of the closure with a continuation, 
+which is inlined from @new_cont@ and will invoke @aeval@ on @e@ with the current continuation of the @let@ expression.
+
+As we can see, by using the @nd@ operator, the evaluation function can be simplified to a 
+single-layer continuation-passing style.
+The only continuation of @aeval@ plays the same role as the one in a concrete CPS interpreter.
+
 \subsubsection{Values}
+
+Up to now, the abstract interpreter still collects and returns a set of configurations.
+But our target, the abstract definitional interpreters, should return a set of final \textit{values}
+instead of collected configurations; the intermediate values are simply discarded.
 To properly represent final values, we introduce the case class @VS@ which
@@
 \subsubsection{Cache-Passing Style}
-% why cache? problem of set of states.
+
 Recall that the latent assumption of the work-list algorithm in classical AAM is that
@@
 
-We can now transform our abstract interpreter into cache-passing style. We begin by defining the
+We can now transform our abstract interpreter into cache-passing style. We begin by redefining the
 return type @Ans@ as a case class which contains a @VS@ set and a cache. @Ans@ also implements
-its own accumulating operation @++@ which defines as an union of @Set[VS]@s and a join of caches.
+its own accumulating operation @++@ which defines as a union of @Set[VS]@s and a join of caches.
 The join operation of caches is defined as joining its two store components.
@@
 
+Accordingly, the @nd@ operator is slightly changed to cache-passing style. 
+The argument of type @Ans@ in the continuation @k@ is replaced by an argument of type @Cache@;
+every time when we apply @k@, the latest cache @acc.cache@ is provided.
+
+\begin{lstlisting}
+  def nd[T](ts: Iterable[T], acc: Ans, k: (T, Cache) => Ans): Ans =
+    if (ts.isEmpty) acc else nd(ts.tail, acc ++ k(ts.head, acc.cache), k)
+\end{lstlisting}
+
 As previously mentioned, the components of configurations are lifted as arguments to
-@aeval@. Given the cache and the simplified one-layer continuation (as we will see later),
-the shape of function @aeval@ now looks as follows:
-
-\begin{lstlisting}
-  type Cont = Ans => Ans
+@aeval@. Given the cache and the simplified one-layer continuation, the shape of function 
+@aeval@ now looks as follows:
+
+\begin{lstlisting}
   def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, continue: Cont): Ans = {
@@
 The @out@ cache of this iteration will be used as the @in@ cache for the next iteration.
-
-\subsubsection{Nondeterminism Abstraction}
-
-We observe that in the refunctionalized AAM, @makeMCont@ is invoked with the tail part of
-the closure set, and actually repeats the code that creates the new environment and
-store for the body expression of a closure.
-Based on this observation, we present a nondeterministic operator that abstracts this pattern:
-
-\begin{lstlisting}
-  def nd[T](ts: Iterable[T], acc: Ans, k: (T, Cache) => Ans): Ans =
-    if (ts.isEmpty) acc else nd(ts.tail, acc ++ k(ts.head, acc.cache), k)
-\end{lstlisting}
-
-Given a collection of elements of type @T@, an initial accumulator of
-type @Ans@, and a function @k@ that works on elements of type @T@, @nd@
-chooses an element in
-the collection and applies @k@ to it iteratively, and accumulates the values that
-returned from @k@ when recurring on the rest of the collection. Eventually, @nd@
-returns the accumulated value @acc@ when there is no element in the collection.
-The latest accumulated cache @acc.cache@ is exposed to @k@ when applied.
 
@@
 
-Now, as we can see, by using the @nd@ operator which abstracts over the meta-continuation,
-the evaluation function can be simplified to a single-layer continuation-passing style.
-The use of two layers of continuations are unchained to two functions,
-in that each of them has one continuation.
-The only continuation of @aeval@ plays the same role as the one in a
-concrete CPS interpreter.
-Furthermore, we add a cache along with the abstract interpreter and the @nd@ operator.
+Now, we have a big-step abstract interpreter written in plain continuation-passing style which
+utilizes the @nd@ operator and guarantees termination by caching.
 

### 2018-07-07 src/main/scala/refunc/DisentangledLinearSmallStepUB.scala
update code

@@ -86,13 +86,8 @@ object DisLinearSmallStepUBStack {
     }
   }
 
-  def analyze(e: Expr): Set[State] = {
+  def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore): Set[State] = {
     trace = List()
-    drive_step(inject(e), Set())
-  }
-
-  def analyze(e: Expr, env: Env, bstore: BStore): Set[State] = {
-    trace = List()
-    drive_step(inject(e, env, bstore), Set())
+    drive_step(inject(e, env, store), Set())
   }
 }

### 2018-07-07 src/main/scala/refunc/RefuncCPS.scala

@@ -135,8 +135,8 @@ object RefuncCPSNoCache {
   }
   var trace: List[Expr] = List()
 
-  def analyze(e: Expr, env: Env, bstore: BStore): Set[Config] = {
+  def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore): Set[Config] = {
     trace = List()
-    aeval(inject(e, env, bstore), Set(), { (c, seen) => seen })
+    aeval(inject(e, env, store), Set(), { (c, seen) => seen })
   }
 }

### 2018-07-07 src/main/scala/refunc/RefuncECPS.scala

@@ -66,13 +66,8 @@ object RefuncECPS {
     }
   }
 
-  def analyze(e: Expr): Set[Config] = {
+  def analyze(e: Expr, env: Env = mtEnv, store: BStore = mtStore): Set[Config] = {
     trace = List()
-    aeval(inject(e), Set(), { (c, seen, m) => m(c, seen) }, { (c, seen) => seen })
-  }
-
-  def analyze(e: Expr, env: Env, bstore: BStore): Set[Config] = {
-    trace = List()
-    aeval(inject(e, env, bstore), Set(), { (c, seen, m) => m(c, seen) }, { (c, seen) => seen })
+    aeval(inject(e, env, store), Set(), { (c, seen, m) => m(c, seen) }, { (c, seen) => seen })
   }
 }

### 2018-07-07 paper/paper.tex
small fixes + grant numbers

@@
   In this paper, we explain their functional correspondence and demonstrate how to
-  syntactically transform small-step abstract abstract machines into big-step abstract
+  systematically transform small-step abstract abstract machines into big-step abstract
   definitional interpreters.
@@
 
-\Citet{van2012systematic,van2010abstracting} proposed the Abstracting abstract machines (AAM)
+\Citet{van2012systematic,van2010abstracting} proposed the abstracting abstract machines (AAM)
 methodology which provides a recipe for constructing sound abstract interpreters
@@
 Applying the same idea to monadic definitional interpreters, \citet{darais2017abstracting}
-built monadic abstract definitional interpreters (ADIs) in big-step style.
+showed how to build monadic abstract definitional interpreters (ADIs) in big-step style.
 One of the advantages of a monadic interpreter is
@@
 interpreters, respectively. An interesting question, and the subject of this paper, is
-how one can derive one abstract semantic artifact given its counterpart.
+how one can derive one abstract semantic artifact given the other.
 The main contribution of this paper is to explain the functional correspondence
@@
 In ANF, an expression is either an atomic expression or a @let@ expression.
-A restriction exists which states that all function applications must be administered
+A restriction exists which states that all function applications (and only those) must be administered
 within a @let@ expression and then bound to a variable name under the current environment.
@@
 
-The abstract syntax in Scala is shown as follows. We assume that the source program conforms
+The abstract syntax is represented in Scala as follows. We assume that the source program conforms
 to the ANF convention, and as such, do not enforce it in the term structure of Scala constructs.
@@
 machines \cite{van2012systematic, van2010abstracting}. An abstract abstract machine
-implements computable abstract semantics which approximates the run-time behaviors of
+implements a computable abstract semantics which approximates the run-time behaviors of
 programs.
@@
 
-Similar to the CESK machine, the machine state of an AAM has a control, an environment, a store,
-and continuation, as well as a timestamp. However, there are several notable differences
+Similar to the concrete CESK machine, the machine state of the derived AAM has a control component, an environment, a store,
+a continuation, as well as a timestamp. However, there are several notable differences
 between an AAM's store and the CESK machine's store. In AAM, the store maps addresses to sets of
@@
 
-  This work was supported in part by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm} and by the
-  \grantsponsor{GS100000001}{DOE}{XXX} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn}.  \note{TODO: grants}
+  This work was supported in part by 
+  NSF awards 1553471 and 1564207, and DOE award DE-SC0018050.
+  % \grantsponsor{GS100000001}{National Science
+  %   Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
+  % No.~\grantnum{GS100000001}{nnnnnnn} and Grant
+  % No.~\grantnum{GS100000001}{mmmmmmm} and by the
+  % \grantsponsor{GS100000001}{DOE}{XXX} under Grant
+  % No.~\grantnum{GS100000001}{nnnnnnn}.  \note{TODO: grants}
 \end{acks}

### 2018-07-08 paper/paper.tex
misc

@@
 %% use defaults for review submission.
-\setcopyright{acmlicensed} %TODO
+\setcopyright{rightsretained}
 \acmPrice{}
@@
     case hd::tl if seen.contains(hd) => drive(tl, seen)
-    case hd::tl => drive(step(hd).toList ++ tl, seen + hd)
+    case hd::tl => drive(step(hd) ++ tl, seen + hd)
   }
@@
 
-Function @f1@ and @f2@ mutually invoke each other, so the frames of @f1@ and @f@ will be
+Function @f1@ and @f2@ mutually invoke each other, so the frames of @f1@ and @f2@ will be
 alternately pushed onto the top of the current stack. However, no two existing stack
@@
 
-Function @ndcontinue@ is identical to the dispatching logic in @continue@ when @konts@ is empty,
+Function @ndcontinue@ is identical to the foregoing dispatching logic of meta-continuations when @konts@ is empty,
 except that the first application of @drive_step@ is reduced to @ndcontinue@ (we can safely do this because 
@@
 Then we apply @aeval@ on the body expression of the closure with a continuation, 
-which is inlined from @new_cont@ and will invoke @aeval@ on @e@ with the current continuation of the @let@ expression.
+which is inlined from @new_cont@ and will apply @aeval@ to @e@ with the current continuation of the @let@ expression.
 

### 2018-07-08 src/main/scala/refunc/SmallStep.scala

@@ -44,7 +44,7 @@ object SmallStep {
   def drive(todo: List[State], seen: Set[State]): Set[State] = todo match {
     case Nil => seen
     case hd::tl if seen.contains(hd) => drive(tl, seen)
-    case hd::tl => drive(step(hd).toList ++ tl, seen + hd)
+    case hd::tl => drive(step(hd) ++ tl, seen + hd)
   }
 
   def analyze(e: Expr): Set[State] = drive(List(inject(e)), Set())

### 2018-07-08 paper/paper.tex
remove useless author info
replace list with set

@@
 \email{tiark@purdue.edu}          %% \email is recommended
-
-\iffalse
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-\fi
 
@@
 The structure of function @step@ is similar to the concrete CESK machine,
-except the nondeterminism which makes @step@ return a list of reachable successor states.
+except the nondeterminism which makes @step@ return a set of reachable successor states.
 We have two cases to consider (code shown below):
@@
     a new frame @Frame(x, e, env)@ paired with the current continuation address @kaddr@ is
-    merged into @new_kaddr@. Finally, a list of successor states is generated.
+    merged into @new_kaddr@. Finally, a set of successor states is generated.
 
@@
 \begin{lstlisting}
-  def step(s: State): List[State] = {
+  def step(s: State): Set[State] = {
     val new_time = s.tick
@@
       case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) =>
-        val closures = atomicEval(f, env, bstore).toList
+        val closures = atomicEval(f, env, bstore)
         for (Clos(Lam(v, body), env_c) <- closures) yield {
@@
       case State(ae, env, bstore, kstore, kaddr, time) if isAtomic(ae) =>
-        val conts = kstore(kaddr).toList
-        for (Cont(Frame(x, e, env_f), f_kaddr) <- conts) yield {
+        for (Cont(Frame(x, e, env_f), f_kaddr) <- kstore(kaddr)) yield {
           val baddr = allocBind(x, new_time)
@@
     case hd::tl if seen.contains(hd) => drive(tl, seen)
-    case hd::tl => drive(step(hd) ++ tl, seen + hd)
+    case hd::tl => drive(step(hd).toList ++ tl, seen + hd)
   }
@@
 
-The state transition function @step@ shown below is still
-nondeterministic, but only when dereferencing the callee @f@ from the
-function application @App(f, ae)@.
-
-\begin{lstlisting}
-  def step(s: State): List[State] = {
+The state transition function @step@ shown below is still nondeterministic, but only when 
+dereferencing the callee @f@ from the function application @App(f, ae)@.
+
+\begin{lstlisting}
+  def step(s: State): Set[State] = {
     val new_time = s.tick
@@
       case State(Let(x, App(f, ae), e), env, bstore, konts, time) if isAtomic(ae) =>
-        for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore).toList) yield {
+        for (Clos(Lam(v, body), env_c) <- atomicEval(f, env, bstore)) yield {
           val frame = Frame(x, e, env)
@@
         konts match {
-          case Nil => List()
+          case Nil => Set()
           case Frame(x, e, env_f)::konts =>
@@
             val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-            List(State(e, new_env, new_store, konts, new_time))
+            Set(State(e, new_env, new_store, konts, new_time))
         } } }

### 2018-07-08 src/main/scala/refunc/SmallStep.scala

@@ -5,24 +5,24 @@ import refunc.ast._
 import ANFAAM._
 
 object SmallStep {
-  def step(s: State): List[State] = {
+  def step(s: State): Set[State] = {
     val new_time = s.tick
     s match {
       case State(Let(x, ae, e), env, bstore, kstore, kaddr, time) if isAtomic(ae) =>
         val baddr = allocBind(x, new_time)
         val new_env = env + (x -> baddr)
         val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-        List(State(e, new_env, new_store, kstore, kaddr, new_time))
+        Set(State(e, new_env, new_store, kstore, kaddr, new_time))
 
       case State(Letrec(bds, body), env, bstore, kstore, kaddr, time) =>
         val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => 
           accenv + (bd.x -> allocBind(bd.x, new_time)))
         val new_store = bds.foldLeft(bstore)((accbst: BStore, bd: B) => 
           accbst.update(allocBind(bd.x, new_time), atomicEval(bd.e, new_env, accbst)))
-        List(State(body, new_env, new_store, kstore, kaddr, time))
+        Set(State(body, new_env, new_store, kstore, kaddr, time))
 
       case State(Let(x, App(f, ae), e), env, bstore, kstore, kaddr, time) if isAtomic(ae)=>
-        for (Clos(Lam(v, body), c_env) <- atomicEval(f, env, bstore).toList) yield {
+        for (Clos(Lam(v, body), c_env) <- atomicEval(f, env, bstore)) yield {
           val baddr = allocBind(v, new_time)
           val new_env = c_env + (v -> baddr)
           val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
@@ -32,7 +32,7 @@ object SmallStep {
         }
 
       case State(ae, env, bstore, kstore, kaddr, time) if isAtomic(ae)=>
-        for (Cont(Frame(x, e, f_env), f_kaddr) <- kstore(kaddr).toList) yield {
+        for (Cont(Frame(x, e, f_env), f_kaddr) <- kstore(kaddr)) yield {
           val baddr = allocBind(x, new_time)
           val new_env = f_env + (x -> baddr)
           val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
@@ -44,7 +44,7 @@ object SmallStep {
[…]

### 2018-07-08 src/main/scala/refunc/SmallStepUB.scala

@@ -12,24 +12,24 @@ object SmallStepUBStack {
   def inject(e: Expr, env: Env = Map(), bstore: Store[BAddr, Storable] = Store[BAddr, Storable](Map())): State =
     State(e, env, bstore, List(), List())
 
-  def step(s: State): List[State] = {
+  def step(s: State): Set[State] = {
     val new_time = s.tick
     s match {
       case State(Let(x, ae, e), env, bstore, konts, time) if isAtomic(ae) =>
         val baddr = allocBind(x, new_time)
         val new_env = env + (x -> baddr)
         val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-        List(State(e, new_env, new_store, konts, new_time))
+        Set(State(e, new_env, new_store, konts, new_time))
 
       case State(Letrec(bds, body), env, bstore, konts, time) =>
         val new_env = bds.foldLeft(env)((accenv: Env, bd: B) => 
           accenv + (bd.x -> allocBind(bd.x, new_time)))
         val newBStore = bds.foldLeft(bstore)((accbst: BStore, bd: B) => 
           accbst.update(allocBind(bd.x, new_time), atomicEval(bd.e, new_env, accbst)))
-        List(State(body, new_env, newBStore, konts, time))
+        Set(State(body, new_env, newBStore, konts, time))
 
       case State(Let(x, App(f, ae), e), env, bstore, konts, time) =>
-        for (Clos(Lam(v, body), c_env) <- atomicEval(f, env, bstore).toList) yield {
+        for (Clos(Lam(v, body), c_env) <- atomicEval(f, env, bstore)) yield {
           val frame = Frame(x, e, env)
           val baddr = allocBind(v, new_time)
           val new_env = c_env + (v -> baddr)
@@ -39,12 +39,12 @@ object SmallStepUBStack {
 
       case State(ae, env, bstore, konts, time) if isAtomic(ae) =>
         konts match {
-          case Nil => List()
+          case Nil => Set()
           case Frame(x, e, f_env)::konts =>
             val baddr = allocBind(x, new_time)
             val new_env = f_env + (x -> baddr)
             val new_store = bstore.update(baddr, atomicEval(ae, env, bstore))
-            List(State(e, new_env, new_store, konts, new_time))
+            Set(State(e, new_env, new_store, konts, new_time))
         }
     }
   }
@@ -52,7 +52,7 @@ object SmallStepUBStack {
[…]

### 2018-07-08 paper/paper.tex
more fixes

@@
 
-The pushdown AAM with an unbounded stack is also described by \citeauthor{van2012systematic}
-\cite{van2012systematic}.
+The pushdown AAM with an unbounded stack is also described by \citet{van2012systematic}.
 We show it here is to maintain consistency of abstract semantics during
@@
 \textit{Note on Termination.}
-Unfortunately, even though other components in the states are finite, this version of AAM with an unbounded
+Unfortunately, even though other components in the state are finite, this version of AAM with an unbounded
 stack may still diverge when analyzing some programs.
@@
 technique that converts the data types representing meta-continuations (failure continuations)
-back to lazy list representation, where the elements in the list are plain machine states.
+back to a lazy list representation, where the elements in the list are plain machine states.
 
@@
 
-With fusing complete, our AAM appears similar to a ``big-step'' interpreter, although it
+With fusion complete, our AAM appears similar to a ``big-step'' interpreter, although it
 still has the machine state representation inside. The inverse transformation of this step
@@
 continuation-passing style~\cite{Danvy:1990:AC:91556.91622}.
-This transformation can be also regarded as making the control flows of the defined
-language be part of the control flows of the meta-language.
+This transformation can be also regarded as expressing the control flow of the defined
+language through control flow of the defining language.
 In this section, we first present a direct refunctionalization of the continuations.
@@
 of this step is known as \textit{de}functionalization, which converts the higher-order functions
-representing continuations back to first-order data
+representing continuations to first-order data
 types~\cite{Danvy:2001:DW:773184.773202, Danvy:2008:DIP:1411204.1411206}.
@@
 \textit{The Problem with Return-flows.}
-Pushdown control-flow is a property in the analysis that precisely models the run-time
+Pushdown control-flow is a property of the analysis that precisely models the run-time
 call structure of the analyzed program. A pushdown control-flow analysis provides an
@@
 \textit{Call/Return Matching through Refunctionalization and Caching.}
-The refunctionalized AAM with caching has the perfect match on return-flows even though
+The refunctionalized AAM with caching precisely matches return-flows even though
 we do not have a stack model. The higher-order functions representing continuations
@@
     assignments and mutations to update the cache, and then achieve the same definitional
-    interpreter with pushdown control-flows. \Citet{friedman-ai}, in a tutorial on abstract
+    interpreter with pushdown control-flow. \Citet{friedman-ai}, in a tutorial on abstract
     interpreters, use this style to model caches and updates.
@@
 Eisenberg:2014:PFT:2633357.2633361, fourtounis2014modular, 10.1007/3-540-46425-5_4, consel1993tour}.
-Refunctionalization~\cite{Danvy:2006:RW:2171265.2171268} as a
+Refunctionalization as a
 left inverse of defunctionalization, was introduced
-by \citeauthor{Danvy:2006:RW:2171265.2171268,DANVY2009534}.
+by \citet{Danvy:2006:RW:2171265.2171268,DANVY2009534}.
 
 \Citet{Ager:2003:FCE:888251.888254} observed that defunctionalization and refunctionalization
-can be used to show the functional correspondence between interpreters and abstract machines
-\cite{Ager:2003:FCE:888251.888254}. A number of independently designed small-step semantic artifacts can
+can be used to show the functional correspondence between interpreters and abstract machines. A number of independently designed small-step semantic artifacts can
 be obtained by closure-converting, CPS-transforming, and then defunctionalizing their big-step counterparts,
@@
 Lightweight fusion as one of the transformation in our paper was presented by
-\citeauthor{Ohori:2007:LFF:1190216.1190241} \cite{Ohori:2007:LFF:1190216.1190241}.
-As shown by \citeauthor{DANVY2008100}, lightweight fusion can be used to show the
-equivalence between small-step and big-step abstract machines \cite{DANVY2008100}.\\
+\citet{Ohori:2007:LFF:1190216.1190241}.
+As shown by \citet{DANVY2008100}, lightweight fusion can be used to show the
+equivalence between small-step and big-step abstract machines.\\
 
 \textit{Abstract Interpretation and Control-Flow Analysis.}
-\citeauthor{cousot1977abstract} discovered abstract interpretation as a sound approach to
-approximate a program's run-time behavior \cite{cousot1977abstract}. Control-flow analysis
-is one instance of abstract interpretation on higher-order functional programs that can be traced to
-Jones \cite{jones1981flow}. \Citet{Shivers:1988:CFA:53990.54007,Shivers:1991:SSC:115865.115884}
+\Citet{cousot1977abstract} discovered abstract interpretation as a sound approach to
+approximate a program's run-time behavior. Control-flow analysis
+is one instance of abstract interpretation on higher-order functional programs that can be traced to \citet{jones1981flow}. \Citet{Shivers:1988:CFA:53990.54007,Shivers:1991:SSC:115865.115884}
 introduced $k$-CFA which uses $k$ recent calling contexts as program contours that
@@
 methodology \cite{van2012systematic, van2010abstracting} which forms the starting point
-of this paper. Modern programming languages such as Java \cite{might2010resolving} and
-Racket \cite{Tobin-Hochstadt:2012:HSE:2384616.2384655} can also be modeled by the abstracting
-abstract machines approach. \\
+of this paper. The AAM approach has been successfully applied to programming languages such as Java \cite{might2010resolving} and
+Racket \cite{Tobin-Hochstadt:2012:HSE:2384616.2384655}. \\
 

### 2018-07-11 paper/paper.tex
fix title/subtitle

@@
 %% Title information
-\title{Refunctionalization of Abstract Abstract Machines (Functional Pearl)}         %% [Short Title] is optional;
+\title{Refunctionalization of Abstract Abstract Machines}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
                                         %% contents suppressed with 'anonymous'
-\subtitle{Bridging the Gap Between Abstract Abstract Machines and Abstract Definitional Interpreters}        %% \subtitle is optional
+\subtitle{Bridging the Gap between Abstract Abstract Machines and Abstract Definitional Interpreters (Functional Pearl)}        %% \subtitle is optional
 %\subtitlenote{with subtitle note}       %% \subtitlenote is optional;

### 2018-08-05 paper/paper.tex
Revised on CFA2

@@
 CFA2 is the first solution that solves the return-flow problem~\cite{vardoulakis2010cfa2},
-but CFA2 has several limitations: it only works on continuation-passing style programs, and
-does not support polyvariant analysis, in addition to having an exponential time complexity.
+which is formulated as a monovariant analysis on programs written in continuation-passing style.
+The original presentation of CFA2 has an exponential runtime complexity, 
+and several improvements are proposed by \citet{vardoulakis-diss12} afterwards.
+But implementing CFA2 is considered requiring significant engineering efforts.
+
 Pushdown control-flow analysis (PDCFA) is a mechanism which maintains this precision through

### 2018-08-20 paper/paper.tex
misc

@@
 The original presentation of CFA2 has an exponential runtime complexity, 
-and several improvements are proposed by \citet{vardoulakis-diss12} afterwards.
+and several variants and improvements are proposed by \citet{vardoulakis-diss12} afterwards.
 But implementing CFA2 is considered requiring significant engineering efforts.

### 2020-12-30 paper/paper.tex
misc

@@
 with a wide range of uses. For example, \citet{Danvy:2001:DW:773184.773202} showed that
-defunctionalization is dual to Church encoding, and further case studies where known
+defunctionalization is dual to Church encoding, and further case studies show that known
 higher-order and first-order implementations can be related through defunctionalization
-and refunctionalization include backtracking algorithms (e.g., regular expression matcher),
+and refunctionalization, including backtracking algorithms (e.g., regular expression matcher),
 Dyck word recognizer, Dijkstra's shunting-yard algorithm, and even Quicksort
